### 2018-11-14 writing/draft1/draftMonika.tex
adding

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{a4wide}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage{colortbl}
\usepackage{footnote}
\makesavenoteenv{tabular}
\usepackage[flushleft]{threeparttable}
\usepackage[hyphens]{url}
\usepackage{placeins}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{changepage}  
\usepackage{mathrsfs}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{tabularx}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage[labelfont=bf]{caption}
\usepackage{titlesec}
\usepackage{endnotes}
\usepackage{csquotes}
\usepackage{epstopdf}
\usepackage{euscript}
\usepackage{textcomp}

\date{\normalsize{October 2018}}
\title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens}
\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}

\parindent 0pt
\parskip 0.5em
\newcommand\starred[1]{\accentset{~~~~~\star}{#1}}

\newcounter{magicrownumbers}
\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}

\begin{document}

\newdateformat{monthyeardate}{%
  \monthname[\THEMONTH], \THEYEAR}

  \interfootnotelinepenalty=10000
 \newcolumntype{d}{D{.}{.}{-1}}
 \newcolumntype{e}{D{+}{\,\pm\,}{6,2}}

\makeatletter
\def\hlinewd#1{%
\noalign{\ifnum0=`}\fi\hrule \@height #1 %
\futurelet\reserved@a\@xhline}
\makeatother

\maketitle
\vfill

\doublespacing

\begin{abstract}
\noindent \textcolor{red}{..the full abstract to be written..} This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security. Our dataset consists of yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
\\
...Applying the linear mixed effects models, we found that...
\end{abstract}

\noindent \textbf{Keywords:}  Drought, Early warning systems, Food security, Food systems, Kenya, Mixed effects models\\

\newpage
\sloppy

\section{Introduction and literature review}\label{Introduction}

Consequences of droughts and extreme heat waves on agriculture have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}). Agricultural drought and soil moister are important factors influencing plant health but they are also essential for land-atmospheric feedback and for temperature variability \citep{nicolai2017}.

Frequent occurrence of severe droughts have been a major problem in Kenya and in other Sub-Saharan countries \citep{Compendium, WorldBank2015, Nicholson2017}. The agricultural sector is the main contributor to the economy of Kenya, hence the effects of drought are especially damaging in this country \citep{MendelsohnDinar2000}. Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment. In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}. Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.  As a consequence of climate change, the situation is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}. Strong downwards trend in precipitation has already been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.

Droughts and climate variability do not only affect agriculture but they have widespread consequences on food security in general. \cite{Ubilava2018} has found a connection between commodity prices and phases of the ENSO cycle. \cite{robinson2010} have used a Computable General Equilibrium (CGE) model for Ethiopia and they have shown that production shocks and resulting price increases have negative effects on farmers' income in drought-prone areas while moderate price increases can be beneficial for farmers in less drought-prone areas. \cite{OxfamIDS} have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM and they have estimated that climate change will result in decline in food productivity while food prices will increase. \cite{BROWN2015} have investigated food prices in $554$ local commodity markets in Africa, South Asia and Latin America and they concluded that almost 20\% of local market food prices had been affected by weather disturbances.

A large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa. A frequent finding is that increase in temperature affects crop revenue negatively while increase in precipitation has positive effects.  \cite{RicardianBello} and  \cite{Ochieng2016} found these results by the means of Ricardian analysis. \cite{Ochieng2016} focused on Kenya while \cite{RicardianBello} addressed the situation in Niger.

Other studies have investigated the relationship of climate and yields in the context of climate change. \cite{MendelsohnDinar2000} have developed a simulation model to evaluate impacts of climate change on African agriculture. According to their study, agricultural GDP and climate change exhibit a hill-shaped relationship. Hence, global warming should be beneficial in high latitudes but harmful in low latitudes. The estimated loss of the Africa as continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}. \cite{kabubo2007} have shown that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial. Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}. \cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa. They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful. \cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania. The authors have shown that by 2050 the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather. \cite{Kabubo2015} have estimated effects of climate change on food security in Kenya. They found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped. Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
\cite{Mendelsohn2008} reviewed several studies on effects of climate change on agriculture in developing countries and based on the revision he confirmed that tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture. \cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia. The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions. According to another meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming. However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.

A number of important studies have focused on responses to drought related disasters or emergency situations. \cite{Sandstorm2017} have questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They have analysed causes of the food crisis identified by the humanitarian appeal documents. Majority of these documents have found food availability and food production to be the major causes of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts. According to \cite{Sandstorm2017}, a large share of humanitarian response budget has been focused on emergency food aid in comparison to the share focused on interventions to build resilience. The authors have suggested to increase the budget share focused on agricultural and livestock production to build resilience. Also other authors have promoted shifting from reactive to proactive approach in disaster risk management (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).  \cite{Nicholson2017} has recommended forecast based financing as it can avoid significant disaster losses. \cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses. Furthermore, risk assessment is essential for disaster risk management \citep{IPCC2012ch1}. In the light of these arguments, the main goal of the present study is to develop a model for assessing the risk of drought based on weather and climate data. This should help to improve focus on proactive rather than reactive approach in drought disaster risk management.

In this study, we focus on Kenya utilizing an yearly panel dataset of $47$ counties over the period of $1981-2017$. As a measure of food security we use agricultural maize yields and as a measure of drought we use precipitation and temperature data.

To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya. Hence, this is the first time that effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country. To the best of our knowledge, this is also the first study which has utilised as comprehensive and detailed weather data  ...resolution...source
% Kabubo-Mariara and Karanja (2015) similar study> but they use the more econometric approach (fixed effects models and random effects models as referred to by econometricians
\section{Methodology}\label{Method}

\subsection{Measures of food security and drought}

Before beginning our research, we had to answer two essential % fundamental??
questions: `How should we measure food security?' and `How should we measure drought?' Taking into account the aim of this study, previous literature (see Section~\ref{Introduction}) and data availability, we were considering three main approaches to measuring food security. In particular, we considered using health and utility indicators, food prices or agricultural crop yields as measures of food security. Although some health and utility indicators including the Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (see \citealp{CSI}) or the NDMA early warning phases are available in the NDMA Early Warning Bulletins, the data are only available since July $2013$. Furthermore, the scales of indicators do not seem to be consistent over the counties and time and the values of indicators or even entire reports are frequently missing. Therefore, we decided not to use the health and utility indicators.

Using food prices as a measure of food security turned out be problematic as well. 
One problem is data availability as the food prices are usually available for market towns rather than at county level. Furthermore, the market data which are available in the publications of \cite{KNBS} (KNBS) only cover several years. Another problem is that prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless the model explicitly accounts for the spatial autocorrelation structure. Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and also additional data, which may not be available. Therefore we opted for agricultural crop yields as a measure of food security and we decided to estimate a single reduced form equation with crop yields as the dependent variable. In particular, we focus on maize yields as this crop is the principal staple food in Kenya and it is grown at 90\% of farms in Kenya \citep{FAO}.

The other important question to answer before starting our analysis was: `How is drought defined and which definition and measure of drought would be the best for the purpose of this study?' Various definitions of drought and their role have been reviewed by \cite{wilhite1985} and \cite{wilhite2000}. For an extensive overview of drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}. \cite{keyantash2002} have quantified, evaluated and compared number of drought indices for meteorological, hydrological and agricultural forms of drought. For more details about types, measures and indices of drought see Appendix~$1$.

In the recent period, remote sensing data have been used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}. Based on these data, various measures such as the vegetation condition index (VCI) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.

We were considering to use the Standardised Precipitation Evapotranspiration Index  (SPEI) or VCI as measures of drought in our analysis. However, we did not find the county level SPEI to be as good predictor of food security as other precipitation and temperature measures and computing the SPEI for the entire high resolution weather dataset would be inadequately demanding. We did not use the VCI either as we were not confident enough about the reliability of the available data.\footnote{The values of the VCI which we obtained from the NDMA were not the same as those available by the \cite{BOKU} (BOKU). Furthermore, the VCI values did not exactly correspond to the early warning phases as defined by the \cite{NDMA} (NDMA).}

Thus, after conducting a rigorous literature research on measures of drought, we decided to utilize the raw daily precipitation data from...CHIRPS and...resolution... The frequency of the maize yield data is yearly while the weather data are daily. Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data. There are many possible ways how to aggregate daily weather data. Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}). \cite{Adejuwon2004} analysed relationship of crop yields and three measures of precipitation. The measures of precipitation include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August). Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa \citep{Adejuwon2004}. Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010}) or annual totals or means \citep{BLIGNAUT2009}.  Some studies are based on simulated daily extremes, averages, daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}. Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}). The definition of wet and dry days and rainy season vary across the literature. For example, \cite{BenMohamed2002} has assumed that rainy season begins when the amount of rainfall in three consecutive days reach at least 25mm and no dry spell of more than seven days occurs in the following thirty days. According to this study, the end of the rainy season is defined as that rainy day after which rain recorded during 20 days is less than 5mm. \cite{BenMohamed2002} has also found sea surface temperature anomalies at various locations and amount of rainfall in July, August and September to be significant for millet crops in Niger. The author has also considered the maximum air temperature in the hottest month (April) and the minimum air temperature in the coldest month (January) as possible predictors of crops in Niger, but he did not find them significant.

An important group of studies that have analysed relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).

\textcolor{red}{Erin Lentz, can we cite her paper?She has not responded to my email}

Based on the literature research above and complexity of deriving the measures, we short-listed the following aggregates of the weather data:

\begin{itemize}

\item \textbf{Precipitation:}

\begin{itemize}
\item Seasonal cumulative rainfall
\item Seasonal standard deviation 
\item Seasonal coefficient of variation 
\item Maximum length of dry spell in number of days
\item ...\textcolor{blue}{the list to be finished}
\item...
\end{itemize}
\item Temperature
\begin{itemize}
\item
\item\textcolor{blue}{the list to be finished}..
\item
\end{itemize}
\end{itemize}
\subsection{Data}

\subsection{Linear mixed models}
\sloppy
Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}. As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties. Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}. Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset. These types of models include both fixed affects and random effects. Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}. On the other hand, random effect are unobserved random variables. There are at least three benefits of treating a set of parameters as a random sample from some distribution. \textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9,KERYch12}).

Formally, a linear mixed model can be described by the distribution of two vectors of random variables: the response $\mathscr{Y}$ and the vector of random effects $\mathscr{B}$. The distribution of $\mathscr{B}$ is multivariate normal and the conditional distribution of $\mathscr{Y}$ given $\mathscr{B}=\mathbf{b}$ is multivariate normal of a form (\citealp{bates2010lme4, KERYch9}):

\begin{equation}\label{MixedGeneral}
\begin{array}{lcl}

(\mathscr{Y}|\mathscr{B}=\mathbf{b})& \sim & \mathit{N}(\mathbf{X}\mathbf{\beta}+\mathbf{Z}\mathbf{b},\sigma^2\mathbf{I}),

\end{array}
\end{equation}

where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variable $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ a scale factor. The distribution of $\mathscr{B}$ can be written as:

\begin{equation}\label{ranefDist}
\mathscr{B} \sim \mathit{N}(0,\mathbf{\Sigma}),
\end{equation}

where $\mathbf{\Sigma}$ is a $q \times q$ positive semi-definite variance-covariance matrix

\FloatBarrier
	\section{Results and discussion}\label{Results}

\large Findings:
\begin{itemize}
\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important

\item Mar-Sept last year temperature very important for maize yields
\item SD temperature last year positive and significant
\item dry spell 20 MAM last year important (but not dry spell MAM10)
\item interesting. Precipitation 2 months MAM last year very significant and positive
\item mean temp last year negative and significant, hill shaped 
\end{itemize}

\underline{New findings:}

\begin{itemize}
\item The yields seem to be more responsive to weather on west than on east
\end{itemize}
\normalsize
\FloatBarrier
\pagebreak

{\centering
\begin{threeparttable}

\singlespacing
\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather}}
% Peggy45ln can be found in foodSystems/Rcodes/Lags/..
 % Yield in tonnes per hectare
\label{Peggy45ln} 
\centering
\begin{small}
\begin{tabular}{lrl} 
\hline \vspace{-0.2cm} \\

\vspace{-0.2cm} \\

  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{1}{c}{\textit{Estimate}} &\multicolumn{1}{c}{\textit{p-value}}\\
 \hline 
\hline
\\
\vspace{-0.2cm}Intercept&$0.158$&$0.086^{\bullet}$\\
  \\
\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.066$&$0.008^{**}$\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$-0.006$&$0.861$\\
  \\
  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.036$&$0.292$\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.081$&$0.008^{**}$\\
  \\

      \vspace{-0.2cm}Prec. max OND, east&$0.081$&$0.056^{\bullet}$\\
  \\
        \vspace{-0.2cm}Prec. max OND, west&$0.108$&$0.009^{**}$\\
  \\
    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.101$&$0.003^{**}$\\
  \\
      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, west&$0.142$&$6\times10^{-7}$ $^{***}$\\
  \\
  \hline
\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
\vspace{-0.2cm}
\\
\hline
\\
  \vspace{-0.2cm}Intercept\\
  \\
  \vspace{-0.2cm}Prec. cum. MAM+OND lag 1\\
  \\
  \vspace{-0.2cm}Prec. max OND\\
  \\
    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1\\
  \vspace{-0.1cm} \\ 
  \hline

\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.05cm}$584$ observations
        \begin{adjustwidth}{1cm}{} 
    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\pagebreak

%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT

%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT

\makeatletter 
\renewcommand{\thesection}{\hspace*{-1.0em}}
\newpage
\linespread{1}
\bibliographystyle{ChicagoM}
%\bibliographystyle{apa}
\bibliography{referencesFS}

\newpage

\setcounter{table}{0} 
\makeatletter 
\renewcommand{\thetable}{A\@arabic \c@table} 
\FloatBarrier

\section{Appendix 1 Drought: definitions, measures and indices}
According to the international meteorological community, drought can be defined in several ways. In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
 \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater. The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).

\cite{wilhite1985} and \cite{wilhite2000} have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational. Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}. Operational definitions are essential for an effective early warning system. An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.

In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary. However, as a result of a large disagreement about a definition of drought, there is no single universal drought index. Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).

Examples of early measures of drought are \cite{wilhite1985}, \cite{munger1916}, \cite{blumenstock1942} or \cite{mcquigg1954}. \cite{munger1916} suggested to use length of period without $24$-h precipitation of $1.27$ mm. \cite{wilhite1985} is based on a measure of precipitation over a given time period. \cite{blumenstock1942} proposed to measure severity of drought as a length of drought in days where the end of a drought is defined by occurrence of $2.54$ mm of precipitation in $48$ hours. \cite{mcquigg1954} developed the Antecedent Precipitation Index (API) which is based on amount and timing of precipitation and it was used for forecasting of floods. Hence, the API is a reverse drought index.

The study of \cite{palmer1965} was a significant milestone in the history of quantification of drought severity. \cite{palmer1965} developed the Palmer Drought Severity Index (PDSI) using a complex water balance model. The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand (\citealp{Heim2002,palmer1965}). As the PDSI suffers from several weaknesses (for details see e.g. \citealt{Heim2002}), other indices were developed in the following decades. These include the standardized precipitation index (SPI) developed by \cite{SPI} and the standardized precipitation evapotranspiration index (SPEI) developed by \cite{SPEI}. The SPI specifies observed precipitation as a standardised departure from a chosen probability distribution which models the precipitation data. Values of SPI can be viewed as a multiple of standard deviations by which the observed amount of rainfall deviates from the long-term mean \citep{SPIonline}.\footnote{Can be created for various periods of 1-36 months, usually using monthly data.} The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of evapotranspiration (which captures increased temperature). It is based on water balance, therefore it can be compared to the self-calibrated PDSI \citep{SPEI}.

\section{Appendix 3 Tables}

\end{document}

### 2018-11-14 writing/draft1/draftMonika1.tex
adding

--- 2018-11-14 writing/draft1/draftMonika.tex
@@
 Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
-As a consequence of climate change, the situation is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
+As a consequence of climate change, the situation is likely to get worse.
+Dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
 Strong downwards trend in precipitation has already been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
 During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
-Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.
+Furthermore, increased frequency of heavy rain events has been observed in the areas with decline in total rainfall \citep{IPCCtrenberth}.
 
@@
 \cite{robinson2010} have used a Computable General Equilibrium (CGE) model for Ethiopia and they have shown that production shocks and resulting price increases have negative effects on farmers' income in drought-prone areas while moderate price increases can be beneficial for farmers in less drought-prone areas.
-\cite{OxfamIDS} have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM and they have estimated that climate change will result in decline in food productivity while food prices will increase.
+\cite{OxfamIDS} have focused on sub-Saharan Africa, central America, North Africa and other low-income countries selected by Oxfam and he has concluded that climate change will result in decline in food productivity while food prices will increase.
 \cite{BROWN2015} have investigated food prices in $554$ local commodity markets in Africa, South Asia and Latin America and they concluded that almost 20\% of local market food prices had been affected by weather disturbances.
@@
 A large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa.
-A frequent finding is that increase in temperature affects crop revenue negatively while increase in precipitation has positive effects.
-\cite{RicardianBello} and  \cite{Ochieng2016} found these results by the means of Ricardian analysis.
-\cite{Ochieng2016} focused on Kenya while \cite{RicardianBello} addressed the situation in Niger.
+Increase in temperature has typically been found to affect crop revenue negatively while increase in precipitation has mostly been associated with positive effects in the literature.
+\cite{RicardianBello} and  \cite{Ochieng2016} have found these results by the means of Ricardian analysis.
+\cite{Ochieng2016} have focused on Kenya while \cite{RicardianBello} addressed the situation in Niger.
 
@@
 Hence, global warming should be beneficial in high latitudes but harmful in low latitudes.
-The estimated loss of the Africa as continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}.
+The estimated loss of the Africa as a continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}.
 \cite{kabubo2007} have shown that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
@@
 \cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa.
-They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
+They argued that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
 \cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania.
@@
 \cite{Kabubo2015} have estimated effects of climate change on food security in Kenya.
-They found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped.
+They have found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped.
 Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
-\cite{Mendelsohn2008} reviewed several studies on effects of climate change on agriculture in developing countries and based on the revision he confirmed that tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture.
+\cite{Mendelsohn2008} has reviewed several studies on effects of climate change on agriculture in developing countries.
+According to the literature review, the tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture \citep{Mendelsohn2008}.
 \cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia.
 The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions.
-According to another meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming.
+According to a meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming.
 However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.
@@
 \cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses.
-Furthermore, risk assessment is essential for disaster risk management \citep{IPCC2012ch1}.
-In the light of these arguments, the main goal of the present study is to develop a model for assessing the risk of drought based on weather and climate data.
-This should help to improve focus on proactive rather than reactive approach in drought disaster risk management.
+Also \cite{IPCC2012ch1} have argued that risk assessment is essential for disaster risk management.
 
-In this study, we focus on Kenya utilizing an yearly panel dataset of $47$ counties over the period of $1981-2017$.
-As a measure of food security we use agricultural maize yields and as a measure of drought we use precipitation and temperature data.
+In the light of these arguments, the goal of the present study is to develop a model which will utilize weather data to assess risk of drought and food security.
+In addition, we want to find out which particular features of climate or weather are the most important factors affecting the yields and food security.
+Are the average weather conditions the most important or is it the weather variability or length and number of dry spells during the growing season what matters the most?
+Answering these and other similar questions should help to improve food security and shift the focus from reactive to proactive approach in drought disaster risk management.
+We focus on Kenya utilizing an yearly panel dataset of $47$ counties over the period of $1981-2017$.
 
-To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
-Hence, this is the first time that effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country.
+The main contribution of this study is utilization of the most detailed publicly available weather datasets while accounting for the real life crop calendar as practised by Kenyan farmers.
+Furthermore, we believe that this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
+Hence, this is the first time that effects of drought and weather on food security have been estimated taking into account the crop calendar, spatially-heterogeneous nature of Kenya and allowing for the effects of weather to vary across the country.
 To the best of our knowledge, this is also the first study which has utilised as comprehensive and detailed weather data  ...resolution...source

### 2018-11-14 writing/draft1/draftMonika2.tex
adding

--- 2018-11-14 writing/draft1/draftMonika1.tex
@@
 We focus on Kenya utilizing an yearly panel dataset of $47$ counties over the period of $1981-2017$.
+% mozna hodit tuhle vetu na zacatek 2.1.
+Measures of food security and drough
 
-The main contribution of this study is utilization of the most detailed publicly available weather datasets while accounting for the real life crop calendar as practised by Kenyan farmers.
-Furthermore, we believe that this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
-Hence, this is the first time that effects of drought and weather on food security have been estimated taking into account the crop calendar, spatially-heterogeneous nature of Kenya and allowing for the effects of weather to vary across the country.
-To the best of our knowledge, this is also the first study which has utilised as comprehensive and detailed weather data  ...resolution...source
+The main contribution of this study is utilization of the most detailed publicly available weather datasets while accounting for the real life crop calendar and different cropping seasons across the country as practised by Kenyan farmers.
+Furthermore, we believe that this is the first study which has described application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
+Hence, this is the first time that effects of drought and weather on food security have been estimated taking into account the real crop calendar, spatially-heterogeneous nature of Kenya and allowing for the effects of weather to vary across the country.
 % Kabubo-Mariara and Karanja (2015) similar study> but they use the more econometric approach (fixed effects models and random effects models as referred to by econometricians
@@
 questions: `How should we measure food security?' and `How should we measure drought?' Taking into account the aim of this study, previous literature (see Section~\ref{Introduction}) and data availability, we were considering three main approaches to measuring food security.
-In particular, we considered using health and utility indicators, food prices or agricultural crop yields as measures of food security.
-Although some health and utility indicators including the Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (see \citealp{CSI}) or the NDMA early warning phases are available in the NDMA Early Warning Bulletins, the data are only available since July $2013$.
-Furthermore, the scales of indicators do not seem to be consistent over the counties and time and the values of indicators or even entire reports are frequently missing.
-Therefore, we decided not to use the health and utility indicators.
+In particular, we considered health and utility indicators, food prices and agricultural crop yields as measures of food security.
+Although some health and utility indicators including the Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (see \citealp{CSI}) or the NDMA early warning phases are available in the NDMA Early Warning Bulletins, these data are only available since July $2013$.
+The other problem is that the scale of indicators does not seem to be consistent over the counties and time.
+Furthermore, the values of indicators are often missing and for some months the reports are not present at all.
 
 Using food prices as a measure of food security turned out be problematic as well. 
-One problem is data availability as the food prices are usually available for market towns rather than at county level.
+One problem is data availability as the food prices are usually available for market towns but not as county averages.
 Furthermore, the market data which are available in the publications of \cite{KNBS} (KNBS) only cover several years.
-Another problem is that prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless the model explicitly accounts for the spatial autocorrelation structure.
+Another problem is that prices are usually strongly spatially autocorrelated and this could cause the regression estimates to be inconsistent and inefficient unless the model accounts for the spatial autocorrelation structure.
 Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and also additional data, which may not be available.
@@
 For an extensive overview of drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
-\cite{keyantash2002} have quantified, evaluated and compared number of drought indices for meteorological, hydrological and agricultural forms of drought.
+\cite{keyantash2002} have quantified, evaluated and compared number of drought indices used for measuring of severity of meteorological, hydrological and agricultural types of drought.
 For more details about types, measures and indices of drought see Appendix~$1$.
@@
 In the recent period, remote sensing data have been used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}.
-Based on these data, various measures such as the vegetation condition index (VCI) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.
+Based on these data, various measures (for example the vegetation condition index) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.
 
-We were considering to use the Standardised Precipitation Evapotranspiration Index  (SPEI) or VCI as measures of drought in our analysis.
+We have considered the Standardised Precipitation Evapotranspiration Index  (SPEI) and the vegetation condition index (VCI) as measures of drought in our analysis.
 However, we did not find the county level SPEI to be as good predictor of food security as other precipitation and temperature measures and computing the SPEI for the entire high resolution weather dataset would be inadequately demanding.
@@
 
-Thus, after conducting a rigorous literature research on measures of drought, we decided to utilize the raw daily precipitation data from...CHIRPS and...resolution...
+Thus, after conducting a rigorous literature research on measures of drought previously used in the relevant context, we decided to utilize various aggregates of daily precipitation and temeprature.
 The frequency of the maize yield data is yearly while the weather data are daily.
-Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
-There are many possible ways how to aggregate daily weather data.
-Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
-\cite{Adejuwon2004} analysed relationship of crop yields and three measures of precipitation.
-The measures of precipitation include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August).
+Hence, the weather data needed to be aggregated in order to obtain a dataset conformable with the yield data.
+There are many possible ways of aggregation.
+Commonly used measures include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
+For example, \cite{Adejuwon2004} has analysed relationship of crop yields and three measures of precipitation in Sub-Saharan West Africa.
+These include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August).
 Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa \citep{Adejuwon2004}.
-Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010}) or annual totals or means \citep{BLIGNAUT2009}.
+Other studies have utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010}) or annual totals or means \citep{BLIGNAUT2009}.
+\cite{BenMohamed2002} have found sea surface temperature anomalies at various locations and amount of rainfall in July, August and September to be significant for millet crops in Niger.
+The authors have also considered the maximum air temperature in the hottest month (April) and the minimum air temperature in the coldest month (January) as possible predictors of crops in Niger, but they did not find them significant.
 Some studies are based on simulated daily extremes, averages, daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}.
-Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
-The definition of wet and dry days and rainy season vary across the literature.
-For example, \cite{BenMohamed2002} has assumed that rainy season begins when the amount of rainfall in three consecutive days reach at least 25mm and no dry spell of more than seven days occurs in the following thirty days.
-According to this study, the end of the rainy season is defined as that rainy day after which rain recorded during 20 days is less than 5mm.
-\cite{BenMohamed2002} has also found sea surface temperature anomalies at various locations and amount of rainfall in July, August and September to be significant for millet crops in Niger.
-The author has also considered the maximum air temperature in the hottest month (April) and the minimum air temperature in the coldest month (January) as possible predictors of crops in Niger, but he did not find them significant.
+Other measures used to explain variability of maize yield include counts of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
+Definitions of wet and dry days and rainy season vary.
+For example, \cite{BenMohamed2002} have assumed that rainy season begins when the amount of rainfall in three consecutive days reaches at least 25~mm and no dry spell of more than seven days occurs in the following thirty days.
+According to this study, the end of the rainy season is defined as that rainy day after which rain recorded during 20 days is less than 5 mm.
 
@@
 \textcolor{red}{Erin Lentz, can we cite her paper?She has not responded to my email}
+
+\textcolor{blue}{maybe following into data section??}
 
@@
 \subsection{Data}
-
+\textcolor{blue}{source of data, resolution and plots/barplots of yield, precip and temp when final data available.}\\
+\textcolor{red}{maybe a table of descriptive statistics or ref to the tale to appendix}
+\textcolor{blue}{also how the weather data were aggregated spatially over counties}
 \subsection{Linear mixed models}

### 2018-11-14 writing/draft1/draftMonika3.tex
adding

--- 2018-11-14 writing/draft1/draftMonika2.tex
@@
 Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset.
-These types of models include both fixed affects and random effects.
+These types of models include both fixed effects parameters and random effects.
 Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}.
@@
 There are at least three benefits of treating a set of parameters as a random sample from some distribution.
-\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9,KERYch12}).
+\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9, KERYch12}).
 
@@
   \\
-\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.066$&$0.008^{**}$\\
+\vspace{-0.2cm}Precipitation total&$0.066$&$0.008^{**}$\\
   \\
-  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$-0.006$&$0.861$\\
+  \vspace{-0.2cm}Precipitation total sq.&$-0.006$&$0.861$\\
   \\
-  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.036$&$0.292$\\
+  \vspace{-0.2cm}Temperature - average&$-0.036$&$0.292$\\
   \\
-    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.081$&$0.008^{**}$\\
+    \vspace{-0.2cm}Precipitation $\times$ temperature& $-0.081$&$0.008^{**}$\\
   \\
 
-      \vspace{-0.2cm}Prec. max OND, east&$0.081$&$0.056^{\bullet}$\\
+      \vspace{-0.2cm}Prec. c. of variation&$0.081$&$0.056^{\bullet}$\\
   \\
-        \vspace{-0.2cm}Prec. max OND, west&$0.108$&$0.009^{**}$\\
-  \\
-    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.101$&$0.003^{**}$\\
-  \\
-      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, west&$0.142$&$6\times10^{-7}$ $^{***}$\\
+        \vspace{-0.2cm}Temp. std. d.&$0.108$&$0.009^{**}$\\
   \\
@@
   \vspace{-0.2cm}Intercept\\
+  \hline
 \\
-  \vspace{-0.2cm}Prec. cum. MAM+OND lag 1\\
-  \\
-  \vspace{-0.2cm}Prec. max OND\\
-  \\
-    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1\\
-  \vspace{-0.1cm} \\ 
-  \hline
-
 \end{tabular} 

### 2018-11-14 writing/draft1/draftMonika4.tex
adding

--- 2018-11-14 writing/draft1/draftMonika3.tex
@@
 \usepackage{mathrsfs}
+\usepackage{gensymb}
 
@@
 \begin{threeparttable}
-
 \singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather}}
-% Peggy45ln can be found in foodSystems/Rcodes/Lags/..
- % Yield in tonnes per hectare
-\label{Peggy45ln} 
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(0,0) errors}}
+% AmBEST00, AmBEST, AmBEST99
+\label{Models1} 
 \centering
 \begin{small}
-\begin{tabular}{lrl} 
+\begin{tabular}{lccc} 
 \hline \vspace{-0.2cm} \\
@@
 
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{1}{c}{\textit{Estimate}} &\multicolumn{1}{c}{\textit{p-value}}\\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{1}{c}{\textit{Model 1}} &\multicolumn{1}{c}{\textit{Model 2}} &\multicolumn{1}{c}{\textit{Model 3}}\\
  \hline 
@@
 \\
-\vspace{-0.2cm}Intercept&$0.158$&$0.086^{\bullet}$\\
+\vspace{-0.2cm}Intercept&$0.298$ $(0.066)^{***}$&$0.256$ $(0.061)^{***}$&$0.243$ $(0.076)^{**}$\\
   \\
-\vspace{-0.2cm}Precipitation total&$0.066$&$0.008^{**}$\\
+\vspace{-0.2cm}Precipitation total&$0.289$ $(0.045)^{***}$&$0.131$ $(0.046)^{**}$&$0.137$ $(0.125)^{}$\\
   \\
-  \vspace{-0.2cm}Precipitation total sq.&$-0.006$&$0.861$\\
+  \vspace{-0.2cm}Precipitation total sq.&$-0.071$ $(0.028)^{*}$&$-0.075$ $(0.026)^{**}$&$-0.073$ $(0.039)^{\bullet}$\\
   \\
-  \vspace{-0.2cm}Temperature - average&$-0.036$&$0.292$\\
+  \vspace{-0.2cm}Temperature - average&$-0.202$ $(0.062)^{**}$&$-0.188$ $(0.058)^{**}$&$-0.464$ $(0.144)^{**}$\\
   \\
-    \vspace{-0.2cm}Precipitation $\times$ temperature& $-0.081$&$0.008^{**}$\\
+    \vspace{-0.2cm}Precip. $\times$ temperature&---&$-0.135$ $(0.037)^{***}$&$-0.108$ $(0.051)^{*}$\\
   \\
-
-      \vspace{-0.2cm}Prec. c. of variation&$0.081$&$0.056^{\bullet}$\\
+      \vspace{-0.2cm}Prec. c. of variation&---&$-0.198$ $(0.035)^{***}$&$-0.181$ $(0.075)^{*}$\\
   \\
-        \vspace{-0.2cm}Temp. std. d.&$0.108$&$0.009^{**}$\\
+        \vspace{-0.2cm}Temp. std. dev.&---&$0.064$ $(0.021)^{**}$&$0.058$ $(0.027)^{*}$\\
+  \\
+          \vspace{-0.2cm}Precip. first 2 months of season&---&---&$-0.003$ $(0.040)^{}$\\
+  \\
+         \vspace{-0.2cm}Dry spell - max length&---&---&$-0.050$ $(0.031)^{}$\\
+  \\
+        \vspace{-0.2cm}Dry spells $>9$ days&---&---&$0.018$ $(0.032)^{}$\\
+  \\
+          \vspace{-0.2cm}Max precipitation&---&---&$-0.008$ $(0.064)^{}$\\
+  \\
+          \vspace{-0.2cm}Degree days ($10-30$\degree C)&---&---&$-0.016$ $(0.092)^{}$\\
+  \\
+          \vspace{-0.2cm}Heat wave days ($>35$\degree C)&---&---&$0.032$ $(0.039)^{}$\\
+  \\
+          \vspace{-0.2cm}Max temperature&---&---&$0.278$ $(0.136)^{*}$\\
   \\
@@
   \vspace{-0.2cm}Intercept\\
+  \vspace{-0.1cm} \\ 
   \hline
-\\
+    \vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{AIC:}}&$2114.98$&$2144.88$&$2118.15$\\
+
+\hline
+\vspace{-0.2cm}
 \end{tabular} 
@@
   \begin{footnotesize}
-    \item \textit{Notes:} \hspace{0.05cm}$584$ observations
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
+        \begin{adjustwidth}{1cm}{} 
+    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
+
+{\centering
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(2,1) errors}}
+%AmArE2100, AmArE21
+\label{ARMA21} 
+\centering
+\begin{small}
+\begin{tabular}{lccc} 
+\hline \vspace{-0.2cm} \\
+
+\vspace{-0.2cm} \\
+
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{1}{c}{\textit{Model 1}} &\multicolumn{1}{c}{\textit{Model 2}} &\multicolumn{1}{c}{\textit{Model 3}}\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$0.299$ $(0.064)^{***}$&$0.240$ $(0.064)^{***}$&$0.270$ $(0.074)^{***}$\\
+  \\
+\vspace{-0.2cm}Precipitation total&$0.273$ $(0.042)^{***}$&$0.159$ $(0.051)^{**}$&$0.219$ $(0.117)^{\bullet}$\\
+  \\
+  \vspace{-0.2cm}Precipitation total sq.&$-0.066$ $(0.026)^{*}$&$-0.079$ $(0.028)^{**}$&$-0.094$ $(0.037)^{*}$\\
+  \\
+  \vspace{-0.2cm}Temperature - average&$-0.177$ $(0.061)^{**}$&$-0.191$ $(0.062)^{**}$&$-0.380$ $(0.133)^{**}$\\
+  \\
+    \vspace{-0.2cm}Precip. $\times$ temperature&---&$-0.135$ $(0.037)^{**}$&$-0.101$ $(0.049)^{*}$\\
+  \\
+      \vspace{-0.2cm}Prec. c. of variation&---&$-0.139$ $(0.040)^{***}$&$-0.091$ $(0.070)^{}$\\
+  \\
+        \vspace{-0.2cm}Temp. std. dev.&---&$0.038$ $(0.024)^{}$&$0.038$ $(0.026)^{}$\\
+  \\
+          \vspace{-0.2cm}Precip. first 2 months of season&---&---&$-0.009$ $(0.037)^{}$\\
+  \\
+         \vspace{-0.2cm}Dry spell - max length&---&---&$-0.045$ $(0.030)^{}$\\
+  \\
+        \vspace{-0.2cm}Dry spells $>9$ days&---&---&$0.008$ $(0.031)^{}$\\
+  \\
+          \vspace{-0.2cm}Max precipitation&---&---&$-0.035$ $(0.060)^{}$\\
+  \\
+          \vspace{-0.2cm}Degree days ($10-30$\degree C)&---&---&$-0.015$ $(0.088)^{}$\\
+  \\
+          \vspace{-0.2cm}Heat wave days ($>35$\degree C)&---&---&$0.040$ $(0.037)^{}$\\
+  \\
+          \vspace{-0.2cm}Max temperature&---&---&$0.207$ $(0.123)^{\bullet}$\\
+  \\
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+  \vspace{-0.1cm} \\ 
+  \hline
+  \vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{AIC:}}&$2064.54$&$2064.99$&$2099.90$\\
+
+\hline
+\vspace{-0.2cm}
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
         \begin{adjustwidth}{1cm}{} 

### 2018-11-14 writing/draft1/draftMonika5.tex
adding

--- 2018-11-14 writing/draft1/draftMonika3.tex
@@
 
-{\centering
+{\hspace{-1.5cm}
 \begin{threeparttable}
@@
 \singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather}}
-% Peggy45ln can be found in foodSystems/Rcodes/Lags/..
- % Yield in tonnes per hectare
-\label{Peggy45ln} 
-\centering
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(0,0) errors}}
+% AmBEST00, AmBEST, AmBEST99
+
+\label{Models1}
+
 \begin{small}
-\begin{tabular}{lrl} 
+\begin{tabular}{lcrcrcr}
+
 \hline \vspace{-0.2cm} \\
-
 \vspace{-0.2cm} \\
-
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{1}{c}{\textit{Estimate}} &\multicolumn{1}{c}{\textit{p-value}}\\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{Model 1}} &\multicolumn{2}{c}{\textit{Model 2}} &\multicolumn{2}{c}{\textit{Model 3}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&Estimate&Sum sq.&Estimate&Sum sq.&Estimate&Sum sq.\\
  \hline 
@@
 \\
-\vspace{-0.2cm}Intercept&$0.158$&$0.086^{\bullet}$\\
+\vspace{-0.2cm}Intercept&$0.298$ $(0.066)^{***}$&---&$0.256$ $(0.061)^{***}$&---&$0.243$ $(0.076)^{**}$&---\\
   \\
-\vspace{-0.2cm}Precipitation total&$0.066$&$0.008^{**}$\\
+\vspace{-0.2cm}Prec. total&$0.289$ $(0.045)^{***}$&$11.9$&$0.131$ $(0.046)^{**}$&$1.3$&$0.137$ $(0.125)^{}$&$0.3$\\
   \\
-  \vspace{-0.2cm}Precipitation total sq.&$-0.006$&$0.861$\\
+  \vspace{-0.2cm}Prec. total sq.&$-0.071$ $(0.028)^{*}$&$1.8$&$-0.075$ $(0.026)^{**}$&$1.5$&$-0.073$ $(0.039)^{\bullet}$&$1.0$\\
   \\
-  \vspace{-0.2cm}Temperature - average&$-0.036$&$0.292$\\
+  \vspace{-0.2cm}Temp. - average&$-0.202$ $(0.062)^{**}$&$3.1$&$-0.188$ $(0.058)^{**}$&$2.9$&$-0.464$ $(0.144)^{**}$&$2.9$\\
   \\
-    \vspace{-0.2cm}Precipitation $\times$ temperature& $-0.081$&$0.008^{**}$\\
+    \vspace{-0.2cm}Prec. $\times$ temp.&---&---&$-0.135$ $(0.037)^{***}$&$2.5$&$-0.108$ $(0.051)^{*}$&$1.3$\\
   \\
-
-      \vspace{-0.2cm}Prec. c. of variation&$0.081$&$0.056^{\bullet}$\\
+      \vspace{-0.2cm}Prec. c. of variation&---&---&$-0.198$ $(0.035)^{***}$&$7.6$&$-0.181$ $(0.075)^{*}$&$1.7$\\
   \\
-        \vspace{-0.2cm}Temp. std. d.&$0.108$&$0.009^{**}$\\
+        \vspace{-0.2cm}Temp. std. dev.&---&---&$0.064$ $(0.021)^{**}$&$1.3$&$0.058$ $(0.027)^{*}$&$1.3$\\
+  \\
+          \vspace{-0.2cm}Prec. first 2 months&---&---&---&---&$-0.003$ $(0.040)^{}$&$0.001$\\
+  \\
+         \vspace{-0.2cm}Dry spell -length&---&---&---&---&$-0.050$ $(0.031)^{}$&$0.7$\\
+  \\
+        \vspace{-0.2cm}Dry spells $>9$ days&---&---&---&---&$0.018$ $(0.032)^{}$&$0.09$\\
+  \\
+          \vspace{-0.2cm}Max precipitation&---&---&---&---&$-0.008$ $(0.064)^{}$&$0.005$\\
+  \\
+          \vspace{-0.2cm}Degree days&---&---&---&---&$-0.016$ $(0.092)^{}$&$0.009$\\
+  \\
+          \vspace{-0.2cm}Heat wave d. ($>35$\degree C)&---&---&---&---&$0.032$ $(0.039)^{}$&$0.2$\\
+  \\
+          \vspace{-0.2cm}Max temperature&---&---&---&---&$0.278$ $(0.136)^{*}$&$1.2$\\
   \\
@@
   \vspace{-0.2cm}Intercept\\
+  \vspace{-0.1cm} \\ 
   \hline
-\\
+    \vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{AIC:}}&$2114.98$&&$2144.88$&&$2118.15$\\
+
+\hline
+\vspace{-0.2cm}
 \end{tabular} 
@@
   \begin{footnotesize}
-    \item \textit{Notes:} \hspace{0.05cm}$584$ observations
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
+        \begin{adjustwidth}{1cm}{} 
+    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
+
+{\hspace{-1.5cm}
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(2,1) errors}}
+%AmArE2100, AmArE21
+\label{ARMA21} 
+\begin{footnotesize}
+
+\begin{tabular}{lcrcrcr}
+
+\hline \vspace{-0.2cm} \\
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{Model 1}} &\multicolumn{2}{c}{\textit{Model 2}} &\multicolumn{2}{c}{\textit{Model 3}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&Estimate&F-value&Estimate&F-value&Estimate&F-value\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$0.299$ $(0.064)^{***}$&$13.9$&$0.240$ $(0.064)^{***}$&$14.0$&$0.270$ $(0.074)^{***}$&$13.8$\\
+  \\
+\vspace{-0.2cm}Prec. total&$0.273$ $(0.042)^{***}$&$63.6$&$0.159$ $(0.051)^{**}$&$64.7$&$0.219$ $(0.117)^{\bullet}$&$65.2$\\
+  \\
+  \vspace{-0.2cm}Prec. total sq.&$-0.066$ $(0.026)^{*}$&$9.9$&$-0.079$ $(0.028)^{**}$&$10.1$&$-0.094$ $(0.037)^{*}$&$10.2$\\
+  \\
+  \vspace{-0.2cm}Temp. - average&$-0.177$ $(0.061)^{**}$&$8.4$&$-0.191$ $(0.062)^{**}$&$11.8$&$-0.380$ $(0.133)^{**}$&$8.7$\\
+  \\
+    \vspace{-0.2cm}Prec. $\times$ temp.&---&---&$-0.135$ $(0.037)^{**}$&$4.6$&$-0.101$ $(0.049)^{*}$&$7.5$\\
+  \\
+      \vspace{-0.2cm}Prec. c. of var.&---&---&$-0.139$ $(0.040)^{***}$&$11$&$-0.091$ $(0.070)^{}$&$11.1$\\
+  \\
+        \vspace{-0.2cm}Temp. std. dev.&---&---&$0.038$ $(0.024)^{}$&$2.4$&$0.038$ $(0.026)^{}$&$2.4$\\
+  \\
+          \vspace{-0.2cm}Prec. 2 months&---&---&---&---&$-0.009$ $(0.037)^{}$&$0.3$\\
+  \\
+         \vspace{-0.2cm}Dry spell -length&---&---&---&---&$-0.045$ $(0.030)^{}$&$2.2$\\
+  \\
+        \vspace{-0.2cm}Dry spells $>9$ days&---&---&---&---&$0.008$ $(0.031)^{}$&$0.1$\\
+  \\
+          \vspace{-0.2cm}Max precipitation&---&---&---&---&$-0.035$ $(0.060)^{}$&$0.2$\\
+  \\
+          \vspace{-0.2cm}Degree days&---&---&---&---&$-0.015$ $(0.088)^{}$&$0.6$\\
+  \\
+          \vspace{-0.2cm}Heat wave d.&---&---&---&---&$0.040$ $(0.037)^{}$&$3.1$\\
+  \\
+          \vspace{-0.2cm}Max temperature&---&---&---&---&$0.207$ $(0.123)^{\bullet}$&$2.8$\\
+  \\
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+  \vspace{-0.1cm} \\ 
+  \hline
+  \vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{AIC:}}&$2064.54$&&$2064.99$&&$2099.90$\\
+
+\hline
+\vspace{-0.2cm}
+\end{tabular} 
+\end{footnotesize}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
         \begin{adjustwidth}{1cm}{} 

### 2018-11-14 writing/draft1/manuscript1.tex
adding

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{a4wide}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage{colortbl}
\usepackage{footnote}
\makesavenoteenv{tabular}
\usepackage[flushleft]{threeparttable}
\usepackage[hyphens]{url}
\usepackage{placeins}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{changepage}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{tabularx}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage[labelfont=bf]{caption}
\usepackage{titlesec}
\usepackage{endnotes}
\usepackage{csquotes}
\usepackage{epstopdf}

\date{\normalsize{September 2018}}
\title{\Large \bf What Does the Weather Say about Yields in Kenya}
\author{Monika Novackova and Pedram Rowhani and Martin Todd}
\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}

\parindent 0pt
\parskip 0.5em
\newcommand\starred[1]{\accentset{~~~~~\star}{#1}}

\newcounter{magicrownumbers}
\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}

\begin{document}

\newdateformat{monthyeardate}{%
  \monthname[\THEMONTH], \THEYEAR}

  \interfootnotelinepenalty=10000
 \newcolumntype{d}{D{.}{.}{-1}}
 \newcolumntype{e}{D{+}{\,\pm\,}{6,2}}

\makeatletter
\def\hlinewd#1{%
\noalign{\ifnum0=`}\fi\hrule \@height #1 %
\futurelet\reserved@a\@xhline}
\makeatother

\maketitle
\vfill

\doublespacing

\begin{abstract}
\noindent We explore an unprecedented dataset of almost $6,000$ observations to identify main predictors of climate knowledge, climate risk perception and willingness to pay for climate change mitigation. Among nearly $70$ potential explanatory variables we detect the most important ones using multisplit lasso estimator. Importantly, we test significance of individuals' preferences about time, risk and equity. Our study is innovative as these behavioural characteristics were recorded by including experimental methods into a live sample survey. This unique way of data collection combines advantages of survey and experiments. The most important predictors of environmental attitudes are numeracy, cognitive ability, ideological world-view and inequity aversion. \\
\end{abstract}

\noindent \textbf{JEL classification:} Q54, Q58, D80\\
\noindent \textbf{Keywords:} Climate change, climate knowledge, climate policy, lasso, risk perception, willingness to pay\\

\newpage
\sloppy

\section{Introduction}\label{Introduction}

According to scientific consensus, climate change does exist and it is human caused \citep{oreskes2004, mccarthy2001impacts, Cook2013COnsensus}. However, public opinions about climate change are far away from consensus \citep{leiserowitz2012, Pew2012}. Without taking actions to prevent or mitigate climate change and its consequences, the effects of global warming could be disastrous \citep{IPCC5WG1, IPCC5, Seneviratne2012}.

Public attitudes towards natural hazards and risk perception are important drivers of policy decision making (\citealp{BritishColumbia, bookDisastersPrep}). Whether and how well climate change will be tackled depends on public opinion heavily. But what are the main factors influencing climate change perception and awareness among general population? Large literature examines role of personal characteristics, demographics and behavioural variables in climate change awareness and risk perception by means of survey or experimental methods. Recent survey-based studies include \citet{Leiserowitz2015}, who exploits the Gallup World Poll data and conclude that civic engagement, communication access and education are the most important predictors of climate change awareness while beliefs about causes and perception of local temperature changes are main predictors of climate risk perception. Another example of a survey-based analysis of environmental attitudes is \citet{Morrison2015}, who use ordered logit models to investigate relationship between religion and climate change attitudes and behaviour. They conclude that Buddhists, atheists and agnostics are the most engaged with climate change while Christian literalists are the least engaged. Also \citet{Carlsson2013} use survey data to estimate preferences of redistribution of burden caused by CO\textsubscript{$2$} emissions in China and in the United States. Understandingly, both Chines and Americans prefer rules of redistribution which are less costly for their country. However, these rules differ for the two countries. US respondents prefer the current emissions rule while in China, the historical emissions rule is preferred.

As for papers based on experiments, for instance \citet{Braaten2014} analyses motivation of good feeling from giving in contrast to 'pure altruism' in climate change context. He concludes that 'warm glow' is important for motivating of environmentally friendly behaviour. Another example of a study which uses experimantal methods is \citet{GlenkColombo2013}, who conduct a choice experiment in Scotland. Based on the results, they claim that an outcome related risk is an important attribute in choice of land-based climate change mitigation project.

The above mentioned studies are just a small portion of recent papers focused on climate concerns and climate knowledge. The previous studies of public attitudes and knowledge about environment are usually based either on survey or on experimental methods. Surveys can be conducted over large, representative samples and experiments are powerful tools to infer parameters of utility functions, measures of risk and time preferences or social value orientation (\citealp{ifcher2011, murphy2011SVO, tanaka2010riskTime}). Each of these methods, however, suffers from serious drawbacks. Surveys often lead to hypothetical bias while experimental data tend to be affected by artificial settings and small, non-representative samples which often consist of students. We contribute by overcoming some of these shortcomings as we use a dataset which was created by surveying a large representative sample of respondents in an experimental, interactive and dynamic way. Experiments are usually computer based and their participants respond to various situations on a screen. We replicate this set-up in a live sample survey by including the experimental methods as a part of the survey \citep{SurveyUK}. The experimental set-up covers attitudes towards risk, attitudes towards equity including altruism and time preferences. We will explore effects of these attitudes on stances towards climate change and climate knowledge.

We also investigate influence of other characteristics on environmental attitudes including standard demographic data such as age, sex, race, ethnicity, religion, education, sector, occupation, date of birth, siblings, questions about assets, debts and family income. We further examine role of financial literacy and numeracy and we also investigate preferences regarding government spending and income redistribution, cultural and political ideology and world-view. As there is no general consensus on what are the main determinants of climate change perception and climate literacy, we start the explanatory analysis using a least absolute shrinkage and selection operator (lasso) to select significant predictors from almost $70$ candidates.

We further contribute by showing substantially different results from those of \citet{Newell2015}, who experimentally measure individual discount rates and analyse their role in energy efficiency decisions of US households. They find a negative and significant effect. More specifically, \citet{Newell2015} conclude that willingness to pay (WTP) for annual operating energy cost savings decrease in discount rates. This disagrees with our results as we do not find any evidence of significant effects of individual discount rates on any of our dependent variables, including WTP. We argue that our estimates are more precise than those of~\citet{Newell2015} as our sample size is substantially larger.\footnote{Our estimation samples include between $5659$ and $5749$ respondents while the estimation sample of~\citet{Newell2015} has $879$ observations.} We also cover much broader spectrum of potential predictors and we use more precise estimator, in concrete multisplit lasso with resampling.

Our additional contribution is a partial replication of~\citet{Kahan2015}. Consistently with his results, we find that climate knowledge measured by the 'ordinary climate science intelligence' (OCSI) instrument\footnote{For a more detailed description of the OCSI instrument see Section~\ref{ClimateVars}.} does not depend on measures of personal ideology and cultural world-view as opposed to other, previously used measures of climate knowledge (\citealp{Hamilton2011, Kahan2012, Kellstedt2008}). In accordance with \cite{Kahan2015}, our results show that climate knowledge is positively correlated with numeracy. We also detect association between our measure of climate knowledge and gender.

In accordance with previous research (\citealp{KahanEtAl2012, Kahan2015, Kellstedt2008}) we find that stated climate change risk perception does not increase with numeracy and financial literacy (we use these variables as proxies for ability of analytical reasoning and capacity to make use of quantitative information although they can also be interpreted as tests of the respondents' attentiveness during the survey) as one may intuitively assume \citep{Weber2011}. As a matter of fact, we find that individuals' concerns about climate change  decline as numeracy and financial literacy increase and it is also closely related to respondents' gender and cultural and ideological world-view. This is consistent with previous literature (\citealp{Kahan2015, Kellstedt2008, WHITMARSH2011}). We particularly show that the respondents, who agree with the statement that 'Government should redistribute income from the better off to those who are less well off' (we will further refer to this statement as 'government should redistribute income', 'degree of agreement with income redistribution' or simply 'income redistribution'), which we use as a measure of cultural or ideological world-view, are more likely to take climate change more seriously and have higher WTP for its mitigation than those not agreeing with this statement. Consistently with recent literature (e.g.~\citealp{Kahan2015,KahanEtAl2012,Hamilton2011, HamiltonKeim2009}), we find evidence suggesting that the ideological polarization over climate change is stronger among people who are more proficient in numeracy and comprehension of quantitative information.

We detect other significant predictors of WTP for climate change mitigation by means of gas and electricity tax. These are age, inequity aversion, perception of equality of intergenerational allocation of resources and risk assessment consistency. Expectedly, the respondents who consider themselves to be more affected by climate change than by climate policy have higher WTP than those who feel to be more affected by climate policy. Consistently with previous literature, we find negative and significant effect of age \citep{Hamilton2011, Kellstedt2008, Hayes2001gender}. The impacts of inequity aversion are mixed and WTP is significantly higher for those respondents who think that their income and standards of living are about the same compared to their parents (at the same stage of their lives) and their children (at the same stage of their lives as respondent currently is). We estimate an analogous model for WTP by means of transport fuel duty as a robustness test and the results are very comparable. The estimates are robust.

Perhaps surprisingly, we did not find the behavioural characteristics to be significant predictors of our climate perception or climate knowledge measures. The only exception is inequity aversion, which has significant effects on WTP for climate change mitigation.

The paper proceeds as follows. In Section~\ref{Method} we discuss methods, in particular multisplit lasso and jackknife ordinary least squares (OLS). Section~\ref{Data} describes the dataset used for our analysis and how the important variables were obtained. In Section~\ref{Results} we present and discuss the results. More specifically, Section~\ref{ResKnow} is focused on climate knowledge, in Section~\ref{ResPerc} we present estimates of climate seriousness perception models and Section~\ref{ResWTP} describes models for WTP. In Section~\ref{Robust} we estimate alternative specifications for each dependent variable to verify robustness of our results. We summarise our findings in Section~\ref{Summary} and in Section~\ref{Caveats} we discuss caveats. Section~\ref{Implications} includes policy implications and concluding remarks.

\FloatBarrier
\section{Methodology}\label{Method}

Prior empirical studies detected large number of miscellaneous predictors of climate change knowledge and concerns \citetext{e.g. \citealt*{Leiserowitz2015, Hamilton2011, McCright2010, Morrison2015}}. There is, however, a lack of consensus about which are the most important ones. Since our dataset includes almost $70$ potential predictors, we decided to start with an explanatory regression analysis using a model selection estimator.

Stepwise-like procedures were found to be problematic as it was shown that large portion of selected variables is often noise and the adjusted $R^2$ is biased upwards \citep{Flack1987}. There are also other problems with these methods. For example, a forward stepwise regression selects in each step the predictor having largest absolute correlation with the response $y$, say $x_{j1}$. Then a simple linear regression of $y$ on $x_{j1}$ is performed and a residual vector from this regression is considering to be the new response variable. Then the procedure is repeated and we eventually end up with a set of selected predictors $x_{j1}, x_{j2},...,x_{jk}$ after $k$ steps. This method can, however, eliminate a good predictor in second step if it happens to be correlated with $x_{j1}$.  Furthermore, these methods frequently fail to identify the correct data generating process, even in large samples \citep{Austin2008}. A possible alternative is the best subset selection approach. Given a collection of possible predictors, the best subset approach compares all possible subsets of predictors based on some well-defined objective criterion, usually having the largest adjusted $R^2$. However, besides being excessively computationally demanding, also this method often fails to identify the true predictors \citep{Flack1987}. On the other hand, sparse estimators such as lasso \citep{tibshirani96} are usually more stable than stepwise procedures and they are commonly better in prediction accuracy \citep{statisHighDimData}. Because lasso has been shown to be very powerful for high-dimensional variable selection in general \citep{pValsLasso}, we opt for this estimator.

Using the same notation as \citet{Friedman2010}, our dependent variable is $Y \in \mathbb{R}$ and our vector of explanatory variables is $X \in \mathbb{R}^p$. We assume that the relationship between them can be approximated by a linear regression model $E(Y|X=x) = \beta_0+x^T\beta$. Lasso estimator selects the predictors by setting some of the coefficients $\beta_j$ to be equal to zero.

We consider four distinct models for the four response variables and one additional model as a robustness test. The dependent variables are: $(i)$~Knowledge about climate change $(ii)$~Perceived seriousness of climate change $(iii)$~Perception of effects of climate change policy relatively to effects of climate change and $(iv)$~WTP for climate change mitigation, which we measure by preferred tax rates on gas and electricity. We also estimate an additional model for petrol duty as a robustness test for the WTP model. How we measure the dependent variables is described in Section~\ref{ClimateVars}. The potential predictors included in $x$, which are not the behavioral variables and which were not selected into any model by multisplit lasso are listed in Tables~\ref{PotentialPredictors} and~\ref{PotentialPredictors2} in Appendix~$3$. How we measure the behavioural variables is discussed in Section~\ref{BehaviouralVars} and their descriptive statistics are summarised in Table~\ref{Descriptive} in Appendix~$3$ with the exemption of inequity aversion as this variable is considered as categorical and its frequencies are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. The predictors, which were selected into some model can be found in a table of estimates of the relevant models and their descriptive statistics or frequencies are summarised in Tables~\ref{Descriptive},~\ref{FreqiencyCat},~\ref{FreqiencyCat2},~and~\ref{FreqiencyBin} in Appendix~$3$.

\sloppy
The estimation function can be written as \citep{Friedman2010}:

\begin{equation}\label{Lasso}
\begin{array}{lcll}

 \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \boldsymbol{R}_{\lambda}(\beta_0, \beta)&=&
   \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \bigg[ \frac{1}{2N}\sum_{i=1}^{N}(y_i - \beta_0-x_i^{\intercal}\beta)^2 + \lambda\sum_{j=1}^{p}(|\beta_j|) \bigg], \\
\end{array}
\end{equation}

%https://www.stata.com/support/faqs/statistics/stepwise-regression-problems/

where $y_i$ is the value of one of our four dependent variables for an individual~$i$, $x_i$ includes potential predictors listed in Tables~\ref{PotentialPredictors}~to~\ref{FreqiencyBin} in Appendix~$3$, $N$ is the number of observations and $\lambda \geq 0$ is the penalty parameter. Without loss of generality, we assume that the potential predictors in~\eqref{Lasso} are standardized: ${\sum_{i=1}^{N}x_{ij} =0}$, ${\frac{1}{N}\sum_{i=1}^{N}x_{ij}^2=1}$, for $j=1,...,p$.~\footnote{Both $x_{ij}$ and $y_j$ are standardized automatically in the implementation of the algorithm  we use. However, the estimated coefficients are always returned and presented on the original scale.}

In line with common practice, we compute estimator~\eqref{Lasso} for a series of~$\lambda$ and then we choose a preferred value of~$\lambda$ using cross-validation \citep{statisHighDimData}. In particular, we use a sequence of $100$ values of~$\lambda$ and $10$\nobreakdash-fold cross validation.\footnote{For estimation of lasso~\eqref{Lasso} we use function cv.glmnet in the \textbf{\textsf{R}}~programming system \citep{RRRR} and we use default settings and values of arguments, unless otherwise stated.} We opt for the value of $\lambda$ which is recommended by \citet{Friedman2010} and it is probably the most common choice. More specifically, we use the largest value of $\lambda$  such that the mean cross-validated error (CVM) is still within one standard error of its minimum.\footnote{In case of WTP we use the value of $\lambda$ which minimises the CVM. This value is also  suggested by \citet{Friedman2010}. The only difference from the model estimated using the one standard error based~$\lambda$ is that for the latter, a dummy variable for male becomes significant and gets into the model. The effect of male is positive and this contradicts predominant conclusions in previous relevant literature \citetext{e.g. \citealt*{Hamilton2011, McCright2010, HamiltonKeim2009, Flynn1994}}.}

Determining significance levels is problematic with lasso. Classical \textit{p}-values are not valid and there is no simple approximation. Therefore, we adopt a concept of \citet{pValsLasso}, who introduce an approach based on multiple random splits of data, repeated estimation and aggregated inference. In particular, \citet{pValsLasso} build on the proposal of \citet{Wasserman2009}, who suggest to split the dataset randomly into two subsets. One of the subsets is used for variable selection via lasso and the other one is for estimating OLS with the predictors selected by lasso and calculating their \textit{p}-values in a usual way. This procedure allows asymptotic error control under minimal conditions. The problem is that the results depend on a one-time arbitrary split and they are therefore irreproducible. \citet{pValsLasso} further develop the single-split method. They suggest to split the sample repeatedly, obtain a set of \textit{p}-values for each split and then aggregate them. In each split, the \textit{p}-values of the variables which are not selected are considered to be equal to one and the \textit{p}-values of the selected variables are multiplied by the number of variables selected in the current split. If a \textit{p}-value multiplied by the number of selected variables happens to be larger than one, it is considered to be equal to one. Let's assume that we have $h=1,...,H$ splits. A \textit{p}-value for predictor $j$ obtained in split $h$ adjusted as described above will be further denoted $P^{(h)}_j$. \citet{pValsLasso} suggest to aggregate the adjusted \textit{p}-values using quantiles. In particular, a suitable aggregated \textit{p}-value is defined for any predictor $j$ and for any fixed $0<\gamma<1$ as

\begin{equation}\label{pvals}
Q_j(\gamma)=min \left \{1,q_{\gamma} (   \{ P^{(h)}_j/\gamma; h=1,...,H\}) \right \},
\end{equation}

 where and $q_{\gamma}(\tiny{\cdot})$ is the (empirical) $\gamma$-quantile function. We will further refer to this procedure as a multisplit lasso.

\citet{pValsLasso} show that for any predefined value of $\gamma \in(0,1)$, the \textit{p}-values defined in~\eqref{pvals} can be used for control of family-wise error rate\footnote{Probability of making at least one incorrect rejection of a true null hypothesis (type 1 error).} and also for regulation of false discovery rate.\footnote{Expected proportion of incorrect rejections of a true null hypothesis (type 1 errors). False discovery rate controlling procedures are less stringent than family-wise error rate controlling methods.} Moreover, the multisplit method improves the power of estimates.

For simplicity, we set $\gamma$ in~\eqref{pvals} to be equal to $0.5$ for every application of a multisplit lasso in this study. Each time we perform $H=100$ splits (we believe that this number is sufficient as \citet{pValsLasso} use $50$ sample splits per simulation) and we always use one third of the sample for the variable selection using lasso and the rest for the OLS estimation and obtaining \textit{p}-values.

A large fraction of our potential predictors are categorical variables because large part of the survey data was collected by multiple choice questions. However, the multisplit lasso selects individual predictors rather than groups of variables. Therefore, it can happen that a model specified by a multisplit lasso includes a dummy variable for one category of a particular categorical variable but it does not include dummy variables for its remaining categories. An obvious way how to overcome this issue would be to add the remaining dummy variables and use an \textit{F}-test to determine the joint significance of the group. If the \textit{F}-test implies that the categories are jointly significant, they should all stay in the model and they should be left out otherwise. However, the solution is not so straightforward with a multisplit lasso as it is not obvious on which subsample we should perform the \textit{F}-test. Therefore, for each model specified by a multisplit lasso, we decided to perform a following procedure which is sometimes called jackknife resampling. We will further refer to the procedure as a jackknife OLS. We again randomly split the dataset into two subsamples. The bigger subsample has size of two thirds of the original sample and it is used for OLS estimation and calculation of \textit{p}-values of the model with predictors selected by multisplit lasso.\footnote{Sample splitting can generally result in loss of efficiency. We, however estimated all models also for the whole sample as a robustness check and the results do not differ in signs or significance levels.} In addition, if the model specified by multisplit lasso includes a binary indicator which represents a category of a nominal variable, we include also all other categories of this variable among the set of predictors. Besides individual \textit{t}-tests we perform an \textit{F}-test of joint significance of the categories of the nominal variable. Similarly as in the case of multisplit lasso, we repeat the resampling and OLS estimation $100$-times. Each time we perform \textit{t}-tests and also a joint \textit{F}-test for each group of dummy variables representing one categorical variable. The \textit{p}-values of the \textit{t}-tests are then aggregated in the same way as in case of multisplit lasso (see above). Further, we calculate mean and median of \textit{p}-values of each joint \textit{F}-test over the $100$ subsamples and according to these statistics we determine whether the dummy indicators of the particular categorical variable should be included. It turns out that every time when a dummy variable representing a category of a nominal variable is chosen by a multiple lasso, both average and median \textit{p}-values of the corresponding \textit{F}-tests are below the significance level (${\alpha=0.05}$). Hence, we include the dummy variables for categories of each nominal variable selected by lasso (see Section~\ref{Results}). % parada. vsechny f testy jsem otestovala u vsech modelu v tabulkach. 27.8. vse v poradku,takze super

\FloatBarrier
\section{Data}\label{Data}
All data used in this study except of predicted income and population density, which we use in robustness tests,  were collected in the survey conducted by \cite{SurveyUK}.

In Section~\ref{Robust} we use an alternative measure of income as a robustness test. In particular, this estimated income is obtained from a regression model based on data from Annual Survey of Hours and Earnings (ASHE). More specifically, the predicted income is based on age, gender, occupation, sector and education.

We use two measures of population density, in particular average density per Lower Layer Super Output Areas (LSOA) estimated by the Office for National Statistics for year $2015$ and average density for Local Authority Districts (LAD) obtained from the $2011$ Census.

The online survey (\citealp{SurveyUK}) ran from $9$ September to $14$
October $2015$ and $6,000$ respondents were selected to answer the questionnaire which included the climate change domain.\footnote{We had to exclude some observations from various parts of analysis as they included missing values for some important variables. However, we have at least $5500$ observations for each model.} Descriptive statistics, methodology of the survey, the survey itself and a detailed description of its administration can be found in~\citet{SurveyUK}.

The survey is reasonably geogrephically representative taking into account population density in the UK \citep{SurveyUK}.\footnote{For map with location of respondents see Figure~$1$ in \cite{SurveyUK}} As the survey was conducted online, the initial sample is representative for UK adults with internet access rather than for the entire UK population.

In Table~\ref{SexAge} we compare distribution of our sample over sex and age  with the distribution of the UK population. The age data are only available as a categorical variable in our survey. As we can see in Table~\ref{SexAge}, the youngest category is slightly over-sampled while the two categories of the highest age are slightly under-sampled, probably because the survey was conducted online. Otherwise the distributions are very comparable.

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Sex and age distribution} \\ of the sample and the population}}
\label{SexAge} 
\begin{small}
\begin{tabular}{|l|rr|rr|} 
\hline	
  \multicolumn{1}{|l|}{ } & \multicolumn{2}{c}{\bf{Sample}} & \multicolumn{2}{|c|}{\bf{UK population\tnote{a}}} \\ %sample pubpolM2 used here
    \multicolumn{1}{|l|}{\vspace{0.1cm}\textbf{Age range}}  &  \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c}{\bf{Female}} & \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c|}{\bf{Female}}\\ 
\hline 
$18-24$&$9.8\%$&$9.4\%$&$6.2\%$&$6.0\%$\\
$25-34$&$10.0\%$&$10.3\%$&$9.0\%$&$9.1\%$\\
$35-44$&$7.8\%$&$8.3\%$&$8.6\%$&$8.8\%$\\
$45-54$&$8.1\%$&$9.4\%$&$9.3\%$&$9.6\%$\\
$55-64$&$7.4\%$&$8.4\%$&$7.5\%$&$7.8\%$\\
$65-74$&$4.1\%$&$4.8\%$&$6.2\%$&$6.7\%$\\
$75-80$&$0.1\%$&$0.1\%$&$2.4\%$&$2.8\%$\\
\hline
\hline
\end{tabular} 
  \begin{tablenotes}
  \begin{footnotesize}
  \singlespacing
     \item[a]Population data are from the Office of National Statistics, Population Estimates of UK, England and
Wales, Scotland and Northern Ireland Mid $2014$, Table MYE2.
\singlespacing
  \end{footnotesize}
\end{tablenotes}
\end{small}
  \end{threeparttable} \par}

\hspace{1.5cm}

Although the survey questionnaire was designed such that more difficult questions were at different pages, we observe that most respondents who did not finish the survey dropped out on pages with more difficult questions. Hence, the final sample is biased towards those who are not afraid of hard questions \citep{SurveyUK}.\footnote{One way how to deal with sample selection is to use sampling weights. We, however decided not use weights given the modest nature of our bias. Weighting usually increases standard errors and leads to less precise estimates and there is lack of consensus on whether or not to use weights in regression methods (\citealp{gelman2007, KottWeight, winship94}). \citet{winship94} for example recommend not to use weights if they are solely a function of independent variables.}

In the rest of this section we focus on how we obtained the data for our climate (dependent) variables and the behavioural characteristics..

\subsection{Climate variables}\label{ClimateVars}

Descriptive statistics of our climate variables are summarised in Table~\ref{DepDes}.

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Dependent variables:} Descriptive statistics}}
\label{DepDes} 
\centering
\begin{small}
\begin{tabular}{lrrrr} 
\hline	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  &  \multicolumn{1}{c}{\bf{Mean}} & \multicolumn{1}{c}{\bf{St. dev.}} & \textbf{Min} & \textbf{Max}\\ 
\hline \vspace{-0.3cm} \\ 
  \vspace{0.15cm}Climate change knowledge&$3.851$&$1.266$&$1$&$8$\\
    \vspace{0.15cm}Climate change seriousness perception&$6.622$&$2.249$&$0$&$10$\\
        \vspace{0.15cm}Climate versus policy effects perception&$5.370$&$2.315$&$0$&$10$\\
\vspace{0.15cm}WTP - gas and electricity tax (\textsterling~per year)&$123.900$&$105.459$&$0$&$500$\\
 \vspace{0.15cm}WTP - duty on transport fuel (pence per year)&$20.530$&$22.518$&$0$&$100$\\
\hline
\hline
\end{tabular} 
\end{small}
  \end{threeparttable} 
\par}

\hspace{1.5cm}

It was previously shown, that questions which are intended to measure climate science comprehension often measure who people are rather than what they know about climate change as the strongest predictor is often respondents' ideology and cultural and political world-view (\citealp{Hamilton2011, KahanEtAl2012, Kahan2015}). To avoid picking of effect of cultural or political world-view instead of climate knowledge, we use questions from the OCSI instrument developed by \citet{Kahan2015} as a measure of climate knowledge. \citet{Kahan2015} shows that these questions are indeed a measure of climate science comprehension rather than an indicator of who one is. The values of climate knowledge are integers from $0$ to $8$ and they stand for counts of correctly answered questions about climate change \citep{Kahan2015}. An example of one of the $8$ climate questions is: 'Climate scientists believe that if the North Pole icecap melted as a result of human-caused global warming, global sea levels would rise. Is this statement true or false?' The list of all climate questions can be found in Appendix~$1$. The relative frequencies of counts of the correctly answered questions are summarised in Table~\ref{DepFreq}.

To investigate opinions about seriousness of climate change, the respondents were asked the following question: 'How serious a problem do you think climate change is at this moment?' Using an interactive slider, the respondents answered an integer value between $0$~and~$10$ where min~$=0$ and max~$=10$ (as it was noted just below the slider). In a similar way, the respondents were asked if they feel to be more affected by climate change or by climate policy. The wording of the question was: 'Which affects you and your way of life more, climate change or policies to reduce greenhouse gas emissions?' Again, the respondents provided answers on an integer scale from $0$ (climate policy) to $10$ (climate change) using a slider. Relative frequencies of climate seriousness perception and climate versus policy perception are summarised in Table~\ref{DepFreq}.

\vspace{1cm}

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Dependent variables:} Relative frequencies (\%)}}
\label{DepFreq} 
\centering
\begin{small}
\begin{tabular}{lrrrrrrrrrrrr} 
\hline 
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  & \multicolumn{1}{c}{\boldsymbol{$0$}}& \multicolumn{1}{c}{\boldsymbol{$1$}}& \multicolumn{1}{c}{\boldsymbol{$2$}}& \multicolumn{1}{c}{\boldsymbol{$3$}}& \multicolumn{1}{c}{\boldsymbol{$4$}}&\multicolumn{1}{c}{\boldsymbol{$5$}}&\multicolumn{1}{c}{\boldsymbol{$6$}}&\multicolumn{1}{c}{\boldsymbol{$7$}}&\multicolumn{1}{c}{\boldsymbol{$8$}}&\multicolumn{1}{c}{\boldsymbol{$9$}}&\multicolumn{1}{c}{\boldsymbol{$10$}}&\\ 
\hline \vspace{-0.3cm} \\ 
  \vspace{0.15cm}Climate knowledge&$0.0$&$1.7$&$11.4$&$30.4$&$25.4$&$20.9$&$8.6$&$1.6$&$0.1$&N/A&N/A\\
\vspace{-0.3cm}Climate seriousness&\multirow{2}{*}{$3.3$}&\multirow{2}{*}{$2.8$}&\multirow{2}{*}{$5.4$}&\multirow{2}{*}{$8.1$}&\multirow{2}{*}{$8.9$}&\multirow{2}{*}{$27.2$}&\multirow{2}{*}{$14.0$}&\multirow{2}{*}{$12.8$}&\multirow{2}{*}{$8.3$}&\multirow{2}{*}{$4.1$}&\multirow{2}{*}{$5.0$}\\
        \vspace{-0.15cm}\hspace{0.5cm}perception&\\
\vspace{-0.3cm}Climate vs. policy&\multirow{2}{*}{$2.1$}&\multirow{2}{*}{$1.5$}&\multirow{2}{*}{$2.5$}&\multirow{2}{*}{$3.8$}&\multirow{2}{*}{$4.6$}&\multirow{2}{*}{$9.8$}&\multirow{2}{*}{$18.5$}&\multirow{2}{*}{$21.7$}&\multirow{2}{*}{$16.7$}&\multirow{2}{*}{$8.6$}&\multirow{2}{*}{$10.4$}\\
\vspace{0.15cm}\hspace{0.5cm}perception\tnote{a}&\\
\hline
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
  \singlespacing
     \item[~]\textit{Notes:} Total number of observations: $5749$
       \begin{adjustwidth}{1cm}{} 
     \vspace{-0.3cm} \item[a]~Higher number means greater concern about climate change, lesser concern about climate policy.
         \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\vspace{1cm}

Regarding the preferred gas and electricity tax rates, the respondents were first asked how much the current tax was. In particular, the question was as follows: 'The average household pays \textsterling$1,369$ per year for gas and electricity. Government intervention has raised the price to encourage people to use less and so reduce greenhouse house gas emissions. How much of that \textsterling$1,369$ is for climate policy?' They indicated the response on a slider with a minimum of $-50$ and a maximum of $500$. We include this variable on right hand site as a robustness test (see Table~\ref{GasInOut}). We refer to it as 'How much is tax gas and electricity'. After this, the respondents were told the correct answer and they were asked about they preferred tax rates: 'Actually, climate policy adds about \textsterling$89$ per year to the gas and electricity bill of the average household. How much do you think climate policy should add to this bill?' The respondents expressed their opinion on a slider from $0$ to $500$. The answer to this question is the dependent variable which we refer to as 'WTP - gas and electricity' and we use it as a proxy for WTP for climate change mitigation. Analogously, we inquired about the fuel duty. The only difference is that the slider for the actual fuel duty is limited from $0$ to $60$ and the one for the preferred fuel duty is from $0$ to $100$ as the actual fuel duty is $3$ pence per litre. Descriptive statistics of the respondents' estimates of actual tax rates can be found in Table~\ref{Descriptive} in Appendix~$3$ and the descriptive statistics of the preferred tax rates are in Table~\ref{DepDes}.

\subsection{Behavioural variables}\label{BehaviouralVars}

One of our goals is to investigate effects of behavioural variables on climate knowledge and concerns about climate change. The behavioural variables that we consider in our study are social value orientation, time preferences, risk preferences, and attitudes towards inequality.

To estimate the social value orientation, respondents played six dictator games with the same questions as in \citet{murphy2011SVO}. The ring measure of social value orientation which we use in our models is defined as

\begin{equation}\label{RingMeasure}
\begin{array}{lcll}

R&=&\arctan \frac{\sum_{i=1}^{N} P_O -50N}{\sum_{i=1}^{N} P_S -50N}, \\
\end{array}
\end{equation}

where $P_O$ is the pay-off given to the other party, $P_S$ is the pay-off taken by the player herself and $N$ is the number of games played (in our case $6$).

As one may notice in Table~\ref{PotentialPredictors} in Appendix~$3$, we also include dummy variables for four types of social value orientation (i.e. altruist, prosocial, individualist, competitive) among potential predictors in the lasso estimators. These types are defined based on ring measure~\eqref{RingMeasure}. Each dummy variable corresponds to one of four non-overlapping intervals of the values of ring measure~\eqref{RingMeasure}. None of these dummy variables was selected by lasso into any of our models, therefore we do not discuss them in more detail.

As a basic measure of time preferences we use derived annual discount rates (in percentage), for investing now for one year from now and we refer to this variable as 'Discount rate year from now' in the present study. To obtain the data which would allow us to infer the discount rates, the respondents played games and answered questions informed by \citet{voors2012violent}, \citet{ifcher2011} and \citet{tanaka2010riskTime}. How the time preferences were derived is described in \cite{SurveyUK}. Besides using discount rates for investing now and getting returns in one year, we also implied other types of discount rates.  These are discount rates for $(i)$~ investing now for getting returns in five years $(ii)$~investing in one year for getting returns in two years from now and $(iii)$~ investing in one year from now for getting return in six years from now. None of them was found to be significant, thus we do not further discuss them.

We use two parameters which describe inequity aversion (\citealp{bergson1938, bergson1954, samuelson1956}), in particular the rate of inequity aversion and the subsistence or reserve income. 
To infer these parameters, respondents were choosing from various distributions of income between three hypothetical people. One of them was higher on average but more unequal and the other was lower on average but more equal. The respondents were asked two sets of choice questions. In one of them, the income distribution was centred on the $70^{th}$ percentile of the UK income distribution and in the other the distribution was centred on $40^{th}$ percentile of the UK income distribution. Given the respondents' answers to the two choice-sets, the rate of inequity aversion and subsistence was obtained for each respondent based on equations~\eqref{InequityEq}:

\begin{equation}\label{InequityEq}
\begin{array}{lcl}

\sum_{i=1}^{3}\frac{(Y^H_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^H_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma} \\

\vspace{0.2cm}\\
\sum_{i=1}^{3}\frac{(Y^L_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^L_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma}
\end{array}
\end{equation}

where $\gamma$ is the rate of inequity aversion, $\underline{Y}$ is the subsistence or reserve income and $Y^H_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $70^{th}$ percentile of the UK income distribution. Analogously, $Y^L_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $40^{th}$ percentile of the UK income distribution. To obtain the inequity parameters, equations~\eqref{InequityEq} were solved for $\gamma$ and $\underline{Y}$ while minimizing distance of $\underline{Y}$ to zero.

In theory, the rate of inequity aversion is a continuous measure. However, we consider it as a categorical one as in our dataset it is equal to one of $16$ distinct values for each respondent.\footnote{We also estimated variants of models where the rate of inequity aversion is considered as scale for completeness but we do not present them to save space. However, the results do not differ substantially from those presented here.} These $16$ values and the corresponding frequencies can be found in Table~\ref{FreqiencyCat2} in Appendix~$3$. The subsistence parameter was not selected by lasso into any of our models thus we do not discuss it in more detail.

To test significance of risk aversion, we use various risk aversion coefficients which were estimated for each person from four different utility functions using Bayesian inference \citep{Balcombe2015}. The utility functions are power, logarithmic, exponential and quadratic and we use the estimates of their means and medians. None of them is significant or chosen by lasso in any of our models. For the economy of space we only present models with median or mean of power function. The estimates are very similar when we use other risk aversion coefficients.

As the behavioural variables are not significant in our study, we described their measures only briefly. For more detailed description see \citet{SurveyUK}. The descriptive statistics of these variables (except of inequity aversion) can be found in Table~\ref{Descriptive} in Appendix~$3$. As explained above, we consider the inequity aversion rate as a categorical variable and its frequencies are in Table~\ref{FreqiencyCat2} in Appendix~$3$.

\FloatBarrier
	\section{Results and discussion}\label{Results}
In this section we describe our results and discuss their interpretation.

In the tables which summarise the estimates of lasso below, \textit{p}-values of some of the explanatory variables are equal to one. These variables were not selected by the lasso in most of the sample splits. They are, however, included in the tables because they represent either a category of a nominal variable whose other category was selected by the lasso or a linear term of a variable whose quadratic term was selected by the lasso.

\subsection{Climate change knowledge}\label{ResKnow}

Table~\ref{Know435} summarizes estimates of the predictors of climate change knowledge which we found to be important by means of multisplit lasso estimator. In particular, three predictors are chosen by lasso (see first column in Table~\ref{Know435}). Total score on financial literacy is the number of correct answers out of three finance related mathematical problems \citep{SurveyUK}.\footnote{We also consider answers to each of the three problems separately as individual potential predictors. Two of them are labelled understands inflation and understands compound interest and they are identified as important predictors in other models later.} However, when we re-estimate the model using jackknife OLS with all relevant dummy variables, all categories of total score on financial literacy are insignificant. Furthermore, the model suffers from multicollinearity as the coefficient of correlation between cognitive reflection and total score of financial literacy is equal to $0.343$ and its \textit{p}-value is smaller than $2\times10^{-8}$. For illustration, estimates of jackknife OLS with all explanatory variables listed in Table~\ref{Know435} including financial literacy are shown in Table~\ref{Know435se} in Appendix~$3$. The last column of Table~\ref{Know435se} includes variance inflation factors (VIF) which confirm the presence of multicollinearity. Because of the multicollinearity and insignificance of total score on financial literacy we do not further consider this variable as a predictor of climate knowledge.

The estimates of jackknife OLS without the financial literacy are summarised in the last two columns of Table~\ref{Know435}. The other two variables which were found to be important in explaining climate knowledge are gender and cognitive reflection test~\citep{Frederick2005}. We use the latter as a measure of numeracy and ability of analytical reasoning.

The cognitive reflection test is fully described in~\citet{Frederick2005} and it consists of three numerical problems. The value of our variable is the number of correct answers out of the three questions.\footnote{The possible values are integers and half-integers between zero and three including zero and three as we also recognise if respondent solves half of a problem. Hence, if a respondent answers for example one and half problems correctly, her score is~$1.5$.} The frequencies of values of this variable are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. To account for plausible non-linear relationship between the test score and cognitive ability we treat the variable as categorical with the base category zero. As it is apparent from Table~\ref{Know435}, the respondents who solved all three problems correctly have significantly higher level of climate knowledge compared to those who did not solve any of them. According to the jackknife OLS, climate knowledge is on average higher also for respondents who answered two problems correctly. However, the effect is larger for three correctly answered problems. Expectedly, the effect of numeracy is positive.

{\centering
\begin{threeparttable}

% table checked 3.10.2017

\singlespacing
\caption{\textit{\textbf{Climate change knowledge:} Multisplit lasso and jackknife OLS}}
% LassoKnow43se3.R      soubor Know435.R 
%musim lasso predelat. predtim jsem pouzivala lambda min. ale alex ma pravdu. je treba u vsech modelu pouzivat stejne pravidlo, takze lambda se.zmena 30.9.2017

%soubor Know435se.R
% finlit2 prestane byt signifikantni, takze vlastne se nic skoro nemeni, asi do prilohy tabulku s Know435se.R

\label{Know435} 
\centering
\begin{small}
\begin{tabular}{lclrcl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Gender = male&$<2\times 10^{-8}$&$^{***}$&$0.733$&$<2\times10^{-8}$&$^{***}$\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=0$~\tnote{a}&$0.038$&$^{*}$ &\multicolumn{3}{c}{\textit{Not included - base cat.}}\\
  \\
\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.000$& &$2.071$&$1.000$&\\
  \\
\vspace{-0.2cm}Cognitive reflection $=1$&$1.000$& &$0.283$&$0.129$& \\
  \\
\vspace{-0.2cm}Cognitive reflection $=1.5$&$1.000$& &$0.895$&$1.000$ \\
  \\
\vspace{-0.2cm}Cognitive reflection $=2$&$1.000$& &$0.628$&$1\times 10^{-5}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2.5$&$1.000$& &$1.098$&$1.000$&\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=3$&$0.046$&$^{*}$&$1.033$&$<2\times10^{-8}$&$^{***}$\\
\\
\vspace{-0.2cm}Financial literacy total score $=0.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\vspace{-0.2cm}Financial literacy total score $=1$&$1.000$&& \multicolumn{3}{c}{\textit{Not included}} \\
  \\
\vspace{-0.2cm}Financial literacy total score $=1.5$&$1.000$&&\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\vspace{-0.2cm}Financial literacy total score $=2$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\vspace{-0.2cm}Financial literacy total score $=2.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
  \\
  \vspace{-0.2cm}Financial literacy total score $=3$&$2\times10^{-5}$&$^{***}$&\multicolumn{3}{c}{\textit{Not included}}\\
  \\
\hline
\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
    \item[a]The estimate is negative for cognitive reflection $=0$ while it is positive for cognitive reflection $=3$ in this model.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\pagebreak

We find the positive and strongly significant effect of dummy variable for males quite peculiar. Previous research shows mixed evidence about effects of gender on climate knowledge and comprehension of science in general. For example, \citet{McCright2010} finds that women demonstrate higher level of scientific knowledge of climate change. On the other hand, \citet{Hayes2001gender} shows that men exhibit significantly higher level of scientific knowledge than women, even if controlling for a number of background variables. We perform additional tests to verify whether the positive effect of gender can be a result of sample selection. The tests include proportion tests, model with interactions as additional explanatory variables and a Heckman selection model. We discuss the results in detail in Appendix~$2$.  Based on the outcomes, we conclude that the results are not driven by sample selection.

A possible explanation why our measure of climate knowledge is significantly higher for men is that the climate knowledge test that we use in this study was developed by a man \citep{Kahan2015}, therefore it may be the case that these particular questions are naturally more comprehensible for men. The only way how to test this would be to let a woman design another set of climate knowledge questions and then conduct a survey which would include these woman-designed climate questions. This is, however, beyond the scope of this study.

 To sum up, we find that gender and cognitive ability are significant predictors of climate knowledge.  Climate knowledge increases with higher numeracy which is consistent with \citet{Kahan2015}, who finds the climate knowledge measure to be positively correlated with ordinary science intelligence. Although various measures of climate knowledge were previously find to be correlated with social ideology or partisan identity (\citealp{Hamilton2011, Kahan2012, Kellstedt2008}), our measures of ideology, cultural world-view or their interactions were not chosen as predictors of climate knowledge by the lasso. This is also consistent with \citet{Kahan2015}.

\subsection{Climate change risk perception}\label{ResPerc}

In this section we discuss our estimates of the models which explain individuals' perception of climate change risk. We focus on two measures of climate risk perception, in particular climate change seriousness perception and climate versus policy perception. We present the results of lasso and jackknife OLS with the climate seriousness perception as dependent variable in Table~\ref{Climcare435}. Three predictors were selected, in particular gender, climate knowledge, and degree of agreement with redistribution of income by government. In this case, the effect of being male is negative. This is mostly consistent with results of previous research which typically finds women to take climate risk more seriously than men (\citealp{WHITMARSH2011, McCright2010, Kahan2007}). As we can see in Table~\ref{Climcare435}, degree of agreement with income redistribution affects climate change seriousness perception positively as the base category is 'Strongly disagree'. This is in agreement with previous literature as we consider the degree of agreement with income redistribution as an indicator of political and ideological world-view, which was found to be significantly correlated with climate concern by large number of previous studies (e.g. \citealp{Leiserowitz2013, Kahan2012, WHITMARSH2011}).

We will comment on the significant effects of climate knowledge at the end of Section~\ref{ResPerc}.

{\centering
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Climate change seriousness perception:} Multisplit lasso and jackknife OLS}}
% soubor LassoClimcare43 Climcare435.R

\label{Climcare435} 
\centering
\begin{small}
\begin{tabular}{lclrcl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Gender = male&$0.0002$&$^{***}$&$-0.3658$&$4.45\times 10^{-6}$&$^{***}$\\
  \\
\vspace{-0.2cm}Climate knowledge&$1.0000$& &$0.1380$&$1.0000$&\\
  \\
\vspace{-0.2cm}Climate knowledge - squared&$<2.00\times 10^{-8}$&$^{***}$&$-0.0548$&$0.0209$&$^{*}$\\
  \\
 Redistribution of income:&\multirow{2}{*}{$1.0000$}& &\multirow{2}{*}{$0.1819$}&\multirow{2}{*}{$1.0000$}&\\%-1
      \hspace{0.6cm}disagree\tnote{a}&& &&&\\%-1
    \\
  \vspace{-0.2cm}Redistribution of income: neutral\tnote{a}&$1.0000$& &$0.2789$&$0.8251$&\\%0
    \\
  \vspace{-0.2cm}Redistribution of income: agree\tnote{a}&$<2.00\times 10^{-8}$&$^{***}$&$0.8343$&$8.58\times 10^{-8}$&$^{***}$\\%1
    \\
  Redistribution of income:& \multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$1.0828$}&\multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
\hspace{0.6cm}strongly agree\tnote{a}&& &&&\\%2
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
   \item[~]\textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
  \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.

 \item[a] Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' The base category is 'Strongly disagree'.
     \end{adjustwidth}
 \singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\hspace{1cm}

We now discuss the estimates of the model with dependent variable which answers the question whether respondent feels to be more affected by climate policy ($0$) or by climate change ($10$). The estimates are shown in Table~\ref{Climpol435}. We can see that the selected predictors are climate knowledge, understanding of inflation and risk assessment consistency.

The level of understanding of inflation is based on respondents' answer to the following numerical problem: 'Imagine that the interest rate on your savings account was 1 percent per year and
inflation was 2 percent per year. After 1 year, would you be able to buy...' The respondents should choose one of the three following answers: $(i)$~'More than today with the money in
this account' $(ii)$~'Exactly the same as today with the money in
this account' $(iii)$~'Less than today with the money in
this account'. Correct answer is 'less'. The value of the variable is equal to one if the respondent answers 'less', it is equal to $0.5$ if she answers 'the same' and it is equal to zero if she answers 'more'. The frequencies of answers are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. Since it is quite possible that the effect of our measure of understanding of inflation is non-linear we treat the variable as categorical with the base category zero.

As it is apparent from Table~\ref{Climpol435}, understanding of inflation and risk assessment consistency increase the likelihood of being subjectively more affected by climate policy than by climate change. This is likely to be because the two predictors are highly correlated with financial literacy. The correlation coefficient of understanding of inflation and financial literacy is $0.694$ and the correlation coefficient of risk assessment consistency and financial literacy is $0.145$. Both correlation coefficients are highly significant with \textit{p}-value lower than $2.00\times 10^{-8}$.~~\footnote{Risk assessment consistency is a binary variable so we also run a two sample \textit{t}-test to measure correlation between risk assessment consistency and financial literacy. In particular, we applied a two sample \textit{t}-test to test if mean financial literacy is statistically equal for the respondent who answered risk questions consistently and for those with inconsistent answers to risk questions. The test statistic is highly significant with \textit{p}-value lower than $2.00\times 10^{-8}$. Hence, the mean financial literacy is different in these two groups which is in accordance with the significance of the correlation coefficient} It is intuitive, that the respondents with higher level of financial literacy are more likely to see how their wealth and way of living can be affected by climate policy through environmental tax rates.

It was previously shown that interactions of measures of cognitive ability and ideological and political world-view are strong predictors of attitudes towards climate change rather than cognitive ability or numeracy itself \citep{Kahan2012, KahanEtAl2012, Hamilton2011, HamiltonKeim2009}. In accordance with this (as we discuss in more detail in Section~\ref{ResWTP} below) we detect a significant impact of interactions of an indicator of political and cultural world-view and a measure of numeracy (and ability of analytical, technical reasoning) on WTP for climate change mitigation. Therefore, we also estimate variants of the models presented in this section with the interaction terms included among the predictors but we did not find them to be significant for climate risk perception. We do not present the results in our study to keep its length within reasonable limits.\footnote{It would probably be more revealing to test for significance of interactions of cognitive ability and political orientation, but unfortunately, the respondents were not asked about their political or partisan preferences directly in the survey.}

\vspace{1.5cm}

{\centering
\begin{threeparttable}
\caption{\small{\textit{\textbf{Climate versus policy effects perception:} Multisplit lasso and jackknife OLS}}}
% soubor LassoClimpol43 Climpol435.R
\label{Climpol435} 
\centering
\begin{small}
\begin{tabular}{lclccl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Climate knowledge&$1.0000$& &$0.2158$&$1.0000$&\\
  \\
\vspace{-0.2cm}Climate knowledge - squared&$<2.00\times 10^{-8}$&$^{***}$&$-0.0607$&$0.0124$&$^{*}$\\
  \\
\vspace{-0.2cm}Understands inflation $=0.5$&$1.0000$& &$-0.0394$&$1.0000$& \\
  \\
\vspace{-0.2cm}Understands inflation $=1$&$0.0130$&$^{*}$&$-0.5759$&$6.27\times 10^{-6}$&$^{***}$\\
    \\
  Consistent answers to risk&\multirow{2}{*}{$1.06\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-0.5885$}&\multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
\hspace{0.6cm}questions $(0/1)$&& &&&\\
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
\begin{tablenotes}
  \begin{footnotesize}
   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
  \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
     \end{adjustwidth}
 \singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\vspace{1cm}

As we can see in Tables~\ref{Climcare435} and~\ref{Climpol435}, the linear climate knowledge term is positive (and insignificant), but the squared term is negative and significant for both climate seriousness and climate change versus policy. That means, the effect of climate knowledge is positive but decreasing for low levels of knowledge, while for medium and high degree of climate knowledge the effect is negative. The negative effect of higher levels of climate knowledge may seem to be counter-intuitive, however it is less surprising in the light of previous literature. According to \citet{reynolds2010now}, people who know less about climate change tend to ascribe unrealistic consequences (for example skin cancer) to global warming. The authors argue that it is even possible for any future ecological or political disaster to be viewed as a consequence of global warming by public \citep{WhatPplKnow94}. Also the level of familiarity with causes and basic mechanisms of climate change is quite unsatisfactory. For example, members of public tend to confuse climate and weather\footnote{Actually, some researchers confuse climate and weather too \citep{Hsiang2013, DesGreenstone2007}.} and as a result they mostly agree with the statement that climate changes from year to year \citep{reynolds2010now}. It is therefore understandable that individuals who are weaker in climate knowledge are more concerned about climate change and its consequences.

Another possible explanation of the negative and significant climate knowledge effect can be affective orientation towards global warming \citep{Kahan2015}. It was previously shown that individuals, who believe that climate change is real and caused by humans and who correctly assert, for instance, that usage of fossil fuels is one of the causes of global warming are also likely to affirm other, perhaps false statements which are consistent with higher environmental risks \citep{reynolds2010now}. An example of such a false statement is that atmospheric emissions of sulphur contribute to global warming \citep{Kahan2015}. The OCSI questions which we use to measure climate knowledge are true/false statements and more than half of them is of a same type as the sulphur emissions statement above. That is, the correct answer does not evince concerns about climate change while the incorrect one does. This could explain why the respondents who believe that climate change is quite serious are likely to score lower on climate knowledge. If this is true, the climate concern variables are predictors of climate knowledge and not the other way around. Therefore, if this is true, climate knowledge should not be included in the specifications with estimates summarised in Tables~\ref{Climcare435} and~\ref{Climpol435}. As a robustness test, we estimate the models for climate seriousness perception and climate change versus climate policy not including the climate knowledge as an explanatory variable. In both cases, the estimates of the rest of the explanatory variables and their significance levels are almost the same as in the case with climate knowledge and they are summarised in Table~\ref{ClimCare435eClimpol435f} in Appendix~$3$.

\subsection{Willingness to pay for climate change mitigation}\label{ResWTP}
This section is focused on models explaining preferred gas and electricity tax rates, which we use as a measure of WTP for climate change mitigation. The estimates of the multisplit lasso and the jackknife OLS are summarised in Table~\ref{Gas435cp}.

One of the important selected predictors is age. The age was recorded as a categorical variable with the lowest category $24$ or younger, the second lowest category is $25-34$, the third one is $35-44$ and so on up to the highest category which is  $75$ or older. We use the lowest age group ($24$ or younger) as the base category. Coefficients of all higher categories are negative and with exemption of $35-44$ and $75$ or older they are all significant. Thus, WTP declines with age, perhaps because older people have lower likelihood of experiencing tougher consequences of climate change predicted for more distant future \citep{Hamilton2011}.\footnote{We also estimated the model with interactions of age and number of children and grandchildren as older people who have more offspring can obviously be more concern about future than those who do not have children. However, we did not find the interactions to be significant. We do not present the results in this study to save space.}

{\centering
\begin{threeparttable}
\caption{\small{\textit{\textbf{WTP climate - gas and electricity tax:} Multisplit lasso and jackknife OLS}}}
% soubor Gas435cp.R, LassoGas47b.R
% soubor Gas435cp2.R, LassoGas47b.R

%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
\label{Gas435cp} 
\centering
\begin{small}
\begin{tabular}{lclccl} 
\hline \vspace{-0.35cm} \\
  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}} %LassoGas47b
  & \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\%Gas435cp

\vspace{-0.35cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.35cm}Age\tnote{a} $~25-34$&$1\times10^{-6}$&$^{***}$&$-13.271$&$0.266$& \\
        \\
\vspace{-0.35cm}Age $35-44$&$0.006$&$^{**}$&$-29.494$&$3\times10^{-7}$&$^{***}$\\
        \\ 
 \vspace{-0.35cm}Age $45-54$&$1.000$& &$-34.625$&$<2\times10^{-8}$&$^{***}$\\
        \\ 
  \vspace{-0.35cm}Age $55-64$&$1.000$& &$-40.061$&$<2\times10^{-8}$&$^{***}$\\
        \\       
 \vspace{-0.35cm}Age $65-74$&$1.000$& &$-46.006$&$<2\times10^{-8}$&$^{***}$\\
        \\ 
 \vspace{-0.35cm}Age$~75$ or older&$1.000$& &$-26.071$&$1.000$& \\
        \\       
 Climate versus policy&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$10.408$}&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
\vspace{-0.35cm}\hspace{0.6cm}effects perception&&&\\
 \\

  \vspace{-0.35cm}Inequity aversion (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{*}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
      \\
 Equal intergenerational&\multirow{2}{*}{$0.011$}&\multirow{2}{*}{${*}$}&\multirow{2}{*}{$20.760$}&\multirow{2}{*}{$0.002$}&\multirow{2}{*}{$^{**}$}\\
\vspace{-0.35cm}\hspace{0.6cm}allocation of resources $(0/1)$\tnote{b}&&&\\
 \\  
 Understands compound&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{ }&\multirow{2}{*}{$-2.942$}&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{ }\\
\vspace{-0.35cm}\hspace{0.6cm}interest $=0.5$&&&\\
 \\   
 Understands compound&\multirow{2}{*}{$1\times10^{-5}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-39.381$}&\multirow{2}{*}{$3\times10^{-5}$}&\multirow{2}{*}{$^{***}$}\\
\vspace{-0.35cm}\hspace{0.6cm}interest $=1$&&&\\
  \\  
\vspace{-0.35cm}Understands inflation $=0.5$&$1.000$& &$-15.892$&$0.516$& \\
  \\
\vspace{-0.35cm}Understands inflation $=1$&$6\times10^{-5}$&$^{***}$&$-42.711$&$<2\times10^{-8}$&$^{***}$\\
  \\
 Consistent answers to risk&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-34.861$}&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
 \hspace{0.6cm}  \vspace{-0.35cm} questions $(0/1)$&& &&&\\
\\ Consistent answers within&\multirow{2}{*}{$0.045$}&\multirow{2}{*}{$^{*}$}&\multicolumn{3}{c}{\textit{Not included}}\\
\hspace{0.6cm}investments $(0/1)$\tnote{c}&& &&&
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}

     \item[~]\textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
\vspace{-0.4cm}  \begin{adjustwidth}{0.65cm}{} \item \vspace{-0.1cm}The signs of the significant coefficients in the multisplit lasso are same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.\vspace{0.05cm}
\item[a]Age is only available as categorical with  base category '$24$ or younger'. Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}).
\vspace{0.05cm}
\item[b]This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age). The variable is equal to $0$ for all other respondents.
\vspace{0.05cm}
\item[c]We eventually excluded this variable from the further analysis. Although consistency within investment was selected by lasso,  it is only marginally significant and strongly correlated with risk assessment consistency. Furthermore, even after exclusion of this variable the model includes relatively large number of predictors and their signs and significance levels do not change.
    \end{adjustwidth}

\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\newpage

We can see in Table~\ref{Gas435cp} that another very strong predictor of WTP is perception of climate versus policy effects on ones way of living which we analyse as a dependent variable in Section~\ref{ResPerc}. Expectedly, those who perceive effects of climate change as more serious than effects of climate policy have higher WPT for climate change mitigation. As we will discus below, this variable is a partial mediator of impact of financial literacy.

Another predictor of WTP is inequity aversion. As we explained in Section~\ref{BehaviouralVars}, we treat this variable as categorical. The values of our inequity aversion measure and their frequencies are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$. Although the effect of inequity aversion is largely positive and decreasing, the signs and significance levels vary across the categories without any clear pattern.

It is further apparent from Table~\ref{Gas435cp} that WTP is higher for respondents who believe that their income and standard of living generally is about equal to the income and standard of living of their parents when they were about the respondent's age and they also believe that their income and standard of living is equal to the income and standard of living of their children when they will reach the respondent's age.

We can also see in Table~\ref{Gas435cp} that WTP for climate change mitigation is significantly lower for individuals with higher financial literacy\footnote{Variable 'Understands inflation' is described in Section~\ref{ResPerc}. Values of variable 'Understands compound interest' are based on respondents answer to the following numerical problem: 'Suppose you had \textsterling $100$ in a savings account and the interest rate was $2$ percent per year.
After $5$ years, how much do you think you would have in the account if you left the money to grow?' The respondents should choose the correct answers from the following three options: $(i)$~'More than \textsterling $102$' $(ii)$~'Exactly \textsterling $102$' $(iii)$~'Less than  \textsterling $102$'. Correct answer is 'more'. The value of the variable is equal to one if the respondent answers 'more', it is equal to $0.5$ if she answers 'the same' and it is equal to zero if she answers 'less'. The frequencies of answers are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.} and for those who answered the questions about risk consistently. This is similar to model with climate versus policy perception as dependent variable (see Table~\ref{Climpol435}). It is very likely, that financial literacy and risk assessment consistency are strongly correlated with ability of analytical reasoning and comprehension of quantitative information, hence, the former can be interpreted as a measure of the latter. Because of complexity of the climate system and inherited difficulty of understanding of climate change by the public these findings may seem to be counterintuitive \citep{Weber2011}. More specifically, one may expect the climate concerns to intensify with increasing level of analytical reasoning and numeracy. Our evidence is, however, consistent with previous literature (\citealp{KahanEtAl2012, Kahan2015, Kellstedt2008}).

\citet{TerrorClimate} and \citet{KahanEtAl2012} argue that the risks related to natural hazards caused by climate change are quite abstract and remote compared to other more salient risks such as terrorism. Hence, it is difficult to perceive the climate change risk as a relatively serious one. It was shown that attitudes towards climate change and related risks are indicators of personal world-view or political outlook rather than correlates of numeracy or science comprehension. People, who identify themselves with egalitarian, communitarian ideology tend to take climate change more seriously than those with rather hierarchical, individualistic world-view \citep{Leiserowitz2013, KahanEtAl2012, WHITMARSH2011}. It can be more important for an individual to consider the climate risk questions from cultural identity perspective than from a scientific and collective knowledge acquisition viewpoint \citep{Kahan2015}. Whether an individual is right or wrong has no meaningful impact on climate change. Decisions of a single consumer or voter can hardly make a measurable difference to the natural hazard risks caused by climate change. On the other hand, adopting a position which is not consistent with one's cultural group can have dangerous consequences \citep{Kahan2012}. \citet{KahanEtAl2012} show that the ideological polarization over climate change is higher among people with the highest degrees of numeracy and science literacy. That means, for the individuals who identify themselves with hierarchical, individualistic ideology, the climate concern is negatively correlated with numeracy and science literacy while for the individuals who believe in rather egalitarian, communitarian ideology the correlation is positive. A possible interpretation is that the members of public with higher degree of numeracy and analytical reasoning are using these abilities to protect their cultural identity and they are therefore better in interpreting the scientific facts in a way which is consistent with their cultural group's ideology. Following \citet{Hamilton2011}, we test this hypothesis by including interaction terms of degree of agreement with redistribution of income and financial literacy among the set of explanatory variables. The estimates are summarised in Table~\ref{Gas435dcp} in Appendix~$3$. The interaction is positive and significant, which means that the positive effect of agreeing with income redistribution is much stronger for those who understand inflation. Similarly, the negative effect of not agreeing with income redistribution is larger in magnitude if accompanied with higher level of understanding of inflation. This is in accordance with the theory that the ideological polarization over climate change is higher among people with higher degrees of numeracy and science literacy (\citealt{Kahan2012, KahanEtAl2012, Hamilton2011, HamiltonKeim2009}). Our results are robust, the estimates and their significance levels are almost the same as those of the model without the interaction term in Table~\ref{Gas435cp}.

One may notice that level of understanding of inflation is a significant predictor of both climate versus policy effects perception and WTP for climate change mitigation. In both cases the effect is negative. The two climate variables are also strongly correlated. Hence, we will now focus on disentangling the structure of relationships among these three variables.

We reveal that the measure of climate versus policy perception partially mediates effect of understanding of inflation on WTP. In Table~\ref{Climpol435} we can see that understanding of inflation is a significant (negative) predictor of climate versus policy perception and in Tables~\ref{Gas435cp} and~\ref{Gas435dcp} we can notice that climate versus policy effects perception is a significant (positive) predictor of WTP. In Table~\ref{MediationTIV} we regress WTP on understanding of inflation without the mediator in order to verify whether the basic condition of mediation is satisfied, i.e. whether we can see the significant effect of the predictor when the mediator is not present. Model~$1$ in Table~\ref{MediationTIV} is a regression of WTP solely on understanding of inflation while Model~$2$ includes also the other predictors selected by the multisplit lasso. Even without the mediator, the effect of understanding of inflation is strongly significant. Furthermore, the effect is larger in magnitude than the effect in the regressions which include the mediator (compare with the estimates in Tables~\ref{Gas435cp} and \ref{Gas435dcp}). This finding also supports the occurrence of mediation. Table~\ref{MediationTMV} summarizes estimates of WTP regressed on the mediator without the effect of understanding of inflation. Model~$1$ in Table~\ref{MediationTMV} only includes climate versus policy as explanatory variable while Model~$2$ in Table~\ref{MediationTMV} also includes the other predictors selected by the multisplit lasso. If the mediation is present, the mediator should also be a significant predictor of the dependent variable itself and we can see in Table~\ref{MediationTMV} that this is true in our case.

As we can see in Tables~\ref{Gas435cp} and~\ref{Gas435dcp}, the effect of understanding of inflation is significant if the mediator is present, therefore the mediation is partial.

To verify our conclusion about the presence of mediation, we perform the Sobel test for the effect of understanding of inflation being mediated through climate versus policy variable. The test statistic is strongly significant with \textit{p}-value equal to $4.21\times10^{-22}$, hence the Sobel test supports the occurrence of mediation.

To sum up, people who understand inflation tend to feel to be more affected by climate policy than by climate change and consequently their WTP for climate change mitigation declines. On the other hand, individuals with lower level of understanding of inflation tend to perceive more effects from climate change than from climate policy and therefore their WTP increases.

\vspace{1cm}
{\centering
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{WTP - mediation through climate versus policy perception:}\\{WTP regressed on financial literacy (understands inflation) without the mediator, OLS}}}
% soubor paper3/Rwork/mediation.R

% soubor paper3/Rwork/mediation2.R
%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation

\label{MediationTIV} 
\centering
\begin{small}
\begin{tabular}{lrcrc} 
\hline
 \multicolumn{1}{l}{\multirow{2}{*}{\textbf{Dependent variable:}}}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}
 &\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
  \multicolumn{1}{l}{WTP-gas and electricity tax (\textsterling~/yr.)} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}} \\
\hline  \vspace{-0.2cm}
  \\
\vspace{-0.2cm}\textbf{Understands inflation} $\boldsymbol{=1}$&${-73.648}$&${<2\times10^{-8}~^{***}}$&${-47.827}$&${<2\times10^{-8}~^{***}}$\\
\\
\vspace{-0.2cm}Understands inflation $=0.5$&$1.912$&$0.712$&$-16.889$&$0.0007~^{***}$\\
\\
\vspace{-0.2cm}Age\tnote{a} $~25-34$&\multicolumn{2}{c}{\textit{Not included}}&$-11.520$&$0.003~^{**}$\\
        \\
\vspace{-0.2cm}Age $35-44$&\multicolumn{2}{c}{\textit{Not included}}&$-27.881$&$<2\times10^{-8}~^{***}$\\
        \\ 
 \vspace{-0.2cm}Age $45-54$&\multicolumn{2}{c}{\textit{Not included}}&$-33.602$&$<2\times10^{-8}~^{***}$\\
        \\ 
  \vspace{-0.2cm}Age $55-64$&\multicolumn{2}{c}{\textit{Not included}}&$-43.684$&$<2\times10^{-8}~^{***}$\\
        \\       
 \vspace{-0.2cm}Age $65-74$&\multicolumn{2}{c}{\textit{Not included}}&$-50.493$&$<2\times10^{-8}~^{***}$\\
        \\ 
 \vspace{-0.2cm}Age$~75$ or older&\multicolumn{2}{c}{\textit{Not included}}&$-33.163$&$0.008~^{**}$\\
        \\       
  \vspace{-0.2cm}Inequity aversion (categorical)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\\
      \\
 Equal intergenerational&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$19.471$}&\multirow{2}{*}{$4\times10^{-6}~^{***}$}\\
\vspace{-0.2cm}\hspace{0.6cm}allocation of resources $(0/1)$&\\
 \\  
\vspace{-0.2cm}Understands compound interest $=0.5$&\multicolumn{2}{c}{\textit{Not included}}&$-5.751$&$0.431$ \\
  \\
\vspace{-0.2cm}Understands compound interest $=1$&\multicolumn{2}{c}{\textit{Not included}}&$-44.022$&$<2\times10^{-8}~^{***}$\\
  \\  
  Consistent answers to risk&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$-39.922$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
\hspace{0.6cm}questions $(0/1)$&& &
  \\
\hline  \vspace{-0.2cm}
  \\
Adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.092$}&\multicolumn{2}{c}{$0.212$}\\
 Observations:&\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5749$}\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
  \begin{adjustwidth}{1cm}{}\item[a]Age is only available as categorical with  base category '$24$ or younger'.
\item[b]Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}).
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\vspace{1cm}

\pagebreak

{\centering
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{WTP - mediation through climate versus policy perception:}\\{WTP regressed on the mediator (climate versus policy effects perception) without variable understands inflation, OLS}}}
% soubor  paper3/Rwork/mediation.R

% soubor paper3/Rwork/mediation2.R
%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
\label{MediationTMV} 
\centering
\begin{small}
\begin{tabular}{lrcrc} 
\hline
 \multicolumn{1}{l}{\multirow{2}{*}{\textbf{Dependent variable:}}}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}
 &\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
  \multicolumn{1}{l}{WTP - gas and electricity tax (\textsterling~/yr.)} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}} \\
\hline  \vspace{-0.2cm}
 \\
\textbf{Climate versus policy}&\multirow{2}{*}{$13.847$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}&\multirow{2}{*}{$10.936$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
\vspace{-0.2cm}\hspace{0.6cm}\textbf{effects perception}&\\
 \\  
\vspace{-0.2cm}Age\tnote{a} $~25-34$&\multicolumn{2}{c}{\textit{Not included}}&$-14.672$&$0.0001~^{***}$\\
        \\
\vspace{-0.2cm}Age $35-44$&\multicolumn{2}{c}{\textit{Not included}}&$-32.393$&$<2\times10^{-8}~^{***}$\\
        \\ 
 \vspace{-0.2cm}Age $45-54$&\multicolumn{2}{c}{\textit{Not included}}&$-40.588$&$<2\times10^{-8}~^{***}$\\
        \\ 
  \vspace{-0.2cm}Age $55-64$&\multicolumn{2}{c}{\textit{Not included}}&$-47.725$&$<2\times10^{-8}~^{***}$\\
        \\       
 \vspace{-0.2cm}Age $65-74$&\multicolumn{2}{c}{\textit{Not included}}&$-54.523$&$<2\times10^{-8}~^{***}$\\
        \\ 
 \vspace{-0.2cm}Age$~75$ or older&\multicolumn{2}{c}{\textit{Not included}}&$-32.341$&$0.008~^{**}$\\
        \\       
  \vspace{-0.2cm}Inequity aversion (categorical)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
      \\
 Equal intergenerational&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$21.644$}&\multirow{2}{*}{$1\times10^{-7}~^{***}$}\\
\vspace{-0.2cm}\hspace{0.6cm}allocation of resources $(0/1)$&\\
 \\  
\vspace{-0.2cm}Understands compound interest $=0.5$&\multicolumn{2}{c}{\textit{Not included}}&$0.396$&$0.956$\\
  \\
\vspace{-0.2cm}Understands compound interest $=1$&\multicolumn{2}{c}{\textit{Not included}}&$-40.066$&$<2\times10^{-8}~^{***}$\\
  \\
  Consistent answers to risk&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$-40.935$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
\hspace{0.6cm}questions $(0/1)$&& &
     \\
\hline  \vspace{-0.2cm}
  \\
Adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.092$}&\multicolumn{2}{c}{$0.244$}\\
 Observations:&\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5749$}\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
  \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
  \begin{adjustwidth}{1cm}{}\item[a]Age is only available as categorical with  base category '$24$ or younger'.
\item[b]Inequity aversion is treated as categorical (see Section~\ref{BehaviouralVars}).
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\newpage

\FloatBarrier

%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT

%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT

\makeatletter 
\renewcommand{\thesection}{\hspace*{-1.0em}}
\newpage
\linespread{1}
\bibliographystyle{ChicagoM}
\bibliography{references}

\newpage

\setcounter{table}{0} 
\makeatletter 
\renewcommand{\thetable}{A\@arabic \c@table} 
\FloatBarrier

\section{Appendix 1 Climate knowledge - OCSI instrument}

Below are the questions of the Ordinary Climate-Science Intelligence (OCSI) assessment developed by \cite{Kahan2015} which we use as a measure of climate knowledge. The eight questions are true or false statements. The correct answers are in bold.

\begin{enumerate}

\item Climate scientists believe that if the North Pole icecap melted as a result of human-caused global warming, global sea levels would rise.  \textbf{FALSE}

\item Climate scientists have concluded
that globally averaged surface air
temperatures were higher for the 
first decade of the twenty-first
century (2000-2009) than for the last decade of the twentieth
century (1990-1999).   \textbf{TRUE}

\item Climate scientists believe that
human-caused global warming
will result in flooding of many
coastal regions. \textbf{TRUE}

\item Climate scientists believe that
human-caused global warming
has increased the number and
severity of hurricanes around the
world in recent decades. \textbf{FALSE}

\item Climate scientists believe that
nuclear power generation
contributes to global warming.  \textbf{FALSE}

\item Climate scientists believe that
human-caused global warming
will increase the risk of skin cancer in human beings.  \textbf{FALSE}

\item Climate scientists and economists
predict there will be positive as
well as negative effects from
human-caused global warming. \textbf{TRUE}

\item Climate scientists believe that the
increase of atmospheric carbon
dioxide associated with the
burning of fossil fuels will reduce
photosynthesis by plants.  \textbf{FALSE}

\end{enumerate}

\section{Appendix 2 Climate knowledge and gender}

As discussed in Section~\ref{ResKnow}, we detect a strong evidence that our measure of climate knowledge is significantly higher for men than for women. We find this outcome merits further investigation.

We hypothesise that less educated men can have higher drop out rates from the survey than more educated men or less educated women. In other words, we believe that it can be more likely for men to abandon the whole survey if they find a series of questions to be too difficult to respond while women answer giving their best guess even if they are uncertain and continue with the survey. This could be caused by different opportunity costs, effect of pride or by males perceiving higher pressure to answer scientific questions correctly. If this is the case, our sample of complete cases will exhibit a selection bias as the ratio of less educated women will be bigger for the subsample of complete cases.

To test for presence of the selection bias, we perform a series of following proportion tests. For each category of education (and also for the whole sample) we test whether the proportion of males in the subsamle of complete observations (used observations) is approximately equal to the proportion of males in the subsample of dropped observations. The \textit{p}-values of the corresponding Pearson's chi-square test statistics of the null hypothesis that the proportions are equal are summarised in Table~\ref{Proptests}.

If the selection bias occurs, we would expect for the lower education categories the proportion of males to be significantly higher among the dropped observations than among the used observations. For the higher categories of education, on the other hand, we would expect the proportion of males to be significantly smaller among the dropped observation than among the used observations. However, this is not what we can see in Table~\ref{Proptests}. Although the proportion tests are significant for some GCSE, GCSE and professional, the differences in proportions are opposite to what we would expect. For the lower categories of education (Some GCSE and GCSE) the proportion of males is smaller among the dropped observations than among the used observations while it is the other way around for the category of professionals. Hence, based on the proportion tests, we do not see any evidence of the selection bias.\footnote{We performed analogous series of tests to compare proportions of males among the dropped observations with the proportions of males in the whole sample. Also these tests were performed separately for each category of education. The results are qualitatively equivalent to those of the tests in Table~\ref{Proptests}. The only difference is that the test became marginally insignificant for the category professional but this has no effect on the conclusion.}

{\centering
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Proportion tests - no selection bias:}\\~ \\ Differences between ratio of males in the group of used observations and in the group of dropped observations. The tests were conducted separately for each category of education.}}

% soubor heckmanNew/droppedDescriptive
\label{Proptests} 
\centering
\begin{small}
\begin{tabular}{l|cccl} 
  \multicolumn{5}{l}{ } \\ \hline
  \multicolumn{1}{l}{ } & \multicolumn{2}{|c}{\textbf{Proportion of males}} &  \multicolumn{2}{c}{\boldsymbol{${\tilde{\chi}^2}$}} \\
   \multicolumn{1}{l}{\textbf{Education category}} &  \multicolumn{1}{|l}{\textbf{Used observations}} &  \multicolumn{1}{c}{\textbf{Dropped observations}} & \multicolumn{2}{c}{\textbf{\textit{p}-value}}\\
 \hline 
\hline
\\
\vspace{-0.2cm}Total&$0.4858$&$0.4538$&$0.0058$&$^{**}$\\
  \\
\vspace{-0.2cm}Craft&$0.6173$&$0.6620$&$0.4671$&\\
  \\
\vspace{-0.2cm}Some GCSE&$0.4957$&$0.4105$&$0.0022$&$^{**}$\\
  \\
\vspace{-0.2cm}GCSE&$0.4861$&$0.3792$&$0.0066$&$^{**}$\\
  \\
\vspace{-0.2cm}A levels&$0.4859$&$0.4671$&$0.5383$&\\
  \\
\vspace{-0.2cm}Diploma&$0.4609$&$0.4172$&$0.2306$&\\
  \\
  \vspace{-0.2cm}Bachelors&$0.4662$&$0.4294$&$0.2470$&\\
\\
  \vspace{-0.2cm}Professional&$0.4269$&$0.5391$&$0.0474$&$^{*}$\\
\\
  \vspace{-0.2cm}Masters&$0.4829$&$0.4000$&$0.1240$&\\
  \\
  \vspace{-0.2cm}PhD&$0.6522$&$0.5625$&$0.4889$&\\
\\
  \vspace{-0.2cm}No answer&$0.5435$&$0.4920$&$0.3355$&\\
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
\item[~]\textit{Notes:} \hspace{0.24cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}
\newpage

\vspace{1cm}

To further eliminate occurrence of the selection bias, we include interactions of gender and the education categories for which the proportion tests are significant as explanatory variables besides the predictors selected by the lasso and we test their significance. For the sake of clearer interpretation, we also include the main effects of the education categories. The estimates of the model with the interactions are summarised in Table~\ref{Know4235f}. None of the interactions or education categories are significant and the signs and significance levels of male and gender are the same as without the interactions.\footnote{We also estimated a version of this model which includes a dummy variable for each education category. Adding these dummy variables does not chance signs or significance levels of gender, cognitive reflection or interactions between gender and education categories.}

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Climate change knowledge:} Jackknife OLS \\ With interactions of gender and education}}
% soubor ResKnow4235f 
\label{Know4235f} 
\centering
\begin{small}
\begin{tabular}{lrcl} 
\hline \vspace{-0.2cm} \\
  \multicolumn{1}{l}{} & \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
   \multicolumn{1}{l}{ } &  \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adjusted \textit{p}-value}} \\
 \hline 
\hline
\\
\vspace{-0.2cm}Gender = male&$0.3379$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.1967$&$1.0000$&\\
  \\
\vspace{-0.2cm}Cognitive reflection $=1$&$0.1195$&$0.4384$& \\
  \\
\vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6160$&$1.0000$ \\
  \\
\vspace{-0.2cm}Cognitive reflection $=2$&$0.2664$&$0.0001$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.4405$&$1.0000$&\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4551$&$2.31\times 10^{-8}$&$^{***}$\\
\\
  \vspace{-0.2cm}Education - some GCSE&$-0.0890$&$1.0000$&\\
\\
  \vspace{-0.2cm}Education - GCSE&$0.0033$&$1.0000$&\\
  \\
  \vspace{-0.2cm}Education -  professional&$-0.0219$&$1.0000$&\\
\\
  \vspace{-0.2cm}Male $\times$ education - some GCSE&$-0.0494$&$1.0000$&\\
\\
  \vspace{-0.2cm}Male $\times$ education - GCSE&$-0.0888$&$1.0000$&\\
\\
  \vspace{-0.2cm}Male $\times$ education -  professional&$0.2377$&$1.0000$&\\
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
  \end{threeparttable} 
\par}
\linespread{1}

\vspace{1.5cm}

As an additional verification that our results can not be attributed to a selection we estimate a Heckman correction models for climate knowledge. In particular, we estimate models which are referred to as Tobit-$2$ \citep{HeckmanR}. The exclusion restriction is count of not responded questions out of those which were prior to climate questions in the survey questionnaire.\footnote{Count of previously not responded questions is not expected to affect climate knowledge. In spite of this, we estimated a variant of multisplit lasso with the count of previously not responded questions as a potential predictor to verify whether or not it should be in the climate knowledge equation according to our estimation method. As we expected, count of previously not responded questions was not selected.} The estimates are summarized in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}. The model in Table~\ref{KnowHeckman1} has only one explanatory variable in the selection equation, namely count of non-responded questions. The variety in Table~\ref{KnowHeckman2} includes also gender, education categories for which the proportion test in Table~\ref{Proptests} is significant and their interactions. The outcome equations include the predictors which were selected by the multisplit lasso (see Section~\ref{ResKnow}). We can see, that the male dummy variable is still positive and strongly significant in the outcome equations even if we correct for possible selection bias (see Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}). Parameter $\rho$ is insignificant in the models in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}. This means that the data are consistent with no correlation of the selection and outcome equation.

We can further see in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2} that achieving score $2$ or $3$ in the cognitive reflection test has positive and significant impact on climate knowledge which is consistent with the model presented in Table~\ref{Know435} in Section~\ref{ResKnow} and with the specification in Table~\ref{Know4235f}. In addition, achieving score $1$ is significant in the Heckman models and score of $0.5$ is close to significant.\footnote{Besides the Heckman selection models presented in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}, we also estimated a version which includes all education categories and their interactions with gender as explanatory variables in the selection equation. However, the estimation algorithm was unable to estimate the coefficients with reasonable standard errors.}

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Climate change knowledge:} Heckman selection model}}
% soubor  N:US/paper3/Rwork/heckmanNew/TheBest2.
% model Claudell12
\label{KnowHeckman1} 
\centering
\begin{small}
\begin{tabular}{lrcl} 
\hline \vspace{-0.3cm} \\
  \vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.25cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Estimate}} &  \multicolumn{2}{c}{\textbf{\textit{p}-value}} \\
 \hline 
\hline
 \vspace{-0.25cm} \\
   \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Probit selection equation:}}\\
   \hline
   \\
\vspace{0.1cm}Not responded questions (count)&$-0.1972$ $(0.0083)$ &$<2.00\times 10^{-8}$&$^{***}$ \\
 \hline
  \vspace{-0.25cm} \\
  \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Outcome equation}}\\
   \hline
      \\
      \vspace{-0.2cm}Gender = male&$0.3304$ $(0.0304)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.2188$ $(0.7202)$&$0.0906$&$^{\bullet}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=1$&$0.1472$ $(0.0387)$&$0.0001$&$^{***}$\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6092$ $(0.8821)$&$0.4898$\\
   \\
\vspace{-0.2cm}Cognitive reflection $=2$&$0.3026$ $(0.0448)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.5158$ $(0.4163)$&$0.2153$&\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4790$ $(0.0559)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
 \hline
  \vspace{-0.25cm} \\
  \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Error terms}}\\
   \hline
      \\
      \vspace{-0.2cm}Sigma $\sigma$&$1.2469$ $(0.0107)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Rho $\rho$&$0.0611$ $(0.0659)$&$0.3540$&\\
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$7244$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
\item[]   \begin{adjustwidth}{1cm}{}Standard errors in brackets
     \end{adjustwidth}
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}
\vspace{1cm}

{\centering \singlespacing
\begin{threeparttable}
\caption{\textit{\textbf{Climate change knowledge:} Heckman selection model \\ With interactions of gender and education}}
% soubor  N:US/paper3/Rwork/heckmanNew/TheBest2.
% model Felicia12
\label{KnowHeckman2} 
\centering
\begin{small}
\begin{tabular}{lrcl} 
\hline \vspace{-0.3cm} \\
  \vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.25cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Estimate}} &  \multicolumn{2}{c}{\textbf{\textit{p}-value}} \\
 \hline 
\hline
 \vspace{-0.25cm} \\
   \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Probit selection equation:\tnote{a}}}\\
   \hline
   \\
 \vspace{-0.2cm}Gender = male&$-0.0163$ $(0.0954)$ &$0.8642$&\\
  \\
 \vspace{-0.2cm}Not responded questions (count)&$-0.1983$ $(0.0084)$ &$<2.00\times 10^{-8}$&$^{***}$ \\
\\
 \vspace{-0.2cm}Education - some GCSE&$-0.2250$ $(0.1213)$ &$0.0637$&$^{\bullet}$ \\
\\
  \vspace{-0.2cm}Education - GCSE&$0.0783$ $(0.1658)$ &$0.6369$&\\
  \\
  \vspace{-0.2cm}Education -  professional&$-0.2696$ $(0.2379)$ &$0.2570$&\\
\\
  \vspace{-0.2cm}Male $\times$ education - some GCSE&$0.2926$ $(0.1934)$ &$0.1303$&\\
\\
  \vspace{-0.2cm}Male $\times$ education - GCSE&$-0.3454$ $(0.2571)$ &$0.1790$&\\
\\
  \vspace{-0.2cm}Male $\times$ education -  professional&$0.3659$ $(0.3468)$ &$0.2914$&\\
 \\
   \hline  
  \multicolumn{4}{c}{\textbf{Outcome equation}}\\
     \hline
      \\
      \vspace{-0.2cm}Gender = male&$0.3307$ $(0.0304)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.2199$ $(0.7202)$&$0.0903$&$^{\bullet}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=1$&$0.1476$ $(0.0387)$&$0.0001$&$^{***}$\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6102$ $(0.8821)$&$0.4891$\\
   \\
\vspace{-0.2cm}Cognitive reflection $=2$&$0.3032$ $(0.0448)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.5168$ $(0.4163)$&$0.2144$&\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4796$ $(0.0559)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\ \hline  
  \multicolumn{4}{c}{\textbf{Error terms}}\\
   \hline  
      \\
      \vspace{-0.2cm}Sigma $\sigma$&$1.2469$ $(0.0107)$&$<2.00\times 10^{-8}$&$^{***}$\\
  \\
\vspace{-0.2cm}Rho $\rho$&$0.0737$ $(0.0661)$&$0.2640$&\\
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$7244$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
 \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
\item[]   \begin{adjustwidth}{1cm}{}Standard errors in brackets \vspace{0.2cm}
 \item[a]~We also estimated a variety of this model which includes dummy variables for all education categories in the selection equation. They are all insignificant and the signs and significance levels of the other variables are the same.
     \end{adjustwidth}  
 \singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\section{Appendix 3 Tables}

{\centering
\begin{threeparttable}
\caption{\textit{\vspace{-0.3cm}List of considered (but not selected) predictors in multisplit lasso}}
\label{PotentialPredictors} 
\centering
\begin{small}
\begin{tabular}{ll} 
\hline \vspace{-0.25cm} \\	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}}&   \multicolumn{1}{l}{\textbf{Description}}   \\ 
\hline \vspace{-0.3cm} \\ 
\vspace{0.15cm}Religion& $11$ categories including atheist, no religion and prefer not to say\\
\vspace{0.15cm}Race& $8$ categories including prefer not to answer\\
\multirow{3}{*}{Length in UK}&Question: \textit{How long have you been living in the UK?}\\
&Response = $5$ categories: All life ,more than $10$ years, $5-10$ years,\\
\vspace{0.15cm}& $1-5$ years, less than $1$ year\\
\vspace{0.15cm}Occupation&$14$ categories\\
\vspace{0.15cm}Sector&$18$ categories\\
\vspace{0.15cm}Operating system&$7$ categories\\
\multirow{2}{*}{Social value orientation}&Response = $4$ categories: altruist, prosocial,\\
\vspace{0.15cm}& individualist, competitive\\
\vspace{0.15cm}Discount rate $0$ vs. $5$&Annual, $\%$, invest now for five years from now \\
\vspace{0.15cm}Discount rate $1$ vs. $2$&Annual, $\%$, invest a year from now for two years from now \\
\vspace{0.15cm}Discount rate $1$ vs. $6$&Annual, $\%$, invest a year from now for six years from now \\
\vspace{0.15cm}Degree of present bias&Continuous, preferences on time \\
\vspace{0.15cm}Degree of hyperbolicity&Continuous, preferences on time \\
\vspace{0.15cm}Annual discount rate&Continuous, preferences on time \\
\vspace{0.15cm}Subsistence income (reserve)&Continuous, \citet{bergson1954, bergson1938, samuelson1956}\\
\vspace{0.15cm}Altruist&Dummy (0/1)\\
\vspace{0.15cm}Prosocial&Dummy (0/1)\\
\vspace{0.15cm}Individualist&Dummy (0/1)\\
\vspace{0.15cm}Competitive&Dummy (0/1)\\
\vspace{0.15cm}Egalitarian&Dummy (0/1)\\
\vspace{0.15cm}Ineqaverse&Dummy (0/1)\\
\vspace{0.15cm}Longitude&Longitude of survey response. Degrees \\
\vspace{0.15cm}Latitude&Latitude of survey response. Degrees \\
\vspace{0.15cm}Letter&First letter of surname, A=$1$,B=$2$,...\\
\vspace{0.15cm}Siblings&Number of siblings\\
\vspace{0.15cm}Older&Number of older siblings\\
\vspace{0.15cm}Children&Number of children\\
Grandchildren&Number of grandchildren\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Note:} \vspace{-0.35cm} Variables in this table were not selected by multisplit lasso into any model.
      \\    \item[~]\hfill (\textit{continued})
\singlespacing
  \end{footnotesize}
   \end{tablenotes}
  \end{threeparttable} 
\par}

{\centering
\begin{threeparttable}
\caption{\textit{\vspace{-0.3cm}List of considered (but not selected) predictors in multisplit lasso}}
\label{PotentialPredictors2} 
\centering
\begin{small}
\begin{tabular}{ll} 
\hline \vspace{-0.25cm} \\	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}}&   \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Description}}   \\ 
\hline \vspace{-0.35cm}\\ 
\vspace{0.14cm}Handedness& $0$=right, $1$=left \\
\vspace{0.14cm}Time&Time taken to complete survey, in minutes\\
\vspace{0.14cm}Hour&Hour of survey, $24$ categories\\
\vspace{0.14cm}Day of week&$7$ categories\\
\vspace{0.14cm}Day of the month&Day of survey, $1-31$\\
\multirow{2}{*}{Fair share}&\textit{Ordinary working people do not get their fair share of the nation's wealth.}\\
\vspace{0.14cm}&Degree of agreement with the statement above, $5$ categories\\
\multirow{2}{*}{Hard work}&Question: \textit{How important is hard work for getting ahead in life?}\\
\vspace{0.14cm}&Response = $5$ categories, degree of agreement\\
\multirow{3}{*}{Better off parents}&Question: \textit{Compared with your parents when they were about your age, }\\
&\textit{are you better or worse in your income and standard of living generally?}\\
\vspace{0.14cm}&Response = $5$ categories (degree of agreement)  and \textit{Don't know}\\
\multirow{3}{*}{Better off children}&Q: \textit{Compared with you, do you think that your children, when they reach}\\
&\textit{your age, will be better or worse in their income and standard of living}\\
\vspace{0.14cm}& \textit{generally?}Answer =$5$ categories (degree of agreement) and \textit{Don't know}\\
\vspace{0.14cm}Always up&Dummy (0/1), Children better off me and me better off parents\\
\vspace{0.14cm}Always down&Dummy (0/1), Parents better off me and me better off children\\
\vspace{0.14cm}Up then down&Dummy (0/1), Me better off parents and me better off children\\
\vspace{0.14cm}Down then up&Dummy (0/1), Parents better off me and children better off me\\
\vspace{0.14cm}Financial literacy&$3$ financial problems, no. of correct answers, \citet{lusardi2014}\\
\vspace{0.14cm}Understands portfolio&Dummy (0/1), $1=$ understands\\
\vspace{0.14cm}Incoherent dr.&Dummy (0/1), Incoherent answers between investments ($0=$ coherent)\\
\vspace{0.14cm}Primed attitudes&$1=$ priming questions about time, risk, social were asked, $0=$ not\\
\multirow{2}{*}{Prime climate}&$0=$ shown picture of polar bear on melting ice (negative),\\
\vspace{0.14cm}& $1=$ shown picture of people enjoying beach (positive)\\
\vspace{0.14cm}Prime pension&$0=$ picture of troubled old man, $1=$ picture of happy old man\\
\vspace{0.14cm}Prime school&$0=$ picture of unruly kids, $1=$ picture of well-behaved kids\\
\vspace{0.14cm}Prime NHS&$0=$ picture NHS in crisis, $1=$ picture love NHS\\
\vspace{0.14cm}Female $\times$ handed &Interaction female and handedness\\
\vspace{0.14cm}Female $\times$ children &Interaction female and number of children\\
\vspace{0.14cm}Age $\times$ children &Interaction age and number of children\\
\hline
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Note:} \vspace{-0.35cm} Variables in this table were not selected by multisplit lasso into any model.
      \\    \item[~]\hfill (\textit{continued})
\singlespacing
  \end{footnotesize}
   \end{tablenotes}
  \end{threeparttable} 
\par}

%seems that all frequencies and descriptives are of the file pubpolM2 (climate domain)

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Descriptive statistics:} Continuous variables}}
\label{Descriptive} 
\centering
\begin{small}
\begin{tabular}{lrrrr} 
\hline \vspace{-0.15cm} \\	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  &  \multicolumn{1}{c}{\bf{Mean}} & \multicolumn{1}{c}{\bf{St. dev.}} & \textbf{Min} & \textbf{Max}\\ 
\hline \vspace{-0.3cm} \\ 
  \vspace{0.15cm}Income - predicted (\textsterling~per year)&$27729$&$11719.89$&$3611$&$58326$\\
    \vspace{0.15cm}Net assets - total assets minus total debts (\textsterling)&$152542$&$223612.90$&$-400000$&$2500000$\\
        \vspace{0.15cm}Population (per Km\textsuperscript{2}, LSOA\tnote{a}\hspace{0.4cm}level)&$3336$&$2975.38$&$7$&$25280$\\
          \vspace{0.15cm}Population (per Km\textsuperscript{2}, LAD\tnote{b}\hspace{0.4cm}level)&$3193$&$3164.75$&$10$&$13870$\\

  \vspace{0.15cm}How much is tax gas and electricity (\textsterling/yr.)&\multirow{1}{*}{$144.90$}&\multirow{1}{*}{$111.94$}&\multirow{1}{*}{$-50$}&\multirow{1}{*}{$500$}\\

  \vspace{0.15cm}How much is duty transport fuel (pence/yr.)&\multirow{1}{*}{$25.18$}&\multirow{1}{*}{$13.68$}&\multirow{1}{*}{$0$}&\multirow{1}{*}{$60$}\\ 
    \hline 
     \vspace{-0.35cm}    
\\         \multicolumn{5}{c}{ \vspace{0.05cm} {Behavioural variables}} \\
    \hline
             \vspace{-0.25cm}    
\\  
\vspace{0.15cm}Social value orientation (ring measure)&$26.28$&$15.52$&$-16.26$&$83.93$\\
 Annual discount rate,$\%$,&\multirow{2}{*}{$148.7$}&\multirow{2}{*}{$181.81$}&\multirow{2}{*}{$1$}&\multirow{2}{*}{$500$}\\
\vspace{0.15cm} \hspace{0.4cm}invest now for a year from now\tnote{c}\\

Risk aversion - estimated median&\multirow{2}{*}{$0.33$}&\multirow{2}{*}{$0.01$}&\multirow{2}{*}{$0.29$}&\multirow{2}{*}{$0.38$}\\
\vspace{0.15cm} \hspace{0.4cm}of quadratic utility function\\ 
 Risk aversion - estimated median&\multirow{2}{*}{$1.81$}&\multirow{2}{*}{$1.08$}&\multirow{2}{*}{$0.67$}&\multirow{2}{*}{$4.33$}\\
\vspace{0.15cm} \hspace{0.4cm}of log utility function\\ 
 Risk aversion - estimated median&\multirow{2}{*}{$0.42$}&\multirow{2}{*}{$0.07$}&\multirow{2}{*}{$0.33$}&\multirow{2}{*}{$0.57$}\\
\vspace{0.15cm} \hspace{0.4cm}of power utility function\\ 
  Risk aversion - estimated mean&\multirow{2}{*}{$0.74$}&\multirow{2}{*}{$0.26$}&\multirow{2}{*}{$0.33$}&\multirow{2}{*}{$1.07$}\\
\vspace{0.15cm} \hspace{0.4cm}of power utility function\\

\hline
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Notes:} Total number of observations: $8541$\vspace{-0.35cm}
          \begin{adjustwidth}{0.7cm}{}

      \item[a] Lower Layer Super Output Area
   \item[b] Local Authority District
  \item[c]This variable is called \textit{Discount rate year from now} in the tables with regression estimates 
    \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\pagebreak

{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Frequency tables:} Categorical variables}}
\label{FreqiencyCat} 
\centering
\begin{small}
\begin{tabular}{llrr} 
\hline \vspace{-0.15cm} \\	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}}& \multicolumn{1}{c}{\textbf{Category}} &  \multicolumn{1}{c}{\textbf{Frequency}}  &  \multicolumn{1}{c}{\textbf{Ratio}}   \\ 
\hline \vspace{-0.3cm} \\ 
 \hline  \vspace{-0.15cm}\\
  \vspace{0.15cm}Education&Craft&$338$&$0.040$\\
  \vspace{0.15cm}&Some GCSE&$1452$&$0.170$\\
  \vspace{0.15cm}&GCSE A*-C grades&$814$&$0.095$\\
  \vspace{0.15cm}&A Level&$1579$&$0.185$\\
  \vspace{0.15cm}&Diploma&\multirow{1}{*}{$979$}&\multirow{1}{*}{$0.115$}\\
    \vspace{0.15cm}&Bachelor's degree&\multirow{1}{*}{$1523$}&\multirow{1}{*}{$0.178$}\\
  \vspace{0.15cm}&Professional qualifications&\multirow{1}{*}{$457$}&\multirow{1}{*}{$0.054$}\\
  \vspace{0.15cm}&Master's degree&$564$&$0.066$\\
  \vspace{0.15cm}&PhD, DPhil&$124$&$0.015$\\
  \vspace{0.15cm}&Prefer not to say&$434$&$0.051$\\
    \vspace{0.15cm}&NA&$277$&$0.032$\\
  \hline  \vspace{-0.15cm} \\
    \vspace{0.15cm}Household income &$<11000$&$919$&$0.158$\\
 \vspace{0.15cm}   pounds per year&$11000 - 16000$&$675$&$0.116$\\
\vspace{0.15cm}  before tax &$16000-20000$&$539$&$0.093$\\
 \vspace{0.15cm} self reported&$20000-26000$&$757$&$0.130$\\
 \vspace{0.15cm}  &$26000-32000$&$610$&$0.105$\\
 \vspace{0.15cm}  &$32000-39000$&$666$&$0.115$\\
 \vspace{0.15cm}  &$39000-48000$&$522$&$0.090$\\
 \vspace{0.15cm}  &$48000-60000$&$544$&$0.094$\\
 \vspace{0.15cm}  &$60000-81000$&$324$&$0.056$\\
 \vspace{0.15cm}  &$81000-100000$&$119$&$0.021$\\
 \vspace{0.15cm}  &$>100000$&$128$&$0.022$\\
\hline
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Notes:} Total number of observations: $8541$\vspace{-0.35cm} \begin{adjustwidth}{0.7cm}{}  
  \item[~]\hfill (\textit{continued})
    \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\pagebreak

{\centering
\begin{threeparttable}
\caption{\vspace{-0.27cm} \small{\textit{\textbf{Frequency tables:} Categorical variables}}}
\label{FreqiencyCat2} 
\centering
\begin{small}
\begin{tabular}{l|crr} 
\hline \vspace{-0.27cm} \\	
  \multicolumn{1}{l|}{\textbf{Variable}}& \multicolumn{1}{c}{\textbf{Category}} &  \multicolumn{1}{c}{\textbf{Frequency}}  &  \multicolumn{1}{c}{\textbf{Ratio}}   \\ 
\hline \vspace{-0.27cm}\\
  Inequity aversion&$0.520$&$1557$&$0.182$\\
 \hspace{0.15cm}(rate) &$0.950$&$55$&$0.006$\\
 \hspace{0.15cm}\citet{bergson1954, bergson1938}, &$1.000$&$652$&$0.076$\\
 \hspace{0.15cm}\citet{samuelson1956} &$1.135$&$127$&$0.015$\\
   &$1.160$&$364$&$0.043$\\
    &$1.255$&$202$&$0.024$\\
    &$1.290$&$130$&$0.015$\\ 
    &$1.485$&$385$&$0.045$\\
    &$1.490$&$288$&$0.034$\\
    &$1.500$&$96$&$0.011$\\ 
     &$1.510$&$86$&$0.010$\\ 
    &$1.685$&$202$&$0.024$\\
    &$1.765$&$93$&$0.011$\\
    &$2.120$&$226$&$0.026$\\ 
    &$3.640$&$59$&$0.007$\\
    &$3.710$&$2551$&$0.299$\\
    &NA&$1468$&$0.172$\\ 
 \hline  \vspace{-0.27cm}\\
 Degree of agreement&\multicolumn{1}{l}{Strongly disagree}&$555$&$0.065$\\ 
 \hspace{0.15cm}with the statement:&\multicolumn{1}{l}{Disagree}&$1121$&$0.131$\\ 
  \textit{ Government should redistribute}&\multicolumn{1}{l}{Neutral}&$1952$&$0.229$\\ 
  \textit{ income from the better off}&\multicolumn{1}{l}{Agree}&$2256$&$0.264$\\ 
  \textit{ to those who are less well off.}&\multicolumn{1}{l}{Strongly agree}&$1206$&$0.141$\\ 
&\multicolumn{1}{l}{NA}&$1451$&$0.170$\\ 
  \hline  \vspace{-0.27cm}\\

Cognitive reflection test\tnote{a}&\multicolumn{1}{c}{$0.0$}&$4145$&$0.485$\\ 
 \hspace{0.25cm}= numeracy, \citet{Frederick2005} &\multicolumn{1}{c}{$0.5$}&$3$&$0.0004$\\ 
   \hspace{0.25cm}$3$ numerical problems&\multicolumn{1}{c}{$1.0$}&$1519$&$0.178$\\ 
  \hspace{0.25cm}no. of correct answers&\multicolumn{1}{c}{$1.5$}&$2$&$0.0002$\\ 
&\multicolumn{1}{c}{$2.0$}&$1017$&$0.119$\\ 
&\multicolumn{1}{c}{$2.5$}&$9$&$0.001$\\ 
&\multicolumn{1}{c}{$3.0$}&$614$&$0.072$\\ 
&\multicolumn{1}{c}{NA}&$1232$&$0.144$\\

 \hline  \vspace{-0.27cm}\\
Understands compound interest&\multicolumn{1}{c}{$0.0$}&$291$&$0.034$\\ 
  \hspace{0.25cm}$1$ = Understands&\multicolumn{1}{c}{$0.5$}&$718$&$0.084$\\ 
    \hspace{0.25cm}(treated as categorical)&\multicolumn{1}{c}{$1.0$}&$5713$&$0.669$\\ 
&\multicolumn{1}{c}{NA}&$1819$&$0.213$\\ 
 \hline  \vspace{-0.27cm}\\
Understands inflation&\multicolumn{1}{c}{$0.0$}&$941$&$0.110$\\ 
  \hspace{0.25cm}$1$ = Understands&\multicolumn{1}{c}{$0.5$}&$953$&$0.112$\\ 
    \hspace{0.25cm}(treated as categorical)&\multicolumn{1}{c}{$1.0$}&$4206$&$0.492$\\ 
&\multicolumn{1}{c}{NA}&$2441$&$0.286$\\ 
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Notes:}  \textbf{a} This variable is called \textit{Cognitive reflection} in the tables with regression
          \begin{adjustwidth}{0.7cm}{}  
  estimates and it is treated as categorical \hfill $8541$ observations
    \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\pagebreak
{\centering
\begin{threeparttable}
\caption{\textit{\textbf{Frequency tables:} Binary variables}}
\label{FreqiencyBin} 
\centering
\begin{small}
\begin{tabular}{lrrr} 
\hline \vspace{-0.15cm} \\	
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable} }&\multicolumn{1}{l}{ \textbf{Frequency $\boldsymbol{= 1$}}} &\multicolumn{1}{l}{  \textbf{Ratio $\boldsymbol{= 1$}}}  &\multicolumn{1}{c}{\textbf{NA's}} \\ 
\hline \vspace{-0.3cm} \\ 
  \vspace{0.15cm}Gender = male &$4060$&$0.475$&$0$\\

Equal intergenerational allocation&\multirow{2}{*}{$793$}&\multirow{2}{*}{$0.110$}&\multirow{2}{*}{$1339$}\\
\vspace{0.15cm} \hspace{0.5cm}of resources (agree = 1)\tnote{a}& \\
  \vspace{0.15cm}Consistent answers to risk questions (consistent = 1)&$7153$&$0.837$&$0$\\
   \vspace{0.15cm}Consist. answers within investment (consistent = 1)&$981$&$0.115$&$0$\\
\hline
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Notes:} Total number of observations: $8541$\vspace{-0.35cm} \begin{adjustwidth}{0.7cm}{}  
 \item[a]~This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age). The variable is equal to $0$ for all other respondents.
    \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\vspace{2cm}

\pagebreak

{\centering
\begin{threeparttable}
\singlespacing
\caption{\textit{\textbf{Climate knowledge:} Jackknife OLS with total score on financial literacy}}
% table checked 3.10.2017

%musim lasso predelat. predtim jsem pouzivala lambda min. ale alex ma pravdu. je treba u vsech modelu pouzivat stejne pravidlo, takze lambda se.zmena 30.9.2017

%soubor Know435se.R
% finlit2 prestane byt signifikantni, takze vlastne se nic skoro nemeni, asi do prilohy tabulku s Know435se.R

\label{Know435se} 
\centering
\begin{small}
\begin{tabular}{lclrc} 
   \hline

\vspace{-0.2cm} \\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}}&  \multicolumn{1}{c}{\textbf{Aggregated}}  \\
    \multicolumn{1}{l}{ } &  \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} &  \multicolumn{1}{c}{\textbf{VIF}}\\
 \hline 
\hline
\\
\vspace{-0.2cm}Gender = male&$1.580$&$<2\times10^{-8}$&$^{***}$&$1.030$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=0.5$&$4.611$&$1.000$&&$1.002$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=1$&$0.431$&$1.000$&&$1.156$\\
\\
\vspace{-0.2cm}Cognitive reflection $=1.5$&$1.681$&$1.000$&&$1.002$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2$&$1.074$&$0.006$&$^{**}$&$1.212$\\
  \\
\vspace{-0.2cm}Cognitive reflection $=2.5$&$1.881$&$1.000$& &$1.006$\\
  \\
  \vspace{-0.2cm}Cognitive reflection $=3$&$1.999$&$9\times10^{-7}$&$^{***}$&$1.171$\\
\\
\vspace{-0.2cm}Financial literacy - total score $=0.5$&$1.173$&$1.000$& &$1.504$\\
  \\
\vspace{-0.2cm}Financial literacy - total score $=1$&$0.456$& $1.000$& &$3.972$\\
  \\
\vspace{-0.2cm}Financial literacy - total score $=1.5$&$0.527$& $1.000$&&$2.586$\\
  \\
\vspace{-0.2cm}Financial literacy - total score $=2$&$0.454$& $1.000$&&$5.644$\\
  \\
\vspace{-0.2cm}Financial literacy - total score $=2.5$&$0.802$& $1.000$&&$2.435$\\
  \\
  \vspace{-0.2cm}Financial literacy - total score $=3$&$1.433						$& $0.150$&&$6.925$\\
 \\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{4}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
    \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
     \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}
\linespread{1}

\pagebreak

{\centering
\begin{threeparttable}
\caption{\textit{{Climate seriousness and climate versus policy effects perception:} Jackknife OLS without climate knowledge}}
\label{ClimCare435eClimpol435f} 
\centering
\begin{small}
\vspace{-1.2cm}
\begin{tabular}{lcccc} 
\hline
  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Seriousness}}&\multicolumn{2}{c}{\textbf{Climate vs. policy}}\\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
 \hline 
\vspace{-0.23cm}
\\
\vspace{-0.2cm}Gender = male&$-0.481$&$<2\times10^{-8}~^{***}$&\multicolumn{2}{c}{\textit{Not included}}\\
\\
 Redistribution of income:&\multirow{2}{*}{$0.201$}&\multirow{2}{*}{$1.000$}&  \multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}} \\%-1
    \vspace{-0.2cm}   \hspace{0.6cm}disagree\tnote{a}&&&&\\%-1
    \\
  \vspace{-0.2cm}Redistribution of inc.: neutral\tnote{a}&$0.331$&$0.265$ &\multicolumn{2}{c}{\textit{Not included}}\\%0
    \\
  \vspace{-0.2cm}Redistribution of income: agree\tnote{a}&$0.891$&$<2\times10^{-8}~^{***}$&\multicolumn{2}{c}{\textit{Not included}}\\%1
    \\
  Redistribution of income:& \multirow{2}{*}{$1.172$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}& \multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}} \\
\hspace{0.6cm}strongly agree\tnote{a}&&&\\%2
\vspace{-0.2cm}
 \\
\vspace{-0.2cm}Understands inflation $=0.5$&\multicolumn{2}{c}{\textit{Not included}}&$-0.049$&$1.000$ \\
  \\
\vspace{-0.2cm}Understands inflation $=1$  &\multicolumn{2}{c}{\textit{Not included}}&$-0.643$& $2\times10^{-7}~^{***}$\\
    \\
  Consistent answers to risk& \multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}} &\multirow{2}{*}{$-0.592$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
\hspace{0.6cm}questions $(0/1)$&& &\\
   \\
     \hline
  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5749$}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}

 \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
 \begin{adjustwidth}{0.9cm}{}
  \vspace{-0.3cm}
  \item[a]~Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' The base category is 'Strongly disagree'.

     \end{adjustwidth}    
\singlespacing  
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\vspace{1cm}

\pagebreak

{\centering
\begin{threeparttable}
\caption{\textit{\small{\textbf{\vspace{-0.25cm} WTP climate:} interaction of cultural world-view and financial literacy}}}
% soubor  Rwork/GasSuperNew/Gas435dcq.R
\label{Gas435dcp} 
\centering
\begin{small}
\begin{tabular}{lccl} 
\hline \vspace{-0.15cm} \\
  \multicolumn{1}{l}{} 
  & \multicolumn{3}{c}{\large{\textbf{\small{Jackknife OLS}}}}  \\
\\
  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Dependent variable:}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
  \multicolumn{1}{l}{WTP - gas and electricity tax (\textsterling~per year)} & \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\   
\hline \vspace{-0.3cm} \\ 
 \vspace{0.15cm}Age\tnote{a} $~25-34$&$-11.3886$&$1.0000$&\\
  \vspace{0.15cm}Age $35-44$&$-27.1860$&$4\times10^{-5}$&$^{***}$\\
 \vspace{0.15cm}Age $45-54$ &$-34.3795$&$2\times10^{-8}$&$^{***}$\\
 \vspace{0.15cm}Age $55-64$ &$-37.5114$&$<2\times10^{-8}$&$^{***}$\\
 \vspace{0.15cm}Age $65-74$&$-45.2613$&$<2\times10^{-8}$&$^{***}$\\
 \vspace{0.15cm}Age $74$ or older &$-28.8122$&$1.0000$&\\     
 \vspace{0.15cm}Climate versus policy effects perception&$10.2983$&$<2\times10^{-8}$&$^{***}$\\
\vspace{0.15cm}Inequity aversion (categorical)\tnote{b}& \multicolumn{2}{c}{\textit{negative correlation}}&$^{**}$\\
Equal intergenerational&\multirow{2}{*}{$21.8916$}&\multirow{2}{*}{$0.0034$}&\multirow{2}{*}{$^{*}$}\\
 \vspace{0.15cm}\hspace{0.6cm}allocation of resources $(0/1)$\tnote{c}&\\
 \vspace{0.15cm}Understands compound interest $=0.5$&\multirow{1}{*}{$-4.3717$}&\multirow{1}{*}{$1.0000$}&\multirow{1}{*}{ }\\
 \vspace{0.15cm}Understands compound interest $=1$&\multirow{1}{*}{$-41.6926$}&\multirow{1}{*}{$5\times10^{-5}$}&\multirow{1}{*}{$^{***}$}\\
  \vspace{0.15cm}Understands inflation $=0.5$&$-17.6328$&$0.1616$& \\
 \vspace{0.15cm}Understands inflation $=1$&$-46.0609$&$<2\times10^{-8}$&$^{***}$\\

\vspace{0.15cm}Consistent answers to risk questions $(0/1)$&\multirow{1}{*}{$-35.4773$}&\multirow{1}{*}{$<2\times10^{-8}$}&\multirow{1}{*}{$^{***}$}\\
\vspace{0.15cm}Redistribution of income (degree of agreement)\tnote{d, e}&\multirow{1}{*}{$-9.6596$}&\multirow{1}{*}{$0.0670$}&\multirow{1}{*}{$^{\bullet}$}\\

 \vspace{0.15cm}\textbf{Redistribution of income\tnote{d} \hspace{0.1cm}$\boldsymbol{\times}$\hspace{0.1cm}Understands inflation}&$13.7064$&$0.0043$&$^{**}$
\\
\hline
\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$5749$} \\  \vspace{-0.4cm}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
     \item[~]\textit{Notes:} \vspace{-0.35cm} \begin{adjustwidth}{0.7cm}{}  \hspace{0.7cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$;\hspace{0.2cm} mean adj. \textit{R\textsuperscript2}:\hspace{0.25cm}$0.279$
\item[a]~Age is only available as categorical. The base category is '$24$ or younger'.
\item[b]~Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}). \item[c]~This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age). The variable is equal to $0$ for all other respondents.
\vspace{0.002cm} 
 \item[d]~Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' $-2=$~Strongly disagree, $2=$~Strongly agree.
  \item[e]~For the sake of simplicity we also included the main effect.
    \end{adjustwidth}
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

{\centering
\begin{threeparttable}
\singlespacing
\caption{\small\textit{\textbf{Climate vs. policy perception:} Jackknife OLS - robustness without WTP}}
% soubor Cpol435InOutbNoWTP.R  Cpol535InONoWTP.R
\label{ClimPolIOnoWTP} 
\centering
\begin{small}
\vspace{-1.2cm}
\begin{tabular}{lcccc} 
\hline
  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
\cmidrule(lr){2-3} \cmidrule(lr){4-5}
  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
 \hline 
\vspace{-0.23cm}
\\
\vspace{-0.33cm}Climate knowledge\tnote{a}&$-0.130$&$9\times10^{-5}~^{***}$&$-0.131$&$0.0002~^{***}$\\
  \\
\vspace{-0.33cm}Understands inflation $=0.5$&$0.043$&$1.000$&$0.021$&$1.000$ \\
  \\
\vspace{-0.33cm}Understands inflation $=1$&$-0.450$&$0.001~^{**}$&$-0.455$&$0.001~^{**}$\\
    \\
  Consistent answers to risk&\multirow{2}{*}{$-0.496$}&\multirow{2}{*}{$3\times10^{-7}~^{***}$}&\multirow{2}{*}{$-0.482$}&\multirow{2}{*}{$1\times10^{-6}~^{***}$}\\
\hspace{0.6cm}questions $(0/1)$&& &\\
\hline
  \\
\vspace{-0.33cm}Income- predicted (mill. \textsterling~/yr.)&$3.347$&$1.000$&\multicolumn{2}{c}{\textit{Not included}}\\ %income PREDICTED!!
  \\
\vspace{-0.33cm}Income- reported (mill. \textsterling~/yr.)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\textit{varies}&$1.000$\\ %income PREDICTED!!
  \\
\vspace{-0.33cm}Net assets (million \textsterling)&$-0.014$&$1.000$&$0.076$&$1.000$\\ 
      \\
\vspace{-0.33cm}People per mill. km\textsuperscript{2}-LSOA level&\multirow{1}{*}{$-15.303$}&\multirow{1}{*}{$1.000$}&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}\\%ha is a symbol for hectares
  \\
\vspace{-0.33cm}People per mill. km\textsuperscript{2}-LAD level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$20.682$}&\multirow{1}{*}{$1.000$}\\%ha is a symbol for hectares
  \\
    \vspace{-0.33cm}Climate seriousness perception&$0.427$&$\small{<}2\times10^{-8}~^{***}$&$0.432$&$\small{<}2\times10^{-8}~^{***}$\\
      \\
Social value orientation&\multirow{2}{*}{$0.006$}&\multirow{2}{*}{$0.250$}&\multirow{2}{*}{$0.006$}&\multirow{2}{*}{$0.555$}\\
\vspace{-0.33cm} \hspace{0.6cm}(ring measure)&& &\\%2
  \\% ring measure> ref working paper dolton tol
 \vspace{-0.33cm}\multirow{1}{*}{Inequity aversion (categorical)}&\textit{varies}& \multirow{1}{*}{$1.000$}& \multirow{1}{*}{\textit{varies}}& \multirow{1}{*}{$1.000$}\\      
    \\ \vspace{-0.33cm}Discount rate yr. from now&$-0.002$&$1.000$&$-0.001$&$1.000$\\  
      \\
 \vspace{-0.33cm}Discount rate yr. from now - sq.&\multirow{1}{*}{$3\times10^{-6}$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$3\times10^{-6}$}&\multirow{1}{*}{$1.000$}\\
  \\  \vspace{-0.33cm}Risk aversion coefficient\tnote{c}&\multicolumn{2}{c}{\textit{Not included}}&$-0.569$&$1.000$\\          
   \\
  \vspace{-0.35cm}Redistribution of income (categ.)\tnote{d}&\textit{$+$,varies}&$1.000$&\textit{$+$,varies}&$1.000$\\%1
    \\
     \hline
          Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.226$}&\multicolumn{2}{c}{$0.230$}\\
  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
\\
\hline
\end{tabular} 
\end{small}
 \begin{tablenotes}
  \begin{footnotesize}
 \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
 \begin{adjustwidth}{0.9cm}{}
  \vspace{-0.3cm}
 \item[a]~Squared term of climate knowledge is insignificant in this version, hence it is not included.
\item[b]~Self reported income is only available as categorical.
\item[c]~The risk aversion coefficient is an estimated parameter of a utility function. In this model, the mean of power function is used. We also estimated varieties of this model with different risk aversion coefficients, particularly means or medians of various utility functions. These are power, log, exponential and quadratic. The risk aversion parameter is always insignificant and whether it is included or not (or which one) does not affect sign or significance level of any other parameter. 
\item[d]~A degree of agreement with the statement: 'Government should redistribute income from the better off to those who are less well off.' Included to test for significance of political opinions.
     \end{adjustwidth}   
\singlespacing
  \end{footnotesize}
\end{tablenotes}
  \end{threeparttable} 
\par}

\vspace{0.2cm}

\end{document}

### 2018-11-14 writing/draft1/manuscript10.tex
adding

--- 2018-11-14 writing/draft1/manuscript1.tex
@@
 \usepackage{changepage}  
+\usepackage{mathrsfs}
 
@@
 \usepackage{epstopdf}
-
-\date{\normalsize{September 2018}}
-\title{\Large \bf What Does the Weather Say about Yields in Kenya}
-\author{Monika Novackova and Pedram Rowhani and Martin Todd}
+\usepackage{euscript}
+
+\date{\normalsize{October 2018}}
+\title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
+\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens}
 \affil{\small{Department of Geography, University of Sussex, Falmer, UK}}
@@
 
-According to scientific consensus, climate change does exist and it is human caused \citep{oreskes2004, mccarthy2001impacts, Cook2013COnsensus}.
-However, public opinions about climate change are far away from consensus \citep{leiserowitz2012, Pew2012}.
-Without taking actions to prevent or mitigate climate change and its consequences, the effects of global warming could be disastrous \citep{IPCC5WG1, IPCC5, Seneviratne2012}.
-
-Public attitudes towards natural hazards and risk perception are important drivers of policy decision making (\citealp{BritishColumbia, bookDisastersPrep}).
-Whether and how well climate change will be tackled depends on public opinion heavily.
-But what are the main factors influencing climate change perception and awareness among general population?
-Large literature examines role of personal characteristics, demographics and behavioural variables in climate change awareness and risk perception by means of survey or experimental methods.
-Recent survey-based studies include \citet{Leiserowitz2015}, who exploits the Gallup World Poll data and conclude that civic engagement, communication access and education are the most important predictors of climate change awareness while beliefs about causes and perception of local temperature changes are main predictors of climate risk perception.
-Another example of a survey-based analysis of environmental attitudes is \citet{Morrison2015}, who use ordered logit models to investigate relationship between religion and climate change attitudes and behaviour.
-They conclude that Buddhists, atheists and agnostics are the most engaged with climate change while Christian literalists are the least engaged.
-Also \citet{Carlsson2013} use survey data to estimate preferences of redistribution of burden caused by CO\textsubscript{$2$} emissions in China and in the United States.
-Understandingly, both Chines and Americans prefer rules of redistribution which are less costly for their country.
-However, these rules differ for the two countries.
-US respondents prefer the current emissions rule while in China, the historical emissions rule is preferred.
-
-As for papers based on experiments, for instance \citet{Braaten2014} analyses motivation of good feeling from giving in contrast to 'pure altruism' in climate change context.
-He concludes that 'warm glow' is important for motivating of environmentally friendly behaviour.
-Another example of a study which uses experimantal methods is \citet{GlenkColombo2013}, who conduct a choice experiment in Scotland.
-Based on the results, they claim that an outcome related risk is an important attribute in choice of land-based climate change mitigation project.
-
-The above mentioned studies are just a small portion of recent papers focused on climate concerns and climate knowledge.
-The previous studies of public attitudes and knowledge about environment are usually based either on survey or on experimental methods.
-Surveys can be conducted over large, representative samples and experiments are powerful tools to infer parameters of utility functions, measures of risk and time preferences or social value orientation (\citealp{ifcher2011, murphy2011SVO, tanaka2010riskTime}).
-Each of these methods, however, suffers from serious drawbacks.
-Surveys often lead to hypothetical bias while experimental data tend to be affected by artificial settings and small, non-representative samples which often consist of students.
-We contribute by overcoming some of these shortcomings as we use a dataset which was created by surveying a large representative sample of respondents in an experimental, interactive and dynamic way.
-Experiments are usually computer based and their participants respond to various situations on a screen.
-We replicate this set-up in a live sample survey by including the experimental methods as a part of the survey \citep{SurveyUK}.
-The experimental set-up covers attitudes towards risk, attitudes towards equity including altruism and time preferences.
-We will explore effects of these attitudes on stances towards climate change and climate knowledge.
-
-We also investigate influence of other characteristics on environmental attitudes including standard demographic data such as age, sex, race, ethnicity, religion, education, sector, occupation, date of birth, siblings, questions about assets, debts and family income.
-We further examine role of financial literacy and numeracy and we also investigate preferences regarding government spending and income redistribution, cultural and political ideology and world-view.
-As there is no general consensus on what are the main determinants of climate change perception and climate literacy, we start the explanatory analysis using a least absolute shrinkage and selection operator (lasso) to select significant predictors from almost $70$ candidates.
-
-We further contribute by showing substantially different results from those of \citet{Newell2015}, who experimentally measure individual discount rates and analyse their role in energy efficiency decisions of US households.
-They find a negative and significant effect.
-More specifically, \citet{Newell2015} conclude that willingness to pay (WTP) for annual operating energy cost savings decrease in discount rates.
-This disagrees with our results as we do not find any evidence of significant effects of individual discount rates on any of our dependent variables, including WTP.
-We argue that our estimates are more precise than those of~\citet{Newell2015} as our sample size is substantially larger.\footnote{Our estimation samples include between $5659$ and $5749$ respondents while the estimation sample of~\citet{Newell2015} has $879$ observations.} We also cover much broader spectrum of potential predictors and we use more precise estimator, in concrete multisplit lasso with resampling.
-
-Our additional contribution is a partial replication of~\citet{Kahan2015}.
-Consistently with his results, we find that climate knowledge measured by the 'ordinary climate science intelligence' (OCSI) instrument\footnote{For a more detailed description of the OCSI instrument see Section~\ref{ClimateVars}.} does not depend on measures of personal ideology and cultural world-view as opposed to other, previously used measures of climate knowledge (\citealp{Hamilton2011, Kahan2012, Kellstedt2008}).
-In accordance with \cite{Kahan2015}, our results show that climate knowledge is positively correlated with numeracy.
-We also detect association between our measure of climate knowledge and gender.
-
-In accordance with previous research (\citealp{KahanEtAl2012, Kahan2015, Kellstedt2008}) we find that stated climate change risk perception does not increase with numeracy and financial literacy (we use these variables as proxies for ability of analytical reasoning and capacity to make use of quantitative information although they can also be interpreted as tests of the respondents' attentiveness during the survey) as one may intuitively assume \citep{Weber2011}.
-As a matter of fact, we find that individuals' concerns about climate change  decline as numeracy and financial literacy increase and it is also closely related to respondents' gender and cultural and ideological world-view.
-This is consistent with previous literature (\citealp{Kahan2015, Kellstedt2008, WHITMARSH2011}).
-We particularly show that the respondents, who agree with the statement that 'Government should redistribute income from the better off to those who are less well off' (we will further refer to this statement as 'government should redistribute income', 'degree of agreement with income redistribution' or simply 'income redistribution'), which we use as a measure of cultural or ideological world-view, are more likely to take climate change more seriously and have higher WTP for its mitigation than those not agreeing with this statement.
-Consistently with recent literature (e.g.~\citealp{Kahan2015,KahanEtAl2012,Hamilton2011, HamiltonKeim2009}), we find evidence suggesting that the ideological polarization over climate change is stronger among people who are more proficient in numeracy and comprehension of quantitative information.
-
-We detect other significant predictors of WTP for climate change mitigation by means of gas and electricity tax.
-These are age, inequity aversion, perception of equality of intergenerational allocation of resources and risk assessment consistency.
-Expectedly, the respondents who consider themselves to be more affected by climate change than by climate policy have higher WTP than those who feel to be more affected by climate policy.
-Consistently with previous literature, we find negative and significant effect of age \citep{Hamilton2011, Kellstedt2008, Hayes2001gender}.
-The impacts of inequity aversion are mixed and WTP is significantly higher for those respondents who think that their income and standards of living are about the same compared to their parents (at the same stage of their lives) and their children (at the same stage of their lives as respondent currently is).
-We estimate an analogous model for WTP by means of transport fuel duty as a robustness test and the results are very comparable.
-The estimates are robust.
-
-Perhaps surprisingly, we did not find the behavioural characteristics to be significant predictors of our climate perception or climate knowledge measures.
-The only exception is inequity aversion, which has significant effects on WTP for climate change mitigation.
-
-The paper proceeds as follows.
-In Section~\ref{Method} we discuss methods, in particular multisplit lasso and jackknife ordinary least squares (OLS).
-Section~\ref{Data} describes the dataset used for our analysis and how the important variables were obtained.
-In Section~\ref{Results} we present and discuss the results.
-More specifically, Section~\ref{ResKnow} is focused on climate knowledge, in Section~\ref{ResPerc} we present estimates of climate seriousness perception models and Section~\ref{ResWTP} describes models for WTP.
-In Section~\ref{Robust} we estimate alternative specifications for each dependent variable to verify robustness of our results.
-We summarise our findings in Section~\ref{Summary} and in Section~\ref{Caveats} we discuss caveats.
-Section~\ref{Implications} includes policy implications and concluding remarks.
-
+\large Findings:
+\begin{itemize}
+\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important
+
+\item Mar-Sept last year temperature very important for maize yields
+\item SD temperature last year positive and significant
+\item dry spell 20 MAM last year important (but not dry spell MAM10)
+\item interesting. Precipitation 2 months MAM last year very significant and positive
+\item mean temp last year negative and significant, hill shaped 
+\end{itemize}
+
+\underline{New findings:}
+
+\begin{itemize}
+\item The yields seem to be more responsive to weather on west than on east
+\end{itemize}
+\normalsize
 \FloatBarrier
+\pagebreak
+This aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.
+
+\textcolor{red}{something about droughts in Kenya, how they were recently and what were the effects, find refrences..see my folder literatureFS/references SSRP or the book East African Agriculture and Climate in the folder literatureFS} for example \citep{Nicholson2017}
+
+ Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
+During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
+Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.
+\cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.
+
+Recent literature has promoted shifting from reactive to proactive approach in disaster risk reduction.
+Forecast based financing has been promoted as it can avoid significant disaster losses \citep{Nicholson2017}.
+It has been shown that the investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}.
+
+As a measure of food security we use agricultural maize yields and as measures of drought we use variables created by aggregating daily temperature and precipitation data.
+
+A number of previous studies have focused on analysis of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa.
+Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
+\cite{Adejuwon2004} analysed relationship of crop yields and three measures of precipitation.
+The measures of precipitation include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August).
+Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa.
+Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010} \textcolor{red}{and others,find!!}) or annual totals or means \citep{BLIGNAUT2009}.
+Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}.
+Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
+The definition of wet and dry days and rainy season vary across the literature.
+For example, \cite{BenMohamed2002} has assumed that rainy season begins when the amount of rainfall in three consecutive days reach at least 25mm and no dry spell of more than seven days occurs in the following thirty days.
+According to this study, the end of the rainy season is defined as that rainy day after which rain recorded during 20 days is less than 5mm.
+\cite{BenMohamed2002} has also found sea surface temperature anomalies at various locations and amount of rainfall in July, August and September to be significant for millet crops in Niger.
+The author has also considered the maximum air temperature in the hottest month (April) and the minimum air temperature in the coldest month (January) as possible predictors of crops in Niger, but he did not find them significant.
+
+An important group of studies analysing the relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
+
+(\citealt{ThorntonEtAl2009}) \textcolor{red}{also uses diurnal daily}
+
+\textcolor{red}{Erin Lentz, can we cite her paper???}
+
+Blignaut: annual averages
+Chipanschi: daily maximum and minimum temperatures
+Giannoukopoulos: number of hw days and so...
+Laux et al.: daily Tmax >30C
+Lobell et al. 2008: Monthly temp and prec.
+Lobell and Burke 2010: Temp. growing season average, precip:growing season total
+
+(\citealt{AbrahaSavage2006,Adejuwon2004,BenMohamed2002,BLIGNAUT2009,
+Chipanshi2003,Giannakopoulos2009,LauxEtAl,Leemans1993,LobellEtAl2008,
+LobellBurke2010,sagoe2006,SchulzeEtA1993,ThorntonEtAl2009,TingemEtAl2008,
+TingemEtAl2009,WalkerSchulze2008})
+
+\textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
+
 \section{Methodology}\label{Method}
 
-Prior empirical studies detected large number of miscellaneous predictors of climate change knowledge and concerns \citetext{e.g.
-\citealt*{Leiserowitz2015, Hamilton2011, McCright2010, Morrison2015}}.
-There is, however, a lack of consensus about which are the most important ones.
-Since our dataset includes almost $70$ potential predictors, we decided to start with an explanatory regression analysis using a model selection estimator.
-
-Stepwise-like procedures were found to be problematic as it was shown that large portion of selected variables is often noise and the adjusted $R^2$ is biased upwards \citep{Flack1987}.
-There are also other problems with these methods.
-For example, a forward stepwise regression selects in each step the predictor having largest absolute correlation with the response $y$, say $x_{j1}$.
-Then a simple linear regression of $y$ on $x_{j1}$ is performed and a residual vector from this regression is considering to be the new response variable.
-Then the procedure is repeated and we eventually end up with a set of selected predictors $x_{j1}, x_{j2},...,x_{jk}$ after $k$ steps.
-This method can, however, eliminate a good predictor in second step if it happens to be correlated with $x_{j1}$.
-Furthermore, these methods frequently fail to identify the correct data generating process, even in large samples \citep{Austin2008}.
-A possible alternative is the best subset selection approach.
-Given a collection of possible predictors, the best subset approach compares all possible subsets of predictors based on some well-defined objective criterion, usually having the largest adjusted $R^2$.
-However, besides being excessively computationally demanding, also this method often fails to identify the true predictors \citep{Flack1987}.
-On the other hand, sparse estimators such as lasso \citep{tibshirani96} are usually more stable than stepwise procedures and they are commonly better in prediction accuracy \citep{statisHighDimData}.
-Because lasso has been shown to be very powerful for high-dimensional variable selection in general \citep{pValsLasso}, we opt for this estimator.
-
-Using the same notation as \citet{Friedman2010}, our dependent variable is $Y \in \mathbb{R}$ and our vector of explanatory variables is $X \in \mathbb{R}^p$.
-We assume that the relationship between them can be approximated by a linear regression model $E(Y|X=x) = \beta_0+x^T\beta$.
-Lasso estimator selects the predictors by setting some of the coefficients $\beta_j$ to be equal to zero.
-
-We consider four distinct models for the four response variables and one additional model as a robustness test.
-The dependent variables are: $(i)$~Knowledge about climate change $(ii)$~Perceived seriousness of climate change $(iii)$~Perception of effects of climate change policy relatively to effects of climate change and $(iv)$~WTP for climate change mitigation, which we measure by preferred tax rates on gas and electricity.
-We also estimate an additional model for petrol duty as a robustness test for the WTP model.
-How we measure the dependent variables is described in Section~\ref{ClimateVars}.
-The potential predictors included in $x$, which are not the behavioral variables and which were not selected into any model by multisplit lasso are listed in Tables~\ref{PotentialPredictors} and~\ref{PotentialPredictors2} in Appendix~$3$.
-How we measure the behavioural variables is discussed in Section~\ref{BehaviouralVars} and their descriptive statistics are summarised in Table~\ref{Descriptive} in Appendix~$3$ with the exemption of inequity aversion as this variable is considered as categorical and its frequencies are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-The predictors, which were selected into some model can be found in a table of estimates of the relevant models and their descriptive statistics or frequencies are summarised in Tables~\ref{Descriptive},~\ref{FreqiencyCat},~\ref{FreqiencyCat2},~and~\ref{FreqiencyBin} in Appendix~$3$.
-
+\subsection{Measures of Yield and climate}
+
+Prior we started  our research, we had to answer the following questions: How do we measure food security?
+And how do we measure drought?
+To answer the first questions, the measure which has been used as a proxy for food security in the literature are as follows:
+
+\vspace{2cm}
+
+The other important questions which need to be answered before starting our research are: How is drought defined?
+What are the ways of measuring drought?
+and How shall we measure drought for the purpose of our study?
+According to the international meteorological community, drought can be defined in several ways.
+In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
+ \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater.
+The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).
+
+Numerous definitions of drought and their role have been reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
+They have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
+Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
+Operational definitions are essential for an effective early warning system.
+An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.
+
+In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
+However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
+Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).
+
+For an extensive overview of various drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
+\cite{keyantash2002} quantify, evaluate and compare number of drought indices for meteorological, hydrological and agricultural forms of drought.
+Based on several criteria they conclude that rainfall deciles and SPI perform the best for meteorological drought.
+
+In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions.
+Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength and severity \citep{KlischAtz2016}.
+
+A description of other drought indices and measures can be found in Appendix~$1$.
+
+\textcolor{red}{We initially analysed SPEI and VCI indices and we tried to use them as input variables for our analysis. However, there were some serious problems with these measures.....}
+
+Therefore we opted for a different approach to gauging drought severity.
+In particular, we decided to utilize daily precipitation data \textcolor{red}{from...CHIRPS??
+with resolution...
+and daily temperature data from..at resolution....}.
+The frequency  of the maize yield data is yearly while we have daily weather data available.
+Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
+There are many possibilities how to aggregate daily weather data.
+Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
+
+\begin{itemize}
+
+\item \textbf{Precipitation:}
+
+\begin{itemize}
+\item Seasonal cumulative rainfall
+\item Seasonal standard deviation \textcolor{red}{actually it is monthly at the moment. I am not sure how I would aggregate this over season??}
+\item Seasonal coefficient of variation \textcolor{red}{actually it is monthly at the moment. I am not sure how I would aggregate this over season??}
+\item Maximum length of dry spell in number of days
+\item 
+\end{itemize}
+\end{itemize}
+\subsection{Data}
+
+\subsection{Linear Mixed Models}
 \sloppy
-The estimation function can be written as \citep{Friedman2010}:
-
-\begin{equation}\label{Lasso}
-\begin{array}{lcll}
-
- \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \boldsymbol{R}_{\lambda}(\beta_0, \beta)&=&
-   \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \bigg[ \frac{1}{2N}\sum_{i=1}^{N}(y_i - \beta_0-x_i^{\intercal}\beta)^2 + \lambda\sum_{j=1}^{p}(|\beta_j|) \bigg], \\
+Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}.
+As a result of the high degree of county-level autonomy, the policies and regulations are likely to differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
+Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (or linear mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
+These types of models include both fixed affects and random effects.
+Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}.
+On the other hand, random effect are unobserved random variables.
+There are at least three benefits of treating a set of parameters as a random sample from some distribution.
+\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9,KERYch12}).
+
+Formally, a linear mixed model can be described by the distribution of two vectors of random variables: the response $\mathscr{Y}$ and the vector of random effects $\mathscr{B}$.
+The distribution of $\mathscr{B}$ is multivariate normal and the conditional distribution of $\mathscr{Y}$ given $\mathscr{B}=\mathbf{b}$ is multivariate normal of a form (\citealp{bates2010lme4, KERYch9}):
+
+\begin{equation}\label{MixedGeneral}
+\begin{array}{lcl}
+
+(\mathscr{Y}|\mathscr{B}=\mathbf{b})& \sim & \mathit{N}(\mathbf{X}\mathbf{\beta}+\mathbf{Z}\mathbf{b},\sigma^2\mathbf{I}),
+
 \end{array}
@@
 
-%https://www.stata.com/support/faqs/statistics/stepwise-regression-problems/
-
-where $y_i$ is the value of one of our four dependent variables for an individual~$i$, $x_i$ includes potential predictors listed in Tables~\ref{PotentialPredictors}~to~\ref{FreqiencyBin} in Appendix~$3$, $N$ is the number of observations and $\lambda \geq 0$ is the penalty parameter.
-Without loss of generality, we assume that the potential predictors in~\eqref{Lasso} are standardized: ${\sum_{i=1}^{N}x_{ij} =0}$, ${\frac{1}{N}\sum_{i=1}^{N}x_{ij}^2=1}$, for $j=1,...,p$.~\footnote{Both $x_{ij}$ and $y_j$ are standardized automatically in the implementation of the algorithm  we use.
-However, the estimated coefficients are always returned and presented on the original scale.}
-
-In line with common practice, we compute estimator~\eqref{Lasso} for a series of~$\lambda$ and then we choose a preferred value of~$\lambda$ using cross-validation \citep{statisHighDimData}.
-In particular, we use a sequence of $100$ values of~$\lambda$ and $10$\nobreakdash-fold cross validation.\footnote{For estimation of lasso~\eqref{Lasso} we use function cv.glmnet in the \textbf{\textsf{R}}~programming system \citep{RRRR} and we use default settings and values of arguments, unless otherwise stated.} We opt for the value of $\lambda$ which is recommended by \citet{Friedman2010} and it is probably the most common choice.
-More specifically, we use the largest value of $\lambda$  such that the mean cross-validated error (CVM) is still within one standard error of its minimum.\footnote{In case of WTP we use the value of $\lambda$ which minimises the CVM.
-This value is also  suggested by \citet{Friedman2010}.
-The only difference from the model estimated using the one standard error based~$\lambda$ is that for the latter, a dummy variable for male becomes significant and gets into the model.
-The effect of male is positive and this contradicts predominant conclusions in previous relevant literature \citetext{e.g.
-\citealt*{Hamilton2011, McCright2010, HamiltonKeim2009, Flynn1994}}.}
-
-Determining significance levels is problematic with lasso.
-Classical \textit{p}-values are not valid and there is no simple approximation.
-Therefore, we adopt a concept of \citet{pValsLasso}, who introduce an approach based on multiple random splits of data, repeated estimation and aggregated inference.
-In particular, \citet{pValsLasso} build on the proposal of \citet{Wasserman2009}, who suggest to split the dataset randomly into two subsets.
-One of the subsets is used for variable selection via lasso and the other one is for estimating OLS with the predictors selected by lasso and calculating their \textit{p}-values in a usual way.
-This procedure allows asymptotic error control under minimal conditions.
-The problem is that the results depend on a one-time arbitrary split and they are therefore irreproducible.
-\citet{pValsLasso} further develop the single-split method.
-They suggest to split the sample repeatedly, obtain a set of \textit{p}-values for each split and then aggregate them.
-In each split, the \textit{p}-values of the variables which are not selected are considered to be equal to one and the \textit{p}-values of the selected variables are multiplied by the number of variables selected in the current split.
-If a \textit{p}-value multiplied by the number of selected variables happens to be larger than one, it is considered to be equal to one.
-Let's assume that we have $h=1,...,H$ splits.
-A \textit{p}-value for predictor $j$ obtained in split $h$ adjusted as described above will be further denoted $P^{(h)}_j$.
-\citet{pValsLasso} suggest to aggregate the adjusted \textit{p}-values using quantiles.
-In particular, a suitable aggregated \textit{p}-value is defined for any predictor $j$ and for any fixed $0<\gamma<1$ as
-
-\begin{equation}\label{pvals}
-Q_j(\gamma)=min \left \{1,q_{\gamma} (   \{ P^{(h)}_j/\gamma; h=1,...,H\}) \right \},
+where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variable $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ a scale factor.
+The distribution of $\mathscr{B}$ can be written as:
+
+\begin{equation}\label{ranefDist}
+\mathscr{B} \sim \mathit{N}(0,\mathbf{\Sigma}),
 \end{equation}
 
- where and $q_{\gamma}(\tiny{\cdot})$ is the (empirical) $\gamma$-quantile function. We will further refer to this procedure as a multisplit lasso.
-
-\citet{pValsLasso} show that for any predefined value of $\gamma \in(0,1)$, the \textit{p}-values defined in~\eqref{pvals} can be used for control of family-wise error rate\footnote{Probability of making at least one incorrect rejection of a true null hypothesis (type 1 error).} and also for regulation of false discovery rate.\footnote{Expected proportion of incorrect rejections of a true null hypothesis (type 1 errors).
-False discovery rate controlling procedures are less stringent than family-wise error rate controlling methods.} Moreover, the multisplit method improves the power of estimates.
-
-For simplicity, we set $\gamma$ in~\eqref{pvals} to be equal to $0.5$ for every application of a multisplit lasso in this study.
-Each time we perform $H=100$ splits (we believe that this number is sufficient as \citet{pValsLasso} use $50$ sample splits per simulation) and we always use one third of the sample for the variable selection using lasso and the rest for the OLS estimation and obtaining \textit{p}-values.
-
-A large fraction of our potential predictors are categorical variables because large part of the survey data was collected by multiple choice questions.
-However, the multisplit lasso selects individual predictors rather than groups of variables.
-Therefore, it can happen that a model specified by a multisplit lasso includes a dummy variable for one category of a particular categorical variable but it does not include dummy variables for its remaining categories.
-An obvious way how to overcome this issue would be to add the remaining dummy variables and use an \textit{F}-test to determine the joint significance of the group.
-If the \textit{F}-test implies that the categories are jointly significant, they should all stay in the model and they should be left out otherwise.
-However, the solution is not so straightforward with a multisplit lasso as it is not obvious on which subsample we should perform the \textit{F}-test.
-Therefore, for each model specified by a multisplit lasso, we decided to perform a following procedure which is sometimes called jackknife resampling.
-We will further refer to the procedure as a jackknife OLS.
-We again randomly split the dataset into two subsamples.
-The bigger subsample has size of two thirds of the original sample and it is used for OLS estimation and calculation of \textit{p}-values of the model with predictors selected by multisplit lasso.\footnote{Sample splitting can generally result in loss of efficiency.
-We, however estimated all models also for the whole sample as a robustness check and the results do not differ in signs or significance levels.} In addition, if the model specified by multisplit lasso includes a binary indicator which represents a category of a nominal variable, we include also all other categories of this variable among the set of predictors.
-Besides individual \textit{t}-tests we perform an \textit{F}-test of joint significance of the categories of the nominal variable.
-Similarly as in the case of multisplit lasso, we repeat the resampling and OLS estimation $100$-times.
-Each time we perform \textit{t}-tests and also a joint \textit{F}-test for each group of dummy variables representing one categorical variable.
-The \textit{p}-values of the \textit{t}-tests are then aggregated in the same way as in case of multisplit lasso (see above).
-Further, we calculate mean and median of \textit{p}-values of each joint \textit{F}-test over the $100$ subsamples and according to these statistics we determine whether the dummy indicators of the particular categorical variable should be included.
-It turns out that every time when a dummy variable representing a category of a nominal variable is chosen by a multiple lasso, both average and median \textit{p}-values of the corresponding \textit{F}-tests are below the significance level (${\alpha=0.05}$).
-Hence, we include the dummy variables for categories of each nominal variable selected by lasso (see Section~\ref{Results}).
-% parada.
-vsechny f testy jsem otestovala u vsech modelu v tabulkach.
-27.8.
-vse v poradku,takze super
+where $\mathbf{\Sigma}$ is a $q \times q$ positive semi-definite variance-covariance matrix
 
 \FloatBarrier
-\section{Data}\label{Data}
-All data used in this study except of predicted income and population density, which we use in robustness tests,  were collected in the survey conducted by \cite{SurveyUK}.
-
-In Section~\ref{Robust} we use an alternative measure of income as a robustness test.
-In particular, this estimated income is obtained from a regression model based on data from Annual Survey of Hours and Earnings (ASHE).
-More specifically, the predicted income is based on age, gender, occupation, sector and education.
-
-We use two measures of population density, in particular average density per Lower Layer Super Output Areas (LSOA) estimated by the Office for National Statistics for year $2015$ and average density for Local Authority Districts (LAD) obtained from the $2011$ Census.
-
-The online survey (\citealp{SurveyUK}) ran from $9$ September to $14$
-October $2015$ and $6,000$ respondents were selected to answer the questionnaire which included the climate change domain.\footnote{We had to exclude some observations from various parts of analysis as they included missing values for some important variables.
-However, we have at least $5500$ observations for each model.} Descriptive statistics, methodology of the survey, the survey itself and a detailed description of its administration can be found in~\citet{SurveyUK}.
-
-The survey is reasonably geogrephically representative taking into account population density in the UK \citep{SurveyUK}.\footnote{For map with location of respondents see Figure~$1$ in \cite{SurveyUK}} As the survey was conducted online, the initial sample is representative for UK adults with internet access rather than for the entire UK population.
-
-In Table~\ref{SexAge} we compare distribution of our sample over sex and age  with the distribution of the UK population.
-The age data are only available as a categorical variable in our survey.
-As we can see in Table~\ref{SexAge}, the youngest category is slightly over-sampled while the two categories of the highest age are slightly under-sampled, probably because the survey was conducted online.
-Otherwise the distributions are very comparable.
+	\section{Results and discussion}\label{Results}
 
@@
 \begin{threeparttable}
-\caption{\textit{\textbf{Sex and age distribution} \\ of the sample and the population}}
-\label{SexAge} 
-\begin{small}
-\begin{tabular}{|l|rr|rr|} 
-\hline	
-  \multicolumn{1}{|l|}{ } & \multicolumn{2}{c}{\bf{Sample}} & \multicolumn{2}{|c|}{\bf{UK population\tnote{a}}} \\ %sample pubpolM2 used here
-    \multicolumn{1}{|l|}{\vspace{0.1cm}\textbf{Age range}}  &  \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c}{\bf{Female}} & \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c|}{\bf{Female}}\\ 
-\hline 
-$18-24$&$9.8\%$&$9.4\%$&$6.2\%$&$6.0\%$\\
-$25-34$&$10.0\%$&$10.3\%$&$9.0\%$&$9.1\%$\\
-$35-44$&$7.8\%$&$8.3\%$&$8.6\%$&$8.8\%$\\
-$45-54$&$8.1\%$&$9.4\%$&$9.3\%$&$9.6\%$\\
-$55-64$&$7.4\%$&$8.4\%$&$7.5\%$&$7.8\%$\\
-$65-74$&$4.1\%$&$4.8\%$&$6.2\%$&$6.7\%$\\
-$75-80$&$0.1\%$&$0.1\%$&$2.4\%$&$2.8\%$\\
-\hline
-\hline
-\end{tabular} 
-  \begin{tablenotes}
-  \begin{footnotesize}
-  \singlespacing
-     \item[a]Population data are from the Office of National Statistics, Population Estimates of UK, England and
-Wales, Scotland and Northern Ireland Mid $2014$, Table MYE2.
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-\end{small}
-  \end{threeparttable} \par}
-
-\hspace{1.5cm}
-
-Although the survey questionnaire was designed such that more difficult questions were at different pages, we observe that most respondents who did not finish the survey dropped out on pages with more difficult questions.
-Hence, the final sample is biased towards those who are not afraid of hard questions \citep{SurveyUK}.\footnote{One way how to deal with sample selection is to use sampling weights.
-We, however decided not use weights given the modest nature of our bias.
-Weighting usually increases standard errors and leads to less precise estimates and there is lack of consensus on whether or not to use weights in regression methods (\citealp{gelman2007, KottWeight, winship94}).
-\citet{winship94} for example recommend not to use weights if they are solely a function of independent variables.}
-
-In the rest of this section we focus on how we obtained the data for our climate (dependent) variables and the behavioural characteristics..
-
-\subsection{Climate variables}\label{ClimateVars}
-
-Descriptive statistics of our climate variables are summarised in Table~\ref{DepDes}.
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Dependent variables:} Descriptive statistics}}
-\label{DepDes} 
+
+\singlespacing
+\caption{\textit{\textbf{Linear mixed effects models:} Maize yield and weather}}
+% Peggy45 & Peggy453   can be found in foodSystems/Rcodes/Lags/..
+  % Yield in tonnes per hectare
+\label{Peggy45} 
 \centering
 \begin{small}
-\begin{tabular}{lrrrr} 
-\hline	
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  &  \multicolumn{1}{c}{\bf{Mean}} & \multicolumn{1}{c}{\bf{St. dev.}} & \textbf{Min} & \textbf{Max}\\ 
-\hline \vspace{-0.3cm} \\ 
-  \vspace{0.15cm}Climate change knowledge&$3.851$&$1.266$&$1$&$8$\\
-    \vspace{0.15cm}Climate change seriousness perception&$6.622$&$2.249$&$0$&$10$\\
-        \vspace{0.15cm}Climate versus policy effects perception&$5.370$&$2.315$&$0$&$10$\\
-\vspace{0.15cm}WTP - gas and electricity tax (\textsterling~per year)&$123.900$&$105.459$&$0$&$500$\\
- \vspace{0.15cm}WTP - duty on transport fuel (pence per year)&$20.530$&$22.518$&$0$&$100$\\
-\hline
-\hline
-\end{tabular} 
-\end{small}
-  \end{threeparttable} 
-\par}
-
-\hspace{1.5cm}
-
-It was previously shown, that questions which are intended to measure climate science comprehension often measure who people are rather than what they know about climate change as the strongest predictor is often respondents' ideology and cultural and political world-view (\citealp{Hamilton2011, KahanEtAl2012, Kahan2015}).
-To avoid picking of effect of cultural or political world-view instead of climate knowledge, we use questions from the OCSI instrument developed by \citet{Kahan2015} as a measure of climate knowledge.
-\citet{Kahan2015} shows that these questions are indeed a measure of climate science comprehension rather than an indicator of who one is.
-The values of climate knowledge are integers from $0$ to $8$ and they stand for counts of correctly answered questions about climate change \citep{Kahan2015}.
-An example of one of the $8$ climate questions is: 'Climate scientists believe that if the North Pole icecap melted as a result of human-caused global warming, global sea levels would rise.
-Is this statement true or false?' The list of all climate questions can be found in Appendix~$1$.
-The relative frequencies of counts of the correctly answered questions are summarised in Table~\ref{DepFreq}.
-
-To investigate opinions about seriousness of climate change, the respondents were asked the following question: 'How serious a problem do you think climate change is at this moment?' Using an interactive slider, the respondents answered an integer value between $0$~and~$10$ where min~$=0$ and max~$=10$ (as it was noted just below the slider).
-In a similar way, the respondents were asked if they feel to be more affected by climate change or by climate policy.
-The wording of the question was: 'Which affects you and your way of life more, climate change or policies to reduce greenhouse gas emissions?' Again, the respondents provided answers on an integer scale from $0$ (climate policy) to $10$ (climate change) using a slider.
-Relative frequencies of climate seriousness perception and climate versus policy perception are summarised in Table~\ref{DepFreq}.
-
-\vspace{1cm}
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Dependent variables:} Relative frequencies (\%)}}
-\label{DepFreq} 
-\centering
-\begin{small}
-\begin{tabular}{lrrrrrrrrrrrr} 
-\hline 
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  & \multicolumn{1}{c}{\boldsymbol{$0$}}& \multicolumn{1}{c}{\boldsymbol{$1$}}& \multicolumn{1}{c}{\boldsymbol{$2$}}& \multicolumn{1}{c}{\boldsymbol{$3$}}& \multicolumn{1}{c}{\boldsymbol{$4$}}&\multicolumn{1}{c}{\boldsymbol{$5$}}&\multicolumn{1}{c}{\boldsymbol{$6$}}&\multicolumn{1}{c}{\boldsymbol{$7$}}&\multicolumn{1}{c}{\boldsymbol{$8$}}&\multicolumn{1}{c}{\boldsymbol{$9$}}&\multicolumn{1}{c}{\boldsymbol{$10$}}&\\
-\hline \vspace{-0.3cm} \\ 
-  \vspace{0.15cm}Climate knowledge&$0.0$&$1.7$&$11.4$&$30.4$&$25.4$&$20.9$&$8.6$&$1.6$&$0.1$&N/A&N/A\\
-\vspace{-0.3cm}Climate seriousness&\multirow{2}{*}{$3.3$}&\multirow{2}{*}{$2.8$}&\multirow{2}{*}{$5.4$}&\multirow{2}{*}{$8.1$}&\multirow{2}{*}{$8.9$}&\multirow{2}{*}{$27.2$}&\multirow{2}{*}{$14.0$}&\multirow{2}{*}{$12.8$}&\multirow{2}{*}{$8.3$}&\multirow{2}{*}{$4.1$}&\multirow{2}{*}{$5.0$}\\
-        \vspace{-0.15cm}\hspace{0.5cm}perception&\\
-\vspace{-0.3cm}Climate vs.
-policy&\multirow{2}{*}{$2.1$}&\multirow{2}{*}{$1.5$}&\multirow{2}{*}{$2.5$}&\multirow{2}{*}{$3.8$}&\multirow{2}{*}{$4.6$}&\multirow{2}{*}{$9.8$}&\multirow{2}{*}{$18.5$}&\multirow{2}{*}{$21.7$}&\multirow{2}{*}{$16.7$}&\multirow{2}{*}{$8.6$}&\multirow{2}{*}{$10.4$}\\
-\vspace{0.15cm}\hspace{0.5cm}perception\tnote{a}&\\
-\hline
+\begin{tabular}{lrccc} 
+\hline \vspace{-0.2cm} \\
+
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{2}{c}{\textbf{Scaled}}& \multicolumn{2}{c}{{\textbf{Unscaled\tnote{a}}}} \\
+
+    \multicolumn{1}{l}{\vspace{0.1cm}}  &\textit{Estimate}&\textit{p-value}&\multicolumn{1}{c}{\textit{Estimate}}& \multicolumn{1}{c}{\textit{p-value}} \\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$-0.059$&$0.646$&$1.379$&$1\times10^{-7}$$^{***}$\\
+  \\
+\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.054$&$0.041^{*}$&$3\times10^{-4}$&$0.177$\\
+  \\
+  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$&$0.477$&$3\times10^{-4}$&$0.197$\\
+  \\
+  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.045$&$0.244$&$-0.029$&$0.001^{**}$\\
+  \\
+    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.130$&$0.0001^{***}$&$-0.027$&$0.001^{**}$\\
+  \\
+
+      \vspace{-0.2cm}Prec. max OND, east&$0.133$&$0.013^{*}$&$0.004$&$0.085^{\bullet}$\\
+  \\
+        \vspace{-0.2cm}Prec. max OND, west&$0.144$&$0.004^{**}$&$0.011$&$9\times10^{-6}$$^{***}$\\
+  \\
+    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.049$&$0.199$&$0.075$&$0.369$\\
+  \\
+      Temp. sd. Oc.-Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$&$0.494$&$7\times10^{-12}$ $^{***}$\\
+    \vspace{-0.1cm} \\ 
+  \hline
+
+\vspace{-0.1cm}\\\multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
+\\\hline \vspace{-0.1cm} \\
+
+\vspace{0.2cm}Intercept\\
+\vspace{0.2cm}Prec. cum. MAM+OND lag 1\\
+\vspace{0.2cm}Prec. max OND\\
+\vspace{0.2cm}Temp. avg. Mar.-Sep. lag 1
+  \\
   \hline
@@
   \begin{footnotesize}
-  \singlespacing
-     \item[~]\textit{Notes:} Total number of observations: $5749$
+    \item \textit{Notes:} \hspace{0.15cm}$584$ observations; \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
     \begin{adjustwidth}{1cm}{} 
-     \vspace{-0.3cm} \item[a]~Higher number means greater concern about climate change, lesser concern about climate policy.
-         \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\vspace{1cm}
-
-Regarding the preferred gas and electricity tax rates, the respondents were first asked how much the current tax was.
-In particular, the question was as follows: 'The average household pays \textsterling$1,369$ per year for gas and electricity.
-Government intervention has raised the price to encourage people to use less and so reduce greenhouse house gas emissions.
-How much of that \textsterling$1,369$ is for climate policy?' They indicated the response on a slider with a minimum of $-50$ and a maximum of $500$.
-We include this variable on right hand site as a robustness test (see Table~\ref{GasInOut}).
-We refer to it as 'How much is tax gas and electricity'.
-After this, the respondents were told the correct answer and they were asked about they preferred tax rates: 'Actually, climate policy adds about \textsterling$89$ per year to the gas and electricity bill of the average household.
-How much do you think climate policy should add to this bill?' The respondents expressed their opinion on a slider from $0$ to $500$.
-The answer to this question is the dependent variable which we refer to as 'WTP - gas and electricity' and we use it as a proxy for WTP for climate change mitigation.
-Analogously, we inquired about the fuel duty.
-The only difference is that the slider for the actual fuel duty is limited from $0$ to $60$ and the one for the preferred fuel duty is from $0$ to $100$ as the actual fuel duty is $3$ pence per litre.
-Descriptive statistics of the respondents' estimates of actual tax rates can be found in Table~\ref{Descriptive} in Appendix~$3$ and the descriptive statistics of the preferred tax rates are in Table~\ref{DepDes}.
-
-\subsection{Behavioural variables}\label{BehaviouralVars}
-
-One of our goals is to investigate effects of behavioural variables on climate knowledge and concerns about climate change.
-The behavioural variables that we consider in our study are social value orientation, time preferences, risk preferences, and attitudes towards inequality.
-
-To estimate the social value orientation, respondents played six dictator games with the same questions as in \citet{murphy2011SVO}.
-The ring measure of social value orientation which we use in our models is defined as
-
-\begin{equation}\label{RingMeasure}
-\begin{array}{lcll}
-
-R&=&\arctan \frac{\sum_{i=1}^{N} P_O -50N}{\sum_{i=1}^{N} P_S -50N}, \\
-\end{array}
-\end{equation}
-
-where $P_O$ is the pay-off given to the other party, $P_S$ is the pay-off taken by the player herself and $N$ is the number of games played (in our case $6$).
-
-As one may notice in Table~\ref{PotentialPredictors} in Appendix~$3$, we also include dummy variables for four types of social value orientation (i.e.
-altruist, prosocial, individualist, competitive) among potential predictors in the lasso estimators.
-These types are defined based on ring measure~\eqref{RingMeasure}.
-Each dummy variable corresponds to one of four non-overlapping intervals of the values of ring measure~\eqref{RingMeasure}.
-None of these dummy variables was selected by lasso into any of our models, therefore we do not discuss them in more detail.
-
-As a basic measure of time preferences we use derived annual discount rates (in percentage), for investing now for one year from now and we refer to this variable as 'Discount rate year from now' in the present study.
-To obtain the data which would allow us to infer the discount rates, the respondents played games and answered questions informed by \citet{voors2012violent}, \citet{ifcher2011} and \citet{tanaka2010riskTime}.
-How the time preferences were derived is described in \cite{SurveyUK}.
-Besides using discount rates for investing now and getting returns in one year, we also implied other types of discount rates.
-These are discount rates for $(i)$~ investing now for getting returns in five years $(ii)$~investing in one year for getting returns in two years from now and $(iii)$~ investing in one year from now for getting return in six years from now.
-None of them was found to be significant, thus we do not further discuss them.
-
-We use two parameters which describe inequity aversion (\citealp{bergson1938, bergson1954, samuelson1956}), in particular the rate of inequity aversion and the subsistence or reserve income. 
-To infer these parameters, respondents were choosing from various distributions of income between three hypothetical people.
-One of them was higher on average but more unequal and the other was lower on average but more equal.
-The respondents were asked two sets of choice questions.
-In one of them, the income distribution was centred on the $70^{th}$ percentile of the UK income distribution and in the other the distribution was centred on $40^{th}$ percentile of the UK income distribution.
-Given the respondents' answers to the two choice-sets, the rate of inequity aversion and subsistence was obtained for each respondent based on equations~\eqref{InequityEq}:
-
-\begin{equation}\label{InequityEq}
-\begin{array}{lcl}
-
-\sum_{i=1}^{3}\frac{(Y^H_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^H_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma} \\
-
-\vspace{0.2cm}\\
-\sum_{i=1}^{3}\frac{(Y^L_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^L_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma}
-\end{array}
-\end{equation}
-
-where $\gamma$ is the rate of inequity aversion, $\underline{Y}$ is the subsistence or reserve income and $Y^H_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $70^{th}$ percentile of the UK income distribution.
-Analogously, $Y^L_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $40^{th}$ percentile of the UK income distribution.
-To obtain the inequity parameters, equations~\eqref{InequityEq} were solved for $\gamma$ and $\underline{Y}$ while minimizing distance of $\underline{Y}$ to zero.
-
-In theory, the rate of inequity aversion is a continuous measure.
-However, we consider it as a categorical one as in our dataset it is equal to one of $16$ distinct values for each respondent.\footnote{We also estimated variants of models where the rate of inequity aversion is considered as scale for completeness but we do not present them to save space.
-However, the results do not differ substantially from those presented here.} These $16$ values and the corresponding frequencies can be found in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-The subsistence parameter was not selected by lasso into any of our models thus we do not discuss it in more detail.
-
-To test significance of risk aversion, we use various risk aversion coefficients which were estimated for each person from four different utility functions using Bayesian inference \citep{Balcombe2015}.
-The utility functions are power, logarithmic, exponential and quadratic and we use the estimates of their means and medians.
-None of them is significant or chosen by lasso in any of our models.
-For the economy of space we only present models with median or mean of power function.
-The estimates are very similar when we use other risk aversion coefficients.
-
-As the behavioural variables are not significant in our study, we described their measures only briefly.
-For more detailed description see \citet{SurveyUK}.
-The descriptive statistics of these variables (except of inequity aversion) can be found in Table~\ref{Descriptive} in Appendix~$3$.
-As explained above, we consider the inequity aversion rate as a categorical variable and its frequencies are in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-
-\FloatBarrier
-	\section{Results and discussion}\label{Results}
-In this section we describe our results and discuss their interpretation.
-
-In the tables which summarise the estimates of lasso below, \textit{p}-values of some of the explanatory variables are equal to one.
-These variables were not selected by the lasso in most of the sample splits.
-They are, however, included in the tables because they represent either a category of a nominal variable whose other category was selected by the lasso or a linear term of a variable whose quadratic term was selected by the lasso.
-
-\subsection{Climate change knowledge}\label{ResKnow}
-
-Table~\ref{Know435} summarizes estimates of the predictors of climate change knowledge which we found to be important by means of multisplit lasso estimator.
-In particular, three predictors are chosen by lasso (see first column in Table~\ref{Know435}).
-Total score on financial literacy is the number of correct answers out of three finance related mathematical problems \citep{SurveyUK}.\footnote{We also consider answers to each of the three problems separately as individual potential predictors.
-Two of them are labelled understands inflation and understands compound interest and they are identified as important predictors in other models later.} However, when we re-estimate the model using jackknife OLS with all relevant dummy variables, all categories of total score on financial literacy are insignificant.
-Furthermore, the model suffers from multicollinearity as the coefficient of correlation between cognitive reflection and total score of financial literacy is equal to $0.343$ and its \textit{p}-value is smaller than $2\times10^{-8}$.
-For illustration, estimates of jackknife OLS with all explanatory variables listed in Table~\ref{Know435} including financial literacy are shown in Table~\ref{Know435se} in Appendix~$3$.
-The last column of Table~\ref{Know435se} includes variance inflation factors (VIF) which confirm the presence of multicollinearity.
-Because of the multicollinearity and insignificance of total score on financial literacy we do not further consider this variable as a predictor of climate knowledge.
-
-The estimates of jackknife OLS without the financial literacy are summarised in the last two columns of Table~\ref{Know435}.
-The other two variables which were found to be important in explaining climate knowledge are gender and cognitive reflection test~\citep{Frederick2005}.
-We use the latter as a measure of numeracy and ability of analytical reasoning.
-
-The cognitive reflection test is fully described in~\citet{Frederick2005} and it consists of three numerical problems.
-The value of our variable is the number of correct answers out of the three questions.\footnote{The possible values are integers and half-integers between zero and three including zero and three as we also recognise if respondent solves half of a problem.
-Hence, if a respondent answers for example one and half problems correctly, her score is~$1.5$.} The frequencies of values of this variable are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-To account for plausible non-linear relationship between the test score and cognitive ability we treat the variable as categorical with the base category zero.
-As it is apparent from Table~\ref{Know435}, the respondents who solved all three problems correctly have significantly higher level of climate knowledge compared to those who did not solve any of them.
-According to the jackknife OLS, climate knowledge is on average higher also for respondents who answered two problems correctly.
-However, the effect is larger for three correctly answered problems.
-Expectedly, the effect of numeracy is positive.
-
-{\centering
-\begin{threeparttable}
-
-% table checked 3.10.2017
-
-\singlespacing
-\caption{\textit{\textbf{Climate change knowledge:} Multisplit lasso and jackknife OLS}}
-% LassoKnow43se3.R      soubor Know435.R 
-%musim lasso predelat. predtim jsem pouzivala lambda min. ale alex ma pravdu. je treba u vsech modelu pouzivat stejne pravidlo, takze lambda se.zmena 30.9.2017
-
-%soubor Know435se.R
-% finlit2 prestane byt signifikantni, takze vlastne se nic skoro nemeni, asi do prilohy tabulku s Know435se.R
-
-\label{Know435} 
-\centering
-\begin{small}
-\begin{tabular}{lclrcl} 
-\hline \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\
-
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Gender = male&$<2\times 10^{-8}$&$^{***}$&$0.733$&$<2\times10^{-8}$&$^{***}$\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=0$~\tnote{a}&$0.038$&$^{*}$ &\multicolumn{3}{c}{\textit{Not included - base cat.}}\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.000$& &$2.071$&$1.000$&\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$1.000$& &$0.283$&$0.129$& \\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1.5$&$1.000$& &$0.895$&$1.000$ \\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$1.000$& &$0.628$&$1\times 10^{-5}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$1.000$& &$1.098$&$1.000$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.046$&$^{*}$&$1.033$&$<2\times10^{-8}$&$^{***}$\\
-\\
-\vspace{-0.2cm}Financial literacy total score $=0.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=1$&$1.000$&& \multicolumn{3}{c}{\textit{Not included}} \\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=1.5$&$1.000$&&\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=2$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=2.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-  \vspace{-0.2cm}Financial literacy total score $=3$&$2\times10^{-5}$&$^{***}$&\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\hline
-\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-    \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
-    \item[a]The estimate is negative for cognitive reflection $=0$ while it is positive for cognitive reflection $=3$ in this model.
+    \item[a] The variety with unscaled variables fails to converge.
      \end{adjustwidth}
@@
 
-We find the positive and strongly significant effect of dummy variable for males quite peculiar.
-Previous research shows mixed evidence about effects of gender on climate knowledge and comprehension of science in general.
-For example, \citet{McCright2010} finds that women demonstrate higher level of scientific knowledge of climate change.
-On the other hand, \citet{Hayes2001gender} shows that men exhibit significantly higher level of scientific knowledge than women, even if controlling for a number of background variables.
-We perform additional tests to verify whether the positive effect of gender can be a result of sample selection.
-The tests include proportion tests, model with interactions as additional explanatory variables and a Heckman selection model.
-We discuss the results in detail in Appendix~$2$.
-Based on the outcomes, we conclude that the results are not driven by sample selection.
-
-A possible explanation why our measure of climate knowledge is significantly higher for men is that the climate knowledge test that we use in this study was developed by a man \citep{Kahan2015}, therefore it may be the case that these particular questions are naturally more comprehensible for men.
-The only way how to test this would be to let a woman design another set of climate knowledge questions and then conduct a survey which would include these woman-designed climate questions.
-This is, however, beyond the scope of this study.
-
- To sum up, we find that gender and cognitive ability are significant predictors of climate knowledge.
-Climate knowledge increases with higher numeracy which is consistent with \citet{Kahan2015}, who finds the climate knowledge measure to be positively correlated with ordinary science intelligence.
-Although various measures of climate knowledge were previously find to be correlated with social ideology or partisan identity (\citealp{Hamilton2011, Kahan2012, Kellstedt2008}), our measures of ideology, cultural world-view or their interactions were not chosen as predictors of climate knowledge by the lasso.
-This is also consistent with \citet{Kahan2015}.
+{\centering
+\begin{threeparttable}
+
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather}}
+% Peggy45ln can be found in foodSystems/Rcodes/Lags/..
+ % Yield in tonnes per hectare
+\label{Peggy45ln} 
+\centering
+\begin{small}
+\begin{tabular}{lrl} 
+\hline \vspace{-0.2cm} \\
+
+\vspace{-0.2cm} \\
+
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{1}{c}{\textit{Estimate}} &\multicolumn{1}{c}{\textit{p-value}}\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$0.158$&$0.086^{\bullet}$\\
+  \\
+\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.066$&$0.008^{**}$\\
+  \\
+  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$-0.006$&$0.861$\\
+  \\
+  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.036$&$0.292$\\
+  \\
+    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.081$&$0.008^{**}$\\
+  \\
+
+      \vspace{-0.2cm}Prec. max OND, east&$0.081$&$0.056^{\bullet}$\\
+  \\
+        \vspace{-0.2cm}Prec. max OND, west&$0.108$&$0.009^{**}$\\
+  \\
+    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.101$&$0.003^{**}$\\
+  \\
+      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, west&$0.142$&$6\times10^{-7}$ $^{***}$\\
+  \\
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+  \\
+  \vspace{-0.2cm}Prec. cum. MAM+OND lag 1\\
+  \\
+  \vspace{-0.2cm}Prec. max OND\\
+  \\
+    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1\\
+  \vspace{-0.1cm} \\ 
+  \hline
+
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} \hspace{0.05cm}$584$ observations
+        \begin{adjustwidth}{1cm}{} 
+    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
 
@@
 
-We now discuss the estimates of the model with dependent variable which answers the question whether respondent feels to be more affected by climate policy ($0$) or by climate change ($10$).
-The estimates are shown in Table~\ref{Climpol435}.
-We can see that the selected predictors are climate knowledge, understanding of inflation and risk assessment consistency.
-
-The level of understanding of inflation is based on respondents' answer to the following numerical problem: 'Imagine that the interest rate on your savings account was 1 percent per year and
-inflation was 2 percent per year. After 1 year, would you be able to buy...' The respondents should choose one of the three following answers: $(i)$~'More than today with the money in
-this account' $(ii)$~'Exactly the same as today with the money in
-this account' $(iii)$~'Less than today with the money in
-this account'.
-Correct answer is 'less'.
-The value of the variable is equal to one if the respondent answers 'less', it is equal to $0.5$ if she answers 'the same' and it is equal to zero if she answers 'more'.
-The frequencies of answers are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-Since it is quite possible that the effect of our measure of understanding of inflation is non-linear we treat the variable as categorical with the base category zero.
-
-As it is apparent from Table~\ref{Climpol435}, understanding of inflation and risk assessment consistency increase the likelihood of being subjectively more affected by climate policy than by climate change.
-This is likely to be because the two predictors are highly correlated with financial literacy.
-The correlation coefficient of understanding of inflation and financial literacy is $0.694$ and the correlation coefficient of risk assessment consistency and financial literacy is $0.145$.
-Both correlation coefficients are highly significant with \textit{p}-value lower than $2.00\times 10^{-8}$.~~\footnote{Risk assessment consistency is a binary variable so we also run a two sample \textit{t}-test to measure correlation between risk assessment consistency and financial literacy.
-In particular, we applied a two sample \textit{t}-test to test if mean financial literacy is statistically equal for the respondent who answered risk questions consistently and for those with inconsistent answers to risk questions.
-The test statistic is highly significant with \textit{p}-value lower than $2.00\times 10^{-8}$.
-Hence, the mean financial literacy is different in these two groups which is in accordance with the significance of the correlation coefficient} It is intuitive, that the respondents with higher level of financial literacy are more likely to see how their wealth and way of living can be affected by climate policy through environmental tax rates.
-
-It was previously shown that interactions of measures of cognitive ability and ideological and political world-view are strong predictors of attitudes towards climate change rather than cognitive ability or numeracy itself \citep{Kahan2012, KahanEtAl2012, Hamilton2011, HamiltonKeim2009}.
-In accordance with this (as we discuss in more detail in Section~\ref{ResWTP} below) we detect a significant impact of interactions of an indicator of political and cultural world-view and a measure of numeracy (and ability of analytical, technical reasoning) on WTP for climate change mitigation.
-Therefore, we also estimate variants of the models presented in this section with the interaction terms included among the predictors but we did not find them to be significant for climate risk perception.
-We do not present the results in our study to keep its length within reasonable limits.\footnote{It would probably be more revealing to test for significance of interactions of cognitive ability and political orientation, but unfortunately, the respondents were not asked about their political or partisan preferences directly in the survey.}
-
-\vspace{1.5cm}
-
-{\centering
-\begin{threeparttable}
-\caption{\small{\textit{\textbf{Climate versus policy effects perception:} Multisplit lasso and jackknife OLS}}}
-% soubor LassoClimpol43 Climpol435.R
-\label{Climpol435} 
-\centering
-\begin{small}
-\begin{tabular}{lclccl} 
-\hline \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\
-
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Climate knowledge&$1.0000$& &$0.2158$&$1.0000$&\\
-  \\
-\vspace{-0.2cm}Climate knowledge - squared&$<2.00\times 10^{-8}$&$^{***}$&$-0.0607$&$0.0124$&$^{*}$\\
-  \\
-\vspace{-0.2cm}Understands inflation $=0.5$&$1.0000$& &$-0.0394$&$1.0000$& \\
-  \\
-\vspace{-0.2cm}Understands inflation $=1$&$0.0130$&$^{*}$&$-0.5759$&$6.27\times 10^{-6}$&$^{***}$\\
-    \\
-  Consistent answers to risk&\multirow{2}{*}{$1.06\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-0.5885$}&\multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
-\hspace{0.6cm}questions $(0/1)$&& &&&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
-\begin{tablenotes}
-  \begin{footnotesize}
-   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-  \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
-     \end{adjustwidth}
- \singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\vspace{1cm}
-
-As we can see in Tables~\ref{Climcare435} and~\ref{Climpol435}, the linear climate knowledge term is positive (and insignificant), but the squared term is negative and significant for both climate seriousness and climate change versus policy.
-That means, the effect of climate knowledge is positive but decreasing for low levels of knowledge, while for medium and high degree of climate knowledge the effect is negative.
-The negative effect of higher levels of climate knowledge may seem to be counter-intuitive, however it is less surprising in the light of previous literature.
-According to \citet{reynolds2010now}, people who know less about climate change tend to ascribe unrealistic consequences (for example skin cancer) to global warming.
-The authors argue that it is even possible for any future ecological or political disaster to be viewed as a consequence of global warming by public \citep{WhatPplKnow94}.
-Also the level of familiarity with causes and basic mechanisms of climate change is quite unsatisfactory.
-For example, members of public tend to confuse climate and weather\footnote{Actually, some researchers confuse climate and weather too \citep{Hsiang2013, DesGreenstone2007}.} and as a result they mostly agree with the statement that climate changes from year to year \citep{reynolds2010now}.
-It is therefore understandable that individuals who are weaker in climate knowledge are more concerned about climate change and its consequences.
-
-Another possible explanation of the negative and significant climate knowledge effect can be affective orientation towards global warming \citep{Kahan2015}.
-It was previously shown that individuals, who believe that climate change is real and caused by humans and who correctly assert, for instance, that usage of fossil fuels is one of the causes of global warming are also likely to affirm other, perhaps false statements which are consistent with higher environmental risks \citep{reynolds2010now}.
-An example of such a false statement is that atmospheric emissions of sulphur contribute to global warming \citep{Kahan2015}.
-The OCSI questions which we use to measure climate knowledge are true/false statements and more than half of them is of a same type as the sulphur emissions statement above.
-That is, the correct answer does not evince concerns about climate change while the incorrect one does.
-This could explain why the respondents who believe that climate change is quite serious are likely to score lower on climate knowledge.
-If this is true, the climate concern variables are predictors of climate knowledge and not the other way around.
-Therefore, if this is true, climate knowledge should not be included in the specifications with estimates summarised in Tables~\ref{Climcare435} and~\ref{Climpol435}.
-As a robustness test, we estimate the models for climate seriousness perception and climate change versus climate policy not including the climate knowledge as an explanatory variable.
-In both cases, the estimates of the rest of the explanatory variables and their significance levels are almost the same as in the case with climate knowledge and they are summarised in Table~\ref{ClimCare435eClimpol435f} in Appendix~$3$.
-
-\subsection{Willingness to pay for climate change mitigation}\label{ResWTP}
-This section is focused on models explaining preferred gas and electricity tax rates, which we use as a measure of WTP for climate change mitigation.
-The estimates of the multisplit lasso and the jackknife OLS are summarised in Table~\ref{Gas435cp}.
-
-One of the important selected predictors is age.
-The age was recorded as a categorical variable with the lowest category $24$ or younger, the second lowest category is $25-34$, the third one is $35-44$ and so on up to the highest category which is  $75$ or older.
-We use the lowest age group ($24$ or younger) as the base category.
-Coefficients of all higher categories are negative and with exemption of $35-44$ and $75$ or older they are all significant.
-Thus, WTP declines with age, perhaps because older people have lower likelihood of experiencing tougher consequences of climate change predicted for more distant future \citep{Hamilton2011}.\footnote{We also estimated the model with interactions of age and number of children and grandchildren as older people who have more offspring can obviously be more concern about future than those who do not have children.
-However, we did not find the interactions to be significant.
-We do not present the results in this study to save space.}
-
-{\centering
-\begin{threeparttable}
-\caption{\small{\textit{\textbf{WTP climate - gas and electricity tax:} Multisplit lasso and jackknife OLS}}}
-% soubor Gas435cp.R, LassoGas47b.R
-% soubor Gas435cp2.R, LassoGas47b.R
-
-%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
-\label{Gas435cp} 
-\centering
-\begin{small}
-\begin{tabular}{lclccl} 
-\hline \vspace{-0.35cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}} %LassoGas47b
-  & \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\%Gas435cp
-
-\vspace{-0.35cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.35cm}Age\tnote{a} $~25-34$&$1\times10^{-6}$&$^{***}$&$-13.271$&$0.266$& \\
-        \\
-\vspace{-0.35cm}Age $35-44$&$0.006$&$^{**}$&$-29.494$&$3\times10^{-7}$&$^{***}$\\
-        \\ 
- \vspace{-0.35cm}Age $45-54$&$1.000$& &$-34.625$&$<2\times10^{-8}$&$^{***}$\\
-        \\ 
-  \vspace{-0.35cm}Age $55-64$&$1.000$& &$-40.061$&$<2\times10^{-8}$&$^{***}$\\
-        \\       
- \vspace{-0.35cm}Age $65-74$&$1.000$& &$-46.006$&$<2\times10^{-8}$&$^{***}$\\
-        \\ 
- \vspace{-0.35cm}Age$~75$ or older&$1.000$& &$-26.071$&$1.000$& \\
-        \\       
- Climate versus policy&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$10.408$}&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
-\vspace{-0.35cm}\hspace{0.6cm}effects perception&&&\\
- \\
-
-  \vspace{-0.35cm}Inequity aversion (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{*}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
-      \\
- Equal intergenerational&\multirow{2}{*}{$0.011$}&\multirow{2}{*}{${*}$}&\multirow{2}{*}{$20.760$}&\multirow{2}{*}{$0.002$}&\multirow{2}{*}{$^{**}$}\\
-\vspace{-0.35cm}\hspace{0.6cm}allocation of resources $(0/1)$\tnote{b}&&&\\
- \\  
- Understands compound&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{ }&\multirow{2}{*}{$-2.942$}&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{ }\\
-\vspace{-0.35cm}\hspace{0.6cm}interest $=0.5$&&&\\
- \\   
- Understands compound&\multirow{2}{*}{$1\times10^{-5}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-39.381$}&\multirow{2}{*}{$3\times10^{-5}$}&\multirow{2}{*}{$^{***}$}\\
-\vspace{-0.35cm}\hspace{0.6cm}interest $=1$&&&\\
-  \\  
-\vspace{-0.35cm}Understands inflation $=0.5$&$1.000$& &$-15.892$&$0.516$& \\
-  \\
-\vspace{-0.35cm}Understands inflation $=1$&$6\times10^{-5}$&$^{***}$&$-42.711$&$<2\times10^{-8}$&$^{***}$\\
-  \\
- Consistent answers to risk&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$-34.861$}&\multirow{2}{*}{$<2\times10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
- \hspace{0.6cm}  \vspace{-0.35cm} questions $(0/1)$&& &&&\\
-\\ Consistent answers within&\multirow{2}{*}{$0.045$}&\multirow{2}{*}{$^{*}$}&\multicolumn{3}{c}{\textit{Not included}}\\
-\hspace{0.6cm}investments $(0/1)$\tnote{c}&& &&&
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-
-     \item[~]\textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\vspace{-0.4cm}  \begin{adjustwidth}{0.65cm}{} \item \vspace{-0.1cm}The signs of the significant coefficients in the multisplit lasso are same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.\vspace{0.05cm}
-\item[a]Age is only available as categorical with  base category '$24$ or younger'. Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}).
-\vspace{0.05cm}
-\item[b]This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age).
-The variable is equal to $0$ for all other respondents.
-\vspace{0.05cm}
-\item[c]We eventually excluded this variable from the further analysis.
-Although consistency within investment was selected by lasso,  it is only marginally significant and strongly correlated with risk assessment consistency.
-Furthermore, even after exclusion of this variable the model includes relatively large number of predictors and their signs and significance levels do not change.
-    \end{adjustwidth}
-
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\newpage
-
-We can see in Table~\ref{Gas435cp} that another very strong predictor of WTP is perception of climate versus policy effects on ones way of living which we analyse as a dependent variable in Section~\ref{ResPerc}.
-Expectedly, those who perceive effects of climate change as more serious than effects of climate policy have higher WPT for climate change mitigation.
-As we will discus below, this variable is a partial mediator of impact of financial literacy.
-
-Another predictor of WTP is inequity aversion.
-As we explained in Section~\ref{BehaviouralVars}, we treat this variable as categorical.
-The values of our inequity aversion measure and their frequencies are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-Although the effect of inequity aversion is largely positive and decreasing, the signs and significance levels vary across the categories without any clear pattern.
-
-It is further apparent from Table~\ref{Gas435cp} that WTP is higher for respondents who believe that their income and standard of living generally is about equal to the income and standard of living of their parents when they were about the respondent's age and they also believe that their income and standard of living is equal to the income and standard of living of their children when they will reach the respondent's age.
-
-We can also see in Table~\ref{Gas435cp} that WTP for climate change mitigation is significantly lower for individuals with higher financial literacy\footnote{Variable 'Understands inflation' is described in Section~\ref{ResPerc}.
-Values of variable 'Understands compound interest' are based on respondents answer to the following numerical problem: 'Suppose you had \textsterling $100$ in a savings account and the interest rate was $2$ percent per year.
-After $5$ years, how much do you think you would have in the account if you left the money to grow?' The respondents should choose the correct answers from the following three options: $(i)$~'More than \textsterling $102$' $(ii)$~'Exactly \textsterling $102$' $(iii)$~'Less than  \textsterling $102$'.
-Correct answer is 'more'.
-The value of the variable is equal to one if the respondent answers 'more', it is equal to $0.5$ if she answers 'the same' and it is equal to zero if she answers 'less'.
-The frequencies of answers are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.} and for those who answered the questions about risk consistently.
-This is similar to model with climate versus policy perception as dependent variable (see Table~\ref{Climpol435}).
-It is very likely, that financial literacy and risk assessment consistency are strongly correlated with ability of analytical reasoning and comprehension of quantitative information, hence, the former can be interpreted as a measure of the latter.
-Because of complexity of the climate system and inherited difficulty of understanding of climate change by the public these findings may seem to be counterintuitive \citep{Weber2011}.
-More specifically, one may expect the climate concerns to intensify with increasing level of analytical reasoning and numeracy.
-Our evidence is, however, consistent with previous literature (\citealp{KahanEtAl2012, Kahan2015, Kellstedt2008}).
-
-\citet{TerrorClimate} and \citet{KahanEtAl2012} argue that the risks related to natural hazards caused by climate change are quite abstract and remote compared to other more salient risks such as terrorism.
-Hence, it is difficult to perceive the climate change risk as a relatively serious one.
-It was shown that attitudes towards climate change and related risks are indicators of personal world-view or political outlook rather than correlates of numeracy or science comprehension.
-People, who identify themselves with egalitarian, communitarian ideology tend to take climate change more seriously than those with rather hierarchical, individualistic world-view \citep{Leiserowitz2013, KahanEtAl2012, WHITMARSH2011}.
-It can be more important for an individual to consider the climate risk questions from cultural identity perspective than from a scientific and collective knowledge acquisition viewpoint \citep{Kahan2015}.
-Whether an individual is right or wrong has no meaningful impact on climate change.
-Decisions of a single consumer or voter can hardly make a measurable difference to the natural hazard risks caused by climate change.
-On the other hand, adopting a position which is not consistent with one's cultural group can have dangerous consequences \citep{Kahan2012}.
-\citet{KahanEtAl2012} show that the ideological polarization over climate change is higher among people with the highest degrees of numeracy and science literacy.
-That means, for the individuals who identify themselves with hierarchical, individualistic ideology, the climate concern is negatively correlated with numeracy and science literacy while for the individuals who believe in rather egalitarian, communitarian ideology the correlation is positive.
-A possible interpretation is that the members of public with higher degree of numeracy and analytical reasoning are using these abilities to protect their cultural identity and they are therefore better in interpreting the scientific facts in a way which is consistent with their cultural group's ideology.
-Following \citet{Hamilton2011}, we test this hypothesis by including interaction terms of degree of agreement with redistribution of income and financial literacy among the set of explanatory variables.
-The estimates are summarised in Table~\ref{Gas435dcp} in Appendix~$3$.
-The interaction is positive and significant, which means that the positive effect of agreeing with income redistribution is much stronger for those who understand inflation.
-Similarly, the negative effect of not agreeing with income redistribution is larger in magnitude if accompanied with higher level of understanding of inflation.
-This is in accordance with the theory that the ideological polarization over climate change is higher among people with higher degrees of numeracy and science literacy (\citealt{Kahan2012, KahanEtAl2012, Hamilton2011, HamiltonKeim2009}).
-Our results are robust, the estimates and their significance levels are almost the same as those of the model without the interaction term in Table~\ref{Gas435cp}.
-
-One may notice that level of understanding of inflation is a significant predictor of both climate versus policy effects perception and WTP for climate change mitigation.
-In both cases the effect is negative.
-The two climate variables are also strongly correlated.
-Hence, we will now focus on disentangling the structure of relationships among these three variables.
-
-We reveal that the measure of climate versus policy perception partially mediates effect of understanding of inflation on WTP.
-In Table~\ref{Climpol435} we can see that understanding of inflation is a significant (negative) predictor of climate versus policy perception and in Tables~\ref{Gas435cp} and~\ref{Gas435dcp} we can notice that climate versus policy effects perception is a significant (positive) predictor of WTP.
-In Table~\ref{MediationTIV} we regress WTP on understanding of inflation without the mediator in order to verify whether the basic condition of mediation is satisfied, i.e.
-whether we can see the significant effect of the predictor when the mediator is not present.
-Model~$1$ in Table~\ref{MediationTIV} is a regression of WTP solely on understanding of inflation while Model~$2$ includes also the other predictors selected by the multisplit lasso.
-Even without the mediator, the effect of understanding of inflation is strongly significant.
-Furthermore, the effect is larger in magnitude than the effect in the regressions which include the mediator (compare with the estimates in Tables~\ref{Gas435cp} and \ref{Gas435dcp}).
-This finding also supports the occurrence of mediation.
-Table~\ref{MediationTMV} summarizes estimates of WTP regressed on the mediator without the effect of understanding of inflation.
-Model~$1$ in Table~\ref{MediationTMV} only includes climate versus policy as explanatory variable while Model~$2$ in Table~\ref{MediationTMV} also includes the other predictors selected by the multisplit lasso.
-If the mediation is present, the mediator should also be a significant predictor of the dependent variable itself and we can see in Table~\ref{MediationTMV} that this is true in our case.
-
-As we can see in Tables~\ref{Gas435cp} and~\ref{Gas435dcp}, the effect of understanding of inflation is significant if the mediator is present, therefore the mediation is partial.
-
-To verify our conclusion about the presence of mediation, we perform the Sobel test for the effect of understanding of inflation being mediated through climate versus policy variable.
-The test statistic is strongly significant with \textit{p}-value equal to $4.21\times10^{-22}$, hence the Sobel test supports the occurrence of mediation.
-
-To sum up, people who understand inflation tend to feel to be more affected by climate policy than by climate change and consequently their WTP for climate change mitigation declines.
-On the other hand, individuals with lower level of understanding of inflation tend to perceive more effects from climate change than from climate policy and therefore their WTP increases.
-
-\vspace{1cm}
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{WTP - mediation through climate versus policy perception:}\\{WTP regressed on financial literacy (understands inflation) without the mediator, OLS}}}
-% soubor paper3/Rwork/mediation.R
-
-% soubor paper3/Rwork/mediation2.R
-%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
-
-\label{MediationTIV} 
-\centering
-\begin{small}
-\begin{tabular}{lrcrc} 
-\hline
- \multicolumn{1}{l}{\multirow{2}{*}{\textbf{Dependent variable:}}}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}
- &\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
-\cmidrule(lr){2-3} \cmidrule(lr){4-5}
-  \multicolumn{1}{l}{WTP-gas and electricity tax (\textsterling~/yr.)} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}} \\
-\hline  \vspace{-0.2cm}
-  \\
-\vspace{-0.2cm}\textbf{Understands inflation} $\boldsymbol{=1}$&${-73.648}$&${<2\times10^{-8}~^{***}}$&${-47.827}$&${<2\times10^{-8}~^{***}}$\\
-\\
-\vspace{-0.2cm}Understands inflation $=0.5$&$1.912$&$0.712$&$-16.889$&$0.0007~^{***}$\\
-\\
-\vspace{-0.2cm}Age\tnote{a} $~25-34$&\multicolumn{2}{c}{\textit{Not included}}&$-11.520$&$0.003~^{**}$\\
-        \\
-\vspace{-0.2cm}Age $35-44$&\multicolumn{2}{c}{\textit{Not included}}&$-27.881$&$<2\times10^{-8}~^{***}$\\
-        \\ 
- \vspace{-0.2cm}Age $45-54$&\multicolumn{2}{c}{\textit{Not included}}&$-33.602$&$<2\times10^{-8}~^{***}$\\
-        \\ 
-  \vspace{-0.2cm}Age $55-64$&\multicolumn{2}{c}{\textit{Not included}}&$-43.684$&$<2\times10^{-8}~^{***}$\\
-        \\       
- \vspace{-0.2cm}Age $65-74$&\multicolumn{2}{c}{\textit{Not included}}&$-50.493$&$<2\times10^{-8}~^{***}$\\
-        \\ 
- \vspace{-0.2cm}Age$~75$ or older&\multicolumn{2}{c}{\textit{Not included}}&$-33.163$&$0.008~^{**}$\\
-        \\       
-  \vspace{-0.2cm}Inequity aversion (categorical)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\\
-      \\
- Equal intergenerational&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$19.471$}&\multirow{2}{*}{$4\times10^{-6}~^{***}$}\\
-\vspace{-0.2cm}\hspace{0.6cm}allocation of resources $(0/1)$&\\
- \\  
-\vspace{-0.2cm}Understands compound interest $=0.5$&\multicolumn{2}{c}{\textit{Not included}}&$-5.751$&$0.431$ \\
-  \\
-\vspace{-0.2cm}Understands compound interest $=1$&\multicolumn{2}{c}{\textit{Not included}}&$-44.022$&$<2\times10^{-8}~^{***}$\\
-  \\  
-  Consistent answers to risk&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$-39.922$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
-\hspace{0.6cm}questions $(0/1)$&& &
-  \\
-\hline  \vspace{-0.2cm}
-  \\
-Adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.092$}&\multicolumn{2}{c}{$0.212$}\\
- Observations:&\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5749$}\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-  \begin{adjustwidth}{1cm}{}\item[a]Age is only available as categorical with  base category '$24$ or younger'.
-\item[b]Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}).
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\vspace{1cm}
-
-\pagebreak
-
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{WTP - mediation through climate versus policy perception:}\\{WTP regressed on the mediator (climate versus policy effects perception) without variable understands inflation, OLS}}}
-% soubor  paper3/Rwork/mediation.R
-
-% soubor paper3/Rwork/mediation2.R
-%jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
-\label{MediationTMV} 
-\centering
-\begin{small}
-\begin{tabular}{lrcrc} 
-\hline
- \multicolumn{1}{l}{\multirow{2}{*}{\textbf{Dependent variable:}}}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}
- &\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
-\cmidrule(lr){2-3} \cmidrule(lr){4-5}
-  \multicolumn{1}{l}{WTP - gas and electricity tax (\textsterling~/yr.)} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}} \\
-\hline  \vspace{-0.2cm}
- \\
-\textbf{Climate versus policy}&\multirow{2}{*}{$13.847$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}&\multirow{2}{*}{$10.936$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
-\vspace{-0.2cm}\hspace{0.6cm}\textbf{effects perception}&\\
- \\  
-\vspace{-0.2cm}Age\tnote{a} $~25-34$&\multicolumn{2}{c}{\textit{Not included}}&$-14.672$&$0.0001~^{***}$\\
-        \\
-\vspace{-0.2cm}Age $35-44$&\multicolumn{2}{c}{\textit{Not included}}&$-32.393$&$<2\times10^{-8}~^{***}$\\
-        \\ 
- \vspace{-0.2cm}Age $45-54$&\multicolumn{2}{c}{\textit{Not included}}&$-40.588$&$<2\times10^{-8}~^{***}$\\
-        \\ 
-  \vspace{-0.2cm}Age $55-64$&\multicolumn{2}{c}{\textit{Not included}}&$-47.725$&$<2\times10^{-8}~^{***}$\\
-        \\       
- \vspace{-0.2cm}Age $65-74$&\multicolumn{2}{c}{\textit{Not included}}&$-54.523$&$<2\times10^{-8}~^{***}$\\
-        \\ 
- \vspace{-0.2cm}Age$~75$ or older&\multicolumn{2}{c}{\textit{Not included}}&$-32.341$&$0.008~^{**}$\\
-        \\       
-  \vspace{-0.2cm}Inequity aversion (categorical)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
-      \\
- Equal intergenerational&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$21.644$}&\multirow{2}{*}{$1\times10^{-7}~^{***}$}\\
-\vspace{-0.2cm}\hspace{0.6cm}allocation of resources $(0/1)$&\\
- \\  
-\vspace{-0.2cm}Understands compound interest $=0.5$&\multicolumn{2}{c}{\textit{Not included}}&$0.396$&$0.956$\\
-  \\
-\vspace{-0.2cm}Understands compound interest $=1$&\multicolumn{2}{c}{\textit{Not included}}&$-40.066$&$<2\times10^{-8}~^{***}$\\
-  \\
-  Consistent answers to risk&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$-40.935$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
-\hspace{0.6cm}questions $(0/1)$&& &
-     \\
-\hline  \vspace{-0.2cm}
-  \\
-Adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.092$}&\multicolumn{2}{c}{$0.244$}\\
- Observations:&\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5749$}\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-  \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-  \begin{adjustwidth}{1cm}{}\item[a]Age is only available as categorical with  base category '$24$ or younger'.
-\item[b]Inequity aversion is treated as categorical (see Section~\ref{BehaviouralVars}).
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\newpage
-
-\FloatBarrier
-
 %TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT
@@
 \bibliographystyle{ChicagoM}
-\bibliography{references}
+%\bibliographystyle{apa}
+\bibliography{referencesFS}
 
@@
 
-\section{Appendix 1 Climate knowledge - OCSI instrument}
-
-Below are the questions of the Ordinary Climate-Science Intelligence (OCSI) assessment developed by \cite{Kahan2015} which we use as a measure of climate knowledge.
-The eight questions are true or false statements.
-The correct answers are in bold.
-
-\begin{enumerate}
-
-\item Climate scientists believe that if the North Pole icecap melted as a result of human-caused global warming, global sea levels would rise.  \textbf{FALSE}
-
-\item Climate scientists have concluded
-that globally averaged surface air
-temperatures were higher for the 
-first decade of the twenty-first
-century (2000-2009) than for the last decade of the twentieth
-century (1990-1999).   \textbf{TRUE}
-
-\item Climate scientists believe that
-human-caused global warming
-will result in flooding of many
-coastal regions. \textbf{TRUE}
-
-\item Climate scientists believe that
-human-caused global warming
-has increased the number and
-severity of hurricanes around the
-world in recent decades. \textbf{FALSE}
-
-\item Climate scientists believe that
-nuclear power generation
-contributes to global warming.  \textbf{FALSE}
-
-\item Climate scientists believe that
-human-caused global warming
-will increase the risk of skin cancer in human beings.  \textbf{FALSE}
-
-\item Climate scientists and economists
-predict there will be positive as
-well as negative effects from
-human-caused global warming. \textbf{TRUE}
-
-\item Climate scientists believe that the
-increase of atmospheric carbon
-dioxide associated with the
-burning of fossil fuels will reduce
-photosynthesis by plants.  \textbf{FALSE}
-
-\end{enumerate}
-
-\section{Appendix 2 Climate knowledge and gender}
-
-As discussed in Section~\ref{ResKnow}, we detect a strong evidence that our measure of climate knowledge is significantly higher for men than for women.
-We find this outcome merits further investigation.
-
-We hypothesise that less educated men can have higher drop out rates from the survey than more educated men or less educated women.
-In other words, we believe that it can be more likely for men to abandon the whole survey if they find a series of questions to be too difficult to respond while women answer giving their best guess even if they are uncertain and continue with the survey.
-This could be caused by different opportunity costs, effect of pride or by males perceiving higher pressure to answer scientific questions correctly.
-If this is the case, our sample of complete cases will exhibit a selection bias as the ratio of less educated women will be bigger for the subsample of complete cases.
-
-To test for presence of the selection bias, we perform a series of following proportion tests.
-For each category of education (and also for the whole sample) we test whether the proportion of males in the subsamle of complete observations (used observations) is approximately equal to the proportion of males in the subsample of dropped observations.
-The \textit{p}-values of the corresponding Pearson's chi-square test statistics of the null hypothesis that the proportions are equal are summarised in Table~\ref{Proptests}.
-
-If the selection bias occurs, we would expect for the lower education categories the proportion of males to be significantly higher among the dropped observations than among the used observations.
-For the higher categories of education, on the other hand, we would expect the proportion of males to be significantly smaller among the dropped observation than among the used observations.
-However, this is not what we can see in Table~\ref{Proptests}.
-Although the proportion tests are significant for some GCSE, GCSE and professional, the differences in proportions are opposite to what we would expect.
-For the lower categories of education (Some GCSE and GCSE) the proportion of males is smaller among the dropped observations than among the used observations while it is the other way around for the category of professionals.
-Hence, based on the proportion tests, we do not see any evidence of the selection bias.\footnote{We performed analogous series of tests to compare proportions of males among the dropped observations with the proportions of males in the whole sample.
-Also these tests were performed separately for each category of education.
-The results are qualitatively equivalent to those of the tests in Table~\ref{Proptests}.
-The only difference is that the test became marginally insignificant for the category professional but this has no effect on the conclusion.}
-
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Proportion tests - no selection bias:}\\~ \\ Differences between ratio of males in the group of used observations and in the group of dropped observations.
-The tests were conducted separately for each category of education.}}
-
-% soubor heckmanNew/droppedDescriptive
-\label{Proptests} 
-\centering
-\begin{small}
-\begin{tabular}{l|cccl} 
-  \multicolumn{5}{l}{ } \\ \hline
-  \multicolumn{1}{l}{ } & \multicolumn{2}{|c}{\textbf{Proportion of males}} &  \multicolumn{2}{c}{\boldsymbol{${\tilde{\chi}^2}$}} \\
-   \multicolumn{1}{l}{\textbf{Education category}} &  \multicolumn{1}{|l}{\textbf{Used observations}} &  \multicolumn{1}{c}{\textbf{Dropped observations}} & \multicolumn{2}{c}{\textbf{\textit{p}-value}}\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Total&$0.4858$&$0.4538$&$0.0058$&$^{**}$\\
-  \\
-\vspace{-0.2cm}Craft&$0.6173$&$0.6620$&$0.4671$&\\
-  \\
-\vspace{-0.2cm}Some GCSE&$0.4957$&$0.4105$&$0.0022$&$^{**}$\\
-  \\
-\vspace{-0.2cm}GCSE&$0.4861$&$0.3792$&$0.0066$&$^{**}$\\
-  \\
-\vspace{-0.2cm}A levels&$0.4859$&$0.4671$&$0.5383$&\\
-  \\
-\vspace{-0.2cm}Diploma&$0.4609$&$0.4172$&$0.2306$&\\
-  \\
-  \vspace{-0.2cm}Bachelors&$0.4662$&$0.4294$&$0.2470$&\\
-\\
-  \vspace{-0.2cm}Professional&$0.4269$&$0.5391$&$0.0474$&$^{*}$\\
-\\
-  \vspace{-0.2cm}Masters&$0.4829$&$0.4000$&$0.1240$&\\
-  \\
-  \vspace{-0.2cm}PhD&$0.6522$&$0.5625$&$0.4889$&\\
-\\
-  \vspace{-0.2cm}No answer&$0.5435$&$0.4920$&$0.3355$&\\
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-\item[~]\textit{Notes:} \hspace{0.24cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-\newpage
-
-\vspace{1cm}
-
-To further eliminate occurrence of the selection bias, we include interactions of gender and the education categories for which the proportion tests are significant as explanatory variables besides the predictors selected by the lasso and we test their significance.
-For the sake of clearer interpretation, we also include the main effects of the education categories.
-The estimates of the model with the interactions are summarised in Table~\ref{Know4235f}.
-None of the interactions or education categories are significant and the signs and significance levels of male and gender are the same as without the interactions.\footnote{We also estimated a version of this model which includes a dummy variable for each education category.
-Adding these dummy variables does not chance signs or significance levels of gender, cognitive reflection or interactions between gender and education categories.}
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Climate change knowledge:} Jackknife OLS \\ With interactions of gender and education}}
-% soubor ResKnow4235f 
-\label{Know4235f} 
-\centering
-\begin{small}
-\begin{tabular}{lrcl} 
-\hline \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\
-
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-   \multicolumn{1}{l}{ } &  \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adjusted \textit{p}-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Gender = male&$0.3379$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.1967$&$1.0000$&\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.1195$&$0.4384$& \\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6160$&$1.0000$ \\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$0.2664$&$0.0001$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.4405$&$1.0000$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4551$&$2.31\times 10^{-8}$&$^{***}$\\
-\\
-  \vspace{-0.2cm}Education - some GCSE&$-0.0890$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Education - GCSE&$0.0033$&$1.0000$&\\
-  \\
-  \vspace{-0.2cm}Education -  professional&$-0.0219$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - some GCSE&$-0.0494$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - GCSE&$-0.0888$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education -  professional&$0.2377$&$1.0000$&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\vspace{1.5cm}
-
-As an additional verification that our results can not be attributed to a selection we estimate a Heckman correction models for climate knowledge.
-In particular, we estimate models which are referred to as Tobit-$2$ \citep{HeckmanR}.
-The exclusion restriction is count of not responded questions out of those which were prior to climate questions in the survey questionnaire.\footnote{Count of previously not responded questions is not expected to affect climate knowledge.
-In spite of this, we estimated a variant of multisplit lasso with the count of previously not responded questions as a potential predictor to verify whether or not it should be in the climate knowledge equation according to our estimation method.
-As we expected, count of previously not responded questions was not selected.} The estimates are summarized in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}.
-The model in Table~\ref{KnowHeckman1} has only one explanatory variable in the selection equation, namely count of non-responded questions.
-The variety in Table~\ref{KnowHeckman2} includes also gender, education categories for which the proportion test in Table~\ref{Proptests} is significant and their interactions.
-The outcome equations include the predictors which were selected by the multisplit lasso (see Section~\ref{ResKnow}).
-We can see, that the male dummy variable is still positive and strongly significant in the outcome equations even if we correct for possible selection bias (see Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}).
-Parameter $\rho$ is insignificant in the models in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}.
-This means that the data are consistent with no correlation of the selection and outcome equation.
-
-We can further see in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2} that achieving score $2$ or $3$ in the cognitive reflection test has positive and significant impact on climate knowledge which is consistent with the model presented in Table~\ref{Know435} in Section~\ref{ResKnow} and with the specification in Table~\ref{Know4235f}.
-In addition, achieving score $1$ is significant in the Heckman models and score of $0.5$ is close to significant.\footnote{Besides the Heckman selection models presented in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}, we also estimated a version which includes all education categories and their interactions with gender as explanatory variables in the selection equation.
-However, the estimation algorithm was unable to estimate the coefficients with reasonable standard errors.}
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Climate change knowledge:} Heckman selection model}}
-% soubor  N:US/paper3/Rwork/heckmanNew/TheBest2.
-% model Claudell12
-\label{KnowHeckman1} 
-\centering
-\begin{small}
-\begin{tabular}{lrcl} 
-\hline \vspace{-0.3cm} \\
-  \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.25cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Estimate}} &  \multicolumn{2}{c}{\textbf{\textit{p}-value}} \\
- \hline 
-\hline
- \vspace{-0.25cm} \\
-   \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Probit selection equation:}}\\
-   \hline
-   \\
-\vspace{0.1cm}Not responded questions (count)&$-0.1972$ $(0.0083)$ &$<2.00\times 10^{-8}$&$^{***}$ \\
- \hline
-  \vspace{-0.25cm} \\
-  \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Outcome equation}}\\
-   \hline
-      \\
-      \vspace{-0.2cm}Gender = male&$0.3304$ $(0.0304)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.2188$ $(0.7202)$&$0.0906$&$^{\bullet}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.1472$ $(0.0387)$&$0.0001$&$^{***}$\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6092$ $(0.8821)$&$0.4898$\\
-   \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$0.3026$ $(0.0448)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.5158$ $(0.4163)$&$0.2153$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4790$ $(0.0559)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
- \hline
-  \vspace{-0.25cm} \\
-  \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Error terms}}\\
-   \hline
-      \\
-      \vspace{-0.2cm}Sigma $\sigma$&$1.2469$ $(0.0107)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Rho $\rho$&$0.0611$ $(0.0659)$&$0.3540$&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$7244$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\item[]   \begin{adjustwidth}{1cm}{}Standard errors in brackets
-     \end{adjustwidth}
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-\vspace{1cm}
-
-{\centering \singlespacing
-\begin{threeparttable}
-\caption{\textit{\textbf{Climate change knowledge:} Heckman selection model \\ With interactions of gender and education}}
-% soubor  N:US/paper3/Rwork/heckmanNew/TheBest2.
-% model Felicia12
-\label{KnowHeckman2} 
-\centering
-\begin{small}
-\begin{tabular}{lrcl} 
-\hline \vspace{-0.3cm} \\
-  \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.25cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Estimate}} &  \multicolumn{2}{c}{\textbf{\textit{p}-value}} \\
- \hline 
-\hline
- \vspace{-0.25cm} \\
-   \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Probit selection equation:\tnote{a}}}\\
-   \hline
-   \\
- \vspace{-0.2cm}Gender = male&$-0.0163$ $(0.0954)$ &$0.8642$&\\
-  \\
- \vspace{-0.2cm}Not responded questions (count)&$-0.1983$ $(0.0084)$ &$<2.00\times 10^{-8}$&$^{***}$ \\
-\\
- \vspace{-0.2cm}Education - some GCSE&$-0.2250$ $(0.1213)$ &$0.0637$&$^{\bullet}$ \\
-\\
-  \vspace{-0.2cm}Education - GCSE&$0.0783$ $(0.1658)$ &$0.6369$&\\
-  \\
-  \vspace{-0.2cm}Education -  professional&$-0.2696$ $(0.2379)$ &$0.2570$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - some GCSE&$0.2926$ $(0.1934)$ &$0.1303$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - GCSE&$-0.3454$ $(0.2571)$ &$0.1790$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education -  professional&$0.3659$ $(0.3468)$ &$0.2914$&\\
- \\
-   \hline  
-  \multicolumn{4}{c}{\textbf{Outcome equation}}\\
-     \hline
-      \\
-      \vspace{-0.2cm}Gender = male&$0.3307$ $(0.0304)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.2199$ $(0.7202)$&$0.0903$&$^{\bullet}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.1476$ $(0.0387)$&$0.0001$&$^{***}$\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6102$ $(0.8821)$&$0.4891$\\
-   \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$0.3032$ $(0.0448)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.5168$ $(0.4163)$&$0.2144$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4796$ $(0.0559)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\ \hline  
-  \multicolumn{4}{c}{\textbf{Error terms}}\\
-   \hline  
-      \\
-      \vspace{-0.2cm}Sigma $\sigma$&$1.2469$ $(0.0107)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Rho $\rho$&$0.0737$ $(0.0661)$&$0.2640$&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$7244$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
- \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\item[]   \begin{adjustwidth}{1cm}{}Standard errors in brackets \vspace{0.2cm}
- \item[a]~We also estimated a variety of this model which includes dummy variables for all education categories in the selection equation.
-They are all insignificant and the signs and significance levels of the other variables are the same.
-     \end{adjustwidth}  
- \singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
+\section{Appendix 1 Drought indices and measures}
+
+Examples of early measures of drought are \cite{wilhite1985}, \cite{munger1916}, \cite{blumenstock1942} or \cite{mcquigg1954}.
+\cite{munger1916} suggested to use length of period without $24$-h precipitation of $1.27$ mm.
+\cite{wilhite1985} is based on a measure of precipitation over a given time period.
+\cite{blumenstock1942} proposed to measure severity of drought as a length of drought in days where the end of a drought is defined by occurrence of $2.54$ mm of precipitation in $48$ hours.
+\cite{mcquigg1954} developed the Antecedent Precipitation Index (API) which is based on amount and timing of precipitation and it was used for forecasting of floods.
+Hence, the API is a reverse drought index.
+
+The study of \cite{palmer1965} was a significant milestone in the history of quantification of drought severity.
+\cite{palmer1965} developed the Palmer Drought Severity Index (PDSI) using a complex water balance model.
+The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand (\citealp{Heim2002,palmer1965}).
+As the PDSI suffers from several weaknesses (for details see e.g.
+\citealt{Heim2002}), other indices were developed in the following decades.
+These include the standardized precipitation index (SPI) developed by \cite{SPI} and the standardized precipitation evapotranspiration index (SPEI) developed by \cite{SPEI}.
+The SPI specifies observed precipitation as a standardised departure from a chosen probability distribution which models the precipitation data.
+Values of SPI can be viewed as a multiple of standard deviations by which the observed amount of rainfall deviates from the long-term mean \citep{SPIonline}.\footnote{Can be created for various periods of 1-36 months, usually using monthly data.} The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of evapotranspiration (which captures increased temperature).
+It is based on water balance, therefore it can be compared to the self-calibrated PDSI \citep{SPEI}.
 
@@
 
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Frequency tables:} Categorical variables}}
-\label{FreqiencyCat} 
-\centering
-\begin{small}
-\begin{tabular}{llrr} 
-\hline \vspace{-0.15cm} \\	
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}}& \multicolumn{1}{c}{\textbf{Category}} &  \multicolumn{1}{c}{\textbf{Frequency}}  &  \multicolumn{1}{c}{\textbf{Ratio}}   \\ 
-\hline \vspace{-0.3cm} \\ 
- \hline  \vspace{-0.15cm}\\
-  \vspace{0.15cm}Education&Craft&$338$&$0.040$\\
-  \vspace{0.15cm}&Some GCSE&$1452$&$0.170$\\
-  \vspace{0.15cm}&GCSE A*-C grades&$814$&$0.095$\\
-  \vspace{0.15cm}&A Level&$1579$&$0.185$\\
-  \vspace{0.15cm}&Diploma&\multirow{1}{*}{$979$}&\multirow{1}{*}{$0.115$}\\
-    \vspace{0.15cm}&Bachelor's degree&\multirow{1}{*}{$1523$}&\multirow{1}{*}{$0.178$}\\
-  \vspace{0.15cm}&Professional qualifications&\multirow{1}{*}{$457$}&\multirow{1}{*}{$0.054$}\\
-  \vspace{0.15cm}&Master's degree&$564$&$0.066$\\
-  \vspace{0.15cm}&PhD, DPhil&$124$&$0.015$\\
-  \vspace{0.15cm}&Prefer not to say&$434$&$0.051$\\
-    \vspace{0.15cm}&NA&$277$&$0.032$\\
-  \hline  \vspace{-0.15cm} \\
-    \vspace{0.15cm}Household income &$<11000$&$919$&$0.158$\\
- \vspace{0.15cm}   pounds per year&$11000 - 16000$&$675$&$0.116$\\
-\vspace{0.15cm}  before tax &$16000-20000$&$539$&$0.093$\\
- \vspace{0.15cm} self reported&$20000-26000$&$757$&$0.130$\\
- \vspace{0.15cm}  &$26000-32000$&$610$&$0.105$\\
- \vspace{0.15cm}  &$32000-39000$&$666$&$0.115$\\
- \vspace{0.15cm}  &$39000-48000$&$522$&$0.090$\\
- \vspace{0.15cm}  &$48000-60000$&$544$&$0.094$\\
- \vspace{0.15cm}  &$60000-81000$&$324$&$0.056$\\
- \vspace{0.15cm}  &$81000-100000$&$119$&$0.021$\\
- \vspace{0.15cm}  &$>100000$&$128$&$0.022$\\
-\hline
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-     \item[~]\textit{Notes:} Total number of observations: $8541$\vspace{-0.35cm} \begin{adjustwidth}{0.7cm}{}  
-  \item[~]\hfill (\textit{continued})
-    \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\pagebreak
-
-{\centering
-\begin{threeparttable}
-\caption{\vspace{-0.27cm} \small{\textit{\textbf{Frequency tables:} Categorical variables}}}
-\label{FreqiencyCat2} 
-\centering
-\begin{small}
-\begin{tabular}{l|crr} 
-\hline \vspace{-0.27cm} \\	
-  \multicolumn{1}{l|}{\textbf{Variable}}& \multicolumn{1}{c}{\textbf{Category}} &  \multicolumn{1}{c}{\textbf{Frequency}}  &  \multicolumn{1}{c}{\textbf{Ratio}}   \\ 
-\hline \vspace{-0.27cm}\\
-  Inequity aversion&$0.520$&$1557$&$0.182$\\
- \hspace{0.15cm}(rate) &$0.950$&$55$&$0.006$\\
- \hspace{0.15cm}\citet{bergson1954, bergson1938}, &$1.000$&$652$&$0.076$\\
- \hspace{0.15cm}\citet{samuelson1956} &$1.135$&$127$&$0.015$\\
-   &$1.160$&$364$&$0.043$\\
-    &$1.255$&$202$&$0.024$\\
-    &$1.290$&$130$&$0.015$\\ 
-    &$1.485$&$385$&$0.045$\\
-    &$1.490$&$288$&$0.034$\\
-    &$1.500$&$96$&$0.011$\\ 
-     &$1.510$&$86$&$0.010$\\ 
-    &$1.685$&$202$&$0.024$\\
-    &$1.765$&$93$&$0.011$\\
-    &$2.120$&$226$&$0.026$\\ 
-    &$3.640$&$59$&$0.007$\\
-    &$3.710$&$2551$&$0.299$\\
-    &NA&$1468$&$0.172$\\ 
- \hline  \vspace{-0.27cm}\\
- Degree of agreement&\multicolumn{1}{l}{Strongly disagree}&$555$&$0.065$\\ 
- \hspace{0.15cm}with the statement:&\multicolumn{1}{l}{Disagree}&$1121$&$0.131$\\ 
-  \textit{ Government should redistribute}&\multicolumn{1}{l}{Neutral}&$1952$&$0.229$\\ 
-  \textit{ income from the better off}&\multicolumn{1}{l}{Agree}&$2256$&$0.264$\\ 
-  \textit{ to those who are less well off.}&\multicolumn{1}{l}{Strongly agree}&$1206$&$0.141$\\ 
-&\multicolumn{1}{l}{NA}&$1451$&$0.170$\\ 
-  \hline  \vspace{-0.27cm}\\
-
-Cognitive reflection test\tnote{a}&\multicolumn{1}{c}{$0.0$}&$4145$&$0.485$\\ 
- \hspace{0.25cm}= numeracy, \citet{Frederick2005} &\multicolumn{1}{c}{$0.5$}&$3$&$0.0004$\\ 
-   \hspace{0.25cm}$3$ numerical problems&\multicolumn{1}{c}{$1.0$}&$1519$&$0.178$\\ 
-  \hspace{0.25cm}no. of correct answers&\multicolumn{1}{c}{$1.5$}&$2$&$0.0002$\\ 
-&\multicolumn{1}{c}{$2.0$}&$1017$&$0.119$\\ 
-&\multicolumn{1}{c}{$2.5$}&$9$&$0.001$\\ 
-&\multicolumn{1}{c}{$3.0$}&$614$&$0.072$\\ 
-&\multicolumn{1}{c}{NA}&$1232$&$0.144$\\
-
- \hline  \vspace{-0.27cm}\\
-Understands compound interest&\multicolumn{1}{c}{$0.0$}&$291$&$0.034$\\ 
-  \hspace{0.25cm}$1$ = Understands&\multicolumn{1}{c}{$0.5$}&$718$&$0.084$\\ 
-    \hspace{0.25cm}(treated as categorical)&\multicolumn{1}{c}{$1.0$}&$5713$&$0.669$\\ 
-&\multicolumn{1}{c}{NA}&$1819$&$0.213$\\ 
- \hline  \vspace{-0.27cm}\\
-Understands inflation&\multicolumn{1}{c}{$0.0$}&$941$&$0.110$\\ 
-  \hspace{0.25cm}$1$ = Understands&\multicolumn{1}{c}{$0.5$}&$953$&$0.112$\\ 
-    \hspace{0.25cm}(treated as categorical)&\multicolumn{1}{c}{$1.0$}&$4206$&$0.492$\\ 
-&\multicolumn{1}{c}{NA}&$2441$&$0.286$\\ 
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-     \item[~]\textit{Notes:}  \textbf{a} This variable is called \textit{Cognitive reflection} in the tables with regression
-          \begin{adjustwidth}{0.7cm}{}  
-  estimates and it is treated as categorical \hfill $8541$ observations
-    \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\pagebreak
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Frequency tables:} Binary variables}}
-\label{FreqiencyBin} 
-\centering
-\begin{small}
-\begin{tabular}{lrrr} 
-\hline \vspace{-0.15cm} \\	
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable} }&\multicolumn{1}{l}{ \textbf{Frequency $\boldsymbol{= 1$}}} &\multicolumn{1}{l}{  \textbf{Ratio $\boldsymbol{= 1$}}}  &\multicolumn{1}{c}{\textbf{NA's}} \\
-\hline \vspace{-0.3cm} \\ 
-  \vspace{0.15cm}Gender = male &$4060$&$0.475$&$0$\\
-
-Equal intergenerational allocation&\multirow{2}{*}{$793$}&\multirow{2}{*}{$0.110$}&\multirow{2}{*}{$1339$}\\
-\vspace{0.15cm} \hspace{0.5cm}of resources (agree = 1)\tnote{a}& \\
-  \vspace{0.15cm}Consistent answers to risk questions (consistent = 1)&$7153$&$0.837$&$0$\\
-   \vspace{0.15cm}Consist. answers within investment (consistent = 1)&$981$&$0.115$&$0$\\
-\hline
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-     \item[~]\textit{Notes:} Total number of observations: $8541$\vspace{-0.35cm} \begin{adjustwidth}{0.7cm}{}  
- \item[a]~This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age).
-The variable is equal to $0$ for all other respondents.
-    \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\vspace{2cm}
-
-\pagebreak
-
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Climate knowledge:} Jackknife OLS with total score on financial literacy}}
-% table checked 3.10.2017
-
-%musim lasso predelat. predtim jsem pouzivala lambda min. ale alex ma pravdu. je treba u vsech modelu pouzivat stejne pravidlo, takze lambda se.zmena 30.9.2017
-
-%soubor Know435se.R
-% finlit2 prestane byt signifikantni, takze vlastne se nic skoro nemeni, asi do prilohy tabulku s Know435se.R
-
-\label{Know435se} 
-\centering
-\begin{small}
-\begin{tabular}{lclrc} 
-   \hline
-
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}}&  \multicolumn{1}{c}{\textbf{Aggregated}}  \\
-    \multicolumn{1}{l}{ } &  \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} &  \multicolumn{1}{c}{\textbf{VIF}}\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Gender = male&$1.580$&$<2\times10^{-8}$&$^{***}$&$1.030$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$4.611$&$1.000$&&$1.002$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.431$&$1.000$&&$1.156$\\
-\\
-\vspace{-0.2cm}Cognitive reflection $=1.5$&$1.681$&$1.000$&&$1.002$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$1.074$&$0.006$&$^{**}$&$1.212$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$1.881$&$1.000$& &$1.006$\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$1.999$&$9\times10^{-7}$&$^{***}$&$1.171$\\
-\\
-\vspace{-0.2cm}Financial literacy - total score $=0.5$&$1.173$&$1.000$& &$1.504$\\
-  \\
-\vspace{-0.2cm}Financial literacy - total score $=1$&$0.456$& $1.000$& &$3.972$\\
-  \\
-\vspace{-0.2cm}Financial literacy - total score $=1.5$&$0.527$& $1.000$&&$2.586$\\
-  \\
-\vspace{-0.2cm}Financial literacy - total score $=2$&$0.454$& $1.000$&&$5.644$\\
-  \\
-\vspace{-0.2cm}Financial literacy - total score $=2.5$&$0.802$& $1.000$&&$2.435$\\
-  \\
-  \vspace{-0.2cm}Financial literacy - total score $=3$&$1.433						$& $0.150$&&$6.925$\\
- \\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{4}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-    \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\pagebreak
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{{Climate seriousness and climate versus policy effects perception:} Jackknife OLS without climate knowledge}}
-\label{ClimCare435eClimpol435f} 
-\centering
-\begin{small}
-\vspace{-1.2cm}
-\begin{tabular}{lcccc} 
-\hline
-  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Seriousness}}&\multicolumn{2}{c}{\textbf{Climate vs. policy}}\\
-\cmidrule(lr){2-3} \cmidrule(lr){4-5}
-  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
-    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
-        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
- \hline 
-\vspace{-0.23cm}
-\\
-\vspace{-0.2cm}Gender = male&$-0.481$&$<2\times10^{-8}~^{***}$&\multicolumn{2}{c}{\textit{Not included}}\\
-\\
- Redistribution of income:&\multirow{2}{*}{$0.201$}&\multirow{2}{*}{$1.000$}&  \multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}} \\%-1
-    \vspace{-0.2cm}   \hspace{0.6cm}disagree\tnote{a}&&&&\\%-1
-    \\
-  \vspace{-0.2cm}Redistribution of inc.: neutral\tnote{a}&$0.331$&$0.265$ &\multicolumn{2}{c}{\textit{Not included}}\\%0
-    \\
-  \vspace{-0.2cm}Redistribution of income: agree\tnote{a}&$0.891$&$<2\times10^{-8}~^{***}$&\multicolumn{2}{c}{\textit{Not included}}\\%1
-    \\
-  Redistribution of income:& \multirow{2}{*}{$1.172$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}& \multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}} \\
-\hspace{0.6cm}strongly agree\tnote{a}&&&\\%2
-\vspace{-0.2cm}
- \\
-\vspace{-0.2cm}Understands inflation $=0.5$&\multicolumn{2}{c}{\textit{Not included}}&$-0.049$&$1.000$ \\
-  \\
-\vspace{-0.2cm}Understands inflation $=1$  &\multicolumn{2}{c}{\textit{Not included}}&$-0.643$& $2\times10^{-7}~^{***}$\\
-    \\
-  Consistent answers to risk& \multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}} &\multirow{2}{*}{$-0.592$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
-\hspace{0.6cm}questions $(0/1)$&& &\\
-   \\
-     \hline
-  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5749$}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-
- \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
- \begin{adjustwidth}{0.9cm}{}
-  \vspace{-0.3cm}
-  \item[a]~Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' The base category is 'Strongly disagree'.
-
-     \end{adjustwidth}    
-\singlespacing  
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\vspace{1cm}
-
-\pagebreak
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\small{\textbf{\vspace{-0.25cm} WTP climate:} interaction of cultural world-view and financial literacy}}}
-% soubor  Rwork/GasSuperNew/Gas435dcq.R
-\label{Gas435dcp} 
-\centering
-\begin{small}
-\begin{tabular}{lccl} 
-\hline \vspace{-0.15cm} \\
-  \multicolumn{1}{l}{} 
-  & \multicolumn{3}{c}{\large{\textbf{\small{Jackknife OLS}}}}  \\
-\\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Dependent variable:}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-  \multicolumn{1}{l}{WTP - gas and electricity tax (\textsterling~per year)} & \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\   
-\hline \vspace{-0.3cm} \\ 
- \vspace{0.15cm}Age\tnote{a} $~25-34$&$-11.3886$&$1.0000$&\\
-  \vspace{0.15cm}Age $35-44$&$-27.1860$&$4\times10^{-5}$&$^{***}$\\
- \vspace{0.15cm}Age $45-54$ &$-34.3795$&$2\times10^{-8}$&$^{***}$\\
- \vspace{0.15cm}Age $55-64$ &$-37.5114$&$<2\times10^{-8}$&$^{***}$\\
- \vspace{0.15cm}Age $65-74$&$-45.2613$&$<2\times10^{-8}$&$^{***}$\\
- \vspace{0.15cm}Age $74$ or older &$-28.8122$&$1.0000$&\\     
- \vspace{0.15cm}Climate versus policy effects perception&$10.2983$&$<2\times10^{-8}$&$^{***}$\\
-\vspace{0.15cm}Inequity aversion (categorical)\tnote{b}& \multicolumn{2}{c}{\textit{negative correlation}}&$^{**}$\\
-Equal intergenerational&\multirow{2}{*}{$21.8916$}&\multirow{2}{*}{$0.0034$}&\multirow{2}{*}{$^{*}$}\\
- \vspace{0.15cm}\hspace{0.6cm}allocation of resources $(0/1)$\tnote{c}&\\
- \vspace{0.15cm}Understands compound interest $=0.5$&\multirow{1}{*}{$-4.3717$}&\multirow{1}{*}{$1.0000$}&\multirow{1}{*}{ }\\
- \vspace{0.15cm}Understands compound interest $=1$&\multirow{1}{*}{$-41.6926$}&\multirow{1}{*}{$5\times10^{-5}$}&\multirow{1}{*}{$^{***}$}\\
-  \vspace{0.15cm}Understands inflation $=0.5$&$-17.6328$&$0.1616$& \\
- \vspace{0.15cm}Understands inflation $=1$&$-46.0609$&$<2\times10^{-8}$&$^{***}$\\
-
-\vspace{0.15cm}Consistent answers to risk questions $(0/1)$&\multirow{1}{*}{$-35.4773$}&\multirow{1}{*}{$<2\times10^{-8}$}&\multirow{1}{*}{$^{***}$}\\
-\vspace{0.15cm}Redistribution of income (degree of agreement)\tnote{d, e}&\multirow{1}{*}{$-9.6596$}&\multirow{1}{*}{$0.0670$}&\multirow{1}{*}{$^{\bullet}$}\\
-
- \vspace{0.15cm}\textbf{Redistribution of income\tnote{d} \hspace{0.1cm}$\boldsymbol{\times}$\hspace{0.1cm}Understands inflation}&$13.7064$&$0.0043$&$^{**}$
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-     \item[~]\textit{Notes:} \vspace{-0.35cm} \begin{adjustwidth}{0.7cm}{}  \hspace{0.7cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$;\hspace{0.2cm} mean adj.
-\textit{R\textsuperscript2}:\hspace{0.25cm}$0.279$
-\item[a]~Age is only available as categorical. The base category is '$24$ or younger'.
-\item[b]~Inequity aversion treated as categorical (see Section~\ref{BehaviouralVars}).
-\item[c]~This variable is equal to $1$ for those respondents who believe that their income and standard of living generally is about equal as the income and standard of living of their parents (when they were about the respondent's age) and it is also equal to the income and standard of living of their children (when they will reach the respondent's age).
-The variable is equal to $0$ for all other respondents.
-\vspace{0.002cm} 
- \item[d]~Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' $-2=$~Strongly disagree, $2=$~Strongly agree.
-  \item[e]~For the sake of simplicity we also included the main effect.
-    \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\small\textit{\textbf{Climate vs. policy perception:} Jackknife OLS - robustness without WTP}}
-% soubor Cpol435InOutbNoWTP.R  Cpol535InONoWTP.R
-\label{ClimPolIOnoWTP} 
-\centering
-\begin{small}
-\vspace{-1.2cm}
-\begin{tabular}{lcccc} 
-\hline
-  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
-\cmidrule(lr){2-3} \cmidrule(lr){4-5}
-  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
-    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
-        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
- \hline 
-\vspace{-0.23cm}
-\\
-\vspace{-0.33cm}Climate knowledge\tnote{a}&$-0.130$&$9\times10^{-5}~^{***}$&$-0.131$&$0.0002~^{***}$\\
-  \\
-\vspace{-0.33cm}Understands inflation $=0.5$&$0.043$&$1.000$&$0.021$&$1.000$ \\
-  \\
-\vspace{-0.33cm}Understands inflation $=1$&$-0.450$&$0.001~^{**}$&$-0.455$&$0.001~^{**}$\\
-    \\
-  Consistent answers to risk&\multirow{2}{*}{$-0.496$}&\multirow{2}{*}{$3\times10^{-7}~^{***}$}&\multirow{2}{*}{$-0.482$}&\multirow{2}{*}{$1\times10^{-6}~^{***}$}\\
-\hspace{0.6cm}questions $(0/1)$&& &\\
-\hline
-  \\
-\vspace{-0.33cm}Income- predicted (mill. \textsterling~/yr.)&$3.347$&$1.000$&\multicolumn{2}{c}{\textit{Not included}}\\ %income PREDICTED!!
-  \\
-\vspace{-0.33cm}Income- reported (mill. \textsterling~/yr.)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\textit{varies}&$1.000$\\ %income PREDICTED!!
-  \\
-\vspace{-0.33cm}Net assets (million \textsterling)&$-0.014$&$1.000$&$0.076$&$1.000$\\ 
-      \\
-\vspace{-0.33cm}People per mill.
-km\textsuperscript{2}-LSOA level&\multirow{1}{*}{$-15.303$}&\multirow{1}{*}{$1.000$}&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}\\%ha is a symbol for hectares
-  \\
-\vspace{-0.33cm}People per mill.
-km\textsuperscript{2}-LAD level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$20.682$}&\multirow{1}{*}{$1.000$}\\%ha is a symbol for hectares
-  \\
-    \vspace{-0.33cm}Climate seriousness perception&$0.427$&$\small{<}2\times10^{-8}~^{***}$&$0.432$&$\small{<}2\times10^{-8}~^{***}$\\
-      \\
-Social value orientation&\multirow{2}{*}{$0.006$}&\multirow{2}{*}{$0.250$}&\multirow{2}{*}{$0.006$}&\multirow{2}{*}{$0.555$}\\
-\vspace{-0.33cm} \hspace{0.6cm}(ring measure)&& &\\%2
-  \\% ring measure> ref working paper dolton tol
- \vspace{-0.33cm}\multirow{1}{*}{Inequity aversion (categorical)}&\textit{varies}& \multirow{1}{*}{$1.000$}& \multirow{1}{*}{\textit{varies}}& \multirow{1}{*}{$1.000$}\\      
-    \\ \vspace{-0.33cm}Discount rate yr. from now&$-0.002$&$1.000$&$-0.001$&$1.000$\\  
-      \\
- \vspace{-0.33cm}Discount rate yr. from now - sq.&\multirow{1}{*}{$3\times10^{-6}$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$3\times10^{-6}$}&\multirow{1}{*}{$1.000$}\\
-  \\  \vspace{-0.33cm}Risk aversion coefficient\tnote{c}&\multicolumn{2}{c}{\textit{Not included}}&$-0.569$&$1.000$\\          
-   \\
-  \vspace{-0.35cm}Redistribution of income (categ.)\tnote{d}&\textit{$+$,varies}&$1.000$&\textit{$+$,varies}&$1.000$\\%1
-    \\
-     \hline
-          Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.226$}&\multicolumn{2}{c}{$0.230$}\\
-  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
- \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
- \begin{adjustwidth}{0.9cm}{}
-  \vspace{-0.3cm}
- \item[a]~Squared term of climate knowledge is insignificant in this version, hence it is not included.
-\item[b]~Self reported income is only available as categorical.
-\item[c]~The risk aversion coefficient is an estimated parameter of a utility function.
-In this model, the mean of power function is used.
-We also estimated varieties of this model with different risk aversion coefficients, particularly means or medians of various utility functions.
-These are power, log, exponential and quadratic.
-The risk aversion parameter is always insignificant and whether it is included or not (or which one) does not affect sign or significance level of any other parameter.
-\item[d]~A degree of agreement with the statement: 'Government should redistribute income from the better off to those who are less well off.' Included to test for significance of political opinions.
-     \end{adjustwidth}   
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-
-\vspace{0.2cm}
-
 \end{document}

### 2018-11-14 writing/draft1/manuscript11.tex
adding

--- 2018-11-14 writing/draft1/manuscript10.tex
@@
 \pagebreak
-This aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.
+The aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.
 
@@
 \cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.
+
+\textcolor{red}{Then maybe something about agriculture in Kenya. for example:}
+ \\
+ \cite{Kabubo2015} \\
+ \cite{WorldBank2015} \\
+  \cite{Lesk2016}: Global, but estimate of national cereal production losses from extreme weather disasters
+ \\
+
+\textcolor{red}{Then maybe something about proactive approach:}
 

### 2018-11-14 writing/draft1/manuscript12.tex
adding

--- 2018-11-14 writing/draft1/manuscript11.tex
@@
 \begin{abstract}
-\noindent We explore an unprecedented dataset of almost $6,000$ observations to identify main predictors of climate knowledge, climate risk perception and willingness to pay for climate change mitigation.
-Among nearly $70$ potential explanatory variables we detect the most important ones using multisplit lasso estimator.
-Importantly, we test significance of individuals' preferences about time, risk and equity.
-Our study is innovative as these behavioural characteristics were recorded by including experimental methods into a live sample survey.
-This unique way of data collection combines advantages of survey and experiments.
-The most important predictors of environmental attitudes are numeracy, cognitive ability, ideological world-view and inequity aversion.
+\noindent This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security. 
 \\
@@
 
-\noindent \textbf{JEL classification:} Q54, Q58, D80\\
 \noindent \textbf{Keywords:} Climate change, climate knowledge, climate policy, lasso, risk perception, willingness to pay\\
@@
 \pagebreak
-The aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.
 
-\textcolor{red}{something about droughts in Kenya, how they were recently and what were the effects, find refrences..see my folder literatureFS/references SSRP or the book East African Agriculture and Climate in the folder literatureFS} for example \citep{Nicholson2017}
-
+It is well known that drought and extreme heat waves affect agriculture and agricultural production mostly in negative ways (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017}) \textcolor{blue}{I can go into details about these studies if not long enough}.
+As a consequence of climate change, the situation is likely to get worse as dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014}.
 Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
@@
 Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.
-\cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.
+
+Drought has been one of the most serious problem in Kenya as well as in other Sub-Saharan countries.
+
+\textcolor{red}{something about droughts in Kenya, how they were recently and what were the effects, find refrences..see my folder literatureFS/references SSRP or the book East African Agriculture and Climate in the folder literatureFS} for example \citep{Nicholson2017}
 
@@
 
-\textcolor{red}{Then maybe something about proactive approach:}
-
-Recent literature has promoted shifting from reactive to proactive approach in disaster risk reduction.
-Forecast based financing has been promoted as it can avoid significant disaster losses \citep{Nicholson2017}.
-It has been shown that the investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}.
+Shifting from reactive to proactive approach in disaster risk reduction has been promoted in recent literature (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).
+Forecast based financing has been recommended as it can avoid significant disaster losses \citep{Nicholson2017}.
+It has been shown that investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}.
+Risk assessment is essential for disaster risk reduction \citep{IPCC2012ch1}.
+Therefore, the main purpose of this paper is to develop a model for assessing the risk of drought based on weather data.
 
@@
 \textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
+\textcolor{green}{There has been a lot of literature about impacts of climate change on agriculture..Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
+}
 
@@
 And how do we measure drought?
-To answer the first questions, the measure which has been used as a proxy for food security in the literature are as follows:
+To answer the first questions, the measure which has been used as a proxy for food security in the literature are as follows:\\
 
+\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...}
 \vspace{2cm}

### 2018-11-14 writing/draft1/manuscript13.tex
adding

--- 2018-11-14 writing/draft1/manuscript12.tex
@@
 
-\section{Introduction}\label{Introduction}
+\section{Introduction  and literature research}\label{Introduction}
 
@@
 
-It is well known that drought and extreme heat waves affect agriculture and agricultural production mostly in negative ways (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017}) \textcolor{blue}{I can go into details about these studies if not long enough}.
-As a consequence of climate change, the situation is likely to get worse as dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014}.
+Effects of drought and extreme heat waves on agriculture, which are mostly negative, have been well documented (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017, schwalm2017}).
+\textcolor{blue}{I can go into details about these studies if not long enough}.
+Agricultural drought and soil moister are important factors for plant health but also for land-atmospheric feedback and temperature variability \citep{nicolai2017}.
+As a consequence of climate change, the situation about droughts is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
 Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
@@
 
-Drought has been one of the most serious problem in Kenya as well as in other Sub-Saharan countries.
+Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Nicholson2017}.
+Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment.
+In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
+Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
 
-\textcolor{red}{something about droughts in Kenya, how they were recently and what were the effects, find refrences..see my folder literatureFS/references SSRP or the book East African Agriculture and Climate in the folder literatureFS} for example \citep{Nicholson2017}
+Large literature has focused on relationship of agricultural production and climate.
 
@@
  \\
- \cite{Kabubo2015} \\
+ \cite{Kabubo2015} \textcolor{green}{have a look at this study. Looks like very much similar to what I want to estimate. Kenya, panel, fixed and random effects...}\\
  \cite{WorldBank2015} \\
@@
  \\
+
+Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
+They analyse causes of the food crisis identified in the humanitarian appeal documents.
+Majority of these documents consider food availability and food production as a major cause of food crisis.
+\cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.
+\cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience.
+They suggest to increase the budget share focused on agricultural and livestock production to build resilience.
 
@@
 
-An important group of studies analysing the relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
+An important group of studies that analyse relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
 
@@
 \textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
-\textcolor{green}{There has been a lot of literature about impacts of climate change on agriculture..Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
+\textcolor{green}{It is worth mentioning at this place that large literature has investigated impacts of climate change on agriculture..worldwide or in Kenya??Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
 }
@@
 
-\subsection{Measures of Yield and climate}
+\subsection{Measures of yield and climate}
 
@@
 
-\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...}
+\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...maybe some references to those who also considered food prices??}
 \vspace{2cm}
@@
 
-In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions.
-Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength and severity \citep{KlischAtz2016}.
+In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}.
+Based on these data, various measures such as the vegetation condition index (VCI) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.
 
@@
 
+\textcolor{blue}{maybe also describe the two rainy seasons and why we created the monthly variables? or maybe this in intro?}
 \begin{itemize}
@@
 
-\subsection{Linear Mixed Models}
+\subsection{Linear mixed models}
 \sloppy
@@
 \pagebreak
-
-\subsection{Climate change risk perception}\label{ResPerc}
-
-In this section we discuss our estimates of the models which explain individuals' perception of climate change risk.
-We focus on two measures of climate risk perception, in particular climate change seriousness perception and climate versus policy perception.
-We present the results of lasso and jackknife OLS with the climate seriousness perception as dependent variable in Table~\ref{Climcare435}.
-Three predictors were selected, in particular gender, climate knowledge, and degree of agreement with redistribution of income by government.
-In this case, the effect of being male is negative.
-This is mostly consistent with results of previous research which typically finds women to take climate risk more seriously than men (\citealp{WHITMARSH2011, McCright2010, Kahan2007}).
-As we can see in Table~\ref{Climcare435}, degree of agreement with income redistribution affects climate change seriousness perception positively as the base category is 'Strongly disagree'.
-This is in agreement with previous literature as we consider the degree of agreement with income redistribution as an indicator of political and ideological world-view, which was found to be significantly correlated with climate concern by large number of previous studies (e.g.
-\citealp{Leiserowitz2013, Kahan2012, WHITMARSH2011}).
-
-We will comment on the significant effects of climate knowledge at the end of Section~\ref{ResPerc}.
-
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Climate change seriousness perception:} Multisplit lasso and jackknife OLS}}
-% soubor LassoClimcare43 Climcare435.R
-
-\label{Climcare435} 
-\centering
-\begin{small}
-\begin{tabular}{lclrcl} 
-\hline \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\
-
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Gender = male&$0.0002$&$^{***}$&$-0.3658$&$4.45\times 10^{-6}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Climate knowledge&$1.0000$& &$0.1380$&$1.0000$&\\
-  \\
-\vspace{-0.2cm}Climate knowledge - squared&$<2.00\times 10^{-8}$&$^{***}$&$-0.0548$&$0.0209$&$^{*}$\\
-  \\
- Redistribution of income:&\multirow{2}{*}{$1.0000$}& &\multirow{2}{*}{$0.1819$}&\multirow{2}{*}{$1.0000$}&\\%-1
-      \hspace{0.6cm}disagree\tnote{a}&& &&&\\%-1
-    \\
-  \vspace{-0.2cm}Redistribution of income: neutral\tnote{a}&$1.0000$& &$0.2789$&$0.8251$&\\%0
-    \\
-  \vspace{-0.2cm}Redistribution of income: agree\tnote{a}&$<2.00\times 10^{-8}$&$^{***}$&$0.8343$&$8.58\times 10^{-8}$&$^{***}$\\%1
-    \\
-  Redistribution of income:& \multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}&\multirow{2}{*}{$1.0828$}&\multirow{2}{*}{$<2.00\times 10^{-8}$}&\multirow{2}{*}{$^{***}$}\\
-\hspace{0.6cm}strongly agree\tnote{a}&& &&&\\%2
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-   \item[~]\textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-  \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
-
- \item[a] Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' The base category is 'Strongly disagree'.
-     \end{adjustwidth}
- \singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\hspace{1cm}
 

### 2018-11-14 writing/draft1/manuscript14.tex
adding

--- 2018-11-14 writing/draft1/manuscript13.tex
@@
 
-Effects of drought and extreme heat waves on agriculture, which are mostly negative, have been well documented (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017, schwalm2017}).
+Consequences of drought and extreme heat waves on agriculture have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}).
 \textcolor{blue}{I can go into details about these studies if not long enough}.
 Agricultural drought and soil moister are important factors for plant health but also for land-atmospheric feedback and temperature variability \citep{nicolai2017}.
-As a consequence of climate change, the situation about droughts is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
+
+Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Compendium,WorldBank2015, Nicholson2017}.
+Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment.
+In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
+Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
+As a consequence of climate change, the situation is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
 Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
@@
 
-Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Nicholson2017}.
-Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment.
-In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
-Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
+Large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa.
+A frequent finding is that increase in temperature affects crop revenue negatively while increase in precipitation has positive effects.
+\cite{RicardianBello} and  \cite{Ochieng2016} found these results by the means of Ricardian analysis.
+\cite{Ochieng2016} focused on Kenya while \cite{RicardianBello} address the situation on Niger.
+\cite{WorldBank2015} \textcolor{red}{identify and assess risks to agriculture in Kenya.}
 
-Large literature has focused on relationship of agricultural production and climate.
+Other studies investigate the relationship of climate and yields in the context of climate change.
+\cite{kabubo2007} argue that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
+Furthermore, fall and summer precipitation is positively correlated with net revenue.
+\cite{KMendelsohn2008} and \cite{SeoMendelsohn} investigate the distribution of climate change impacts on agriculture in Africa.
+They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
+\cite{Mendelsohn2008} reviewed several papers on effects of climate change on agriculture in developing countries and confirms that tropical and sub-tropical counties are much more sensitive to climate change than temperate agriculture.
+Another meta-study of climate change impacts of crop productivity in Africa in South Asia was conducted by \cite{Knox2012} who conclude that the projected mean change in yield is -$8\%$ by the 2050s in both regions.
 
@@
 
-Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
-They analyse causes of the food crisis identified in the humanitarian appeal documents.
+Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
+They analysed causes of the food crisis identified in the humanitarian appeal documents.
 Majority of these documents consider food availability and food production as a major cause of food crisis.

### 2018-11-14 writing/draft1/manuscript15.tex
adding

--- 2018-11-14 writing/draft1/manuscript13.tex
@@
 \usepackage{euscript}
+\usepackage{textcomp}
 
@@
 
-\section{Introduction  and literature research}\label{Introduction}
+\section{Introduction  and literature review}\label{Introduction}
 
-\large Findings:
-\begin{itemize}
-\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important
+Agricultural consequences of droughts and extreme heat waves have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}).
+\textcolor{blue}{I can go into details about these studies if not long enough}.
+Agricultural drought and soil moister are important factors influencing plant health but they are also essential for land-atmospheric feedback and for temperature variability \citep{nicolai2017}.
 
-\item Mar-Sept last year temperature very important for maize yields
-\item SD temperature last year positive and significant
-\item dry spell 20 MAM last year important (but not dry spell MAM10)
-\item interesting. Precipitation 2 months MAM last year very significant and positive
-\item mean temp last year negative and significant, hill shaped 
-\end{itemize}
-
-\underline{New findings:}
-
-\begin{itemize}
-\item The yields seem to be more responsive to weather on west than on east
-\end{itemize}
-\normalsize
-\FloatBarrier
-\pagebreak
-
-Effects of drought and extreme heat waves on agriculture, which are mostly negative, have been well documented (\citealp{Deschenes2007Ric,Lesk2016,Mehrabi2017, schwalm2017}).
-\textcolor{blue}{I can go into details about these studies if not long enough}.
-Agricultural drought and soil moister are important factors for plant health but also for land-atmospheric feedback and temperature variability \citep{nicolai2017}.
-As a consequence of climate change, the situation about droughts is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
-Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
+Frequent occurrence of severe droughts have been a major problem in Kenya and in other Sub-Saharan countries \citep{Compendium, WorldBank2015, Nicholson2017}.
+The agricultural sector is the main contributor to the economy of Kenya, hence the effects of drought are especially damaging in this country \citep{MendelsohnDinar2000}.
+Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment.
+In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
+Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
+As a consequence of climate change, the situation is likely to get worse since dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
+Strong downwards trend in precipitation has already been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
 During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
@@
 
-Frequent occurrence of severe drought has been a major problem in Kenya as well as in other Sub-Saharan countries \citep{Nicholson2017}.
-Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment.
-In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
-Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
+A large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa.
+A frequent finding is that increase in temperature affects crop revenue negatively while increase in precipitation has positive effects.
+\cite{RicardianBello} and  \cite{Ochieng2016} found these results by the means of Ricardian analysis.
+\cite{Ochieng2016} focused on Kenya while \cite{RicardianBello} addressed the situation in Niger.
 
-Large literature has focused on relationship of agricultural production and climate.
+Other studies have investigated the relationship of climate and yields in the context of climate change.
+\cite{MendelsohnDinar2000} have developed a simulation model to evaluate impacts of climate change on African agriculture.
+According to their study, agricultural GDP and climate change exhibit a hill-shaped relationship.
+Hence, global warming should be beneficial in high latitudes but harmful in low latitudes.
+The estimated loss of the Africa as continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}.
+\cite{kabubo2007} show that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
+Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}.
+\cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa.
+They argue that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
+\cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania.
+The authors have shown that by 2050 the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather.
+\cite{Kabubo2015} have estimated effects of climate change on food security in Kenya.
+They found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped.
+Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
+\cite{Mendelsohn2008} reviewed several studies on effects of climate change on agriculture in developing countries and based on the revision he confirmed that tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture.
+\cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia.
+The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions.
+According to another meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming.
+However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.
 
-\textcolor{red}{Then maybe something about agriculture in Kenya. for example:}
- \\
- \cite{Kabubo2015} \textcolor{green}{have a look at this study. Looks like very much similar to what I want to estimate. Kenya, panel, fixed and random effects...}\\
- \cite{WorldBank2015} \\
-  \cite{Lesk2016}: Global, but estimate of national cereal production losses from extreme weather disasters
- \\
-
-Previous studies focused on responses to drought related disasters or emergency situations include \cite{Sandstorm2017}, who question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
-They analyse causes of the food crisis identified in the humanitarian appeal documents.
+Another important group of studies has been focused on responses to drought related disasters or emergency situations.
+\cite{Sandstorm2017} questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
+They analysed causes of the food crisis identified in the humanitarian appeal documents.
 Majority of these documents consider food availability and food production as a major cause of food crisis.
@@
 Therefore, the main purpose of this paper is to develop a model for assessing the risk of drought based on weather data.
+This should in turn help to switch from reactive to proactive approach in drought disaster risk management.
 
@@
 
-\textcolor{blue}{\textbf{Add more studies. Then separate somehow into groups of different topics..}}
-\textcolor{green}{It is worth mentioning at this place that large literature has investigated impacts of climate change on agriculture..worldwide or in Kenya??Mendelsohn, Dinar, Dalfelt 2000 and 2009 and Seo, Mendelsohn, Dinar, Hassan, Kurukulasuriya 2008
-}
+To the best of our knowledge, this is the first study which has applied linear mixed effects models to estimates effects of precipitation and agriculture on crop yields in Kenya.
+Mixed effects models are suitable for analysis of panel data as they account for the panel structure of the dataset.
+That is, the relationship of yield and weather is likely to differ across counties.
+To the best of our knowledge, this is also the first study which has utilised as comprehensive weather dataset as we have used: the frequency of our weather data was daily and the resolution: \textcolor{red}{\textbf{!!!}}
 
@@
 
-{\centering
-\begin{threeparttable}
+\large Findings:
+\begin{itemize}
+\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important
 
-\singlespacing
-\caption{\textit{\textbf{Linear mixed effects models:} Maize yield and weather}}
-% Peggy45 & Peggy453   can be found in foodSystems/Rcodes/Lags/..
-  % Yield in tonnes per hectare
-\label{Peggy45} 
-\centering
-\begin{small}
-\begin{tabular}{lrccc} 
-\hline \vspace{-0.2cm} \\
+\item Mar-Sept last year temperature very important for maize yields
+\item SD temperature last year positive and significant
+\item dry spell 20 MAM last year important (but not dry spell MAM10)
+\item interesting. Precipitation 2 months MAM last year very significant and positive
+\item mean temp last year negative and significant, hill shaped 
+\end{itemize}
 
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{2}{c}{\textbf{Scaled}}& \multicolumn{2}{c}{{\textbf{Unscaled\tnote{a}}}} \\
+\underline{New findings:}
 
-    \multicolumn{1}{l}{\vspace{0.1cm}}  &\textit{Estimate}&\textit{p-value}&\multicolumn{1}{c}{\textit{Estimate}}& \multicolumn{1}{c}{\textit{p-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Intercept&$-0.059$&$0.646$&$1.379$&$1\times10^{-7}$$^{***}$\\
-  \\
-\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.054$&$0.041^{*}$&$3\times10^{-4}$&$0.177$\\
-  \\
-  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$&$0.477$&$3\times10^{-4}$&$0.197$\\
-  \\
-  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.045$&$0.244$&$-0.029$&$0.001^{**}$\\
-  \\
-    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.130$&$0.0001^{***}$&$-0.027$&$0.001^{**}$\\
-  \\
-
-      \vspace{-0.2cm}Prec. max OND, east&$0.133$&$0.013^{*}$&$0.004$&$0.085^{\bullet}$\\
-  \\
-        \vspace{-0.2cm}Prec. max OND, west&$0.144$&$0.004^{**}$&$0.011$&$9\times10^{-6}$$^{***}$\\
-  \\
-    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.049$&$0.199$&$0.075$&$0.369$\\
-  \\
-      Temp. sd. Oc.-Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$&$0.494$&$7\times10^{-12}$ $^{***}$\\
-    \vspace{-0.1cm} \\ 
-  \hline
-
-\vspace{-0.1cm}\\\multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
-\\\hline \vspace{-0.1cm} \\
-
-\vspace{0.2cm}Intercept\\
-\vspace{0.2cm}Prec. cum. MAM+OND lag 1\\
-\vspace{0.2cm}Prec. max OND\\
-\vspace{0.2cm}Temp. avg. Mar.-Sep. lag 1
-  \\
-  \hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} \hspace{0.15cm}$584$ observations; \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-    \begin{adjustwidth}{1cm}{} 
-    \item[a] The variety with unscaled variables fails to converge.
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
+\begin{itemize}
+\item The yields seem to be more responsive to weather on west than on east
+\end{itemize}
+\normalsize
+\FloatBarrier
 \pagebreak

### 2018-11-14 writing/draft1/manuscript16.tex
adding

--- 2018-11-14 writing/draft1/manuscript15.tex
@@
 \noindent This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
+Our dataset consists of yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
 \\
+...Applying the linear mixed effects models, we found that...
 \end{abstract}
 
-\noindent \textbf{Keywords:} Climate change, climate knowledge, climate policy, lasso, risk perception, willingness to pay\\
+\noindent \textbf{Keywords:}  Drought, Early warning systems, Food security, Food systems, Kenya, Mixed effects models\\
 
@@
 
-\section{Introduction  and literature review}\label{Introduction}
+\section{Introduction}\label{Introduction}
 
-Agricultural consequences of droughts and extreme heat waves have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}).
+Consequences of droughts and extreme heat waves on agriculture have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}).
 \textcolor{blue}{I can go into details about these studies if not long enough}.
@@
 
-Another important group of studies has been focused on responses to drought related disasters or emergency situations.
-\cite{Sandstorm2017} questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
-They analysed causes of the food crisis identified in the humanitarian appeal documents.
-Majority of these documents consider food availability and food production as a major cause of food crisis.
+A number of important studies have focused on responses to drought related disasters or emergency situations.
+\cite{Sandstorm2017} have questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
+They have analysed causes of the food crisis identified by the humanitarian appeal documents.
+Majority of these documents have found food availability and food production to be the major causes of food crisis.
 \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.
-\cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience.
-They suggest to increase the budget share focused on agricultural and livestock production to build resilience.
+According to \cite{Sandstorm2017}, a large share of humanitarian response budget has been focused on emergency food aid in comparison to the share focused on interventions to build resilience.
+The authors have suggested to increase the budget share focused on agricultural and livestock production to build resilience.
+Also other authors have promoted shifting from reactive to proactive approach in disaster risk management (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).
+\cite{Nicholson2017} has recommended forecast based financing as it can avoid significant disaster losses.
+\cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses.
+Furthermore, risk assessment is essential for disaster risk management \citep{IPCC2012ch1}.
+In the light of these arguments, the main purpose of the present study is to develop a model for assessing the risk of drought based on weather and climate data.
+This should help to improve focus on proactive rather than reactive approach in drought disaster risk management.
 
-Shifting from reactive to proactive approach in disaster risk reduction has been promoted in recent literature (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).
-Forecast based financing has been recommended as it can avoid significant disaster losses \citep{Nicholson2017}.
-It has been shown that investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}.
-Risk assessment is essential for disaster risk reduction \citep{IPCC2012ch1}.
-Therefore, the main purpose of this paper is to develop a model for assessing the risk of drought based on weather data.
-This should in turn help to switch from reactive to proactive approach in drought disaster risk management.
+In this study, we focus on Kenya utilizing an yearly panel dataset of $47$ counties over the period of $1981-2017$.
+As a measure of food security we use agricultural maize yields and as a measure of drought we use precipitation and temperature data.
 
-As a measure of food security we use agricultural maize yields and as measures of drought we use variables created by aggregating daily temperature and precipitation data.
+To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and agriculture on crop yields in Kenya.
+Hence, this is the first time that the effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country.
+To the best of our knowledge, this is also the first study \textcolor{blue}{which has utilised as comprehensive weather dataset as we have used: the frequency of our weather data was daily and the resolution:} \textcolor{red}{\textbf{!!!
+source of data}}
+% Kabubo-Mariara and Karanja (2015) similar study> but they use the more econometric approach (fixed effects models and random effects models as referred to by econometricians
+\section{Methodology}\label{Method}
 
-A number of previous studies have focused on analysis of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa.
+\subsection{Measures of yield and climate}
+
+Before we started our research, we had to answer the following questions: `How should we measure food security?' And `How should we measure drought?' To answer the first question, the measure which has been used as a proxy for food security in the literature are as follows:\\
+
+\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...maybe some references to those who also considered food prices??}
+\vspace{2cm}
+
+The other important question to answer before starting our analysis was: `How is drought defined and which definition and measure of drought would be the best for the purpose of this study?' Various definitions of drought and their role have been reviewed by \cite{wilhite1985} and \cite{wilhite2000}.
+For an extensive overview of drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
+\cite{keyantash2002} have quantified, evaluated and compared number of drought indices for meteorological, hydrological and agricultural forms of drought.
+For more details about types, measures and indices of drought see Appendix~$1$.
+
+In the recent period, remote sensing data have been used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}.
+Based on these data, various measures such as the vegetation condition index (VCI) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.
+
+We were considering to use the Standardised Precipitation Evapotranspiration Index  (SPEI) or VCI as measures of drought in our analysis.
+However, we did not find the county level SPEI to be as good predictor of food security as other precipitation and temperature measures and computing the SPEI for the entire \textcolor{red}{high resolution weather dataset ..}would be inadequately demanding.
+We did not use the VCI either as we were not confident enough about the reliability of the available data.\footnote{The values of the VCI which we obtained from the NDMA were not the same as those available by the \cite{BOKU} (BOKU).
+Furthermore, the VCI values did not exactly correspond to the early warning phases as defined by the \cite{NDMA} (NDMA).}
+
+Thus, after conducting a rigorous literature research on measures of drought, we decided to utilize raw daily precipitation data \textcolor{red}{from...CHIRPS??
+with resolution...
+and daily temperature data from..at resolution....}.
+The frequency of the maize yield data is yearly while the weather data are daily.
+Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
+There are many possibilities how to aggregate daily weather data.
 Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
@@
 The measures of precipitation include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August).
-Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa.
-Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010} \textcolor{red}{and others,find!!}) or annual totals or means \citep{BLIGNAUT2009}.
-Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}.
+Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa \citep{Adejuwon2004}.
+Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010}) or annual totals or means \citep{BLIGNAUT2009}.
+Some studies are based on simulated daily extremes, averages, daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}.
 Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
@@
 
-To the best of our knowledge, this is the first study which has applied linear mixed effects models to estimates effects of precipitation and agriculture on crop yields in Kenya.
-Mixed effects models are suitable for analysis of panel data as they account for the panel structure of the dataset.
-That is, the relationship of yield and weather is likely to differ across counties.
-To the best of our knowledge, this is also the first study which has utilised as comprehensive weather dataset as we have used: the frequency of our weather data was daily and the resolution: \textcolor{red}{\textbf{!!!}}
-
-\section{Methodology}\label{Method}
-
-\subsection{Measures of yield and climate}
-
-Prior we started  our research, we had to answer the following questions: How do we measure food security?
-And how do we measure drought?
-To answer the first questions, the measure which has been used as a proxy for food security in the literature are as follows:\\
-
-\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...maybe some references to those who also considered food prices??}
-\vspace{2cm}
-
-The other important questions which need to be answered before starting our research are: How is drought defined?
-What are the ways of measuring drought?
-and How shall we measure drought for the purpose of our study?
-According to the international meteorological community, drought can be defined in several ways.
-In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
- \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater.
-The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).
-
-Numerous definitions of drought and their role have been reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
-They have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
-Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
-Operational definitions are essential for an effective early warning system.
-An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.
-
-In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
-However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
-Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).
-
-For an extensive overview of various drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
-\cite{keyantash2002} quantify, evaluate and compare number of drought indices for meteorological, hydrological and agricultural forms of drought.
-Based on several criteria they conclude that rainfall deciles and SPI perform the best for meteorological drought.
-
-In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}.
-Based on these data, various measures such as the vegetation condition index (VCI) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.
-
-A description of other drought indices and measures can be found in Appendix~$1$.
-
-\textcolor{red}{We initially analysed SPEI and VCI indices and we tried to use them as input variables for our analysis. However, there were some serious problems with these measures.....}
-
-Therefore we opted for a different approach to gauging drought severity.
-In particular, we decided to utilize daily precipitation data \textcolor{red}{from...CHIRPS??
-with resolution...
-and daily temperature data from..at resolution....}.
-The frequency  of the maize yield data is yearly while we have daily weather data available.
-Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
-There are many possibilities how to aggregate daily weather data.
-Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
+Based on the literature research above and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
 
@@
 Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}.
-As a result of the high degree of county-level autonomy, the policies and regulations are likely to differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
-Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (or linear mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
+As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
+Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
+Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset.
 These types of models include both fixed affects and random effects.
@@
 
-\section{Appendix 1 Drought indices and measures}
+\section{Appendix 1 Drought: definitions, measures and indices}
+According to the international meteorological community, drought can be defined in several ways.
+In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
+ \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater.
+The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).
+
+\cite{wilhite1985} and \cite{wilhite2000} have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
+Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
+Operational definitions are essential for an effective early warning system.
+An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.
+
+In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
+However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
+Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).
 

### 2018-11-14 writing/draft1/manuscript17.tex
adding

--- 2018-11-14 writing/draft1/manuscript16.tex
@@
 
-Before we started our research, we had to answer the following questions: `How should we measure food security?' And `How should we measure drought?' To answer the first question, the measure which has been used as a proxy for food security in the literature are as follows:\\
+Before beginning our research, we had to answer the following two questions: `How should we measure food security?' and `How should we measure drought?' Based on previous literature and data availability, there are three main approaches of how to measure food security.
+These include agricultural crop yields (see literature overview in Sections~\ref{Introduction} and further in this section), food prices or health or utility indicators.
+The relevant studies focused on food prices
+were used to estimate relationship between sea surface temperature and commodity prices in tropics
+
+We focused on maize as this crop is the principal staple food in Kenya and it is grown at 90\% of all farms in Kenya \citep{FAO}.
+
+\cite{brown2014book}: global, food prices and so \\
+\cite{BROWN2015} Weather and international price shocks on food prices in the developing world.  Markets in Africa, South Asia, Latin America (regions that contain large segments of poor) \\
+\cite{OxfamIDS}: GLOBE, CGE sub-Saharan Africa, central America north Africa, other low-income countries selected by OXFAM  \\
+\cite{robinson2010}   Implications of food production and price shocks welfare in Ethiopia: a general equilibrium analysis CGE \\
+\cite{Ubilava2018}: Estimate relationship between sea surface temperature and commodity prices using. tropics?? not sure about geographic focus entirely...
 
@@
 Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
-There are many possibilities how to aggregate daily weather data.
+There are many possible ways how to aggregate daily weather data.
 Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
@@
 
-An important group of studies that analyse relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
-
-(\citealt{ThorntonEtAl2009}) \textcolor{red}{also uses diurnal daily}
+An important group of studies that have analysed relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
 
 \textcolor{red}{Erin Lentz, can we cite her paper???}
-
-Blignaut: annual averages
-Chipanschi: daily maximum and minimum temperatures
-Giannoukopoulos: number of hw days and so...
-Laux et al.: daily Tmax >30C
-Lobell et al. 2008: Monthly temp and prec.
-Lobell and Burke 2010: Temp. growing season average, precip:growing season total
-
-(\citealt{AbrahaSavage2006,Adejuwon2004,BenMohamed2002,BLIGNAUT2009,
-Chipanshi2003,Giannakopoulos2009,LauxEtAl,Leemans1993,LobellEtAl2008,
-LobellBurke2010,sagoe2006,SchulzeEtA1993,ThorntonEtAl2009,TingemEtAl2008,
-TingemEtAl2009,WalkerSchulze2008})
 

### 2018-11-14 writing/draft1/manuscript18.tex
adding

--- 2018-11-14 writing/draft1/manuscript17.tex
@@
 \begin{abstract}
-\noindent This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
+\noindent \textcolor{red}{..the full abstract to be written..} This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
 Our dataset consists of yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
@@
 
-Before beginning our research, we had to answer the following two questions: `How should we measure food security?' and `How should we measure drought?' Based on previous literature and data availability, there are three main approaches of how to measure food security.
+Before beginning our research, we had to answer two essential % fundamental??
+questions: `How should we measure food security?' and `How should we measure drought?' Taking into account previous literature and data availability, we considered three main approaches to measuring food security.
 These include agricultural crop yields (see literature overview in Sections~\ref{Introduction} and further in this section), food prices or health or utility indicators.
-The relevant studies focused on food prices
-were used to estimate relationship between sea surface temperature and commodity prices in tropics
-
+Unfortunately we had to exclude the health and utility indicators because of data availability issues.
+\textcolor{red}{rozvest..} The relevant studies focused on food prices include \cite{Ubilava2018}, who has estimated relationship between sea surface temperature and commodity prices in tropics, \cite{robinson2010} who analysed implications of food production and price shocks welfare in Ethiopia or \cite{OxfamIDS} who have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM.
+\cite{brown2014book} and \cite{BROWN2015} have developed a quantitative models to investigate relationship of  weather and international price shocks on food prices in the developing world.
+\textcolor{red}{or the disadvantage of using pricess is..}We, however, decided not to use food prices as a measure of food security because prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless taking into account the spatial autocorrelation structure.
+Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and additional data which may not be available.
+Therefore we opted for a \textcolor{red}{single} reduced form equation with crop yields as the dependent variable.
 We focused on maize as this crop is the principal staple food in Kenya and it is grown at 90\% of all farms in Kenya \citep{FAO}.
-
-\cite{brown2014book}: global, food prices and so \\
-\cite{BROWN2015} Weather and international price shocks on food prices in the developing world.  Markets in Africa, South Asia, Latin America (regions that contain large segments of poor) \\
-\cite{OxfamIDS}: GLOBE, CGE sub-Saharan Africa, central America north Africa, other low-income countries selected by OXFAM  \\
-\cite{robinson2010}   Implications of food production and price shocks welfare in Ethiopia: a general equilibrium analysis CGE \\
-\cite{Ubilava2018}: Estimate relationship between sea surface temperature and commodity prices using. tropics?? not sure about geographic focus entirely...
-
-\textcolor{red}{We were also considering to use food prices or export and import prices/retail prices, however, there were some problems...maybe some references to those who also considered food prices??}
-\vspace{2cm}
 

### 2018-11-14 writing/draft1/manuscript19.tex
adding

--- 2018-11-14 writing/draft1/manuscript18.tex
@@
 
-\subsection{Measures of yield and climate}
+\subsection{Measures of food security and drought}
 
@@
 questions: `How should we measure food security?' and `How should we measure drought?' Taking into account previous literature and data availability, we considered three main approaches to measuring food security.
-These include agricultural crop yields (see literature overview in Sections~\ref{Introduction} and further in this section), food prices or health or utility indicators.
-Unfortunately we had to exclude the health and utility indicators because of data availability issues.
-\textcolor{red}{rozvest..} The relevant studies focused on food prices include \cite{Ubilava2018}, who has estimated relationship between sea surface temperature and commodity prices in tropics, \cite{robinson2010} who analysed implications of food production and price shocks welfare in Ethiopia or \cite{OxfamIDS} who have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM.
+These include using food prices, health and utility indicators or agricultural crop yields as measures of food security.
+Unfortunately, we had to exclude the health and utility indicators because of data availability issues.
+Although some health and utility indicators, such as Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (\citealp{CSI}) or the early warning phases are available in the NDMA Early Warning Bulletins, they are only available since mid $2013$.
+Furthermore, the scales do not seem to be consistent over all counties and frequently the values are not present at all.
+The relevant studies focused on food prices include \cite{Ubilava2018}, who has estimated relationship between sea surface temperature and commodity prices in tropics, \cite{robinson2010} who analysed implications of food production and price shocks welfare in Ethiopia or \cite{OxfamIDS} who have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM.
 \cite{brown2014book} and \cite{BROWN2015} have developed a quantitative models to investigate relationship of  weather and international price shocks on food prices in the developing world.
@@
 Therefore we opted for a \textcolor{red}{single} reduced form equation with crop yields as the dependent variable.
+We have provided a broad overview of literature concerned with agricultural yields in Sections~\ref{Introduction} and further in this section.
 We focused on maize as this crop is the principal staple food in Kenya and it is grown at 90\% of all farms in Kenya \citep{FAO}.

### 2018-11-14 writing/draft1/manuscript2.tex
adding

--- 2018-11-14 writing/draft1/manuscript1.tex
@@
 
-According to scientific consensus, climate change does exist and it is human caused \citep{oreskes2004, mccarthy2001impacts, Cook2013COnsensus}.
-However, public opinions about climate change are far away from consensus \citep{leiserowitz2012, Pew2012}.
-Without taking actions to prevent or mitigate climate change and its consequences, the effects of global warming could be disastrous \citep{IPCC5WG1, IPCC5, Seneviratne2012}.
-
-Public attitudes towards natural hazards and risk perception are important drivers of policy decision making (\citealp{BritishColumbia, bookDisastersPrep}).
-Whether and how well climate change will be tackled depends on public opinion heavily.
-But what are the main factors influencing climate change perception and awareness among general population?
-Large literature examines role of personal characteristics, demographics and behavioural variables in climate change awareness and risk perception by means of survey or experimental methods.
-Recent survey-based studies include \citet{Leiserowitz2015}, who exploits the Gallup World Poll data and conclude that civic engagement, communication access and education are the most important predictors of climate change awareness while beliefs about causes and perception of local temperature changes are main predictors of climate risk perception.
-Another example of a survey-based analysis of environmental attitudes is \citet{Morrison2015}, who use ordered logit models to investigate relationship between religion and climate change attitudes and behaviour.
-They conclude that Buddhists, atheists and agnostics are the most engaged with climate change while Christian literalists are the least engaged.
-Also \citet{Carlsson2013} use survey data to estimate preferences of redistribution of burden caused by CO\textsubscript{$2$} emissions in China and in the United States.
-Understandingly, both Chines and Americans prefer rules of redistribution which are less costly for their country.
-However, these rules differ for the two countries.
-US respondents prefer the current emissions rule while in China, the historical emissions rule is preferred.
-
-As for papers based on experiments, for instance \citet{Braaten2014} analyses motivation of good feeling from giving in contrast to 'pure altruism' in climate change context.
-He concludes that 'warm glow' is important for motivating of environmentally friendly behaviour.
-Another example of a study which uses experimantal methods is \citet{GlenkColombo2013}, who conduct a choice experiment in Scotland.
-Based on the results, they claim that an outcome related risk is an important attribute in choice of land-based climate change mitigation project.
-
-The above mentioned studies are just a small portion of recent papers focused on climate concerns and climate knowledge.
-The previous studies of public attitudes and knowledge about environment are usually based either on survey or on experimental methods.
-Surveys can be conducted over large, representative samples and experiments are powerful tools to infer parameters of utility functions, measures of risk and time preferences or social value orientation (\citealp{ifcher2011, murphy2011SVO, tanaka2010riskTime}).
-Each of these methods, however, suffers from serious drawbacks.
-Surveys often lead to hypothetical bias while experimental data tend to be affected by artificial settings and small, non-representative samples which often consist of students.
-We contribute by overcoming some of these shortcomings as we use a dataset which was created by surveying a large representative sample of respondents in an experimental, interactive and dynamic way.
-Experiments are usually computer based and their participants respond to various situations on a screen.
-We replicate this set-up in a live sample survey by including the experimental methods as a part of the survey \citep{SurveyUK}.
-The experimental set-up covers attitudes towards risk, attitudes towards equity including altruism and time preferences.
-We will explore effects of these attitudes on stances towards climate change and climate knowledge.
-
-We also investigate influence of other characteristics on environmental attitudes including standard demographic data such as age, sex, race, ethnicity, religion, education, sector, occupation, date of birth, siblings, questions about assets, debts and family income.
-We further examine role of financial literacy and numeracy and we also investigate preferences regarding government spending and income redistribution, cultural and political ideology and world-view.
-As there is no general consensus on what are the main determinants of climate change perception and climate literacy, we start the explanatory analysis using a least absolute shrinkage and selection operator (lasso) to select significant predictors from almost $70$ candidates.
-
-We further contribute by showing substantially different results from those of \citet{Newell2015}, who experimentally measure individual discount rates and analyse their role in energy efficiency decisions of US households.
-They find a negative and significant effect.
-More specifically, \citet{Newell2015} conclude that willingness to pay (WTP) for annual operating energy cost savings decrease in discount rates.
-This disagrees with our results as we do not find any evidence of significant effects of individual discount rates on any of our dependent variables, including WTP.
-We argue that our estimates are more precise than those of~\citet{Newell2015} as our sample size is substantially larger.\footnote{Our estimation samples include between $5659$ and $5749$ respondents while the estimation sample of~\citet{Newell2015} has $879$ observations.} We also cover much broader spectrum of potential predictors and we use more precise estimator, in concrete multisplit lasso with resampling.
-
-Our additional contribution is a partial replication of~\citet{Kahan2015}.
-Consistently with his results, we find that climate knowledge measured by the 'ordinary climate science intelligence' (OCSI) instrument\footnote{For a more detailed description of the OCSI instrument see Section~\ref{ClimateVars}.} does not depend on measures of personal ideology and cultural world-view as opposed to other, previously used measures of climate knowledge (\citealp{Hamilton2011, Kahan2012, Kellstedt2008}).
-In accordance with \cite{Kahan2015}, our results show that climate knowledge is positively correlated with numeracy.
-We also detect association between our measure of climate knowledge and gender.
-
-In accordance with previous research (\citealp{KahanEtAl2012, Kahan2015, Kellstedt2008}) we find that stated climate change risk perception does not increase with numeracy and financial literacy (we use these variables as proxies for ability of analytical reasoning and capacity to make use of quantitative information although they can also be interpreted as tests of the respondents' attentiveness during the survey) as one may intuitively assume \citep{Weber2011}.
-As a matter of fact, we find that individuals' concerns about climate change  decline as numeracy and financial literacy increase and it is also closely related to respondents' gender and cultural and ideological world-view.
-This is consistent with previous literature (\citealp{Kahan2015, Kellstedt2008, WHITMARSH2011}).
-We particularly show that the respondents, who agree with the statement that 'Government should redistribute income from the better off to those who are less well off' (we will further refer to this statement as 'government should redistribute income', 'degree of agreement with income redistribution' or simply 'income redistribution'), which we use as a measure of cultural or ideological world-view, are more likely to take climate change more seriously and have higher WTP for its mitigation than those not agreeing with this statement.
-Consistently with recent literature (e.g.~\citealp{Kahan2015,KahanEtAl2012,Hamilton2011, HamiltonKeim2009}), we find evidence suggesting that the ideological polarization over climate change is stronger among people who are more proficient in numeracy and comprehension of quantitative information.
-
-We detect other significant predictors of WTP for climate change mitigation by means of gas and electricity tax.
-These are age, inequity aversion, perception of equality of intergenerational allocation of resources and risk assessment consistency.
-Expectedly, the respondents who consider themselves to be more affected by climate change than by climate policy have higher WTP than those who feel to be more affected by climate policy.
-Consistently with previous literature, we find negative and significant effect of age \citep{Hamilton2011, Kellstedt2008, Hayes2001gender}.
-The impacts of inequity aversion are mixed and WTP is significantly higher for those respondents who think that their income and standards of living are about the same compared to their parents (at the same stage of their lives) and their children (at the same stage of their lives as respondent currently is).
-We estimate an analogous model for WTP by means of transport fuel duty as a robustness test and the results are very comparable.
-The estimates are robust.
-
-Perhaps surprisingly, we did not find the behavioural characteristics to be significant predictors of our climate perception or climate knowledge measures.
-The only exception is inequity aversion, which has significant effects on WTP for climate change mitigation.
-
-The paper proceeds as follows.
-In Section~\ref{Method} we discuss methods, in particular multisplit lasso and jackknife ordinary least squares (OLS).
-Section~\ref{Data} describes the dataset used for our analysis and how the important variables were obtained.
-In Section~\ref{Results} we present and discuss the results.
-More specifically, Section~\ref{ResKnow} is focused on climate knowledge, in Section~\ref{ResPerc} we present estimates of climate seriousness perception models and Section~\ref{ResWTP} describes models for WTP.
-In Section~\ref{Robust} we estimate alternative specifications for each dependent variable to verify robustness of our results.
-We summarise our findings in Section~\ref{Summary} and in Section~\ref{Caveats} we discuss caveats.
-Section~\ref{Implications} includes policy implications and concluding remarks.
-
+\large Findings:
+\begin{itemize}
+\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important
+
+\item Mar-Sept last year temperature very important for maize yields
+\item SD temperature last year positive and significant
+\item dry spell 20 MAM last year important (but not dry spell MAM10)
+\item interesting. Precipitation 2 months MAM last year very significant and positive
+\item mean temp last year negative and significant, hill shaped 
+\end{itemize}
+
+\underline{New findings:}
+
+\begin{itemize}
+\item The yields seem to be more responsive to weather on west than on east
+\end{itemize}
+\normalsize
 \FloatBarrier
@@
 
-For simplicity, we set $\gamma$ in~\eqref{pvals} to be equal to $0.5$ for every application of a multisplit lasso in this study.
-Each time we perform $H=100$ splits (we believe that this number is sufficient as \citet{pValsLasso} use $50$ sample splits per simulation) and we always use one third of the sample for the variable selection using lasso and the rest for the OLS estimation and obtaining \textit{p}-values.
-
-A large fraction of our potential predictors are categorical variables because large part of the survey data was collected by multiple choice questions.
-However, the multisplit lasso selects individual predictors rather than groups of variables.
-Therefore, it can happen that a model specified by a multisplit lasso includes a dummy variable for one category of a particular categorical variable but it does not include dummy variables for its remaining categories.
-An obvious way how to overcome this issue would be to add the remaining dummy variables and use an \textit{F}-test to determine the joint significance of the group.
-If the \textit{F}-test implies that the categories are jointly significant, they should all stay in the model and they should be left out otherwise.
-However, the solution is not so straightforward with a multisplit lasso as it is not obvious on which subsample we should perform the \textit{F}-test.
-Therefore, for each model specified by a multisplit lasso, we decided to perform a following procedure which is sometimes called jackknife resampling.
-We will further refer to the procedure as a jackknife OLS.
-We again randomly split the dataset into two subsamples.
-The bigger subsample has size of two thirds of the original sample and it is used for OLS estimation and calculation of \textit{p}-values of the model with predictors selected by multisplit lasso.\footnote{Sample splitting can generally result in loss of efficiency.
-We, however estimated all models also for the whole sample as a robustness check and the results do not differ in signs or significance levels.} In addition, if the model specified by multisplit lasso includes a binary indicator which represents a category of a nominal variable, we include also all other categories of this variable among the set of predictors.
-Besides individual \textit{t}-tests we perform an \textit{F}-test of joint significance of the categories of the nominal variable.
-Similarly as in the case of multisplit lasso, we repeat the resampling and OLS estimation $100$-times.
-Each time we perform \textit{t}-tests and also a joint \textit{F}-test for each group of dummy variables representing one categorical variable.
-The \textit{p}-values of the \textit{t}-tests are then aggregated in the same way as in case of multisplit lasso (see above).
-Further, we calculate mean and median of \textit{p}-values of each joint \textit{F}-test over the $100$ subsamples and according to these statistics we determine whether the dummy indicators of the particular categorical variable should be included.
-It turns out that every time when a dummy variable representing a category of a nominal variable is chosen by a multiple lasso, both average and median \textit{p}-values of the corresponding \textit{F}-tests are below the significance level (${\alpha=0.05}$).
-Hence, we include the dummy variables for categories of each nominal variable selected by lasso (see Section~\ref{Results}).
-% parada.
-vsechny f testy jsem otestovala u vsech modelu v tabulkach.
-27.8.
-vse v poradku,takze super
-
 \FloatBarrier
@@
 
-One of our goals is to investigate effects of behavioural variables on climate knowledge and concerns about climate change.
-The behavioural variables that we consider in our study are social value orientation, time preferences, risk preferences, and attitudes towards inequality.
-
 To estimate the social value orientation, respondents played six dictator games with the same questions as in \citet{murphy2011SVO}.
@@
 
-where $P_O$ is the pay-off given to the other party, $P_S$ is the pay-off taken by the player herself and $N$ is the number of games played (in our case $6$).
-
-As one may notice in Table~\ref{PotentialPredictors} in Appendix~$3$, we also include dummy variables for four types of social value orientation (i.e.
-altruist, prosocial, individualist, competitive) among potential predictors in the lasso estimators.
-These types are defined based on ring measure~\eqref{RingMeasure}.
-Each dummy variable corresponds to one of four non-overlapping intervals of the values of ring measure~\eqref{RingMeasure}.
-None of these dummy variables was selected by lasso into any of our models, therefore we do not discuss them in more detail.
-
-As a basic measure of time preferences we use derived annual discount rates (in percentage), for investing now for one year from now and we refer to this variable as 'Discount rate year from now' in the present study.
-To obtain the data which would allow us to infer the discount rates, the respondents played games and answered questions informed by \citet{voors2012violent}, \citet{ifcher2011} and \citet{tanaka2010riskTime}.
-How the time preferences were derived is described in \cite{SurveyUK}.
-Besides using discount rates for investing now and getting returns in one year, we also implied other types of discount rates.
-These are discount rates for $(i)$~ investing now for getting returns in five years $(ii)$~investing in one year for getting returns in two years from now and $(iii)$~ investing in one year from now for getting return in six years from now.
-None of them was found to be significant, thus we do not further discuss them.
-
-We use two parameters which describe inequity aversion (\citealp{bergson1938, bergson1954, samuelson1956}), in particular the rate of inequity aversion and the subsistence or reserve income. 
-To infer these parameters, respondents were choosing from various distributions of income between three hypothetical people.
-One of them was higher on average but more unequal and the other was lower on average but more equal.
-The respondents were asked two sets of choice questions.
-In one of them, the income distribution was centred on the $70^{th}$ percentile of the UK income distribution and in the other the distribution was centred on $40^{th}$ percentile of the UK income distribution.
-Given the respondents' answers to the two choice-sets, the rate of inequity aversion and subsistence was obtained for each respondent based on equations~\eqref{InequityEq}:
-
-\begin{equation}\label{InequityEq}
-\begin{array}{lcl}
-
-\sum_{i=1}^{3}\frac{(Y^H_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^H_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma} \\
-
-\vspace{0.2cm}\\
-\sum_{i=1}^{3}\frac{(Y^L_{i,1}-\underline{Y})^{1-\gamma}}{1-\gamma} & =& \sum_{i=1}^{3}\frac{(Y^L_{i,2}-\underline{Y})^{1-\gamma}}{1-\gamma}
-\end{array}
-\end{equation}
-
-where $\gamma$ is the rate of inequity aversion, $\underline{Y}$ is the subsistence or reserve income and $Y^H_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $70^{th}$ percentile of the UK income distribution.
-Analogously, $Y^L_{i,j}$ is the income of a hypothetical person $i$ according to distribution chosen by respondent in a choice set $j$ which was centred on the $40^{th}$ percentile of the UK income distribution.
-To obtain the inequity parameters, equations~\eqref{InequityEq} were solved for $\gamma$ and $\underline{Y}$ while minimizing distance of $\underline{Y}$ to zero.
-
-In theory, the rate of inequity aversion is a continuous measure.
-However, we consider it as a categorical one as in our dataset it is equal to one of $16$ distinct values for each respondent.\footnote{We also estimated variants of models where the rate of inequity aversion is considered as scale for completeness but we do not present them to save space.
-However, the results do not differ substantially from those presented here.} These $16$ values and the corresponding frequencies can be found in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-The subsistence parameter was not selected by lasso into any of our models thus we do not discuss it in more detail.
-
-To test significance of risk aversion, we use various risk aversion coefficients which were estimated for each person from four different utility functions using Bayesian inference \citep{Balcombe2015}.
-The utility functions are power, logarithmic, exponential and quadratic and we use the estimates of their means and medians.
-None of them is significant or chosen by lasso in any of our models.
-For the economy of space we only present models with median or mean of power function.
-The estimates are very similar when we use other risk aversion coefficients.
-
-As the behavioural variables are not significant in our study, we described their measures only briefly.
-For more detailed description see \citet{SurveyUK}.
-The descriptive statistics of these variables (except of inequity aversion) can be found in Table~\ref{Descriptive} in Appendix~$3$.
-As explained above, we consider the inequity aversion rate as a categorical variable and its frequencies are in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-
 \FloatBarrier
@@
 \subsection{Climate change knowledge}\label{ResKnow}
-
-Table~\ref{Know435} summarizes estimates of the predictors of climate change knowledge which we found to be important by means of multisplit lasso estimator.
-In particular, three predictors are chosen by lasso (see first column in Table~\ref{Know435}).
-Total score on financial literacy is the number of correct answers out of three finance related mathematical problems \citep{SurveyUK}.\footnote{We also consider answers to each of the three problems separately as individual potential predictors.
-Two of them are labelled understands inflation and understands compound interest and they are identified as important predictors in other models later.} However, when we re-estimate the model using jackknife OLS with all relevant dummy variables, all categories of total score on financial literacy are insignificant.
-Furthermore, the model suffers from multicollinearity as the coefficient of correlation between cognitive reflection and total score of financial literacy is equal to $0.343$ and its \textit{p}-value is smaller than $2\times10^{-8}$.
-For illustration, estimates of jackknife OLS with all explanatory variables listed in Table~\ref{Know435} including financial literacy are shown in Table~\ref{Know435se} in Appendix~$3$.
-The last column of Table~\ref{Know435se} includes variance inflation factors (VIF) which confirm the presence of multicollinearity.
-Because of the multicollinearity and insignificance of total score on financial literacy we do not further consider this variable as a predictor of climate knowledge.
-
-The estimates of jackknife OLS without the financial literacy are summarised in the last two columns of Table~\ref{Know435}.
-The other two variables which were found to be important in explaining climate knowledge are gender and cognitive reflection test~\citep{Frederick2005}.
-We use the latter as a measure of numeracy and ability of analytical reasoning.
-
-The cognitive reflection test is fully described in~\citet{Frederick2005} and it consists of three numerical problems.
-The value of our variable is the number of correct answers out of the three questions.\footnote{The possible values are integers and half-integers between zero and three including zero and three as we also recognise if respondent solves half of a problem.
-Hence, if a respondent answers for example one and half problems correctly, her score is~$1.5$.} The frequencies of values of this variable are summarised in Table~\ref{FreqiencyCat2} in Appendix~$3$.
-To account for plausible non-linear relationship between the test score and cognitive ability we treat the variable as categorical with the base category zero.
-As it is apparent from Table~\ref{Know435}, the respondents who solved all three problems correctly have significantly higher level of climate knowledge compared to those who did not solve any of them.
-According to the jackknife OLS, climate knowledge is on average higher also for respondents who answered two problems correctly.
-However, the effect is larger for three correctly answered problems.
-Expectedly, the effect of numeracy is positive.
 
@@
 \FloatBarrier
-
-\section{Appendix 1 Climate knowledge - OCSI instrument}
-
-Below are the questions of the Ordinary Climate-Science Intelligence (OCSI) assessment developed by \cite{Kahan2015} which we use as a measure of climate knowledge.
-The eight questions are true or false statements.
-The correct answers are in bold.
-
-\begin{enumerate}
-
-\item Climate scientists believe that if the North Pole icecap melted as a result of human-caused global warming, global sea levels would rise.  \textbf{FALSE}
-
-\item Climate scientists have concluded
-that globally averaged surface air
-temperatures were higher for the 
-first decade of the twenty-first
-century (2000-2009) than for the last decade of the twentieth
-century (1990-1999).   \textbf{TRUE}
-
-\item Climate scientists believe that
-human-caused global warming
-will result in flooding of many
-coastal regions. \textbf{TRUE}
-
-\item Climate scientists believe that
-human-caused global warming
-has increased the number and
-severity of hurricanes around the
-world in recent decades. \textbf{FALSE}
-
-\item Climate scientists believe that
-nuclear power generation
-contributes to global warming.  \textbf{FALSE}
-
-\item Climate scientists believe that
-human-caused global warming
-will increase the risk of skin cancer in human beings.  \textbf{FALSE}
-
-\item Climate scientists and economists
-predict there will be positive as
-well as negative effects from
-human-caused global warming. \textbf{TRUE}
-
-\item Climate scientists believe that the
-increase of atmospheric carbon
-dioxide associated with the
-burning of fossil fuels will reduce
-photosynthesis by plants.  \textbf{FALSE}
-
-\end{enumerate}
-
-\section{Appendix 2 Climate knowledge and gender}
-
-As discussed in Section~\ref{ResKnow}, we detect a strong evidence that our measure of climate knowledge is significantly higher for men than for women.
-We find this outcome merits further investigation.
-
-We hypothesise that less educated men can have higher drop out rates from the survey than more educated men or less educated women.
-In other words, we believe that it can be more likely for men to abandon the whole survey if they find a series of questions to be too difficult to respond while women answer giving their best guess even if they are uncertain and continue with the survey.
-This could be caused by different opportunity costs, effect of pride or by males perceiving higher pressure to answer scientific questions correctly.
-If this is the case, our sample of complete cases will exhibit a selection bias as the ratio of less educated women will be bigger for the subsample of complete cases.
-
-To test for presence of the selection bias, we perform a series of following proportion tests.
-For each category of education (and also for the whole sample) we test whether the proportion of males in the subsamle of complete observations (used observations) is approximately equal to the proportion of males in the subsample of dropped observations.
-The \textit{p}-values of the corresponding Pearson's chi-square test statistics of the null hypothesis that the proportions are equal are summarised in Table~\ref{Proptests}.
-
-If the selection bias occurs, we would expect for the lower education categories the proportion of males to be significantly higher among the dropped observations than among the used observations.
-For the higher categories of education, on the other hand, we would expect the proportion of males to be significantly smaller among the dropped observation than among the used observations.
-However, this is not what we can see in Table~\ref{Proptests}.
-Although the proportion tests are significant for some GCSE, GCSE and professional, the differences in proportions are opposite to what we would expect.
-For the lower categories of education (Some GCSE and GCSE) the proportion of males is smaller among the dropped observations than among the used observations while it is the other way around for the category of professionals.
-Hence, based on the proportion tests, we do not see any evidence of the selection bias.\footnote{We performed analogous series of tests to compare proportions of males among the dropped observations with the proportions of males in the whole sample.
-Also these tests were performed separately for each category of education.
-The results are qualitatively equivalent to those of the tests in Table~\ref{Proptests}.
-The only difference is that the test became marginally insignificant for the category professional but this has no effect on the conclusion.}
-
-{\centering
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Proportion tests - no selection bias:}\\~ \\ Differences between ratio of males in the group of used observations and in the group of dropped observations.
-The tests were conducted separately for each category of education.}}
-
-% soubor heckmanNew/droppedDescriptive
-\label{Proptests} 
-\centering
-\begin{small}
-\begin{tabular}{l|cccl} 
-  \multicolumn{5}{l}{ } \\ \hline
-  \multicolumn{1}{l}{ } & \multicolumn{2}{|c}{\textbf{Proportion of males}} &  \multicolumn{2}{c}{\boldsymbol{${\tilde{\chi}^2}$}} \\
-   \multicolumn{1}{l}{\textbf{Education category}} &  \multicolumn{1}{|l}{\textbf{Used observations}} &  \multicolumn{1}{c}{\textbf{Dropped observations}} & \multicolumn{2}{c}{\textbf{\textit{p}-value}}\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Total&$0.4858$&$0.4538$&$0.0058$&$^{**}$\\
-  \\
-\vspace{-0.2cm}Craft&$0.6173$&$0.6620$&$0.4671$&\\
-  \\
-\vspace{-0.2cm}Some GCSE&$0.4957$&$0.4105$&$0.0022$&$^{**}$\\
-  \\
-\vspace{-0.2cm}GCSE&$0.4861$&$0.3792$&$0.0066$&$^{**}$\\
-  \\
-\vspace{-0.2cm}A levels&$0.4859$&$0.4671$&$0.5383$&\\
-  \\
-\vspace{-0.2cm}Diploma&$0.4609$&$0.4172$&$0.2306$&\\
-  \\
-  \vspace{-0.2cm}Bachelors&$0.4662$&$0.4294$&$0.2470$&\\
-\\
-  \vspace{-0.2cm}Professional&$0.4269$&$0.5391$&$0.0474$&$^{*}$\\
-\\
-  \vspace{-0.2cm}Masters&$0.4829$&$0.4000$&$0.1240$&\\
-  \\
-  \vspace{-0.2cm}PhD&$0.6522$&$0.5625$&$0.4889$&\\
-\\
-  \vspace{-0.2cm}No answer&$0.5435$&$0.4920$&$0.3355$&\\
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-\item[~]\textit{Notes:} \hspace{0.24cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-\newpage
-
-\vspace{1cm}
-
-To further eliminate occurrence of the selection bias, we include interactions of gender and the education categories for which the proportion tests are significant as explanatory variables besides the predictors selected by the lasso and we test their significance.
-For the sake of clearer interpretation, we also include the main effects of the education categories.
-The estimates of the model with the interactions are summarised in Table~\ref{Know4235f}.
-None of the interactions or education categories are significant and the signs and significance levels of male and gender are the same as without the interactions.\footnote{We also estimated a version of this model which includes a dummy variable for each education category.
-Adding these dummy variables does not chance signs or significance levels of gender, cognitive reflection or interactions between gender and education categories.}
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Climate change knowledge:} Jackknife OLS \\ With interactions of gender and education}}
-% soubor ResKnow4235f 
-\label{Know4235f} 
-\centering
-\begin{small}
-\begin{tabular}{lrcl} 
-\hline \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\
-
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-   \multicolumn{1}{l}{ } &  \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adjusted \textit{p}-value}} \\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Gender = male&$0.3379$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.1967$&$1.0000$&\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.1195$&$0.4384$& \\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6160$&$1.0000$ \\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$0.2664$&$0.0001$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.4405$&$1.0000$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4551$&$2.31\times 10^{-8}$&$^{***}$\\
-\\
-  \vspace{-0.2cm}Education - some GCSE&$-0.0890$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Education - GCSE&$0.0033$&$1.0000$&\\
-  \\
-  \vspace{-0.2cm}Education -  professional&$-0.0219$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - some GCSE&$-0.0494$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - GCSE&$-0.0888$&$1.0000$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education -  professional&$0.2377$&$1.0000$&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$5749$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\vspace{1.5cm}
-
-As an additional verification that our results can not be attributed to a selection we estimate a Heckman correction models for climate knowledge.
-In particular, we estimate models which are referred to as Tobit-$2$ \citep{HeckmanR}.
-The exclusion restriction is count of not responded questions out of those which were prior to climate questions in the survey questionnaire.\footnote{Count of previously not responded questions is not expected to affect climate knowledge.
-In spite of this, we estimated a variant of multisplit lasso with the count of previously not responded questions as a potential predictor to verify whether or not it should be in the climate knowledge equation according to our estimation method.
-As we expected, count of previously not responded questions was not selected.} The estimates are summarized in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}.
-The model in Table~\ref{KnowHeckman1} has only one explanatory variable in the selection equation, namely count of non-responded questions.
-The variety in Table~\ref{KnowHeckman2} includes also gender, education categories for which the proportion test in Table~\ref{Proptests} is significant and their interactions.
-The outcome equations include the predictors which were selected by the multisplit lasso (see Section~\ref{ResKnow}).
-We can see, that the male dummy variable is still positive and strongly significant in the outcome equations even if we correct for possible selection bias (see Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}).
-Parameter $\rho$ is insignificant in the models in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}.
-This means that the data are consistent with no correlation of the selection and outcome equation.
-
-We can further see in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2} that achieving score $2$ or $3$ in the cognitive reflection test has positive and significant impact on climate knowledge which is consistent with the model presented in Table~\ref{Know435} in Section~\ref{ResKnow} and with the specification in Table~\ref{Know4235f}.
-In addition, achieving score $1$ is significant in the Heckman models and score of $0.5$ is close to significant.\footnote{Besides the Heckman selection models presented in Tables~\ref{KnowHeckman1} and~\ref{KnowHeckman2}, we also estimated a version which includes all education categories and their interactions with gender as explanatory variables in the selection equation.
-However, the estimation algorithm was unable to estimate the coefficients with reasonable standard errors.}
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Climate change knowledge:} Heckman selection model}}
-% soubor  N:US/paper3/Rwork/heckmanNew/TheBest2.
-% model Claudell12
-\label{KnowHeckman1} 
-\centering
-\begin{small}
-\begin{tabular}{lrcl} 
-\hline \vspace{-0.3cm} \\
-  \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.25cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Estimate}} &  \multicolumn{2}{c}{\textbf{\textit{p}-value}} \\
- \hline 
-\hline
- \vspace{-0.25cm} \\
-   \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Probit selection equation:}}\\
-   \hline
-   \\
-\vspace{0.1cm}Not responded questions (count)&$-0.1972$ $(0.0083)$ &$<2.00\times 10^{-8}$&$^{***}$ \\
- \hline
-  \vspace{-0.25cm} \\
-  \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Outcome equation}}\\
-   \hline
-      \\
-      \vspace{-0.2cm}Gender = male&$0.3304$ $(0.0304)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.2188$ $(0.7202)$&$0.0906$&$^{\bullet}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.1472$ $(0.0387)$&$0.0001$&$^{***}$\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6092$ $(0.8821)$&$0.4898$\\
-   \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$0.3026$ $(0.0448)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.5158$ $(0.4163)$&$0.2153$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4790$ $(0.0559)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
- \hline
-  \vspace{-0.25cm} \\
-  \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Error terms}}\\
-   \hline
-      \\
-      \vspace{-0.2cm}Sigma $\sigma$&$1.2469$ $(0.0107)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Rho $\rho$&$0.0611$ $(0.0659)$&$0.3540$&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$7244$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-   \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\item[]   \begin{adjustwidth}{1cm}{}Standard errors in brackets
-     \end{adjustwidth}
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-\vspace{1cm}
-
-{\centering \singlespacing
-\begin{threeparttable}
-\caption{\textit{\textbf{Climate change knowledge:} Heckman selection model \\ With interactions of gender and education}}
-% soubor  N:US/paper3/Rwork/heckmanNew/TheBest2.
-% model Felicia12
-\label{KnowHeckman2} 
-\centering
-\begin{small}
-\begin{tabular}{lrcl} 
-\hline \vspace{-0.3cm} \\
-  \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.25cm}\textbf{Variable}} & \multicolumn{1}{l}{\textbf{Estimate}} &  \multicolumn{2}{c}{\textbf{\textit{p}-value}} \\
- \hline 
-\hline
- \vspace{-0.25cm} \\
-   \multicolumn{4}{c}{\vspace{0.1cm}\textbf{Probit selection equation:\tnote{a}}}\\
-   \hline
-   \\
- \vspace{-0.2cm}Gender = male&$-0.0163$ $(0.0954)$ &$0.8642$&\\
-  \\
- \vspace{-0.2cm}Not responded questions (count)&$-0.1983$ $(0.0084)$ &$<2.00\times 10^{-8}$&$^{***}$ \\
-\\
- \vspace{-0.2cm}Education - some GCSE&$-0.2250$ $(0.1213)$ &$0.0637$&$^{\bullet}$ \\
-\\
-  \vspace{-0.2cm}Education - GCSE&$0.0783$ $(0.1658)$ &$0.6369$&\\
-  \\
-  \vspace{-0.2cm}Education -  professional&$-0.2696$ $(0.2379)$ &$0.2570$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - some GCSE&$0.2926$ $(0.1934)$ &$0.1303$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education - GCSE&$-0.3454$ $(0.2571)$ &$0.1790$&\\
-\\
-  \vspace{-0.2cm}Male $\times$ education -  professional&$0.3659$ $(0.3468)$ &$0.2914$&\\
- \\
-   \hline  
-  \multicolumn{4}{c}{\textbf{Outcome equation}}\\
-     \hline
-      \\
-      \vspace{-0.2cm}Gender = male&$0.3307$ $(0.0304)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.2199$ $(0.7202)$&$0.0903$&$^{\bullet}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$0.1476$ $(0.0387)$&$0.0001$&$^{***}$\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=1.5$&$0.6102$ $(0.8821)$&$0.4891$\\
-   \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$0.3032$ $(0.0448)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$0.5168$ $(0.4163)$&$0.2144$&\\
-  \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.4796$ $(0.0559)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\ \hline  
-  \multicolumn{4}{c}{\textbf{Error terms}}\\
-   \hline  
-      \\
-      \vspace{-0.2cm}Sigma $\sigma$&$1.2469$ $(0.0107)$&$<2.00\times 10^{-8}$&$^{***}$\\
-  \\
-\vspace{-0.2cm}Rho $\rho$&$0.0737$ $(0.0661)$&$0.2640$&\\
-\\
-\hline
-\vspace{-0.4cm} \\ Observations:&    \multicolumn{3}{c}{$7244$} \\  \vspace{-0.4cm}
-\\
-\hline
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
- \item \textit{Notes:} \hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-\item[]   \begin{adjustwidth}{1cm}{}Standard errors in brackets \vspace{0.2cm}
- \item[a]~We also estimated a variety of this model which includes dummy variables for all education categories in the selection equation.
-They are all insignificant and the signs and significance levels of the other variables are the same.
-     \end{adjustwidth}  
- \singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
 

### 2018-11-14 writing/draft1/manuscript20.tex
adding

--- 2018-11-14 writing/draft1/manuscript18.tex
@@
 
+Droughts and climate variability do not only affect agriculture but they have widespread consequences on food security in general.
+\cite{Ubilava2018} has found a connection between commodity prices and phases of the ENSO cycle.
+\cite{robinson2010} have used a Computable General Equilibrium (CGE) model for Ethiopia and they have shown that production shocks and resulting price increases have negative effects on farmers' income in drought-prone areas while moderate price increases can be beneficial for farmers in less drought-prone areas.
+\cite{OxfamIDS} have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM and they have estimated that climate change will result in decline in food productivity while food prices will increase.
+\cite{BROWN2015} have investigated food prices in $554$ local commodity markets in Africa, South Asia and Latin America and they concluded that almost 20\% of local market food prices had been affected by weather disturbances.
+
 A large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa.
@@
 The estimated loss of the Africa as continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}.
-\cite{kabubo2007} show that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
+\cite{kabubo2007} have shown that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
 Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}.
@@
 Furthermore, risk assessment is essential for disaster risk management \citep{IPCC2012ch1}.
-In the light of these arguments, the main purpose of the present study is to develop a model for assessing the risk of drought based on weather and climate data.
+In the light of these arguments, the main goal of the present study is to develop a model for assessing the risk of drought based on weather and climate data.
 This should help to improve focus on proactive rather than reactive approach in drought disaster risk management.
@@
 
-To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and agriculture on crop yields in Kenya.
-Hence, this is the first time that the effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country.
+To the best of our knowledge, this is the first study which describes application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
+Hence, this is the first time that effects of drought and weather on food security were estimated taking into account the spatially-heterogeneous nature of Kenya and   allowing for the effects of weather to vary across the country.
 To the best of our knowledge, this is also the first study \textcolor{blue}{which has utilised as comprehensive weather dataset as we have used: the frequency of our weather data was daily and the resolution:} \textcolor{red}{\textbf{!!!
@@
 
-\subsection{Measures of yield and climate}
+\subsection{Measures of food security and drought}
 
 Before beginning our research, we had to answer two essential % fundamental??
-questions: `How should we measure food security?' and `How should we measure drought?' Taking into account previous literature and data availability, we considered three main approaches to measuring food security.
-These include agricultural crop yields (see literature overview in Sections~\ref{Introduction} and further in this section), food prices or health or utility indicators.
-Unfortunately we had to exclude the health and utility indicators because of data availability issues.
-\textcolor{red}{rozvest..} The relevant studies focused on food prices include \cite{Ubilava2018}, who has estimated relationship between sea surface temperature and commodity prices in tropics, \cite{robinson2010} who analysed implications of food production and price shocks welfare in Ethiopia or \cite{OxfamIDS} who have focused on sub-Saharan Africa, central America north Africa and other low-income countries selected by OXFAM.
-\cite{brown2014book} and \cite{BROWN2015} have developed a quantitative models to investigate relationship of  weather and international price shocks on food prices in the developing world.
-\textcolor{red}{or the disadvantage of using pricess is..}We, however, decided not to use food prices as a measure of food security because prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless taking into account the spatial autocorrelation structure.
-Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and additional data which may not be available.
-Therefore we opted for a \textcolor{red}{single} reduced form equation with crop yields as the dependent variable.
-We focused on maize as this crop is the principal staple food in Kenya and it is grown at 90\% of all farms in Kenya \citep{FAO}.
+questions: `How should we measure food security?' and `How should we measure drought?' Taking into account the aim of this study, previous literature (see Section~\ref{Introduction}) and data availability, we were considering three main approaches to measuring food security.
+In particular, we considered using health and utility indicators, food prices or agricultural crop yields as measures of food security.
+Although some health and utility indicators including the Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (see \citealp{CSI}) or the NDMA early warning phases are available in the NDMA Early Warning Bulletins, the data are only available since July $2013$.
+Furthermore, the scales of indicators do not seem to be consistent over the counties and time and the values of indicators or even entire reports are frequently missing.
+Therefore, we decided not to use the health and utility indicators.
+
+Using food prices as a measure of food security turned out be problematic as well. 
+One problem is data availability as the food prices are usually available for market towns rather than at county level.
+Furthermore, the market data which are available in the publications of \cite{KNBS} (KNBS) only cover several years.
+Another problem is that prices are usually strongly spatially autocorrelated across counties and this could cause the regression estimates to be inconsistent and inefficient unless the model explicitly accounts for the spatial autocorrelation structure.
+Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and also additional data, which may not be available.
+Therefore we opted for agricultural crop yields as a measure of food security and we decided to estimate a \textcolor{red}{single} reduced form equation with crop yields as the dependent variable.
+In particular, we focus on maize yields as this crop is the principal staple food in Kenya and it is grown at 90\% of farms in Kenya \citep{FAO}.
 

### 2018-11-14 writing/draft1/manuscript3.tex
adding

--- 2018-11-14 writing/draft1/manuscript2.tex
@@
 
-\date{\normalsize{September 2018}}
-\title{\Large \bf What Does the Weather Say about Yields in Kenya}
-\author{Monika Novackova and Pedram Rowhani and Martin Todd}
+\date{\normalsize{October 2018}}
+\title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
+\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens}
 \affil{\small{Department of Geography, University of Sussex, Falmer, UK}}
@@
 
-% table checked 3.10.2017
+\singlespacing
+\caption{\textit{\textbf{Relationship of weather and maize yield:} Mixed models}}
+% Peggy45,    can be found in foodSystems/Rcodes/Lags/..
 
-\singlespacing
-\caption{\textit{\textbf{Climate change knowledge:} Multisplit lasso and jackknife OLS}}
-% LassoKnow43se3.R      soubor Know435.R 
-%musim lasso predelat. predtim jsem pouzivala lambda min. ale alex ma pravdu. je treba u vsech modelu pouzivat stejne pravidlo, takze lambda se.zmena 30.9.2017
-
-%soubor Know435se.R
-% finlit2 prestane byt signifikantni, takze vlastne se nic skoro nemeni, asi do prilohy tabulku s Know435se.R
-
-\label{Know435} 
+\label{Peggy45} 
 \centering
 \begin{small}
-\begin{tabular}{lclrcl} 
+\begin{tabular}{lclrlcl} 
 \hline \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{} & \multicolumn{2}{c}{\large{\textbf{Multisplit lasso}}}& \multicolumn{3}{c}{\large{\textbf{Jackknife OLS}}}  \\
+  &\multicolumn{6}{c}{\textit{Estimated coefficients and p-values}}  \\
 
 \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable}} & \multicolumn{2}{c}{\textbf{Aggregated}}& \multicolumn{1}{c}{\textbf{Aggregated}} &  \multicolumn{2}{c}{\textbf{Aggregated}} \\
-    \multicolumn{1}{l}{ } & \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}}& \multicolumn{1}{c}{\textbf{coefficient}} &  \multicolumn{2}{c}{\textbf{adj. \textit{p}-value}} \\
+
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects}}  &\multicolumn{2}{c}{{\textbf{Linear - unscaled}}}& \multicolumn{2}{c}{{\textbf{Linear-scaled}}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\
+
  \hline 
@@
 \\
-\vspace{-0.2cm}Gender = male&$<2\times 10^{-8}$&$^{***}$&$0.733$&$<2\times10^{-8}$&$^{***}$\\
+\vspace{-0.2cm}Intercept&$-0.059$& $0.646$&\\
   \\
-  \vspace{-0.2cm}Cognitive reflection $=0$~\tnote{a}&$0.038$&$^{*}$ &\multicolumn{3}{c}{\textit{Not included - base cat.}}\\
+\vspace{-0.2cm}Prec. cum. MAM + OND lag 1, east&$0.054$& $0.041^{*}$&\\
   \\
-\vspace{-0.2cm}Cognitive reflection $=0.5$&$1.000$& &$2.071$&$1.000$&\\
+  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$& $0.477$&\\
   \\
-\vspace{-0.2cm}Cognitive reflection $=1$&$1.000$& &$0.283$&$0.129$& \\
+  \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, east&$-0.045$& $0.244$&\\
   \\
-\vspace{-0.2cm}Cognitive reflection $=1.5$&$1.000$& &$0.895$&$1.000$ \\
+    \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, west&$-0.045$& $0.244$&\\
   \\
-\vspace{-0.2cm}Cognitive reflection $=2$&$1.000$& &$0.628$&$1\times 10^{-5}$&$^{***}$\\
+
+      \vspace{-0.2cm}Prec. max OND, east&$0.0.133$& $0.013^{*}$&\\
   \\
-\vspace{-0.2cm}Cognitive reflection $=2.5$&$1.000$& &$1.098$&$1.000$&\\
+        \vspace{-0.2cm}Prec. max OND, west&$0.144$& $0.004^{**}$&\\
   \\
-  \vspace{-0.2cm}Cognitive reflection $=3$&$0.046$&$^{*}$&$1.033$&$<2\times10^{-8}$&$^{***}$\\
+    \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, east&$0.049$& $0.199$&\\
   \\
-\vspace{-0.2cm}Financial literacy total score $=0.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=1$&$1.000$&& \multicolumn{3}{c}{\textit{Not included}} \\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=1.5$&$1.000$&&\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=2$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-\vspace{-0.2cm}Financial literacy total score $=2.5$&$1.000$& &\multicolumn{3}{c}{\textit{Not included}}\\
-  \\
-  \vspace{-0.2cm}Financial literacy total score $=3$&$2\times10^{-5}$&$^{***}$&\multicolumn{3}{c}{\textit{Not included}}\\
+      \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$\\
+
   \\
 \hline
-\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$5749$} \\  \vspace{-0.4cm}
+\vspace{-0.4cm} \\ Observations:&  \multicolumn{5}{c}{$584$} \\  \vspace{-0.4cm}
 \\
@@
     \item \textit{Notes:} \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-    \begin{adjustwidth}{1cm}{} \item For the significant predictors, the signs of the coefficients of the multisplit lasso are the same as those of the jackknife OLS and also size of most of the coefficients is very comparable for these two models.
-    \item[a]The estimate is negative for cognitive reflection $=0$ while it is positive for cognitive reflection $=3$ in this model.
+    \begin{adjustwidth}{1cm}{} \item a comment
+    \item[a]another comment
      \end{adjustwidth}

### 2018-11-14 writing/draft1/manuscript5.tex
adding

--- 2018-11-14 writing/draft1/manuscript3.tex
@@
 
-\subsection{Climate change knowledge}\label{ResKnow}
+{\centering
+\begin{threeparttable}
+
+\singlespacing
+\caption{\textit{\textbf{Linear mixed effects models:} Maize yield and weather}}
+% Peggy45 & Peggy453   can be found in foodSystems/Rcodes/Lags/..
+
+\label{Peggy45} 
+\centering
+\begin{small}
+\begin{tabular}{lrccc} 
+\hline \vspace{-0.2cm} \\
+
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}  &\multicolumn{2}{c}{\textbf{Scaled}}& \multicolumn{2}{c}{{\textbf{Unscaled\tnote{a}}}} \\
+
+    \multicolumn{1}{l}{\vspace{0.1cm}}  &\textit{Estimate}&\textit{p-value}&\multicolumn{1}{c}{\textit{Estimate}}& \multicolumn{1}{c}{\textit{p-value}} \\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$-0.059$&$0.646$&$1.379$&$1\times10^{-7}$$^{***}$\\
+  \\
+\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, east&$0.054$&$0.041^{*}$&$3\times10^{-4}$&$0.177$\\
+  \\
+  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$&$0.477$&$3\times10^{-4}$&$0.197$\\
+  \\
+  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, east&$-0.045$&$0.244$&$-0.029$&$0.001^{**}$\\
+  \\
+    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, west&$-0.130$&$0.0001^{***}$&$-0.027$&$0.001^{**}$\\
+  \\
+
+      \vspace{-0.2cm}Prec. max OND, east&$0.133$&$0.013^{*}$&$0.004$&$0.085^{\bullet}$\\
+  \\
+        \vspace{-0.2cm}Prec. max OND, west&$0.144$&$0.004^{**}$&$0.011$&$9\times10^{-6}$$^{***}$\\
+  \\
+    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, east&$0.049$&$0.199$&$0.075$&$0.369$\\
+  \\
+      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$&$0.494$&$7\times10^{-12}$ $^{***}$\\
+  \\
+  \hline
+\vspace{-0.3cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.3cm}
+\\
+\hline
+Intercept\\
+Prec. cum. MAM+OND lag 1\\
+ Prec. max OND\\
+Temp. avg. Mar.-Sep. lag 1
+  \\
+  \hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} \hspace{0.15cm}$584$ observations; \hspace{0.15cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+    \begin{adjustwidth}{1cm}{} 
+    \item[a] The variety with unscaled variables fails to converge.
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
 
@@
 \singlespacing
-\caption{\textit{\textbf{Relationship of weather and maize yield:} Mixed models}}
-% Peggy45,    can be found in foodSystems/Rcodes/Lags/..
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather}}
+% Peggy45ln can be found in foodSystems/Rcodes/Lags/..
 
-\label{Peggy45} 
+\label{Peggy45ln} 
 \centering
 \begin{small}
-\begin{tabular}{lclrlcl} 
+\begin{tabular}{lrlclcl} 
 \hline \vspace{-0.2cm} \\
-  &\multicolumn{6}{c}{\textit{Estimated coefficients and p-values}}  \\
 
@@
 
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects}}  &\multicolumn{2}{c}{{\textbf{Linear - unscaled}}}& \multicolumn{2}{c}{{\textbf{Linear-scaled}}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\
+  \multicolumn{1}{l}{\vspace{0.1cm}}  &\multicolumn{4}{c}{\textbf{Linear}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\
 
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects}}  &\multicolumn{2}{c}{{\textbf{scaled}}}& \multicolumn{2}{c}{{\textbf{unscaled\tnote{a}}}}& \multicolumn{2}{c}{{\textbf{Log-linear}}}  \\
  \hline 
@@
 \\
-\vspace{-0.2cm}Intercept&$-0.059$& $0.646$&\\
+\vspace{-0.2cm}Intercept&$-0.059$&$0.646$&$1.379$&$1\times10^{-7}$$^{***}$&$0.158$&$0.086^{\bullet}$\\
   \\
-\vspace{-0.2cm}Prec. cum. MAM + OND lag 1, east&$0.054$& $0.041^{*}$&\\
+\vspace{-0.2cm}Prec. cum. MAM+OND lag 1, e.&$0.054$&$0.041^{*}$&$3\times10^{-4}$&$0.177$&$0.066$&$0.008^{**}$\\
   \\
-  \vspace{-0.2cm}Prec. cum. MAM lag 1, west&$0.025$& $0.477$&\\
+  \vspace{-0.2cm}Prec. cum. MAM lag 1, w.&$0.025$&$0.477$&$3\times10^{-4}$&$0.197$&$-0.006$&$0.861$\\
   \\
-  \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, east&$-0.045$& $0.244$&\\
+  \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, e.&$-0.045$&$0.244$&$-0.029$&$0.001^{**}$&$-0.036$&$0.292$\\
   \\
-    \vspace{-0.2cm}Temp. avg. Mar. - Sep. lag 1, west&$-0.045$& $0.244$&\\
+    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1, w.&$-0.130$&$0.0001^{***}$&$-0.027$&$0.001^{**}$&$-0.081$&$0.008^{**}$\\
   \\
 
-      \vspace{-0.2cm}Prec. max OND, east&$0.0.133$& $0.013^{*}$&\\
+      \vspace{-0.2cm}Prec. max OND, e.&$0.133$&$0.013^{*}$&$0.004$&$0.085^{\bullet}$&$0.081$&$0.056^{\bullet}$\\
   \\
-        \vspace{-0.2cm}Prec. max OND, west&$0.144$& $0.004^{**}$&\\
+        \vspace{-0.2cm}Prec. max OND, w.&$0.144$&$0.004^{**}$&$0.011$& $9\times10^{-6}$$^{***}$&$0.108$&$0.009^{**}$\\
   \\
-    \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, east&$0.049$& $0.199$&\\
+    \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, e.&$0.049$&$0.199$&$0.075$&$0.369$&$0.101$&$0.003^{**}$\\
   \\
-      \vspace{-0.2cm}Temp. sd. Oct. - Mar. lag 1, west&$0.225$& $4\times10^{-12}$ $^{***}$\\
+      \vspace{-0.2cm}Temp. sd. Oc.-Mar. lag 1, w.&$0.225$& $4\times10^{-12}$$^{***}$&$0.494$&$7\times10^{-12}$$^{***}$&$0.142$&$6\times10^{-7}$ $^{***}$\\
+  \\
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+  \\
+  \vspace{-0.2cm}Prec. cum. MAM+OND lag 1\\
+  \\
+  \vspace{-0.2cm}Prec. max OND\\
+  \\
+    \vspace{-0.2cm}Temp. avg. Mar.-Sep. lag 1\\
 
@@
     \begin{adjustwidth}{1cm}{} \item a comment
-    \item[a]another comment
+    \item[a] The variety with unscaled variables fails to converge.
      \end{adjustwidth}

### 2018-11-14 writing/draft1/manuscript7.tex
adding

--- 2018-11-14 writing/draft1/manuscript10.tex
@@
 \usepackage{changepage}
-\usepackage{mathrsfs}
 
@@
 \usepackage{epstopdf}
-\usepackage{euscript}
 
@@
 \FloatBarrier
-\pagebreak
-This aim of this paper is to contribute to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security in Kenya.
 
-\textcolor{red}{something about droughts in Kenya, how they were recently and what were the effects, find refrences..see my folder literatureFS/references SSRP or the book East African Agriculture and Climate in the folder literatureFS} for example \citep{Nicholson2017}
+The purpose of this paper is to contribute to better understanding of effects of drought on food security in Kenya and to improve early warning systems.
+\textcolor{red}{shifting from reactive to proactive approach} As a measure of food security we use agricultural maize yields and as measures of drought we use variables created by aggregating daily temperature and precipitation data.
 
@@
 
-Recent literature has promoted shifting from reactive to proactive approach in disaster risk reduction.
-Forecast based financing has been promoted as it can avoid significant disaster losses \citep{Nicholson2017}.
-It has been shown that the investment into disaster risk reduction had usually been outweighed by avoided losses \citep{Mechler2005}.
-
-As a measure of food security we use agricultural maize yields and as measures of drought we use variables created by aggregating daily temperature and precipitation data.
-
-A number of previous studies have focused on analysis of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa.
-Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
-\cite{Adejuwon2004} analysed relationship of crop yields and three measures of precipitation.
-The measures of precipitation include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August).
-Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa.
-Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010} \textcolor{red}{and others,find!!}) or annual totals or means \citep{BLIGNAUT2009}.
-Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}.
-Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
-The definition of wet and dry days and rainy season vary across the literature.
-For example, \cite{BenMohamed2002} has assumed that rainy season begins when the amount of rainfall in three consecutive days reach at least 25mm and no dry spell of more than seven days occurs in the following thirty days.
-According to this study, the end of the rainy season is defined as that rainy day after which rain recorded during 20 days is less than 5mm.
-\cite{BenMohamed2002} has also found sea surface temperature anomalies at various locations and amount of rainfall in July, August and September to be significant for millet crops in Niger.
-The author has also considered the maximum air temperature in the hottest month (April) and the minimum air temperature in the coldest month (January) as possible predictors of crops in Niger, but he did not find them significant.
-
-An important group of studies analysing the relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
-
-(\citealt{ThorntonEtAl2009}) \textcolor{red}{also uses diurnal daily}
-
+A number of previous studies were based on estimation of relationship of yield (or other measures of food security) and precipitation, temperature or other variables derived from higher frequency weather data.
+Many of them were spatially focused on Kenya or other countries of Sub-Saharan Africa.
 \textcolor{red}{Erin Lentz, can we cite her paper???}
@@
 Chipanshi2003,Giannakopoulos2009,LauxEtAl,Leemans1993,LobellEtAl2008,
-LobellBurke2010,sagoe2006,SchulzeEtA1993,ThorntonEtAl2009,TingemEtAl2008,
-TingemEtAl2009,WalkerSchulze2008})
+LobellBurke2010})
 
@@
 \section{Methodology}\label{Method}
-
-\subsection{Measures of Yield and climate}
 
@@
 
-Numerous definitions of drought and their role have been reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
-They have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
+Numerous definitions of drought and their role are reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
+They distinguish two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
 Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
@@
 
-In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
+A numerical measure is needed to compare severity of drought across different time periods or geographical locations.
 However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
-Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).
+Instead of that a number of measures of drought has been developed (\citealp{Heim2002, wilhite1985, wilhite2000}).
 
-For an extensive overview of various drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
+For an extensive overview of various drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{svoboda2016} or \cite{zargar2011}.
 \cite{keyantash2002} quantify, evaluate and compare number of drought indices for meteorological, hydrological and agricultural forms of drought.
@@
 In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions.
-Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength and severity \citep{KlischAtz2016}.
+Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength \citep{KlischAtz2016}.
 
-A description of other drought indices and measures can be found in Appendix~$1$.
+For description of other drought indices and measures see Appendix~$1$.
 
@@
 
-Therefore we opted for a different approach to gauging drought severity.
+Therefore we opted for a different approach of measuring drought severity.
 In particular, we decided to utilize daily precipitation data \textcolor{red}{from...CHIRPS??
@@
 and daily temperature data from..at resolution....}.
-The frequency  of the maize yield data is yearly while we have daily weather data available.
-Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
+The frequency  of the maize yield data is yearly while the frequency of the weather data is daily.
+Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the maize data.
 There are many possibilities how to aggregate daily weather data.
-Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
+Based on the literature research in Section~\ref{Introduction}, we decided to create a set of the following weather aggregates:
 
-\begin{itemize}
+We then searched for the best subset of predictors of yield among the variables listed...
 
-\item \textbf{Precipitation:}
+\sloppy
+The estimation function can be written as \citep{Friedman2010}:
 
-\begin{itemize}
-\item Seasonal cumulative rainfall
-\item Seasonal standard deviation \textcolor{red}{actually it is monthly at the moment. I am not sure how I would aggregate this over season??}
-\item Seasonal coefficient of variation \textcolor{red}{actually it is monthly at the moment. I am not sure how I would aggregate this over season??}
-\item Maximum length of dry spell in number of days
-\item 
-\end{itemize}
-\end{itemize}
-\subsection{Data}
+\begin{equation}\label{Lasso}
+\begin{array}{lcll}
 
-\subsection{Linear Mixed Models}
-\sloppy
-Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}.
-As a result of the high degree of county-level autonomy, the policies and regulations are likely to differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
-Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (or linear mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
-These types of models include both fixed affects and random effects.
-Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}.
-On the other hand, random effect are unobserved random variables.
-There are at least three benefits of treating a set of parameters as a random sample from some distribution.
-\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9,KERYch12}).
-
-Formally, a linear mixed model can be described by the distribution of two vectors of random variables: the response $\mathscr{Y}$ and the vector of random effects $\mathscr{B}$.
-The distribution of $\mathscr{B}$ is multivariate normal and the conditional distribution of $\mathscr{Y}$ given $\mathscr{B}=\mathbf{b}$ is multivariate normal of a form (\citealp{bates2010lme4, KERYch9}):
-
-\begin{equation}\label{MixedGeneral}
-\begin{array}{lcl}
-
-(\mathscr{Y}|\mathscr{B}=\mathbf{b})& \sim & \mathit{N}(\mathbf{X}\mathbf{\beta}+\mathbf{Z}\mathbf{b},\sigma^2\mathbf{I}),
-
+ \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \boldsymbol{R}_{\lambda}(\beta_0, \beta)&=&
+   \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \bigg[ \frac{1}{2N}\sum_{i=1}^{N}(y_i - \beta_0-x_i^{\intercal}\beta)^2 + \lambda\sum_{j=1}^{p}(|\beta_j|) \bigg], \\
 \end{array}
@@
 
-where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variable $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ a scale factor.
-The distribution of $\mathscr{B}$ can be written as:
+In line with common practice, we compute estimator~\eqref{Lasso} for a series of~$\lambda$ and then we choose a preferred value of~$\lambda$ using cross-validation \citep{statisHighDimData}.
 
-\begin{equation}\label{ranefDist}
-\mathscr{B} \sim \mathit{N}(0,\mathbf{\Sigma}),
+\begin{equation}\label{pvals}
+Q_j(\gamma)=min \left \{1,q_{\gamma} (   \{ P^{(h)}_j/\gamma; h=1,...,H\}) \right \},
 \end{equation}
 
-where $\mathbf{\Sigma}$ is a $q \times q$ positive semi-definite variance-covariance matrix
+ where and $q_{\gamma}(\tiny{\cdot})$ is the (empirical) $\gamma$-quantile function. We will further refer to this procedure as a multisplit lasso.
+
+\FloatBarrier
+\section{Data}\label{Data}
+All data used in this study except of predicted income and population density, which we use in robustness tests,  were collected in the survey conducted by \cite{SurveyUK}.
+
+In Section~\ref{Robust} we use an alternative measure of income as a robustness test.
+In particular, this estimated income is obtained from a regression model based on data from Annual Survey of Hours and Earnings (ASHE).
+More specifically, the predicted income is based on age, gender, occupation, sector and education.
+
+{\centering
+\begin{threeparttable}
+\caption{\textit{\textbf{Sex and age distribution} \\ of the sample and the population}}
+\label{SexAge} 
+\begin{small}
+\begin{tabular}{|l|rr|rr|} 
+\hline	
+  \multicolumn{1}{|l|}{ } & \multicolumn{2}{c}{\bf{Sample}} & \multicolumn{2}{|c|}{\bf{UK population\tnote{a}}} \\ %sample pubpolM2 used here
+    \multicolumn{1}{|l|}{\vspace{0.1cm}\textbf{Age range}}  &  \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c}{\bf{Female}} & \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c|}{\bf{Female}}\\ 
+\hline 
+$18-24$&$9.8\%$&$9.4\%$&$6.2\%$&$6.0\%$\\
+$25-34$&$10.0\%$&$10.3\%$&$9.0\%$&$9.1\%$\\
+$35-44$&$7.8\%$&$8.3\%$&$8.6\%$&$8.8\%$\\
+$45-54$&$8.1\%$&$9.4\%$&$9.3\%$&$9.6\%$\\
+$55-64$&$7.4\%$&$8.4\%$&$7.5\%$&$7.8\%$\\
+$65-74$&$4.1\%$&$4.8\%$&$6.2\%$&$6.7\%$\\
+$75-80$&$0.1\%$&$0.1\%$&$2.4\%$&$2.8\%$\\
+\hline
+\hline
+\end{tabular} 
+  \begin{tablenotes}
+  \begin{footnotesize}
+  \singlespacing
+     \item[a]Population data are from the Office of National Statistics, Population Estimates of UK, England and
+Wales, Scotland and Northern Ireland Mid $2014$, Table MYE2.
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+\end{small}
+  \end{threeparttable} \par}
+
+\hspace{1.5cm}
+
+\subsection{Climate variables}\label{ClimateVars}
+
+Descriptive statistics of our climate variables are summarised in Table~\ref{DepDes}.
+
+{\centering
+\begin{threeparttable}
+\caption{\textit{\textbf{Dependent variables:} Descriptive statistics}}
+\label{DepDes} 
+\centering
+\begin{small}
+\begin{tabular}{lrrrr} 
+\hline	
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  &  \multicolumn{1}{c}{\bf{Mean}} & \multicolumn{1}{c}{\bf{St. dev.}} & \textbf{Min} & \textbf{Max}\\ 
+\hline \vspace{-0.3cm} \\ 
+  \vspace{0.15cm}Climate change knowledge&$3.851$&$1.266$&$1$&$8$\\
+    \vspace{0.15cm}Climate change seriousness perception&$6.622$&$2.249$&$0$&$10$\\
+        \vspace{0.15cm}Climate versus policy effects perception&$5.370$&$2.315$&$0$&$10$\\
+\vspace{0.15cm}WTP - gas and electricity tax (\textsterling~per year)&$123.900$&$105.459$&$0$&$500$\\
+ \vspace{0.15cm}WTP - duty on transport fuel (pence per year)&$20.530$&$22.518$&$0$&$100$\\
+\hline
+\hline
+\end{tabular} 
+\end{small}
+  \end{threeparttable} 
+\par}
+
+\hspace{1.5cm}
+
+To investigate opinions about seriousness of climate change, the respondents were asked the following question: 'How serious a problem do you think climate change is at this moment?' Using an interactive slider, the respondents answered an integer value between $0$~and~$10$ where min~$=0$ and max~$=10$ (as it was noted just below the slider).
+In a similar way, the respondents were asked if they feel to be more affected by climate change or by climate policy.
+The wording of the question was: 'Which affects you and your way of life more, climate change or policies to reduce greenhouse gas emissions?' Again, the respondents provided answers on an integer scale from $0$ (climate policy) to $10$ (climate change) using a slider.
+Relative frequencies of climate seriousness perception and climate versus policy perception are summarised in Table~\ref{DepFreq}.
+
+\vspace{1cm}
+
+{\centering
+\begin{threeparttable}
+\caption{\textit{\textbf{Dependent variables:} Relative frequencies (\%)}}
+\label{DepFreq} 
+\centering
+\begin{small}
+\begin{tabular}{lrrrrrrrrrrrr} 
+\hline 
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  & \multicolumn{1}{c}{\boldsymbol{$0$}}& \multicolumn{1}{c}{\boldsymbol{$1$}}& \multicolumn{1}{c}{\boldsymbol{$2$}}& \multicolumn{1}{c}{\boldsymbol{$3$}}& \multicolumn{1}{c}{\boldsymbol{$4$}}&\multicolumn{1}{c}{\boldsymbol{$5$}}&\multicolumn{1}{c}{\boldsymbol{$6$}}&\multicolumn{1}{c}{\boldsymbol{$7$}}&\multicolumn{1}{c}{\boldsymbol{$8$}}&\multicolumn{1}{c}{\boldsymbol{$9$}}&\multicolumn{1}{c}{\boldsymbol{$10$}}&\\
+\hline \vspace{-0.3cm} \\ 
+  \vspace{0.15cm}Climate knowledge&$0.0$&$1.7$&$11.4$&$30.4$&$25.4$&$20.9$&$8.6$&$1.6$&$0.1$&N/A&N/A\\
+\vspace{-0.3cm}Climate seriousness&\multirow{2}{*}{$3.3$}&\multirow{2}{*}{$2.8$}&\multirow{2}{*}{$5.4$}&\multirow{2}{*}{$8.1$}&\multirow{2}{*}{$8.9$}&\multirow{2}{*}{$27.2$}&\multirow{2}{*}{$14.0$}&\multirow{2}{*}{$12.8$}&\multirow{2}{*}{$8.3$}&\multirow{2}{*}{$4.1$}&\multirow{2}{*}{$5.0$}\\
+        \vspace{-0.15cm}\hspace{0.5cm}perception&\\
+\vspace{-0.3cm}Climate vs.
+policy&\multirow{2}{*}{$2.1$}&\multirow{2}{*}{$1.5$}&\multirow{2}{*}{$2.5$}&\multirow{2}{*}{$3.8$}&\multirow{2}{*}{$4.6$}&\multirow{2}{*}{$9.8$}&\multirow{2}{*}{$18.5$}&\multirow{2}{*}{$21.7$}&\multirow{2}{*}{$16.7$}&\multirow{2}{*}{$8.6$}&\multirow{2}{*}{$10.4$}\\
+\vspace{0.15cm}\hspace{0.5cm}perception\tnote{a}&\\
+\hline
+\hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+  \singlespacing
+     \item[~]\textit{Notes:} Total number of observations: $5749$
+       \begin{adjustwidth}{1cm}{} 
+     \vspace{-0.3cm} \item[a]~Higher number means greater concern about climate change, lesser concern about climate policy.
+         \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+
+\vspace{1cm}
+
+\subsection{Behavioural variables}\label{BehaviouralVars}
+
+To estimate the social value orientation, respondents played six dictator games with the same questions as in \citet{murphy2011SVO}.
+The ring measure of social value orientation which we use in our models is defined as
+
+\begin{equation}\label{RingMeasure}
+\begin{array}{lcll}
+
+R&=&\arctan \frac{\sum_{i=1}^{N} P_O -50N}{\sum_{i=1}^{N} P_S -50N}, \\
+\end{array}
+\end{equation}
 
@@
 	\section{Results and discussion}\label{Results}
+In this section we describe our results and discuss their interpretation.
+
+In the tables which summarise the estimates of lasso below, \textit{p}-values of some of the explanatory variables are equal to one.
+These variables were not selected by the lasso in most of the sample splits.
+They are, however, included in the tables because they represent either a category of a nominal variable whose other category was selected by the lasso or a linear term of a variable whose quadratic term was selected by the lasso.
 

### 2018-11-14 writing/draft1/manuscript8.tex
adding

--- 2018-11-14 writing/draft1/manuscript7.tex
@@
 
-A number of previous studies were based on estimation of relationship of yield (or other measures of food security) and precipitation, temperature or other variables derived from higher frequency weather data.
-Many of them were spatially focused on Kenya or other countries of Sub-Saharan Africa.
+A number of previous studies were focused on estimation of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa.
+Common measures of precipitation and temperature are monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
+\cite{Adejuwon2004} focused on the Sub-Saharan West Africa and he used total rainfall for: \textit{(i)} The first month of the period from sowing to harvesting (June) \textit{(ii)}: The first two months of the period from sowing to harvesting (June and July) and \textit{(iii)}: The first three months of the period from sowing to harvesting (June, July and August).
+Based on his results, June and July are the most important for crop yield.
+Other studies are based on seasonal totals or means..(..)  Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{Chipanshi2003,AbrahaSavage2006,SchulzeEtA1993}).
+Another measure which has been used to model relationship with maize yields in the past literature is the number of wet and dry day counts per a defined period, usually a month or season (\citealt{AbrahaSavage2006}) The definition of wet and dry days vary across the studies.
+
+ describes simulate potential maize yields using generated weather data and generated weather a
 \textcolor{red}{Erin Lentz, can we cite her paper???}
@@
 Chipanshi2003,Giannakopoulos2009,LauxEtAl,Leemans1993,LobellEtAl2008,
-LobellBurke2010})
+LobellBurke2010,sagoe2006,SchulzeEtA1993,ThorntonEtAl2009,TingemEtAl2008,
+TingemEtAl2009,WalkerSchulze2008})
 
@@
 
-Numerous definitions of drought and their role are reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
-They distinguish two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
+Numerous definitions of drought and their role have been reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
+They have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
 Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
@@
 
-A numerical measure is needed to compare severity of drought across different time periods or geographical locations.
+In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
 However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
@@
 In the recent period, remote sensing data have been collected and used increasingly to monitor levels of greenness and closely related vegetation conditions.
-Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength \citep{KlischAtz2016}.
+Based on these data, the vegetation condition index (VCI) has been developed for quantifying drought strength and severity \citep{KlischAtz2016}.
 
-For description of other drought indices and measures see Appendix~$1$.
+A description of other drought indices and measures can be found in Appendix~$1$.
 
@@
 
-Therefore we opted for a different approach of measuring drought severity.
+Therefore we opted for a different approach to gauging drought severity.
 In particular, we decided to utilize daily precipitation data \textcolor{red}{from...CHIRPS??
@@
 and daily temperature data from..at resolution....}.
-The frequency  of the maize yield data is yearly while the frequency of the weather data is daily.
-Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the maize data.
+The frequency  of the maize yield data is yearly while we have daily weather data available.
+Hence, the weather data need to be aggregated in order to obtain a dataset conformable with the yield data.
 There are many possibilities how to aggregate daily weather data.
-Based on the literature research in Section~\ref{Introduction}, we decided to create a set of the following weather aggregates:
+Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed following aggregates of the weather data:
+
+ we decided to create a set of the following weather aggregates:
 

### 2018-11-14 writing/draft1/manuscript9.tex
adding

--- 2018-11-14 writing/draft1/manuscript8.tex
@@
 A number of previous studies were focused on estimation of relationship of agricultural yield and measures of precipitation and temperature in Kenya or in other countries in Sub-Saharan Africa.
-Common measures of precipitation and temperature are monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
-\cite{Adejuwon2004} focused on the Sub-Saharan West Africa and he used total rainfall for: \textit{(i)} The first month of the period from sowing to harvesting (June) \textit{(ii)}: The first two months of the period from sowing to harvesting (June and July) and \textit{(iii)}: The first three months of the period from sowing to harvesting (June, July and August).
-Based on his results, June and July are the most important for crop yield.
-Other studies are based on seasonal totals or means..(..)  Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{Chipanshi2003,AbrahaSavage2006,SchulzeEtA1993}).
-Another measure which has been used to model relationship with maize yields in the past literature is the number of wet and dry day counts per a defined period, usually a month or season (\citealt{AbrahaSavage2006}) The definition of wet and dry days vary across the studies.
+Commonly used measures of precipitation and temperature include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
+\cite{Adejuwon2004} analysed relationship of crop yields and three measures of precipitation.
+The three measures include total rainfall during: \textit{(i)} The first month of the period from sowing to harvesting (June) \textit{(ii)} The first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} The first three months of the period from sowing to harvesting (June, July and August).
+Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa.
+Other studies utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010} \textcolor{red}{and others,find!!})  Some authors have utilized simulated daily extremes, averages or daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}).
+Another measures which have been proposed for modelling the variability of maize yield are numbers of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009})   or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
+The definition of wet and dry days and rainy season vary across the literature.
+For example, \citep{BenMohamed2002} has assumed that rainy season begins when the amount of rainfall in three consecutive days is at least 25mm and no dry spell of more than seven days duration occurs in the following thirty days and the end of the rainy season is that rainy day after which rain recorded during 20 days is less than 5mm.
 
@@
 \section{Methodology}\label{Method}
+
+\subsection{Measures of Yield and climate}
 
@@
 There are many possibilities how to aggregate daily weather data.
-Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed following aggregates of the weather data:
+Based on the literature research in Section~\ref{Introduction} and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
 
@@
 
+\subsection{Linear Mixed Models}
+
 \sloppy
-The estimation function can be written as \citep{Friedman2010}:
-
-\begin{equation}\label{Lasso}
-\begin{array}{lcll}
-
- \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \boldsymbol{R}_{\lambda}(\beta_0, \beta)&=&
-   \underset{(\beta_0, \beta) \in \mathbb{R}^{(p+1)}}{min} \bigg[ \frac{1}{2N}\sum_{i=1}^{N}(y_i - \beta_0-x_i^{\intercal}\beta)^2 + \lambda\sum_{j=1}^{p}(|\beta_j|) \bigg], \\
-\end{array}
-\end{equation}
-
-In line with common practice, we compute estimator~\eqref{Lasso} for a series of~$\lambda$ and then we choose a preferred value of~$\lambda$ using cross-validation \citep{statisHighDimData}.
-
-\begin{equation}\label{pvals}
-Q_j(\gamma)=min \left \{1,q_{\gamma} (   \{ P^{(h)}_j/\gamma; h=1,...,H\}) \right \},
-\end{equation}
-
- where and $q_{\gamma}(\tiny{\cdot})$ is the (empirical) $\gamma$-quantile function. We will further refer to this procedure as a multisplit lasso.
 
@@
 
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Sex and age distribution} \\ of the sample and the population}}
-\label{SexAge} 
-\begin{small}
-\begin{tabular}{|l|rr|rr|} 
-\hline	
-  \multicolumn{1}{|l|}{ } & \multicolumn{2}{c}{\bf{Sample}} & \multicolumn{2}{|c|}{\bf{UK population\tnote{a}}} \\ %sample pubpolM2 used here
-    \multicolumn{1}{|l|}{\vspace{0.1cm}\textbf{Age range}}  &  \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c}{\bf{Female}} & \multicolumn{1}{|c}{\bf{Male}} & \multicolumn{1}{c|}{\bf{Female}}\\ 
-\hline 
-$18-24$&$9.8\%$&$9.4\%$&$6.2\%$&$6.0\%$\\
-$25-34$&$10.0\%$&$10.3\%$&$9.0\%$&$9.1\%$\\
-$35-44$&$7.8\%$&$8.3\%$&$8.6\%$&$8.8\%$\\
-$45-54$&$8.1\%$&$9.4\%$&$9.3\%$&$9.6\%$\\
-$55-64$&$7.4\%$&$8.4\%$&$7.5\%$&$7.8\%$\\
-$65-74$&$4.1\%$&$4.8\%$&$6.2\%$&$6.7\%$\\
-$75-80$&$0.1\%$&$0.1\%$&$2.4\%$&$2.8\%$\\
-\hline
-\hline
-\end{tabular} 
-  \begin{tablenotes}
-  \begin{footnotesize}
-  \singlespacing
-     \item[a]Population data are from the Office of National Statistics, Population Estimates of UK, England and
-Wales, Scotland and Northern Ireland Mid $2014$, Table MYE2.
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-\end{small}
-  \end{threeparttable} \par}
-
-\hspace{1.5cm}
-
 \subsection{Climate variables}\label{ClimateVars}
@@
 Descriptive statistics of our climate variables are summarised in Table~\ref{DepDes}.
-
-{\centering
-\begin{threeparttable}
-\caption{\textit{\textbf{Dependent variables:} Descriptive statistics}}
-\label{DepDes} 
-\centering
-\begin{small}
-\begin{tabular}{lrrrr} 
-\hline	
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Variable:}}  &  \multicolumn{1}{c}{\bf{Mean}} & \multicolumn{1}{c}{\bf{St. dev.}} & \textbf{Min} & \textbf{Max}\\ 
-\hline \vspace{-0.3cm} \\ 
-  \vspace{0.15cm}Climate change knowledge&$3.851$&$1.266$&$1$&$8$\\
-    \vspace{0.15cm}Climate change seriousness perception&$6.622$&$2.249$&$0$&$10$\\
-        \vspace{0.15cm}Climate versus policy effects perception&$5.370$&$2.315$&$0$&$10$\\
-\vspace{0.15cm}WTP - gas and electricity tax (\textsterling~per year)&$123.900$&$105.459$&$0$&$500$\\
- \vspace{0.15cm}WTP - duty on transport fuel (pence per year)&$20.530$&$22.518$&$0$&$100$\\
-\hline
-\hline
-\end{tabular} 
-\end{small}
-  \end{threeparttable} 
-\par}
-
-\hspace{1.5cm}
 
@@
 
-To estimate the social value orientation, respondents played six dictator games with the same questions as in \citet{murphy2011SVO}.
-The ring measure of social value orientation which we use in our models is defined as
-
-\begin{equation}\label{RingMeasure}
-\begin{array}{lcll}
-
-R&=&\arctan \frac{\sum_{i=1}^{N} P_O -50N}{\sum_{i=1}^{N} P_S -50N}, \\
-\end{array}
-\end{equation}
-
 \FloatBarrier
 	\section{Results and discussion}\label{Results}
-In this section we describe our results and discuss their interpretation.
-
-In the tables which summarise the estimates of lasso below, \textit{p}-values of some of the explanatory variables are equal to one.
-These variables were not selected by the lasso in most of the sample splits.
-They are, however, included in the tables because they represent either a category of a nominal variable whose other category was selected by the lasso or a linear term of a variable whose quadratic term was selected by the lasso.
 

### 2018-11-14 writing/draft1/manuscriptAll.tex
adding

--- 2018-11-14 writing/draft1/manuscript1.tex
@@
 
-\date{\normalsize{September 2018}}
-\title{\Large \bf What Does the Weather Say about Yields in Kenya}
-\author{Monika Novackova and Pedram Rowhani and Martin Todd}
-\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}
+\date{\normalsize{December 2017}}
+\title{\Large \bf Climate Change Awareness and Willingness to Pay for its Mitigation: \\ Evidence from the UK}
+\author{Monika Novackova and Richard Tol}
+\affil{\small{Department of Economics, University of Sussex, Falmer, UK}}
 
@@
 \FloatBarrier
-\section{Methodology}\label{Method}
+\section{Econometric Methodology}\label{Method}
 
@@
 \FloatBarrier
-\section{Data}\label{Data}
+\section{Data and Survey Methodology}\label{Data}
 All data used in this study except of predicted income and population density, which we use in robustness tests,  were collected in the survey conducted by \cite{SurveyUK}.
@@
 \FloatBarrier
+\section{Robustness}\label{Robust}
+
+One of the important objectives of this study is to examine effects of behavioural variables on climate knowledge and climate change perception.
+Except of inequity aversion, none of the behavioural variables was chosen by lasso into any of our models.
+In spite of this, for each dependent variable we estimate a jackknife OLS with all the behavioural measures included as explanatory variables (besides the predictors selected by lasso) as robustness tests and to investigate possible changes in signs and significance levels of the previously
+selected predictors.
+We also add some other potentially confounding variables including population density, climate variables other than the dependent variable, degree of agreement with income redistribution, net assets and predicted income as an alternative to the income recorded in our survey~\citep{SurveyUK} since the income obtained from the survey is categorical rather than continuous.
+\footnote{We include these variables in the robustness tests although they were not originally chosen by lasso.}
+
+The population density serves as a proxy for rural-urban classification.
+We include it in the robustness tests because we believe that whether one lives in rural or urban area can have considerable impact on attitudes towards climate change.
+
+In each model we include attitude towards income redistribution as an indicator of political and ideological world-views as the ideological opinions were found to be especially important for explaining of attitudes towards climate change and measures of climate knowledge by large number of previous studies (e.g.
+\citet{Hamilton2011}, \citet{Kahan2012}, \citet{Kellstedt2008}).
+
+Controlling for income and net assets is especially important for WTP as both income and assets are very likely to be correlated with WTP.
+However, also other variables, such as discount rates, can be correlated with income and net assets.
+Therefore, we examine whether controlling for them changes our estimates.
+As an alternative measure of income we use a prediction obtained from a regression model estimated using the ASHE data.
+
+We noticed that the relationship between discount rate and each dependent variable exhibit similar, characteristic patterns.
+The values of dependent variables tend to be high for small values of discount rate, they are quite low for medium values of discount rate and they increase again for relatively high values of discount rate.
+Because of this parabolic shape, we include both linear and quadratic terms of discount rate as explanatory variables.
+The estimates of all coefficients are almost the same if the squared discount rate is omitted, but we only present the results of the models with both linear and quadratic term to keep the length of our paper within reasonable limit.
+
+The descriptive statistics of the additional explanatory variables can be found in Table~\ref{Descriptive} in Appendix~$3$.
+
+For WTP models we also perform other robustness tests.
+
+\subsection{Climate change knowledge}\label{RobKnow}
+
+In this section we estimate the climate knowledge model specified in Section~\ref{ResKnow} including the behavioural variables and additional potential confounders as discussed at the beginning of this Section~\ref{Robust}.
+The estimates are summarised in Table~\ref{ClimKnowInOut}.
+As we can see, all additional covariates and behavioural variables with except for climate seriousness perception and climate versus policy perception are insignificant.
+
+Climate seriousness perception and climate versus policy perception are negative and strongly significant predictors of climate knowledge (see Table~\ref{ClimKnowInOut}).
+This is not surprising given the fact that climate knowledge is negative and significant when included as an explanatory variable for both of these climate concern variables (see Tables~\ref{Climcare435} and~\ref{Climpol435} in Section~\ref{ResPerc}).
+There are two possible explanations.
+$(i)$ Individuals, who are less educated in climate change tend to believe to incorrect statements and mechanisms which would imply that climate change is much more serious than it actually is \citep{reynolds2010now, WhatPplKnow94}.
+In this case, the direction of dependency would be the other way around.
+$(ii)$ The negative correlation is caused by affective orientation towards global warming \citep{Kahan2015}.
+That is, people who believe in anthropogenic climate change and correctly assert, for example, that usage of fossil fuels is one of the causes of global warming are also likely to affirm other, perhaps false propositions which would imply higher environmental risks, for example that atmospheric emissions of sulphur contribute to global warming \citep{reynolds2010now}.
+
+The estimates of the predictors which were originally chosen by the multisplit lasso are qualitatively the same as those in Section~\ref{ResKnow} (see Table~\ref{Know435}). The estimates are robust.
+
+{\centering
+\begin{threeparttable}
+\singlespacing
+\caption{\vspace{-0.2cm}\textit{\textbf{Climate change knowledge:} Jackknife OLS - robustness}}
+% soubor Know435InOut.R  Know535InOutKm2.R 
+\label{ClimKnowInOut} 
+\centering
+\begin{small}
+\vspace{-1.35cm}
+\begin{tabular}{lcccc} 
+\hline
+  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
+        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
+ \hline \vspace{-0.3cm}
+\\
+% ResKnmow435io
+\vspace{-0.33cm}Gender = male&$0.280$&$1\times10^{-8}~^{***}$&$0.282$&$<2\times10^{-8}~^{***}$\\
+  \\
+\vspace{-0.33cm}Cognitive reflection $=0.5$&$0.863$&$1.000$&$0.934$&$1.000$\\
+  \\
+  \vspace{-0.33cm}Cognitive reflection $=1$&$0.107$&$1.000$ &$0.112$&$1.000$\\
+    \\
+  \vspace{-0.33cm}Cognitive reflection $=1.5$&$0.432$&$1.000$&$0.461$&$1.000$\\
+\\
+\vspace{-0.33cm}Cognitive reflection $=2$&$0.258$&$0.001~^{**}$&$0.255$&$0.003~^{**}$\\
+  \\
+  \vspace{-0.33cm}Cognitive reflection $=2.5$&$0.520$&$1.000$ &$0.537$&$1.000$\\
+    \\
+ Cognitive reflection $=3$&$0.450$&$1\times10^{-7}~^{***}$&$0.448$&$5\times10^{-7}~^{***}$\\
+\hline \vspace{-0.33cm}
+  \\
+\vspace{-0.33cm}Income - predicted (mill. \textsterling~/yr.)&$-0.273$&$1.000$&\multicolumn{2}{c}{\textit{Not included}}\\ %income PREDICTED!!
+  \\
+\vspace{-0.33cm}Income - reported (mill. \textsterling~/yr.)\tnote{a}&\multicolumn{2}{c}{\textit{Not included}}&\textit{varies}&$1.000$\\ %income PREDICTED!!
+  \\
+\vspace{-0.33cm}Net assets (mill. \textsterling)&$0.026$&$1.000$&$0.020$&$1.000$\\ 
+      \\
+\vspace{-0.33cm}People per mill.
+km\textsuperscript{2}-LSOA level&\multirow{1}{*}{$1.416$}&\multirow{1}{*}{$1.000$}&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}\\%ha is a symbol for hectares
+  \\
+\vspace{-0.33cm}People per mill.
+km\textsuperscript{2}-LAD level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$0.254$}&\multirow{1}{*}{$1.000$}\\%ha is a symbol for hectares
+  \\
+  \vspace{-0.33cm}WTP-gas and electricity tax (\textsterling~/ yr.)&$-0.0003$&$1.000$&$-0.0001$&$1.000$\\
+      \\
+  \vspace{-0.33cm}WTP-duty on transport fuel (pence/ yr.)&$0.001$&$1.000$&$0.0006$&$1.000$\\
+    \\
+    \vspace{-0.33cm}Climate seriousness perception&$-0.077$&$<2\times10^{-8}~^{***}$&$-0.076$&$<2\times10^{-8}~^{***}$\\
+  \\
+    \vspace{-0.33cm}Climate vs. policy effects perception&$-0.043$&$0.0009~^{***}$&$-0.044$&$0.001~^{**}$\\
+      \\
+\vspace{-0.33cm}Social value orientation (ring meas.)&\multirow{1}{*}{$0.0008$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$0.0007$}&\multirow{1}{*}{$1.000$}\\
+  \\ \vspace{-0.33cm}\multirow{1}{*}{Inequity aversion (categorical)}&\textit{varies}& \multirow{1}{*}{$1.000$}& \multirow{1}{*}{\textit{varies}}& \multirow{1}{*}{$1.000$}\\      
+    \\ \vspace{-0.33cm}Discount rate yr. from now&$-0.001$&$1.000$&$-0.001$&$1.000$\\  
+      \\
+ \vspace{-0.33cm}Discount rate yr. from now - sq.\tnote{b}&\multirow{1}{*}{$2\times10^{-6}$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$2\times10^{-6}$}&\multirow{1}{*}{$1.000$}\\
+  \\  \vspace{-0.33cm}Risk aversion coefficient\tnote{c}&\multicolumn{2}{c}{\textit{Not included}}&$-0.390$&$1.000$\\          
+    \\
+Redistribution of income (cat.)\tnote{d}&$-$,\textit{varies}&$1.000$&$-$,\textit{varies}&$1.000$\\
+     \hline
+     Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.071$}&\multicolumn{2}{c}{$0.072$}\\
+  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
+\\
+\hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+ \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+ \begin{adjustwidth}{0.9cm}{}
+  \vspace{-0.3cm}
+ \item[a]~Self reported income is only available as categorical.
+\item[b]~If squared discount rate is omitted, linear discount rate remains insignificant and the estimates of the other covariates are almost the same as those presented in this table.
+\item[c]~The risk aversion coefficient is an estimated parameter of a utility function.
+In this model, the mean of power function is used.
+We also estimated varieties of this model with different risk aversion coefficients, particularly means or medians of various utility functions.
+These are power, log, exponential and quadratic.
+The risk aversion parameter is always insignificant and whether it is included or not (or which one) does not affect sign or significance level of any other parameter.
+  \item[d]~A degree of agreement with the statement: 'Government should redistribute income from the better off to those who are less well off.' Included to test for significance of political opinions.
+     \end{adjustwidth}    
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+
+\subsection{Climate change risk perception}\label{RobPerc}
+
+In this section we discuss robustness of the climate risk perception models.
+The models which were specified using lasso in Section~\ref{ResPerc} are re-estimated with behavioural variables and additional potential confounders among explanatory variables.
+The estimates of the climate seriousness perception models are summarised in Table~\ref{ClimCareInOut} and the estimates of the models for climate change versus policy perception can be found in Table~\ref{ClimPolInOut}.
+
+As apparent from Table~\ref{ClimCareInOut}, all the behavioural variables and the additional potential confounders are insignificant for climate seriousness, with exemption of the climate variables.
+The significance of the climate variables is in accordance with our expectation and it confirms that our climate measures are valid.
+The estimates of the predictors selected using lasso are qualitatively equivalent to the estimates in Table~\ref{Climcare435}, hence our results are robust.
+
+Table~\ref{ClimPolInOut} summarises the estimates of models with climate change versus policy as a dependent variable.
+Alike in the case of the climate seriousness model, all the behavioural variables, income, assets and population density are insignificant while the climate variables are significant.
+This is what we expected for the climate measures to be valid.
+However, understanding of inflation is not significant in Table~\ref{ClimPolInOut} while it is strongly significant in the model specified using lasso (see Table~\ref{Climpol435}).
+This is probably a result of the mediation relationship structure among climate versus policy, financial literacy and WTP discussed in Section~\ref{ResWTP}.
+Level of understanding of inflation is correlated with WTP.
+If we remove WTP for gas and electricity and WTP for transport fuel keeping all other variables in, understanding of inflation becomes significant.
+Estimates of this variety can be found in Table~\ref{ClimPolIOnoWTP} in Appendix~$3$.
+This evidence is consistence with our mediation hypothesis.
+The effects of climate knowledge and risk assessment consistency are negative and significant as in the original model summarised in Table~\ref{Climpol435}. The results are, on the whole, robust.
+
+{\centering
+\begin{threeparttable}
+\singlespacing
+\caption{\vspace{-0.35cm}\textit{\textbf{Climate change seriousness perception:} Jackknife OLS - robustness}}
+% soubor Ccare435InOutb.R  Ccare535InOutbKM2.R
+\label{ClimCareInOut} 
+\centering
+\begin{small}
+\vspace{-1.2cm}
+\begin{tabular}{lcccc} 
+\hline
+  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
+        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
+ \hline 
+\vspace{-0.23cm}
+\\
+\vspace{-0.33cm}Gender = male&$-0.302$&$0.001~^{*}$&$-0.285$&$0.001~^{**}$\\
+  \\
+\vspace{-0.33cm}Climate knowledge\tnote{a}&$-0.187$&$<2\times10^{-8}~^{***}$&$-0.182$&$<2\times10^{-8}~^{***}$\\
+  \\
+ Redistribution of income:&\multirow{2}{*}{$0.297$}&\multirow{2}{*}{$1.000$}&  \multirow{2}{*}{$0.298$}&\multirow{2}{*}{$1.000$}  \\%-1
+    \vspace{-0.33cm}   \hspace{0.6cm}disagree\tnote{b}&&&&\\%-1
+    \\
+  \vspace{-0.33cm}Redistribution of inc.: neutral\tnote{b}&$0.256$&$1.000$ &$0.263$&$1.000$\\%0
+    \\
+  \vspace{-0.35cm}Redistribution of income: agree\tnote{b}&$0.778$&$2\times10^{-7}~^{***}$&$0.806$&$8\times10^{-8}~^{***}$\\%1
+    \\
+  Redistribution of income:& \multirow{2}{*}{$0.937$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}&\multirow{2}{*}{$0.939$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
+\hspace{0.6cm}strongly agree\tnote{b}&&&\\%2
+\hline\vspace{-0.33cm}
+  \\
+\vspace{-0.33cm}Income- predicted (mill. \textsterling~/yr.)&$2.760$&$1.000$&\multicolumn{2}{c}{\textit{Not included}}\\ %income PREDICTED!!
+  \\
+\vspace{-0.33cm}Income- reported (mill. \textsterling~/yr.)\tnote{c}&\multicolumn{2}{c}{\textit{Not included}}&\textit{varies}&$1.000$\\ %income PREDICTED!!
+  \\
+\vspace{-0.33cm}Net assets (million \textsterling)&$-0.163$&$1.000$&$-0.310$&$1.000$\\ 
+      \\
+\vspace{-0.33cm}People per mill.
+km\textsuperscript{2}-LSOA level&\multirow{1}{*}{$9.838$}&\multirow{1}{*}{$1.000$}&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}\\%ha is a symbol for hectares
+  \\
+  \vspace{-0.33cm}People per mill.
+km\textsuperscript{2}-LAD level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$-6.520$}&\multirow{1}{*}{$1.000$}\\%ha is a symbol for hectares
+  \\
+  \vspace{-0.33cm}WTP-gas and elec. tax (\textsterling~/ yr.)&$0.002$&$0.0003~^{***}$&$0.002$&$0.0005~^{***}$\\
+      \\
+  \vspace{-0.33cm}WTP-duty on transp. fuel (p./yr.)&$0.005$&$1.000$& $0.005$&$1.000$\\
+\\
+    \vspace{-0.33cm}Climate vs. policy effects perc.&$0.356$&$<2\times10^{-8}~^{***}$&$0.359$&$<2\times10^{-8}~^{***}$\\
+      \\
+Social value orientation&\multirow{2}{*}{$0.003$}&\multirow{2}{*}{$1.000$}&\multirow{2}{*}{$0.003$}&\multirow{2}{*}{$1.000$}\\
+\vspace{-0.33cm} \hspace{0.6cm}(ring measure)&& &\\%2
+  \\% ring measure> ref working paper dolton tol
+    \vspace{-0.33cm}\multirow{1}{*}{Inequity aversion (categorical)}&\textit{$+$, varies}& \multirow{1}{*}{$1.000$}& \multirow{1}{*}{\textit{$+$, varies}}& \multirow{1}{*}{$1.000$}\\      
+    \\ \vspace{-0.33cm}Discount rate yr. from now&$0.0003$&$1.000$&$4\times10^{-5}$&$1.000$\\  
+      \\
+ \vspace{-0.33cm}Discount rate yr. from now - sq.\tnote{d}&\multirow{1}{*}{$-2\times10^{-6}$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$-1\times10^{-6}$}&\multirow{1}{*}{$1.000$}\\
+  \\  \vspace{-0.33cm}Risk aversion coefficient\tnote{e}&\multicolumn{2}{c}{\textit{Not included}}&$0.731$&$1.000$\\          
+   \\
+     \hline
+          Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.247$}&\multicolumn{2}{c}{$0.254$}\\
+  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
+\\
+\hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+
+ \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+ \begin{adjustwidth}{0.9cm}{}
+  \vspace{-0.3cm}
+ \item[a]~Squared term of climate knowledge is insignificant in this version, hence it is not included.
+  \item[b]~Degree of agreement with the following statement: 'Government should redistribute income from the better off to those who are less well off.' The base category is 'Strongly disagree'.
+  \item[c]~Self reported income is only available as categorical.
+  \item[d]~If squared discount rate is omitted, linear discount rate remains insignificant and the estimates of the other covariates are almost the same as those presented in this table.
+  \item[e]~The risk aversion coefficient is an estimated parameter of a utility function.
+In this model, the median of power function is used.
+We also estimated varieties of this model with different risk aversion coefficients, particularly means or medians of various utility functions.
+These are power, log, exponential and quadratic.
+The risk aversion parameter is always insignificant and whether it is included or not (or which one) does not affect sign or significance level of any other parameter.
+     \end{adjustwidth}    
+\singlespacing  
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+
+\pagebreak
+
+{\centering
+\begin{threeparttable}
+\singlespacing
+\caption{\vspace{-0.2cm}\textit{\textbf{Climate versus policy effects perception:} Jackknife OLS - robustness}}
+% soubor Cpol435InOutb.R  Cpol535InOutbKM2.R
+\label{ClimPolInOut} 
+\centering
+\begin{small}
+\vspace{-1.2cm}
+\begin{tabular}{lcccc} 
+\hline
+  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
+    \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adjusted}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adjusted}} \\
+        \multicolumn{1}{l}{\textbf{Variable}}&\multicolumn{1}{c}{}& \multicolumn{1}{c}{\textbf{\textit{p}-value}}&\multicolumn{1}{c}{} &  \multicolumn{1}{c}{\textbf{\textit{p}-value}}\\
+ \hline 
+\vspace{-0.23cm}
+\\
+\vspace{-0.33cm}Climate knowledge\tnote{a}&$-0.127$&$8\times10^{-5}~^{***}$&$-0.128$&$0.0001~^{***}$\\
+  \\
+\vspace{-0.33cm}Understands inflation $=0.5$&$0.051$&$1.000$&$0.028$&$1.000$ \\
+  \\
+\vspace{-0.33cm}Understands inflation $=1$&$-0.219$&$1.000$&$-0.243$&$1.000$\\
+    \\
+  Consistent answers to risk&\multirow{2}{*}{$-0.310$}&\multirow{2}{*}{$0.019~^{*}$}&\multirow{2}{*}{$-0.305$}&\multirow{2}{*}{$0.038~^{*}$}\\
+\hspace{0.6cm}questions $(0/1)$&& &\\
+\hline
+  \\
+\vspace{-0.33cm}Income- predicted (mill. \textsterling~/yr.)&$2.446$&$1.000$&\multicolumn{2}{c}{\textit{Not included}}\\ %income PREDICTED!!
+  \\
+\vspace{-0.33cm}Income- reported (mill. \textsterling~/yr.)\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\textit{varies}&$1.000$\\ %income PREDICTED!!
+  \\
+\vspace{-0.33cm}Net assets (million \textsterling)&$-0.070$&$1.000$&$0.027$&$1.000$\\ 
+      \\
+\vspace{-0.33cm}People per mill.
+km\textsuperscript{2}-LSOA level&\multirow{1}{*}{$-16.683$}&\multirow{1}{*}{$1.000$}&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}\\%ha is a symbol for hectares
+  \\
+\vspace{-0.33cm}People per mill.
+km\textsuperscript{2}-LAD level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$15.548$}&\multirow{1}{*}{$1.000$}\\%ha is a symbol for hectares
+  \\
+  \vspace{-0.33cm}WTP- gas and electric. tax (\textsterling~/ yr.)&$0.002$&$9\times10^{-5}~^{***}$&$0.002$&$9\times10^{-5}~^{***}$\\
+      \\
+  \vspace{-0.33cm}WTP-duty on transport fuel (p./yr.)&$0.011$&$9\times10^{-5}~^{***}$& $0.010$&$0.001~^{**}$\\
+  \\
+    \vspace{-0.33cm}Climate seriousness perception&$0.382$&$\small{<}2\times10^{-8}~^{***}$&$0.387$&$\small{<}2\times10^{-8}~^{***}$\\
+      \\
+Social value orientation&\multirow{2}{*}{$0.006$}&\multirow{2}{*}{$0.396$}&\multirow{2}{*}{$0.005$}&\multirow{2}{*}{$0.726$}\\
+\vspace{-0.33cm} \hspace{0.6cm}(ring measure)&& &\\%2
+  \\% ring measure> ref working paper dolton tol
+ \vspace{-0.33cm}\multirow{1}{*}{Inequity aversion (categorical)}&\textit{varies}& \multirow{1}{*}{$1.000$}& \multirow{1}{*}{\textit{varies}}& \multirow{1}{*}{$1.000$}\\      
+    \\ \vspace{-0.33cm}Discount rate yr. from now&$-0.001$&$1.000$&$-0.001$&$1.000$\\  
+      \\
+ \vspace{-0.33cm}Discount rate yr. from now - sq.\tnote{c}&\multirow{1}{*}{$2\times10^{-6}$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$2\times10^{-6}$}&\multirow{1}{*}{$1.000$}\\
+  \\  \vspace{-0.33cm}Risk aversion coefficient\tnote{d}&\multicolumn{2}{c}{\textit{Not included}}&$-0.523$&$1.000$\\          
+   \\
+  \vspace{-0.35cm}Redistribution of income (categ.)\tnote{e}&\textit{$+$,varies}&$1.000$&\textit{$+$,varies}&$1.000$\\%1
+    \\
+     \hline
+          Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.254$}&\multicolumn{2}{c}{$0.257$}\\
+  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
+\\
+\hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+ \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+ \begin{adjustwidth}{0.9cm}{}
+  \vspace{-0.3cm}
+ \item[a]~Squared term of climate knowledge is insignificant in this version, hence it is not included.
+\item[b]~Self reported income is only available as categorical.
+\item[c]~If squared discount rate is omitted, linear discount rate remains insignificant and the estimates of the other covariates are almost the same as those presented in this table.
+\item[d]~The risk aversion coefficient is an estimated parameter of a utility function.
+In this model, the mean of power function is used.
+We also estimated varieties of this model with different risk aversion coefficients, particularly means or medians of various utility functions.
+These are power, log, exponential and quadratic.
+The risk aversion parameter is always insignificant and whether it is included or not (or which one) does not affect sign or significance level of any other parameter.
+\item[e]~A degree of agreement with the statement: 'Government should redistribute income from the better off to those who are less well off.' Included to test for significance of political opinions.
+     \end{adjustwidth}   
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+
+\vspace{0.2cm}
+
+\subsection{Willingness to pay for climate change mitigation}\label{RobWTP}
+
+In this section we verify robustness of the models with dependent variable  WTP for climate change mitigation.
+Table~\ref{GasInOut} summarises estimates of two varieties of the models specified in Section~\ref{ResWTP} with additional behavioural variables, climate change variables and other potential confounders discussed at the beginning of Section~\ref{Robust}.
+The estimates of the predictors which were originally selected by the multisplit lasso (in the first $11$ rows of Table~\ref{GasInOut}) are qualitatively the same as those of the models in Table~\ref{Gas435cp}.
+The only exception is the dummy variable for scoring $0.5$ on understanding of compound interest as its estimate is negative in Table~\ref{Gas435cp} while it is positive in both models in Table~\ref{GasInOut}.
+However, the estimate of this dummy variable is insignificant in each of these models so the difference in signs does not imply that the estimates are not robust.
+
+Similarly as in the case of the robustness model for climate knowledge (Table~\ref{ClimKnowInOut}) and the robustness models for climate change risk perception (Tables~\ref{ClimCareInOut} and~\ref{ClimPolInOut}) the climate change variables are significant (with the exemption of climate knowledge) while the other potential confounders and behavioural variables in the second half of Table~\ref{GasInOut} are insignificant.
+An exemption is predicted income, which has a significant impact on WTP.\footnote{Interestingly, the categorical income recorded in the survey \citep{SurveyUK} is not significant.
+We suspect that it can be due to inaccuracy of the income variable recorded in the survey as relatively big number of participants are students and it is not clear if they stated their own income or income of their parents.
+The possible inaccuracy of the income measure is one of the reasons why we also use the alternative predicted income.} The significance of income can be explained by its diminishing marginal utility.
+For people with higher income, the utility of amount of money paid as climate tax is lower than for people with lower income.
+Therefore, preferred tax rates are higher for higher income groups.
+Whether or not income or the other additional explanatory variables are included does not have any significant impact on estimates of the predictors in Table~\ref{GasInOut} which were chosen by the means of the multisplit lasso in Section~\ref{ResWTP}.
+
+Pride may have played role in the strong significance of current tax rates estimates.
+As explained in Section~\ref{ClimateVars}, the respondents were informed about the correct tax rates after they gave their estimates and before they were asked about their preferences regarding the climate policies.
+It is possible that respondents tended to give preferred tax rates which were close to their estimate of the current tax rates feeling that their estimate should be the preferred one.
+
+{\centering
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{\small{WTP climate - gas and electricity tax:}} Jackknife OLS - robustness}}
+% soubor Gas435InOut3.R a %Gas535InOutKm2b.R
+% % soubor Gas435InOut5.R a %Gas535InOutKm2c.R jeste to odhaduju znovu a davam tam ty kategoricky jako scale, jen abych vedela jestli muzu napsat negative/positive correlation
+
+\label{GasInOut} 
+\centering
+\begin{small}
+\vspace{-1cm}
+\begin{tabular}{lcccc} 
+\vspace{-0.8cm}
+\\
+\hline
+  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}% Gas435InOut3.R
+  &\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\%Gas535InOutKm2b.R
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
+ \multicolumn{1}{l}{\textbf{Variable}} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adj.
+\textit{p}-value}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adj.
+\textit{p}-value}} \\
+\hline  \vspace{-0.345cm}
+\\
+\vspace{-0.345cm}Age (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
+  \\
+\vspace{-0.345cm}Climate vs. policy effects perc.&$6.654$&$<2\times10^{-8}~^{***}$&$6.534$&$<2\times10^{-8}~^{***}$\\
+  \\
+  \vspace{-0.345cm}Inequity aversion (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{**}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{**}$}\\
+      \\
+ Equal intergenerational&\multirow{2}{*}{$18.843$}&\multirow{2}{*}{${0.004}~^{**}$}&\multirow{2}{*}{$19.646$}&\multirow{2}{*}{${0.004}~^{**}$}\\
+\vspace{-0.345cm}\hspace{0.6cm}allocation of resources $(0/1)$&\\
+ \\  
+\vspace{-0.345cm}Understands comp. interest $=0.5$&$1.276$&$1.000$&$3.870$&$1.000$ \\
+  \\
+\vspace{-0.345cm}Understands comp. interest $=1$&$-34.032$&$0.0003~^{***}$&$-32.084$&$0.001~^{***}$\\
+  \\  
+\vspace{-0.345cm}Understands inflation$=0.5$&$-12.364$&$1.000$&$-11.779$&$1.000$\\
+  \\
+\vspace{-0.345cm}Understands inflation$=1$&$-31.737$&$<2\times10^{-8}~^{***}$&$-29.101$&$<2\times10^{-8}~^{***}$\\
+    \\
+Consistent answers to risk $(0/1)$&\multirow{1}{*}{$-28.046$}&\multirow{1}{*}{$<2\times10^{-8}~^{***}$}&\multirow{1}{*}{$-27.714$}&\multirow{1}{*}{$<2\times10^{-8}~^{***}$}\\
+\hline \vspace{-0.345cm}\\
+\vspace{-0.345cm}Income- predicted (thousand \textsterling/yr.)&$0.703$&$3\times10^{-5}~^{***}$&\multicolumn{2}{c}{\textit{Not included}}\\ %income PREDICTED!!
+  \\
+\vspace{-0.345cm}Income- reported (mill. \textsterling/yr.)\tnote{a}&\multicolumn{2}{c}{\textit{Not included}}&$+$,\textit{varies}&$1.000$\\ %income PREDICTED!!
+  \\
+\vspace{-0.345cm}Net assets (million \textsterling)&$18.033$&$0.133$&$20.072$&$0.151$\\ 
+      \\
+\vspace{-0.345cm}People per mill.
+km\textsuperscript{2}-LSOA level&\multirow{1}{*}{$768.974$}&\multirow{1}{*}{$1.000$}&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}\\%ha is a symbol for hectares
+  \\
+\vspace{-0.345cm}People per mill.
+km\textsuperscript{2}-LAD level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$594.369$}&\multirow{1}{*}{$1.000$}%ha is a symbol for hectares
+  \\  \\
+\vspace{-0.345cm}Climate knowledge&$0.020$&$1.000$&$0.556$&$1.000$\\
+      \\
+  \vspace{-0.345cm}How much is tax gas and el.\small{(\textsterling~/yr.)}&$0.252$&$\small{<}2\times10^{-8}~^{***}$&$0.251$&$\small{<}2\times10^{-8}~^{***}$\\
+      \\
+    \vspace{-0.345cm}Climate seriousness perception&$7.158$&$\small{<2\times10^{-8}}~^{***}$&$7.282$&$\small{<}2\times10^{-8}~^{***}$\\
+      \\
+\vspace{-0.345cm}Social val.
+orientation (ring meas.)&\multirow{1}{*}{$0.048$}&\multirow{1}{*}{$1.000$}&\multirow{1}{*}{$0.009$}&\multirow{1}{*}{$1.000$}\\ % ring measure> ref working paper dolton tol
+    \\ \vspace{-0.34cm}Discount rate yr. from now&$-0.078$&$1.000$&$-0.084$&$1.000$\\  
+      \\
+ \vspace{-0.345cm}Discount rate yr. from now - sq.\tnote{b}&\multirow{1}{*}{$0.0002$}&\multirow{1}{*}{$0.992$}&\multirow{1}{*}{$0.0002$}&\multirow{1}{*}{$0.908$}\\
+  \\  \vspace{-0.345cm}Risk aversion coefficient\tnote{c}&\multicolumn{2}{c}{\textit{Not included}}&$-0.261$&$1.000$\\          
+   \\
+  \vspace{-0.35cm}Redistribution of income (cat.)\tnote{d}&\textit{varies}&$1.000$&\textit{varies}&$1.000$\\
+    \\
+     \hline
+Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.370$}&\multicolumn{2}{c}{$0.364$}\\
+  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
+\\
+\hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+ \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$;\hspace{0.4cm} 
+  \vspace{-0.4cm}\large{\textsuperscript{\textbf{a}}}~\footnotesize{Age and reported income only available}\\ \begin{adjustwidth}{0.7cm}{} \footnotesize{as  categorical.
+Inequity av.
+treated as categorical (see Sec.~\ref{Method})\hspace{0.4cm}}
+  \large{\textsuperscript{\textbf{b}}}~\footnotesize{If squared discount rate is omitted, linear discount rate remains insignificant and the estimates of the other covariates are almost the same as those presented in this table.\hspace{0.4cm}}
+  \large{\textsuperscript{\textbf{c}}}~\footnotesize{The risk aversion coefficient is an estimated parameter of a utility function.
+In this model, the mean of power function is used.
+We also estimated varieties of this model with different risk aversion coefficients, particularly means or medians of power, log, exponential or quadratic utility function.
+The risk aversion parameter is always insignificant and whether it is included or not (or which one) does not affect sign or significance level of any other parameter.\hspace{0.4cm}}\normalsize{\textsuperscript{\textbf{d}}}~\footnotesize{A degree of agreement with the statement: 'Government should redistribute income from the better off to those who are less well off.'}
+     \end{adjustwidth}     
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+
+\pagebreak
+
+As another robustness test we use preferred tax rates on transport fuel as a measure of WTP.
+The estimates are summarised in Table~\ref{FuelInOut}.
+Model~$1$ in Table~\ref{FuelInOut} only includes explanatory variables which were chosen for WTP using the lasso in Section~\ref{ResWTP}.
+Model~$2$ in Table~\ref{FuelInOut} also includes climate variables, behavioural variables and additional potential confounders to test their significance and to examine whether or not their inclusion changes signs or significance levels of the main predictors.
+Comparing Model~$1$ in Table~\ref{FuelInOut} with the estimates of the models in Table~\ref{Gas435cp} (Section~\ref{ResWTP}), we can see that the signs and significance levels of the coefficients are the same in these two models.
+Also their magnitude is comparable considering the different scales of the two dependent variables.
+Regarding Model~$2$ in Table~\ref{FuelInOut}, we can see that inclusion of the additional explanatory variables does not change signs or significance levels of the main predictors (the variables in the first $11$ rows of Table~\ref{FuelInOut}).
+An exemption is dummy variable for scoring $0.5$ on understanding of inflation as in Model~$1$ the coefficient is negative while in Model~$2$ it is positive.
+However, this change is unremarkable as these coefficients of understanding of inflation are insignificant in both models.
+The additional explanatory variables in Model~$2$ in Table~\ref{FuelInOut} are insignificant except of the estimate of current fuel duty and climate seriousness perception.
+The signs of all the additional variables are the same as their signs in the model for WTP on gas and electricity in Table~\ref{GasInOut}.
+Interestingly, income does not have any significant impact on preferred fuel duty although its impact on gas and electricity tax is significant.
+Our explanation is that individuals with lower income exhibit lower WTP through gas and electricity tax, but they are less likely to own a car (thus they are less likely to be eligible for paying transport fuel duty).
+Thus, when asked about their preferred rates on fuel duty, they actually impose the duty on those owning a car rather than on themselves.
+It is understandable that preferred tax rates imposed on others are higher than preferred tax rates paid by ourselves.
+Hence, people with lower income may exhibit higher preferred fuel duty as they impose it on others rather than on yourself.
+This may offset the significant positive impact of income, which we detected in the model for gas and electricity tax.
+Even lower income groups are eligible for gas and electricity tax, therefore the positive influence of income is significant as it is not offset by the fact that lower income groups impose the tax on others rather than on themselves.
+
+We can conclude that our models are reasonable robust.
+
+{\centering
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{\vspace{-0.6cm} WTP climate - duty on transport fuel:} Jackknife OLS}}
+% soubor FuelGas4cp.R a %.FuelGas4bInOut.R % soubor FuelGas5cp.R  %.FuelGas4bInOut4.R
+\label{FuelInOut} 
+\centering
+\begin{small}
+\vspace{-0.8cm}
+\begin{tabular}{lcccc} 
+\\
+\hline
+  \multicolumn{1}{l}{}&\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{1}$}% FuelGas4cp.R
+  &\multicolumn{2}{c}{\textbf{Model} $\boldsymbol{2}$}\\%.FuelGas4bInOut.R
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+  \multicolumn{1}{l}{} & \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}}& \multicolumn{1}{c}{\textbf{Aggreg.}} &  \multicolumn{1}{c}{\textbf{Aggreg.}} \\
+ \multicolumn{1}{l}{\textbf{Variable}} & \multicolumn{1}{c}{\textbf{coef.}}& \multicolumn{1}{c}{\textbf{adj.
+\textit{p}-val.}}& \multicolumn{1}{c}{\textbf{coef.}} &  \multicolumn{1}{c}{\textbf{adj.
+\textit{p}-val.}} \\
+\hline  \vspace{-0.32cm} 
+\\ 
+\vspace{-0.32cm}Age (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{***}$}\\
+  \\
+\vspace{-0.32cm}Climate vs. policy effects perception&$2.192$&$<2\times10^{-8}~^{***}$&$1.494$&$<2\times10^{-8}~^{***}$\\
+  \\
+  \vspace{-0.32cm}Inequity aversion (categorical)\tnote{a}&\multicolumn{2}{c}{\textit{negative cor.~}$^{**}$}&\multicolumn{2}{c}{\textit{negative cor.~}$^{**}$}\\
+      \\
+ Equal intergenerational&\multirow{2}{*}{$4.461$}&\multirow{2}{*}{${0.002}~^{**}$}&\multirow{2}{*}{$4.824$}&\multirow{2}{*}{${0.0004}~^{***}$}\\
+\vspace{-0.32cm}\hspace{0.6cm}allocation of resources $(0/1)$&\\
+ \\  
+\vspace{-0.32cm}Understands compound interest$=0.5$&$1.703$&$1.000$&$1.818$&$1.000$ \\
+  \\
+\vspace{-0.32cm}Understands compound interest$=1$&$-7.363$&$0.001~^{***}$&$-7.091$&$0.0008~^{***}$\\
+  \\
+\vspace{-0.32cm}Understands inflation$=0.5$&$-0.602$&$1.000$&$0.135$&$1.000$\\
+  \\
+\vspace{-0.32cm}Understands inflation$=1$&$-7.458$&$<2\times10^{-8}~^{***}$&$-5.834$&$4\times10^{-7}~^{***}$\\
+    \\
+  Consistent answers to risk&\multirow{2}{*}{$-6.294$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}&\multirow{2}{*}{$-5.591$}&\multirow{2}{*}{$<2\times10^{-8}~^{***}$}\\
+\hspace{0.6cm}questions $(0/1)$&& &\\
+\hline \vspace{-0.3cm}\\
+\vspace{-0.32cm}Income - predicted (thousands \textsterling~/yr.)&\multicolumn{2}{c}{\textit{Not included}}&$88.788$&$0.344$ \\ %income PREDICTED!!
+  \\
+\vspace{-0.32cm}Net assets (million pounds \textsterling)&\multicolumn{2}{c}{\textit{Not included}}&$3.168$&$0.936$\\ 
+      \\
+\vspace{-0.32cm}People per mill.
+km\textsuperscript{2} - LSOA level&\multicolumn{2}{c}{\multirow{1}{*}{\textit{Not included}}}&\multirow{1}{*}{$43.595$}&\multirow{1}{*}{$1.000$}\\%ha is a symbol for hectares
+  \\
+\vspace{-0.32cm}Climate knowledge&\multicolumn{2}{c}{\textit{Not included}}&$0.210$&$1.000$\\
+      \\
+  \vspace{-0.32cm}How much is duty transp. fuel (p./yr.)&\multicolumn{2}{c}{\textit{Not included}}&$0.311$&$<2\times10^{-8}~^{***}$\\
+      \\
+    \vspace{-0.32cm}Climate seriousness perception&\multicolumn{2}{c}{\textit{Not included}}&$1.327$&$<2\times10^{-8}~^{***}$\\
+           \\
+Social value orientation&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{$0.004$}&\multirow{2}{*}{$1.000$}\\
+\vspace{-0.32cm} \hspace{0.6cm}(ring measure)&&&\\ 
+      \\
+ \vspace{-0.32cm}Discount rate year from now&\multicolumn{2}{c}{\textit{Not included}}&\multirow{1}{*}{$-0.011$}&\multirow{1}{*}{$1.000$}\\     \\     
+ \vspace{-0.32cm}Discount rate year from now - sq.\tnote{b}&\multicolumn{2}{c}{\textit{Not included}}&\multirow{1}{*}{$3\times10^{-5}$}&\multirow{1}{*}{$1.000$}\\     
+      \\
+Redistribution of income&\multicolumn{2}{c}{\multirow{2}{*}{\textit{Not included}}}&\multirow{2}{*}{\textit{varies}}&\multirow{2}{*}{$1.000$}\\
+\vspace{-0.32cm} \hspace{0.6cm}(categorical)\tnote{c}&&&\\
+  \\    
+     \hline
+Mean adjusted \textit{R\textsuperscript2}:&\multicolumn{2}{c}{$0.244$}&\multicolumn{2}{c}{$0.309$}\\    
+  Observations: &\multicolumn{2}{c}{$5749$}&\multicolumn{2}{c}{$5659$}
+\\
+\hline
+\end{tabular} 
+\end{small}
+ \begin{tablenotes}
+  \begin{footnotesize}
+ \item \textit{Notes:}\hspace{0.2cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+ \begin{adjustwidth}{0.9cm}{}
+  \vspace{-0.2cm}
+\item[a]~Age is only available as categorical. Inequity aversion is treated as categorical (see Section~\ref{BehaviouralVars}).
+\item[b]~If squared discount rate is omitted, linear discount rate remains insignificant and the estimates of the other covariates are almost the same as those presented in this table.
+\item[c]~A degree of agreement with the statement: 'Government should redistribute income from the better off to those who are less well off.' Included to test for significance of political opinions.
+     \end{adjustwidth}       
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+
+\pagebreak
+
+\section{Summary}\label{Summary}
+We exploited a unique dataset of nearly $6000$ observations which combines advantages of survey and experimental methods \citep{SurveyUK}.
+Among almost $70$ explanatory variables and much more interactions we identified the main predictors of climate knowledge, climate risk perception and WTP for climate change mitigation.
+An important part of our analysis was testing of effects of four behavioural variables, in particular social value orientation, time preferences, risk preferences and attitudes towards
+inequality on the climate change variables. The measures of these variables were implied using data from the survey conducted recently in the UK 
+\citep{SurveyUK}.
+
+Using a multisplit lasso we helped to understand the relationship structure among the climate variables and the behavioural measures.
+The lasso estimator has been shown to be very powerful in high-dimensional context as it yields sparse and interpretable results \citep{pValsLasso}.
+We used \textit{p}-values proposed by~\citet{pValsLasso} which were shown to be a good tool for control of both family-wise error and false discovery rate.
+
+With exemption of inequity aversion, we did not find any of the behavioural measures to have significant effect on the climate variables.
+According to our results, the rate of inequity aversion has a significant impact on WTP for climate change mitigation.
+The value of our inequity aversion measure is equal to one of $16$ distinct values for all respondents in our dataset (given the way of its construction and the data recorded in our survey and experiments).
+Therefore, we treat the inequity aversion rate as categorical.
+We can say that its effects on WTP are largely positive and decreasing, although the impacts vary in magnitude and significance over the categories without clear pattern.
+Besides rate of inequity aversion, the most important predictors of WTP are age, perception of intergenerational allocation of resources, financial literacy which we also use as a proxy for numeracy, climate versus policy effects perception which is a partial mediator of understanding of inflation.
+Consistently with previous literature the impact of age and numeracy is negative \citep{Hamilton2011}.
+Expectedly, WTP is higher for individuals who feel to be more affected by climate change than by climate policy.
+
+We found that climate knowledge is higher for men and for individuals with higher level of cognitive ability.
+
+We considered two measures of climate risk perception.
+The first one is based on the respondents' opinion on level of climate change seriousness.
+The other one is referred to as climate versus policy perception and it records respondents' opinions on whether they are more affected by climate change or by climate policy.
+The first question is meant to be answered immediately without thinking by most people while we expected the respondents to take more time before  answering the latter one.
+For climate seriousness, the important predictors are gender, climate knowledge and opinion on whether or not government should redistribute income from the better off to those who are less well off, which we consider as a proxy for ideological or political world-views.
+The drivers of climate versus policy perception are climate knowledge, degree of understanding of inflation and risk assessment consistency.
+It is noticeable that with exemption of climate knowledge (which is common to both) the predictors of the first climate measure which is meant to be answered without thinking and therefore more intuitively are likely to be correlated with personality traits.
+On the other hand, the predictors of the latter climate risk question, which is considered to need more time and thinking before giving an answer, are likely to be correlated with cognitive ability and analytical reasoning.
+
+As robustness tests, we used alternative measure of WTP and for each dependent variable we re-estimated the models with additional potential confounders.
+Based on them we can conclude that our estimates are robust.
+
+% with regard to climate knowledge?? but also relevant for climate concern...
+
+%Another important question is, what actually is 'scientific consensus'?
+\citet{Cook2013COnsensus} and \citet{Kahan2015} argue, that an opinion or a position, which is agreed on by most articles on a given topic published in peer-reviewed scientific journals is a scientific consensus.
+But is scientific consensus actually the best available evidence or a theory which is closest to absolute objective truth?
+Is it reasonable to assume that individuals with high degree of numeracy, analytical reasoning and scientific literacy will tend to believe to the scientific consensus, if considered from purely objective point of view, independently of what stance toward the topic (in our case climate change) express their identity?
+Publication bias casts doubt on this assumption.
+Publication bias typically occurs when a result of a study affects the decision whether to publish the results or not.
+In particular, it can arise if there is a preference to publish statistically significant results, which is usually the case.
+Therefore, every average or counts of significant versus insignificant studies is likely to be biased and misleading \citep{PubBias}.
+\citet{Cook2013COnsensus} review abstracts of nearly $12000$ scientific articles about global climate change published in peer-reviewed journals during period $1991-2011$.
+\citet{Cook2013COnsensus} argue that $64.6\%$ of these abstracts express no position on anthropogenic global warming and among the rest of them, $97.1\%$ agree that global warming is caused by humans.
+Hence, anthropogenic global warming is a scientific consensus.
+However, the scientific consensus does not take into account publication bias, therefore the results can be biased.`
+
+\section{Limitations and further research}\label{Caveats}
+
+We conducted our research within some unavoidable constrains, mostly related to our dataset and time restriction.
+As a result, the study has some limitations.
+Caveats are discussed in this section.
+We also outline suggestions for further research in this section.
+
+The sample is not exactly representative of the UK adult population as it was conducted online.
+The highest age category is slightly under-sampled and the lowest two age categories are slightly over-sampled.
+However, as we can see in Table~\ref{SexAge}, the age and sex distribution in our sample is relatively comparable to the age and sex distribution of the UK population.
+Also, more respondents tended to drop out on more complicated questions.
+Thus, the sample of respondents who finished the survey is slightly biased towards those who are unafraid of hard questions.
+
+As a measure of climate knowledge we adopted the OCSI instrument developed by \citet{Kahan2015} and we found it to be significantly higher for men.
+For further research, it would be particularly interesting to let a woman propose a set of similar questions which would constitute an alternative climate knowledge measure and to test whether we would find  a significant effect of gender on the alternative climate knowledge measure.
+
+The data on WTP were enquired in the survey as follows.
+The respondents were first asked how much they think the climate duty currently is.
+Then they were told the correct answer and then they were asked about their preferred tax rates.
+We believe that for further analysis it would be useful to ask half of the respondents about their preferred tax rates before telling them the correct answer or both before and after informing them about the actual tax rates.
+This could enable us to estimate effect of information.
+
+Unfortunately, the survey which collected our data did not include direct questions about political opinions.
+Hence, our next suggestion for future research is to collect data about political opinions directly and test their significance as well as significance of their interactions with measures of numeracy and cognitive ability on climate knowledge, climate risk perception and WTP for climate change mitigation.
+Also, if a similar survey will be conducted in future, we would like to suggest including questions about method of travel to work.
+This information could be particularly helpful in identifying who answered the desired duty question assuming that the rate will apply to themselves and who, on the other hand, implied the duty on others.
+The respondents travelling by car are probably assuming that the fuel duty will apply to themselves, while those travelling by train or bike are likely to be implying the tax on others.
+If these data were available, it would also be interesting to examine the gap between desired duty imposed on others and desired climate duty implied on oneself.
+
+Other possible direction of further research would be to verify whether some natural disaster happened during the period of survey and if so, how did it affect public environmental attitudes.
+This could be also analysed using data from different surveys.
+Scrutinising weather data at the time of survey and investigating their possible effects on stated environmental attitudes could be another useful approach.
+
+An interesting and relevant question is what actually 'scientific consensus' is.
+\citet{Kahan2015} and \citet{Cook2013COnsensus} argue that a scientific consensus is an opinion or a position, which is agreed on by most articles on a given topic published in peer-reviewed scientific journals during given time period.
+According to this definition, anthropogenic climate change is a scientific consensus.
+But is scientific consensus always the best available evidence?
+Publication bias\footnote{Publication bias typically occurs when a result of a study affects the decision whether to publish the results or not.}, which often occurs as there is usually a preference to publish statistically significant results, casts doubt on this assumption.
+Counts of significant versus insignificant studies are likely to be biased and misleading \citep{PubBias}.
+This should be taken into account when interpreting results about environmental knowledge and concerns.
+
+\section{Concluding remarks and policy implications}\label{Implications}
+
+Consistently with previous literature, we reveal that people who incline to communitarian, pro-social world-view tend to be more concerned about climate change than those who incline towards competitive, individualistic ideology (\citealp{Leiserowitz2013, Kahan2012, WHITMARSH2011}).
+This polarization, which increases with higher degree of numeracy and cognitive ability, hinders efforts to mitigate climate change and its consequences.
+Reducing the association between ideological or political opinions and attitudes towards global warming would be a good step towards alleviation of climate change.
+In practice, this could be achieved for example by more cautious utilization of ideological and political polarization over climate change in political campaigns.
+
+We found that cognitive reflection and financial literacy are among the most important factors affecting the climate knowledge and climate perception.
+Therefore, we would like to empathise importance of education in these areas.
+Our results further suggest that the effect of age on climate concerns is negative and significant.
+Possible explanation is that much less was known about climate change back in times when older people were in education system.
+Hence, they were not exposed to the same level of information about climate change as younger generations.
+In addition, older people are less likely to be proficient in working with internet which is probably the most common source of information today.
+Therefore, we believe that more information provided through media which are easily accessible also for elderly people would help them to improve understanding of consequences of global warming for them and their offspring.
+We would like to encourage policy makers to facilitate support towards educating middle age and older people about climate change and towards motivating them to further educate themselves in this area.
+
+According to our results, income has significant positive effect on preferred gas and electricity tax rates.
+We suggest that this should be further examined and taken into account when making decisions about climate tax rates.
+In particular, we suggest to investigate possibility of introducing different tax rates for different income groups.
 

### 2018-11-15 Rcodes/November/KenARe11.R
working hardly

@@ -1,0 +1,43 @@
+rm(list=ls())
+library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+# KenARe11 is the model which I presented on the meeting 14.11.2018
+# arma(11) and ar(2) the best. AIC a bit better for 11>> claiming KenARe11 the best !!!
+
+
+KenARe11<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+              +AvgTemp + SDTemp  + HWDays, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+              data=ScaledTS,na.action=na.exclude); summary(KenARe11)
+acf(summary(KenARe11)$resid); pacf(summary(KenARe11)$resid)
+anova(KenARe11)
+vif(KenARe11)
+exp(summary(KenARe11)$coef[[1]])
+
+#----------------------------------------------------------------------------------------------------------------------------------------------
+
+# so now try changes suggested during the meeting:
+
+KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+              +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+              data=ScaledTS,na.action=na.exclude); summary(KEN11a)
+nobs(KEN11a); anova(KEN11a)
+summary(KEN11a)$tTable[,1]
+summary(KEN11a)$tTable[,2]
+
+
+                    KENDAL11a<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                                    +AvgTemp + SDTemp  +(1|ID1),data=ScaledTS) 
+                    summary(KENDAL11a)
+                    anova(KENDAL11a,type="I")
+                          # Just playing:
[…]

### 2018-11-15 writing/draft1/draftMonika6.tex
adding
working hardly
working

--- 2018-11-14 writing/draft1/draftMonika5.tex
@@
 \begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11a
+\label{KenARe11} 
+\begin{footnotesize}
+
+\begin{tabular}{llrllll}
+
+\hline \vspace{-0.2cm} \\
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{All counties}} &\multicolumn{2}{c}{\textit{ASAL}} &\multicolumn{2}{c}{\textit{non-ASAL}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&$exp(\beta)$&F-value%> anova(KEN11a, type='marginal')
+    &$exp(\beta)$&F-value&$exp(\beta)$&F-value\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$1.296$ $(1.060)^{***}$&$19.916$&$1.276$&$5.230$&$1.410^{}$&\\
+  \\
+\vspace{-0.2cm}Prec. total&$1.081$ $(1.034)^{*}$&$5.402$&$1.006^{**}$&$0.022$&$1.278^{**}$&\\
+  \\
+  \vspace{-0.2cm}Prec. total sq.&$0.973$ $(1.013)^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{\bullet}$&\\
+    \\
+      \vspace{-0.2cm}Prec. c. of var.&$0.924$ $(1.044)^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$\\
+  \\
+
+         \vspace{-0.2cm}Dry spell -length&$0.935$ $(1.031)^{*}$&$4.810$&$0.833$ ${}$&$6.969$&$ 0.988^{}$&\\
+  \\
+        \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939$ $(1.029)^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&\\
+  \\
+          \vspace{-0.2cm}Temp. - average&$0.819$ $(1.059)^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&\\
+  \\
+
+          \vspace{-0.2cm}Temp. std. dev.&$1.043$ $(1.024)^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&\\
+  \\
+
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+ \\ 
+ \hline
+\vspace{-0.2cm} \\
+\textit{Number of observations:}  &$1300$&$698$& \\
+\vspace{-0.2cm}
+\\
+
+  \hline
+  \vspace{-0.2cm} \\
+
+\hline
+\vspace{-0.2cm}
+\end{tabular} 
+\end{footnotesize}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
+        \begin{adjustwidth}{1cm}{} 
+    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
+
+{\hspace{-1.5cm}
+\begin{threeparttable}
 

### 2018-11-15 Rcodes/November/KenARe11.R

@@ -24,14 +24,14 @@ KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
               +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
               data=ScaledTS,na.action=na.exclude); summary(KEN11a)
 nobs(KEN11a); anova(KEN11a)
-summary(KEN11a)$tTable[,1]
-summary(KEN11a)$tTable[,2]
-
+exp(summary(KEN11a)$tTable[,1])
+exp(summary(KEN11a)$tTable[,2])
+anova(KEN11a, type='marginal')
 
                     KENDAL11a<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                                     +AvgTemp + SDTemp  +(1|ID1),data=ScaledTS) 
                     summary(KENDAL11a)
-                    anova(KENDAL11a,type="I")
+                    anova(KENDAL11a);     anova(KENDAL11a,type="I")
                           # Just playing:
                                 KAL11a<-lm(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp+SDTemp,data=ScaledTS)
                                 anova(KAL11a)
@@ -41,3 +41,15 @@ summary(KEN11a)$tTable[,2]
 KEN11b<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + I(sign(AvgTemp)*(AvgTemp^2))+SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
             data=ScaledTS,na.action=na.exclude); summary(KEN11b)  # ok, square not significant
+
+#----------------------------------------------------------------------------------------------------------------------------------------------
+
+# and subsamples
+
+KEN11a_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
[…]

### 2018-11-15 LiteratureFS/MyReview/LitReviewDrought.tex
blll

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{a4wide}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage{colortbl}
\usepackage{footnote}
\makesavenoteenv{tabular}
\usepackage[flushleft]{threeparttable}
\usepackage[hyphens]{url}
\usepackage{placeins}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{changepage}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{tabularx}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage[labelfont=bf]{caption}
\usepackage{titlesec}
\usepackage{endnotes}
\usepackage{csquotes}
\usepackage{epstopdf}

\date{\normalsize{September 2017}}
\title{\Large \bf Climate Change Awareness and Willingness to Pay for its Mitigation: \\ Evidence from the UK}
\author{Monika Novackova and Richard Tol}
\affil{\small{Department of Economics, University of Sussex, Falmer, UK}}

\parindent 0pt
\parskip 0.5em
\newcommand\starred[1]{\accentset{~~~~~\star}{#1}}

\newcounter{magicrownumbers}
\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}

\begin{document}
\sloppy
\section*{Definition of drought}
According to the international meteorological community, drought can be defined in several ways. In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
 \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater . The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).

Numerous definitions of drought and their role are reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}. They distinguish two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational. Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}. Operational definitions are essential for an effective early warning system. An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.

\subsubsection*{Drought indices}
A numerical measure is needed to compare severity of drought across different time periods or geographical locations. However, as a result of a large disagreement about a definition of drought, there is no single universal drought index. Instead of that a number of measures of drought has been developed (\citealp{Heim2002, wilhite1985, wilhite2000}).

Examples of early measures of drought are \cite{wilhite1985}, \cite{munger1916}, \cite{blumenstock1942} or \cite{mcquigg1954}. \cite{munger1916} suggested to use length of period without $24$-h precipitation of $1.27$ mm. \cite{wilhite1985} is based on a measure of precipitation over a given time period. \cite{blumenstock1942} proposed to measure severity of drought as a length of drought in days where the end of a drought is defined by occurrence of $2.54$ mm of precipitation in $48$ hours. \cite{mcquigg1954} developed the Antecedent Precipitation Index (API) which is based on amount and timing of precipitation and it was used for forecasting of floods. Hence, the API is a reverse drought index.

The study of \cite{palmer1965} was a significant milestone in the history of quantification of drought severity. \cite{palmer1965} developed the Palmer Drought Severity Index (PDSI) using a complex water balance model. The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand (\citealp{Heim2002,palmer1965}). As the PDSI suffers from several weaknesses (for details see e.g. \citealt{Heim2002}), other indices were developed in the following decades. These include the standardized precipitation index (SPI) developed by \cite{SPI} and the standardized precipitation evapotranspiration index (SPEI) developed by \cite{SPEI}. The SPI specifies observed precipitation as a standardised departure from a chosen probability distribution which models the precipitation data. Values of SPI can be viewed as a multiple of standard deviations by which the observed amount of rainfall deviates from the long-term mean \citep{SPIonline}.\footnote{Can be created for various periods of 1-36 months, usually using monthly data.} The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of evapotranspiration (which captures increased temperature). It is based on water balance, therefore it can be compared to the self-calibrated PDSI \citep{SPEI}.

For an extensive overview of various drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{svoboda2016} or \cite{zargar2011}. \cite{keyantash2002} quantify, evaluate and compare number of drought indices for meteorological, hydrological and agricultural forms of drought. Based on several criteria they conclude that rainfall deciles and SPI perform the best for meteorological drought.
%\cite{keyantash2002} further provides evaluation of the indecies...

\cite{Trenberth2014} discuss various formulations of the PDSI (\citealt{Sheffield2012, Dai2011}).  \cite{Sheffield2012} argue that the way  how PDSI is formulated can have substantial effect on estimation of changes in severity of droughts over time. According to other studies, differences in formulation of the PDSI do not play a big role and it is more important to formulate the index such that the required data are available and reliable (\citealp{Trenberth2014,vanderSchrier2011,Wang2012}). \cite{Trenberth2014} attribute the differences between the results of \cite{Sheffield2012} and \cite{Trenberth2014} to disparities among various rainfall datasets and to different baseline periods they use rather than to different formulations of the PDSI.

%\textcolor{green}{Precist HEIM A Review of Twentieth Century Drought Indices Used in the United States 2002}
\subsubsection*{Extreme events, disaster and hazards}
\cite{IPCC2012ch1} define extreme events as 'the occurrence of a value of a weather or climate variable above (or below) a threshold value near the upper (or lower) ends of the range of observed values of the variable'. Some authors define extreme events only in terms of meteorological phenomena (\citealp{easterling2000, Jentsch2007}), others include also consequential physiological impacts or other effects on humans and ecosystems (\citealp{IPCC2012ch1, young2002}).

According to \cite{IPCC2012ch1}, disasters are defined as a 'severe alternations in the normal functioning of community or a society due to hazardous physical events interacting with vulnerable social conditions, leading to widespread adverse human, material, economic or environmental effects that require immediate emergency response to satisfy critical human needs and that may acquire external support for recovery.' The hazardous physical events may be of natural, socio-natural, or purely anthropogenic origin (\citealp{IPCC2012ch1, wisner2004risk}).

Hazard can be defined as 'the potential occurrence of a natural or human induced
physical event that may cause loss of life, injury, or other
health impacts, as well as damage and loss to property, infrastructure,
livelihoods, service provision, and environmental resources' \citet{IPCC2012ch1}.
 \section*{Effects of Drought on Economy}

 \cite{Pedram2016} estimate national production losses per disaster worldwide during ${1964-2007}$ using superposed epoch analysis. Besides drought, they focus on extreme heat, cold and flood events. They conclude that on average $10.1\%$ reduction of cereal production can be linked to droughts and $9.1\%$ reduction is attributable to extreme heat. They did not find any significant effect of extreme cold and floods on production. \cite{Mehrabi2017} estimate cumulative crop production losses resulting from heat and drought disasters over the same time period (${1964-2007}$). Their estimates are almost half of those of \cite{Pedram2016}. The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA. \cite{HLAVINKA2009} analyse effects of drought on major cereal yields in the Czech Republic. They also define crop specific thresholds below which moisture deficit leads to severe negative impact on yields. \cite{HLAVINKA2009} find a link between severe droughts and reduction of crop yields.

 \cite{edwards2009} estimate economic and financial impact of drought in rural areas of Australia. They use data from a telephone interviews with $8000$ people. \cite{edwards2009} find that drought has significant effects on experience of financial hardship and detoriation of household with negative economic impact.

\cite{shahid2008} use SPI time series to produce drought hazard maps of western part of Bangladesh. The hazard maps are then combined with maps of socio-economic vulnerability to create drought risk maps.

 \cite{Sandstorm2017} question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia. They analyse causes of the food crisis identified in the humanitarian appeal documents. Majority of these documents consider food availability and food production as a major cause of food crisis. \cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.  \cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience. They suggest to increase the budget share focused on agricultural and livestock production to build resilience.

\subsubsection*{CGE}

\cite{robinson2010} use a Social Accounting Matrix (SAM) based CGE model to simulate impacts of food production and price shocks on household consumption in Ethiopia. They model a drought (or a disease) as a $20\%$ reduction in crop productivity and a $20\%$ decrease in livestock capital in lowland plains and drought-prone highlands in Ethiopia and they conclude that this would cause a real income loss on order of $2.3\%$ of Ethiopia's GDP. They also consider two alternative scenarios, i.e. variations in food aid and change in international prices.

\cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios. According to his results, climate change will lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change. However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citep{OxfamIDS}.
 \section*{Economic Effects of Climate and Weather}

Shifts in staple food demand curve don't tend to be very large. Hence, when staple food become scarce, its price is usually subject to a massive increase (\citealp{Brown2015, brown2014book}). For low income groups, this often leads to a reduction in calorie intake, malnutrition and increased risk of related health problems (\citealp{Golden2011, Handa2006}). Local food prices are therefore a good indicator of food scarcity and insecurity (\citealp{baffes2017, Brown2015}). \cite{Brown2015} investigate effects of weather disturbances and international price changes on local food prices which they use as a measure of food scarcity. They use a Kalman Filter approach (see \cite{KoopmanSJ} for more details) and they focus on regions which contain large segments of low income population including locations in Africa, South Asia and Latin America. They conclude that almost $20\%$ of local market prices are affected by domestic weather disturbances, $9\%$ of them are affected by international price change and $4\%$ are affected by both of them. Based on whether or not international food price and weather shocks are significant in explaining local food prices, \cite{brown2014book} groups food markets in selected developing countries into four categories as follows: significantly affected by both international food prices and weather, significantly affected by weather but not by international food prices, significantly affected by international food prices but not by weather and not significantly affected by either of them. \cite{brown2014book} then discus common characteristics of markets in each of these groups.

\cite{Ochieng2016} estimate effects of climate variability and change on agricultural production\footnote{Measured as value of yields per acre in farm household.} using panel data in Kenya. According to their results, the effects are significant yet different for different crops. Temperature has positive effect on tea and negative effect on production of maize and crop. Furthermore, rainfall affects production of tea negatively. On the other hand, \cite{Vrieling2011} finds positive correlation between precipitation and agricultural productivity.

%\textcolor{blue}{look at many references in \cite{Brown2015} and \cite{Ochieng2016} !!!maybe also look if good references in \cite{OxfamIDS}??}

 \cite{nicolai2017} quantify relation between soil moisture drought and temperature, precipitation, evapotranspiration and vegetation during growing season. Using superposed epoch analysis, they find strong link between soil moisture drought and temperature, precipitation and evapotranspiration. Vegetation indices show a delayed response. \cite{schwalm2017} analyse length of post-draught recovery time in terms of gross primary productivity. According to their results the recovery times are strongly associated with climate and carbon cycle dynamics and the recovery is the longest in the tropics and high northern latitudes. They also conclude that drought impacts have increased over the 20th century. Both \cite{nicolai2017} and \cite{schwalm2017} analyse effects of drought globally.

\subsubsection*{Hedonic approach and an alternative strategy}

A large literature has focused on application of the Ricardian approach in the context of climate change (including changes in precipitation) and its effects on crop revenues in Africa (\citealp{RicardianBello,kabubo2007, KMendelsohn2008, SeoMendelsohn}). According to \cite{SeoMendelsohn} and \cite{KMendelsohn2008}, if the warming will be mild and wet, crop net revenue should increase. In case of more severe and dry scenario, the crop revenue is likely to decrease (\citealp{KMendelsohn2008, SeoMendelsohn}).  \cite{RicardianBello} conclude that increase in temperature will affect net crop revenue negatively while increase in precipitation is likely to increase the revenue. According to \citet{kabubo2007}, global warming will be harmful for crop production.

\cite{Deschenes2007Ric} use an alternative strategy to estimate economic impact of climate change, including impact of change in precipitation, on agricultural profit in the United States.\footnote{The strategy of \cite{Deschenes2007Ric} is different from the hedonic approach and the production function approach.} In particular, they use year-to-year variation in precipitation and temperature which is assumed to be statistically independent of unobserved determinants of agricultural profit. \cite{Deschenes2007Ric} argue that this strategy is preferable to the hedonic approach, as hedonic approach is likely to be plagued with omitted variables bias. They conclude that effect of global warming on agriculture should be positive in the US. In particular, agricultural yields should increase by $1.3$ billion US dollars or $4\%$. However, the estimated $95\%$ confidence interval is from $-0.5$ to $3.1$ billion US dollars, hence, the estimated effect is not very large.

\section*{Trends}

 Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}. \cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter. 
 \pagebreak
\bibliographystyle{apa}
\bibliography{referencesFS}
\end{document}

### 2018-11-15 LiteratureFS/MyReview/Review10.tex
blll

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{a4wide}

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{bm}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage{colortbl}
\usepackage{footnote}
\makesavenoteenv{tabular}
\usepackage[flushleft]{threeparttable}
\usepackage[hyphens]{url}
\usepackage{placeins}
\usepackage{dcolumn}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{changepage}

\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{caption}
\usepackage{array}
\usepackage{pdfpages}
\usepackage{float}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{amsxtra}
\usepackage{amsbsy}
\usepackage{accents}
\usepackage{chngcntr}
\usepackage{tabularx}
\usepackage{dcolumn}
\usepackage[none]{hyphenat}
\usepackage[affil-it]{authblk}
\usepackage{datetime}
\usepackage[labelfont=bf]{caption}
\usepackage{titlesec}
\usepackage{endnotes}
\usepackage{csquotes}
\usepackage{epstopdf}

\date{\normalsize{September 2017}}
\title{\Large \bf Climate Change Awareness and Willingness to Pay for its Mitigation: \\ Evidence from the UK}
\author{Monika Novackova and Richard Tol}
\affil{\small{Department of Economics, University of Sussex, Falmer, UK}}

\parindent 0pt
\parskip 0.5em
\newcommand\starred[1]{\accentset{~~~~~\star}{#1}}

\newcounter{magicrownumbers}
\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}

\begin{document}
\sloppy
\section*{Definition of drought}
How we define drought is very important. Based on specification of drought, we can measure changes in aridity over time. According to international meteorological community, drought can be defined as \textit{ 'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{IPCCtrenberth, Heim2002}). The International Panel for Climate Change discuss three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater (IPCC; \citealp{IPCCtrenberth, Heim2002}). \cite{Trenberth2014} discuss definitions and measures of drought and their relation to contradictory results of two recent studies, in particular \cite{Sheffield2012} and \cite{Dai2011}. \cite{Sheffield2012} argue that drought has not increased much since $1960$ althought incorrect versions of Palmer Drought Severity Index (PDSI) give substantially different results. On the other hand, \cite{Dai2011} conclude that results differ only slightly for different forms of PDSI and all its forms indicates widespread drying. Besides difference in way of calculating the drought index, \cite{Trenberth2014} attribute the contradicting results to disparities among various rainfall datasets and different baseline periods.

As discussed by \cite{Trenberth2014}, drought can be measured in absolute terms (e.g. lake levels or amount of soil moisture) or using relative measures, such as PDSI. Because drought is defined based on one tail of probability distribution function of a drought measure, small decrease in mean can appear as very big increase in frequency of droughts. This has caused confusions and therefore usage of percentiles of soil moisture or streamflow is recommended as a better measure than mean \citep{Trenberth2014}.

Specifying a reliable index which could be used as a basis for definition of drought seems to problematic. The degree of drought does not only depend on precipitation, but also on whether and how fast the moisture is carried away (so the index should also incorporate evapotranspiration, which PDSI does. It also accounts for balance of precipitation.) Thus, besides precipitation, the index should incorporate humidity, wind, solar and long-wave radiation data \citep{vanderSchrier2011}.  However, availability of reliable data for solar radiation is a real problem \citep{Wang2012}.

\subsection*{How do authors define drought in terms of distribution of index?}
\textit{quantile..}
\begin{itemize}

\item \cite{Trenberth2014}:
\end{itemize}`

\subsubsection*{Extreme events, disaster and hazards}
\cite{IPCC2012ch1} define extreme events as 'the occurrence of a value of a weather or climate variable above (or below) a threshold value near the upper (or lower) ends of the range of observed values of the variable'. Some authors define extreme events only in terms of meteorological phenomena (\citealp{easterling2000, Jentsch2007}), others include also consequential physiological impacts or other effects on humans and ecosystems (\citealp{IPCC2012ch1, young2002}).

According to \cite{IPCC2012ch1}, disasters are defined as 'severe alternations in the normal functioning of community or a society due to hazardous physical events interacting with vulnerable social conditions, leading to widespread adverse human, material, economic or environmental effects that require immediate emergency response to satisfy critical human needs and that may acquire external support for recovery.' The hazardous physical events may be of natural, socio-natural, or purely anthropogenic origin (\citealp{IPCC2012ch1, wisner2004risk}).

Hazard can be defined as 'the potential occurrence of a natural or human induced
physical event that may cause loss of life, injury, or other
health impacts, as well as damage and loss to property, infrastructure,
livelihoods, service provision, and environmental resources' \citet{IPCC2012ch1}.
\section*{Changes and trends in drought}

 Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1976/1977$ \citep{IPCCtrenberth}. During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia. Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}. \cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.

 \section*{Economic Effects of Droughts}

Demand for staple food is usually fixed, even if its production change. Therefore, when staple food becomes scarce in some location, its price tend to increase rapidly. For lower income families, this leads to reduction in calorie intake and often to malnutrition and increased risk of health problems (\citealp{Golden2011, Handa2006}). Local food prices are therefore a good indicator of food scarcity and insecurity \citep{Brown2015}. \cite{Brown2015} investigate effects of weather disturbances and international price changes on local food prices which serve as a proxy for food scarcity. They use Kalman Filter approach (for details see \cite{KoopmanSJ}) and they focus on regions which contain large segments of low income population including Africa, South Asia and Latin America. They conclude that almost $20\%$ of local market prices are affected by domestic weather disturbances, $9\%$ of them are affected by international price change and $4\%$by both of them. Based on whether or not international food price and weather shocks are significant in explaining local food prices,\cite{brown2014book} groups food markets in developing countries into four categories: significantly affected by both international food prices and weather, significantly affected by weather but not international food prices, significantly affected by international food prices but not by weather and not significantly affected by either of them. \cite{brown2014book} then discus common characteristics of markets in each of these groups.

\cite{Ochieng2016} estimate effects of climate variability and change on agricultural production \footnote{Mesured as value of yields per acre in farm household} using panel data in Kenya. According to their results the effects are significant, yet different for different crops. Temperature has positive effect on tea and negative effect on production of maize and crop. Further, rainfall affects production of tea negatively.

\textcolor{blue}{look at many references in \cite{Brown2015} and \cite{Ochieng2016} !!!maybe also look if good references in \cite{OxfamIDS}??}

\cite{Pedram2016} estimate national production losses per disaster worldwide during ${1964-2007}$ using a statistical method called superposed epoch analysis. Besides drought, they focus on extreme heat, cold and flood events. They conclude that on average $10.1\%$ reduction of cereal production can be linked to droughts and $9.1\%$ reduction is attributable to extreme heat. They did not find any significant effect of extreme cold and floods on production. \cite{Mehrabi2017} estimate cumulative crop production losses resulting from heat and drought disasters over the same time period (${1964-2007}$). Their estimates are almost half of those estimated by \cite{Pedram2016}. The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA.

\cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios. According to his results, climate change will cause $35\%$ decrease in maize productivity and $18\%$ decrease in productivity of paddy raise, wheat and other crops in south and south-east Africa. Climate change will also lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change. However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citet{OxfamIDS}.

\bibliographystyle{apa}
\bibliography{referencesFS}
\end{document}

### 2018-11-15 LiteratureFS/MyReview/Review11.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/Review10.tex
@@
 
-Demand for staple food is usually fixed, even if its production change.
-Therefore, when staple food becomes scarce in some location, its price tend to increase rapidly.
-For lower income families, this leads to reduction in calorie intake and often to malnutrition and increased risk of health problems (\citealp{Golden2011, Handa2006}).
+Shifts in staple food demand curve are usually not very large; hence, when staple food becomes scarce, its price is usually subject to a massive increase (\citealp{Brown2015, brown2014book}).
+For low income groups, this often leads to reduction in calorie intake and often to malnutrition and increased risk of related health problems (\citealp{Golden2011, Handa2006}).
 Local food prices are therefore a good indicator of food scarcity and insecurity \citep{Brown2015}.
 \cite{Brown2015} investigate effects of weather disturbances and international price changes on local food prices which serve as a proxy for food scarcity.
-They use Kalman Filter approach (for details see \cite{KoopmanSJ}) and they focus on regions which contain large segments of low income population including Africa, South Asia and Latin America.
+They use Kalman Filter approach (see \cite{KoopmanSJ} for more details) and they focus on regions which contain large segments of low income population including locations in Africa, South Asia and Latin America.
 They conclude that almost $20\%$ of local market prices are affected by domestic weather disturbances, $9\%$ of them are affected by international price change and $4\%$by both of them.
-Based on whether or not international food price and weather shocks are significant in explaining local food prices,\cite{brown2014book} groups food markets in developing countries into four categories: significantly affected by both international food prices and weather, significantly affected by weather but not international food prices, significantly affected by international food prices but not by weather and not significantly affected by either of them.
+Based on whether or not international food price and weather shocks are significant in explaining local food prices, \cite{brown2014book} groups food markets in selected developing countries into four categories as follows: significantly affected by both international food prices and weather, significantly affected by weather but not international food prices, significantly affected by international food prices but not by weather and not significantly affected by either of them.
 \cite{brown2014book} then discus common characteristics of markets in each of these groups.
@@
 \cite{Mehrabi2017} estimate cumulative crop production losses resulting from heat and drought disasters over the same time period (${1964-2007}$).
-Their estimates are almost half of those estimated by \cite{Pedram2016}.
+Their estimates are almost half of those of \cite{Pedram2016}.
 The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA.
@@
 \cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios.
-According to his results, climate change will cause $35\%$ decrease in maize productivity and $18\%$ decrease in productivity of paddy raise, wheat and other crops in south and south-east Africa.
-Climate change will also lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change.
-However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citet{OxfamIDS}.
+According to his results, climate change will lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change.
+However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citep{OxfamIDS}.
 

### 2018-11-15 LiteratureFS/MyReview/Review12.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/Review11.tex
@@
 \section*{Definition of drought}
-How we define drought is very important.
-Based on specification of drought, we can measure changes in aridity over time.
  According to international meteorological community, drought can be defined as \textit{ 'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{IPCCtrenberth, Heim2002}).
-The International Panel for Climate Change discuss three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater (IPCC; \citealp{IPCCtrenberth, Heim2002}).
-\cite{Trenberth2014} discuss definitions and measures of drought and their relation to contradictory results of two recent studies, in particular \cite{Sheffield2012} and \cite{Dai2011}.
-\cite{Sheffield2012} argue that drought has not increased much since $1960$ althought incorrect versions of Palmer Drought Severity Index (PDSI) give substantially different results.
-On the other hand, \cite{Dai2011} conclude that results differ only slightly for different forms of PDSI and all its forms indicates widespread drying.
-Besides difference in way of calculating the drought index, \cite{Trenberth2014} attribute the contradicting results to disparities among various rainfall datasets and different baseline periods.
+The International Panel for Climate Change recognise three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater (IPCC; \citealp{IPCCtrenberth, Heim2002}).
 
-As discussed by \cite{Trenberth2014}, drought can be measured in absolute terms (e.g.
-lake levels or amount of soil moisture) or using relative measures, such as PDSI.
-Because drought is defined based on one tail of probability distribution function of a drought measure, small decrease in mean can appear as very big increase in frequency of droughts.
+ \cite{Trenberth2014} discuss different ways of formulation of the Palmer Drought Severity Index (\citealt{Sheffield2012, Dai2011}).
+\cite{Sheffield2012} argue that way of formulation of PDSI can have substantial effect on estimation of changes in severity of droughts over time.
+According to \cite{Trenberth2014}, differences in formulation of the PDSI do not play a big role and it is more important to formulate the index such that the required data are available and reliable.
+\cite{Trenberth2014} attribute the differences in results of  \cite{Sheffield2012} and \cite{Trenberth2014} mostly to disparities among various rainfall datasets and different baseline periods rather than different formulations of the PDSI.
+
+Drought can be measured in absolute terms (e.g.
+lake levels or amount of soil moisture) or in relative measures, which is for example the  PDSI \citep{Trenberth2014}.
+Because drought is defined based on one tail of probability distribution function of a drought measure, small decrease in mean can appear as substantial increase in frequency of droughts.
 This has caused confusions and therefore usage of percentiles of soil moisture or streamflow is recommended as a better measure than mean \citep{Trenberth2014}.
@@
 Specifying a reliable index which could be used as a basis for definition of drought seems to problematic.
-The degree of drought does not only depend on precipitation, but also on whether and how fast the moisture is carried away (so the index should also incorporate evapotranspiration, which PDSI does.
-It also accounts for balance of precipitation.) Thus, besides precipitation, the index should incorporate humidity, wind, solar and long-wave radiation data \citep{vanderSchrier2011}.
-However, availability of reliable data for solar radiation is a real problem \citep{Wang2012}.
-
-\subsection*{How do authors define drought in terms of distribution of index?}
-\textit{quantile..}
-\begin{itemize}
-
-\item \cite{Trenberth2014}:
-\end{itemize}`
+The degree of drought does not only depend on precipitation, but also on whether and how fast the moisture is carried away.
+Therefore, the index should also incorporate evapotranspiration, which PDSI does.
+Hence, \cite{vanderSchrier2011} argue that besides precipitation, the index should incorporate humidity, wind, solar and long-wave radiation data.
+However, availability of reliable data for e.g.
+solar radiation is a real problem, hence, it may be preferable to use a more simplistic index (perhaps not including evapotranspiration) for which reliable data are available \citep{Wang2012}.
 
@@
 livelihoods, service provision, and environmental resources' \citet{IPCC2012ch1}.
-\section*{Changes and trends in drought}
-
- Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1976/1977$ \citep{IPCCtrenberth}.
-During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
-Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.
-\cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.
 
@@
 
-Shifts in staple food demand curve are usually not very large; hence, when staple food becomes scarce, its price is usually subject to a massive increase (\citealp{Brown2015, brown2014book}).
-For low income groups, this often leads to reduction in calorie intake and often to malnutrition and increased risk of related health problems (\citealp{Golden2011, Handa2006}).
-Local food prices are therefore a good indicator of food scarcity and insecurity \citep{Brown2015}.
+Shifts in staple food demand curve are usually not very large.
+Hence, when staple food becomes scarce, its price is usually subject to a massive increase (\citealp{Brown2015, brown2014book}).
+For low income groups, this often leads to reduction in calorie intake, malnutrition and increased risk of related health problems (\citealp{Golden2011, Handa2006}).
+Local food prices are therefore a good indicator of food scarcity and insecurity (\citealp{baffes2017, Brown2015}).
 \cite{Brown2015} investigate effects of weather disturbances and international price changes on local food prices which serve as a proxy for food scarcity.
@@
 Further, rainfall affects production of tea negatively.
+Another study, which finds positive correlation between precipitation and agricultural productivity is \cite{Vrieling2011}.
 
@@
 
+\section*{Trends}
+
+ Strong downwards trend in precipitation has been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1976/1977$ \citep{IPCCtrenberth}.
+During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
+Furthermore, increased frequency of heavy rain events has been observed also in the areas with decline in total rainfall \citep{IPCCtrenberth}.
+\cite{Trenberth2014} argue that as a consequence of global warming, dry areas have strong tendency to get drier while wet areas are getting wetter.
+
 \bibliographystyle{apa}

### 2018-11-15 LiteratureFS/MyReview/Review13.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/Review12.tex
@@
 \cite{Sheffield2012} argue that way of formulation of PDSI can have substantial effect on estimation of changes in severity of droughts over time.
-According to \cite{Trenberth2014}, differences in formulation of the PDSI do not play a big role and it is more important to formulate the index such that the required data are available and reliable.
+According to other studies, differences in formulation of the PDSI do not play a big role and it is more important to formulate the index such that the required data are available and reliable (\citealp{Trenberth2014,vanderSchrier2011,Wang2012}).
 \cite{Trenberth2014} attribute the differences in results of  \cite{Sheffield2012} and \cite{Trenberth2014} mostly to disparities among various rainfall datasets and different baseline periods rather than different formulations of the PDSI.
@@
 
-Specifying a reliable index which could be used as a basis for definition of drought seems to problematic.
-The degree of drought does not only depend on precipitation, but also on whether and how fast the moisture is carried away.
-Therefore, the index should also incorporate evapotranspiration, which PDSI does.
-Hence, \cite{vanderSchrier2011} argue that besides precipitation, the index should incorporate humidity, wind, solar and long-wave radiation data.
-However, availability of reliable data for e.g.
-solar radiation is a real problem, hence, it may be preferable to use a more simplistic index (perhaps not including evapotranspiration) for which reliable data are available \citep{Wang2012}.
-
+%\textcolor{green}{Precist HEIM A Review of Twentieth Century Drought Indices Used in the United States 2002}
 \subsubsection*{Extreme events, disaster and hazards}
@@
 
-\textcolor{blue}{look at many references in \cite{Brown2015} and \cite{Ochieng2016} !!!maybe also look if good references in \cite{OxfamIDS}??}
+%\textcolor{blue}{look at many references in \cite{Brown2015} and \cite{Ochieng2016} !!!maybe also look if good references in \cite{OxfamIDS}??}
 
@@
 
+Large literature uses Ricardian approach to estimate impacts of climate change, including change in precipitation on crop revenue (\citealp{RicardianBello,kabubo2007, KMendelsohn2008, SeoMendelsohn}.
+
 \section*{Trends}

### 2018-11-15 LiteratureFS/MyReview/Review14.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/Review13.tex
@@
 
+Large literature apply Ricardian approach (which was originally developed by \citealp{Mendelsohn94}) to estimate impacts of climate change, including change in precipitation on crop revenues in Africa (\citealp{RicardianBello,kabubo2007, KMendelsohn2008, SeoMendelsohn}).
+According to \cite{SeoMendelsohn} and \cite{KMendelsohn2008}, if the warming will be mild and wet crop net revenue should increase.
+In case of more severe and dry scenario, the crop revenue is likely to decrease (\citealp{KMendelsohn2008, SeoMendelsohn}).
+\cite{RicardianBello} conclude that increase in temperature will affect net crop revenue negatively while increase in precipitation is likely to increase crop revenue.
+\citet{kabubo2007} argues that global warming will be harmful for crop production.
+
+\cite{Deschenes2007Ric} use a production function approach to estimate economic impact of climate change, including impact of change in precipitation on agricultural profit in the United States.
+Production function method is an alternative to the Ricardian approach.
+\cite{Deschenes2007Ric} conclude that effect of global warming on agriculture should be positive in the US.
+In particular, agricultural yields should increase by $1.3$ billion US dollars or $4\%$.
+However, the estimated $95\%$ confidence interval is from $-0.5$ to $3.1$ billion US dollars, hence, the estimated effect is not very large.
+
 \cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios.
@@
 However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citep{OxfamIDS}.
-
-Large literature uses Ricardian approach to estimate impacts of climate change, including change in precipitation on crop revenue (\citealp{RicardianBello,kabubo2007, KMendelsohn2008, SeoMendelsohn}.
 

### 2018-11-15 LiteratureFS/MyReview/Review17.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/LitReviewDrought.tex
@@
 
-Numerous definitions of drought and their role are reviewed and discussed by \cite{wilhite1985} and \cite{wilhite2000}.
-They distinguish two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
-Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
-Operational definitions are essential for an effective early warning system.
-An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.
-
 \subsubsection*{Drought indices}
+\textcolor{blue}{hello pridat referenci  sombreok et al 1982 viz \cite{kabubo2007}}
 A numerical measure is needed to compare severity of drought across different time periods or geographical locations.
 However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
-Instead of that a number of measures of drought has been developed (\citealp{Heim2002, wilhite1985, wilhite2000}).
+Instead of that a number of measures of drought has been developed \citep{Heim2002}.
+For an extensive overview of various drought indices see for example  \cite{Heim2002} or \cite{keyantash2002}.
 
@@
 The study of \cite{palmer1965} was a significant milestone in the history of quantification of drought severity.
-\cite{palmer1965} developed the Palmer Drought Severity Index (PDSI) using a complex water balance model.
-The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand (\citealp{Heim2002,palmer1965}).
+\cite{palmer1965} has developed the Palmer Drought Severity Index (PDSI) using a complex water balance model.
+The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand \citep{Heim2002}.
 As the PDSI suffers from several weaknesses (for details see e.g.
@@
 These include the standardized precipitation index (SPI) developed by \cite{SPI} and the standardized precipitation evapotranspiration index (SPEI) developed by \cite{SPEI}.
-The SPI specifies observed precipitation as a standardised departure from a chosen probability distribution which models the precipitation data.
-Values of SPI can be viewed as a multiple of standard deviations by which the observed amount of rainfall deviates from the long-term mean \citep{SPIonline}.\footnote{Can be created for various periods of 1-36 months, usually using monthly data.} The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of evapotranspiration (which captures increased temperature).
+The SPI is based on cumulative probability of a given rainfall situation on a particular station.
+The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of temperature.
 It is based on water balance, therefore it can be compared to the self-calibrated PDSI \citep{SPEI}.
 
-For an extensive overview of various drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{svoboda2016} or \cite{zargar2011}.
-\cite{keyantash2002} quantify, evaluate and compare number of drought indices for meteorological, hydrological and agricultural forms of drought.
-Based on several criteria they conclude that rainfall deciles and SPI perform the best for meteorological drought.
 %\cite{keyantash2002} further provides evaluation of the indecies...
+
+Drought can be measured in absolute terms (e.g.
+lake levels or amount of soil moisture) or in relative measures, which is for example the  PDSI \citep{Trenberth2014}.
+Because drought is defined based on one tail of probability distribution function of a drought measure, small decrease in mean can appear as a substantial increase in frequency of droughts.
+This has caused confusions and therefore usage of percentiles of soil moisture or streamflow is recommended as a better measure than mean \citep{Trenberth2014}.
 
@@
 \cite{Sheffield2012} argue that the way  how PDSI is formulated can have substantial effect on estimation of changes in severity of droughts over time.
-According to other studies, differences in formulation of the PDSI do not play a big role and it is more important to formulate the index such that the required data are available and reliable (\citealp{Trenberth2014,vanderSchrier2011,Wang2012}).
+According to other studies, differences in formulation of the PDSI do not play big role and it is more important to formulate the index such that the required data are available and reliable (\citealp{Trenberth2014,vanderSchrier2011,Wang2012}).
 \cite{Trenberth2014} attribute the differences between the results of \cite{Sheffield2012} and \cite{Trenberth2014} to disparities among various rainfall datasets and to different baseline periods they use rather than to different formulations of the PDSI.
@@
 livelihoods, service provision, and environmental resources' \citet{IPCC2012ch1}.
- \section*{Effects of Drought on Economy}
 
- \cite{Pedram2016} estimate national production losses per disaster worldwide during ${1964-2007}$ using superposed epoch analysis.
-Besides drought, they focus on extreme heat, cold and flood events.
-They conclude that on average $10.1\%$ reduction of cereal production can be linked to droughts and $9.1\%$ reduction is attributable to extreme heat.
-They did not find any significant effect of extreme cold and floods on production.
-\cite{Mehrabi2017} estimate cumulative crop production losses resulting from heat and drought disasters over the same time period (${1964-2007}$).
-Their estimates are almost half of those of \cite{Pedram2016}.
-The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA.
-\cite{HLAVINKA2009} analyse effects of drought on major cereal yields in the Czech Republic.
-They also define crop specific thresholds below which moisture deficit leads to severe negative impact on yields.
-\cite{HLAVINKA2009} find a link between severe droughts and reduction of crop yields.
-
- \cite{edwards2009} estimate economic and financial impact of drought in rural areas of Australia.
-They use data from a telephone interviews with $8000$ people.
-\cite{edwards2009} find that drought has significant effects on experience of financial hardship and detoriation of household with negative economic impact.
-
-\cite{shahid2008} use SPI time series to produce drought hazard maps of western part of Bangladesh.
-The hazard maps are then combined with maps of socio-economic vulnerability to create drought risk maps.
-
- \cite{Sandstorm2017} question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
-They analyse causes of the food crisis identified in the humanitarian appeal documents.
-Majority of these documents consider food availability and food production as a major cause of food crisis.
-\cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.
-\cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience.
-They suggest to increase the budget share focused on agricultural and livestock production to build resilience.
-
-\subsubsection*{CGE}
-
-\cite{robinson2010} use a Social Accounting Matrix (SAM) based CGE model to simulate impacts of food production and price shocks on household consumption in Ethiopia.
-They model a drought (or a disease) as a $20\%$ reduction in crop productivity and a $20\%$ decrease in livestock capital in lowland plains and drought-prone highlands in Ethiopia and they conclude that this would cause a real income loss on order of $2.3\%$ of Ethiopia's GDP.
-They also consider two alternative scenarios, i.e.
-variations in food aid and change in international prices.
-
-\cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios.
-According to his results, climate change will lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change.
-However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citep{OxfamIDS}.
- \section*{Economic Effects of Climate and Weather}
+ \section*{Economic Effects of Droughts}
 
@@
 
-\cite{Ochieng2016} estimate effects of climate variability and change on agricultural production\footnote{Measured as value of yields per acre in farm household.} using panel data in Kenya.
+\cite{Ochieng2016} estimate effects of climate variability and change on agricultural production\footnote{Mesured as value of yields per acre in farm household} using panel data in Kenya.
 According to their results, the effects are significant yet different for different crops.
@@
 %\textcolor{blue}{look at many references in \cite{Brown2015} and \cite{Ochieng2016} !!!maybe also look if good references in \cite{OxfamIDS}??}
+
+\cite{Pedram2016} estimate national production losses per disaster worldwide during ${1964-2007}$ using superposed epoch analysis.
+Besides drought, they focus on extreme heat, cold and flood events.
+They conclude that on average $10.1\%$ reduction of cereal production can be linked to droughts and $9.1\%$ reduction is attributable to extreme heat.
+They did not find any significant effect of extreme cold and floods on production.
+\cite{Mehrabi2017} estimate cumulative crop production losses resulting from heat and drought disasters over the same time period (${1964-2007}$).
+Their estimates are almost half of those of \cite{Pedram2016}.
+The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA.
 
@@
 
-\subsubsection*{Hedonic approach and an alternative strategy}
+ \cite{Sandstorm2017} question whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
+They analyse causes of the food crisis identified in the humanitarian appeal documents.
+Majority of these documents consider food availability and food production as a major cause of food crisis.
+\cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.
+\cite{Sandstorm2017} also find that a large share of humanitarian response budget is focused on emergency food aid in comparison to the share focused on interventions to build resilience.
+They suggest to increase the budget share focused on agricultural and livestock production to build resilience.
+
+\subsubsection*{Ricardian and production function approach}
 
@@
 
-\cite{Deschenes2007Ric} use an alternative strategy to estimate economic impact of climate change, including impact of change in precipitation, on agricultural profit in the United States.\footnote{The strategy of \cite{Deschenes2007Ric} is different from the hedonic approach and the production function approach.} In particular, they use year-to-year variation in precipitation and temperature which is assumed to be statistically independent of unobserved determinants of agricultural profit.
-\cite{Deschenes2007Ric} argue that this strategy is preferable to the hedonic approach, as hedonic approach is likely to be plagued with omitted variables bias.
-They conclude that effect of global warming on agriculture should be positive in the US.
+\cite{Deschenes2007Ric} use a production function approach (as an alternative to the Ricardian approach) to estimate economic impact of climate change, including impact of change in precipitation on agricultural profit in the United States.
+\cite{Deschenes2007Ric} conclude that effect of global warming on agriculture should be positive in the US.
 In particular, agricultural yields should increase by $1.3$ billion US dollars or $4\%$.
 However, the estimated $95\%$ confidence interval is from $-0.5$ to $3.1$ billion US dollars, hence, the estimated effect is not very large.
+
+\subsubsection*{CGE}
+
+\cite{robinson2010} use a Social Accounting Matrix (SAM) based CGE model to simulate impacts of food production and price shocks on household consumption in Ethiopia.
+They model a drought (or a disease) as a $20\%$ reduction in crop productivity and a $20\%$ decrease in livestock capital in lowland plains and drough-prone highlands in Ethiopia and they conclude that this would cause a real income loss on order of $2.3\%$ of Ethiopia's GDP.
+They also consider two alternative scenarios, i.e.
+variations in food aid and change in international prices.
+
+\cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios.
+According to his results, climate change will lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change.
+However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citep{OxfamIDS}.
 

### 2018-11-15 LiteratureFS/MyReview/Review19.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/LitReviewDrought.tex
@@
 \subsubsection*{Drought indices}
+\textcolor{blue}{hello pridat referenci  sombreok et al 1982 viz \cite{kabubo2007}}
 A numerical measure is needed to compare severity of drought across different time periods or geographical locations.
@@
 
+Drought can be measured in absolute terms (e.g.
+lake levels or amount of soil moisture) or in relative measures, which is for example the  PDSI \citep{Trenberth2014}.
+Because drought is defined based on one tail of probability distribution function of a drought measure, small decrease in mean can appear as a substantial increase in frequency of droughts.
+This has caused confusions and therefore usage of percentiles of soil moisture or streamflow is recommended as a better measure than mean \citep{Trenberth2014}.
+
 \cite{Trenberth2014} discuss various formulations of the PDSI (\citealt{Sheffield2012, Dai2011}).
 \cite{Sheffield2012} argue that the way  how PDSI is formulated can have substantial effect on estimation of changes in severity of droughts over time.
-According to other studies, differences in formulation of the PDSI do not play a big role and it is more important to formulate the index such that the required data are available and reliable (\citealp{Trenberth2014,vanderSchrier2011,Wang2012}).
+According to other studies, differences in formulation of the PDSI do not play big role and it is more important to formulate the index such that the required data are available and reliable (\citealp{Trenberth2014,vanderSchrier2011,Wang2012}).
 \cite{Trenberth2014} attribute the differences between the results of \cite{Sheffield2012} and \cite{Trenberth2014} to disparities among various rainfall datasets and to different baseline periods they use rather than to different formulations of the PDSI.
@@
 The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA.
-\cite{HLAVINKA2009} analyse effects of drought on major cereal yields in the Czech Republic.
-They also define crop specific thresholds below which moisture deficit leads to severe negative impact on yields.
-\cite{HLAVINKA2009} find a link between severe droughts and reduction of crop yields.
-
- \cite{edwards2009} estimate economic and financial impact of drought in rural areas of Australia.
-They use data from a telephone interviews with $8000$ people.
-\cite{edwards2009} find that drought has significant effects on experience of financial hardship and detoriation of household with negative economic impact.
-
-\cite{shahid2008} use SPI time series to produce drought hazard maps of western part of Bangladesh.
-The hazard maps are then combined with maps of socio-economic vulnerability to create drought risk maps.
 
@@
 They suggest to increase the budget share focused on agricultural and livestock production to build resilience.
+
+ \textcolor{red}{TO READ 12.12.2017:}
+\begin{itemize}
+  \item   \textcolor{red}{  \cite{bailey2012} maybe}
+  \item   \textcolor{red}{\cite{AMS1997} maybe}
+ \end{itemize}
 
@@
 
-\subsubsection*{Hedonic approach and an alternative strategy}
+\subsubsection*{Ricardian and production function approach}
 
@@
 
-\cite{Deschenes2007Ric} use an alternative strategy to estimate economic impact of climate change, including impact of change in precipitation, on agricultural profit in the United States.\footnote{The strategy of \cite{Deschenes2007Ric} is different from the hedonic approach and the production function approach.} In particular, they use year-to-year variation in precipitation and temperature which is assumed to be statistically independent of unobserved determinants of agricultural profit.
-\cite{Deschenes2007Ric} argue that this strategy is preferable to the hedonic approach, as hedonic approach is likely to be plagued with omitted variables bias.
-They conclude that effect of global warming on agriculture should be positive in the US.
+\cite{Deschenes2007Ric} use a production function approach (as an alternative to the Ricardian approach) to estimate economic impact of climate change, including impact of change in precipitation on agricultural profit in the United States.
+\cite{Deschenes2007Ric} conclude that effect of global warming on agriculture should be positive in the US.
 In particular, agricultural yields should increase by $1.3$ billion US dollars or $4\%$.

### 2018-11-15 LiteratureFS/MyReview/Review9.tex
blll

--- 2018-11-15 LiteratureFS/MyReview/Review11.tex
@@
 \sloppy
-\section*{Definition of drought}
+\section*{What is drought?}
 How we define drought is very important.
@@
 \subsection*{How do authors define drought in terms of distribution of index?}
-\textit{quantile..}
 \begin{itemize}
@@
 
-\subsubsection*{Extreme events, disaster and hazards}
-\cite{IPCC2012ch1} define extreme events as 'the occurrence of a value of a weather or climate variable above (or below) a threshold value near the upper (or lower) ends of the range of observed values of the variable'.
-Some authors define extreme events only in terms of meteorological phenomena (\citealp{easterling2000, Jentsch2007}), others include also consequential physiological impacts or other effects on humans and ecosystems (\citealp{IPCC2012ch1, young2002}).
-
-According to \cite{IPCC2012ch1}, disasters are defined as 'severe alternations in the normal functioning of community or a society due to hazardous physical events interacting with vulnerable social conditions, leading to widespread adverse human, material, economic or environmental effects that require immediate emergency response to satisfy critical human needs and that may acquire external support for recovery.' The hazardous physical events may be of natural, socio-natural, or purely anthropogenic origin (\citealp{IPCC2012ch1, wisner2004risk}).
-
-Hazard can be defined as 'the potential occurrence of a natural or human induced
-physical event that may cause loss of life, injury, or other
-health impacts, as well as damage and loss to property, infrastructure,
-livelihoods, service provision, and environmental resources' \citet{IPCC2012ch1}.
-\section*{Changes and trends in drought}
+\section*{trends..}
 
@@
 
-Shifts in staple food demand curve are usually not very large; hence, when staple food becomes scarce, its price is usually subject to a massive increase (\citealp{Brown2015, brown2014book}).
-For low income groups, this often leads to reduction in calorie intake and often to malnutrition and increased risk of related health problems (\citealp{Golden2011, Handa2006}).
-Local food prices are therefore a good indicator of food scarcity and insecurity \citep{Brown2015}.
-\cite{Brown2015} investigate effects of weather disturbances and international price changes on local food prices which serve as a proxy for food scarcity.
-They use Kalman Filter approach (see \cite{KoopmanSJ} for more details) and they focus on regions which contain large segments of low income population including locations in Africa, South Asia and Latin America.
-They conclude that almost $20\%$ of local market prices are affected by domestic weather disturbances, $9\%$ of them are affected by international price change and $4\%$by both of them.
-Based on whether or not international food price and weather shocks are significant in explaining local food prices, \cite{brown2014book} groups food markets in selected developing countries into four categories as follows: significantly affected by both international food prices and weather, significantly affected by weather but not international food prices, significantly affected by international food prices but not by weather and not significantly affected by either of them.
-\cite{brown2014book} then discus common characteristics of markets in each of these groups.
-
-\cite{Ochieng2016} estimate effects of climate variability and change on agricultural production \footnote{Mesured as value of yields per acre in farm household} using panel data in Kenya.
-According to their results the effects are significant, yet different for different crops.
-Temperature has positive effect on tea and negative effect on production of maize and crop.
-Further, rainfall affects production of tea negatively.
-
-\textcolor{blue}{look at many references in \cite{Brown2015} and \cite{Ochieng2016} !!!maybe also look if good references in \cite{OxfamIDS}??}
-
-\cite{Pedram2016} estimate national production losses per disaster worldwide during ${1964-2007}$ using a statistical method called superposed epoch analysis.
+\cite{Pedram2016} estimate national production losses per disaster worldwide during $1964-2007$.
 Besides drought, they focus on extreme heat, cold and flood events.
@@
 They did not find any significant effect of extreme cold and floods on production.
-\cite{Mehrabi2017} estimate cumulative crop production losses resulting from heat and drought disasters over the same time period (${1964-2007}$).
-Their estimates are almost half of those of \cite{Pedram2016}.
-The biggest losses are in Botswana, Paraguay, Nigeria, Angola and USA.
-
-\cite{OxfamIDS} uses the GLOBE Computable General Equilibrium model of the global economy to estimate food prices for various $2030$ scenarios.
-According to his results, climate change will lead to substantial increase in both domestic and world market crop prices in comparison to baseline scenario in the absence of climate change.
-However, the increase in prices can be substantially mitigated if appropriate adaptation measures will be taken in sub-Saharan Africa \citep{OxfamIDS}.
-
 \bibliographystyle{apa}

### 2018-11-15 writing/draft1/draftMonika7.tex
kasflawbkueyfhli WORKING

--- 2018-11-14 writing/draft1/draftMonika5.tex
@@
 
+{
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11a
+\label{KenARe11} 
+\begin{footnotesize}
+
+\begin{tabular}{llrlllr}
+
+\hline \vspace{-0.2cm} \\
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{All counties}} &\multicolumn{2}{c}{\textit{ASAL}} &\multicolumn{2}{c}{\textit{non-ASAL}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&$exp(\beta)$&F-value%> anova(KEN11a, type='marginal')
+    &$exp(\beta)$&F-value&$exp(\beta)$&F-value\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$1.296$ $(1.060)^{***}$&$19.916$&$1.276$&$5.230$&$1.410^{}$&$10.061$\\
+  \\
+\vspace{-0.2cm}Prec. total&$1.081$ $(1.034)^{*}$&$5.402$&$1.006^{**}$&$0.022$&$1.278^{**}$&$19.386$\\
+  \\
+  \vspace{-0.2cm}Prec. total sq.&$0.973$ $(1.013)^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{\bullet}$&$23.747$\\
+    \\
+      \vspace{-0.2cm}Prec. c. of var.&$0.924$ $(1.044)^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
+  \\
+
+         \vspace{-0.2cm}Dry spell -length&$0.935$ $(1.031)^{*}$&$4.810$&$0.833$ ${}$&$6.969$&$ 0.988^{}$&$0.163$\\
+  \\
+        \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939$ $(1.029)^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
+  \\
+          \vspace{-0.2cm}Temp. - average&$0.819$ $(1.059)^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
+  \\
+
+          \vspace{-0.2cm}Temp. std. dev.&$1.043$ $(1.024)^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
+  \\
+
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+ \\ 
+ \hline
+\vspace{-0.2cm} \\
+\textit{Number of observations:}  &$1300$&$698$& \\
+\vspace{-0.2cm}
+\\
+
+  \hline
+  \vspace{-0.2cm} \\
+
+\hline
+\vspace{-0.2cm}
+\end{tabular} 
+\end{footnotesize}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
+        \begin{adjustwidth}{1cm}{} 
+    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
+
 {\hspace{-1.5cm}

### 2018-11-15 writing/draft1/draftMonika8.tex
kasflawbkueyfhli WORKING

--- 2018-11-14 writing/draft1/draftMonika5.tex
@@
 
+{
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11a  KEN11a_ASAL   
+\label{KenARe11} 
+\begin{footnotesize}
+
+\begin{tabular}{llrlllr}
+
+\hline \vspace{-0.2cm} \\
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{All counties}} &\multicolumn{2}{c}{\textit{ASAL}} &\multicolumn{2}{c}{\textit{non-ASAL}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&$exp(\beta)$&F-value\tnote{a}%> anova(KEN11a, type='marginal')
+    &$exp(\beta)$&F-value\tnote{a}&$exp(\beta)$&F-value\tnote{a}\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
+  \\
+\vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
+  \\
+  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
+    \\
+      \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
+  \\
+
+         \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
+  \\
+        \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
+  \\
+          \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
+  \\
+
+          \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
+  \\
+
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+ \\ 
+ \hline
+\vspace{-0.2cm} \\
+\textit{Number of observations:}  &$1300$&$698$&$602$ \\
+\vspace{-0.2cm}
+\\
+
+  \hline
+  \vspace{-0.2cm} \\
+
+\hline
+\vspace{-0.2cm}
+\end{tabular} 
+\end{footnotesize}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
+        \begin{adjustwidth}{1cm}{} 
+    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$\\
+    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+\pagebreak
+
 {\hspace{-1.5cm}

### 2018-11-15 Rcodes/November/KenARe11.R

@@ -53,3 +53,12 @@ nobs(KEN11a_ASAL); anova(KEN11a_ASAL)
 exp(summary(KEN11a_ASAL)$tTable[,1])
 exp(summary(KEN11a_ASAL)$tTable[,2])
 anova(KEN11a_ASAL, type='marginal')
+
+
+KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                 +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                 data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL)
+nobs(KEN11a_nonASAL); anova(KEN11a_nonASAL)
+exp(summary(KEN11a_nonASAL)$tTable[,1])
+exp(summary(KEN11a_nonASAL)$tTable[,2])
+anova(KEN11a_nonASAL, type='marginal')

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation1.tex
adding files, initial

%\documentclass[notes]{beamer}       % print frame + notes
%\documentclass[notes=only]{beamer}   % only notes
\documentclass{beamer}              % only frames
  \usepackage[english]{babel}
\usepackage[round]{natbib}
%\documentclass[hyperref={pdfpagelabels=false}]{beamer}
\usepackage{lmodern}
\usepackage{marvosym} % \MVRIGHTarrow

\usepackage{amsmath}  
\usepackage{multirow,graphicx,array}
\usepackage{multicol}
\usepackage{hhline}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{threeparttable,booktabs}
\usepackage{chngcntr}
\usepackage{dcolumn}
\usepackage{caption}
\usepackage{fixltx2e}
\usepackage{rotating}
\usepackage{amssymb}

\usetheme{Singapore}
\useoutertheme{miniframes}
\AtBeginSection[]{\subsection{}}

\newcommand{\RowColor}{\rowcolor{gray} \cellcolor{white}}
\newcommand{\RowColorY}{\rowcolor{yellow} \cellcolor{white}}

\newcommand{\RN}[1]{%
  \textup{\uppercase\expandafter{\romannumeral#1}}%
}

\newcommand\Wider[2][3em]{%
\makebox[\linewidth][c]{%
  \begin{minipage}{\dimexpr\textwidth+#1\relax}
  \raggedright#2
  \end{minipage}%
  }%
}

\title{Drought and its Impacts on Economy in East Africa}  
\author{Monika Novackova} 
\date{December 2017} 
\institute{University of Sussex}

\setbeamercolor{block title}{bg=purple!30,fg=black}

\begin{document}

\definecolor{LightCyan}{rgb}{0.88,1,1}
\definecolor{BeigeNadine}{rgb}{1,0.94,0.86}
\definecolor{DarkGreen}{rgb}{0.1803922,0.545098,0.3411765}
\definecolor{BeigeNadine}{rgb}{1,0.94,0.86}
\definecolor{PaleGreen}{rgb}{ 0.5960784,0.9843137, 0.5960784}
\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
\definecolor{bored}{rgb}{0.8, 0.0, 0.0}
\definecolor{babypink}{rgb}{0.96, 0.76, 0.76}

 \newcolumntype{d}{D{.}{.}{-1}}
 \newcolumntype{e}{D{+}{\,\pm\,}{6,2}}

\begin{frame}
\titlepage
\end{frame} `

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\normalsize

\section{Defining drought}

\begin{frame}\label{Definition of Drought}
\frametitle{Definition of Drought} 
 \begin{itemize}

\item Prolonged absence or marked deficiency of precipitation  
\item Deficiency of precipitation that results in water shortage for some activity
or for some group
\item Period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance 
\end{itemize}

(\citealp{heim2002,IPCCtrenberth})
\end{frame}

\note{Heim = a review of the 20tieth century indices, Trenberth = IPCC report, summary for policy makers \\
\vspace{0.3cm}f  \\
\vspace{0.3cm} f\\
\vspace{0.3cm}

}

\begin{frame}\label{Definition of Drought}
\frametitle{Categories of Definition of Drought} 
 \begin{itemize}

\item \textbf{Conceptual definitions:} dictionary types, usually defining boundaries of the concept of drought \\ 
e.g. \textit{An extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region \citep{schneider1996}}
\item \textbf{Operational definitions:} Foundation for an effective early warning system \\ e.g. SPI, PDSI

\end{itemize}

(\citealp{wilhite1985,wilhite2000})
\end{frame}

\note{An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000} besides PDSI SPI\\
\vspace{0.3cm}f  \\
\vspace{0.3cm} f\\
\vspace{0.3cm}

}

\begin{frame}
\frametitle{Types of Drought}
\textit{operational drought} \\
  \begin{enumerate}
      \item \textbf{Agricultural drought:} moister deficits in upper layer of
soil up to about one meter depth
      \item \textbf{Meteorological drought:} which refers to prolonged
deficit of precipitation
      \item \textbf{Hydrological drought:} relates to low stream flow, lake and levels of groundwater
      \item \textbf{Socioeconomic drought:} associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought
  \end{enumerate}
  (\citealp{heim2002,IPCCtrenberth,AMS2013})
\end{frame}

\note{first three types in AMS 1997, then later the fourth added? \\
\vspace{0.3cm}first three types physical, the last monetary..(KEYANTASH and dracup 2000, evaluation of drough indecies) \\
\vspace{0.3cm} f\\
\vspace{0.3cm}

}

\section{Drought indices}

\begin{frame}\label{Indices1}
\frametitle{Drought Indices} 
\begin{block}{Early measures of drought}
\begin{itemize}
\item Length of period without $24$-h precipitation of $1.27$mm \citep{munger1916}

\item Length of drought in days, end of drought defined as $2.54$mm of precipitation in $48$ hours \citep{blumenstock1942}
\item Measure of precipitation over a given time period \citep{wilhite1985}
\item Antecedent Precipitation Index (API) based on amount and timing of precipitation, inverse drought index - for flood forecasting \citep{mcquigg1954}
\end{itemize}
\end{block}

\end{frame}

\note{these are some examples of early drought indices\\
\vspace{0.3cm} A numerical measure is needed to compare severity of drought across different time periods
or geographical locations. However, as a result of a large disagreement about a definition of drought, there is no single universal drought index. Instead of that a number of measures of drought has been developed\\
\vspace{0.3cm} f\\
\vspace{0.3cm}

}

\begin{frame}\label{Indices2}
\frametitle{Drought Indices}

\begin{itemize}
\item \textbf{Palmer Drought Severity Index} (PDSI, \citealp{palmer1965})

\begin{itemize}

\item Significant milestone in history of drought severity quantification
\item Based on a hydrological accounting system
\item Incorporate antecedent precipitation, moisture supply and moisture demand
\end{itemize}
\item \textbf{Standardized Precipitation Index} (SPI, \citealp{SPI})
\begin{itemize}
\item Can be interpreted as the number of standard deviations by which the observed value differs from the long-term mean
\item Standardised departure of observed precipitation from a chosen probability distribution function which models the precipitation data \citep{SPIonline}

\end{itemize}

\end{itemize}

\end{frame}

\note{On short timescales, the SPI is closely related to soil moisture, while at longer timescales, the SPI can be related to groundwater and reservoir storage.\\
\vspace{0.3cm}SPI: usuallygamma or a Pearson Type III distribution \\
\vspace{0.3cm} can be used across regions with very different climates\\
\vspace{0.3cm}SPEI no assumptions about underlining system
\\
\vspace{0.3cm} requires more data than SPI
\\
\vspace{0.3cm} SPEI sensitive to method that calculate potential evapotranspiration

\\
\vspace{0.3cm}
Evapotranspiration (ET) is the sum of evaporation and plant transpiration from the Earth's land and ocean surface to the atmosphere. Evaporation accounts for the movement of water to the air from sources such as the soil, canopy interception, and waterbodies. 
}

\begin{frame}\label{Indices3}
\frametitle{Drought Indices}

 \begin{itemize}
 \item \textbf{Standardized Precipitation Evapotranspiration Index} (SPEI, \citealp{SPEI})
\begin{itemize}
\item Extention of SPI
\item Accounts for potential evapotranspiration (hence captures impacts of increased temperature on water demand)
\end{itemize}
\item Number of other drought indicators and indices exist
\begin{itemize}
\item E.g.: Percent of Normal Precipitation, Drought Area Index (DAI), Soil Moisture Anomaly (SMA), Standardized Water-level Index (SWI), Normalized Difference Vegetation Index (NDVI)
\end{itemize}
\item  \citealp{monacelli2005, svoboda2016, zargar2011}: Overview of number of drought indices

\item Quantification and evaluation of drought indices for each of the three physical types of drought \citep{keyantash2002}
\end{itemize}

\end{frame}

\note{ the examples of indices are from the following categories:meteorology, soil moisture, hydrology, remote sensing \citep{svoboda2016}
\vspace{0.3cm}f  \\
\vspace{0.3cm} f\\
\vspace{0.3cm}}

\begin{frame}

\frametitle{Data}\label{data}

mozna taky trochu popsat data?? nebo az u napadu, co budu delat??

\end{frame}

\section{Effects of Drought on Economy}

\begin{frame}

\frametitle{Effects of drought on economy}\label{Effects} 
\begin{block}{\textbf{Generalized Computable Equilibrium (CGE)} Models }

\begin{itemize}
\item \underline{\textbf{\cite{OxfamIDS}}}
\begin{itemize}
\item Exploring range of scenarios for food price increase in 2030
\begin{itemize}
\item 1. Baseline 2. Climate change 3. Climate change with adaptation 4. Adaptation only in sub-Saharan Africa
\end{itemize}
\item Global coverage, set of individual country models, linked through international trade
\item Effects of climate change (incl. drought) modelled as changes in factor productivity (usually negative)

\end{itemize}
\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item Ethiopia
\item Social Accounting Matrix (SAM)
\end{itemize}
\end{itemize}
\end{block}
\end{frame}

\note{Computable general equilibrium (CGE) models are a class of economic models that use actual economic data to estimate how an economy might react to changes in policy, technology or other external factors.\\
\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
\vspace{0.3cm}a\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a

\\
\vspace{0.3cm}
Evapotranspiration (ET) is the sum of evaporation and plant transpiration from the Earth's land and ocean surface to the atmosphere. Evaporation accounts for the movement of water to the air from sources such as the soil, canopy interception, and waterbodies. 
}

\begin{frame}{\small Recent Relavant Literature}\label{Previous research}
\Wider[2em]{
\vspace{0.2cm}
\linespread{0.9}
\begin{itemize}
\small

\item \textbf{\cite{robinson2010}}: CGE
\begin{itemize}
\item Ethiopia. Udelat tabulku, v jednom sloupci co je fixed, v jednom co determined by model
\item ktere promenne se tam dosadi a ktere vylezou
\end{itemize}
\item \textbf{\cite{OxfamIDS}} CGE global
\begin{itemize}
\item probably also describe this CGE, how does it work, inputs, outputs
maybe it is like not that production is exogenous and price endo. but interdependent. and we have production shock as input $\rightarrow$ it gives price shock as output $\rightarrow$ it has again effect on production volumes
\end{itemize}

\item \textbf{\cite{Carlsson2013}}
\begin{itemize}
\item Analyse preferences of distributing the burden of reducing CO\textsubscript{2}
\item Higher educated people have higher WTP for distribution that is less costly for their country
\end{itemize}

\end{itemize}}
\end{frame}

\begin{frame}\label{Data3}
\frametitle{\small Data}

\vspace{-0.3cm}
      \begin{columns}
        \column{.15\linewidth}
        n=$4592$
              \column{.85\linewidth}
\begin{table}
\footnotesize
\label{Demographics} 
\begin{tabular}{llllllll} 
\hline
\rowcolor{PaleGreen} \textbf{Age} & 20 & 30 & 40 & 50 & 60 & 70 & 80 \\
 \hline
 \textbf{Frequency}  &$805$ & $1113$ & $812$ & $791$ & $657$ & $361$ & $53$\\
 \hline
\end{tabular}
\end{table}
 \end{columns}

      \begin{columns}
        \column{.335\linewidth}
\begin{table}
\footnotesize
\label{Sex} 
\vspace{-0.7cm}
\begin{tabular}{lll} 
\hline
\rowcolor{PaleGreen} \textbf{Gender} & \textbf{Freq.}  \\
 \hline
 Male & $2250$ \\
 Female& $2310$ \\
 \hline
\end{tabular}
\end{table}

         \column{.65\linewidth}
         \vspace{-1.1cm}
         \begin{figure}
         \caption*{\footnotesize{\vspace{-0.05cm} Education}}
           \vspace{-0.2cm}
 % \includegraphics[width = 0.9\textwidth]{Education.png} 
    \end{figure}
      \end{columns}

 \begin{figure}
 \vspace{-1.2cm}

 \caption*{\footnotesize{\vspace{-0.2cm}  \hspace{-2cm} Occupation}}
  \hspace{-2.3cm}
  \vspace{-0.55cm}  
%\includegraphics[width = 0.6\textwidth]{Occupation.png}  
\end{figure}

\end{frame}

\section{Methods}

\linespread{1}

\note{2 hlavni skupiny prediktoru: demographics a tak, a popisne...
 \\
\vspace{1cm}Nemela jsem moc hypotezu a literatura rika ruzne veci, v kazdem clanku neco jineho. takze selection methods...}

\begin{frame}

\frametitle{Methods}\label{MethodsLasso} 
\textbf{Lasso} = Least Absolute Shrinkage and Selection Operator
\begin{itemize}
\item In literature preferred to stepwise procedures (Tibshirani, 1996; Efron et. al, 2003; Yuan and Lin, 2004; Shah, 2012) 
\begin{equation}\label{Lasso}
\begin{array}{lcll}

\widehat{\beta}(\lambda)&=& \underset{\beta}{arg~min} \bigg( \sum_{i=1}^{n}(Y_i - \boldsymbol{X} \beta)_i \bigg)^2 + \lambda \sum_{j=1}^p|\beta_j|, \\
\end{array}
\end{equation}

\begin{itemize}
\item $i$ ... index of observations
\item $j$ ... index of explanatory variables
\item $\lambda \geq 0$...penalty parameter
\end{itemize}

\item The estimator does variable selection in the sense that $\widehat{\beta_j}(\lambda) = 0 $ for some $j$'s
\item Cross-validation of a series of models with different values of $\lambda$
\item I choose the largest value of $\lambda$ such that the mean square error is within 1 standard error of the minimum MSE
\end{itemize}
\end{frame}

\note{\tiny
Why lasso prefered to stepwise??

why stepwise bad:
Efron at all: Stepwise can eliminate a good valid predictor in the second step if it happens to be correlted with the first selected predictor \\

http://www.stata.com/support/faqs/statistics/stepwise-regression-problems/\\

 It yields R-squared values that are badly biased to be high.\\
The F and chi-squared tests quoted next to each variable on the printout do not have the claimed distribution.\\
The method yields confidence intervals for effects and predicted values that are falsely narrow; see Altman and Andersen (1989).\\
It yields p-values that do not have the proper meaning, and the proper correction for them is a difficult problem.\\
It gives biased regression coefficients that need shrinkage (the coefficients for remaining variables are too large; see Tibshirani [1996]).\\
It has severe problems in the presence of collinearity.\\
It is based on methods (e.g., F tests for nested models) that were intended to be used to test prespecified hypotheses. \\

“The degree of correlation between the predictor variables affected the frequency with which authentic predictor variables found their way into the final model.” \\
 p is the number of parameters	 \\
\hspace{1cm} the selection depends on lambda
\hspace{1cm} R cross validates various models for series of different lamdas.
\hspace{1cm} }

\begin{frame}

\frametitle{Cumulative logits model}\label{Results-Logit2}

 $Y$ is a categorical response with $M$ categories $(m=1...M)$

\begin{equation}\label{PCumLog}
\begin{array}{lclcl}
logit  \big[P( Y \leq m| x) \big] &=& \alpha_m + \boldsymbol{\beta'x}, ~~m=1,...,M-1
\end{array}
\end{equation}

\begin{align*}\label{Stickin}
\begin{array}{lcl}
\text{Substitute estimates into:} \hspace{1cm}\hat{P}(Y\leq m) &=& \frac{exp(\hat{\alpha_m}+ \hat{\beta} x)}{1+exp(\hat{\alpha_m}+ \hat{\beta} x)}
\end{array}
\end{align*}

\vspace{0.5cm}
\begin{itemize}
\item \textbf{Proportional odds assumption}
\begin{itemize}
\item Likelihood ratio test for comparison with multinomial logit insignificant $\boldsymbol{\Rightarrow}$ assumption satisfied

\end{itemize}

\end{itemize}

\end{frame}

\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc ok}

\section{Results-Knowledge}\label{Results-Knowledge}

\section{Results-WTP}

\begin{frame}[plain]
\Wider[3.5em]{\vspace{-0.5cm}
\frametitle{\small Results: Willingness to pay}\label{ResultsWTP1} 
\begin{table}
%  bodie38  bodie70 Dunquan98  Dunquan80
\label{WTP1} 
\begin{footnotesize}

\begin{tabular}{m{0.2cm}m{0.05cm}|l|cc|ccc|} 
\hline
\rowcolor{PaleGreen} \multicolumn{3}{c|}{\textbf{OLS after selection by lasso}} &  \multicolumn{2}{c|}{\textbf{Tax gas}} &\multicolumn{2}{c}{\textbf{Tax fuel}} \\
\rowcolor{PaleGreen} \multicolumn{3}{c|}{Robust std. errors HC\textsubscript{0}, n=4178}& \textbf{(\RN{1})}&\textbf{(\RN{2})} & \textbf{(\RN{1})}&\textbf{(\RN{2}) } \\
\hline
 & &Female& + & +& -&+ \\ 
& & Age& -  *** & -  ***& -  ***& -  ***  \\ 
& & Debts& +  *** &+  ***&  +  ***&  +  *** \\ 
\RowColor \cellcolor{white} \multirow{-5}{*}{\hspace{-0.5cm} \vspace{-0.5cm} \rotatebox[origin=c]{55}{\textbf{\scriptsize{Demographics}}}}  &\cellcolor{white} & \cellcolor{white}Occupation & \multicolumn{2}{c}{\cellcolor{white} vary} &  \multicolumn{2}{|c}{ \textcolor{red}{Not included}    } \\
\hline
& & Understands interest & -  *** &-  ***  &-  ** &-  **\\ 
& & Understands inflation & -  *** &-  ***  &-  ***&-  ***\\

 \RowColor \cellcolor{white} \multirow{-3}{*}{\hspace{-0.3cm}\rotatebox[origin=c]{60}{\textbf{\scriptsize{Numeracy}}}} & \cellcolor{white} & \cellcolor{white}Cognitive reflection test&\multicolumn{2}{c}{\textcolor{red}{Not included}} &  \cellcolor{white} -   **&  \cellcolor{white} -   **\\

\hline

&  &Not getting fair share& \multirow{2}{*}{- ***}& \multirow{2}{*}{- ***} &\multirow{2}{*}{- **} & \multirow{2}{*}{- **}\\ 
& & \hspace{0.3cm} of nation's wealth& & \\ 
& & Hard work is important& -  ***& -  ***& -  ***& -  ***\\ 
\vspace{-0.2cm}& & Parents + kids have same&\multirow{2}{*}{+ ***}&  \multirow{2}{*}{+ ***} &  \multirow{2}{*}{+ ***}&  \multirow{2}{*}{+ ***} \\ 
\multirow{-5}{*}{\rotatebox[origin=c]{90}{\textbf{\textbf{Attitudes,}}}} & \hspace{-0.5cm} \multirow{-5}{*}{\rotatebox[origin=c]{90}{\textbf{\textbf{beliefs}}}} & \hspace{0.1cm} living standards as me& & \\

\hline

\RowColor \cellcolor{white} &\cellcolor{white} &\cellcolor{white}Clim. change risk perception&  \cellcolor{white}+ *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}&  \cellcolor{white}+ *** &\textcolor{red}{Not incl.}\\

\RowColor \cellcolor{white} &\cellcolor{white}&\cellcolor{white}Clim. change vs policies risk&  \cellcolor{white}+ *** & \cellcolor{white}+ *** &  \cellcolor{white}+ *** &  \cellcolor{white}+ ***\\

\RowColor \cellcolor{white}    & \cellcolor{white}    &  \cellcolor{white}How much is gas/fuel tax & \cellcolor{white} + *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}& \cellcolor{white} + *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}\\

 \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{Climate}}}     &  \hspace{-0.5cm} \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{change}}}    &  Female $\times$ Clim. policies & - ** & - * &  - & -\\

\hline
\RowColor \cellcolor{white} & \cellcolor{white} & \cellcolor{white}Inequity aversion index& \cellcolor{white} \scriptsize + {\tiny then} -  ***& \multicolumn{1}{c}{\textcolor{red}{Not incl.}} &\cellcolor{white} \scriptsize + {\tiny then} -  ***& \multicolumn{1}{c}{\textcolor{red}{Not incl.}} \\ 
\RowColor \cellcolor{white}&\cellcolor{white}&\cellcolor{white}Discount rate yr. 0 vs.1 &\cellcolor{white} + * &\multicolumn{1}{c}{\textcolor{red}{Not incl.}} &\cellcolor{white} + $\bullet$ &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}\\
&  & Consistent discount rate & + *** &+ ***&+ ***&+ *** \\
\RowColor \cellcolor{white} \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{\scriptsize{Behavioral}}}}& \cellcolor{white} &\cellcolor{white}Egalitarian & \multicolumn{2}{c}{\textcolor{red}{Not included}} & \cellcolor{white}- ***& \cellcolor{white}-*** \\
\hline
\end{tabular} 
\end{footnotesize}
\end{table} }
%checked 31.3.2017
\end{frame}

\note{Consistent answers within investments==0 consdr \\
climpol a climcare popsat otazky, mozna udelat o nich slide
\\

occupation: homemaker unemployed negative, otherwise insignificant \\

 to je taky duvod, proc si myslim, ze numeracy a literacy...ze to muze byt v politickych nazorech?? ze levicaci nemaji technicke vzdelani???}

\begin{frame}
% bodie53, Dunquan68lm
\frametitle{Results: Willingness to pay - robustness}\label{ResultsLassoGas}

\begin{itemize}
\item OLS after selection by lasso
\item With additional control variables
\end{itemize}

\begin{table}
\label{LassoBigSample} 
\begin{small}
\begin{tabular}{l|cc|cc} 
\hline
\rowcolor{PaleGreen} Dependent variable: &  \multicolumn{2}{c|}{\textbf{Tax gas}} &\multicolumn{2}{c}{\textbf{Tax fuel}} \\
\hline
Demographics & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust} \\ 
Numeracy & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
Attitudes, beliefs & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
Climate change & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
Behavioral & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
\hline

  Income & -& & - &**\\
    Climate knowledge &- & $\bullet$ & +\\
Social value orientation & + & &+&*\\

 Risk aversion Kehneman-Tversky & - & & + & \\
  Government should redistribute income & + &$\bullet$ &+& $\bullet$\\
\hline
 \multicolumn{1}{l|}{Robust standard errors HC\textsubscript{0}}&  \multicolumn{2}{l|}{n=$4178$}&  \multicolumn{2}{l}{n=$4184$}\\
\end{tabular} 
\end{small}
\end{table}

\end{frame}

\note{ redistribute jsem dala jako politickou. Mozna ukazat bar??

}

\begin{frame}

         \textbf{Average preferred tax rate for \colorbox{babypink}{gas}} grouped by:
\begin{itemize}       
    \item    \textbf{\small Understands compound interest}

     \begin{figure}
      \begin{columns}
        \column{.4\linewidth}

  \caption*{ \hspace{0.6cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.6cm} \textcolor{red}{$\blacksquare$} Female
  \\ \vspace{0.2cm} \hspace{0.6cm} $1$ = Understands}
        \column{.6\linewidth}

 %   \vspace{-0.2cm}   \hspace{0.6cm}   \includegraphics[width = 0.8\textwidth]{GasInterest.png}  %565*238
 \vspace{-0.3cm}   \caption*{\footnotesize \hspace{0.6cm}$0$ \hspace{1cm}$0.5$ \hspace{1cm}$1$ }
      \end{columns}
    \end{figure}

\vspace{-0.3cm}
    \item    \textbf{\small Hard work is important}
\end{itemize}

\vspace{-0.2cm}
    \begin{figure}
      \begin{columns}
        \column{.5\linewidth }
        \vspace{-1.5cm}
  \caption*{ \hspace{0.8cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.8cm} \textcolor{red}{$\blacksquare$} Female
  \\ \vspace{0.2cm} \hspace{0.8cm} $1$ = Not important at all  \\ \hspace{0.8cm} $5$ = Essential}
  \vspace{-1.5cm}        
        \column{.5\linewidth}

  % \includegraphics[width = 0.85\textwidth]{GasHardwork.png}  %565*238
 \vspace{-0.3cm}   \caption*{\footnotesize  $1$ \hspace{0.5cm}$2$ \hspace{0.5cm}$3$ \hspace{0.5cm}$4$ \hspace{0.5cm}$5$  }
      \end{columns}
    \end{figure}

\end{frame}

\note{hello}

\begin{frame}

    \textbf{Average preferred tax rate for \colorbox{babypink}{fuel}} grouped by:
    \begin{itemize}       
    \item \textbf{\small Cognitive Reflection Test}

    \begin{figure}
      \begin{columns}
        \column{.5\linewidth}
  \caption*{ \hspace{0.4cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.7cm} \textcolor{red}{$\blacksquare$} Female}
        \column{.5\linewidth}

  %    \vspace{-0.5cm}  \includegraphics[width = 1\textwidth]{FuelNumeracy.png}  %565*238
 \vspace{-0.3cm}   \caption*{\footnotesize \hspace{-4.5cm} Number of correct answers:\hspace{1.1cm} $0$\hspace{0.3cm}$0.5$ \hspace{0.2cm}$1$ \hspace{0.2cm}$1.5$ \hspace{0.2cm}$2$ \hspace{0.2cm}$2.5$\hspace{0.2cm}$3$ }
      \end{columns}
    \end{figure}

\vspace{-0.15cm} 
 \item \textbf{\small Ordinary working people do no get their fair share of the nation's wealth}
\end{itemize}
\vspace{-0.5cm}
  \begin{figure}
      \begin{columns}

        \column{.5\linewidth}
  \caption*{ \hspace{0.8cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.8cm} \textcolor{red}{$\blacksquare$} Female
  \\ \vspace{0.2cm} \hspace{0.8cm} $-2$ = Strongly disagree  \\ \hspace{1.1cm} $2$ = Strongly agree}
\vspace{-1cm}        
        \column{.5\linewidth}
   \vspace{-0.2cm}     \hspace{-0.8cm} 
  % \includegraphics[width = 0.9\textwidth]{FuelFairshare.png} \\ %565*238
  \caption*{\footnotesize \hspace{-1.6cm}$-2$ \hspace{0.5cm}$-1$ \hspace{0.5cm}$0$ \hspace{0.5cm}$1$ \hspace{0.5cm}$2$  }
      \end{columns}
    \end{figure}

\end{frame}

\note{hello}

\section{Conclusion}

\normalsize
\begin{frame}
\linespread{1.2}
\Wider[3.5em]{
\frametitle{Conclusion}
\begin{itemize}
\item The key factors associated with \textbf{climate knowledge}:
\begin{itemize}
\item Female: negative effect
\item Financial literacy: positive effect
\item Score on the Cognitive Reflection Test: positive effect 
\end{itemize}

\item The key factors associated with \textbf{WTP for climate change mitigation}:

\begin{itemize}

\item Age: negative effect
\item Numeracy, financial literacy: negative effect
\item Attitudes and beliefs about redistribution of income and wealth
\item Climate change risk perception: positive effect (robustness check)
\end{itemize}

\end{itemize}}
\linespread{1}
\end{frame}

\note{ Interesting for climate knowledge: behavioral variabless not significant, not even opinions. education ordinal.pops out for the smaller sample Lasso \\
Interesting for climate knowledge: some evidencem that also climcare in 100 years, but not good for proportional odds thing\\
Discount rate mildly significant, SVO little bit significant for fuel\\
WTP: HOW MUCH FEMALE WANT TO PAY DOES NOT DEPEND ON CLIMPOL SO MUCH}

\section*{~~}

\begin{frame}
\centering
\huge{Thank you for attention}
\end{frame}

\note{hello}

\begin{frame}\label{ClimpolClimcare}
\Wider[0.9em]{
\frametitle{Climate change and policies risk perception}
\linespread{1.2}
\textbf{Climate care} \\
\vspace{0.3cm}
How serious a problem do you think climate change is at this moment?
\\

\vspace{0.3cm}
{\footnotesize

\begin{tabular}{lrrrrrrrrrrr} 
\hline
\rowcolor{PaleGreen}Response&$0$&$1$&$2$&$3$&$4$&$5$&$6$&$7$&$8$&$9$&$10$\\
 \hline
 \cellcolor{PaleGreen}Frequency&$77$&$68$&$112$&$171$&$199$&$430$&$858$&$1025$&$786$&$395$&$471$\\
\hline
\end{tabular}}

}
\Wider[1.8em]{
\vspace{0.6cm}
\textbf{Climate policy} \\
\vspace{0.3cm}
Which affects you and your way of life more, climate change (10) or policies (0) to reduce
greenhouse gas emissions?
\\

\vspace{0.3cm}
{\footnotesize

\begin{tabular}{lrrrrrrrrrrr} 
\hline
\rowcolor{PaleGreen}Response&$0$&$1$&$2$&$3$&$4$&$5$&$6$&$7$&$8$&$9$&$10$\\
 \hline
 \cellcolor{PaleGreen}Frequency&$130$&$127$&$256$&$374$&$415$&$1174$&$670$&$621$&$397$&$205$&$223$\\
\hline
\end{tabular}}
}

\vspace{0.6cm}
n=$4592$

\end{frame}
\note{The same for seriousness in 10 and 100 years and policy for children and grandchildren}

\begin{frame}\label{Data3}
\frametitle{Data} 
n=$4592$

 \begin{figure}
 \caption*{Occupation}
%\includegraphics[width = 1\textwidth]{Occupation.png}  
\end{figure}

\end{frame}

\note{ hello}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Financial literacy} \\
\hspace{1cm} Number of correct answers out of three financial questions
\\
\vspace{0.3cm}
\begin{tabular}{lllllll} 
\hline
\multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{1.5} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{2.5} & \multicolumn{1}{c}{3}\\
\hline
\end{tabular}

\end{frame}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Understands interest} \\
\hspace{1cm}1 = Understands compound interest
\\
\vspace{0.3cm}
\begin{tabular}{llll} 
\hline
\multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1} \\
\hline
\end{tabular}

\vspace{1.3cm}

\textbf{Understands inflation} \\
\hspace{1cm}1 = Understands inflation
\\
\vspace{0.3cm}
\begin{tabular}{cccc} 
\hline
\multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1} \\
\hline
\end{tabular}

\end{frame}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Cognitive Reflection Test} \\
\begin{itemize}
 \item In survey called numeracy
 \item Frederick (2005)
 \item Number of correct answers - max is 3
\end{itemize}

\vspace{0.3cm}
\begin{tabular}{lllllll} 
\hline
\multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{1.5} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{2.5} & \multicolumn{1}{c}{3}\\
\hline
\end{tabular}

\vspace{1.3cm}

\end{frame}

\note{The Cognitive reflection test measures the ability or disposition to reflect on a question and resist reporting the first response that comes to mind}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Not getting fair share of nation's wealth} \\
\hspace{1cm}Ordinary working people do not get their fair share of the nation's wealth
\\
\vspace{0.3cm}
\begin{tabular}{lllll} 
\hline
Strongly disagree & Disagree & Neutral & Agree & Strongly agree \\
\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
\hline
\end{tabular}

\vspace{1.3cm}
\textbf{Hard work is important} \\
\hspace{1cm}How important is hard work for getting ahead in life?
\\
\vspace{0.3cm}
\begin{tabular}{lllll}

\hline
 Not important & Not very  & Fairly  & Very  & Essential \\
~~~at all & important & important & important & Essential \\
\multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5} \\
\hline
\end{tabular}

\end{frame}

\begin{frame}\label{Questions}

\frametitle{Survey questions and coding}
\linespread{1.2}
\textbf{Government should redistribute income} \\
\hspace{1cm}Government should redistribute income from the better off to those who are less well off
\\
\vspace{0.3cm}
\begin{tabular}{lllll} 
\hline
Strongly disagree & Disagree & Neutral & Agree & Strongly agree \\
\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
\hline
\end{tabular}

\end{frame}

\begin{frame}\label{Questions}
\Wider[1.5em]{\small
\frametitle{Survey questions and coding}
\linespread{1}
\textbf{Better of parents} \\
\hspace{1cm}Compared with your parents when they were about your age, are you better or worse in
your income and standard of living generally? 
\\
\vspace{0.45cm}
\begin{tabular}{lllll} 
\hline
\small Much worse off & \small Worse off &\small About equal &\small Better off &\small Much better off \\
\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
\hline
\end{tabular}

\vspace{0.45cm}
\textbf{Better of kids} \\
\hspace{1cm}Compared with you, do you think that your children, when they reach your age, will be
better or worse in their income and standard of living generally?
\\
\vspace{0.45cm}
\begin{tabular}{lllll} 
\hline
\small Much worse off & \small Worse off &\small About equal &\small Better off &\small Much better off \\
\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
\hline
\end{tabular}

\vspace{0.45cm}
\textbf{My parents and kids have about same living standards as me} \\
\hspace{1cm}1= Both questions above answered\textit{\textbf{ about equal }}(0) \\
\hspace{1cm}0= Otherwise
}
\end{frame}

\begin{frame}

\frametitle{Methods}\label{MethodsLogit} 
\textbf{The Logistic Regression Model:}
\begin{itemize}
\item Binary response variable $Y$, explanatory variable $X$
\item $\pi(x)~~~ =~~~ P(Y=1|X=x) ~~~=~~~ 1-P(Y=0|X=x)$

\begin{equation}\label{LRegression}
\begin{array}{lcl}

\pi(x) &=& \frac{exp(\alpha+ \beta x)}{1+exp(\alpha+ \beta x)}
\end{array}
\end{equation}

\item \textbf{Logit} = Log odds

\begin{equation}\label{Logit}
\begin{array}{lclcl}

logit \big[ \pi(x) \big] &=& log \frac{ \pi(x)}{1- \pi(x)} & \\
\\
\end{array}
\end{equation}

After substituting \eqref{LRegression} into \eqref{Logit} and working out:

\begin{equation}\label{Logit2}
\begin{array}{lclcl}

logit \big[ \pi(x) \big] &=& log \frac{ \pi(x)}{1- \pi(x)}  &=& \alpha + \beta x \\
\end{array}
\end{equation}

\item Substituting $\hat{\alpha} + \hat{\beta} x$ into \eqref{LRegression} we get probability that $y$ is equal to $1$ for particular $x$

\end{itemize}

\end{frame}

\note{ hello}

\begin{frame}

\frametitle{Methods}\label{MethodsCumLog1} 
\textbf{Cumulative Logit Models}
\begin{itemize}
\item Logits of cumulative response probabilities
\item Logistic regression generalized for ordinal response variable
\item $Y$ is a categorical response with $M$ categories $(m=1..M)$

\begin{equation}\label{PCumLog}
\begin{array}{lclcl}
P( Y \leq m| x) &=& \pi_1(x)+...+ \pi_m(x), ~~m=1,...,M
\end{array}
\end{equation}

\item The \textbf{cumulative logits} are defined as:
\begin{equation}\label{PCumLog}
\begin{array}{lclcl}
logit  \big[P( Y \leq m| x) \big] &=& log\frac{ P( Y \leq m| x)}{1- P( Y \leq m| x)}         \\
\\
 &=& log \frac{\pi_1(x)+...+ \pi_m(x)}{\pi_{m+1}(x)+...+ \pi_M(x)}, ~~m=1,...,M-1

\end{array}
\end{equation}

\end{itemize}

\end{frame}

\note{For nominal dependent variables, better model is MULTICATEGORY LOGIT model \\
each cumulative logit uses all J response categories
\\ amodel for logit[P<=j)[ alone is an ordinary logit model for a binary response where categories 1to j form one outcome and categories j+1 to J form the second.
\\
models can use all J-1 models in a single parsimonious model

}

\begin{frame}

\frametitle{Methods}\label{MethodsCumLog2} 
\textbf{Proportional Odds Model}
A model, that simultaneously uses all cumulative logits is:
\begin{equation}\label{PCumLog}
\begin{array}{lclcl}
logit  \big[P( Y \leq m| x) \big] &=& \alpha_m + \boldsymbol{\beta'x}, ~~m=1,...,M-1
\end{array}
\end{equation}
\hspace{-0.5cm}
\begin{itemize}
\item Each cumulative logit has its own intercept $\alpha_m$ increasing in $m$ 
\item The same effect $\boldsymbol{\beta}$ for each logit
\item The odds of response $\leq m$ at $\boldsymbol{x} = \boldsymbol{x_1}$ are $exp[\boldsymbol{\beta}'(\boldsymbol{x_1}-\boldsymbol{x_2})]$ times the odds at $\boldsymbol{x} = \boldsymbol{x_2}$
\\
\begin{itemize}
\item[$\boldsymbol{\hookrightarrow}$] Thus, log of odds ratio of cumulative probabilities is proportional to distance between $\boldsymbol{x_1}$ and $\boldsymbol{x_2}$ $\Rightarrow$ \textcolor{DarkGreen}{\textbf{PROPORTIONAL ODDS MODEL}}
\end{itemize}
\item Must satisfy the \textcolor{DarkGreen}{\textbf{proportional odds assumption}}
\begin{itemize}
\item Can be tested e.g. by likelihood ratio test against unconstrained multinomial logit
\end{itemize}
\end{itemize}
\end{frame}

\note{Model (5) proportional odds model constrains the J-1 response curves to have the same shape. Thus it is not the same as fitting separate logit model for each J
\\
test of proportional odds assumption: multinoial logit has much more parameters, separate beta forJ1 categories and alpha as well
}

\begin{frame}
\Wider[1em]{
\frametitle{\small Results: Climate knowledge - Cumulative logits model}\label{Results-Logit} 
\begin{table}\label{Results-LogitT1} %Avon233
\small
\begin{tabular}{l|cl|cl|cl} 
\hline
\rowcolor{blue} \multicolumn{7}{c}{\textbf{\textcolor{white}{Marginal effects, n=4592}}}  \\
\hline
\multicolumn{7}{c}{\vspace{-0.2cm}} \\

\hline
\rowcolor{LightCyan} \multicolumn{1}{l|}{\textbf{Correct answers:}} &\multicolumn{2}{c|}{$\boldsymbol{0}$} &\multicolumn{2}{c|}{$\boldsymbol{2}$}&\multicolumn{2}{c}{$\boldsymbol{4}$}\\

\hline
Female& ${4\times10^{-4}}$& *& $0.034$ &*** &$-0.013$&***\\

Financial literacy&$-1\times10^{-4}$&$\bullet$ & $-0.014$ &*** &$0.005$& ***\\

Cognitive reflection 1&$-3\times10^{-5}$ & &$-0.003$& &$0.001$& \\
Cognitive reflection 2&$-2\times10^{-4}$ &$\bullet$ &$-0.020$&** &$0.006$& ***\\
Cognitive reflection 3&${-4\times10^{-4}}$&* &$-0.043$ &*** &$0.005$&$\bullet$  \\

Climate care&$1\times10^{-4}$&$\bullet$  &$0.012$ &*** &$-0.005$&*** \\
Climate policy&$5\times10^{-5}$&* &$0.005$ & ***&$-0.002$& ***\\

\hline
\end{tabular}
%
\end{table}}

\vspace{-0.3cm}
\begin{table}\label{Results-LogitT1} %Avon233

\small
\begin{tabular}{l|cl|cl} 
\hline
\rowcolor{LightCyan} \multicolumn{1}{l|}{\textbf{Correct answers:}}&\multicolumn{2}{c|}{$\boldsymbol{6}$} &\multicolumn{2}{c}{$\boldsymbol{8}$}\\
\hline
Female&$-0.024$&***&$-3\times10^{-4}$&$\bullet$\\
Financial literacy&$0.010$&***&$1\times10^{-4}$&$\bullet$\\
Cognitive reflection 1&$0.002$& &$2\times10^{-5}$&\\
Cognitive reflection 2&$0.016$&**&$2\times10^{-4}$&\\
Cognitive reflection 3&$0.044$&***&$-5\times10^{-4}$&$\bullet$\\
Climate care&$-0.009$&***&$-1\times10^{-4}$&*\\
Climate policy&$-0.003$&***&$-4\times10^{-5}$&$\bullet$\\
\hline
\end{tabular}
%
\end{table}

\end{frame}

       \linespread{1}
\note{  HODNE ZAJIMAVE. REDISTRIBUTE.. Dan Cone, ktery napsal clanek na ty climate knowledge questions, mu vychazelo, ze politicke nazory signifikantni. Me taky ale opacne. u redistribut, kdyz pridam, tak ti co souhlasi maji zapornu koeficient!!!!!je to myslim proto, ze lide, kteri jsou levicaci spis budoumit vzdelani humanitni, sociologicke nez technicke. to je taky duvod, proc si myslim, ze numeracy a literacy...ze to muze byt v politickych nazorech?? ze levicaci nemaji technicke vzdelani???}

%\begin{frame}\label{Questions}

%\textcolor{red}{Social Value orientation explain}

%\textcolor{green}{Climcare a Climpol urcite}

%\textcolor{red}{results of stepwise as a backup at least?}

%\vspace{0.45cm}

%\end{frame}

\normalsize

\begin{frame}{Previous research}\label{Previous research}
\begin{itemize}
\item Worldwide: \textbf{Education} is the strongest predictor of climate change awareness

\item  \cite{Leiserowitz2015}

\begin{itemize}
\item{Data from 119 countries - global assessment}

\item{Predictors of climate change awareness:}
\begin{itemize}
\item Civic engagement
\item Communication access
\item Education
\item Geographic location
\item Household income
\end{itemize}

\item{Predictors of climate change risk perception:}
\begin{itemize}
\item Believes about \textbf{anthropogenic cause} of climate change
\item Perception of \textbf{local temperature changes}
\item Attitudes towards government efforts for environmental preservation 
\item Local water and air quality
\end{itemize}
\end{itemize}
\end{itemize}
\end{frame}

\note{Gallup World Poll}

\begin{frame}\label{Previous research2}
\begin{itemize}
\small

 \item \cite{Braaten2014}
\begin{itemize}
 \item Warm glow is important for motivating environmentally friendly behavior
 \item Motivation of good feeling from giving as contrast to 'pure altruism'
 \item Substantial share of contributions to carbon abatement not dependent on any direct climate effect
 \end{itemize}

  \item \cite{Carlsson2013}
\begin{itemize}
\item Preferences of distributing the burden of reducing CO\textsubscript{2}
\item Discrete choice experiment in USA and Chine
\item WTP varies with socioeconomic characteristics and attitudes
\item Educated people higher WTP for rules less costly for their country
\item Less costly rules for respondents' country preferred
 \end{itemize}

  \end{itemize} 
  \end{frame}

\begin{frame}\label{Previous research3}
\begin{itemize}
\small
 \item \cite{GlenkColombo2013}
\begin{itemize}
 \item Focus on risk aversion and willingness to pay for emission reduction 
 \item If uncertainty over outcomes exists, it is important to consider people’s risk preferences.
  \item Choice experiment in Scotland 
   \item Comparison of different specifications arising different assumptions about the way respondent process information on outcome-related risk
   \item Significant differences between different specifications can arise in terms of model fit and WTP estimates
   \item Outcome related risk matters to respondents
 \end{itemize}
  \end{itemize}
  \end{frame}

\begin{frame}\label{Previous research4}
\begin{itemize}
\small

  \item \cite{Kahan2015}
\begin{itemize}
 \item Use the same 8 climate knowledge questions mostly to point out problem that what is designed to measure knowledge often measures something else, like cultural identification
 \item Strong correlation between risk perception from global warming and political outlook (right = small risk)
  \item Correlation between beief in human cause of global warming and political outlook (left vs right)
 \end{itemize}

 \item \cite{Morrison2015}
\begin{itemize}
 \item Relationship between religion and climate change attitudes and behavior
 \item Differences across religious groups found in terms of:
\begin{itemize} 
 \item human induced climate change
 \item the level of consensus among scientists 
  \item their own efficacy 
  \item the need for policy responses
  \end{itemize}

 \end{itemize} 
  \item Ordered logits model
   \item Buddhists, atheists, and agnostics the most engaged with climate issues, Christian literalists the least engaged

  \end{itemize}
\end{frame}

\begin{frame}
\small
\frametitle{References}
\bibliographystyle{apa}

\bibliography{referencesFS}
\end{frame}

\end{document}

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation2.tex
adding files, initial

--- 2018-11-16 presentations/dec2017BMEcSeminar/presentation1.tex
@@
 
-\title{Drought and its Impacts on Economy in East Africa}  
+\title{Drought and its Effects on Economy}  
 \author{Monika Novackova} 
@@
 
-\begin{frame}
-\titlepage
-\end{frame} `
+\maketitle
+
+\normalsize
+
+\begin{frame}\label{My Interest}
+
+\large{My research interest:} \\
+\vspace{0.7cm}
+
+\large{\hspace{1cm}\textbf{Effects of droughts on economy in Kenya}}
+\end{frame}
 
@@
 
-\normalsize
-
-\section{Defining drought}
+\section{Definition of Drought}
 
@@
 
-\section{Drought indices}
+\section{Drought Indices}
 
@@
 \end{itemize}
-\item  \citealp{monacelli2005, svoboda2016, zargar2011}: Overview of number of drought indices
+\item  \textit{For a detailed overview of drought indices see e.g.:} \cite{monacelli2005}, \cite{svoboda2016} or \cite{zargar2011}
 
-\item Quantification and evaluation of drought indices for each of the three physical types of drought \citep{keyantash2002}
+\item \cite{keyantash2002}: The quantification of drought: an evaluation of drought indices
 \end{itemize}
@@
 
-\section{Effects of Drought on Economy}
+\section{Effects of Droughts}
 
@@
 
-\frametitle{Effects of drought on economy}\label{Effects} 
-\begin{block}{\textbf{Generalized Computable Equilibrium (CGE)} Models }
+\frametitle{Effects of droughts on economy}\label{Effects} 
+\begin{block}{\textbf{Computable General Equilibrium (CGE)} Models }
+\begin{itemize}
 
-\begin{itemize}
+\item Consists of equations describing model variables and a database (input-output tables, SAM matrix)
+\item Assuming optimizing behaviour (cost minimizing producers, optimal households demands)
+
 \item \underline{\textbf{\cite{OxfamIDS}}}
@@
 \item Global coverage, set of individual country models, linked through international trade
-\item Effects of climate change (incl. drought) modelled as changes in factor productivity (usually negative)
+\item Climate change (incl. drought) modelled as changes in factor productivity (usually negative)
 
 \end{itemize}
+
+\end{itemize}
+\end{block}
+\end{frame}
+
+\begin{frame}
+
+\frametitle{Effects of droughts on economy}\label{Effects} 
+\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
+\begin{itemize}
+
+\item[~] \underline{\textbf{\cite{OxfamIDS}}}
+\begin{itemize}
+\item Solves the within-country models and between-country trade
+relationships simultaneously
+\item 19 region, 12 sector/commodity group, 8 commodity groups
+\end{itemize}
+\begin{table}
+\begin{footnotesize}
+\begin{tabular}{lll} 
+\hline
+\rowcolor{PaleGreen}  & \textbf{Determined by model}  \\
+\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{(outputs)}  \\
+ \hline
+Agricultural productivity growth	& Production volumes of \\
+&  \hspace{0.5cm}food commodities \\
+Commodity shares in hh. expenditure &Production vol. per capita \\
+Shares of import in demand (commodities) & World market  \\
+GDP growth rates&\hspace{0.5cm} food prices (change)\\
+Population growth& Domestic food price (change)\\
+& Volumes of global trade\\
+Shares of food in hh. expenditure  & Aggregate index of
+\\
+ & \hspace{0.5cm}consumption p.capita
+\\
+ \hline
+\end{tabular}
+\end{footnotesize}
+\end{table}
+
+\end{itemize}
+\end{block}
+\end{frame}
+
+\note{\\
+\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
+\vspace{0.3cm}a\\
+\vspace{0.3cm}
+\vspace{0.3cm}
+}
+
+\begin{frame}
+
+\frametitle{Effects of droughts on economy}\label{Effects} 
+\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
+
+\begin{itemize}
 \item \underline{\textbf{\cite{robinson2010}}}
 \begin{itemize}
-\item Ethiopia
-\item Social Accounting Matrix (SAM)
+\item Ethiopia, Social Accounting Matrix (SAM)
+
+\item Model drought as a $20\%$ reduction in crop productivity and $20$ reduction of livestock capital
+\item Effects of shocks in production on cereal prices and food consumption
+\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
 \end{itemize}
@@
 
-\note{Computable general equilibrium (CGE) models are a class of economic models that use actual economic data to estimate how an economy might react to changes in policy, technology or other external factors.\\
-\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
-\vspace{0.3cm}a\\
+\note{
+\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP.
+Also models price transition between international and domestic prices..
+\\
+\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
+\\
+\vspace{0.3cm}a
+
+\\
+\vspace{0.3cm}
+}
+
+\begin{frame}
+
+\frametitle{Effects of droughts on economy}\label{Effects} 
+\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
+
+\begin{itemize}
+
+\item \underline{\textbf{\cite{robinson2010}}}
+\begin{itemize}
+\item 5 agro-ecological zones,46 production activities (incl. 35 zone specific agricultural production sectors), 22 commodity groups,  15 primary factors of production
+\end{itemize}
+
+\begin{table}
+\begin{footnotesize}
+\begin{tabular}{lll}
+
+\hline
+\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{Determined by model (outputs)}  \\
+ \hline
+Capital stock& Domestic price of each commodity \\
+Land (by region) & Land allocated across crops \\
+Supply of labor per skill type& Real wages\\
+Foreign capital inflow	& Real exchange rate\\
+Trade balance	& \\
+ \hline
+\end{tabular}
+\end{footnotesize}
+\end{table}
+
+\begin{footnotesize}
+
+\item The simulation use a 'balanced' macro closure in which aggregate investment, government demand, and consumption are fixed shares of total absorption
+\item Intermediate inputs into production are determined as fixed shares
+of the quantity of output
+
+\end{footnotesize}
+\end{itemize}
+\end{block}
+\end{frame}
+\note{\\
+
+\vspace{0.3cm}
 \vspace{0.3cm}a
@@
 \vspace{0.3cm}
-Evapotranspiration (ET) is the sum of evaporation and plant transpiration from the Earth's land and ocean surface to the atmosphere.
-Evaporation accounts for the movement of water to the air from sources such as the soil, canopy interception, and waterbodies.
 }
 
-\begin{frame}{\small Recent Relavant Literature}\label{Previous research}
-\Wider[2em]{
-\vspace{0.2cm}
-\linespread{0.9}
+\begin{frame}
+
+\frametitle{Effects of droughts on economy}\label{Effects}
+
 \begin{itemize}
-\small
 
-\item \textbf{\cite{robinson2010}}: CGE
+\item \textbf{\cite{Pedram2016}}
+
 \begin{itemize}
-\item Ethiopia. Udelat tabulku, v jednom sloupci co je fixed, v jednom co determined by model
-\item ktere promenne se tam dosadi a ktere vylezou
+\item National production losses per disaster (droughts, floods and extreme temperatures)
+\item Worldwide, superposed epoch analysis
+\item On average $10.1\%$ reduction of cereal production can be linked to drought
 \end{itemize}
-\item \textbf{\cite{OxfamIDS}} CGE global
+\item \textbf{\cite{Mehrabi2017}}
 \begin{itemize}
-\item probably also describe this CGE, how does it work, inputs, outputs
-maybe it is like not that production is exogenous and price endo.
-but interdependent.
-and we have production shock as input $\rightarrow$ it gives price shock as output $\rightarrow$ it has again effect on production volumes
+\item Cumulative production losses linked to extreme heat and drought events
+\item Per years (53), countries (131), commodities (6)
 \end{itemize}
 
-\item \textbf{\cite{Carlsson2013}}
-\begin{itemize}
-\item Analyse preferences of distributing the burden of reducing CO\textsubscript{2}
-\item Higher educated people have higher WTP for distribution that is less costly for their country
 \end{itemize}
 
-\end{itemize}}
 \end{frame}
+\note{Lesk: 9.1 percent reduction is attributable to extreme heat.
+They did not find any significant effect of extreme cold and 
+oods on production.
 
-\begin{frame}\label{Data3}
-\frametitle{\small Data}
+\vspace{0.3cm} mehrabi: estimate counter factual production in disaster years
+\vspace{0.3cm}mehrabi:units are a bit like a panel - years x commodity x country x disaster (disaster years and not disaster years
+\\mehrabi: difference in production of disaster years and non-disaster year average. Cumulative impact: sum per countries and disasters. Loss relatively to (divided by) production + loss of production
+\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola
+\\
+\vspace{0.3cm}a
 
-\vspace{-0.3cm}
-      \begin{columns}
-        \column{.15\linewidth}
-        n=$4592$
-              \column{.85\linewidth}
-\begin{table}
-\footnotesize
-\label{Demographics} 
-\begin{tabular}{llllllll} 
-\hline
-\rowcolor{PaleGreen} \textbf{Age} & 20 & 30 & 40 & 50 & 60 & 70 & 80 \\
- \hline
- \textbf{Frequency}  &$805$ & $1113$ & $812$ & $791$ & $657$ & $361$ & $53$\\
- \hline
-\end{tabular}
-\end{table}
- \end{columns}
+\\
+\vspace{0.3cm}
+}
 
-      \begin{columns}
-        \column{.335\linewidth}
-\begin{table}
-\footnotesize
-\label{Sex} 
-\vspace{-0.7cm}
-\begin{tabular}{lll} 
-\hline
-\rowcolor{PaleGreen} \textbf{Gender} & \textbf{Freq.}  \\
- \hline
- Male & $2250$ \\
- Female& $2310$ \\
- \hline
-\end{tabular}
-\end{table}
+\section{Effects of Climate}
 
-         \column{.65\linewidth}
-         \vspace{-1.1cm}
-         \begin{figure}
-         \caption*{\footnotesize{\vspace{-0.05cm} Education}}
-           \vspace{-0.2cm}
- % \includegraphics[width = 0.9\textwidth]{Education.png} 
-    \end{figure}
-      \end{columns}
+\begin{frame}
 
- \begin{figure}
- \vspace{-1.2cm}
+\frametitle{Effects of climate and weather on economy}\label{EffectsCLimate}
 
- \caption*{\footnotesize{\vspace{-0.2cm}  \hspace{-2cm} Occupation}}
-  \hspace{-2.3cm}
-  \vspace{-0.55cm}  
-%\includegraphics[width = 0.6\textwidth]{Occupation.png}  
-\end{figure}
+\begin{itemize}
+
+\item Effects on food prices
+
+\begin{itemize}
+
+\item \textbf{\cite{Brown2015}}
+\begin{itemize}
+\item  Climate measured by NDVI vegetation index 
+\item  Effects of NDVI and world food prices on local food prices
+\item $20\%$ prices affected by weather, $9\%$ by international prices
+\end{itemize}
+
+\end{itemize}
+
+\item Effects on production (production function approach)
+\begin{itemize}
+\item \textbf{\cite{Ochieng2016}}
+\begin{itemize}
+\item Effect of climate variability and change on crop revenue
+\item Augmented production function, Kenya
+\item  Household FE
+\end{itemize}
+\item \textbf{\cite{Deschenes2007Ric}:} US, positive effect of global warming
+\end{itemize}
+\item Effects on land prices or yields/acre (Ricardian analysis)
+\begin{itemize}
+\item  \textbf{\citet{kabubo2007}:} Kenya, warming harmful
+
+\item  \cite{SeoMendelsohn} and \cite{KMendelsohn2008}: mild and wet warming - positive, dry more severe warming - negative
+
+\end{itemize}
+
+\end{itemize}
+\end{frame}
+\note{brown, Khsingar - KALMAN FILTER
+
+\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	
+\\
+\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones
+\\
+\vspace{0.3cm}a
+
+\\
+\vspace{0.3cm}
+}
+% tady mozna pridat nejake ricardian studie a rozdelit to na dva slidy
+
+\section{My Idea}
+
+\begin{frame}
+
+\frametitle{My idea - panel estimation}\label{MySuggestion}
+
+My interest: \textbf{Effects of drought on economy in Kenya}
+\begin{itemize}
+
+\item \textbf{Outcome variable}
+
+\begin{itemize}
+\item Volumes of production (crop specific, total)
+\item Alternatively food prices, yields/per acre
+\end{itemize}
+
+\item \textbf{Units of analysis}
+\begin{itemize}
+\item Counties in Kenya $\times$ year
+\
+\end{itemize}
+
+\item \textbf{Explanatory variable of interest}
+
+\begin{itemize}
+\item Dummy variable (0/1) drought occurred in a particular county and year or not
+\item Several varieties - various specifications of drought:
+
+\end{itemize}
+
+\end{itemize}
 
@@
 
-\section{Methods}
+\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
+\\
+\vspace{0.3cm}a
 
-\linespread{1}
-
-\note{2 hlavni skupiny prediktoru: demographics a tak, a popisne...
 \\
-\vspace{1cm}Nemela jsem moc hypotezu a literatura rika ruzne veci, v kazdem clanku neco jineho. takze selection methods...}
+\vspace{0.3cm}
+}
 
@@
 
-\frametitle{Methods}\label{MethodsLasso} 
-\textbf{Lasso} = Least Absolute Shrinkage and Selection Operator
+\frametitle{My idea - panel estimation}\label{MySuggestion}
+
+My interest: \textbf{Effects of drought on economy in Kenya}
 \begin{itemize}
-\item In literature preferred to stepwise procedures (Tibshirani, 1996; Efron et. al, 2003; Yuan and Lin, 2004; Shah, 2012) 
-\begin{equation}\label{Lasso}
-\begin{array}{lcll}
 
-\widehat{\beta}(\lambda)&=& \underset{\beta}{arg~min} \bigg( \sum_{i=1}^{n}(Y_i - \boldsymbol{X} \beta)_i \bigg)^2 + \lambda \sum_{j=1}^p|\beta_j|, \\
-\end{array}
-\end{equation}
+\item \textbf{Explanatory variable of interest}
 
 \begin{itemize}
-\item $i$ ... index of observations
-\item $j$ ... index of explanatory variables
-\item $\lambda \geq 0$...penalty parameter
+\item Dummy variable (0/1) drought occurred in a particular county and year or not
+\item Several varieties - various specifications of drought:
+
+\begin{itemize}
+\item Drought index monthly - NDVI, SPI, SPEI 
+\item Drought index below a specific value ($-1$, $-2$ or $-3$)
+ \begin{description}[abc]  % for indentation of length of abc
+ \item[$\bullet$] At least for one month during a specific year
+ \item[$\bullet$] At least for one month during growing season in a specific year
+ \item[$\bullet$] At least for one month during long rains in a specific year
+  \item[$\bullet$] At least for two (three) consecutive months during a growing season in a specific year
+ \item[$\bullet$] At least for one month during two (three) consecutive growing seasons
+ \end{description}
+
 \end{itemize}
 
-\item The estimator does variable selection in the sense that $\widehat{\beta_j}(\lambda) = 0 $ for some $j$'s
-\item Cross-validation of a series of models with different values of $\lambda$
-\item I choose the largest value of $\lambda$ such that the mean square error is within 1 standard error of the minimum MSE
 \end{itemize}
+\end{itemize}
+
 \end{frame}
 
-\note{\tiny
-Why lasso prefered to stepwise??
+\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
+\\
+\vspace{0.3cm}a
 
-why stepwise bad:
-Efron at all: Stepwise can eliminate a good valid predictor in the second step if it happens to be correlted with the first selected predictor \\
+\\
+\vspace{0.3cm}
+}
 
-http://www.stata.com/support/faqs/statistics/stepwise-regression-problems/\\
+\begin{frame}
 
- It yields R-squared values that are badly biased to be high.\\
-The F and chi-squared tests quoted next to each variable on the printout do not have the claimed distribution.\\
-The method yields confidence intervals for effects and predicted values that are falsely narrow; see Altman and Andersen (1989).\\
-It yields p-values that do not have the proper meaning, and the proper correction for them is a difficult problem.\\
-It gives biased regression coefficients that need shrinkage (the coefficients for remaining variables are too large; see Tibshirani [1996]).\\
-It has severe problems in the presence of collinearity.\\
-It is based on methods (e.g., F tests for nested models) that were intended to be used to test prespecified hypotheses. \\
+\frametitle{My idea - panel estimation}\label{MySuggestion}
 
-“The degree of correlation between the predictor variables affected the frequency with which authentic predictor variables found their way into the final model.” \\
- p is the number of parameters	 \\
-\hspace{1cm} the selection depends on lambda
-\hspace{1cm} R cross validates various models for series of different lamdas.
-\hspace{1cm} }
+My interest: \textbf{Effects of drought on economy in Kenya}
+\begin{itemize}
+
+\item \textbf{Explanatory variable of interest}
+
+\begin{itemize}
+\item Dummy variable (0/1) drought occurred in a particular county and year or not
+
+\end{itemize}
+
+\item \textbf{Units of analysis:} Counties in Kenya $\times$ year
+
+\item \textbf{Control variables ?}
+
+\begin{itemize}
+\item GDP, population, land area, soil characteristics, climate (2 seasons)?
+\end{itemize}
+
+\item \textbf{Estimation methods}
+
+\begin{itemize}
+\item Fixed effects, SURE, Kalman filter, Box-Jenkins
+\end{itemize}
+
+\end{itemize}
+
+\end{frame}
+
+\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
+\\
+\vspace{0.3cm}a
+
+\\
+\vspace{0.3cm}
+}
 
@@
 
-\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc ok}
-
-\section{Results-Knowledge}\label{Results-Knowledge}
-
-\section{Results-WTP}
-
-\begin{frame}[plain]
-\Wider[3.5em]{\vspace{-0.5cm}
-\frametitle{\small Results: Willingness to pay}\label{ResultsWTP1} 
-\begin{table}
-%  bodie38  bodie70 Dunquan98  Dunquan80
-\label{WTP1} 
-\begin{footnotesize}
-
-\begin{tabular}{m{0.2cm}m{0.05cm}|l|cc|ccc|} 
-\hline
-\rowcolor{PaleGreen} \multicolumn{3}{c|}{\textbf{OLS after selection by lasso}} &  \multicolumn{2}{c|}{\textbf{Tax gas}} &\multicolumn{2}{c}{\textbf{Tax fuel}} \\
-\rowcolor{PaleGreen} \multicolumn{3}{c|}{Robust std. errors HC\textsubscript{0}, n=4178}& \textbf{(\RN{1})}&\textbf{(\RN{2})} & \textbf{(\RN{1})}&\textbf{(\RN{2}) } \\
-\hline
- & &Female& + & +& -&+ \\ 
-& & Age& -  *** & -  ***& -  ***& -  ***  \\ 
-& & Debts& +  *** &+  ***&  +  ***&  +  *** \\ 
-\RowColor \cellcolor{white} \multirow{-5}{*}{\hspace{-0.5cm} \vspace{-0.5cm} \rotatebox[origin=c]{55}{\textbf{\scriptsize{Demographics}}}}  &\cellcolor{white} & \cellcolor{white}Occupation & \multicolumn{2}{c}{\cellcolor{white} vary} &  \multicolumn{2}{|c}{ \textcolor{red}{Not included}    } \\
-\hline
-& & Understands interest & -  *** &-  ***  &-  ** &-  **\\ 
-& & Understands inflation & -  *** &-  ***  &-  ***&-  ***\\
-
- \RowColor \cellcolor{white} \multirow{-3}{*}{\hspace{-0.3cm}\rotatebox[origin=c]{60}{\textbf{\scriptsize{Numeracy}}}} & \cellcolor{white} & \cellcolor{white}Cognitive reflection test&\multicolumn{2}{c}{\textcolor{red}{Not included}} &  \cellcolor{white} -   **&  \cellcolor{white} -   **\\
-
-\hline
-
-&  &Not getting fair share& \multirow{2}{*}{- ***}& \multirow{2}{*}{- ***} &\multirow{2}{*}{- **} & \multirow{2}{*}{- **}\\ 
-& & \hspace{0.3cm} of nation's wealth& & \\ 
-& & Hard work is important& -  ***& -  ***& -  ***& -  ***\\ 
-\vspace{-0.2cm}& & Parents + kids have same&\multirow{2}{*}{+ ***}&  \multirow{2}{*}{+ ***} &  \multirow{2}{*}{+ ***}&  \multirow{2}{*}{+ ***} \\ 
-\multirow{-5}{*}{\rotatebox[origin=c]{90}{\textbf{\textbf{Attitudes,}}}} & \hspace{-0.5cm} \multirow{-5}{*}{\rotatebox[origin=c]{90}{\textbf{\textbf{beliefs}}}} & \hspace{0.1cm} living standards as me& & \\
-
-\hline
-
-\RowColor \cellcolor{white} &\cellcolor{white} &\cellcolor{white}Clim.
-change risk perception&  \cellcolor{white}+ *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}&  \cellcolor{white}+ *** &\textcolor{red}{Not incl.}\\
-
-\RowColor \cellcolor{white} &\cellcolor{white}&\cellcolor{white}Clim. change vs policies risk&  \cellcolor{white}+ *** & \cellcolor{white}+ *** &  \cellcolor{white}+ *** &  \cellcolor{white}+ ***\\
-
-\RowColor \cellcolor{white}    & \cellcolor{white}    &  \cellcolor{white}How much is gas/fuel tax & \cellcolor{white} + *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}& \cellcolor{white} + *** &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}\\
-
- \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{Climate}}}     &  \hspace{-0.5cm} \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{change}}}    &  Female $\times$ Clim.
-policies & - ** & - * &  - & -\\
-
-\hline
-\RowColor \cellcolor{white} & \cellcolor{white} & \cellcolor{white}Inequity aversion index& \cellcolor{white} \scriptsize + {\tiny then} -  ***& \multicolumn{1}{c}{\textcolor{red}{Not incl.}} &\cellcolor{white} \scriptsize + {\tiny then} -  ***& \multicolumn{1}{c}{\textcolor{red}{Not incl.}} \\
-\RowColor \cellcolor{white}&\cellcolor{white}&\cellcolor{white}Discount rate yr.
-0 vs.1 &\cellcolor{white} + * &\multicolumn{1}{c}{\textcolor{red}{Not incl.}} &\cellcolor{white} + $\bullet$ &\multicolumn{1}{c}{\textcolor{red}{Not incl.}}\\
-&  & Consistent discount rate & + *** &+ ***&+ ***&+ *** \\
-\RowColor \cellcolor{white} \multirow{-4}{*}{\rotatebox[origin=c]{90}{\textbf{\scriptsize{Behavioral}}}}& \cellcolor{white} &\cellcolor{white}Egalitarian & \multicolumn{2}{c}{\textcolor{red}{Not included}} & \cellcolor{white}- ***& \cellcolor{white}-*** \\
-\hline
-\end{tabular} 
-\end{footnotesize}
-\end{table} }
-%checked 31.3.2017
-\end{frame}
-
-\note{Consistent answers within investments==0 consdr \\
-climpol a climcare popsat otazky, mozna udelat o nich slide
-\\
-
-occupation: homemaker unemployed negative, otherwise insignificant \\
-
- to je taky duvod, proc si myslim, ze numeracy a literacy...ze to muze byt v politickych nazorech?? ze levicaci nemaji technicke vzdelani???}
-
-\begin{frame}
-% bodie53, Dunquan68lm
-\frametitle{Results: Willingness to pay - robustness}\label{ResultsLassoGas}
-
-\begin{itemize}
-\item OLS after selection by lasso
-\item With additional control variables
-\end{itemize}
-
-\begin{table}
-\label{LassoBigSample} 
-\begin{small}
-\begin{tabular}{l|cc|cc} 
-\hline
-\rowcolor{PaleGreen} Dependent variable: &  \multicolumn{2}{c|}{\textbf{Tax gas}} &\multicolumn{2}{c}{\textbf{Tax fuel}} \\
-\hline
-Demographics & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust} \\ 
-Numeracy & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
-Attitudes, beliefs & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
-Climate change & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
-Behavioral & \multicolumn{2}{c|}{Robust}& \multicolumn{2}{c}{Robust}  \\ 
-\hline
-
-  Income & -& & - &**\\
-    Climate knowledge &- & $\bullet$ & +\\
-Social value orientation & + & &+&*\\
-
- Risk aversion Kehneman-Tversky & - & & + & \\
-  Government should redistribute income & + &$\bullet$ &+& $\bullet$\\
-\hline
- \multicolumn{1}{l|}{Robust standard errors HC\textsubscript{0}}&  \multicolumn{2}{l|}{n=$4178$}&  \multicolumn{2}{l}{n=$4184$}\\
-\end{tabular} 
-\end{small}
-\end{table}
-
-\end{frame}
-
-\note{ redistribute jsem dala jako politickou. Mozna ukazat bar??
-
-}
-
-\begin{frame}
-
-         \textbf{Average preferred tax rate for \colorbox{babypink}{gas}} grouped by:
-\begin{itemize}       
-    \item    \textbf{\small Understands compound interest}
-
-     \begin{figure}
-      \begin{columns}
-        \column{.4\linewidth}
-
-  \caption*{ \hspace{0.6cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.6cm} \textcolor{red}{$\blacksquare$} Female
-  \\ \vspace{0.2cm} \hspace{0.6cm} $1$ = Understands}
-        \column{.6\linewidth}
-
- %   \vspace{-0.2cm}   \hspace{0.6cm}   \includegraphics[width = 0.8\textwidth]{GasInterest.png}  %565*238
- \vspace{-0.3cm}   \caption*{\footnotesize \hspace{0.6cm}$0$ \hspace{1cm}$0.5$ \hspace{1cm}$1$ }
-      \end{columns}
-    \end{figure}
-
-\vspace{-0.3cm}
-    \item    \textbf{\small Hard work is important}
-\end{itemize}
-
-\vspace{-0.2cm}
-    \begin{figure}
-      \begin{columns}
-        \column{.5\linewidth }
-        \vspace{-1.5cm}
-  \caption*{ \hspace{0.8cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.8cm} \textcolor{red}{$\blacksquare$} Female
-  \\ \vspace{0.2cm} \hspace{0.8cm} $1$ = Not important at all  \\ \hspace{0.8cm} $5$ = Essential}
-  \vspace{-1.5cm}        
-        \column{.5\linewidth}
-
-  % \includegraphics[width = 0.85\textwidth]{GasHardwork.png}  %565*238
- \vspace{-0.3cm}   \caption*{\footnotesize  $1$ \hspace{0.5cm}$2$ \hspace{0.5cm}$3$ \hspace{0.5cm}$4$ \hspace{0.5cm}$5$  }
-      \end{columns}
-    \end{figure}
-
-\end{frame}
+\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc o}
 
 \note{hello}
-
-\begin{frame}
-
-    \textbf{Average preferred tax rate for \colorbox{babypink}{fuel}} grouped by:
-    \begin{itemize}       
-    \item \textbf{\small Cognitive Reflection Test}
-
-    \begin{figure}
-      \begin{columns}
-        \column{.5\linewidth}
-  \caption*{ \hspace{0.4cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.7cm} \textcolor{red}{$\blacksquare$} Female}
-        \column{.5\linewidth}
-
-  %    \vspace{-0.5cm}  \includegraphics[width = 1\textwidth]{FuelNumeracy.png}  %565*238
- \vspace{-0.3cm}   \caption*{\footnotesize \hspace{-4.5cm} Number of correct answers:\hspace{1.1cm} $0$\hspace{0.3cm}$0.5$ \hspace{0.2cm}$1$ \hspace{0.2cm}$1.5$ \hspace{0.2cm}$2$ \hspace{0.2cm}$2.5$\hspace{0.2cm}$3$ }
-      \end{columns}
-    \end{figure}
-
-\vspace{-0.15cm} 
- \item \textbf{\small Ordinary working people do no get their fair share of the nation's wealth}
-\end{itemize}
-\vspace{-0.5cm}
-  \begin{figure}
-      \begin{columns}
-
-        \column{.5\linewidth}
-  \caption*{ \hspace{0.8cm} \textcolor{blue}{$\blacksquare$} Male\\ \hspace{0.8cm} \textcolor{red}{$\blacksquare$} Female
-  \\ \vspace{0.2cm} \hspace{0.8cm} $-2$ = Strongly disagree  \\ \hspace{1.1cm} $2$ = Strongly agree}
-\vspace{-1cm}        
-        \column{.5\linewidth}
-   \vspace{-0.2cm}     \hspace{-0.8cm} 
-  % \includegraphics[width = 0.9\textwidth]{FuelFairshare.png} \\ %565*238
-  \caption*{\footnotesize \hspace{-1.6cm}$-2$ \hspace{0.5cm}$-1$ \hspace{0.5cm}$0$ \hspace{0.5cm}$1$ \hspace{0.5cm}$2$  }
-      \end{columns}
-    \end{figure}
-
-\end{frame}
-
-\note{hello}
-
-\section{Conclusion}
-
-\normalsize
-\begin{frame}
-\linespread{1.2}
-\Wider[3.5em]{
-\frametitle{Conclusion}
-\begin{itemize}
-\item The key factors associated with \textbf{climate knowledge}:
-\begin{itemize}
-\item Female: negative effect
-\item Financial literacy: positive effect
-\item Score on the Cognitive Reflection Test: positive effect 
-\end{itemize}
-
-\item The key factors associated with \textbf{WTP for climate change mitigation}:
-
-\begin{itemize}
-
-\item Age: negative effect
-\item Numeracy, financial literacy: negative effect
-\item Attitudes and beliefs about redistribution of income and wealth
-\item Climate change risk perception: positive effect (robustness check)
-\end{itemize}
-
-\end{itemize}}
-\linespread{1}
-\end{frame}
-
-\note{ Interesting for climate knowledge: behavioral variabless not significant, not even opinions. education ordinal.pops out for the smaller sample Lasso \\
-Interesting for climate knowledge: some evidencem that also climcare in 100 years, but not good for proportional odds thing\\
-Discount rate mildly significant, SVO little bit significant for fuel\\
-WTP: HOW MUCH FEMALE WANT TO PAY DOES NOT DEPEND ON CLIMPOL SO MUCH}
 
@@
 
-\begin{frame}\label{ClimpolClimcare}
-\Wider[0.9em]{
-\frametitle{Climate change and policies risk perception}
-\linespread{1.2}
-\textbf{Climate care} \\
-\vspace{0.3cm}
-How serious a problem do you think climate change is at this moment?
-\\
-
-\vspace{0.3cm}
-{\footnotesize
-
-\begin{tabular}{lrrrrrrrrrrr} 
-\hline
-\rowcolor{PaleGreen}Response&$0$&$1$&$2$&$3$&$4$&$5$&$6$&$7$&$8$&$9$&$10$\\
- \hline
- \cellcolor{PaleGreen}Frequency&$77$&$68$&$112$&$171$&$199$&$430$&$858$&$1025$&$786$&$395$&$471$\\
-\hline
-\end{tabular}}
-
-}
-\Wider[1.8em]{
-\vspace{0.6cm}
-\textbf{Climate policy} \\
-\vspace{0.3cm}
-Which affects you and your way of life more, climate change (10) or policies (0) to reduce
-greenhouse gas emissions?
-\\
-
-\vspace{0.3cm}
-{\footnotesize
-
-\begin{tabular}{lrrrrrrrrrrr} 
-\hline
-\rowcolor{PaleGreen}Response&$0$&$1$&$2$&$3$&$4$&$5$&$6$&$7$&$8$&$9$&$10$\\
- \hline
- \cellcolor{PaleGreen}Frequency&$130$&$127$&$256$&$374$&$415$&$1174$&$670$&$621$&$397$&$205$&$223$\\
-\hline
-\end{tabular}}
-}
-
-\vspace{0.6cm}
-n=$4592$
-
-\end{frame}
-\note{The same for seriousness in 10 and 100 years and policy for children and grandchildren}
-
-\begin{frame}\label{Data3}
-\frametitle{Data} 
-n=$4592$
-
- \begin{figure}
- \caption*{Occupation}
-%\includegraphics[width = 1\textwidth]{Occupation.png}  
-\end{figure}
-
-\end{frame}
-
-\note{ hello}
-
-\begin{frame}\label{Questions}
-
-\frametitle{Survey questions and coding}
-\linespread{1.2}
-\textbf{Financial literacy} \\
-\hspace{1cm} Number of correct answers out of three financial questions
-\\
-\vspace{0.3cm}
-\begin{tabular}{lllllll} 
-\hline
-\multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{1.5} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{2.5} & \multicolumn{1}{c}{3}\\
-\hline
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}\label{Questions}
-
-\frametitle{Survey questions and coding}
-\linespread{1.2}
-\textbf{Understands interest} \\
-\hspace{1cm}1 = Understands compound interest
-\\
-\vspace{0.3cm}
-\begin{tabular}{llll} 
-\hline
-\multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1} \\
-\hline
-\end{tabular}
-
-\vspace{1.3cm}
-
-\textbf{Understands inflation} \\
-\hspace{1cm}1 = Understands inflation
-\\
-\vspace{0.3cm}
-\begin{tabular}{cccc} 
-\hline
-\multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1} \\
-\hline
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}\label{Questions}
-
-\frametitle{Survey questions and coding}
-\linespread{1.2}
-\textbf{Cognitive Reflection Test} \\
-\begin{itemize}
- \item In survey called numeracy
- \item Frederick (2005)
- \item Number of correct answers - max is 3
-\end{itemize}
-
-\vspace{0.3cm}
-\begin{tabular}{lllllll} 
-\hline
-\multicolumn{1}{c}{0} & \multicolumn{1}{c}{0.5} & \multicolumn{1}{c}{1}  & \multicolumn{1}{c}{1.5} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{2.5} & \multicolumn{1}{c}{3}\\
-\hline
-\end{tabular}
-
-\vspace{1.3cm}
-
-\end{frame}
-
-\note{The Cognitive reflection test measures the ability or disposition to reflect on a question and resist reporting the first response that comes to mind}
-
-\begin{frame}\label{Questions}
-
-\frametitle{Survey questions and coding}
-\linespread{1.2}
-\textbf{Not getting fair share of nation's wealth} \\
-\hspace{1cm}Ordinary working people do not get their fair share of the nation's wealth
-\\
-\vspace{0.3cm}
-\begin{tabular}{lllll} 
-\hline
-Strongly disagree & Disagree & Neutral & Agree & Strongly agree \\
-\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
-\hline
-\end{tabular}
-
-\vspace{1.3cm}
-\textbf{Hard work is important} \\
-\hspace{1cm}How important is hard work for getting ahead in life?
-\\
-\vspace{0.3cm}
-\begin{tabular}{lllll}
-
-\hline
- Not important & Not very  & Fairly  & Very  & Essential \\
-~~~at all & important & important & important & Essential \\
-\multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} & \multicolumn{1}{c}{3} & \multicolumn{1}{c}{4} & \multicolumn{1}{c}{5} \\
-\hline
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}\label{Questions}
-
-\frametitle{Survey questions and coding}
-\linespread{1.2}
-\textbf{Government should redistribute income} \\
-\hspace{1cm}Government should redistribute income from the better off to those who are less well off
-\\
-\vspace{0.3cm}
-\begin{tabular}{lllll} 
-\hline
-Strongly disagree & Disagree & Neutral & Agree & Strongly agree \\
-\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
-\hline
-\end{tabular}
-
-\end{frame}
-
-\begin{frame}\label{Questions}
-\Wider[1.5em]{\small
-\frametitle{Survey questions and coding}
-\linespread{1}
-\textbf{Better of parents} \\
-\hspace{1cm}Compared with your parents when they were about your age, are you better or worse in
-your income and standard of living generally? 
-\\
-\vspace{0.45cm}
-\begin{tabular}{lllll} 
-\hline
-\small Much worse off & \small Worse off &\small About equal &\small Better off &\small Much better off \\
-\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
-\hline
-\end{tabular}
-
-\vspace{0.45cm}
-\textbf{Better of kids} \\
-\hspace{1cm}Compared with you, do you think that your children, when they reach your age, will be
-better or worse in their income and standard of living generally?
-\\
-\vspace{0.45cm}
-\begin{tabular}{lllll} 
-\hline
-\small Much worse off & \small Worse off &\small About equal &\small Better off &\small Much better off \\
-\multicolumn{1}{c}{-2} & \multicolumn{1}{c}{-1} & \multicolumn{1}{c}{0} & \multicolumn{1}{c}{1} & \multicolumn{1}{c}{2} \\
-\hline
-\end{tabular}
-
-\vspace{0.45cm}
-\textbf{My parents and kids have about same living standards as me} \\
-\hspace{1cm}1= Both questions above answered\textit{\textbf{ about equal }}(0) \\
-\hspace{1cm}0= Otherwise
-}
-\end{frame}
-
-\begin{frame}
-
-\frametitle{Methods}\label{MethodsLogit} 
-\textbf{The Logistic Regression Model:}
-\begin{itemize}
-\item Binary response variable $Y$, explanatory variable $X$
-\item $\pi(x)~~~ =~~~ P(Y=1|X=x) ~~~=~~~ 1-P(Y=0|X=x)$
-
-\begin{equation}\label{LRegression}
-\begin{array}{lcl}
-
-\pi(x) &=& \frac{exp(\alpha+ \beta x)}{1+exp(\alpha+ \beta x)}
-\end{array}
-\end{equation}
-
-\item \textbf{Logit} = Log odds
-
-\begin{equation}\label{Logit}
-\begin{array}{lclcl}
-
-logit \big[ \pi(x) \big] &=& log \frac{ \pi(x)}{1- \pi(x)} & \\
-\\
-\end{array}
-\end{equation}
-
-After substituting \eqref{LRegression} into \eqref{Logit} and working out:
-
-\begin{equation}\label{Logit2}
-\begin{array}{lclcl}
-
-logit \big[ \pi(x) \big] &=& log \frac{ \pi(x)}{1- \pi(x)}  &=& \alpha + \beta x \\
-\end{array}
-\end{equation}
-
-\item Substituting $\hat{\alpha} + \hat{\beta} x$ into \eqref{LRegression} we get probability that $y$ is equal to $1$ for particular $x$
-
-\end{itemize}
-
-\end{frame}
-
-\note{ hello}
-
-\begin{frame}
-
-\frametitle{Methods}\label{MethodsCumLog1} 
-\textbf{Cumulative Logit Models}
-\begin{itemize}
-\item Logits of cumulative response probabilities
-\item Logistic regression generalized for ordinal response variable
-\item $Y$ is a categorical response with $M$ categories $(m=1..M)$
-
-\begin{equation}\label{PCumLog}
-\begin{array}{lclcl}
-P( Y \leq m| x) &=& \pi_1(x)+...+ \pi_m(x), ~~m=1,...,M
-\end{array}
-\end{equation}
-
-\item The \textbf{cumulative logits} are defined as:
-\begin{equation}\label{PCumLog}
-\begin{array}{lclcl}
-logit  \big[P( Y \leq m| x) \big] &=& log\frac{ P( Y \leq m| x)}{1- P( Y \leq m| x)}         \\
-\\
- &=& log \frac{\pi_1(x)+...+ \pi_m(x)}{\pi_{m+1}(x)+...+ \pi_M(x)}, ~~m=1,...,M-1
-
-\end{array}
-\end{equation}
-
-\end{itemize}
-
-\end{frame}
-
-\note{For nominal dependent variables, better model is MULTICATEGORY LOGIT model \\
-each cumulative logit uses all J response categories
-\\ amodel for logit[P<=j)[ alone is an ordinary logit model for a binary response where categories 1to j form one outcome and categories j+1 to J form the second.
-\\
-models can use all J-1 models in a single parsimonious model
-
-}
-
-\begin{frame}
-
-\frametitle{Methods}\label{MethodsCumLog2} 
-\textbf{Proportional Odds Model}
-A model, that simultaneously uses all cumulative logits is:
-\begin{equation}\label{PCumLog}
-\begin{array}{lclcl}
-logit  \big[P( Y \leq m| x) \big] &=& \alpha_m + \boldsymbol{\beta'x}, ~~m=1,...,M-1
-\end{array}
-\end{equation}
-\hspace{-0.5cm}
-\begin{itemize}
-\item Each cumulative logit has its own intercept $\alpha_m$ increasing in $m$ 
-\item The same effect $\boldsymbol{\beta}$ for each logit
-\item The odds of response $\leq m$ at $\boldsymbol{x} = \boldsymbol{x_1}$ are $exp[\boldsymbol{\beta}'(\boldsymbol{x_1}-\boldsymbol{x_2})]$ times the odds at $\boldsymbol{x} = \boldsymbol{x_2}$
-\\
-\begin{itemize}
-\item[$\boldsymbol{\hookrightarrow}$] Thus, log of odds ratio of cumulative probabilities is proportional to distance between $\boldsymbol{x_1}$ and $\boldsymbol{x_2}$ $\Rightarrow$ \textcolor{DarkGreen}{\textbf{PROPORTIONAL ODDS MODEL}}
-\end{itemize}
-\item Must satisfy the \textcolor{DarkGreen}{\textbf{proportional odds assumption}}
-\begin{itemize}
-\item Can be tested e.g. by likelihood ratio test against unconstrained multinomial logit
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\note{Model (5) proportional odds model constrains the J-1 response curves to have the same shape. Thus it is not the same as fitting separate logit model for each J
-\\
-test of proportional odds assumption: multinoial logit has much more parameters, separate beta forJ1 categories and alpha as well
-}
-
-\begin{frame}
-\Wider[1em]{
-\frametitle{\small Results: Climate knowledge - Cumulative logits model}\label{Results-Logit} 
-\begin{table}\label{Results-LogitT1} %Avon233
-\small
-\begin{tabular}{l|cl|cl|cl} 
-\hline
-\rowcolor{blue} \multicolumn{7}{c}{\textbf{\textcolor{white}{Marginal effects, n=4592}}}  \\
-\hline
-\multicolumn{7}{c}{\vspace{-0.2cm}} \\
-
-\hline
-\rowcolor{LightCyan} \multicolumn{1}{l|}{\textbf{Correct answers:}} &\multicolumn{2}{c|}{$\boldsymbol{0}$} &\multicolumn{2}{c|}{$\boldsymbol{2}$}&\multicolumn{2}{c}{$\boldsymbol{4}$}\\
-
-\hline
-Female& ${4\times10^{-4}}$& *& $0.034$ &*** &$-0.013$&***\\
-
-Financial literacy&$-1\times10^{-4}$&$\bullet$ & $-0.014$ &*** &$0.005$& ***\\
-
-Cognitive reflection 1&$-3\times10^{-5}$ & &$-0.003$& &$0.001$& \\
-Cognitive reflection 2&$-2\times10^{-4}$ &$\bullet$ &$-0.020$&** &$0.006$& ***\\
-Cognitive reflection 3&${-4\times10^{-4}}$&* &$-0.043$ &*** &$0.005$&$\bullet$  \\
-
-Climate care&$1\times10^{-4}$&$\bullet$  &$0.012$ &*** &$-0.005$&*** \\
-Climate policy&$5\times10^{-5}$&* &$0.005$ & ***&$-0.002$& ***\\
-
-\hline
-\end{tabular}
-%
-\end{table}}
-
-\vspace{-0.3cm}
-\begin{table}\label{Results-LogitT1} %Avon233
-
-\small
-\begin{tabular}{l|cl|cl} 
-\hline
-\rowcolor{LightCyan} \multicolumn{1}{l|}{\textbf{Correct answers:}}&\multicolumn{2}{c|}{$\boldsymbol{6}$} &\multicolumn{2}{c}{$\boldsymbol{8}$}\\
-\hline
-Female&$-0.024$&***&$-3\times10^{-4}$&$\bullet$\\
-Financial literacy&$0.010$&***&$1\times10^{-4}$&$\bullet$\\
-Cognitive reflection 1&$0.002$& &$2\times10^{-5}$&\\
-Cognitive reflection 2&$0.016$&**&$2\times10^{-4}$&\\
-Cognitive reflection 3&$0.044$&***&$-5\times10^{-4}$&$\bullet$\\
-Climate care&$-0.009$&***&$-1\times10^{-4}$&*\\
-Climate policy&$-0.003$&***&$-4\times10^{-5}$&$\bullet$\\
-\hline
-\end{tabular}
-%
-\end{table}
-
-\end{frame}
-
-       \linespread{1}
-\note{  HODNE ZAJIMAVE.
-REDISTRIBUTE..
-Dan Cone, ktery napsal clanek na ty climate knowledge questions, mu vychazelo, ze politicke nazory signifikantni.
-Me taky ale opacne.
-u redistribut, kdyz pridam, tak ti co souhlasi maji zapornu koeficient!!!!!je to myslim proto, ze lide, kteri jsou levicaci spis budoumit vzdelani humanitni, sociologicke nez technicke.
-to je taky duvod, proc si myslim, ze numeracy a literacy...ze to muze byt v politickych nazorech??
-ze levicaci nemaji technicke vzdelani???}
-
-%\begin{frame}\label{Questions}
-
-%\textcolor{red}{Social Value orientation explain}
-
-%\textcolor{green}{Climcare a Climpol urcite}
-
-%\textcolor{red}{results of stepwise as a backup at least?}
-
-%\vspace{0.45cm}
-
-%\end{frame}
-
-\normalsize
-
-\begin{frame}{Previous research}\label{Previous research}
-\begin{itemize}
-\item Worldwide: \textbf{Education} is the strongest predictor of climate change awareness
-
-\item  \cite{Leiserowitz2015}
-
-\begin{itemize}
-\item{Data from 119 countries - global assessment}
-
-\item{Predictors of climate change awareness:}
-\begin{itemize}
-\item Civic engagement
-\item Communication access
-\item Education
-\item Geographic location
-\item Household income
-\end{itemize}
-
-\item{Predictors of climate change risk perception:}
-\begin{itemize}
-\item Believes about \textbf{anthropogenic cause} of climate change
-\item Perception of \textbf{local temperature changes}
-\item Attitudes towards government efforts for environmental preservation 
-\item Local water and air quality
-\end{itemize}
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\note{Gallup World Poll}
-
-\begin{frame}\label{Previous research2}
-\begin{itemize}
-\small
-
- \item \cite{Braaten2014}
-\begin{itemize}
- \item Warm glow is important for motivating environmentally friendly behavior
- \item Motivation of good feeling from giving as contrast to 'pure altruism'
- \item Substantial share of contributions to carbon abatement not dependent on any direct climate effect
- \end{itemize}
-
-  \item \cite{Carlsson2013}
-\begin{itemize}
-\item Preferences of distributing the burden of reducing CO\textsubscript{2}
-\item Discrete choice experiment in USA and Chine
-\item WTP varies with socioeconomic characteristics and attitudes
-\item Educated people higher WTP for rules less costly for their country
-\item Less costly rules for respondents' country preferred
- \end{itemize}
-
-  \end{itemize} 
-  \end{frame}
-
-\begin{frame}\label{Previous research3}
-\begin{itemize}
-\small
- \item \cite{GlenkColombo2013}
-\begin{itemize}
- \item Focus on risk aversion and willingness to pay for emission reduction 
- \item If uncertainty over outcomes exists, it is important to consider people’s risk preferences.
-  \item Choice experiment in Scotland 
-   \item Comparison of different specifications arising different assumptions about the way respondent process information on outcome-related risk
-   \item Significant differences between different specifications can arise in terms of model fit and WTP estimates
-   \item Outcome related risk matters to respondents
- \end{itemize}
-  \end{itemize}
-  \end{frame}
-
-\begin{frame}\label{Previous research4}
-\begin{itemize}
-\small
-
-  \item \cite{Kahan2015}
-\begin{itemize}
- \item Use the same 8 climate knowledge questions mostly to point out problem that what is designed to measure knowledge often measures something else, like cultural identification
- \item Strong correlation between risk perception from global warming and political outlook (right = small risk)
-  \item Correlation between beief in human cause of global warming and political outlook (left vs right)
- \end{itemize}
-
- \item \cite{Morrison2015}
-\begin{itemize}
- \item Relationship between religion and climate change attitudes and behavior
- \item Differences across religious groups found in terms of:
-\begin{itemize} 
- \item human induced climate change
- \item the level of consensus among scientists 
-  \item their own efficacy 
-  \item the need for policy responses
-  \end{itemize}
-
- \end{itemize} 
-  \item Ordered logits model
-   \item Buddhists, atheists, and agnostics the most engaged with climate issues, Christian literalists the least engaged
-
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
+\frame[allowframebreaks]{ 
 \small
@@
 \bibliographystyle{apa}
-
 \bibliography{referencesFS}
-\end{frame}
+}
 

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation3.tex
adding files, initial

--- 2018-11-16 presentations/dec2017BMEcSeminar/presentation2.tex
@@
 \end{frame}
+\note{A}
 
@@
 \end{frame}
+
+\note{A}
 
@@
 \note{Heim = a review of the 20tieth century indices, Trenberth = IPCC report, summary for policy makers \\
-\vspace{0.3cm}f  \\
-\vspace{0.3cm} f\\
-\vspace{0.3cm}
 
@@
 \vspace{0.3cm} SPEI sensitive to method that calculate potential evapotranspiration
-
-\\
 \vspace{0.3cm}
@@
 
-\begin{frame}
-
-\frametitle{Data}\label{data}
-
-mozna taky trochu popsat data?? nebo az u napadu, co budu delat??
-
-\end{frame}
-
 \section{Effects of Droughts}
@@
 
-\note{\\
+\note{
 \vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
-\vspace{0.3cm}a\\
 \vspace{0.3cm}
@@
 \vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
-\\
-\vspace{0.3cm}a
-
-\\
-\vspace{0.3cm}
-}
+\\qe
+\vspace{0.3cm}a\\}
 
@@
 \end{frame}
-\note{\\
+\note{
 
@@
 \vspace{0.3cm}a
-
-\\
-\vspace{0.3cm}
 }
@@
 oods on production.
-
 \vspace{0.3cm} mehrabi: estimate counter factual production in disaster years
@@
 \\mehrabi: difference in production of disaster years and non-disaster year average. Cumulative impact: sum per countries and disasters. Loss relatively to (divided by) production + loss of production
-\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola
-\\
+\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola\\
 \vspace{0.3cm}a
-
-\\
 \vspace{0.3cm}
@@
 \end{itemize}
-\item \textbf{\cite{Deschenes2007Ric}:} US, positive effect of global warming
+\item \textbf{\cite{Deschenes2007Ric}:} US county level, positive effect of global warming
 \end{itemize}
@@
 
-\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	
-\\
-\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones
-\\
-\vspace{0.3cm}a
-
-\\
-\vspace{0.3cm}
-}
+\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	\\
+\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones\\
+\vspace{0.3cm}a\\
+\vspace{0.3cm}}
 % tady mozna pridat nejake ricardian studie a rozdelit to na dva slidy
 
-\section{My Idea}
+\section{My Suggestion}
 
@@
 
-\frametitle{My idea - panel estimation}\label{MySuggestion}
+\frametitle{My suggestion - panel estimation}\label{MySuggestion1}
 
@@
 
-\item \textbf{Outcome variable}
+\item \textbf{Response variable}
 
@@
 \item Volumes of production (crop specific, total)
-\item Alternatively food prices, yields/per acre
+\item Profit per acre \normalsize{\citep{Deschenes2007Ric}}
+
+\begin{itemize}
+\item \footnotesize{(Value of agricultural products - prod. expanses)/acres (crops, pasture, grazing)}\\ 
 \end{itemize}
-
+\end{itemize}
 \item \textbf{Units of analysis}
@@
 
-\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county
-\\
-\vspace{0.3cm}a
-
-\\
-\vspace{0.3cm}
-}
+\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
+\vspace{0.3cm}a\\
+\vspace{0.3cm}}
 
@@
 
-\frametitle{My idea - panel estimation}\label{MySuggestion}
+\frametitle{My suggestion - panel estimation}\label{MySuggestion2}
 
@@
 
-\frametitle{My idea - panel estimation}\label{MySuggestion}
+\frametitle{My suggestion - panel estimation}\label{MySuggestion3}
 
@@
 
-\item \textbf{Control variables ?}
+\item \textbf{Control variables ? - subject to availability}
 
 \begin{itemize}
-\item GDP, population, land area, soil characteristics, climate (2 seasons)?
+\item GDP, soil quality data, population,average land area, average value assets,  climate, degree days?
 \end{itemize}
@@
 
-\frametitle{Cumulative logits model}\label{Results-Logit2}
+\frametitle{My idea - panel estimation}\label{MySuggestion4}
 
- $Y$ is a categorical response with $M$ categories $(m=1...M)$
-
-\begin{equation}\label{PCumLog}
-\begin{array}{lclcl}
-logit  \big[P( Y \leq m| x) \big] &=& \alpha_m + \boldsymbol{\beta'x}, ~~m=1,...,M-1
-\end{array}
-\end{equation}
-
-\begin{align*}\label{Stickin}
+\begin{align*}
 \begin{array}{lcl}
-\text{Substitute estimates into:} \hspace{1cm}\hat{P}(Y\leq m) &=& \frac{exp(\hat{\alpha_m}+ \hat{\beta} x)}{1+exp(\hat{\alpha_m}+ \hat{\beta} x)}
+ Y_{i,t} &=& \alpha_i +\gamma_t + \delta D'_{i,t} +  \boldsymbol{\beta X_{i,t}}+ \epsilon_{i,t}
 \end{array}
@@
 
-\vspace{0.5cm}
-\begin{itemize}
-\item \textbf{Proportional odds assumption}
-\begin{itemize}
-\item Likelihood ratio test for comparison with multinomial logit insignificant $\boldsymbol{\Rightarrow}$ assumption satisfied
+\begin{table}
+\begin{small}
+\begin{tabular}{lll}
 
-\end{itemize}
-
-\end{itemize}
+$Y_{i,t}$ &$=$& Response variable (food production/price), county $i$ in year $t$\\
+$\alpha_i$ &$=$& Fixed effects, county $i$\\
+{\color{darkblue}$\delta$ }&{\color{darkblue}$=$}&{\color{darkblue}{Effect of drought on economy}}\\
+$D_{i,t}$ &$=$& Indicator variable\\
+& &  $D=1$ if drought in county $i$ in year $t$, $D=0$ otherwise\\
+$\boldsymbol{\beta}$ &$=$& Vector of effects of other covariates\\
+$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
+$\epsilon_{i,t}$ &$=$& Error term \\
+{\color{DarkGreen}$\gamma_t$} & {\color{DarkGreen}$=$}& {\color{DarkGreen}Year specific indicator?}
+\end{tabular}
+\end{small}
+\end{table}
 
@@
 
-\section*{~~}
+\section*{}
 
-\begin{frame}
+\frame[allowframebreaks]{ 
 \centering
 \huge{Thank you for attention}
-\end{frame}
+}
 

### 2018-11-16 presentations/dec2017BMEcSeminar/presentation4.tex
adding files, initial

--- 2018-11-16 presentations/dec2017BMEcSeminar/presentation3.tex
@@
 e.g. \textit{An extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region \citep{schneider1996}}
-\item \textbf{Operational definitions:} Foundation for an effective early warning system \\ e.g. SPI, PDSI
-
+\item \textbf{Operational definitions:} Foundation for an effective early warning system \begin{itemize}
+\item e.g.  Standardized Precipitation Index (SPI), Palmer Drought Severity Index (PDSI)
+\end{itemize} 
 \end{itemize}
@@
 \item Can be interpreted as the number of standard deviations by which the observed value differs from the long-term mean
-\item Standardised departure of observed precipitation from a chosen probability distribution function which models the precipitation data \citep{SPIonline}
+\item Standardized departure of observed precipitation from a chosen probability distribution function which models the precipitation data \citep{SPIonline}
 
@@
 \frametitle{Effects of droughts on economy}\label{Effects} 
-\begin{block}{\textbf{Computable General Equilibrium (CGE)} Models }
+\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
+
 \begin{itemize}
@@
 \item Assuming optimizing behaviour (cost minimizing producers, optimal households demands)
+
+\vspace{0.3cm}
+
+\item \underline{\textbf{\cite{robinson2010}}}
+\begin{itemize}
+\item Ethiopia, Social Accounting Matrix (SAM)
+
+\item Model drought as a $20\%$ reduction in crop productivity and $20\%$ reduction of livestock capital
+\item Effects of shocks in production on cereal prices and food consumption
+\item Results in $2.3\%$ decrease in total GDP
+\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
+\end{itemize}
+\end{itemize}
+\end{block}
+\end{frame}
+
+\note{
+\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP.
+Also models price transition between international and domestic prices..
+\\
+\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
+\\qe
+\vspace{0.3cm}a\\}
+
+\begin{frame}
+
+\frametitle{Effects of droughts on economy}\label{Effects} 
+\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
+
+\begin{itemize}
+
+\item \underline{\textbf{\cite{robinson2010}}}
+\begin{itemize}
+\item 5 agro-ecological zones,46 production activities (incl. 35 zone specific agricultural production sectors), 22 commodity groups,  15 primary factors of production
+\end{itemize}
+
+\begin{table}
+\begin{footnotesize}
+\begin{tabular}{lll}
+
+\hline
+\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{Determined by model (outputs)}  \\
+ \hline
+Capital stock& Domestic price of each commodity \\
+Land (by region) & Land allocated across crops \\
+Supply of labor per skill type& Real wages\\
+Foreign capital inflow	& Real exchange rate\\
+Trade balance	& \\
+ \hline
+\end{tabular}
+\end{footnotesize}
+\end{table}
+
+\begin{footnotesize}
+
+\item The simulation use a 'balanced' macro closure in which aggregate \textbf{investment, government demand, and consumption are fixed shares of total absorption}
+\item Intermediate inputs into production are determined as fixed shares
+of the quantity of output
+
+\end{footnotesize}
+\end{itemize}
+\end{block}
+\end{frame}
+\note{
+
+\vspace{0.3cm}
+\vspace{0.3cm}a
+\\
+\vspace{0.3cm}a
+\\
+\vspace{0.3cm}a
+}
+
+\begin{frame}
+
+\frametitle{Effects of droughts on economy}\label{Effects} 
+\begin{block}{\textbf{Computable General Equilibrium (CGE)} Models }
+\begin{itemize}
 
@@
 \frametitle{Effects of droughts on economy}\label{Effects}
-\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
-
-\begin{itemize}
-\item \underline{\textbf{\cite{robinson2010}}}
-\begin{itemize}
-\item Ethiopia, Social Accounting Matrix (SAM)
-
-\item Model drought as a $20\%$ reduction in crop productivity and $20$ reduction of livestock capital
-\item Effects of shocks in production on cereal prices and food consumption
-\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
-\end{itemize}
-\end{itemize}
-\end{block}
-\end{frame}
-
-\note{
-\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP.
-Also models price transition between international and domestic prices..
-\\
-\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
-\\qe
-\vspace{0.3cm}a\\}
-
-\begin{frame}
-
-\frametitle{Effects of droughts on economy}\label{Effects} 
-\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
-
-\begin{itemize}
-
-\item \underline{\textbf{\cite{robinson2010}}}
-\begin{itemize}
-\item 5 agro-ecological zones,46 production activities (incl. 35 zone specific agricultural production sectors), 22 commodity groups,  15 primary factors of production
-\end{itemize}
-
-\begin{table}
-\begin{footnotesize}
-\begin{tabular}{lll}
-
-\hline
-\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{Determined by model (outputs)}  \\
- \hline
-Capital stock& Domestic price of each commodity \\
-Land (by region) & Land allocated across crops \\
-Supply of labor per skill type& Real wages\\
-Foreign capital inflow	& Real exchange rate\\
-Trade balance	& \\
- \hline
-\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\begin{footnotesize}
-
-\item The simulation use a 'balanced' macro closure in which aggregate investment, government demand, and consumption are fixed shares of total absorption
-\item Intermediate inputs into production are determined as fixed shares
-of the quantity of output
-
-\end{footnotesize}
-\end{itemize}
-\end{block}
-\end{frame}
-\note{
-
-\vspace{0.3cm}
-\vspace{0.3cm}a
-\\
-\vspace{0.3cm}a
-\\
-\vspace{0.3cm}a
-}
-
-\begin{frame}
-
-\frametitle{Effects of droughts on economy}\label{Effects}
 
@@
 \item Per years (53), countries (131), commodities (6)
+\item Most severe losses in Botswana ($5.7\%$), Paraguay ($5.5\%$), Nigeria ($4.8\%$)
 \end{itemize}

### 2018-11-16 monika.R

@@ -1,0 +1,52 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+
+emissions<-read.csv( "emissions.csv",header=TRUE,dec=".")
+
+#ooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+summary(emissions)
+
+#1
+plot(emissions$noise_level)
+which(emissions$noise_level<20)
+sum(which(emissions$noise_level<20))
+
+
+#2
+plot(emissions$extra_urban_metric)
+emissions[emissions$extra_urban_metric>80 & !is.na(emissions$extra_urban_metric),]
+emissions$model[emissions$extra_urban_metric>80 & !is.na(emissions$extra_urban_metric)]
+
+#3
+names(emissions)
+lm1<-lm(  (emissions$combined_metric-emissions$extra_urban_metric)~ I(emissions$urban_metric-emissions$extra_urban_metric )  -1)
+summary(lm1)
+
+#4
+i<-which(emissions$extra_urban_metric>80 & !is.na(emissions$extra_urban_metric))
+w<-summary(lm1)$coef[1]
+
+newVal<-(emissions$combined_metric[i]-w*emissions$urban_metric[i])/(1-w)
+
+#5
+par(mar=c(5,5,5,5))
+plot(log(emissions$urban_imperial)~log(emissions$urban_metric))
+
+#7
+
+aggregate(emissions$noise_level ,list( emissions$manufacturer), mean)
+
+min(aggregate(emissions$noise_level ,list( emissions$manufacturer), mean)[2],na.rm=TRUE)
+
+#8
+
+lm2<-lm(co2~engine_capacity,data=emissions)
+summary(lm2)
+a=lm2$coeff[1]+lm2$coefficients[2]*4000
+a

### 2018-11-16 presentations/codesSSRPday/Data2mergeMaize.R

@@ -1,0 +1,349 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+source("/its/home/mn301/foodSystems/Rcodes/Eq1ProdFun4/Data1ClimateShape.R")
+
+# or just load the data:
+
+load("Main/climate13.RData")
+load("Main/CrMaize14.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# first I need to separate counties into eastern and western
+
+zones<-read.csv(file='zonesHarvest/WestEast.csv')
+CrMaize15<-merge(CrMaize14,zones,all.x=TRUE,sort=FALSE)
+CrMaize15<-CrMaize15[order(CrMaize15["ID"],CrMaize15["Year"]),]
+
+CrMaize16<-droplevels(subset(CrMaize15, Year %in% 2000:2014, select=c("ID","county","Year","Yield","MT","Area","west1")))
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+CrMaize16$PreMean<-NA
+CrMaize16$PreMeanCV<-NA
+CrMaize16$PreMed<-NA
+CrMaize16$PreMedCV<-NA
+
+CrMaize16$PreMeanZ<-NA
[…]

### 2018-11-16 presentations/codesSSRPday/PlotsPresentationSeptember.R

@@ -1,0 +1,140 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+cor.test(CrMaize16$PreMed,CrMaize16$Yield)
+
+par(mar=c(4.2,4.2,3.5,1))
+plot(Yield~PreMed,data=CrMaize16,xlab="Average monthly precipitation, growing season",
+     ylab="Maize Yield, tonnes per hectare",main=c("+100 mm in average monthly precipitation -> + 0.78 t/hectare"),cex.lab=1.2)
+abline(lm (Yield~PreMed,data=CrMaize16),lwd=2)
+summary(lm (Yield~PreMed,data=CrMaize16))
+lines(c(156.59112,156.59112),c(1,1.77512),col="red",lwd=3,lty=2)
+lines(c(56.59112,156.59112),c(1,1),col="red",lwd=3,lty=2)
+
+legend(160,1.8,legend = c("0.78 tonnes/ hectare"),text.col=c("red")
+       ,bg="beige",box.col="white",  cex=1)
+legend(100,0.90,legend = c("100mm"),bg = "beige" ,box.col="white",pch=NA,
+       text.col=c("red"))
+
[…]

### 2018-11-16 presentations/codesSSRPday/forPresentation.R

@@ -1,0 +1,61 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+Keith7s<-lmer(Yield~scale(PreMed) + I(scale(PreMed)^2)+ scale(TemMed) +I(scale(TemMed)^2) + scale(PreMedCV)+I(scale(PreMedCV)^2) +
+                scale(TemMedCV) +I(scale(TemMedCV)^2) 
+              +(scale(PreMed)+I(scale(PreMed)^2)  + scale(TemMed) +I(scale(TemMed)^2) |ID),data=CrMaize16ts)
+summary(Keith7s) 
+
+# omg. realized that this is nonsense as scaled precipitation can be negative>>adding squares is a bit nonsense..
+#----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+Astrid1<-lmer(log(Yield)~scale(PreMed) + I(scale(PreMed,center=FALSE)^2)+ scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + scale(PreMedCV)+
+                scale(TemMedCV)  
+              +(scale(PreMed)+I(scale(PreMed,center=FALSE)^2)
[…]

### 2018-11-16 presentations/codesSSRPday/forPresentation2.R

@@ -1,0 +1,50 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+
+Rachel3<-lmer(log(Yield)~PreMedZ + I(PreMedZ^2)+ TemMedZ +I(TemMedZ^2) + PreMedCVz+I(PreMedCVz^2) +
+                TemMedCVz +I(TemMedCVz^2)  +(PreMedZ+I(PreMedZ^2)  + TemMedZ +I(TemMedZ^2) |ID),data=CrMaize16ts)
+summary(Rachel3)
+
+R3step<-step(Rachel3)
+R3step1<-get_model(R3step) 
+summary(R3step1)
+ranef(R3step1)
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+
+
+
+
+hist(CrMaize16ts$PreMedz)
+
+summary(CrMaize16$PreMedz)
+
+
+CrMaize16ts[CrMaize16ts$PreMedZ<(-1),]
+
+CrMaize16ts[CrMaize16ts$PreMedZ<(-1),]

### 2018-11-16 presentations/codesSSRPday/forPresentation6.R

@@ -1,0 +1,59 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+
+Astrid2<-lmer(Yield~scale(PreMed) +  scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + scale(PreMedCV)+
+              +(scale(PreMed)
+                + scale(TemMed) |ID),data=CrMaize16ts)
+summary(Astrid2) 
+
+A2step<-step(Astrid2)
+A21<-get_model(A2step) 
+summary(A21)
+
+
+A22<-lmer(Yield~scale(PreMed) +  scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + (scale(TemMed) |ID),data=CrMaize16ts)
+summary(A22) 
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
[…]

### 2018-11-16 presentations/codesSSRPday/forPresentation6a.R

@@ -1,0 +1,84 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+Keith7s<-lmer(Yield~scale(PreMed) + I(scale(PreMed)^2)+ scale(TemMed) +I(scale(TemMed)^2) + scale(PreMedCV)+I(scale(PreMedCV)^2) +
+                scale(TemMedCV) +I(scale(TemMedCV)^2) 
+              +(scale(PreMed)+I(scale(PreMed)^2)  + scale(TemMed) +I(scale(TemMed)^2) |ID),data=CrMaize16ts)
+summary(Keith7s) 
+
+
+# omg. realized that this is nonsense as scaled precipitation can be negative>>adding squares is a bit nonsense..
+#----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+Astrid1<-lmer(Yield~scale(PreMed) + I(scale(PreMed,center=FALSE)^2)+ scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + scale(PreMedCV)+
+                scale(TemMedCV)  
+              +(scale(PreMed)+I(scale(PreMed,center=FALSE)^2)
[…]

### 2018-11-16 presentations/codesSSRPday/forPresentation6b.R

@@ -1,0 +1,144 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+
+A22<-lmer(Yield~scale(PreMed) +  scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + (scale(TemMed) |ID),data=CrMaize16ts)
+summary(A22) 
+ranef(A22) 
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+hist(CrMaize16ts$PreMed)
+hist(scale(CrMaize16ts$PreMed))
[…]

### 2018-11-16 presentations/codesSSRPday/forPresentation7.R

@@ -1,0 +1,146 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+
+A22<-lmer(Yield~scale(PreMed) +  scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + (scale(TemMed) |ID),data=CrMaize16ts)
+summary(A22) 
+ranef(A22) 
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+hist(CrMaize16ts$PreMed)
+hist(scale(CrMaize16ts$PreMed))
[…]

### 2018-11-16 presentations/codesSSRPday/forPresentationPokus.R

@@ -1,0 +1,61 @@
+rm(list=ls())
+
+WDuni<-c("/its/home/mn301/foodSystems/dataFS") # uni
+WDhome<-c("/home/trennion/foodSystems/dataFS") # doma
+
+setwd(WDuni)
+setwd(WDhome)
+
+library('dplyr')
+library('tseries')
+library(plm)
+library(lme4)
+library(lattice)
+library(car)
+library(lmerTest)
+
+
+
+load("Main/CrMaize16.RData")
+CrMaize16ts<-pdata.frame(CrMaize16,index=c("ID","Year"))
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+Keith7s<-lmer(Yield~scale(PreMed) + I(scale(PreMed^2))+ scale(TemMed) +I(scale(TemMed)^2) + scale(PreMedCV)+I(scale(PreMedCV)^2) +
+                scale(TemMedCV) +I(scale(TemMedCV)^2) 
+              +(scale(PreMed)+I(scale(PreMed^2))  + scale(TemMed) +I(scale(TemMed)^2) |ID),data=CrMaize16ts)
+summary(Keith7s) 
+
+# omg. realized that this is nonsense as scaled precipitation can be negative>>adding squares is a bit nonsense..
+#----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+
+Astrid1<-lmer(log(Yield)~scale(PreMed) + I(scale(PreMed,center=FALSE)^2)+ scale(TemMed) 
+              +I(scale(TemMed,center=FALSE)^2) 
+              + scale(PreMedCV)+
+                scale(TemMedCV)  
+              +(scale(PreMed)+I(scale(PreMed,center=FALSE)^2)
[…]

### 2018-11-16 writing/draft1/draftMonika9.tex
working hardly
writing bullet points of the future paper..

--- 2018-11-15 writing/draft1/draftMonika8.tex
@@
 \subsection{Data}
-\textcolor{blue}{source of data, resolution and plots/barplots of yield, precip and temp when final data available.}\\
-\textcolor{red}{maybe a table of descriptive statistics or ref to the tale to appendix}
-\textcolor{blue}{also how the weather data were aggregated spatially over counties}
+\begin{itemize}
+\item \textcolor{blue}{Source of data}
+\item \textcolor{blue} {Verbal description of the data, resolution, period, frequency}
+\item \textcolor{blue} {Plot(s)/barplots of yield, precip and temp when final data available}
+\item \textcolor{blue}{Maybe a table of descriptive statistics or ref to the tale to appendix}
+\item \textcolor{blue}{Description how the weather data were aggregated spatially over counties}
+\item \textcolor{red}{Description how the weather data were aggregated temporally - seasons}
+\end{itemize}
 \subsection{Linear mixed models}
@@
 
-\large Findings:
+\color{blue}
 \begin{itemize}
-\item OND last year dry spell, max rain very important for Maize, but cumulative precipitation for the same period not so important
 
-\item Mar-Sept last year temperature very important for maize yields
-\item SD temperature last year positive and significant
-\item dry spell 20 MAM last year important (but not dry spell MAM10)
-\item interesting. Precipitation 2 months MAM last year very significant and positive
-\item mean temp last year negative and significant, hill shaped 
+\item Tables of estimates of:
+
+\begin{enumerate}
+\item Basic model which only includes seasonal precipitation and average seasonal temperature
+\item The preferred specification which includes the significant weather aggregates (see the preliminary results in Table \ref{KenARe11})
+
+\begin{itemize}
+\item For all counties and also separately for Arid or semi-araid (ASAL) and non-ASAL
 \end{itemize}
 
-\underline{New findings:}
+\end{enumerate}
+\item Verbal description and interpretation of the results in table
+\item Discussion and Conclusion:
+	\begin{itemize}
+	\item Stressing the importance of not only focusing on prediction of rain totals, but also other characteristics of the weather - particularly the variables found significant
+	\end{itemize}
+	\item \textcolor{red}{Caveats}
+\end{itemize}
+\color{black}
 
-\begin{itemize}
-\item The yields seem to be more responsive to weather on west than on east
-\end{itemize}
-\normalsize
 \FloatBarrier
@@
 
-{\hspace{-1.5cm}
-\begin{threeparttable}
-
-\singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(0,0) errors}}
-% AmBEST00, AmBEST, AmBEST99
-
-\label{Models1}
-
-\begin{small}
-\begin{tabular}{lcrcrcr}
-
-\hline \vspace{-0.2cm} \\
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{Model 1}} &\multicolumn{2}{c}{\textit{Model 2}} &\multicolumn{2}{c}{\textit{Model 3}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&Estimate&Sum sq.&Estimate&Sum sq.&Estimate&Sum sq.\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Intercept&$0.298$ $(0.066)^{***}$&---&$0.256$ $(0.061)^{***}$&---&$0.243$ $(0.076)^{**}$&---\\
-  \\
-\vspace{-0.2cm}Prec. total&$0.289$ $(0.045)^{***}$&$11.9$&$0.131$ $(0.046)^{**}$&$1.3$&$0.137$ $(0.125)^{}$&$0.3$\\
-  \\
-  \vspace{-0.2cm}Prec. total sq.&$-0.071$ $(0.028)^{*}$&$1.8$&$-0.075$ $(0.026)^{**}$&$1.5$&$-0.073$ $(0.039)^{\bullet}$&$1.0$\\
-  \\
-  \vspace{-0.2cm}Temp. - average&$-0.202$ $(0.062)^{**}$&$3.1$&$-0.188$ $(0.058)^{**}$&$2.9$&$-0.464$ $(0.144)^{**}$&$2.9$\\
-  \\
-    \vspace{-0.2cm}Prec. $\times$ temp.&---&---&$-0.135$ $(0.037)^{***}$&$2.5$&$-0.108$ $(0.051)^{*}$&$1.3$\\
-  \\
-      \vspace{-0.2cm}Prec. c. of variation&---&---&$-0.198$ $(0.035)^{***}$&$7.6$&$-0.181$ $(0.075)^{*}$&$1.7$\\
-  \\
-        \vspace{-0.2cm}Temp. std. dev.&---&---&$0.064$ $(0.021)^{**}$&$1.3$&$0.058$ $(0.027)^{*}$&$1.3$\\
-  \\
-          \vspace{-0.2cm}Prec. first 2 months&---&---&---&---&$-0.003$ $(0.040)^{}$&$0.001$\\
-  \\
-         \vspace{-0.2cm}Dry spell -length&---&---&---&---&$-0.050$ $(0.031)^{}$&$0.7$\\
-  \\
-        \vspace{-0.2cm}Dry spells $>9$ days&---&---&---&---&$0.018$ $(0.032)^{}$&$0.09$\\
-  \\
-          \vspace{-0.2cm}Max precipitation&---&---&---&---&$-0.008$ $(0.064)^{}$&$0.005$\\
-  \\
-          \vspace{-0.2cm}Degree days&---&---&---&---&$-0.016$ $(0.092)^{}$&$0.009$\\
-  \\
-          \vspace{-0.2cm}Heat wave d. ($>35$\degree C)&---&---&---&---&$0.032$ $(0.039)^{}$&$0.2$\\
-  \\
-          \vspace{-0.2cm}Max temperature&---&---&---&---&$0.278$ $(0.136)^{*}$&$1.2$\\
-  \\
-  \hline
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
-\vspace{-0.2cm}
-\\
-\hline
-\\
-  \vspace{-0.2cm}Intercept\\
-  \vspace{-0.1cm} \\ 
-  \hline
-    \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{AIC:}}&$2114.98$&&$2144.88$&&$2118.15$\\
-
-\hline
-\vspace{-0.2cm}
-\end{tabular} 
-\end{small}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
-        \begin{adjustwidth}{1cm}{} 
-    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\pagebreak
-
-{\hspace{-1.5cm}
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(2,1) errors}}
-%AmArE2100, AmArE21
-\label{ARMA21} 
-\begin{footnotesize}
-
-\begin{tabular}{lcrcrcr}
-
-\hline \vspace{-0.2cm} \\
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{Model 1}} &\multicolumn{2}{c}{\textit{Model 2}} &\multicolumn{2}{c}{\textit{Model 3}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&Estimate&F-value&Estimate&F-value&Estimate&F-value\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Intercept&$0.299$ $(0.064)^{***}$&$13.9$&$0.240$ $(0.064)^{***}$&$14.0$&$0.270$ $(0.074)^{***}$&$13.8$\\
-  \\
-\vspace{-0.2cm}Prec. total&$0.273$ $(0.042)^{***}$&$63.6$&$0.159$ $(0.051)^{**}$&$64.7$&$0.219$ $(0.117)^{\bullet}$&$65.2$\\
-  \\
-  \vspace{-0.2cm}Prec. total sq.&$-0.066$ $(0.026)^{*}$&$9.9$&$-0.079$ $(0.028)^{**}$&$10.1$&$-0.094$ $(0.037)^{*}$&$10.2$\\
-  \\
-  \vspace{-0.2cm}Temp. - average&$-0.177$ $(0.061)^{**}$&$8.4$&$-0.191$ $(0.062)^{**}$&$11.8$&$-0.380$ $(0.133)^{**}$&$8.7$\\
-  \\
-    \vspace{-0.2cm}Prec. $\times$ temp.&---&---&$-0.135$ $(0.037)^{**}$&$4.6$&$-0.101$ $(0.049)^{*}$&$7.5$\\
-  \\
-      \vspace{-0.2cm}Prec. c. of var.&---&---&$-0.139$ $(0.040)^{***}$&$11$&$-0.091$ $(0.070)^{}$&$11.1$\\
-  \\
-        \vspace{-0.2cm}Temp. std. dev.&---&---&$0.038$ $(0.024)^{}$&$2.4$&$0.038$ $(0.026)^{}$&$2.4$\\
-  \\
-          \vspace{-0.2cm}Prec. 2 months&---&---&---&---&$-0.009$ $(0.037)^{}$&$0.3$\\
-  \\
-         \vspace{-0.2cm}Dry spell -length&---&---&---&---&$-0.045$ $(0.030)^{}$&$2.2$\\
-  \\
-        \vspace{-0.2cm}Dry spells $>9$ days&---&---&---&---&$0.008$ $(0.031)^{}$&$0.1$\\
-  \\
-          \vspace{-0.2cm}Max precipitation&---&---&---&---&$-0.035$ $(0.060)^{}$&$0.2$\\
-  \\
-          \vspace{-0.2cm}Degree days&---&---&---&---&$-0.015$ $(0.088)^{}$&$0.6$\\
-  \\
-          \vspace{-0.2cm}Heat wave d.&---&---&---&---&$0.040$ $(0.037)^{}$&$3.1$\\
-  \\
-          \vspace{-0.2cm}Max temperature&---&---&---&---&$0.207$ $(0.123)^{\bullet}$&$2.8$\\
-  \\
-  \hline
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
-\vspace{-0.2cm}
-\\
-\hline
-\\
-  \vspace{-0.2cm}Intercept\\
-  \vspace{-0.1cm} \\ 
-  \hline
-  \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{AIC:}}&$2064.54$&&$2064.99$&&$2099.90$\\
-
-\hline
-\vspace{-0.2cm}
-\end{tabular} 
-\end{footnotesize}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
-        \begin{adjustwidth}{1cm}{} 
-    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\pagebreak
-
 %TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT

### 2018-11-16 writing/draft1/draftMonika10.tex
workiing more and more
writing bullet points

--- 2018-11-16 writing/draft1/draftMonika9.tex
@@
 \FloatBarrier
-	\section{Results and discussion}\label{Results}
+	\section{\textcolor{black}{Results and discussion}}\label{Results}
 
@@
 
-\item Tables of estimates of:
+\item Tables of estimates:
 
@@
 \item Discussion and Conclusion:
+
+	\end{itemize}
+\section{Conclusion}\label{Conclusion}
+
 	\begin{itemize}
-	\item Stressing the importance of not only focusing on prediction of rain totals, but also other characteristics of the weather - particularly the variables found significant
+	\item Besides seasonal totals (averages), there are many weather/climate characteristics (measures) which are important for yield
+	\item  Current weather and climate forecasts (in the context of disaster management) are mostly focused on totals in precipitation (and average temperature?) 
+	\item There is a strong movement to shift the focus from forecasting merely weather/climate to forecast of weather/climate and its impacts
+
+	\begin{itemize}
+	\item[$\boldsymbol{\rightarrow}$] We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models
 \end{itemize}
+
 	\item \textcolor{red}{Caveats}
+
+	\begin{itemize}
+	\item\textcolor{red}{ Yield data: Is the collection method consistent over time and space?}
+	\item Livestock
+	\item Cropping calendar is likely to differ across the country
+
+	\end{itemize}
 \end{itemize}

### 2018-11-16 writing/draft1/draftMonika10.tex
writing

@@
 \begin{enumerate}
-\item Basic model which only includes seasonal precipitation and average seasonal temperature
+\item Base model which only includes seasonal precipitation and average seasonal temperature
 \item The preferred specification which includes the significant weather aggregates (see the preliminary results in Table \ref{KenARe11})
@@
 \end{enumerate}
-\item Verbal description and interpretation of the results in table
-\item Discussion and Conclusion:
+\item Verbal description and interpretation of the results
+\item Estimates of future impacts
+
+\begin{itemize}
+\item Comparison of estimates accuracy of the predictions from the base model and the preferred specification
+
+		\begin{itemize}
+		\item choose two points in time (past).
+Imagine we are in the earlier of the two time points and we want to make a prediction of yields in the future point.
+Compare the predictions of the base model and the predictions of the preferred specification.
+Here, we assume that we know the forecast of the weather measures- variables which are used in the models.
+We can use the real weather data instead of the forecast.
+This will demonstrate why is it important to not only forecast totals/means but also other characteristics (measures) of weather/climate
+		\item Graphical comparison, possibly using barplots
 
 		\end{itemize}
-\section{Conclusion}\label{Conclusion}
+
+\end{itemize}
+
+	\end{itemize}
+\section{\textcolor{black}{Conclusion}}\label{Conclusion}
 
@@
 	\begin{itemize}
-	\item[$\boldsymbol{\rightarrow}$] We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models
+	\item[$\boldsymbol{\rightarrow}$] \textbf{We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models}
 \end{itemize}

### 2018-11-16 writing/draft1/draftMonika10.tex
writing more and more

--- 2018-11-16 writing/draft1/draftMonika10.tex
@@
 
-where $\mathbf{\Sigma}$ is a $q \times q$ positive semi-definite variance-covariance matrix
+where $\mathbf{\Sigma}$ is a $q \times q$ positive semi-definite variance-covariance matrix.
 
+\color{blue}
+\begin{itemize}
+\item Add description of residual and other diagnostic tests (AIC, VIF, autocorrelation)
+\item Describe the procedure that I have done to find the preferred way of modelling the  error correlation structure (ARMA errors).
+\end{itemize}
+\color{black}
 \FloatBarrier
@@
 
-\item Tables of estimates:
+\item Tables of estimates of two models:
 
 \begin{enumerate}
-\item Basic model which only includes seasonal precipitation and average seasonal temperature
-\item The preferred specification which includes the significant weather aggregates (see the preliminary results in Table \ref{KenARe11})
+\item The base model which only includes seasonal precipitation and average seasonal temperature
+\item The preferred specification which includes the significant weather variables (see the preliminary results in Table \ref{KenARe11})
 
 \begin{itemize}
-\item For all counties and also separately for Arid or semi-araid (ASAL) and non-ASAL
+\item For all counties 
+\item For Arid and semi-araid (ASAL) counties
+\item For non-ASAL counties
 \end{itemize}
@@
 \end{enumerate}
-\item Verbal description and interpretation of the results in table
-\item Discussion and Conclusion:
+\item Verbal description and interpretation of the results
+\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
+
+\begin{itemize}
+
+\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
+Now we want to make a prediction of yields for the later point in time.
+We will make two predictions: one using the base model and the other based on the preferred specification and we will compare their prediction accuracy.
+Here, we assume that we have a forecast of the weather measures (the explanatory variables of our models).
+We can use real weather data instead of the forecast.
+This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
+		\item Accompany with a graph(s), possibly barplot(s)
 
 \end{itemize}
-\section{Conclusion}\label{Conclusion}
+\item \textcolor{red}{Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
+This would give us an approximation of $R^2$ which could be used to compare our models.}
+
+		\begin{itemize}
+			\item \textcolor{red}{As Annemie has advised me, for a more precise approximation of $R^2$ we could estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
+This would give us a percentage of yield variability explained by the weather measures (not including the variability explained by the random intercepts).}
+	\end{itemize}
+		\end{itemize}
+
+\section{\textcolor{black}{Conclusion}}\label{Conclusion}
 
@@
 	\item Cropping calendar is likely to differ across the country 
+	\item \textcolor{red}{Some more??}
 

### 2018-11-16 writing/draft1/draftMonika10.tex
writing
I am writing draft and bullet points, working hard(ly)

@@
 
-\textcolor{blue}{maybe following into data section??}
-
 Based on the literature research above and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
@@
 \begin{itemize}
-\item Seasonal cumulative rainfall
-\item Seasonal standard deviation 
-\item Seasonal coefficient of variation 
+\item Cumulative precipitation and its square
+\item Cumulative precipitation during the first two months of season
+\item \textcolor{red}{I will also test cumulative precipitation  during the first month of each season - Guigma has promised to compute this variable}
+
+\item Coefficient of variation 
 \item Maximum length of dry spell in number of days
-\item ...\textcolor{blue}{the list to be finished}
-\item...
+\item Number of dry spell lasting for at least four days
+\item Number of dry spell lasting for at least ten days
+\item Number of dry spell lasting for at least twenty days
+\item Maximum daily precipitation
+\item \textcolor{red}{Additional indicators of floods - Guigma has promised to compute them over this weekend} 
 \end{itemize}
@@
 \begin{itemize}
-\item
-\item\textcolor{blue}{the list to be finished}..
-\item
+\item Average temperature and its square
+\item Degree days
+\item Heatwave days
 \end{itemize}
@@
 \item \textcolor{blue}{Description how the weather data were aggregated spatially over counties}
-\item \textcolor{red}{Description how the weather data were aggregated temporally - seasons}
+
 \end{itemize}
+
+\textcolor{blue}{\textbf{Description of how the weather data were aggregated temporally - seasons} (this maybe before the list of aggregates??)}
 \subsection{Linear mixed models}
@@
 \begin{itemize}
-\item Add description of residual and other diagnostic tests (AIC, VIF, autocorrelation)
-\item Describe the procedure that I have done to find the preferred way of modelling the  error correlation structure (ARMA errors).
+\item Add a description of residual and other diagnostic tests (AIC, VIF, autocorrelation)
+\item Describe the procedure that I have done to find the preferred way of modelling the  error correlation structure (ARMA errors)
 \end{itemize}

### 2018-11-17 Rcodes/November/KenARe11.R
writing and editing the draft

@@ -62,3 +62,25 @@ nobs(KEN11a_nonASAL); anova(KEN11a_nonASAL)
 exp(summary(KEN11a_nonASAL)$tTable[,1])
 exp(summary(KEN11a_nonASAL)$tTable[,2])
 anova(KEN11a_nonASAL, type='marginal')
+
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# now I have to remove residual outliers
+
+Kendalln3<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                +AvgTemp + SDTemp  + HWDays+(1|ID1),data=ScaledTS) 
+
+which(resid(Kendalln3,type="pearson")<(-6))
+which(resid(Kendalln3,type="pearson")<(-2))
+which(resid(KEN11a,type="pearson")<(-3))
+rem<-which(resid(KEN11a,type="pearson")<(-3))
+ScaledTS$Year[rem]
+
+[!(rownames(isdataScTS) %in% rem),]
+
+KEN16a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS[-rem,],na.action=na.exclude); summary(KEN16a)
+exp(summary(KEN16a)$tTable[,1])

### 2018-11-17 Rcodes/November/Ken_BEST_resTests.R

@@ -1,9 +1,7 @@
 rm(list=ls())
-setwd("foodSystems/dataFS") 
-setwd("dataFS") # home
 library(dplyr); library(tseries); library(plm); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 
-load("Main/DaTS.RData")
+load("dataFS/Main/DaTS.RData")
 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
@@ -26,8 +24,9 @@ sqrt(abs(resid(.)))
 qqmath(Kendalln3,id=0.05)
 
 plot(resid(Kendalln3,type="pearson"))
-which(resid(Kendalln3,type="pearson")<(-6))# yep, the same as in the previous case...
+which(resid(Kendalln3,type="pearson")<(-6))
+which(resid(Kendalln3,type="pearson")<(-2))
+# yep, the same as in the previous case...
 which(resid(AmBEST,type="pearson")<(-6))
-
 # test of AR(1) and MA(1) errors
 pbltest(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + SDTemp  + HWDays,data=ScaledTS,alternative='onesided')#ok, significant

### 2018-11-17 Rcodes/November/Models2Scaled.R

@@ -1,10 +1,10 @@
 rm(list=ls())
-setwd("foodSystems/dataFS") 
-# load("../Rcodes/octoberNew/Models1.RData")
+
+# load("Rcodes/octoberNew/Models1.RData")
 
 library(dplyr); library(tseries); library(plm); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 
-load("Main/DaTS.RData")
+load("dataFS/Main/DaTS.RData")
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
 Emmit0<-lm(Yield~SeasPr+AvgTemp,data=ScaledTS);  summary(Emmit0) 

### 2018-11-17 Rcodes/November/ReadData.R

@@ -1,5 +1,4 @@
 rm(list=ls())
-setwd("foodSystems/dataFS") 
 
 library('dplyr')
 library('purrr')
@@ -7,14 +6,14 @@ library("reshape")
 
 #ooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
-weath<-read.csv("climateNov/Seasonally.csv",na.strings="-999",dec=".")
+weath<-read.csv("dataFS/climateNov/Seasonally.csv",na.strings="-999",dec=".")
 summary(weath)
 testRe<-lm(AvgTemp_MarMay~.,data=weath)
 summary(testRe)
 nobs(testRe)  # grood, all loaded
 #----------------------------------------------------------------------------------------------------------------------------------------------------------------------
 # dry spell 4 in a separate file
-weath<-merge(weath,read.csv("climateNov/Dry_spell4_seasonsM.csv",na.strings="-999",dec=".") )
+weath<-merge(weath,read.csv("dataFS/climateNov/Dry_spell4_seasonsM.csv",na.strings="-999",dec=".") )
 weath<-weath[with(weath,order(code,year)),]
 testRe<-lm(AvgTemp_MarMay~.,data=weath)
 summary(testRe)
@@ -22,7 +21,7 @@ nobs(testRe)
 #----------------------------------------------------------------------------------------------------------------------------------------------------------------------
 # monthly also in separate file
 
-weath<-merge(weath,read.csv("climateNov/MonthlyM.csv",na.strings="-999",dec=".") )
+weath<-merge(weath,read.csv("dataFS/climateNov/MonthlyM.csv",na.strings="-999",dec=".") )
 weath<-weath[with(weath,order(code,year)),]
 
 #----------------------------------------------------------------------------------------------------------------------------------------------------------------------
@@ -40,10 +39,10 @@ setdiff(weather$code,weath$code)
 # ok, seems good. now I will have to match it with yield
 
 # but first merge with county codes
-IDdict<-read.csv("Main/IDdict.csv")
+IDdict<-read.csv("dataFS/Main/IDdict.csv")
[…]

### 2018-11-19 writing/draft1/draftMonika11.tex
draft writing
 bullet points what I will do in the future

--- 2018-11-16 writing/draft1/draftMonika10.tex
@@
 \title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
-\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens}
+\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens, Dominic Kniveton}
 \affil{\small{Department of Geography, University of Sussex, Falmer, UK}}
@@
 
-\item \textbf{Precipitation:}
+\item \textbf{\textbf{Precipitation}:}
 
@@
 \item Cumulative precipitation during the first two months of season
-\item \textcolor{red}{I will also test cumulative precipitation  during the first month of each season - Guigma has promised to compute this variable}
+\item \textcolor{blue}{I will test cumulative precipitation  during the first month of each season - Guigma has has just provided the aggregated variables}
 
@@
 \item Maximum daily precipitation
-\item \textcolor{red}{Additional indicators of floods - Guigma has promised to compute them over this weekend} 
+\color{blue}
+\item Additional indicators of floods:
+
+\begin{itemize}
+\item One of the Climate Extreme Indices described at \url{https://www.climdex.org/indices.html} is the sum of precipitation amount on the days where precipitation is above 99th percentile of precip.
+of the whole period (i.e.
+1981-2017 in our case).
+The 99th percentile is calculated based on the sample of wet days, that is the days where precipitation is above or equal to 1 mm (index no.
+26)
+\item Guigma has just sent me this measure and its varieties with thresholds 95th percentile and 90th percentile instead of the 99th percentile
 \end{itemize}
-\item Temperature
+\color{black}
+
+\end{itemize}
+
+\item \textbf{Temperature}
 \begin{itemize}
-\item Average temperature and its square
-\item Degree days
-\item Heatwave days
+\item Seasonal average (of daily average) and its square
+\item Average maximum daily temperature
+\item Standard deviation
+\item Cumulative degree days $10-30\degree$C
+\item Number of heatwave days $>35\degree$C
 \end{itemize}
 \end{itemize}
+
 \subsection{Data}

### 2018-11-19 writing/draft1/draftMonika11.tex
writing

@@
 We will make two predictions: one using the base model and the other based on the preferred specification and we will compare their prediction accuracy.
-Here, we assume that we have a forecast of the weather measures (the explanatory variables of our models).
+Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
 We can use real weather data instead of the forecast.
@@
 		\item Accompany with a graph(s), possibly barplot(s)
+\end{itemize}
 
-\end{itemize}
+\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately)
+
+\item Show a variant of the model without the outliers. This could be a table here, a table in the Appendix or a table in a footnote.
+
+\item If time, I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
+
 \item \textcolor{red}{Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
@@
 		\begin{itemize}
-			\item \textcolor{red}{As Annemie has advised me, for a more precise approximation of $R^2$ we could estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
-This would give us a percentage of yield variability explained by the weather measures (not including the variability explained by the random intercepts).}
+			\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
+This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
 	\end{itemize}

### 2018-11-19 writing/draft1/draftMonika11.tex
working writing

@@
 
-\date{\normalsize{October 2018}}
+\date{\normalsize{November 2018}}
 \title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
@@
 \noindent \textcolor{red}{..the full abstract to be written..} This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
-Our dataset consists of yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
+Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
 \\
@@
 
-Frequent occurrence of severe droughts have been a major problem in Kenya and in other Sub-Saharan countries \citep{Compendium, WorldBank2015, Nicholson2017}.
-The agricultural sector is the main contributor to the economy of Kenya, hence the effects of drought are especially damaging in this country \citep{MendelsohnDinar2000}.
-Besides compromising food security, drought usually leads to severe damage in water supply, electricity and the environment.
+Repeated severe droughts have been a major problem in Kenya and in other Sub-Saharan countries \citep{Compendium, WorldBank2015, Nicholson2017}.
+The agricultural sector is the main contributor to the economy of Kenya, hence the effects of droughts can be especially damaging in this country \citep{MendelsohnDinar2000}.
+Besides compromising food security, droughts usually lead to severe damage in water supply, electricity and the environment.
 In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
@@
 As a consequence of climate change, the situation is likely to get worse.
-Dry areas have strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
+Dry areas exhibit a strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
 Strong downwards trend in precipitation has already been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
@@
 \cite{RicardianBello} and  \cite{Ochieng2016} have found these results by the means of Ricardian analysis.
-\cite{Ochieng2016} have focused on Kenya while \cite{RicardianBello} addressed the situation in Niger.
+\cite{Ochieng2016} have focused on Kenya while \cite{RicardianBello} have addressed the situation in Niger.
 
@@
 \cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania.
-The authors have shown that by 2050 the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather.
+The authors have shown that by $2050$, the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather.
 \cite{Kabubo2015} have estimated effects of climate change on food security in Kenya.
@@
 Answering these and other similar questions should help to improve food security and shift the focus from reactive to proactive approach in drought disaster risk management.
-We focus on Kenya utilizing an yearly panel dataset of $47$ counties over the period of $1981-2017$.
+We focus on Kenya utilizing an yearly panel of $47$ counties over the period of $1981-2017$.
 % mozna hodit tuhle vetu na zacatek 2.1.
@@
 
-The main contribution of this study is utilization of the most detailed publicly available weather datasets while accounting for the real life crop calendar and different cropping seasons across the country as practised by Kenyan farmers.
-Furthermore, we believe that this is the first study which has described application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
+The main contribution of this study is leveraging the most detailed publicly available weather datasets while accounting for the real life crop calendar and different cropping seasons across the country as practised by Kenyan farmers.
+Furthermore, we believe that this is the first study which has described an application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
 Hence, this is the first time that effects of drought and weather on food security have been estimated taking into account the real crop calendar, spatially-heterogeneous nature of Kenya and allowing for the effects of weather to vary across the country.
@@
 
-Using food prices as a measure of food security turned out be problematic as well. 
+Using food prices as a measure of food security turned out to be problematic as well. 
 One problem is data availability as the food prices are usually available for market towns but not as county averages.
@@
 Another problem is that prices are usually strongly spatially autocorrelated and this could cause the regression estimates to be inconsistent and inefficient unless the model accounts for the spatial autocorrelation structure.
-Modelling the structure of spatial autocorrelation would require developing and estimation of complicated structural equations and also additional data, which may not be available.
+Modelling the structure of spatial autocorrelation would require developing and estimation of complicated model consisting of several structural equations and also additional data, which may not be available.
 Therefore we opted for agricultural crop yields as a measure of food security and we decided to estimate a single reduced form equation with crop yields as the dependent variable.

### 2018-11-19 writing/draft1/draftMonika11.tex
writing

@@
 Another problem is that prices are usually strongly spatially autocorrelated and this could cause the regression estimates to be inconsistent and inefficient unless the model accounts for the spatial autocorrelation structure.
-Modelling the structure of spatial autocorrelation would require developing and estimation of complicated model consisting of several structural equations and also additional data, which may not be available.
+Modelling the structure of spatial autocorrelation would require developing and estimation of a complicated model consisting of several structural equations and additional data, which may not be available.
 Therefore we opted for agricultural crop yields as a measure of food security and we decided to estimate a single reduced form equation with crop yields as the dependent variable.
@@
 
-The other important question to answer before starting our analysis was: `How is drought defined and which definition and measure of drought would be the best for the purpose of this study?' Various definitions of drought and their role have been reviewed by \cite{wilhite1985} and \cite{wilhite2000}.
+The other important question to answer before starting our analysis was: `How is drought defined and which definition and measure of drought is the most suitable for the purpose of this study?' Various definitions of drought and their role have been reviewed by \cite{wilhite1985} and \cite{wilhite2000}.
 For an extensive overview of drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
@@
 We have considered the Standardised Precipitation Evapotranspiration Index  (SPEI) and the vegetation condition index (VCI) as measures of drought in our analysis.
-However, we did not find the county level SPEI to be as good predictor of food security as other precipitation and temperature measures and computing the SPEI for the entire high resolution weather dataset would be inadequately demanding.
+We did not find the county level SPEI to be as good predictor of food security as other precipitation and temperature measures and computing the SPEI for the entire high resolution weather dataset would be inadequately demanding.
 We did not use the VCI either as we were not confident enough about the reliability of the available data.\footnote{The values of the VCI which we obtained from the NDMA were not the same as those available by the \cite{BOKU} (BOKU).
-Furthermore, the VCI values did not exactly correspond to the early warning phases as defined by the \cite{NDMA} (NDMA).}
+Furthermore, the VCI values did not exactly correspond to the early warning phases although they should according to the \cite{NDMA} (NDMA).}
 

### 2018-11-19 writing/draft1/draftMonika11.tex
writting more

@@
 
-Thus, after conducting a rigorous literature research on measures of drought previously used in the relevant context, we decided to utilize various aggregates of daily precipitation and temeprature.
+After conducting a rigorous literature research on measures of drought previously used in the relevant context, we decided to utilize various aggregates of daily precipitation and temeprature.
 The frequency of the maize yield data is yearly while the weather data are daily.
-Hence, the weather data needed to be aggregated in order to obtain a dataset conformable with the yield data.
+Hence, the weather data had to be aggregated in order to obtain a dataset conformable with the yield data.
 There are many possible ways of aggregation.
@@
 
-An important group of studies that have analysed relationship of yield and climate in Sub-Saharan Africa has utilised degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or number of days with temperature above certain level or within defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
+An important group of studies focused on relationship of yield and climate in Sub-Saharan Africa has been focused on degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or on the number of days with temperature above certain level or within a defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
 
-\textcolor{red}{Erin Lentz, can we cite her paper?She has not responded to my email}
+%\textcolor{red}{Erin Lentz, can we cite her paper?She has not responded to my email}
 
@@
 \item Cumulative precipitation during the first two months of season
-\item \textcolor{blue}{I will test cumulative precipitation  during the first month of each season - Guigma has has just provided the aggregated variables}
+\item \textcolor{blue}{Additionally, I will test significance of the cumulative precipitation  during the first month of each season - Guigma has just provided the aggregated variables}
 
@@
 \item Maximum length of dry spell in number of days
-\item Number of dry spell lasting for at least four days
-\item Number of dry spell lasting for at least ten days
-\item Number of dry spell lasting for at least twenty days
+\item Number of dry spells lasting for at least four days
+\item Number of dry spells lasting for at least ten days
+\item Number of dry spells lasting for at least twenty days
 \item Maximum daily precipitation
@@
 \end{itemize}
+\textcolor{blue}{\textbf{Description of how the weather data were aggregated temporally - seasons}}
 
-\subsection{Data}
-\begin{itemize}
-\item \textcolor{blue}{Source of data}
-\item \textcolor{blue} {Verbal description of the data, resolution, period, frequency}
-\item \textcolor{blue} {Plot(s)/barplots of yield, precip and temp when final data available}
-\item \textcolor{blue}{Maybe a table of descriptive statistics or ref to the tale to appendix}
-\item \textcolor{blue}{Description how the weather data were aggregated spatially over counties}
-
-\end{itemize}
-
-\textcolor{blue}{\textbf{Description of how the weather data were aggregated temporally - seasons} (this maybe before the list of aggregates??)}
 \subsection{Linear mixed models}
@@
 As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
-Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
+Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
 Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset.
@@
 Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}.
-On the other hand, random effect are unobserved random variables.
+On the other hand, random effects are unobserved random variables.
 There are at least three benefits of treating a set of parameters as a random sample from some distribution.
@@
 \item Add a description of residual and other diagnostic tests (AIC, VIF, autocorrelation)
-\item Describe the procedure that I have done to find the preferred way of modelling the  error correlation structure (ARMA errors)
+\item Describe the procedure that I have done to find the preferred way of modelling the  correlation structure in errors (ARMA errors)
 \end{itemize}
@@
 \FloatBarrier
+
+\subsection{Data}
+\begin{itemize}
+\item \textcolor{blue}{Source of data}
+\item \textcolor{blue} {Verbal description of the data, resolution, period, frequency}
+\item \textcolor{blue} {Plot(s)/barplots of yield, precip. and temp.}
+\item \textcolor{blue}{Maybe a table of descriptive statistics or ref. to a table in the Appendix}
+\item \textcolor{blue}{Description how the weather data were aggregated spatially over counties}
+\end{itemize}
+
 	\section{\textcolor{black}{Results and discussion}}\label{Results}
@@
 
-\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately)
+\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss
 
-\item Show a variant of the model without the outliers. This could be a table here, a table in the Appendix or a table in a footnote.
+\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
 
@@
 
-\item \textcolor{red}{Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
-This would give us an approximation of $R^2$ which could be used to compare our models.}
+\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
+This would give us an approximation of $R^2$ which could be used to compare our models.
 
@@
 
-\section{\textcolor{black}{Conclusion}}\label{Conclusion}
-
-	\begin{itemize}
-	\item Besides seasonal totals (averages), there are many weather/climate characteristics (measures) which are important for yield
-	\item  Current weather and climate forecasts (in the context of disaster management) are mostly focused on totals in precipitation (and average temperature?) 
-	\item There is a strong movement to shift the focus from forecasting merely weather/climate to forecast of weather/climate and its impacts
-
-	\begin{itemize}
-	\item[$\boldsymbol{\rightarrow}$] \textbf{We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models}
-\end{itemize}
-
-	\item \textcolor{red}{Caveats}
-
-	\begin{itemize}
-	\item\textcolor{red}{ Yield data: Is the collection method consistent over time and space?}
-	\item Livestock
-	\item Cropping calendar is likely to differ across the country 
-	\item \textcolor{red}{Some more??}
-
-	\end{itemize}
-\end{itemize}
 \color{black}
-
-\FloatBarrier
-\pagebreak
 
@@
 
+		\color{blue}
+\section{\textcolor{black}{Conclusion}}\label{Conclusion}
+
+	\begin{itemize}
+	\item There are many weather/climate characteristics (measures) which are important for yield, not only seasonal totals (averages)
+	\item  Current weather and climate forecasts (in the context of disaster management) are mostly focused on totals in precipitation (and average temperature?) 
+	\item There is a strong movement to shift the focus from forecasting merely weather/climate to forecast of weather/climate and its impacts
+
+	\begin{itemize}
+	\item[$\boldsymbol{\rightarrow}$] \textbf{We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models}
+\end{itemize}
+
+	\item \textcolor{red}{Caveats}
+
+	\begin{itemize}
+	\item\textcolor{red}{ Yield data: Is the collection method consistent over time and space?}
+	\item Livestock
+	\item Cropping calendar is likely to differ across the country 
+	\item \textcolor{red}{Some more??}
+
+	\end{itemize}
+\end{itemize}
+\color{black}
+
+\FloatBarrier
+\pagebreak
+
 %TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT

### 2018-11-21 writing/draft1/draftMonika12.tex
writing - adjusting
and other working

--- 2018-11-19 writing/draft1/draftMonika11.tex
@@
 \item Add a description of residual and other diagnostic tests (AIC, VIF, autocorrelation)
-\item Describe the procedure that I have done to find the preferred way of modelling the  correlation structure in errors (ARMA errors)
+\item Describe the procedure that I have applied to find the preferred combination of fixed effects and random effects
+\item Describe the procedure that I have applied to find the preferred way of modelling the  correlation structure in errors (ARMA errors)
 \end{itemize}
@@
 \end{itemize}
-
 \end{enumerate}
-\item Verbal description and interpretation of the results
-\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
-
-\begin{itemize}
-
-\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
-Now we want to make a prediction of yields for the later point in time.
-We will make two predictions: one using the base model and the other based on the preferred specification and we will compare their prediction accuracy.
-Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
-We can use real weather data instead of the forecast.
-This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
-		\item Accompany with a graph(s), possibly barplot(s)
-\end{itemize}
-
-\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss
-
-\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
-
-\item If time, I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
-
-\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
-This would give us an approximation of $R^2$ which could be used to compare our models.
-
-		\begin{itemize}
-			\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
-This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
-	\end{itemize}
-		\end{itemize}
 
@@
 \caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
-%KEN11a  KEN11a_ASAL   
+%KEN11a  KEN11a_ASAL    KEN11a_nonASAL														
 \label{KenARe11} 
 \begin{footnotesize}
-
 \begin{tabular}{llrlllr} 
-
 \hline \vspace{-0.2cm} \\
@@
 \color{blue}
+
+\item Verbal description and interpretation of the results
+\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
+
+\begin{itemize}
+
+\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
+Now we want to make a prediction of yields for the later point in time.
+We will make two predictions: one using the base model and the other based on the preferred specification and we will compare their prediction accuracy.
+Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
+We can use real weather data instead of the forecast.
+This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
+		\item Accompany with a graph(s), possibly barplot(s)
+\end{itemize}
+
+\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss.
+
+\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
+
+\item If time, I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
+
+\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
+This would give us an approximation of $R^2$ which could be used to compare our models.
+
+		\begin{itemize}
+			\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
+This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
+	\end{itemize}
+		\end{itemize}
+
+		\color{blue}
 \section{\textcolor{black}{Conclusion}}\label{Conclusion}
@@
 	\item\textcolor{red}{ Yield data: Is the collection method consistent over time and space?}
-	\item Livestock
+	\item Livestock not explicitly taken into account
 	\item Cropping calendar is likely to differ across the country 

### 2018-11-21 Rcodes/November/Models7ASAL_ranEf.R

@@ -16,7 +16,7 @@ load("Main/DaTS.RData")
                        + SDTemp + DDays + HWDays+MaxT+(I(Prec2m^2)+Prec2m|ASAL),data=ScaledTS) 
           summary(Shiv14); vif(Shiv14); ranova(Shiv14)
                           #but worse vif fo Shiv14...maybe search continues in another script..
-          #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
           
 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo

### 2018-11-21 writing/draft1/draftMonika12.tex
working hard
adding new flood indicators and first months rainfall

@@
 \vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{All counties}} &\multicolumn{2}{c}{\textit{ASAL}} &\multicolumn{2}{c}{\textit{non-ASAL}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&$exp(\beta)$&F-value\tnote{a}%> anova(KEN11a, type='marginal')
-    &$exp(\beta)$&F-value\tnote{a}&$exp(\beta)$&F-value\tnote{a}\\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}%> anova(KEN11a, type='marginal')
+    &\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}&\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
  \hline 
@@
 \vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
-  \\
-\vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
+  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
   \\
   \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
+    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
+  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
+  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
+  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
   \\
-      \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
-  \\
-
-         \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
-  \\
-        \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
-  \\
-          \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
-  \\
-
-          \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
-  \\
-
   \hline
@@
 \vspace{-0.2cm} \\
-\textit{Number of observations:}  &$1300$&$698$&$602$ \\
+\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
 \vspace{-0.2cm}
@@
   \begin{footnotesize}
-    \item \textit{Notes:} Standard errors in brackets, \hspace{0.05cm}$1256$ observations
+    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
         \begin{adjustwidth}{1cm}{} 
-    \item \hspace{0.45cm}$^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$\\
     \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.

### 2018-11-22 Rcodes/NovemberNewVars/ReadNewData.R
Update ReadNewData.R

@@ -35,44 +35,27 @@ names(monthlyRa)[1]<-"Year"
 floods2<-floods; floods<-floods2[-c(3,4)]; floods$Year<-as.factor(floods$Year)
 monthlyRa2<-monthlyRa; monthlyRa<-monthlyRa2[-c(3,4)]; monthlyRa$Year<-as.factor(monthlyRa$Year)
 
-test1<-DaTS
+
 
 DaTS$code<-as.factor(DaTS$code)
-DaTS<-merge(DaTS,floods, all.x=TRUE)
-DaTS<-merge(DaTS,monthlyRa, all.x=TRUE)
-
-which(sapply(DaTS,function(x) is.factor(x)==TRUE))
-test99<-DaTS
-DaTS[[1]]<-as.character(DaTS[[1]])
-DaTS[[1]]<-factor(DaTS[[1]])
-
-
-i<-as.numeric(which(sapply(DaTS,function(x) is.factor(x)==TRUE)))
-
+test1<-DaTS
+DaTS<-merge(DaTS,floods, all.x=TRUE); DaTS<-merge(DaTS,monthlyRa, all.x=TRUE)
+DaTS<-DaTS[with(DaTS,order(ID1,Year) ),]
+# to fix malformed factors, I need to first convert factors to characters and then back to factors:
+test99<-DaTS;                  i<-as.numeric(which(sapply(DaTS,function(x) is.factor(x)==TRUE)))
 DaTS[as.numeric(which(sapply(DaTS,function(x) is.factor(x)==TRUE)))]<-lapply(DaTS[as.numeric(which(sapply(DaTS,function(x) is.factor(x)==TRUE)))],as.character)
 DaTS[i]<-lapply(DaTS[i],as.factor)
+# and test:
 all.equal(test99[-i], DaTS[-i]  )
+all_equal(test1,DaTS[1:124]);          all.equal(test1[-(2:3)],DaTS[c(1,4:124)],check.attributes=FALSE)
 
-
-summary(DaTS)
-head(test99[i]) head(DaTS[i])
+summary(DaTS); sum(complete.cases(DaTS))
 
 all.equal(test99[i],DaTS[i],check.attributes=FALSE)
 
 
-
-
-
-
-
-
-
-
-
-
-
 #----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
[…]

### 2018-11-28 writing/draft2/draftMonika12.tex
wirintg

--- 2018-11-21 writing/draft1/draftMonika12.tex
@@
 \date{\normalsize{November 2018}}
-\title{\Large \bf Relationship of Weather and Maize Yields in Kenya}
-\author{Monika Novackova, Pedram Rowhani, Martin Todd, Annemie Maertens, Dominic Kniveton}
+\title{\Large \bf The Effects of Weather on Maize Yields: New Evidence from Kenya}
+\author{Monika Novackova, Pedram Rowhani, Martin Todd, Dominic Kniveton}
 \affil{\small{Department of Geography, University of Sussex, Falmer, UK}}

### 2018-11-28 writing/draft1/draftMonika12.tex
hello
asd
redoing my draft, bullet points

@@
 
+\begin{itemize}
+\color{blue}
+\item The title is provisional, we will decide about the final title when most of the paper will be written
+\end{itemize}
+
 \begin{abstract}
@@
 
-\section{Introduction and literature review}\label{Introduction}
+\section{Introduction}\label{Introduction}
 
-Consequences of droughts and extreme heat waves on agriculture have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}).
-Agricultural drought and soil moister are important factors influencing plant health but they are also essential for land-atmospheric feedback and for temperature variability \citep{nicolai2017}.
+\begin{itemize}
+\color{blue}
+\item[] \textbf{Paragraph 1}
 
-Repeated severe droughts have been a major problem in Kenya and in other Sub-Saharan countries \citep{Compendium, WorldBank2015, Nicholson2017}.
-The agricultural sector is the main contributor to the economy of Kenya, hence the effects of droughts can be especially damaging in this country \citep{MendelsohnDinar2000}.
-Besides compromising food security, droughts usually lead to severe damage in water supply, electricity and the environment.
-In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
-Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
-As a consequence of climate change, the situation is likely to get worse.
-Dry areas exhibit a strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
-Strong downwards trend in precipitation has already been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
-During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
-Furthermore, increased frequency of heavy rain events has been observed in the areas with decline in total rainfall \citep{IPCCtrenberth}.
+\begin{itemize}
 
-Droughts and climate variability do not only affect agriculture but they have widespread consequences on food security in general.
-\cite{Ubilava2018} has found a connection between commodity prices and phases of the ENSO cycle.
-\cite{robinson2010} have used a Computable General Equilibrium (CGE) model for Ethiopia and they have shown that production shocks and resulting price increases have negative effects on farmers' income in drought-prone areas while moderate price increases can be beneficial for farmers in less drought-prone areas.
-\cite{OxfamIDS} have focused on sub-Saharan Africa, central America, North Africa and other low-income countries selected by Oxfam and he has concluded that climate change will result in decline in food productivity while food prices will increase.
-\cite{BROWN2015} have investigated food prices in $554$ local commodity markets in Africa, South Asia and Latin America and they concluded that almost 20\% of local market food prices had been affected by weather disturbances.
+\item Extreme weather causes disasters $\rightarrow$	 early warning systems have been developed
+\end{itemize}
+\item[] \textbf{Paragraph 2}
+\begin{itemize}
+\item What weather forecasts (measures) have been used in EWS? \textit{ref. litrature}
+\begin{itemize}
+\item Mostly seasonal precip. totals and temperature averages
+\end{itemize}
+\end{itemize}
 
-A large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa.
-Increase in temperature has typically been found to affect crop revenue negatively while increase in precipitation has mostly been associated with positive effects in the literature.
-\cite{RicardianBello} and  \cite{Ochieng2016} have found these results by the means of Ricardian analysis.
-\cite{Ochieng2016} have focused on Kenya while \cite{RicardianBello} have addressed the situation in Niger.
+\item[] \textbf{Paragraph 3} 
+\begin{itemize}
 
-Other studies have investigated the relationship of climate and yields in the context of climate change.
-\cite{MendelsohnDinar2000} have developed a simulation model to evaluate impacts of climate change on African agriculture.
-According to their study, agricultural GDP and climate change exhibit a hill-shaped relationship.
-Hence, global warming should be beneficial in high latitudes but harmful in low latitudes.
-The estimated loss of the Africa as a continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}.
-\cite{kabubo2007} have shown that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
-Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}.
-\cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa.
-They argued that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
-\cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania.
-The authors have shown that by $2050$, the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather.
-\cite{Kabubo2015} have estimated effects of climate change on food security in Kenya.
-They have found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped.
-Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
-\cite{Mendelsohn2008} has reviewed several studies on effects of climate change on agriculture in developing countries.
-According to the literature review, the tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture \citep{Mendelsohn2008}.
-\cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia.
-The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions.
-According to a meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming.
-However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.
+\item Identify difference between hazard and disaster
 
-A number of important studies have focused on responses to drought related disasters or emergency situations.
-\cite{Sandstorm2017} have questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
-They have analysed causes of the food crisis identified by the humanitarian appeal documents.
-Majority of these documents have found food availability and food production to be the major causes of food crisis.
-\cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.
-According to \cite{Sandstorm2017}, a large share of humanitarian response budget has been focused on emergency food aid in comparison to the share focused on interventions to build resilience.
-The authors have suggested to increase the budget share focused on agricultural and livestock production to build resilience.
-Also other authors have promoted shifting from reactive to proactive approach in disaster risk management (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).
-\cite{Nicholson2017} has recommended forecast based financing as it can avoid significant disaster losses.
-\cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses.
-Also \cite{IPCC2012ch1} have argued that risk assessment is essential for disaster risk management.
+\begin{itemize}
+\item Not every hazard turns into disaster
+\item For a hazard to become a disaster it needs to have \textbf{impact}
+\item Here, we identify the key metrics which have impact on yield
+\end{itemize}
+
+\end{itemize}
+
+\item[] \textbf{Paragraph 4}
+
+\begin{itemize}
+\item Crop yield versus climate forecasting
+\end{itemize}
+ \item[] \textbf{Paragraph 5}
+
+\begin{itemize}
+\item Aim of the paper
+\end{itemize}
+
+\end{itemize}
+\color{black}
 
@@
 
-	\section{\textcolor{black}{Results and discussion}}\label{Results}
+	\section{Results}\label{Results}
 
@@
 
-		\color{blue}
-\section{\textcolor{black}{Conclusion}}\label{Conclusion}
+\section{Discussion}\label{Discussion}
 

### 2018-11-28 writing/draft2/draftMonika12.tex
draft bullets

@@
 
-\section{Introduction and literature review}\label{Introduction}
+\section{Introduction}\label{Introduction}
 
-Consequences of droughts and extreme heat waves on agriculture have been well documented (\citealp{Deschenes2007Ric,RicardianBello,Lesk2016,Mehrabi2017, schwalm2017}).
-Agricultural drought and soil moister are important factors influencing plant health but they are also essential for land-atmospheric feedback and for temperature variability \citep{nicolai2017}.
+\begin{itemize}
+\color{blue}
+\item[] \textbf{Paragraph 1}
 
-Repeated severe droughts have been a major problem in Kenya and in other Sub-Saharan countries \citep{Compendium, WorldBank2015, Nicholson2017}.
-The agricultural sector is the main contributor to the economy of Kenya, hence the effects of droughts can be especially damaging in this country \citep{MendelsohnDinar2000}.
-Besides compromising food security, droughts usually lead to severe damage in water supply, electricity and the environment.
-In 2010-2011, the most devastating drought in decades resulted in food crisis in many East African countries including Kenya \citep{Chen2015}.
-Widespread catastrophic droughts occurred also in 1984-1985, 2005 and 2008 \citep{Hastenrath2007, Hastenrath2010, Hastenrath2011, Chen2015}.
-As a consequence of climate change, the situation is likely to get worse.
-Dry areas exhibit a strong tendency to get drier while wet areas are getting wetter \citep{Trenberth2014, Chen2015, Kabubo2015}.
-Strong downwards trend in precipitation has already been observed in the tropics from $10^\circ$N to $10^\circ$S, especially after $1977$ \citep{IPCCtrenberth}.
-During the period $1900-2005$, the climate has become wetter in many parts of the world (eastern parts of America, northern Europe, northern and central Asia) but it has became much drier in Mediterranean, Sahel, southern Africa and parts of Southern Asia.
-Furthermore, increased frequency of heavy rain events has been observed in the areas with decline in total rainfall \citep{IPCCtrenberth}.
+\begin{itemize}
 
-Droughts and climate variability do not only affect agriculture but they have widespread consequences on food security in general.
-\cite{Ubilava2018} has found a connection between commodity prices and phases of the ENSO cycle.
-\cite{robinson2010} have used a Computable General Equilibrium (CGE) model for Ethiopia and they have shown that production shocks and resulting price increases have negative effects on farmers' income in drought-prone areas while moderate price increases can be beneficial for farmers in less drought-prone areas.
-\cite{OxfamIDS} have focused on sub-Saharan Africa, central America, North Africa and other low-income countries selected by Oxfam and he has concluded that climate change will result in decline in food productivity while food prices will increase.
-\cite{BROWN2015} have investigated food prices in $554$ local commodity markets in Africa, South Asia and Latin America and they concluded that almost 20\% of local market food prices had been affected by weather disturbances.
+\item Extreme weather causes disasters $\rightarrow$	 early warning systems have been developed
+\end{itemize}
+\item[] \textbf{Paragraph 2}
+\begin{itemize}
+\item What weather forecasts (measures) have been used in EWS? \textit{ref. litrature}
+\begin{itemize}
+\item Mostly seasonal precip. totals and temperature averages
+\end{itemize}
+\end{itemize}
 
-A large literature has focused on relationship of agricultural production and climate in Sub-Saharan Africa.
-Increase in temperature has typically been found to affect crop revenue negatively while increase in precipitation has mostly been associated with positive effects in the literature.
-\cite{RicardianBello} and  \cite{Ochieng2016} have found these results by the means of Ricardian analysis.
-\cite{Ochieng2016} have focused on Kenya while \cite{RicardianBello} have addressed the situation in Niger.
+\item[] \textbf{Paragraph 3} 
+\begin{itemize}
 
-Other studies have investigated the relationship of climate and yields in the context of climate change.
-\cite{MendelsohnDinar2000} have developed a simulation model to evaluate impacts of climate change on African agriculture.
-According to their study, agricultural GDP and climate change exhibit a hill-shaped relationship.
-Hence, global warming should be beneficial in high latitudes but harmful in low latitudes.
-The estimated loss of the Africa as a continent ranges from \$25 billion to \$194 billion per year with $2$ \textdegree C warming \citep{MendelsohnDinar2000}.
-\cite{kabubo2007} have shown that in Kenya, higher summer temperatures are harmful while higher winter temperatures are beneficial.
-Furthermore, fall and summer precipitation is positively correlated with net revenue \citep{kabubo2007}.
-\cite{KMendelsohn2008} and \cite{SeoMendelsohn} have investigated the distribution of climate change impacts on agriculture in Africa.
-They argued that mild scenario would result in income gains for African farmers while more severe scenario is likely to be harmful.
-\cite{Rowhani2011} have analysed relationship between climate variability and crop production in Tanzania.
-The authors have shown that by $2050$, the climate change will result in significant decrease in crop production and big part of this loss can be attributed to change in intra-seasonal variability of weather.
-\cite{Kabubo2015} have estimated effects of climate change on food security in Kenya.
-They have found that temperature and maize yield exhibit a U-shaped relationship while the relationship of rainfall and maize is hill-shaped.
-Thus, sufficient precipitation is crucial, but excessive rainfall is harmful for maize yields.
-\cite{Mendelsohn2008} has reviewed several studies on effects of climate change on agriculture in developing countries.
-According to the literature review, the tropical and sub-tropical countries are much more sensitive to climate change than temperate agriculture \citep{Mendelsohn2008}.
-\cite{Knox2012} conducted a meta-study of papers focused on effects of climate change impacts on crop productivity in Africa and South Asia.
-The authors concluded that the projected mean change in yield is -$8\%$ by the 2050s in both regions.
-According to a meta-study conducted by \cite{Challinor2014}, losses in global production of wheat, maize and rice are expected with $2$ \textdegree C global warming.
-However, with adaptation the simulated yields are on average $7-12\%$ higher than without adaptation \citep{Challinor2014}.
+\item Identify difference between hazard and disaster
 
-A number of important studies have focused on responses to drought related disasters or emergency situations.
-\cite{Sandstorm2017} have questioned whether drought is actually a major cause of food insecurity in Ethiopia, Kenya and Somalia.
-They have analysed causes of the food crisis identified by the humanitarian appeal documents.
-Majority of these documents have found food availability and food production to be the major causes of food crisis.
-\cite{Sandstorm2017} argue that there is a tendency to explain failure of more complex food systems as 'droughts' and insufficient attention is paid to non-climatic drivers such as food prices or conflicts.
-According to \cite{Sandstorm2017}, a large share of humanitarian response budget has been focused on emergency food aid in comparison to the share focused on interventions to build resilience.
-The authors have suggested to increase the budget share focused on agricultural and livestock production to build resilience.
-Also other authors have promoted shifting from reactive to proactive approach in disaster risk management (\citealp{Mechler2005,IPCC2012ch1,Nicholson2017}).
-\cite{Nicholson2017} has recommended forecast based financing as it can avoid significant disaster losses.
-\cite{Mechler2005} has shown that investments into disaster risk reduction had usually been outweighed by avoided losses.
-Also \cite{IPCC2012ch1} have argued that risk assessment is essential for disaster risk management.
+\begin{itemize}
+\item Not every hazard turns into disaster
+\item For a hazard to become a disaster it needs to have \textbf{impact}
+\item Here, we identify the key metrics which have impact on yield
+\end{itemize}
+
+\end{itemize}
+
+\item[] \textbf{Paragraph 4}
+\begin{itemize}
+\item Crop yield versus climate forecasting
+\end{itemize}
+
+ \item[] \textbf{Paragraph 5}
+\begin{itemize}
+\item Aim of the paper
+\end{itemize}
+
+\end{itemize}
+\color{red}
 
@@
 Measures of food security and drough
+\color{black}
 
-The main contribution of this study is leveraging the most detailed publicly available weather datasets while accounting for the real life crop calendar and different cropping seasons across the country as practised by Kenyan farmers.
-Furthermore, we believe that this is the first study which has described an application of linear mixed effects models to estimate effects of precipitation and temperature on crop yields in Kenya.
-Hence, this is the first time that effects of drought and weather on food security have been estimated taking into account the real crop calendar, spatially-heterogeneous nature of Kenya and allowing for the effects of weather to vary across the country.
-% Kabubo-Mariara and Karanja (2015) similar study> but they use the more econometric approach (fixed effects models and random effects models as referred to by econometricians
 \section{Methodology}\label{Method}
+\color{blue}
+\begin{itemize}
+ \item[] \textbf{Paragraph 1}
+\begin{itemize}
+\item \textit{case study looking at Kenya...}
+\end{itemize}
 
-\subsection{Measures of food security and drought}
+ \item[] \textbf{Paragraph 2}
+\begin{itemize}
+\item Data
+\begin{itemize}
+\item \textit{Just write where the data come from, not much detail needed}
+\end{itemize}
+\end{itemize}
 
-Before beginning our research, we had to answer two essential % fundamental??
-questions: `How should we measure food security?' and `How should we measure drought?' Taking into account the aim of this study, previous literature (see Section~\ref{Introduction}) and data availability, we were considering three main approaches to measuring food security.
-In particular, we considered health and utility indicators, food prices and agricultural crop yields as measures of food security.
-Although some health and utility indicators including the Mid Upper Arm Circumference (MUAC) of children under five, the Coping Strategies Index (see \citealp{CSI}) or the NDMA early warning phases are available in the NDMA Early Warning Bulletins, these data are only available since July $2013$.
-The other problem is that the scale of indicators does not seem to be consistent over the counties and time.
-Furthermore, the values of indicators are often missing and for some months the reports are not present at all.
+ \item[] \textbf{Paragraph 3}
+\begin{itemize}
+\item Statistical approach
+\end{itemize}
 
-Using food prices as a measure of food security turned out to be problematic as well. 
-One problem is data availability as the food prices are usually available for market towns but not as county averages.
-Furthermore, the market data which are available in the publications of \cite{KNBS} (KNBS) only cover several years.
-Another problem is that prices are usually strongly spatially autocorrelated and this could cause the regression estimates to be inconsistent and inefficient unless the model accounts for the spatial autocorrelation structure.
-Modelling the structure of spatial autocorrelation would require developing and estimation of a complicated model consisting of several structural equations and additional data, which may not be available.
-Therefore we opted for agricultural crop yields as a measure of food security and we decided to estimate a single reduced form equation with crop yields as the dependent variable.
-In particular, we focus on maize yields as this crop is the principal staple food in Kenya and it is grown at 90\% of farms in Kenya \citep{FAO}.
-
-The other important question to answer before starting our analysis was: `How is drought defined and which definition and measure of drought is the most suitable for the purpose of this study?' Various definitions of drought and their role have been reviewed by \cite{wilhite1985} and \cite{wilhite2000}.
-For an extensive overview of drought indices see \cite{Heim2002}, \cite{monacelli2005}, \cite{zargar2011} or \cite{svoboda2016}.
-\cite{keyantash2002} have quantified, evaluated and compared number of drought indices used for measuring of severity of meteorological, hydrological and agricultural types of drought.
-For more details about types, measures and indices of drought see Appendix~$1$.
-
-In the recent period, remote sensing data have been used increasingly to monitor levels of greenness and closely related vegetation conditions \citep{nicolai2017}.
-Based on these data, various measures (for example the vegetation condition index) have been developed and used for quantifying drought strength and severity \citep{KlischAtz2016}.
-
-We have considered the Standardised Precipitation Evapotranspiration Index  (SPEI) and the vegetation condition index (VCI) as measures of drought in our analysis.
-We did not find the county level SPEI to be as good predictor of food security as other precipitation and temperature measures and computing the SPEI for the entire high resolution weather dataset would be inadequately demanding.
-We did not use the VCI either as we were not confident enough about the reliability of the available data.\footnote{The values of the VCI which we obtained from the NDMA were not the same as those available by the \cite{BOKU} (BOKU).
-Furthermore, the VCI values did not exactly correspond to the early warning phases although they should according to the \cite{NDMA} (NDMA).}
-
-After conducting a rigorous literature research on measures of drought previously used in the relevant context, we decided to utilize various aggregates of daily precipitation and temeprature.
-The frequency of the maize yield data is yearly while the weather data are daily.
-Hence, the weather data had to be aggregated in order to obtain a dataset conformable with the yield data.
-There are many possible ways of aggregation.
-Commonly used measures include monthly averages (or monthly totals in case of precipitation) and their variances or standard deviations (\citealt{AbrahaSavage2006, LobellEtAl2008, ThorntonEtAl2009}).
-For example, \cite{Adejuwon2004} has analysed relationship of crop yields and three measures of precipitation in Sub-Saharan West Africa.
-These include: \textit{(i)} Total rainfall during the first month of the period from sowing to harvesting (June) \textit{(ii)} Total rainfall during the first two months of the period from sowing to harvesting (June and July) and \textit{(iii)} Total rainfall during the first three months of the period from sowing to harvesting (June, July and August).
-Based on his results, weather during June and July is the most important for crop yield in Sub-Saharan West Africa \citep{Adejuwon2004}.
-Other studies have utilised seasonal totals or means (\citealt{sagoe2006,LobellBurke2010}) or annual totals or means \citep{BLIGNAUT2009}.
-\cite{BenMohamed2002} have found sea surface temperature anomalies at various locations and amount of rainfall in July, August and September to be significant for millet crops in Niger.
-The authors have also considered the maximum air temperature in the hottest month (April) and the minimum air temperature in the coldest month (January) as possible predictors of crops in Niger, but they did not find them significant.
-Some studies are based on simulated daily extremes, averages, daily measures of variance  (\citealt{SchulzeEtA1993,Chipanshi2003,AbrahaSavage2006}) or yearly extremes \citep{sagoe2006}.
-Other measures used to explain variability of maize yield include counts of wet and dry days per a defined period, usually a month or a season (\citealt{BenMohamed2002,AbrahaSavage2006,sagoe2006,Giannakopoulos2009}) or length of rainy season (\citealt{Leemans1993,BenMohamed2002}).
-Definitions of wet and dry days and rainy season vary.
-For example, \cite{BenMohamed2002} have assumed that rainy season begins when the amount of rainfall in three consecutive days reaches at least 25~mm and no dry spell of more than seven days occurs in the following thirty days.
-According to this study, the end of the rainy season is defined as that rainy day after which rain recorded during 20 days is less than 5 mm.
-
-An important group of studies focused on relationship of yield and climate in Sub-Saharan Africa has been focused on degree days (\citealt{SchulzeEtA1993,TingemEtAl2008,WalkerSchulze2008,TingemEtAl2009}) or on the number of days with temperature above certain level or within a defined range (\citealt{Giannakopoulos2009,LauxEtAl}).
-
-%\textcolor{red}{Erin Lentz, can we cite her paper?She has not responded to my email}
-
-Based on the literature research above and complexity of deriving the measures, we short-listed the following aggregates of the weather data:
-
-\begin{itemize}
-
-\item \textbf{\textbf{Precipitation}:}
-
-\begin{itemize}
-\item Cumulative precipitation and its square
-\item Cumulative precipitation during the first two months of season
-\item \textcolor{blue}{Additionally, I will test significance of the cumulative precipitation  during the first month of each season - Guigma has just provided the aggregated variables}
-
-\item Coefficient of variation
-\item Maximum length of dry spell in number of days
-\item Number of dry spells lasting for at least four days
-\item Number of dry spells lasting for at least ten days
-\item Number of dry spells lasting for at least twenty days
-\item Maximum daily precipitation
+\color{red}
+\item No need to write too many details (or a section) about measures of food security
+\item No need to describe all the weather measures/characteristics that we have calculated and/or tried to use and didn't work
 \color{blue}
-\item Additional indicators of floods:
-
-\begin{itemize}
-\item One of the Climate Extreme Indices described at \url{https://www.climdex.org/indices.html} is the sum of precipitation amount on the days where precipitation is above 99th percentile of precip.
-of the whole period (i.e.
-1981-2017 in our case).
-The 99th percentile is calculated based on the sample of wet days, that is the days where precipitation is above or equal to 1 mm (index no.
-26)
-\item Guigma has just sent me this measure and its varieties with thresholds 95th percentile and 90th percentile instead of the 99th percentile
+\item \textit{Just say something like 'We used commonly used measures of weather/drought'}
+\item \textit{Only list the significant weather measures/variables}
 \end{itemize}
 \color{black}
-
-\end{itemize}
-
-\item \textbf{Temperature}
-\begin{itemize}
-\item Seasonal average (of daily average) and its square
-\item Average maximum daily temperature
-\item Standard deviation
-\item Cumulative degree days $10-30\degree$C
-\item Number of heatwave days $>35\degree$C
-\end{itemize}
-\end{itemize}
-\textcolor{blue}{\textbf{Description of how the weather data were aggregated temporally - seasons}}
 

### 2018-11-28 writing/draft2/draftMonika14.tex
draft

--- 2018-11-28 writing/draft2/draftMonika12.tex
@@
 
- \item[] \textbf{Paragraph 2}
+\subsection{Data}\label{Data}
 \begin{itemize}
-\item Data
-\begin{itemize}
-\item \textit{Just write where the data come from, not much detail needed}
-\end{itemize}
+\item Source of the climate data (Boku and Berkeley)
+\item Source of the yield data (Kenya MoA)
 \end{itemize}
@@
 \FloatBarrier
-
-\subsection{Data}
-\begin{itemize}
-\item \textcolor{blue}{Source of data}
-\item \textcolor{blue} {Verbal description of the data, resolution, period, frequency}
-\item \textcolor{blue} {Plot(s)/barplots of yield, precip. and temp.}
-\item \textcolor{blue}{Maybe a table of descriptive statistics or ref. to a table in the Appendix}
-\item \textcolor{blue}{Description how the weather data were aggregated spatially over counties}
-\end{itemize}
 

### 2018-11-28 writing/draft2/draftMonika14.tex
dddrraft

@@
 \color{blue}
-\begin{itemize}
- \item[] \textbf{Paragraph 1}
+
 \begin{itemize}
@@
 \begin{itemize}
-\item Source of the climate data (Boku and Berkeley)
+\item Source of the climate data \cite{BOKU} (BOKU) and \cite{Berkeley}
 \item Source of the yield data (Kenya MoA)
@@
 
- \item[] \textbf{Paragraph 3}
+\color{black}
+\subsection{Statistical approach}
+
+% No need to write too many details (or a section) about measures of food security
+% No need to describe all the weather measures/characteristics that we have calculated and/or tried to use and didn't work
 \begin{itemize}
-\item Statistical approach
-\end{itemize}
-
-\color{red}
-\item No need to write too many details (or a section) about measures of food security
-\item No need to describe all the weather measures/characteristics that we have calculated and/or tried to use and didn't work
 \color{blue}
-\item \textit{Just say something like 'We used commonly used measures of weather/drought'}
-\item \textit{Only list the significant weather measures/variables}
+\item \textit{We used commonly used measures of weather/drought (Only mention the significant weather measures/variables)}
 \end{itemize}
@@
 
-\subsection{Linear mixed models}
 \sloppy
@@
 
-\item Tables of estimates of two models:
-
-\begin{enumerate}
-\item The base model which only includes seasonal precipitation and average seasonal temperature
-\item The preferred specification which includes the significant weather variables (see the preliminary results in Table \ref{KenARe11})
+\item Tables of estimates of the preferred specification which includes the significant weather variables (see the preliminary results in Table \ref{KenARe11})
 
@@
 \end{itemize}
-\end{enumerate}
 

### 2018-11-28 writing/draft2/draftMonika14.tex
redoing draft.
and working hardly

@@
 \usepackage{color}
+\usepackage{xcolor}
 \usepackage{caption}
@@
 \usepackage{graphicx}
-\usepackage{color}
 \usepackage{caption}
@@
 \item \textit{We used commonly used measures of weather/drought (Only mention the significant weather measures/variables)}
+
+\item Describe the temporal aggregation of the weather variables, seasons
 \end{itemize}
@@
 \item Verbal description and interpretation of the results
-\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
-
-\begin{itemize}
-
-\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
-Now we want to make a prediction of yields for the later point in time.
-We will make two predictions: one using the base model and the other based on the preferred specification and we will compare their prediction accuracy.
-Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
-We can use real weather data instead of the forecast.
-This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
-		\item Accompany with a graph(s), possibly barplot(s)
-\end{itemize}
-
-\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss.
-
-\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
-
-\item If time, I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
-
 \item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
@@
 	\end{itemize}
+
+\item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data
 \end{itemize}
 
+\color{black}
 \section{Discussion}\label{Discussion}
+
+\color{red}
 
@@
 
+\section*{If time}\label{iftime:}
+\color{violet}
+\begin{itemize}
+\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
+\begin{itemize}
+\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
+Now we want to make a prediction of yields for the later point in time.
+We will make two predictions: one using the base model (which only includes precipitation and temperature) and the other based on the preferred specification and we will compare their prediction accuracy.
+Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
+We can use real weather data instead of the forecast.
+This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
+		\item Accompany with a graph(s), possibly barplot(s)
+\end{itemize}
+
+\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss.
+
+\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
+
+\item I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
+
+\end{itemize}
+\color{black}
 \FloatBarrier

### 2018-11-29 writing/draft2/draftMonika15.tex
including figure

--- 2018-11-28 writing/draft2/draftMonika14.tex
@@
 \usepackage{csquotes}
-\usepackage{epstopdf}
 \usepackage{euscript}
 \usepackage{textcomp}
+\usepackage{epstopdf}
 
@@
 \vfill
-
+\epstopdfsetup{outdir=./}
 \doublespacing
@@
 
+  \begin{figure}
+   \includegraphics{marginalEffects.pdf}
+\caption{Marginal effects of seasonal rainfall and temperature}\label{bar}
+\end{figure}
+
+\pagebreak
+
 \color{blue}
 
-\item Verbal description and interpretation of the results
-\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
+\item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.
+%\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
 This would give us an approximation of $R^2$ which could be used to compare our models.
 
-		\begin{itemize}
-			\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
+	%	\begin{itemize}%
+		%	\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
 This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
-	\end{itemize}
+	%\end{itemize}
 

### 2018-11-29 Rcodes/NovemberNewVars/ggeffectsImages3.R

@@ -22,7 +22,7 @@ exp(summary(KEN11a)$tTable[,1])
 AllPrec<- ggpredict(KEN11a,terms=c("SeasPr [n=20]"))
 AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
 
-AllTemp<- ggpredict(KEN11a,terms=c("AvgTemp [n=20]"),condition)
+AllTemp<- ggpredict(KEN11a,terms=c("AvgTemp [n=20]"))
 AllTemp[,c(2,4,5)]<-AllTemp[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
 
 ASALPrec<- ggpredict(KEN11a_ASAL,terms=c("SeasPr [n=20]"),condition=c(AvgTemp=0,CVPrec = 0, Spell = 0,Spell4=0,SDTemp =0))
@@ -50,5 +50,12 @@ require(gridExtra)
 pdf("presentations/ggeffects/marginalEffects.pdf")
 grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
+
+
+require(gridExtra)
+cairo_ps("presentations/ggeffects/marginalEffects.eps")
+grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
+dev.off()
+
 
 plot1+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))

### 2018-11-29 writing/draft3/draft0.tex
new draft latex template from ERL

--- 2018-11-28 writing/draft2/draftMonika14.tex
@@
-\documentclass[a4paper,12pt]{article}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{a4wide}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%    INSTITUTE OF PHYSICS PUBLISHING                                   %
+%                                                                      %
+%   `Preparing an article for publication in an Institute of Physics   %
+%    Publishing journal using LaTeX'                                   %
+%                                                                      %
+%    LaTeX source code `ioplau2e.tex' used to generate `author         %
+%    guidelines', the documentation explaining and demonstrating use   %
+%    of the Institute of Physics Publishing LaTeX preprint files       %
+%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
+%                                                                      %
+%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
+%                                                                      %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+% First we have a character check
+%
+% ! exclamation mark    " double quote  
+% # hash                ` opening quote (grave)
+% & ampersand           ' closing quote (acute)
+% $ dollar              % percent       
+% ( open parenthesis    ) close paren.  
+% - hyphen              = equals sign
+% | vertical bar        ~ tilde         
+% @ at sign             _ underscore
+% { open curly brace    } close curly   
+% [ open square         ] close square bracket
+% + plus sign           ; semi-colon    
+% * asterisk            : colon
+% < open angle bracket  > close angle   
+% , comma               . full stop
+% ? question mark       / forward slash 
+% \ backslash           ^ circumflex
+%
+% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
+% abcdefghijklmnopqrstuvwxyz 
+% 1234567890
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\documentclass[12pt]{iopart}
+\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
 
+%Uncomment next line if AMS fonts required
+%\usepackage{iopams}  
 \usepackage{amsfonts}
 \usepackage{amssymb}
-\usepackage{graphicx}
+\usepackage{xcolor}
 \usepackage{color}
-\usepackage{caption}
-\usepackage{array}
 \usepackage{pdfpages}
 \usepackage{float}
-\usepackage[round]{natbib}
 \usepackage{multirow}
 \usepackage{multicol}
-\usepackage{amsxtra}
 \usepackage{amsbsy}
-\usepackage{bm}
-\usepackage{accents}
-\usepackage{chngcntr}
-\usepackage{dcolumn}
 \usepackage[none]{hyphenat}
-\usepackage[affil-it]{authblk}
-\usepackage{datetime}
-\usepackage{colortbl}
 \usepackage{footnote}
-\makesavenoteenv{tabular}
-\usepackage[flushleft]{threeparttable}
+\usepackage{tabularx}
+\usepackage{epstopdf}
+\usepackage{setspace} % for double spacing, maybe remove later
+\usepackage{mathrsfs} 
 \usepackage[hyphens]{url}
 \usepackage{placeins}
-\usepackage{dcolumn}
-\usepackage{longtable}
-\usepackage{booktabs}
-\usepackage{setspace}
-\usepackage{changepage}  
-\usepackage{mathrsfs}
-\usepackage{gensymb}
-
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{color}
-\usepackage{caption}
-\usepackage{array}
-\usepackage{pdfpages}
-\usepackage{float}
-\usepackage[round]{natbib}
-\usepackage{multirow}
-\usepackage{multicol}
-\usepackage{amsxtra}
-\usepackage{amsbsy}
-\usepackage{accents}
-\usepackage{chngcntr}
-\usepackage{tabularx}
-\usepackage{dcolumn}
-\usepackage[none]{hyphenat}
-\usepackage[affil-it]{authblk}
-\usepackage{datetime}
-\usepackage[labelfont=bf]{caption}
-\usepackage{titlesec}
-\usepackage{endnotes}
-\usepackage{csquotes}
-\usepackage{epstopdf}
-\usepackage{euscript}
-\usepackage{textcomp}
-
-\date{\normalsize{November 2018}}
-\title{\Large \bf The Effects of Weather on Maize Yields: New Evidence from Kenya}
-\author{Monika Novackova, Pedram Rowhani, Martin Todd, Dominic Kniveton}
-\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}
-
-\parindent 0pt
-\parskip 0.5em
-\newcommand\starred[1]{\accentset{~~~~~\star}{#1}}
-
-\newcounter{magicrownumbers}
-\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}
+\usepackage[english]{babel}
 
@@
 
-\newdateformat{monthyeardate}{%
-  \monthname[\THEMONTH], \THEYEAR}
+\pdfminorversion=4
+\submitto{\ERL}
 
-  \interfootnotelinepenalty=10000
- \newcolumntype{d}{D{.}{.}{-1}}
- \newcolumntype{e}{D{+}{\,\pm\,}{6,2}}
+\title[ The Effects of Weather on Maize Yields: New Evidence from Kenya]{The Effects of Weather on Maize Yields: New Evidence from Kenya}
 
-\makeatletter
-\def\hlinewd#1{%
-\noalign{\ifnum0=`}\fi\hrule \@height #1 %
-\futurelet\reserved@a\@xhline}
-\makeatother
+\author{Monika Novackova, Pedram Rowhani, Martin Todd and Dominic Kniveton}
 
-\maketitle
-\vfill
+\address{Department of Geography, University of Sussex, Falmer, UK}
+\ead{monika.novac@gmail.com}
+\vspace{10pt}
+\begin{indented}
+\item[]November 2018
+\end{indented}
 
 \doublespacing
-
-\begin{itemize}
-\color{blue}
-\item The title is provisional, we will decide about the final title when most of the paper will be written
-\end{itemize}
-
 \begin{abstract}
-\noindent \textcolor{red}{..the full abstract to be written..} This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
+\textcolor{red}{..the full abstract to be written..} This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
 Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
@@
 
-\noindent \textbf{Keywords:}  Drought, Early warning systems, Food security, Food systems, Kenya, Mixed effects models\\
-
-\newpage
-\sloppy
+%
+% Uncomment for keywords
+%\vspace{2pc}
+%\noindent{\it Keywords}: XXXXXX, YYYYYYYY, ZZZZZZZZZ
+%
+% Uncomment for Submitted to journal title message
+%\submitto{\JPA}
+%
+% Uncomment if a separate title page is required
+%\maketitle
+% 
+% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
+%\ioptwocol
+%
 
@@
 	\begin{itemize}
-\item Aim of the paper
-\end{itemize}
-
-\end{itemize}
+		\item Aim of the paper:
 		\color{red}
-
-In the light of these arguments, the goal of the present study is to develop a model which will utilize weather data to assess risk of drought and food security.
+		\\ In the light of these arguments, the goal of the present study is to develop a model which will utilize weather data to assess risk of drought and food security.
 In addition, we want to find out which particular features of climate or weather are the most important factors affecting the yields and food security.
@@
 We focus on Kenya utilizing an yearly panel of $47$ counties over the period of $1981-2017$.
-% mozna hodit tuhle vetu na zacatek 2.1.
-Measures of food security and drough
 		\color{black}
+		\end{itemize}
 
-\section{Methodology}\label{Method}
+	\end{itemize}
+
+\section{Methods}\label{Methods}
 \color{blue}
@@
 \begin{itemize}
-\item \textit{case study looking at Kenya...}
+\item \textit{A case study looking at Kenya...}
 \end{itemize}
@@
 \begin{itemize}
-\item Source of the climate data \cite{BOKU} (BOKU) and \cite{Berkeley}
+\item Source of the climate data: \cite{BOKU} (BOKU) and \cite{Berkeley}
 \item Source of the yield data (Kenya MoA)
@@
 \item \textit{We used commonly used measures of weather/drought (Only mention the significant weather measures/variables)}
+\item Describe the temporal aggregation of the weather variables, seasons
 \end{itemize}
@@
 \sloppy
-Kenya consists of $47$ counties with semi-autonomous county governments  \citep{Barasa2017}.
+Kenya consists of $47$ counties with semi-autonomous county governments  \cite{Barasa2017}.
 As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
-Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \citep{bates2000mixed}.
+Therefore, following the standard methodology, we estimated a battery of linear mixed effects models (also known as mixed models) commonly used to analyse longitudinal data \cite{bates2000mixed}.
 Mixed models are suitable for analysis of panel data as they account for the panel structure of the dataset.
 These types of models include both fixed effects parameters and random effects.
-Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \citep{bates2010lme4}.
+Fixed effects are analogous to parameters in a classical linear regression model and value of each effect is assumed to be fixed over all counties \cite{bates2010lme4}.
 On the other hand, random effects are unobserved random variables.
 There are at least three benefits of treating a set of parameters as a random sample from some distribution.
-\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation (\citealp{KERYch9, KERYch12}).
+\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation %(\citealp{KERYch9, KERYch12}).
 
 Formally, a linear mixed model can be described by the distribution of two vectors of random variables: the response $\mathscr{Y}$ and the vector of random effects $\mathscr{B}$.
-The distribution of $\mathscr{B}$ is multivariate normal and the conditional distribution of $\mathscr{Y}$ given $\mathscr{B}=\mathbf{b}$ is multivariate normal of a form (\citealp{bates2010lme4, KERYch9}):
+The distribution of $\mathscr{B}$ is multivariate normal and the conditional distribution of $\mathscr{Y}$ given $\mathscr{B}=\mathbf{b}$ is multivariate normal of a form %(\citealp{bates2010lme4, KERYch9}):
 
@@
 
-	\section{Results}\label{Results}
+\section{The title and abstract page} 
+If you use \verb"iopart.cls", the code for setting the title page information is slightly different from
+the normal default in \LaTeX.  If you are using a different class file, you do not need to mimic the appearance of
+an \verb"iopart.cls" title page, but please ensure that all of the necessary information is present.
 
-\color{blue}
-\begin{itemize}
+\clearpage
 
-\item Tables of estimates of the preferred specification which includes the significant weather variables (see the preliminary results in Table \ref{KenARe11})
+\appendix
 
-\begin{itemize}
-\item For all counties 
-\item For Arid and semi-araid (ASAL) counties
-\item For non-ASAL counties
-\end{itemize}
+\section{List of macros for formatting text, figures and tables}
 
-\color{black}
+\begin{table}
+\caption{Macros defined within {\tt iopart.cls}
+for use with figures and tables.}
+\begin{indented}
+\item[]\begin{tabular}{@{}l*{15}{l}}
+\br
+Macro name&Purpose\\
+\mr
+\verb"\Figures"&Heading for list of figure captions\\
+\verb"\Figure{#1}"&Figure caption\\
+\verb"\Tables"&Heading for tables and table captions\\
+\verb"\Table{#1}"&Table caption\\
+\verb"\fulltable{#1}"&Table caption for full width table\\
+\verb"\endTable"&End of table created with \verb"\Table"\\
+\verb"\endfulltab"&End of table created with \verb"\fulltable"\\
+\verb"\endtab"&End of table\\
+\verb"\br"&Bold rule for tables\\
+\verb"\mr"&Medium rule for tables\\
+\verb"\ns"&Small negative space for use in table\\
+\verb"\centre{#1}{#2}"&Centre heading over columns\\
+\verb"\crule{#1}"&Centre rule over columns\\
+\verb"\lineup"&Set macros for alignment in columns\\
+\verb"\m"&Space equal to width of minus sign\\
+\verb"\-"&Left overhanging minus sign\\
+\verb"\0"&Space equal to width of a digit\\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
 
-{
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
-%KEN11a  KEN11a_ASAL    KEN11a_nonASAL														
-\label{KenARe11} 
-\begin{footnotesize}
-\begin{tabular}{llrlllr} 
-\hline \vspace{-0.2cm} \\
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}%> anova(KEN11a, type='marginal')
-    &\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}&\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
-  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
-  \\
-  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
-    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
-  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
-  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
-  \\
-  \hline
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
-\vspace{-0.2cm}
-\\
-\hline
-\\
-  \vspace{-0.2cm}Intercept\\
- \\ 
- \hline
-\vspace{-0.2cm} \\
-\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
-\vspace{-0.2cm}
-\\
-
-  \hline
-  \vspace{-0.2cm} \\
-
-\hline
-\vspace{-0.2cm}
-\end{tabular} 
-\end{footnotesize}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-        \begin{adjustwidth}{1cm}{} 
-    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-
-\pagebreak
-
-\color{blue}
-
-\item Verbal description and interpretation of the results
-\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
-
-\begin{itemize}
-
-\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
-Now we want to make a prediction of yields for the later point in time.
-We will make two predictions: one using the base model and the other based on the preferred specification and we will compare their prediction accuracy.
-Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
-We can use real weather data instead of the forecast.
-This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
-		\item Accompany with a graph(s), possibly barplot(s)
-\end{itemize}
-
-\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss.
-
-\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
-
-\item If time, I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
-
-\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
-This would give us an approximation of $R^2$ which could be used to compare our models.
-
-		\begin{itemize}
-			\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
-This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
-	\end{itemize}
-		\end{itemize}
-
-\section{Discussion}\label{Discussion}
-
-	\begin{itemize}
-	\item There are many weather/climate characteristics (measures) which are important for yield, not only seasonal totals (averages)
-	\item  Current weather and climate forecasts (in the context of disaster management) are mostly focused on totals in precipitation (and average temperature?) 
-	\item There is a strong movement to shift the focus from forecasting merely weather/climate to forecast of weather/climate and its impacts
-
-	\begin{itemize}
-	\item[$\boldsymbol{\rightarrow}$] \textbf{We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models}
-\end{itemize}
-
-	\item \textcolor{red}{Caveats}
-
-	\begin{itemize}
-	\item\textcolor{red}{ Yield data: Is the collection method consistent over time and space?}
-	\item Livestock not explicitly taken into account
-	\item Cropping calendar is likely to differ across the country 
-	\item \textcolor{red}{Some more??}
-
-	\end{itemize}
-\end{itemize}
-\color{black}
-
-\FloatBarrier
-\pagebreak
-
-%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT
-
-%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT
-
-\makeatletter 
-\renewcommand{\thesection}{\hspace*{-1.0em}}
-\newpage
-\linespread{1}
-\bibliographystyle{ChicagoM}
-%\bibliographystyle{apa}
-\bibliography{referencesFS}
-
-\newpage
-
-\setcounter{table}{0} 
-\makeatletter 
-\renewcommand{\thetable}{A\@arabic \c@table} 
-\FloatBarrier
-
-\section{Appendix 1 Drought: definitions, measures and indices}
-According to the international meteorological community, drought can be defined in several ways.
-In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
- \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater.
-The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).
-
-\cite{wilhite1985} and \cite{wilhite2000} have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
-Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
-Operational definitions are essential for an effective early warning system.
-An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.
-
-In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
-However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
-Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).
-
-Examples of early measures of drought are \cite{wilhite1985}, \cite{munger1916}, \cite{blumenstock1942} or \cite{mcquigg1954}.
-\cite{munger1916} suggested to use length of period without $24$-h precipitation of $1.27$ mm.
-\cite{wilhite1985} is based on a measure of precipitation over a given time period.
-\cite{blumenstock1942} proposed to measure severity of drought as a length of drought in days where the end of a drought is defined by occurrence of $2.54$ mm of precipitation in $48$ hours.
-\cite{mcquigg1954} developed the Antecedent Precipitation Index (API) which is based on amount and timing of precipitation and it was used for forecasting of floods.
-Hence, the API is a reverse drought index.
-
-The study of \cite{palmer1965} was a significant milestone in the history of quantification of drought severity.
-\cite{palmer1965} developed the Palmer Drought Severity Index (PDSI) using a complex water balance model.
-The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand (\citealp{Heim2002,palmer1965}).
-As the PDSI suffers from several weaknesses (for details see e.g.
-\citealt{Heim2002}), other indices were developed in the following decades.
-These include the standardized precipitation index (SPI) developed by \cite{SPI} and the standardized precipitation evapotranspiration index (SPEI) developed by \cite{SPEI}.
-The SPI specifies observed precipitation as a standardised departure from a chosen probability distribution which models the precipitation data.
-Values of SPI can be viewed as a multiple of standard deviations by which the observed amount of rainfall deviates from the long-term mean \citep{SPIonline}.\footnote{Can be created for various periods of 1-36 months, usually using monthly data.} The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of evapotranspiration (which captures increased temperature).
-It is based on water balance, therefore it can be compared to the self-calibrated PDSI \citep{SPEI}.
-
-\section{Appendix 3 Tables}
+\clearpage
 

### 2018-11-29 writing/draft3/draft2.tex
asdads

--- 2018-11-29 writing/draft2/draftMonika15.tex
@@
-\documentclass[a4paper,12pt]{article}
-\usepackage[english]{babel}
-\usepackage{amsmath}
-\usepackage{a4wide}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%    INSTITUTE OF PHYSICS PUBLISHING                                   %
+%                                                                      %
+%   `Preparing an article for publication in an Institute of Physics   %
+%    Publishing journal using LaTeX'                                   %
+%                                                                      %
+%    LaTeX source code `ioplau2e.tex' used to generate `author         %
+%    guidelines', the documentation explaining and demonstrating use   %
+%    of the Institute of Physics Publishing LaTeX preprint files       %
+	%    `iopart.cls, iopart12.clo and iopart10.clo'.                      %
+%                                                                      %
+%    `ioplau2e.tex' itself uses LaTeX with `iopart.cls'                %
+%                                                                      %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%
+% First we have a character check
+%
+% ! exclamation mark    " double quote  
+% # hash                ` opening quote (grave)
+% & ampersand           ' closing quote (acute)
+% $ dollar              % percent       
+% ( open parenthesis    ) close paren.  
+% - hyphen              = equals sign
+% | vertical bar        ~ tilde         
+% @ at sign             _ underscore
+% { open curly brace    } close curly   
+% [ open square         ] close square bracket
+% + plus sign           ; semi-colon    
+% * asterisk            : colon
+% < open angle bracket  > close angle   
+% , comma               . full stop
+% ? question mark       / forward slash 
+% \ backslash           ^ circumflex
+%
+% ABCDEFGHIJKLMNOPQRSTUVWXYZ 
+% abcdefghijklmnopqrstuvwxyz 
+% 1234567890
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+\documentclass[12pt]{iopart}
+\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
 
+%Uncomment next line if AMS fonts required
+%\usepackage{iopams}  
 \usepackage{amsfonts}
 \usepackage{amssymb}
-\usepackage{graphicx}
+\usepackage{xcolor}
 \usepackage{color}
-\usepackage{xcolor}
-\usepackage{caption}
-\usepackage{array}
 \usepackage{pdfpages}
 \usepackage{float}
-\usepackage[round]{natbib}
 \usepackage{multirow}
 \usepackage{multicol}
-\usepackage{amsxtra}
 \usepackage{amsbsy}
-\usepackage{bm}
-\usepackage{accents}
-\usepackage{chngcntr}
-\usepackage{dcolumn}
 \usepackage[none]{hyphenat}
-\usepackage[affil-it]{authblk}
-\usepackage{datetime}
-\usepackage{colortbl}
 \usepackage{footnote}
-\makesavenoteenv{tabular}
+\usepackage{tabularx}
+\usepackage{epstopdf}
+\usepackage{setspace} % for double spacing, maybe remove later
+\usepackage{mathrsfs} 
+\usepackage[hyphens]{url}
+\usepackage[english]{babel}
+\usepackage{harvard}
 \usepackage[flushleft]{threeparttable}
-\usepackage[hyphens]{url}
-\usepackage{placeins}
-\usepackage{dcolumn}
-\usepackage{longtable}
-\usepackage{booktabs}
-\usepackage{setspace}
 \usepackage{changepage}
-\usepackage{mathrsfs}
-\usepackage{gensymb}
-
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{array}
-\usepackage{pdfpages}
-\usepackage{float}
-\usepackage[round]{natbib}
-\usepackage{multirow}
-\usepackage{multicol}
-\usepackage{amsxtra}
-\usepackage{amsbsy}
-\usepackage{accents}
-\usepackage{chngcntr}
-\usepackage{tabularx}
-\usepackage{dcolumn}
-\usepackage[none]{hyphenat}
-\usepackage[affil-it]{authblk}
-\usepackage{datetime}
-\usepackage[labelfont=bf]{caption}
-\usepackage{titlesec}
-\usepackage{endnotes}
-\usepackage{csquotes}
-\usepackage{euscript}
-\usepackage{textcomp}
-\usepackage{epstopdf}
-
-\date{\normalsize{November 2018}}
-\title{\Large \bf The Effects of Weather on Maize Yields: New Evidence from Kenya}
-\author{Monika Novackova, Pedram Rowhani, Martin Todd, Dominic Kniveton}
-\affil{\small{Department of Geography, University of Sussex, Falmer, UK}}
-
-\parindent 0pt
-\parskip 0.5em
-\newcommand\starred[1]{\accentset{~~~~~\star}{#1}}
-
-\newcounter{magicrownumbers}
-\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}
 
@@
 
-\newdateformat{monthyeardate}{%
-  \monthname[\THEMONTH], \THEYEAR}
+\pdfminorversion=4
+\submitto{\ERL}
+\makesavenoteenv{tabular}
 
-  \interfootnotelinepenalty=10000
- \newcolumntype{d}{D{.}{.}{-1}}
- \newcolumntype{e}{D{+}{\,\pm\,}{6,2}}
+\title[ The Effects of Weather on Maize Yields: New Evidence from Kenya]{The Effects of Weather on Maize Yields: New Evidence from Kenya}
 
-\makeatletter
-\def\hlinewd#1{%
-\noalign{\ifnum0=`}\fi\hrule \@height #1 %
-\futurelet\reserved@a\@xhline}
-\makeatother
+\author{Monika Novackova, Pedram Rowhani, Martin Todd and Dominic Kniveton}
 
-\maketitle
-\vfill
-\epstopdfsetup{outdir=./}
+\address{Department of Geography, University of Sussex, Falmer, UK}
+\ead{monika.novac@gmail.com}
+\vspace{10pt}
+\begin{indented}
+\item[]November 2018
+\end{indented}
+
 \doublespacing
-
-\begin{itemize}
-\color{blue}
-\item The title is provisional, we will decide about the final title when most of the paper will be written
-\end{itemize}
-
 \begin{abstract}
-\noindent \textcolor{red}{..the full abstract to be written..} This paper contributes to better understanding of effects of drought on food security in Kenya which should lead to improving of early warning systems and food security.
-Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
-\\
-...Applying the linear mixed effects models, we found that...
+...Applying the linear mixed effects models, we found that...\\
+\textcolor{red}{.. to be written later..}
 \end{abstract}
 
-\noindent \textbf{Keywords:}  Drought, Early warning systems, Food security, Food systems, Kenya, Mixed effects models\\
-
-\newpage
-\sloppy
+%
+% Uncomment for keywords
+%\vspace{2pc}
+%\noindent{\it Keywords}: XXXXXX, YYYYYYYY, ZZZZZZZZZ
+%
+% Uncomment for Submitted to journal title message
+%\submitto{\JPA}
+%
+% Uncomment if a separate title page is required
+%\maketitle
+% 
+% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
+%\ioptwocol
+%
 
@@
 	\end{itemize}
-\color{red}
 
-In the light of these arguments, the goal of the present study is to develop a model which will utilize weather data to assess risk of drought and food security.
-In addition, we want to find out which particular features of climate or weather are the most important factors affecting the yields and food security.
-Are the average weather conditions the most important or is it the weather variability or length and number of dry spells during the growing season what matters the most?
-Answering these and other similar questions should help to improve food security and shift the focus from reactive to proactive approach in drought disaster risk management.
-We focus on Kenya utilizing an yearly panel of $47$ counties over the period of $1981-2017$.
-% mozna hodit tuhle vetu na zacatek 2.1.
-Measures of food security and drough
+\section{Methods}\label{Methods}
+\color{blue}
+
+...a case study looking at Kenya...
 \color{black}
 
-\section{Methodology}\label{Method}
-\color{blue}
-
-\begin{itemize}
-\item \textit{case study looking at Kenya...}
-\end{itemize}
+Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
 
 \subsection{Data}\label{Data}
+\color{blue}
 \begin{itemize}
-\item Source of the climate data \cite{BOKU} (BOKU) and \cite{Berkeley}
+\item Source of the climate data: \cite{BOKU} (BOKU) and \cite{Berkeley}
 \item Source of the yield data (Kenya MoA)
@@
 \item \textit{We used commonly used measures of weather/drought (Only mention the significant weather measures/variables)}
-
 \item Describe the temporal aggregation of the weather variables, seasons
@@
 \color{black}
-\FloatBarrier
 
@@
   \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}%> anova(KEN11a, type='marginal')
-    &\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}&\boldsymbol{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
+    &{$exp(\beta)$}&\textbf{F-value\tnote{a}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
  \hline 
@@
 
-\pagebreak
-
   \begin{figure}
@@
 \end{figure}
-
-\pagebreak
 
@@
 \color{black}
-\section{Discussion}\label{Discussion}
 
-\color{red}
+\clearpage
 
-	\begin{itemize}
-	\item There are many weather/climate characteristics (measures) which are important for yield, not only seasonal totals (averages)
-	\item  Current weather and climate forecasts (in the context of disaster management) are mostly focused on totals in precipitation (and average temperature?) 
-	\item There is a strong movement to shift the focus from forecasting merely weather/climate to forecast of weather/climate and its impacts
-
-	\begin{itemize}
-	\item[$\boldsymbol{\rightarrow}$] \textbf{We can stress the importance of focusing the forecast on the climate characteristics (or measures) which we find significant in our models}
-\end{itemize}
-
-	\item \textcolor{red}{Caveats}
-
-	\begin{itemize}
-	\item\textcolor{red}{ Yield data: Is the collection method consistent over time and space?}
-	\item Livestock not explicitly taken into account
-	\item Cropping calendar is likely to differ across the country 
-	\item \textcolor{red}{Some more??}
-
-	\end{itemize}
-\end{itemize}
-\color{black}
-
-\section*{If time}\label{iftime:}
-\color{violet}
-\begin{itemize}
-\item Estimates of future impacts and comparison of forecast accuracy of the base and preferred model
-\begin{itemize}
-\item The idea is to choose two points in time from within our period of data and imagine that we are in the earlier of the two time points.
-Now we want to make a prediction of yields for the later point in time.
-We will make two predictions: one using the base model (which only includes precipitation and temperature) and the other based on the preferred specification and we will compare their prediction accuracy.
-Here, we assume that predictions of the weather measures are available (the explanatory variables of our models).
-We can use real weather data instead of the forecast.
-This will demonstrate why it is important to not only forecast totals/means but also other characteristics (measures) of weather/climate.
-		\item Accompany with a graph(s), possibly barplot(s)
-\end{itemize}
-
-\item Estimate a variant of the model which separates the effects of the seasons for ASAL counties (OND and MAM separately). If interesting show a table and discuss.
-
-\item Show a variant of the model without the outliers. This could be presented either by a table in this section, a table in the Appendix or a table in a footnote.
-
-\item I can also estimate similar models for different important crops in Kenya (e.g. wheat, rice, sorghum or millet)
-
-\end{itemize}
-\color{black}
-\FloatBarrier
-\pagebreak
-
-%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT
-
-%TTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTTT
-
-\makeatletter 
-\renewcommand{\thesection}{\hspace*{-1.0em}}
-\newpage
-\linespread{1}
-\bibliographystyle{ChicagoM}
-%\bibliographystyle{apa}
+\bibliographystyle{dcu}
 \bibliography{referencesFS}
 
-\newpage
-
-\setcounter{table}{0} 
-\makeatletter 
-\renewcommand{\thetable}{A\@arabic \c@table} 
-\FloatBarrier
-
-\section{Appendix 1 Drought: definitions, measures and indices}
-According to the international meteorological community, drought can be defined in several ways.
-In particular, drought is a \textit{'prolonged absence or marked deficiency of precipitation'}, a \textit{'deficiency of precipitation that results in water shortage for some activity or for some group'} or a \textit{'period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance'} (\citealp{Heim2002, IPCCtrenberth}).
- \cite{AMS1997} has defined three types of droughts: \textit{(i)}~'Agricultural drought' which is defined in terms of moister deficits in upper layer of soil up to about one meter depth~\textit{(ii)} 'meteorological drought' which refers to prolonged deficit of precipitation and~\textit{(iii)} 'hydrological drought' which relates to low streamflow, lake and levels of groundwater.
-The  \cite{AMS1997} policy statement was later replaced by another statement \citep{AMS2013} which besides the three types of drought above, covers also the 'socioeconomic drought' which associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought (\citealt{Heim2002, IPCCtrenberth}).
-
-\cite{wilhite1985} and \cite{wilhite2000} have distinguished two main categories of definitions of drought: \textit{(i)} conceptual and \textit{(ii)} operational.
-Conceptual definitions are dictionary types, usually defining boundaries of the concept of drought\footnote{An example of conceptual definition of drought is an 'extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region' \cite{schneider1996}.}.
-Operational definitions are essential for an effective early warning system.
-An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000}.
-
-In order to compare severity of drought across different time periods or geographical locations a numerical measure turns out to be necessary.
-However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
-Instead of that a number of measures of drought has been developed (\citealp{ wilhite1985, wilhite2000, Heim2002}).
-
-Examples of early measures of drought are \cite{wilhite1985}, \cite{munger1916}, \cite{blumenstock1942} or \cite{mcquigg1954}.
-\cite{munger1916} suggested to use length of period without $24$-h precipitation of $1.27$ mm.
-\cite{wilhite1985} is based on a measure of precipitation over a given time period.
-\cite{blumenstock1942} proposed to measure severity of drought as a length of drought in days where the end of a drought is defined by occurrence of $2.54$ mm of precipitation in $48$ hours.
-\cite{mcquigg1954} developed the Antecedent Precipitation Index (API) which is based on amount and timing of precipitation and it was used for forecasting of floods.
-Hence, the API is a reverse drought index.
-
-The study of \cite{palmer1965} was a significant milestone in the history of quantification of drought severity.
-\cite{palmer1965} developed the Palmer Drought Severity Index (PDSI) using a complex water balance model.
-The PDSI is based on a hydrological accounting system, which incorporate antecedent precipitation, moisture supply and moisture demand (\citealp{Heim2002,palmer1965}).
-As the PDSI suffers from several weaknesses (for details see e.g.
-\citealt{Heim2002}), other indices were developed in the following decades.
-These include the standardized precipitation index (SPI) developed by \cite{SPI} and the standardized precipitation evapotranspiration index (SPEI) developed by \cite{SPEI}.
-The SPI specifies observed precipitation as a standardised departure from a chosen probability distribution which models the precipitation data.
-Values of SPI can be viewed as a multiple of standard deviations by which the observed amount of rainfall deviates from the long-term mean \citep{SPIonline}.\footnote{Can be created for various periods of 1-36 months, usually using monthly data.} The SPEI is similar to SPI, but unlike SPI, the SPEI includes the role of evapotranspiration (which captures increased temperature).
-It is based on water balance, therefore it can be compared to the self-calibrated PDSI \citep{SPEI}.
-
-\section{Appendix 3 Tables}
-
 \end{document}

### 2018-11-29 Rcodes/NovemberNewVars/ggeffectsImages3.R

@@ -48,14 +48,17 @@ plot6<-plot(nonASALTemp)+ylab("Yield")+xlab("Average temperature in multiples of
 
 require(gridExtra)
 pdf("presentations/ggeffects/marginalEffects.pdf")
+pdf("writing/draft3/marginalEffects.pdf")
 grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
 
 
 require(gridExtra)
-cairo_ps("presentations/ggeffects/marginalEffects.eps")
+cairo_ps("writing/draft2/marginalEffects.eps")
 grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
+
+
 
 
 plot1+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))

### 2018-12-03 Rcodes/NovemberNewVars/ggeffectsImages3.R
working hardly

@@ -59,6 +59,12 @@ grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
 
 
+require(gridExtra)
+pdf("writing/draft3/Figure1.pdf")
+grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
+dev.off()
+
+
 
 
 plot1+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))

### 2018-12-03 writing/draft3/draft3.tex
working hardly
formating figure

--- 2018-11-29 writing/draft3/draft2.tex
@@
 ...Applying the linear mixed effects models, we found that...\\
-\textcolor{red}{.. to be written later..}
+\textcolor{blue}{.. to be written later..}
 \end{abstract}
@@
 \begin{itemize}
-\item Source of the climate data: \cite{BOKU} (BOKU) and \cite{Berkeley}
+\item Source of the climate data: BOKU and Berkeley Earth
 \item Source of the yield data (Kenya MoA)
@@
   \begin{figure}
-   \includegraphics{marginalEffects.pdf}
-\caption{Marginal effects of seasonal rainfall and temperature}\label{bar}
+   \includegraphics{Figure1a_1f.pdf}\label{bar}
+\caption{Predicted marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+Precipitation and temperature (x-axis) are in multiples of their standard deviations.}
 \end{figure}
@@
 
-\bibliographystyle{dcu}
+\bibliographystyle{dcuM}
+
+ \nocite{BOKU} 
+ \nocite{Berkeley} 
 \bibliography{referencesFS}

### 2018-12-03 Rcodes/NovemberNewVars/ggeffectsImages3.R

@@ -63,8 +63,19 @@ require(gridExtra)
 pdf("writing/draft3/Figure1.pdf")
 grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# now I need to format the images for the journal:
 
+plot1<-plot(AllPrec)+ylab("Yield")+xlab(NULL)+xlab("(a)")+ggtitle("Seasonal precipitation")+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))
+plot2<-plot(AllTemp)+ylab(NULL)+xlab(NULL)+xlab("(b)")+ggtitle("Average temperature")
+plot3<-plot(ASALPrec)+ylab("Yield")+xlab(NULL)+xlab("(c)")+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))+ggtitle(NULL)
+plot4<-plot(ASALTemp)+ylab(NULL)+xlab(NULL)+xlab("(d)")+ggtitle(NULL)
+plot5<-plot(nonASALPrec)+ylab("Yield")+xlab(NULL)+xlab("(e)")+ggtitle(NULL)
+plot6<-plot(nonASALTemp)+ylab(NULL)+xlab(NULL)+xlab("(f)")+ggtitle(NULL)
 
-
-
-plot1+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))
+require(gridExtra)
+pdf("writing/draft3/Figure1a_1f.pdf")
+grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
+dev.off()

### 2018-12-03 Rcodes/NovemberNewVars/ggeffectsImages3.R
commit

@@ -79,3 +79,8 @@ require(gridExtra)
 pdf("writing/draft3/Figure1a_1f.pdf")
 grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
+
+require(gridExtra)
+cairo_ps("writing/draft3/Figure1a_1f.eps")
+grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
+dev.off()

### 2018-12-04 Rcodes/NovemberNewVars/Models1.R
more plots to write and present

@@ -7,15 +7,15 @@ load("dataFS/Main/DaTS.RData")
 
 KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-            data=ScaledTS,na.action=na.exclude); summary(KEN11a)
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a); exp(summary(KEN11a)$coef[[1]])
 # and subsamples
       KEN11a_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                        +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                       data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11a_ASAL)
+                       data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11a_ASAL); exp(summary(KEN11a_ASAL)$coef[[1]])
       
       KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                           +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                          data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL)
+                          data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL); exp(summary(KEN11a_nonASAL)$coef[[1]])
 #----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
       
 # so I can try to add the new variables:

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImagesAllvars.R

@@ -1,0 +1,72 @@
+rm(list=ls())
+library(ggeffects);library(ggplot2);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+ScaledTS$foo=0.25942373
+KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a)
+# and subsamples
+KEN11a_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                 +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11a_ASAL)
+
+KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL)
[…]

### 2018-12-04 writing/draft3/draft3.tex
commit
more plots to write and present
figures marginal effects

@@
 
-\title[ The Effects of Weather on Maize Yields: New Evidence from Kenya]{The Effects of Weather on Maize Yields: New Evidence from Kenya}
+\title{The Effects of weather on maize yields: New evidence from Kenya}
 
@@
 %
-
+\maketitle
 \section{Introduction}\label{Introduction}
@@
    \includegraphics{Figure1a_1f.pdf}\label{bar}
-\caption{Predicted marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields.
+\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields.
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-Precipitation and temperature (x-axis) are in multiples of their standard deviations.}
+Precipitation and temperature (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
+\end{figure}
+
+  \begin{figure}
+   \includegraphics{Figure2a_2f.pdf}\label{bar}
+\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
 \end{figure}
@@
 
+\appendix
+\section*{Appendix}
+should be at the end of the main text but before list of references
+
 \bibliographystyle{dcuM}

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImages3.R

@@ -20,7 +20,7 @@ KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
 exp(summary(KEN11a)$tTable[,1])
 
 AllPrec<- ggpredict(KEN11a,terms=c("SeasPr [n=20]"))
-AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
+AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
 
 AllTemp<- ggpredict(KEN11a,terms=c("AvgTemp [n=20]"))
 AllTemp[,c(2,4,5)]<-AllTemp[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImagesAllvars.R

@@ -20,7 +20,7 @@ KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
 exp(summary(KEN11a)$tTable[,1])
 
   AllPrec<- ggpredict(KEN11a,terms=c("SeasPr [n=20]"))
-  AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
+  AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
 AllTemp<- ggpredict(KEN11a,terms=c("AvgTemp [n=20]"))
 AllTemp[,c(2,4,5)]<-AllTemp[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
   ASALPrec<- ggpredict(KEN11a_ASAL,terms=c("SeasPr [n=20]"),condition=c(AvgTemp=0,CVPrec = 0, Spell = 0,Spell4=0,SDTemp =0))
@@ -34,9 +34,8 @@ nonASALTemp[,c(2,4,5)]<-nonASALTemp[,c(2,4,5)]/exp(summary(KEN11a_nonASAL)$tTabl
 
 #------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 # the other variables
-
 AllPrec<- ggpredict(KEN11a,terms=c("CVPrec [n=20]"))
-AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
+AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
   AllTemp<- ggpredict(KEN11a,terms=c("SDTemp [n=20]"))
   AllTemp[,c(2,4,5)]<-AllTemp[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
[…]

### 2018-12-04 writing/draft3/draft3.tex
marginal effects
ggplot2 - all variables.

@@
 \caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
-The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
 The effects are multiplicative as the models are in log-linear form.}

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImages3.R

@@ -1,86 +1,0 @@
-rm(list=ls())
-library(ggeffects);library(ggplot2);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
-load("dataFS/Main/DaTS.RData")
-
-#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
-# these models are the best ones for now:
-ScaledTS$foo=0.25942373
-KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
-            data=ScaledTS,na.action=na.exclude); summary(KEN11a)
-# and subsamples
-KEN11a_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-                 +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11a_ASAL)
-
-KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-                    +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL)
-#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
-exp(summary(KEN11a)$tTable[,1])
-
[…]

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImages4.R

@@ -1,0 +1,83 @@
+rm(list=ls())
+library(ggeffects);library(ggplot2);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+ScaledTS$foo=0.25942373
+KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a)
+# and subsamples
+KEN11a_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                 +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11a_ASAL)
+
+KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL)
[…]

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImagesAllvars.R

@@ -1,71 +1,0 @@
-rm(list=ls())
-library(ggeffects);library(ggplot2);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
-load("dataFS/Main/DaTS.RData")
-
-#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
-# these models are the best ones for now:
-ScaledTS$foo=0.25942373
-KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
-            data=ScaledTS,na.action=na.exclude); summary(KEN11a)
-# and subsamples
-KEN11a_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-                 +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11a_ASAL)
-
-KEN11a_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-                    +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11a_nonASAL)
[…]

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImages4.R
working hard

@@ -80,4 +80,37 @@ cairo_ps("writing/draft3/Figure2a_2f.eps")
 grid.arrange(plot1, plot2, plot3, plot4,plot5, plot6,ncol=2)
 dev.off()
 
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# other variables
+
+AllSpell<- ggpredict(KEN11a,terms=c("Spell [n=20]"))
+AllSpell[,c(2,4,5)]<-AllSpell[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
+AllSpell4<- ggpredict(KEN11a,terms=c("Spell4 [n=20]"))
+AllSpell4[,c(2,4,5)]<-AllSpell4[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
+ASALSpell<- ggpredict(KEN11a_ASAL,terms=c("Spell [n=20]"),condition=c(AvgTemp=0,SeasPr = 0, CVPrec = 0,SDTemp=0,Spell4 =0))
+ASALSpell[,c(2,4,5)]<-ASALSpell[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
+ASALSpell4<- ggpredict(KEN11a_ASAL,terms=c("Spell4 [n=20]"))
+ASALSpell4[,c(2,4,5)]<-ASALSpell4[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
+nonASALSpell<- ggpredict(KEN11a_nonASAL,terms=c("Spell [n=20]"),condition=c(AvgTemp=0,SeasPr = 0, Spell4 = 0, CVPrec = 0,SDTemp=0))
+nonASALSpell[,c(2,4,5)]<-nonASALSpell[,c(2,4,5)]/exp(summary(KEN11a_nonASAL)$tTable[,1])[["(Intercept)"]]
[…]

### 2018-12-04 Rcodes/NovemberNewVars/ggeffectsImages4.R
working hard

@@ -37,7 +37,7 @@ plot2<-plot(AllTemp)+ylab(NULL)+xlab(NULL)+xlab("(b)")+ggtitle("Average temperat
 plot3<-plot(ASALPrec)+ylab("Yield")+xlab(NULL)+xlab("(c)")+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))+ggtitle(NULL)
 plot4<-plot(ASALTemp)+ylab(NULL)+xlab(NULL)+xlab("(d)")+ggtitle(NULL)
 plot5<-plot(nonASALPrec)+ylab("Yield")+xlab(NULL)+xlab("(e)")+ggtitle(NULL)
-plot6<-plot(nonASALTemp)+ylab(NULL)+xlab(NULL)+xlab("(f)")+ggtitle(NULL)
+plot6<-plot(nonASALTemp)+ylab(NULL)+xlab(NULL)+xlab("(f)")+ggtitle(NULL)+scale_x_continuous(breaks = c(-2,-1.5,-1,-0.5,0,0.5))
 
 require(gridExtra)
 pdf("writing/draft3/Figure1a_1f.pdf")
@@ -96,12 +96,12 @@ nonASALSpell[,c(2,4,5)]<-nonASALSpell[,c(2,4,5)]/exp(summary(KEN11a_nonASAL)$tTa
 nonASALSpell4<- ggpredict(KEN11a_nonASAL,terms=c("Spell4 [n=20]"),condition=c(SeasPr=0,Spell = 0,  CVPrec = 0,SDTemp=0,AvgTemp =0))
 nonASALSpell4[,c(2,4,5)]<-nonASALSpell4[,c(2,4,5)]/exp(summary(KEN11a_nonASAL)$tTable[,1])[["(Intercept)"]]
 #------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-plot1<-plot(AllSpell)+ylab("All")+xlab(NULL)+xlab("(a)")+ggtitle("Effects of precipitation variability on yields")+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))+theme(plot.title=element_text(size=11))
-plot2<-plot(AllSpell4)+ylab(NULL)+xlab(NULL)+xlab("(b)")+ggtitle("Effects of temperature variability on yields")+theme(plot.title=element_text(size=11))
[…]

### 2018-12-04 writing/draft3/draft4.tex
redo table of estimates

--- 2018-12-04 writing/draft3/draft3.tex
@@
 \singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} \\ Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11a  KEN11a_ASAL    KEN11a_nonASAL % in the file Models1.R in folder RCodes\NovemberNewVars												
+\label{KenARe11} 
+\begin{footnotesize}
+\lineup
+\begin{tabular}{@{}lllllll} 
+\br \vspace{-0.2cm} \\
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&\textbf{Estimate}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
+    &\textbf{Estimate}&\textbf{F-value\tnote{a}}&\textbf{Estimate}&\textbf{F-value\tnote{a}}\\
+\mr
+\\
+\vspace{-0.2cm}Intercept&$\m0.259^{***}$&$19.916$&$\m0.243^{*}$&$5.230$&$\m0.344^{**}$&$10.061$\\
+  \\ \vspace{-0.2cm}Prec. total&$\m0.078^{*}$&$\05.402$&$\m0.006^{}$&$0.022$&$\m0.246^{***}$&$19.386$\\
+  \\
+  \vspace{-0.2cm}Prec. total sq.&$-0.028^{*}$&$\04.289$&$\m0.004$&$0.051$&$-0.128^{***}$&$23.747$\\
+    \\ \vspace{-0.2cm}Prec. c. of var.&$-0.079^{\bullet}$&$\03.277$&$-0.031$ &$0.246$&$-0.095$&$\02.231$\\
+  \\  \vspace{-0.2cm}Dry spell -length&$-0.067^{*}$&$\04.810$&$-0.183^{**}$&$6.969$&$-0.012$&$\00.163$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.826$&$-0.156^{**}$&$8.065$&$-0.011$&$\00.096$\\
+  \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.127$&$-0.213^{*}$&$5.376$&$-0.130$&$\01.580$\\
+  \\  \vspace{-0.2cm}Temp. std. dev.&$\m0.042^{\bullet}$&$\03.125$&$\m0.038$&$0.558$&$\m0.057$ ${*}$&$\05.640$\\
+  \\
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+ \\ 
+ \hline
+\vspace{-0.2cm} \\
+\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
+\vspace{-0.2cm}
+\\  
+\br
+\end{tabular} 
+\end{footnotesize}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+        \begin{adjustwidth}{1cm}{} 
+    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
+
+  \begin{figure}
+   \includegraphics{Figure1a_1f.pdf}\label{MarEff1}
+\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+Precipitation and temperature (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
+\end{figure}
+
+  \begin{figure}
+   \includegraphics{Figure2a_2f.pdf}\label{MarEff2}
+\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
+\end{figure}
+
+  \begin{figure}
+   \includegraphics{Figure3a_3f.pdf}\label{MarEff3}
+\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+Maximum length of dry spell and number of dry spells (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
+\end{figure}
+
+\color{blue}
+
+\item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.
+%\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
+This would give us an approximation of $R^2$ which could be used to compare our models.
+
+	%	\begin{itemize}%
+		%	\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
+This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
+	%\end{itemize}
+
+\item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data
+\end{itemize}
+
+\color{black}
+
+\clearpage
+
+\appendix
+\section*{Appendix}
+\textcolor{blue}{\textit{should be at the end of the main text but before list of references}}
+
+\vspace{1cm}
+\textcolor{blue}{Possibly include a table of all values which I get from the lme or lme4 models summary in R, that is. correlation of the fixed effects etcetera}
+
+\vspace{1cm}
+\textcolor{blue}{Maybe also a table with standard errors here}
+
+{
+\begin{threeparttable}
+\singlespacing
 \caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
 %KEN11a  KEN11a_ASAL    KEN11a_nonASAL														
-\label{KenARe11} 
+\label{KenARe11_exponents} 
 \begin{footnotesize}
-\begin{tabular}{llrlllr} 
+\begin{tabular}{@{}llrlllr} 
 \hline \vspace{-0.2cm} \\
@@
 \linespread{1}
+{
+\begin{threeparttable}
+\singlespacing
+\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11a  KEN11a_ASAL    KEN11a_nonASAL														
+\label{KenARe11} 
+\begin{footnotesize}
+\begin{tabular}{@{}llrlllr} 
+\hline \vspace{-0.2cm} \\
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
+    &{$exp(\beta)$}&\textbf{F-value\tnote{a}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
+ \hline 
+\hline
+\\
+\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
+  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
+  \\
+  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
+    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
+  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
+  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
+  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
+  \\
+  \hline
+\vspace{-0.2cm} \\
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
+\vspace{-0.2cm}
+\\
+\hline
+\\
+  \vspace{-0.2cm}Intercept\\
+ \\ 
+ \hline
+\vspace{-0.2cm} \\
+\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
+\vspace{-0.2cm}
+\\
 
-  \begin{figure}
-   \includegraphics{Figure1a_1f.pdf}\label{bar}
-\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields.
-The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-Precipitation and temperature (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
-\end{figure}
+  \hline
+  \vspace{-0.2cm} \\
 
-  \begin{figure}
-   \includegraphics{Figure2a_2f.pdf}\label{bar}
-\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
-The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
-\end{figure}
-
-\color{blue}
-
-\item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.
-%\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
-This would give us an approximation of $R^2$ which could be used to compare our models.
-
-	%	\begin{itemize}%
-		%	\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
-This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
-	%\end{itemize}
-
-\item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data
-\end{itemize}
-
-\color{black}
-
+\hline
+\vspace{-0.2cm}
+\end{tabular} 
+\end{footnotesize}
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+        \begin{adjustwidth}{1cm}{} 
+    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
+     \end{adjustwidth}
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\par}
+\linespread{1}
 \clearpage
 
-\appendix
-\section*{Appendix}
-should be at the end of the main text but before list of references
-
+\section*{References}
 \bibliographystyle{dcuM}
@@
  \nocite{Berkeley}
+
 \bibliography{referencesFS}

### 2018-12-04 writing/draft3/draft4.tex
formating tablewith estimates
in the draft

@@
 \begin{tabular}{@{}lllllll} 
-\br \vspace{-0.2cm} \\
+\br 
 \vspace{-0.2cm} \\
@@
   \\  \vspace{-0.2cm}Dry spell -length&$-0.067^{*}$&$\04.810$&$-0.183^{**}$&$6.969$&$-0.012$&$\00.163$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.826$&$-0.156^{**}$&$8.065$&$-0.011$&$\00.096$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.826$&$-0.157^{**}$&$8.065$&$-0.011$&$\00.096$\\
   \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.127$&$-0.213^{*}$&$5.376$&$-0.130$&$\01.580$\\
@@
 \singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
+\caption{\textit{\textbf{Mixed  effects model:} \\ Log of maize yield and weather, ARMA(1,1) errors}}
 %KEN11a  KEN11a_ASAL    KEN11a_nonASAL										
@@
 \begin{footnotesize}
+\lineup
 \begin{tabular}{@{}llrlllr} 
-\hline \vspace{-0.2cm} \\
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
-    &{$exp(\beta)$}&\textbf{F-value\tnote{a}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
- \hline 
-\hline
+\br   &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
+    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&F-value\tnote{a}% anova(KEN11a, type='marginal')
+    &{$exp(\beta)$}&F-value\tnote{a}&{$exp(\beta)$}&F-value\tnote{a}\\
+\mr
 \\
 \vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
-  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
+  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$\05.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
   \\
-  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
-    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
-  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
-  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
+  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$\04.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
+    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$\03.277$&$0.969$ &$0.246$&$0.909^{}$&$\02.231$\\
+  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$\04.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$\00.163$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$\04.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$\00.096$\\
+  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$\01.580$\\
+  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$\03.125$&$1.039$&$0.558$&$1.059$ ${*}$&$\05.640$\\
   \\
@@
 \\  
-
-  \hline
-  \vspace{-0.2cm} \\
-
-\hline
-\vspace{-0.2cm}
-\end{tabular} 
-\end{footnotesize}
- \begin{tablenotes}
-  \begin{footnotesize}
-    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-        \begin{adjustwidth}{1cm}{} 
-    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
-     \end{adjustwidth}
-\singlespacing
-  \end{footnotesize}
-\end{tablenotes}
-  \end{threeparttable} 
-\par}
-\linespread{1}
-{
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} Log of maize yield and weather, ARMA(1,1) errors}}
-%KEN11a  KEN11a_ASAL    KEN11a_nonASAL														
-\label{KenARe11} 
-\begin{footnotesize}
-\begin{tabular}{@{}llrlllr} 
-\hline \vspace{-0.2cm} \\
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\vspace{0.1cm}\textbf{ }}  &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}% anova(KEN11a, type='marginal')
-    &{$exp(\beta)$}&\textbf{F-value\tnote{a}}&{$exp(\beta)$}&\textbf{F-value\tnote{a}}\\
- \hline 
-\hline
-\\
-\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
-  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$5.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
-  \\
-  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$4.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
-    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$3.277$&$0.969$ &$0.246$&$0.909^{}$&$2.231$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$4.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$0.163$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$4.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$0.096$\\
-  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$1.580$\\
-  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$3.125$&$1.039$&$0.558$&$1.059$ ${*}$&$5.640$\\
-  \\
-  \hline
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
-\vspace{-0.2cm}
-\\
-\hline
-\\
-  \vspace{-0.2cm}Intercept\\
- \\ 
- \hline
-\vspace{-0.2cm} \\
-\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
-\vspace{-0.2cm}
-\\
-
-  \hline
-  \vspace{-0.2cm} \\
-
-\hline
-\vspace{-0.2cm}
+\br
 \end{tabular} 

### 2018-12-04 writing/draft3/draft4.tex
stareting to write..
again. first methods and data descriotion..

@@
 
-Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
+\subsection{Data}\label{Data}
 
-\subsection{Data}\label{Data}
+In this study, we analyzed the relationship between maize yields and climate. Our dataset consisted of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
+	We acquired the county level yield data from the Famine Early Warning Systems Network (FEWS NET).
+As for the weather data, we utilized $0.25^\circ$ resolution daily precipitation and temperature gridded data.
+The precipitation data come the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth.
+
+	In order to obtain a conformable dataset of yield, precipitation and temperature (that is all data should have same frequency and spatial resolution), the gridded weather data were first averaged over the counties of Kenya so that we obtained daily time series of each county.
+In the next step, we aggregated the resulting county level time series temporally to obtain yearly time series corresponding to the frequency of the yield data.
+We only considered weather data during the rainy seasons which we defined as follows, using the classification of...
+
 \color{blue}
@@
 \par}
-\linespread{1}
+
 \clearpage
@@
 \bibliographystyle{dcuM}
-
- \nocite{BOKU} 
+\nocite{FEWSNET}
+\nocite{CHIRPS} 
 \nocite{Berkeley}

### 2018-12-04 writing/draft3/draft4.tex
write more

@@
 
-	In order to obtain a conformable dataset of yield, precipitation and temperature (that is all data should have same frequency and spatial resolution), the gridded weather data were first averaged over the counties of Kenya so that we obtained daily time series of each county.
-In the next step, we aggregated the resulting county level time series temporally to obtain yearly time series corresponding to the frequency of the yield data.
+	In order to proceed with the analysis, the daily weather data had to be converted to an yearly, county-level panel to correspond with the yield data.
+The first step of this procedure was averaging the gridded weather data over the counties of Kenya resulting in daily time series for each county.
+In the next step, we aggregated the resulting county level time series temporally to obtain yearly time series.
 We only considered weather data during the rainy seasons which we defined as follows, using the classification of...
 
-\color{blue}
-\begin{itemize}
-\item Source of the climate data: BOKU and Berkeley Earth
-\item Source of the yield data (Kenya MoA)
-\end{itemize}
-
-\color{black}
 \subsection{Statistical approach}

### 2018-12-05 writing/draft3/draft5.tex
writing about data

--- 2018-12-04 writing/draft3/draft4.tex
@@
 
-Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of 1981-2017.
+\subsection{Data}\label{Data}
 
-\subsection{Data}\label{Data}
-\color{blue}
-\begin{itemize}
-\item Source of the climate data: BOKU and Berkeley Earth
-\item Source of the yield data (Kenya MoA)
-\end{itemize}
+In this study, we analyzed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
+	We acquired the county level yield data from the Famine Early Warning Systems Network (FEWS NET).
+As for the weather data, we utilized $0.25^\circ$ resolution precipitation and temperature gridded datasets.
+The precipitation data come from the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth.
 
-\color{black}
+	Both precipitation and temperature datasets are daily while the yield data are yearly by its nature.
+Hence, in order to proceed with the analysis, the daily precipitation and temperature data had to be converted
+into a county-level yearly panel.
+The first step of this procedure was averaging the gridded weather data over the counties of Kenya resulting in a single daily time series for each county.
+In the next step, we aggregated the resulting county level time series temporally to obtain yearly time series using the seasonal crop calendar which is available at the website of FEWS~NET.
+We separated the counties of Kenya into two groups, such that the crop calendar can be considered homogeneous within each group.
+The groups are: \textit{(i)} Arid and semi arid (ASAL) counties and \textit{(ii)} non-ASAL as defined by the ASAL Stakeholders Forum (ASF).
+In particular, the ASAL counties include ..see appendix..\footnote{The crop calendar is probably not exactly the same for all counties within each group, however, it is similar enough so that it can be considered homogeneous within the groups.
+To a certain extent, the crop calendar is likely to differ even within a single county.
+However, considering the information provided by ASF and FEWS NET and our intermediate results, we opted for the two groups as a compromise between accuracy and complexity.}
+
+ In particular, we   We only considered weather data during the rainy seasons which we defined as follows, using the classification of...
+
 \subsection{Statistical approach}
@@
 \section*{References}
-\bibliographystyle{dcuM}
-
- \nocite{BOKU} 
- \nocite{Berkeley}
 
@@
 
+\bibliographystyle{dcuM}
+\nocite{ASF} 
+\nocite{Berkeley} 
+\nocite{CHIRPS} 
+\nocite{FEWSNET}
+
 \end{document}

### 2018-12-05 writing/draft3/draft5.tex
bibtex sorting..
tableees

@@
 %
+
 \documentclass[12pt]{iopart}
@@
 We separated the counties of Kenya into two groups, such that the crop calendar can be considered homogeneous within each group.
-The groups are: \textit{(i)} Arid and semi arid (ASAL) counties and \textit{(ii)} non-ASAL as defined by the ASAL Stakeholders Forum (ASF).
+The groups are: \textit{(i)} Arid and semi-arid (ASAL) counties and \textit{(ii)} non-ASAL as defined by the ASAL Stakeholders Forum (ASF).
 In particular, the ASAL counties include ..see appendix..\footnote{The crop calendar is probably not exactly the same for all counties within each group, however, it is similar enough so that it can be considered homogeneous within the groups.
@@
 \section*{Appendix}
+
+\begin{table}[H]
+\caption{\label{ASAL}\textbf{List of Arid and semi-arid (ASAL) and non-ASAL counties}}
+
+\begin{indented}
+\item[]\begin{tabular}{@{}lp{10cm}}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{ASAL:}&Baringo, Embu, Garissa, Isiolo, Kajiado, Kilifi Kitui, Kwale, Laikipia, Lamu, Makueni, Mandera, Marsabit, Meru, Mombasa, Narok, Nyeri, Samburu, Taita-Taveta, Tana River, Tharaka Nithi, Turkana, Wajir, West Pokot\\
+\mr
+\textbf{non-ASAL:}&Bomet, Bungoma, Busia, Elgeyo Marakwet, Homa Bay, Kakamega, Kericho         Kiambu, Kirinyaga, Kisii, Kisumu, Machakos, Migori, Murang'a, Nakuru, Nyamira,   Nyandarua, Siaya, Trans Nzoia, Uasin Gishu, Vihiga\\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
+
 \textcolor{blue}{\textit{should be at the end of the main text but before list of references}}
@@
 \lineup
-\begin{tabular}{@{}llrlllr} 
+\begin{indented}
+\item[]\begin{tabular}{@{}llrlllr} 
 \br   &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
@@
   \\
+  \hline\\[-1em]
+  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\ 
+  \\[-1em]
 \hline
-\vspace{-0.2cm} \\
-  \multicolumn{1}{l}{\textbf{Random effects:}}  & \\
-\vspace{-0.2cm}
-\\
-\hline
-\\
-  \vspace{-0.2cm}Intercept\\
- \\ 
- \hline
-\vspace{-0.2cm} \\
-\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}\\
-\vspace{-0.2cm}
+\\[-1em]Intercept\\
+ \\[-1em] \hline
+\\[-1em]
+\textit{Number of observations:}  &\multicolumn{2}{c}{$1300$}&\multicolumn{2}{c}{$698$}&\multicolumn{2}{c}{$602$}
 \\
@@
 \end{tabular} 
+\end{indented}
 \end{footnotesize}
@@
 \bibliographystyle{dcuM}
+
 \nocite{ASF} 

### 2018-12-05 writing/draft3/draft6.tex
writing
working hardly

--- 2018-12-05 writing/draft3/draft5.tex
@@
 We separated the counties of Kenya into two groups, such that the crop calendar can be considered homogeneous within each group.
-The groups are: \textit{(i)} Arid and semi-arid (ASAL) counties and \textit{(ii)} non-ASAL as defined by the ASAL Stakeholders Forum (ASF).
-In particular, the ASAL counties include ..see appendix..\footnote{The crop calendar is probably not exactly the same for all counties within each group, however, it is similar enough so that it can be considered homogeneous within the groups.
+The groups are: \textit{(i)} Arid and semi-arid (ASAL) counties and \textit{(ii)} non-ASAL counties as defined by the ASAL Stakeholders Forum (ASF).
+For the list of counties in each of the groups see table~\ref{ASAL} in the appendix.
+\footnote{The crop calendar is probably not exactly the same for all counties within each group, however, it is similar enough so that it can be considered homogeneous within the groups.
 To a certain extent, the crop calendar is likely to differ even within a single county.
-However, considering the information provided by ASF and FEWS NET and our intermediate results, we opted for the two groups as a compromise between accuracy and complexity.}
+However, considering the information provided by ASF, FEWS NET and accounting for our intermediate results, we opted for the two groups as a compromise between accuracy and complexity.}.
 
- In particular, we   We only considered weather data during the rainy seasons which we defined as follows, using the classification of...
+In terms of weather, the most relevant periods for yield are the rainy seasons.
+Therefore, we only considered weather during the rainy seasons for our analysis.
+According to FEWS~NET, the ASAL counties have two distinct rainy seasons while the non-ASAL counties have one longer rainy seasons.
+The two rainy seasons of the ASAL counties are: \textit{(i)} March-May and \textit{(ii)} October-December.
+In the non-ASAL area, the rainy season is in March-August.
+
+Hence, for the ASAL counties, we aggregated daily precipitation and temperature over October, November and December of the previous year and May, April, May of the current year  to get single value of the weather variables for each county and year.
+For the non-ASAL counties we aggregated the daily weather values over March, April, May, June, July and August to obtain yearly values.
 
@@
 
+\textcolor{blue}{\textit{should be at the end of the main text but before list of references}}
+
 \begin{table}[H]
@@
 \end{table}
-
-\textcolor{blue}{\textit{should be at the end of the main text but before list of references}}
 

### 2018-12-05 writing/draft3/draft7.tex
writing

--- 2018-12-05 writing/draft3/draft5.tex
@@
 In this study, we analyzed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
-	We acquired the county level yield data from the Famine Early Warning Systems Network (FEWS NET).
+	We acquired the county level yearly yield data from the Famine Early Warning Systems Network (FEWS NET).
 As for the weather data, we utilized $0.25^\circ$ resolution precipitation and temperature gridded datasets.
-The precipitation data come from the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth.
+The precipitation data were obtained from the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth.
+We calculated spatial averages of the gridded data over the counties resulting in a single value for each county and each point in time.
+The frequency of precipitation and temperature data was daily, therefore, further aggregation was needed to obtain yearly values corresponding to the yearly frequency of yield data.
+Hence, each weather characteristic was further aggregated over years resulting in a county-level panel conformable with the yield data.
 
-	Both precipitation and temperature datasets are daily while the yield data are yearly by its nature.
-Hence, in order to proceed with the analysis, the daily precipitation and temperature data had to be converted
-into a county-level yearly panel.
-The first step of this procedure was averaging the gridded weather data over the counties of Kenya resulting in a single daily time series for each county.
-In the next step, we aggregated the resulting county level time series temporally to obtain yearly time series using the seasonal crop calendar which is available at the website of FEWS~NET.
-We separated the counties of Kenya into two groups, such that the crop calendar can be considered homogeneous within each group.
-The groups are: \textit{(i)} Arid and semi-arid (ASAL) counties and \textit{(ii)} non-ASAL as defined by the ASAL Stakeholders Forum (ASF).
-In particular, the ASAL counties include ..see appendix..\footnote{The crop calendar is probably not exactly the same for all counties within each group, however, it is similar enough so that it can be considered homogeneous within the groups.
-To a certain extent, the crop calendar is likely to differ even within a single county.
-However, considering the information provided by ASF and FEWS NET and our intermediate results, we opted for the two groups as a compromise between accuracy and complexity.}
+	There are two predominant rainfall regimes in Kenya.
+The Arid and semi-arid (ASAL) counties mostly exhibit bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October to December.
+In the non-ASAL counties, the single rainy season usually starts in March and lasts until August.
+Following the precipitation regimes and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.
 
- In particular, we   We only considered weather data during the rainy seasons which we defined as follows, using the classification of...
-
+Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature. Only some of them were found significant. These include:
 \subsection{Statistical approach}
@@
 % No need to describe all the weather measures/characteristics that we have calculated and/or tried to use and didn't work
-\begin{itemize}
-\color{blue}
-\item \textit{We used commonly used measures of weather/drought (Only mention the significant weather measures/variables)}
-\item Describe the temporal aggregation of the weather variables, seasons
-\end{itemize}
-\color{black}
 

### 2018-12-06 writing/draft3/draft8.tex
starting to write and copying literature...

--- 2018-12-05 writing/draft3/draft7.tex
@@
 We calculated spatial averages of the gridded data over the counties resulting in a single value for each county and each point in time.
-The frequency of precipitation and temperature data was daily, therefore, further aggregation was needed to obtain yearly values corresponding to the yearly frequency of yield data.
+The frequency of precipitation and temperature data was daily, therefore, further aggregation was needed to obtain yearly values corresponding to the yearly frequency of the yield data.
 Hence, each weather characteristic was further aggregated over years resulting in a county-level panel conformable with the yield data.
@@
 	There are two predominant rainfall regimes in Kenya.
-The Arid and semi-arid (ASAL) counties mostly exhibit bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October to December.
+The Arid and semi-arid (ASAL) counties mostly exhibit bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October and December.
 In the non-ASAL counties, the single rainy season usually starts in March and lasts until August.
-Following the precipitation regimes and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.
+Following the precipitation patterns and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.
 
-Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature. Only some of them were found significant. These include:
+Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature.
+Only some of them were found significant predictors of maize yield.
+These include: Total seasonal precipitation and
 \subsection{Statistical approach}
@@
 \item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.
+
+\item relative importance of the individual variables and its measure
 %\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.

### 2018-12-06 writing/draft3/draft8.tex
copying the literature. Most of it had somewhiat dissappeared. a
also working

@@
 Only some of them were found significant predictors of maize yield.
-These include: Total seasonal precipitation and
+These include: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, standard deviation of average seasonal temperature,  maximum length of dry spell in number of days, the number of dry spell lasting for four days or more
+
+We did not include squared values of the average seasonal temperature as it was not significant.
 \subsection{Statistical approach}
@@
 
-\item relative importance of the individual variables and its measure
+\item relative importance of the individual variables and its measures
 %\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.

### 2018-12-06 Rcodes/November/InterestingFoundOnTheWay/bestOfPlots.R

@@ -13,3 +13,9 @@ hist(DaTS$SeasPr[DaTS$ASAL==1],40)
 hist(DaTS$SeasPr[DaTS$ASAL==0],40)
 
 par(mfrow=c(1,1))
+
+
+plot(Yield~AvgTemp, data=DaTS)
+plot(Yield~I(AvgTemp^2), data=DaTS)
+plot(Yield~AvgTemp, data=DaTS)
+lines(Yield~I(AvgTemp^2), data=DaTS,col=2,type='p') 

### 2018-12-06 writing/draft3/draft8.tex
more literature and writing

--- 2018-12-06 writing/draft3/draft8.tex
@@
 
-Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature.
-Only some of them were found significant predictors of maize yield.
-These include: Total seasonal precipitation and
+Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods.
+Only some of them were found significant in our models.
+The significant measures include: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, standard deviation of average seasonal temperature \footnote{In case of temperature, we used standard deviation rather than coefficient of variation because temperature is measured on an interval scale and the coefficient of variation does not have any meaning for this type of data.
+},  maximum length of dry spell in number of days and the number of dry spell lasting for four days or more (a dry day is defined as day with precipitation less than $1$mm).
+We did not include squared values of the average seasonal temperature as it was not significant.
 \subsection{Statistical approach}
@@
 \begin{itemize}
-\item Add a description of residual and other diagnostic tests (AIC, VIF, autocorrelation)
+\item Add a description of residual and other diagnostic tests (AIC, autocorrelation)
 \item Describe the procedure that I have applied to find the preferred combination of fixed effects and random effects
@@
 
+\item show VIF
 \item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data

### 2018-12-06 writing/draft3/draft8.tex
writing
and mainly moving more literature

@@
 The significant measures include: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, standard deviation of average seasonal temperature \footnote{In case of temperature, we used standard deviation rather than coefficient of variation because temperature is measured on an interval scale and the coefficient of variation does not have any meaning for this type of data.
-},  maximum length of dry spell in number of days and the number of dry spell lasting for four days or more (a dry day is defined as day with precipitation less than $1$mm).
+},  maximum length of dry spell in number of days and the number of dry spell lasting for four days or more (a dry day has been defined as a day when precipitation didn't exceed $1$mm).
 We did not include squared values of the average seasonal temperature as it was not significant.
@@
 \color{black}
+
+To find the most suitable set of predictors, we adopted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-test \cite{lmerTest}.
+That is, we started by including of all possible weather measures in both fixed effects and random effects and we consequently removed the insignificant ones until we got model with only significant fixed and random effects.
+For both fixed and random effects we considered the level of significance $\alpha=0.05$.
 

### 2018-12-06 Rcodes/November/InterestingFoundOnTheWay/bestOfPlots.R

@@ -18,4 +18,7 @@ par(mfrow=c(1,1))
 plot(Yield~AvgTemp, data=DaTS)
 plot(Yield~I(AvgTemp^2), data=DaTS)
 plot(Yield~AvgTemp, data=DaTS)
-lines(Yield~I(AvgTemp^2), data=DaTS,col=2,type='p') 
+
+
+plot(Yield~I(AvgTemp^2), data=DaTS,col=4,type='p',xlim=c(0,800)) 
+lines(Yield~AvgTemp, data=DaTS,col=1,type='p') 

### 2018-12-06 writing/draft3/draft8.tex
writing

@@
 \usepackage{harvard}
+
 \usepackage[flushleft]{threeparttable}
 \usepackage{changepage}
+
+\makeatletter
+\newif\ifnobrackets
+\renewcommand\@cite[2]{\ifnobrackets\else[\fi{#1\if@tempswa , #2\fi}\ifnobrackets\else]\fi\nobracketsfalse}
+\newcommand\nbcite{\nobracketstrue\cite}
+\makeatother
 
@@
 \item Add a description of residual and other diagnostic tests (AIC, autocorrelation)
-\item Describe the procedure that I have applied to find the preferred combination of fixed effects and random effects
 \item Describe the procedure that I have applied to find the preferred way of modelling the  correlation structure in errors (ARMA errors)
@@
 
-To find the most suitable set of predictors, we adopted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-test \cite{lmerTest}.
-That is, we started by including of all possible weather measures in both fixed effects and random effects and we consequently removed the insignificant ones until we got model with only significant fixed and random effects.
+To find the most suitable set of predictors, we adopted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}.
+That is, we started with a model with all possible weather measures in both fixed effects and random effects and we consequently removed the insignificant variables until we left with a model with only significant predictors in fixed and random effects.
 For both fixed and random effects we considered the level of significance $\alpha=0.05$.
+
+According to the conditional Lagrange multiplier (LM) test for serial correlation in the idiosyncratic component of the errors (developed by Baltagi \& Li 1991 and Baltagi \& Li 1995), the errors do exhibit a within group correlation structure (the p-value of the LM test is $9.48\times10^{-14}$).
+To investigate the autocorrelation structure of errors, we estimated all variants of the preferred model with autocorrelation structure of a form ARMA($p$,$q$) such that $p\leq2$ and $q\geq2$.
+Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found out that the most appropriate correlation structure of the errors is ARMA($1$,$1$).
+The AIC criterion see table~\ref{ARMA} in the appendix..
 
@@
 
+\begin{table}[H]
+\caption{\label{ARMA}Comparison of models with different error autcorrelation structure}
+
+\begin{indented}
+\item[]\begin{tabular}{@{}llll}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Error autocorrelation structure}&\textbf{AIC}&\textbf{Likelihood ratio}&\textbf{p-value}\\
+\mr
+\textit{None}& $2205.1$& & \\
+ARMA($1$,$0$)& $2129.2$& $77.9$&$<0.0001$ \\
+ARMA($0$,$1$)&  $2144.9$& & \\
+\textbf{ARMA($\boldsymbol{1}$,$\boldsymbol{1}$)}& $\boldsymbol{2122.2}$& $\boldsymbol{24.7}$&$\boldsymbol{<0.0001}$ \\
+ARMA($2$,$1$)& $2124.0$& $0.1$&$0.7066$ \\
+ARMA($1$,$2$)&  $2124.2$& & \\
+ARMA($2$,$2$)&$2125.8$& $0.3$&$0.5681$ \\
+ARMA($3$,$2$)& $2128.0$& $0.1$&$0.7312$ \\
+ARMA($2$,$3$)&  $2127.6$& & \\
+ARMA($3$,$3$)&\multicolumn{3}{c}{\textit{Failed to converge}} \\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
+
 \vspace{1cm}
@@
 \bibliographystyle{dcuM}
-
+\nocite{Baltagi91}
+\nocite{Baltagi95}
 \nocite{ASF} 

### 2018-12-06 Rcodes/November/Ken_BEST_ARMAerrs112noHWd.R

@@ -1,0 +1,59 @@
+rm(list=ls())
+library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# arma(11) and ar(2) the best. AIC a bit better for 11>> claiming KenARe11 the best !!!
+
+KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a); exp(summary(KEN11a)$coef[[1]])
+
+KEN00a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,
+            data=ScaledTS,na.action=na.exclude); summary(KEN00a); exp(summary(KEN00a)$coef[[1]])
+
+anova(KEN11a,KEN00a)
+
+KEN10a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=0),
+            data=ScaledTS,na.action=na.exclude);
+
+KEN01a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=0,q=1),
+            data=ScaledTS,na.action=na.exclude);
+
+KEN02a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
[…]

### 2018-12-06 Rcodes/November/Ken_BEST_resTests.R

@@ -30,3 +30,5 @@ which(resid(Kendalln3,type="pearson")<(-2))
 which(resid(AmBEST,type="pearson")<(-6))
 # test of AR(1) and MA(1) errors
 pbltest(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + SDTemp  + HWDays,data=ScaledTS,alternative='onesided')#ok, significant
+
+pbltest(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + SDTemp  ,data=ScaledTS,alternative='onesided')#ok, significant

### 2018-12-07 writing/draft3/draft9.tex
working hard
writing editing

--- 2018-12-06 writing/draft3/draft8.tex
@@
 \usepackage{xcolor}
-\usepackage{color}
+\usepackage{color, colortbl}
 \usepackage{pdfpages}
@@
 \usepackage{harvard}
+
 \usepackage[flushleft]{threeparttable}
 \usepackage{changepage}
+
+\makeatletter
+\newif\ifnobrackets
+\renewcommand\@cite[2]{\ifnobrackets\else[\fi{#1\if@tempswa , #2\fi}\ifnobrackets\else]\fi\nobracketsfalse}
+\newcommand\nbcite{\nobracketstrue\cite}
+\makeatother
+
+\definecolor{Gray}{gray}{0.9}
 
@@
 	There are two predominant rainfall regimes in Kenya.
-The Arid and semi-arid (ASAL) counties mostly exhibit bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October and December.
+The arid and semi-arid (ASAL) counties exhibit mostly bi-modal precipitation patterns with long rains lasting from March to May and short rains occurring between October and December.
 In the non-ASAL counties, the single rainy season usually starts in March and lasts until August.
@@
 
-To find the most suitable set of predictors, we adopted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-test \cite{lmerTest}.
-That is, we started by including of all possible weather measures in both fixed effects and random effects and we consequently removed the insignificant ones until we got model with only significant fixed and random effects.
+To find the most suitable set of predictors, we adapted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}.
+To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
+In the following steps, we were sequentially removing the insignificant variables until we were left with a model with only significant predictors in fixed and random effects.
 For both fixed and random effects we considered the level of significance $\alpha=0.05$.
+
+According to the conditional Lagrange multiplier (LM) test for serial correlation in the idiosyncratic component of the errors (developed by Baltagi \& Li 1991 and Baltagi \& Li 1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $9.48\times10^{-14}$).
+To investigate the autocorrelation structure of the errors, we estimated several variants of our preferred specification with the autocorrelation structure of a form ARMA($p$,$q$).
+In particular, we estimated all varieties of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$.
+Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate correlation structure of the errors was ARMA($1$,$1$).
+The AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC criterion of the model without any correlation structure was $2205.1$ and the AIC criterion of the model with ARMA($1$,$0$) errors was $2129.2$.
+The likelihood ratio test of comparison of
 
@@
 \begin{table}[H]
-\caption{\label{ASAL}\textbf{List of Arid and semi-arid (ASAL) and non-ASAL counties}}
+\caption{\label{ASAL}\textbf{List of arid and semi-arid (ASAL) and non-ASAL counties}}
 
@@
 \textbf{non-ASAL:}&Bomet, Bungoma, Busia, Elgeyo Marakwet, Homa Bay, Kakamega, Kericho         Kiambu, Kirinyaga, Kisii, Kisumu, Machakos, Migori, Murang'a, Nakuru, Nyamira,   Nyandarua, Siaya, Trans Nzoia, Uasin Gishu, Vihiga\\
+\br
+\end{tabular}
+\end{indented}
+\end{table}
+
+\begin{table}[H]
+\caption{\label{ARMA}Comparison of models with different error autcorrelation structure}
+
+\begin{indented}
+\lineup
+\item[]\begin{tabular}{@{}llll}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Error autocorrelation structure}&\textbf{AIC}&\textbf{Likelihood ratio}&\multicolumn{1}{c}{\textbf{p-value}}\\
+\mr
+\textit{None}& $2205.1$& & \\
+ARMA($1$,$0$)& $2129.2$& $77.9$&$<1\times10^{-4}$ \\
+ARMA($0$,$1$)&  $2144.9$& & \\
+\rowcolor{Gray}ARMA(${1}$,${1}$)& ${2122.2}$&${24.7}$&${<1\times10^{-4}}$ \\
+ARMA($2$,$1$)& $2124.0$&$\00.1$&$\0\00.7066$ \\
+ARMA($1$,$2$)&  $2124.2$& & \\
+ARMA($2$,$2$)&$2125.8$& $\00.3$&$\0\00.5681$ \\
+ARMA($3$,$2$)& $2128.0$& $\00.1$&$\0\00.7312$ \\
+ARMA($2$,$3$)&  $2127.6$& & \\
+ARMA($3$,$3$)&\multicolumn{3}{c}{\textit{Failed to converge}} \\
 \br

### 2018-12-07 writing/draft3/draft10.tex
writing  and tables..

--- 2018-12-07 writing/draft3/draft9.tex
@@
 
-\color{blue}
-\begin{itemize}
-\item Add a description of residual and other diagnostic tests (AIC, autocorrelation)
-\item Describe the procedure that I have applied to find the preferred way of modelling the  correlation structure in errors (ARMA errors)
-\end{itemize}
-\color{black}
-
 To find the most suitable set of predictors, we adapted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}.
@@
 Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate correlation structure of the errors was ARMA($1$,$1$).
-The AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC criterion of the model without any correlation structure was $2205.1$ and the AIC criterion of the model with ARMA($1$,$0$) errors was $2129.2$.
-The likelihood ratio test of comparison of
+The AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while of the model with ARMA($1$,$0$) errors was $2129.2$.
+The p-value of the likelihood ratio test of comparison of the with ARMA($1$,$1$) errors and the one with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) errors structure is preferable.
+The AIC criteria and likelihood ratio test statics of all models with ARMA($p$,$q$) errors can be found in table~\ref{ARMA} in the appendix.
 
@@
 
-\begin{table}[H]
+{\begin{center}
+\begin{threeparttable}[H]
 \caption{\label{ARMA}Comparison of models with different error autcorrelation structure}
+\lineup
+\begin{indented}
 
-\begin{indented}
-\lineup
 \item[]\begin{tabular}{@{}llll}
@@
 \br
-\\[-1em]
-\textbf{Error autocorrelation structure}&\textbf{AIC}&\textbf{Likelihood ratio}&\multicolumn{1}{c}{\textbf{p-value}}\\
+&&\centre{2}{\textbf{Likelihood ratio} vs. ARMA($1$,$1$)\tnote{a}}\\
+\textbf{Error autocorrelation structure}&\textbf{AIC}&\textbf{Statistic}&\multicolumn{1}{l}{\textbf{p-value}}\\
 \mr
-\textit{None}& $2205.1$& & \\
-ARMA($1$,$0$)& $2129.2$& $77.9$&$<1\times10^{-4}$ \\
-ARMA($0$,$1$)&  $2144.9$& & \\
-\rowcolor{Gray}ARMA(${1}$,${1}$)& ${2122.2}$&${24.7}$&${<1\times10^{-4}}$ \\
-ARMA($2$,$1$)& $2124.0$&$\00.1$&$\0\00.7066$ \\
-ARMA($1$,$2$)&  $2124.2$& & \\
-ARMA($2$,$2$)&$2125.8$& $\00.3$&$\0\00.5681$ \\
-ARMA($3$,$2$)& $2128.0$& $\00.1$&$\0\00.7312$ \\
-ARMA($2$,$3$)&  $2127.6$& & \\
-ARMA($3$,$3$)&\multicolumn{3}{c}{\textit{Failed to converge}} \\
+\textit{None}&$2205.1$&$86.94$&$<1\times10^{-4}$\\
+ARMA($1$,$0$)&$2129.2$&$\09.01$&$0.0027$\\
+ARMA($0$,$1$)&$2144.9$&${24.72}$&${<1\times10^{-4}}$ \\
+\rowcolor{Gray}ARMA(${1}$,${1}$)&${2122.2}$&$---$&$---$\\
+ARMA($2$,$1$)&$2124.0$&$\00.14$&$\0\00.7066$\\
+ARMA($1$,$2$)& $2124.2$&$\00.01$&$\0\00.9143$\\
+ARMA($2$,$2$)&$2125.8$& $\00.31$&$\0\00.8545$\\
 \br
 \end{tabular}
+
 \end{indented}
-\end{table}
+\begin{tablenotes}
+  \begin{footnotesize}
+  \item[a] Likelihood ratio test of comparison with the ARMA(${1}$,${1}$) error structure model.
+ARMA(${1}$,${1}$) error structure seems to be the most suitable as all lower-order correlation structure models are rejected against ARMA(${1}$,${1}$) while ARMA(${1}$,${1}$) is not rejected against the higher order structures.
+\singlespacing
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable} 
+\end{center}
 

### 2018-12-07 Rcodes/November/Ken_BEST_ARMAerrs112noHWd.R

@@ -45,6 +45,17 @@ KEN22a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=2,q=2),
             data=ScaledTS,na.action=na.exclude);
 
+
+KEN31a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=2,q=1),
+            data=ScaledTS,na.action=na.exclude);
+
+
+KEN13a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=2),
+            data=ScaledTS,na.action=na.exclude);
+
+
 KEN32a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=3,q=2),
             data=ScaledTS,na.action=na.exclude);
@@ -57,3 +68,8 @@ KEN33a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=3,q=3),
             data=ScaledTS,na.action=na.exclude);
 anova(KEN00a,KEN10a,KEN01a,KEN11a,KEN21a,KEN12a,KEN22a,KEN23a,KEN32a)
+
+anova(KEN00a,KEN11a); anova(KEN10a,KEN11a);anova(KEN01a,KEN11a)
+anova(KEN21a,KEN11a); anova(KEN12a,KEN11a);anova(KEN22a,KEN11a);
+anova(KEN31a,KEN11a); anova(KEN13a,KEN11a);
+anova(KEN23a,KEN23a)

### 2018-12-07 writing/draft3/draft10.tex
writing and editing

@@
 
-According to the conditional Lagrange multiplier (LM) test for serial correlation in the idiosyncratic component of the errors (developed by Baltagi \& Li 1991 and Baltagi \& Li 1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $9.48\times10^{-14}$).
-To investigate the autocorrelation structure of the errors, we estimated several variants of our preferred specification with the autocorrelation structure of a form ARMA($p$,$q$).
-In particular, we estimated all varieties of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$.
-Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate correlation structure of the errors was ARMA($1$,$1$).
-The AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while of the model with ARMA($1$,$0$) errors was $2129.2$.
-The p-value of the likelihood ratio test of comparison of the with ARMA($1$,$1$) errors and the one with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) errors structure is preferable.
-The AIC criteria and likelihood ratio test statics of all models with ARMA($p$,$q$) errors can be found in table~\ref{ARMA} in the appendix.
+According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $9.48\times10^{-14}$).
+To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with ARMA($p$,$q$) error structure.
+In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$.
+Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$).
+The value of the AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.
+The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) error structure turned out to be a better fit.
+The AIC criteria and the likelihood ratio tests of all the models with ARMA($p$,$q$) errors are summarised in table~\ref{ARMA} in the appendix.
 

### 2018-12-07 writing/draft3/draft10.tex
writing. at least tryiing

@@
 To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the following steps, we were sequentially removing the insignificant variables until we were left with a model with only significant predictors in fixed and random effects.
+In the following steps, the insignificant variables were removed one by one until only significant predictors remained.
 For both fixed and random effects we considered the level of significance $\alpha=0.05$.
+
+According to a histogram, the shape of our maize yield data was closer to log-normal than normal distribution.
+Therefore, following the common practice, the  maize yields values were log-transformed before conducting the analysis.
+Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.
 
@@
 
-\color{blue}
-\begin{itemize}
+Besides our main model we estimated the same specification for two additional subsamples of Kenyan counties.
+In particular, we estimated the model for the subsample of ASAL counties and for the subsample of non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate.
+The estimates of all three models are summarised in table \ref{MainEst}.
+The first two columns represent the model for all counties,  the third and fourth columns include estimates based on the ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties.
+Besides the estimates of coefficients we present measures of relative importance of the predictors.
+These can be found in the columns denoted F-value in table~\ref{MainEst}.
 
-\item Tables of estimates of the preferred specification which includes the significant weather variables (see the preliminary results in Table \ref{KenARe11})
+table \ref{MainEst} includes measures of relative importance of the fixed effects variables.
 
-\begin{itemize}
-\item For all counties 
-\item For Arid and semi-araid (ASAL) counties
-\item For non-ASAL counties
-\end{itemize}
-
-\color{black}
+Consistently with our assumptions, 
+Besides seasonal precipitation and average temperature, the length and number of dry spells
 
@@
 %KEN11a  KEN11a_ASAL    KEN11a_nonASAL % in the file Models1.R in folder RCodes\NovemberNewVars												
-\label{KenARe11} 
+\label{MainEst} 
 \begin{footnotesize}
@@
 
+\begin{itemize}
 \color{blue}
-
 \item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.

### 2018-12-07 Rcodes/November/Ken_BEST_resTestsNoHWdays.R

@@ -6,9 +6,21 @@ load("dataFS/Main/DaTS.RData")
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
 Kendalln3<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-                +AvgTemp + SDTemp  + HWDays+(1|ID1),data=ScaledTS) 
+                +AvgTemp + SDTemp +(1|ID1),data=ScaledTS) 
 summary(Kendalln3); anova(Kendalln3)
 vif(Kendalln3)
+
+      KenUSl3<-lmer(log(Yield)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                      +AvgTemp + SDTemp +(1|ID1),data=DaTS) 
+      summary(KenUSl3); anova(KenUSl3)
+      
+      KenUSl4<-lmer(log(Yield)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + SDTemp +(1|ID1),data=DaTS[DaTS$ASAL==1,]) 
+      summary(KenUSl4); anova(KenUSl4)
+      
+      KenUSl5<-lmer(log(Yield)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + SDTemp +(1|ID1),data=DaTS[DaTS$ASAL==0,]) 
+      summary(KenUSl5); 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
 anova(Kendalln3)

### 2018-12-10 writing/draft3/draft11.tex
getting back to it- re reading my work

--- 2018-12-07 writing/draft3/draft10.tex
@@
 	We acquired the county level yearly yield data from the Famine Early Warning Systems Network (FEWS NET).
-As for the weather data, we utilized $0.25^\circ$ resolution precipitation and temperature gridded datasets.
+Regarding the weather data, we exploited $0.25^\circ$ resolution precipitation and temperature gridded datasets.
 The precipitation data were obtained from the Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS) while the temperature data are available at the website of the Berkeley Earth.
-We calculated spatial averages of the gridded data over the counties resulting in a single value for each county and each point in time.
+We calculated spatial averages of the gridded data over the counties to get a single value for each county and each point in time.
 The frequency of precipitation and temperature data was daily, therefore, further aggregation was needed to obtain yearly values corresponding to the yearly frequency of the yield data.
@@
 The significant measures include: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, standard deviation of average seasonal temperature \footnote{In case of temperature, we used standard deviation rather than coefficient of variation because temperature is measured on an interval scale and the coefficient of variation does not have any meaning for this type of data.
-},  maximum length of dry spell in number of days and the number of dry spell lasting for four days or more (a dry day has been defined as a day when precipitation didn't exceed $1$mm).
+},  maximum length of dry spell in number of days and the number of dry spells lasting for four days or more (we considered every day when precipitation didn't exceed $1$ mm as a dry day).
 We did not include squared values of the average seasonal temperature as it was not significant.
@@
 
-According to a histogram, the shape of our maize yield data was closer to log-normal than normal distribution.
-Therefore, following the common practice, the  maize yields values were log-transformed before conducting the analysis.
+According to a histogram, the shape of our maize yield data was closer to a log-normal than to a normal distribution.
+Therefore, following the common practice, the  maize yield values were log-transformed before conducting the analysis.
 Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.
@@
 
-Besides our main model we estimated the same specification for two additional subsamples of Kenyan counties.
+Besides our main model, we estimated the same specification for two additional subsamples of Kenyan counties.
 In particular, we estimated the model for the subsample of ASAL counties and for the subsample of non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate.
 The estimates of all three models are summarised in table \ref{MainEst}.
-The first two columns represent the model for all counties,  the third and fourth columns include estimates based on the ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties.
-Besides the estimates of coefficients we present measures of relative importance of the predictors.
+The first two columns represent the model for all counties,  the third and fourth columns include estimates based on the ASAL counties and the last two columns describe the estimates based on the subsample of the non-ASAL counties.
+Besides the coefficient estimates, we present measures of relative importance of the predictors.
 These can be found in the columns denoted F-value in table~\ref{MainEst}.
-
-table \ref{MainEst} includes measures of relative importance of the fixed effects variables.
 
@@
   \begin{footnotesize}
-  \item[a] Likelihood ratio test of comparison with the ARMA(${1}$,${1}$) error structure model.
-ARMA(${1}$,${1}$) error structure seems to be the most suitable as all lower-order correlation structure models are rejected against ARMA(${1}$,${1}$) while ARMA(${1}$,${1}$) is not rejected against the higher order structures.
+  \item[a] Likelihood ratio test of a comparison of the model in each row against the ARMA(${1}$,${1}$) error structure model.
+ARMA(${1}$,${1}$) error structure seems to be the most suitable as all lower-order correlation structure models are rejected against ARMA(${1}$,${1}$) while ARMA(${1}$,${1}$) is not rejected against any of the higher order structures.
 \singlespacing

### 2018-12-10 writing/draft3/draft11.tex
writing f-tests results

@@
 We did not include squared values of the average seasonal temperature as it was not significant.
-\subsection{Statistical approach}
+\subsection{Statistical approach}\label{stats}
 
@@
 The first two columns represent the model for all counties,  the third and fourth columns include estimates based on the ASAL counties and the last two columns describe the estimates based on the subsample of the non-ASAL counties.
-Besides the coefficient estimates, we present measures of relative importance of the predictors.
-These can be found in the columns denoted F-value in table~\ref{MainEst}.
+
+Since all the predictors were standardised, the estimates and the corresponding p-values can be used as measures of relative importance of the predictors.
+As an additional measure of relative importance, table~\ref{MainEst} includes F-values of the type \textit{III} analysis of variance.
+That is, the F-tests are conducted using the marginal sum of squares rather than sequential.
+For each variable, the F-statistic corresponds to the additional sum of of squares explained by that particular variable in comparison to the variance explained by the model which includes all other predictors but the one being tested.
+
+, we present measures of relative importance of the predictors. These can be found in the columns denoted F-value in table~\ref{MainEst}.
 

### 2018-12-10 writing/draft3/draft11.tex
asd

--- 2018-12-10 writing/draft3/draft11.tex
@@
 In particular, we estimated the model for the subsample of ASAL counties and for the subsample of non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate.
-The estimates of all three models are summarised in table \ref{MainEst}.
-The first two columns represent the model for all counties,  the third and fourth columns include estimates based on the ASAL counties and the last two columns describe the estimates based on the subsample of the non-ASAL counties.
-Besides the coefficient estimates, we present measures of relative importance of the predictors.
-These can be found in the columns denoted F-value in table~\ref{MainEst}.
+The estimates of all three models are summarised in table~\ref{MainEst}.
+The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties.
+All the predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance.
+As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance in table~\ref{MainEst}.
+The F-tests were conducted using the marginal rather than sequential sum of squares.
+That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable.
+In other worlds, the F-value is a test statistic of comparison of the preferred model (with all variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors of the preferred model.
 
-Consistently with our assumptions, 
+The models have a log-linear functional form, therefore estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effect on yields is multiplicative.
+As in the classical linear regression model, a negative sign implies a negative effect and positive sign is consistent with positive effect on maize yield.
+We will first discuss the estimates for the sample of all counties, that is the first two columns of table~\ref{MainEst}, then we will describe the models of the two subsamples.
+Consistently with our assumptions, seasonal precipitation has positive and significant effect on maize yields while average temperature has negative effects on yields for the case of all counties.
+The squared seasonal precipitation has negative effect, but...marginal total effect
+
+affects the
+
 Besides seasonal precipitation and average temperature, the length and number of dry spells

### 2018-12-11 writing/draft3/draft11.tex
writing results

@@
 That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable.
-In other worlds, the F-value is a test statistic of comparison of the preferred model (with all variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors of the preferred model.
+In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.
 
-The models have a log-linear functional form, therefore estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effect on yields is multiplicative.
-As in the classical linear regression model, a negative sign implies a negative effect and positive sign is consistent with positive effect on maize yield.
-We will first discuss the estimates for the sample of all counties, that is the first two columns of table~\ref{MainEst}, then we will describe the models of the two subsamples.
-Consistently with our assumptions, seasonal precipitation has positive and significant effect on maize yields while average temperature has negative effects on yields for the case of all counties.
-The squared seasonal precipitation has negative effect, but...marginal total effect
+The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects on yield are multiplicative rather than additive.
+As in the classical linear regression model, a negative sign of a coefficient implies a negative effect and positive sign implies a positive effect.
+The estimates of the exponents of the coefficients, which can be interpreted as
+additive marginal effects on yield are in table~\ref{KenARe11_exponents} in the appendix.
+
+We will now focus on the estimates for the sample of all counties, that is the first two columns in table~\ref{MainEst}.
+Consistently with our assumptions, seasonal precipitation has mostly positive and significant effect on maize yields while average temperature has negative effects on yields for the case of all counties.
+We can see in table~\ref{MainEst} that the squared seasonal precipitation is negative while the linear precipitation is posititve,
 
@@
 \singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} \\ Log of maize yield and weather, ARMA(1,1) errors}}
+\caption{\textit{\textbf{Mixed  effects model:} exponents of the coefficients \\ Log of maize yield and weather, ARMA(1,1) errors}}
 %KEN11a  KEN11a_ASAL    KEN11a_nonASAL										

### 2018-12-11 writing/draft3/draft11.tex
witing results

@@
 The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties.
-All the predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance.
+All predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance.
+
 As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance in table~\ref{MainEst}.
@@
 
-The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects on yield are multiplicative rather than additive.
-As in the classical linear regression model, a negative sign of a coefficient implies a negative effect and positive sign implies a positive effect.
-The estimates of the exponents of the coefficients, which can be interpreted as
-additive marginal effects on yield are in table~\ref{KenARe11_exponents} in the appendix.
+The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and the effects on yield are multiplicative.
+The exponents of the estimates (which can be interpreted as additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.
 
-We will now focus on the estimates for the sample of all counties, that is the first two columns in table~\ref{MainEst}.
-Consistently with our assumptions, seasonal precipitation has mostly positive and significant effect on maize yields while average temperature has negative effects on yields for the case of all counties.
-We can see in table~\ref{MainEst} that the squared seasonal precipitation is negative while the linear precipitation is posititve,
-
-affects the
-
-Besides seasonal precipitation and average temperature, the length and number of dry spells
+First, we will focus on the estimates based on the sample of all counties, that is the first two columns in table~\ref{MainEst}.
+Consistently with our assumptions, seasonal precipitation has mostly positive and significant effect on maize yield while average temperature has negative effects on yields.
 
@@
 
+We can see in table~\ref{MainEst} that the squared seasonal precipitation is negative while the linear precipitation is posititve,
+
+affects the
+
+Besides seasonal precipitation and average temperature, the length and number of dry spells
+
   \begin{figure}
@@
 Precipitation and temperature (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
-\end{figure}
-
-  \begin{figure}
-   \includegraphics{Figure2a_2f.pdf}\label{MarEff2}
-\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
-The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
-\end{figure}
-
-  \begin{figure}
-   \includegraphics{Figure3a_3f.pdf}\label{MarEff3}
-\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yields.
-The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-Maximum length of dry spell and number of dry spells (x-axis) are in multiples of their standard deviations.
 The effects are multiplicative as the models are in log-linear form.}
@@
 \singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} exponents of the coefficients \\ Log of maize yield and weather, ARMA(1,1) errors}}
+\caption{\textit{\textbf{Mixed  effects model:} exponents of the coefficient estimates\\ Log of maize yield and weather, ARMA(1,1) errors}}
 %KEN11a  KEN11a_ASAL    KEN11a_nonASAL										
@@
 
+  \begin{figure}
+   \includegraphics{Figure2a_2f.pdf}\label{MarEff2}
+\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
+\end{figure}
+
+  \begin{figure}
+   \includegraphics{Figure3a_3f.pdf}\label{MarEff3}
+\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yields.
+The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
+Maximum length of dry spell and number of dry spells (x-axis) are in multiples of their standard deviations.
+The effects are multiplicative as the models are in log-linear form.}
+\end{figure}
+
 \section*{References}

### 2018-12-11 Rcodes/NovemberNewVars/ggeffectsImages4.R
Update ggeffectsImages4.R

@@ -23,9 +23,11 @@ AllPrec<- ggpredict(KEN11a,terms=c("SeasPr [n=20]"))
   AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
 AllTemp<- ggpredict(KEN11a,terms=c("AvgTemp [n=20]"))
   AllTemp[,c(2,4,5)]<-AllTemp[,c(2,4,5)]/exp(summary(KEN11a)$tTable[,1])[["(Intercept)"]]
-ASALPrec<- ggpredict(KEN11a_ASAL,terms=c("SeasPr [n=20]"),condition=c(AvgTemp=0,CVPrec = 0, Spell = 0,Spell4=0,SDTemp =0))
+
+  ASALPrec<- ggpredict(KEN11a_ASAL,terms=c("SeasPr [n=20]"),condition=c(AvgTemp=0,CVPrec = 0, Spell = 0,Spell4=0,SDTemp =0))
   ASALPrec[,c(2,4,5)]<-ASALPrec[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
-ASALTemp<- ggpredict(KEN11a_ASAL,terms=c("AvgTemp [n=20]"))
+
+ASALTemp<- ggpredict(KEN11a_ASAL,terms=c("AvgTemp [n=20]"),condition=c(SeasP=0,CVPrec = 0, Spell = 0,Spell4=0,SDTemp =0) )
   ASALTemp[,c(2,4,5)]<-ASALTemp[,c(2,4,5)]/exp(summary(KEN11a_ASAL)$tTable[,1])[["(Intercept)"]]
 nonASALPrec<- ggpredict(KEN11a_nonASAL,terms=c("SeasPr [n=20]"),condition=c(AvgTemp=0,CVPrec = 0, Spell = 0,Spell4=0,SDTemp =0))
   nonASALPrec[,c(2,4,5)]<-nonASALPrec[,c(2,4,5)]/exp(summary(KEN11a_nonASAL)$tTable[,1])[["(Intercept)"]]
@@ -57,7 +59,7 @@ AllTempSD<- ggpredict(KEN11a,terms=c("SDTemp [n=20]"))
[…]

### 2018-12-16 Rcodes/December18/Models1step2.R
AIC and so

@@ -1,0 +1,62 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11a<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + SDTemp, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a); exp(summary(KEN11a)$coef[[1]])
+
+fullishLM<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+                 + CVTempK + DDays, random= ~1 | ID1,method="ML",
+            data=ScaledTS,na.action=na.exclude); summary(fullishLM); exp(summary(fullishLM)$coef[[1]])
+
+
+
+fullish<-lmer(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+              + CVTempK + DDays +(1|ID1),data=ScaledTS) #
+
+Cary<-stepAIC(fullishLM)
+
+fullish<-lmer(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+              + CVTempK + DDays +(1|ID1),data=ScaledTS) #
+
+Cary<-stepAIC(fullish)
+
+
+summary(get_model(Cary))
+
+drop1(fullish)
+summary(get_model(Kalinda))
+summary(get_model(Cary))
+extractAIC(get_model(Cary));extractAIC(get_model(Kalinda)) # WOW, Kalinda lower AIC here
+
[…]

### 2018-12-16 Rcodes/December18/Modelsss1.R

@@ -1,0 +1,49 @@
+rm(list=ls())
+library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d); exp(summary(KEN11d)$coef[[1]])
+# and subsamples
+KEN11d_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                 +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11d_ASAL); exp(summary(KEN11d_ASAL)$coef[[1]])
+
+KEN11d_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11d_nonASAL); exp(summary(KEN11d_nonASAL)$coef[[1]])
+#----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+KEN11h<-update(KEN11d,.~.,method="ML")
+KEN11h_ASAL<-update(KEN11d_ASAL,.~.,method="ML")
+KEN11h_nonASAL<-update(KEN11d_nonASAL,.~.,method="ML")
+
+
+drop1(KEN11d)
+drop1(KEN11h)
+ranova(KEN11h)
+
+roza1<-step(KEN11h)
+
[…]

### 2018-12-16 writing/draft3/draft13.tex
knj

--- 2018-12-11 writing/draft3/draft11.tex
@@
 
-\pdfminorversion=4
+%\pdfminorversion=4
 \submitto{\ERL}
@@
 
-\author{Monika Novackova, Pedram Rowhani, Martin Todd and Dominic Kniveton}
+\author{Monika Novackova, Martin Todd, Dominic Kniveton and Pedram Rowhani}
 
@@
 
-According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $9.48\times10^{-14}$).
+According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
 To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with ARMA($p$,$q$) error structure.
@@
 \mr
-\textit{None}&$2205.1$&$86.94$&$<1\times10^{-4}$\\
-ARMA($1$,$0$)&$2129.2$&$\09.01$&$0.0027$\\
-ARMA($0$,$1$)&$2144.9$&${24.72}$&${<1\times10^{-4}}$ \\
+\textit{None}&$2205.4$&$87.17$&$<1\times10^{-4}$\\
+ARMA($1$,$0$)&$2129.2$&$\08.99$&$0.0027$\\
+ARMA($0$,$1$)&$2145.0$&${24.73}$&${<1\times10^{-4}}$ \\
 \rowcolor{Gray}ARMA(${1}$,${1}$)&${2122.2}$&$---$&$---$\\
-ARMA($2$,$1$)&$2124.0$&$\00.14$&$\0\00.7066$\\
-ARMA($1$,$2$)& $2124.2$&$\00.01$&$\0\00.9143$\\
-ARMA($2$,$2$)&$2125.8$& $\00.31$&$\0\00.8545$\\
+ARMA($2$,$1$)&$2124.1$&$\00.15$&$\0\00.6990$\\
+ARMA($1$,$2$)& $2124.2$&$\00.01$&$\0\00.9181$\\
+ARMA($2$,$2$)&$2125.9$& $\00.31$&$\0\00.8549$\\
 \br

### 2018-12-17 Rcodes/December18/ggeffectsImages.R
work

@@ -4,7 +4,7 @@ load("dataFS/Main/DaTS.RData")
 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 # these models are the best ones for now:
-ScaledTS$foo=0.25942373
+
 KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
             data=ScaledTS,na.action=na.exclude); summary(KEN11d)
@@ -17,7 +17,7 @@ KEN11d_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                     +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
                     data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11d_nonASAL)
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
-exp(summary(KEN11d)$tTable[,1])
+exp(summary(KEN11d)$tTable[,1]); exp(summary(KEN11d_ASAL)$tTable[,1]);exp(summary(KEN11d_nonASAL)$tTable[,1])
 
 AllPrec<- ggpredict(KEN11d,terms=c("SeasPr [n=20]"))
   AllPrec[,c(2,4,5)]<-AllPrec[,c(2,4,5)]/exp(summary(KEN11d)$tTable[,1])[["(Intercept)"]]
@@ -98,11 +98,11 @@ nonASALSpell[,c(2,4,5)]<-nonASALSpell[,c(2,4,5)]/exp(summary(KEN11d_nonASAL)$tTa
[…]

### 2018-12-17 writing/draft3/draft13.tex
work
the last bits of today ()or yesterday

@@
 
+\textcolor{blue}{we checked this approach by drop1 =anova testing and it turns out that our model is better than each model without one of the variables..Also the automatic step procedure, which we run for comparison gave us almost the same model (forcing the seasonal precip and average temperature).
+The only varibale which was extra was MaxT and we decided to remove it because of high VIF - or correlation with average temp...Also mention add1}
+
 According to a histogram, the shape of our maize yield data was closer to a log-normal than to a normal distribution.
@@
 
-  \begin{figure}
+  \begin{figure}% new plots checked and good 17.dec.2018
    \includegraphics{Figure1a_1f.pdf}\label{MarEff1}

### 2018-12-17 writing/draft3/draft13.tex
working out
steps, AIC, model selection...

@@
 
-%\pdfminorversion=4
+\pdfminorversion=4
 \submitto{\ERL}
@@
 Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods.
-Only some of them were found significant in our models.
-The significant measures include: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, standard deviation of average seasonal temperature \footnote{In case of temperature, we used standard deviation rather than coefficient of variation because temperature is measured on an interval scale and the coefficient of variation does not have any meaning for this type of data.
-},  maximum length of dry spell in number of days and the number of dry spells lasting for four days or more (we considered every day when precipitation didn't exceed $1$ mm as a dry day).
-We did not include squared values of the average seasonal temperature as it was not significant.
+The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, coefficient of variation of average seasonal temperature (converted into Kelvin because other scales of temperature such us Celsius are interval and coefficient of variation does not have any meaning for data measured on an interval scale),  maximum length of dry spell in number of days and the number of dry spells lasting for four days or more (we considered every day when precipitation didn't exceed $1$ mm as a dry day).
+We did not include squared values of the average seasonal temperature as it was not significant in our models and there is not much evidence of non-linear relationship between yield and temperature in the past literature.
 \subsection{Statistical approach}\label{stats}
@@
 
-As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance in table~\ref{MainEst}.
+As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance (ANOVA) in table~\ref{MainEst}.
 The F-tests were conducted using the marginal rather than sequential sum of squares.
@@
 In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.
+The type~\textit{III} ANOVA does not depend on the order in which the effects are entered in the model.
 

### 2018-12-17 Rcodes/DecemberNew/KEN11d_step.R

@@ -1,0 +1,67 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a); exp(summary(KEN11a)$coef[[1]])
+
+fullishML<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+                 + CVTempK + DDays, random= ~1 | ID1,method="ML",
+            data=ScaledTS,na.action=na.exclude); summary(fullishLM); exp(summary(fullishLM)$coef[[1]])
+
+    CaryML_stepAIC<-stepAIC(fullishML)
+    summary(CaryML_stepAIC)
+          CaryML_step<-step(fullishML) # step not working for lme??
+          
+                                      # both step and stepAIC undefined for REML fit
+                                              fullishRELM<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+                                                           + CVTempK + DDays, random= ~1 | ID1,method="REML",
+                                                           data=ScaledTS,na.action=na.exclude); summary(fullishRELM); exp(summary(fullishRELM)$coef[[1]])
+                                              CaryREML_stepAIC<-step(fullishRELM); CaryREML_stepAIC<-step(fullishRELM)
+                                      # both step and stepAIC undefined for REML fit
[…]

### 2018-12-17 writing/draft3/draft13.tex
writing

@@
 
-To find the most suitable set of predictors, we adapted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}.
-To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the following steps, the insignificant variables were removed one by one until only significant predictors remained.
-For both fixed and random effects we considered the level of significance $\alpha=0.05$.
-
 \textcolor{blue}{we checked this approach by drop1 =anova testing and it turns out that our model is better than each model without one of the variables..Also the automatic step procedure, which we run for comparison gave us almost the same model (forcing the seasonal precip and average temperature).
@@
 
+To find the most suitable set of fixed effects, we adapted a stepwise selection procedure to choose a model with the smallest value of the Akaike’s Information Criterion (AIC).
+We started  with a model which included the complete set of our weather measures in the fixed effects and random intercepts and we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (removing and adding predictors).
+We decided to not to include the number of heatwave days and average maximum daily temperature although they slightly improve the AIC if present.
+The reason for not including them is that they are strongly correlated with average seasonal temperature (the value of correlation coefficient is $0.981$ for the maximum daily temperature and $0.517$ for the number of heatwave days and the p-values of both of them are smaller than $1\times10^{-6}$) and they are not significant at the level of significance $\alpha=0.05$ (their p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
+Furthermore, including them only improved AIC by $4$ (when estimated by maximum likelihood as required for a meaningful comparison) and the signs of these variables were positive in contrary to our assumption.
+Hence, we concluded that the decrease in AIC and positive signs are likely to be result of their correlation with average temperature or other omitted variable and we did not include them in our model.
+
+We further applied a series of likelihood ratio tests to test for significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only included intercepts in random effects.
+
+ allowed for both are moving and adding 
+model building approach using  To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
+In the following steps, the insignificant random effects and subsequently insignificant fixed effects were removed one by one until only significant predictors remained.
+For random effects we considered the level of significance $\alpha=0.1$ and for fixed effects we opted for $\alpha=0.05$~\cite{lmerTest}.
+
+\textcolor{blue}{the Satterthwaite's method to determine the p-values of the individual t-tests} \cite{lmerTest}
 	\section{Results}\label{Results}

### 2018-12-17 Rcodes/DecemberNew/KEN11d_stepDrop1Add1.R

@@ -28,7 +28,7 @@ drop1(KENd_lmer)
 anova(KENd_lmer)
 add1(KENd_lmer,scope=scopefullish)
 
-#----  ---  --   ----      -----     -----    ---  ----   ----       ----      ------        ---- -- - - - - -   - -- - - - - -- - - - - -- - - -- - - - -- - - - -- - -- -- - - --- - -- - - - -- - - 
+#- - - - - - - - - - - - - -- - -- -- -- -    ---  ----   ----       ----      ------        ---- -- - - - - -   - -- - - - - -- - - - - -- - - -- - - - -- - - - -- - -- -- - - --- - -- - - - -- - - 
 # now add1 full..
 scopefull=formula(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                      +AvgTemp + CVTempK+I(AvgTemp^2)+ Prec2m +MaxP + DDays+HWDays+MaxT,data=ScaleTS)

### 2018-12-17 Rcodes/DecemberNew/KEN11d_stepNofix.R

@@ -1,0 +1,75 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11a); exp(summary(KEN11a)$coef[[1]])
+
+fullishML<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+                 + CVTempK + DDays, random= ~1 | ID1,method="ML",
+            data=ScaledTS,na.action=na.exclude); summary(fullishML); exp(summary(fullishML)$coef[[1]])
+
+    CaryML_stepAIC<-stepAIC(fullishML);   summary(CaryML_stepAIC);CaryML_stepAIC;CaryML_stepAIC$anova
+          CaryML_step<-step(fullishML) # step not working for lme??
+          
+                                      # both step and stepAIC undefined for REML fit
+                                              fullishRELM<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+ Prec2m+CVPrec+Spell+Spell4 +MaxP
+                                                           + CVTempK + DDays, random= ~1 | ID1,method="REML",
+                                                           data=ScaledTS,na.action=na.exclude); summary(fullishRELM); exp(summary(fullishRELM)$coef[[1]])
+                                              CaryREML_stepAIC<-step(fullishRELM); CaryREML_stepAIC<-step(fullishRELM)
[…]

### 2018-12-17 Rcodes/DecemberNew/changeHWdaysMaxT.R

@@ -1,0 +1,19 @@
+
+KEN11d_ML<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+               +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+               data=ScaledTS,na.action=na.exclude,method="ML"); summary(KEN11d_ML)
+
+
+KEN11l_ML<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+               +AvgTemp + CVTempK+HWDays+MaxT, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+               data=ScaledTS,na.action=na.exclude,method="ML"); summary(KEN11l_ML)
+
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+               +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+               data=ScaledTS,na.action=na.exclude); summary(KEN11d)
+
+
+KEN11l<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d)

### 2018-12-17 writing/draft3/draft13.tex
asdf

@@
 
- allowed for both are moving and adding 
-model building approach using  To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the following steps, the insignificant random effects and subsequently insignificant fixed effects were removed one by one until only significant predictors remained.
+To verify our results, we adapted an alternative step-down model building
+approach using the Satterthwaite’s method to determine the p-values of the individual
+t-tests \cite{lmerTest}.
+To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
+In the first part of this procedure the insignificant random effects were removed one by one until only significant predictors remained.
+After this, the insignificant fixed effects were removed one by one until only significant predictors remained.
 For random effects we considered the level of significance $\alpha=0.1$ and for fixed effects we opted for $\alpha=0.05$~\cite{lmerTest}.
+Using this method we chose the same model as according to the AIC-based stepwise method described above.
 
-\textcolor{blue}{the Satterthwaite's method to determine the p-values of the individual t-tests} \cite{lmerTest}
 	\section{Results}\label{Results}

### 2018-12-17 Rcodes/DecemberNew/changeHWdaysMaxT.R

@@ -7,7 +7,7 @@ KEN11d_ML<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
 KEN11l_ML<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                +AvgTemp + CVTempK+HWDays+MaxT, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
                data=ScaledTS,na.action=na.exclude,method="ML"); summary(KEN11l_ML)
-
+anova(KEN11d_ML,KEN11l_ML)
 
 KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
@@ -15,5 +15,17 @@ KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
 
 
 KEN11l<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
-            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            +AvgTemp + CVTempK+HWDays+MaxT, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11l)
+
+vif(KEN11l)
+
+cor.test(ScaledTS$AvgTemp,ScaledTS$MaxT)
+cor.test(ScaledTS$AvgTemp,ScaledTS$HWDays)
+
+KEN11l<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK+HWDays, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
             data=ScaledTS,na.action=na.exclude); summary(KEN11d)
+
+vif(KEN11l)
+anova(KEN11d,KEN11l)

### 2018-12-17 writing/draft3/draft14.tex
write and so..

--- 2018-12-17 writing/draft3/draft13.tex
@@
 \sloppy
-Kenya consists of $47$ counties with semi-autonomous county governments  \cite{Barasa2017}.
+Kenya consists of $47$ counties with semi-autonomous county governments.
 As a result of the high degree of county-level autonomy, the policies and regulations often differ across the counties, hence the effects of weather on crop yield are likely to be different across the counties.
@@
 
-To find the most suitable set of predictors, we adapted a step-down model building approach using the Satterthwaite's method to determine the p-values of the individual t-tests \cite{lmerTest}.
-To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the following steps, the insignificant variables were removed one by one until only significant predictors remained.
-For both fixed and random effects we considered the level of significance $\alpha=0.05$.
-
-\textcolor{blue}{we checked this approach by drop1 =anova testing and it turns out that our model is better than each model without one of the variables..Also the automatic step procedure, which we run for comparison gave us almost the same model (forcing the seasonal precip and average temperature).
-The only varibale which was extra was MaxT and we decided to remove it because of high VIF - or correlation with average temp...Also mention add1}
-
-According to a histogram, the shape of our maize yield data was closer to a log-normal than to a normal distribution.
-Therefore, following the common practice, the  maize yield values were log-transformed before conducting the analysis.
+According to a shape of a histogram of the maize yield data we decided to log-transform the yield data before conducting our analysis.
+Hence, we used a log-linear functional form of our models as is a common practice in agricultural economics.
 Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.
 
-According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
+To find the most suitable set of fixed effects, we adapted a stepwise selection procedure to choose a model which minimized the value of the Akaike’s Information Criterion (AIC).
+We started  with a model which included the complete set of our weather measures in the fixed effects and random intercepts and we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors).
+We decided to not to include the number of heatwave days and average maximum daily temperature although including them in the fixed effects slightly improved the AIC.
+The reasons for not including them were their strong correlation with average seasonal temperature (the value of the correlation coefficient was $0.981$ for the correlation of the average seasonal temperature with the maximum daily temperature and $0.517$ the correlation of the average seasonal temperature with the number of heatwave days and the p-values of both of them are smaller than $1\times10^{-6}$) and the lack of significance of their individual t-tests (their p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
+Furthermore, including them only improved AIC by $4$ (when estimated by maximum likelihood as required for a meaningful comparison) and the signs of these variables were positive in contrary to our expectation.
+Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.
+
+We further applied a series of likelihood ratio tests to test for significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.
+
+To verify our results, we adapted an alternative step-down model building
+approach using the Satterthwaite’s method to determine the p-values of the individual
+t-tests \cite{lmerTest}.
+To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
+In the first part of this procedure the insignificant random effects were removed one by one until only significant predictors remained.
+After this, the insignificant fixed effects were removed one by one until only significant predictors remained.
+For random effects we considered the level of significance $\alpha=0.1$ and for fixed effects we opted for $\alpha=0.05$~\cite{lmerTest}.
+Applying this method we obtained the same model as according to the AIC-based stepwise method described above.
+
+\textcolor{blue}{re-write the tables with estimates  UNITS KELVIN AND CV TIMES THOUSAND; possibly table of all available weather measure in the appendix, maybe a drop1 and add1 tables??}
+
+Based on the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
 To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with ARMA($p$,$q$) error structure.

### 2018-12-18 writing/draft3/draft14.tex
writing random effects, tests and so

@@
 
-We further applied a series of likelihood ratio tests to test for significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.
+We further applied a series of likelihood ratio tests of significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.\footnote{Although including the coefficient of variation of temperature as a random effect instead of the random intercepts does improve the AIC from $2164.7$ to $2163.1$ (these AIC values differ from those presented in table~\ref{ARMA} in the appendix as the current models do not account for the error autocorrelation structure due to convergence issues of the model with the random slope) we decided not to include the coefficient of variation of temperature in the random effects.
+The reason for not including the random slope is that the difference in AIC is negligible and F-test of the comparison of the two models is marginally insignificant favouring the simpler model (p-value$=0.059$).}
 

### 2018-12-18 Rcodes/DecemberNew/KEN11d_re.R

@@ -1,0 +1,65 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_re.RData")
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d); exp(summary(KEN11d)$coef[[1]])
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# drop1 undefined for restricted ML, have to refit with ML
+
+KEN11d_re1<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1+SeasPr+AvgTemp | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            control=lmeControl(maxIter=300,msMaxIter=300, msMaxEval=500),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d_re1) # singular convergence
+
+#ehm..the following seems to be going...
+KEN11d_re1<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~CVTempK | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
[…]

### 2018-12-18 Rcodes/DecemberNew/KEN11d_stepNice.R

@@ -82,13 +82,22 @@ CaryML_stepAIC_re2<-stepAIC(fullML_re2)
 
 
 fullML_re3<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+CVPrec+Spell+Spell4 +MaxP
-            + CVTempK + DDays, random= ~1+SeasPr+AvgTemp+I(SeasPr^2)+CVPrec+Spell+Spell4 +MaxP
+            + CVTempK + DDays, random= ~1+SeasPr+AvgTemp+CVPrec+Spell+Spell4 +MaxP
             + CVTempK + DDays| ID1,method="ML",
             data=ScaledTS,na.action=na.exclude,control=lmeControl(maxIter=300,msMaxIter=300,msMaxEval=500)); summary(fullML_re3); exp(summary(fullML_re3)$coef[[1]])
 CaryML_stepAIC_re3<-stepAIC(fullML_re3)
 
 
-save.image("\\\\smbhome.uscs.susx.ac.uk\\mn301\\FoodSystemGitH\\Rcodes\\DecemberNew\\KEN11d_stepNice3.RData")
+save.image("\\\\smbhome.uscs.susx.ac.uk\\mn301\\FoodSystemGitH\\Rcodes\\DecemberNew\\KEN11d_stepNice4.RData")
+
+fullML_re3<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+CVPrec+Spell+Spell4 +MaxP
+            + CVTempK + DDays, random= ~1+SeasPr+AvgTemp+CVPrec+Spell+Spell4 +MaxP
+            + CVTempK + DDays| ID1,method="ML",
+            data=ScaledTS,na.action=na.exclude,control=lmeControl(maxIter=300,msMaxIter=300,msMaxEval=500)); summary(fullML_re3); exp(summary(fullML_re3)$coef[[1]])
+CaryML_stepAIC_re3<-stepAIC(fullML_re3)
+
+
+save.image("\\\\smbhome.uscs.susx.ac.uk\\mn301\\FoodSystemGitH\\Rcodes\\DecemberNew\\KEN11d_stepNice4.RData")
 
 
 summary(CaryML_stepAIC_re); vif(CaryML_stepAIC_re)

### 2018-12-18 writing/draft3/draft15.tex
estimating testing working hard

--- 2018-12-17 writing/draft3/draft14.tex
@@
 
-We further applied a series of likelihood ratio tests to test for significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.
+We further applied a series of likelihood ratio tests of significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.\footnote{Although adding the random slope of the coefficient of variation slightly improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
+The reason for not including the random slope is that the difference in AIC is negligible and the F-test of the comparison of the two models is insignificant favouring the simpler model (p-value$=0.128$).}
 

### 2018-12-18 Rcodes/DecemberNew/KEN11d_re.R

@@ -29,6 +29,12 @@ KEN11d_re2<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                 +AvgTemp + CVTempK, random= ~CVTempK-1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
                 control=lmeControl(maxIter=300,msMaxIter=300, msMaxEval=500),
                 data=ScaledTS,na.action=na.exclude);summary(KEN11d_re2)
+
+KEN11d_re3<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                +AvgTemp + CVTempK, random= ~SeasPr+CVTempK | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                control=lmeControl(maxIter=300,msMaxIter=300, msMaxEval=500),
+                data=ScaledTS,na.action=na.exclude);summary(KEN11d_re3)
+
 anova(KEN11d,KEN11d_re2)
 ranova(KEN11d,KEN11d_re2,reduce.terms=FALSE)
 

### 2018-12-18 Rcodes/DecemberNew/KEN11d_re_lemrinHand.R

@@ -1,0 +1,59 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_re.RData")
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d); exp(summary(KEN11d)$coef[[1]])
+
+KENd_lmer<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + CVTempK+(1 | ID1),data=ScaledTS,na.action=na.exclude)
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+KENd_lmer_reFullish<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + CVTempK+
+                            (1+SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + CVTempK | ID1),data=ScaledTS,na.action=na.exclude)
+
+anova(KENd_lmer,KENd_lmer_reFullish)
+
+KENd_lmer_re2<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + CVTempK+
[…]

### 2018-12-18 Rcodes/DecemberNew/KEN11d_re_step.R

@@ -1,0 +1,21 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_re.RData")
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d); exp(summary(KEN11d)$coef[[1]])
+
+# - -  --  - - - - - - ------- - - - - - - - -- - - - -- - - - -- - - -- - - -  --  - - - - - - ------- - - - - - - - -- - - - -- - - - -- - - -- - - -  --  - - - - - - ------- - - - - - - - -- - - - -- - - - -- - - -- - 
+KENd_lmer<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + CVTempK+(1 | ID1),data=ScaledTS,na.action=na.exclude)
+step(KENd_lmer)
+
+KENd_lmer_reFull<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+AvgTemp +CVPrec+Spell+Spell4+ CVTempK
+                       +(SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4+AvgTemp + CVTempK +MaxP+MaxT+ Prec2m+DDays +HWDays| ID1)
+                       ,data=ScaledTS,na.action=na.exclude)
+KENd_lmer_reFull_step<-step(KENd_lmer_reFull,reduce.fixed=FALSE)
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+save.image("Rcodes/DecemberNew/KEN11d_re_step.RData")

### 2018-12-18 Rcodes/DecemberNew/KEN11d_stepNice.R

@@ -85,7 +85,7 @@ fullML_re3<-lme(log(Yield0)~SeasPr+AvgTemp+I(SeasPr^2)+CVPrec+Spell+Spell4 +MaxP
             + CVTempK + DDays, random= ~1+SeasPr+AvgTemp+CVPrec+Spell+Spell4 +MaxP
             + CVTempK + DDays| ID1,method="ML",
             data=ScaledTS,na.action=na.exclude,control=lmeControl(maxIter=300,msMaxIter=300,msMaxEval=500)); summary(fullML_re3); exp(summary(fullML_re3)$coef[[1]])
-CaryML_stepAIC_re3<-stepAIC(fullML_re3)
+CaryML_stepAIC_re3<-stepAIC(fullML_re3) # error>> singular
 
 
 save.image("\\\\smbhome.uscs.susx.ac.uk\\mn301\\FoodSystemGitH\\Rcodes\\DecemberNew\\KEN11d_stepNice4.RData")

### 2018-12-18 writing/draft3/draft15.tex
writing hardly

@@
 There are at least three benefits of treating a set of parameters as a random sample from some distribution.
-\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation %(\citealp{KERYch9, KERYch12}).
+\textit{(i)} Extrapolation of inference to a wider population \textit{(ii)} improved accounting for system uncertainty and \textit{(iii)} efficiency of estimation \cite{KERYch9,KERYch12}.
 
@@
 
-According to a shape of a histogram of the maize yield data we decided to log-transform the yield data before conducting our analysis.
-Hence, we used a log-linear functional form of our models as is a common practice in agricultural economics.
+It was apparent from the histogram of the maize yield data that the distribution of maize yield was closer to a log-normal than to a normal distribution.
+Therefore, we opted for a log-linear functional form of our models as is a common practice in agricultural economics.
 Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.
 
-To find the most suitable set of fixed effects, we adapted a stepwise selection procedure to choose a model which minimized the value of the Akaike’s Information Criterion (AIC).
-We started  with a model which included the complete set of our weather measures in the fixed effects and random intercepts and we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors).
-We decided to not to include the number of heatwave days and average maximum daily temperature although including them in the fixed effects slightly improved the AIC.
-The reasons for not including them were their strong correlation with average seasonal temperature (the value of the correlation coefficient was $0.981$ for the correlation of the average seasonal temperature with the maximum daily temperature and $0.517$ the correlation of the average seasonal temperature with the number of heatwave days and the p-values of both of them are smaller than $1\times10^{-6}$) and the lack of significance of their individual t-tests (their p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
-Furthermore, including them only improved AIC by $4$ (when estimated by maximum likelihood as required for a meaningful comparison) and the signs of these variables were positive in contrary to our expectation.
+To find the most suitable set of fixed effects, we adapted a stepwise selection procedure which minimized the Akaike's Information Criterion (AIC).
+The first step of the procedure was the estimation of the model with complete set of our weather measures in the fixed effects and with the random intercepts.
+After this we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors).
+We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects slightly improved the AIC.
+The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
+Furthermore, including them only improved AIC from $2080.877$ to $2076.439$ (these values differ from the AIC values presented
+in table~\ref{ARMA} in the appendix because the final estimates compared in table~\ref{ARMA} were estimated using restricted maximum likelihood (REML), but the models were re-estimating using ML method for the comparison as REML comparisons are not meaningful).
+In addition, the signs of these variables were positive in contrary to our expectation.
 Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.

### 2018-12-18 Rcodes/DecemberNew/KEN11d_finals.R

@@ -1,0 +1,40 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d)
+
+KEN11d2<-lme(log(Yield0)~SeasPr +AvgTemp +I(SeasPr^2)+CVPrec+Spell+Spell4
+           + CVTempK+HWDays+MaxT, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d2)
+
+extractAIC(KEN11d);extractAIC(KEN11d2)
+anova(KEN11d2,KEN11d)
+vif(KEN11d2)
+#-------  have to refit with ML ----------- --------------------
+
+KEN11d_ML<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude,method="ML"); summary(KEN11d_ML)
+
+
+KEN11d2_ML<-lme(log(Yield0)~SeasPr +AvgTemp +I(SeasPr^2)+CVPrec+Spell+Spell4
+                              + CVTempK+HWDays+MaxT, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
[…]

### 2018-12-18 Rcodes/DecemberNew/KEN11d_re_step.R

@@ -17,5 +17,6 @@ KENd_lmer_reFull<-lmer(log(Yield0)~SeasPr+I(SeasPr^2)+AvgTemp +CVPrec+Spell+Spel
                        ,data=ScaledTS,na.action=na.exclude)
 KENd_lmer_reFull_step<-step(KENd_lmer_reFull,reduce.fixed=FALSE)
 
+summary(get_model(KENd_lmer_reFull_step)) # groot, nothing much left in the RE
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 save.image("Rcodes/DecemberNew/KEN11d_re_step.RData")

### 2018-12-18 Rcodes/DecemberNew/KEN11d_stepNice.R

@@ -7,7 +7,7 @@ load("dataFS/Main/DaTS.RData")
 
 KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-            data=ScaledTS,na.action=na.exclude); summary(KEN11a); exp(summary(KEN11a)$coef[[1]])
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d); exp(summary(KEN11d)$coef[[1]])
 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 # what I can do with step and stepAIC (each of them works for some cases, here are ONLY those which work...and then theor varieties which I will use..)

### 2018-12-18 writing/draft3/draft15.tex
writing...

@@
 To find the most suitable set of fixed effects, we adapted a stepwise selection procedure which minimized the Akaike's Information Criterion (AIC).
-The first step of the procedure was the estimation of the model with complete set of our weather measures in the fixed effects and with the random intercepts.
-After this we performed a search through the subsets off the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors).
-We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects slightly improved the AIC.
+The first step of the procedure was the estimation of the model with the complete set of our weather measures in the fixed effects and with the random intercepts.
+After this, we performed a search through the subsets of the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors).
+We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects would slightly improve the AIC.
 The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
-Furthermore, including them only improved AIC from $2080.877$ to $2076.439$ (these values differ from the AIC values presented
-in table~\ref{ARMA} in the appendix because the final estimates compared in table~\ref{ARMA} were estimated using restricted maximum likelihood (REML), but the models were re-estimating using ML method for the comparison as REML comparisons are not meaningful).
-In addition, the signs of these variables were positive in contrary to our expectation.
-Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature or perhaps other omitted variables and we did not include them in our model.
+Furthermore, including the heatwave days and maximum daily temperature would only improve AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for the comparison\footnote{In general, REML is the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however REML criterion is not meaningful for comparison of the models with different fixed effects therefore the models need to be re-estimated by ML for comparison \cite{Zuur2009}.}).
+In addition, the coefficients of both heatwave days and maximum daily temperature were positive, which is opposite to what we expected.
+Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature and so we did not include them in our preferred model.
 

### 2018-12-19 writing/draft3/draft16.tex
writing

--- 2018-12-18 writing/draft3/draft15.tex
@@
 The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
-Furthermore, including the heatwave days and maximum daily temperature would only improve AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for the comparison\footnote{In general, REML is the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however REML criterion is not meaningful for comparison of the models with different fixed effects therefore the models need to be re-estimated by ML for comparison \cite{Zuur2009}.}).
+Furthermore, including the heatwave days and maximum daily temperature would only lead to small change in AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for comparison\footnote{In general, REML is the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however, REML criterion is not meaningful for comparison of the models with different fixed effects.
+Therefore, the models had to be re-estimated by ML for comparison \cite{Zuur2009}.}).
 In addition, the coefficients of both heatwave days and maximum daily temperature were positive, which is opposite to what we expected.
-Hence, we concluded that the decrease in AIC and positive signs are likely to be results of the correlation with average temperature and so we did not include them in our preferred model.
+Hence, we concluded that the decrease in AIC and the positive signs are likely to be results of the correlation with average temperature and we did not include them in our preferred model.
 
-We further applied a series of likelihood ratio tests of significance of random slopes, however no subset of random slopes improved the fit, hence, our final model only includes random intercepts.\footnote{Although adding the random slope of the coefficient of variation slightly improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
+We further applied a series of likelihood ratio tests of significance of random slopes, however no subset of random slopes improved the fit.
+Hence, besides the fixed effects, our final model only included random intercepts.\footnote{Although adding the random slope of the coefficient of variation slightly improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
 The reason for not including the random slope is that the difference in AIC is negligible and the F-test of the comparison of the two models is insignificant favouring the simpler model (p-value$=0.128$).}
@@
 t-tests \cite{lmerTest}.
-To be more specific, we started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the first part of this procedure the insignificant random effects were removed one by one until only significant predictors remained.
-After this, the insignificant fixed effects were removed one by one until only significant predictors remained.
-For random effects we considered the level of significance $\alpha=0.1$ and for fixed effects we opted for $\alpha=0.05$~\cite{lmerTest}.
-Applying this method we obtained the same model as according to the AIC-based stepwise method described above.
+We started with a model which included the complete set of our weather measures in both fixed effects and random effects.
+In the first part of the procedure, the insignificant random effects were removed one by one.
+In each iteration, the insignificant variable with the highest p-value was removed and this was repeated until only the significant predictors remained.
+After this, we applied an analogous procedure to the fixed effects until we were left only with the significant ones.
+As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$~ for fixed effects \cite{lmerTest}.
+This method lead to the same model as the one obtained with our primary AIC-based method described above.
 
@@
 
-Based on the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
-To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with ARMA($p$,$q$) error structure.
+According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
+To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with an ARMA($p$,$q$) error structure.
 In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$.
 Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$).
-The value of the AIC criterion of the model with ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.
+The value of the AIC criterion of the model with the ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.
 The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) error structure turned out to be a better fit.

### 2018-12-19 writing/draft3/draft16.tex
work

@@
 The value of the AIC criterion of the model with the ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.
-The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $1 \times10^{-4}$, hence the ARMA($1$,$1$) error structure turned out to be a better fit.
+The p-value of the likelihood ratio test of the comparison of the model with ARMA($1$,$1$) errors and the model with ARMA($1$,$0$) errors was smaller than $0.003$, hence the ARMA($1$,$1$) error structure turned out to be a better fit.
 The AIC criteria and the likelihood ratio tests of all the models with ARMA($p$,$q$) errors are summarised in table~\ref{ARMA} in the appendix.

### 2018-12-19 Rcodes/December18/KEN11d_ARMAerrsR.R

@@ -20,8 +20,8 @@ anova(KEN11d,KEN00d)
 
 KEN10d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=0),
-            data=ScaledTS,na.action=na.exclude);
-
+            data=ScaledTS,na.action=na.exclude);summary(KEN10d)
+anova(KEN11d,KEN10d)
 KEN01d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=0,q=1),
             data=ScaledTS,na.action=na.exclude);

### 2018-12-19 writing/draft3/draft17.tex
writing at home

--- 2018-12-19 writing/draft3/draft16.tex
@@
 
-We further applied a series of likelihood ratio tests of significance of random slopes, however no subset of random slopes improved the fit.
-Hence, besides the fixed effects, our final model only included random intercepts.\footnote{Although adding the random slope of the coefficient of variation slightly improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
+We further applied a series of likelihood ratio tests to test for significance of random slopes.
+No subset of the random slopes improved the fit, therefore besides the fixed effects, our final model only included the random intercepts.\footnote{Although adding the random slope of the coefficient of variation slightly improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
 The reason for not including the random slope is that the difference in AIC is negligible and the F-test of the comparison of the two models is insignificant favouring the simpler model (p-value$=0.128$).}
@@
 In the first part of the procedure, the insignificant random effects were removed one by one.
-In each iteration, the insignificant variable with the highest p-value was removed and this was repeated until only the significant predictors remained.
+In each iteration, the insignificant variable with the highest p-value was removed .
+This was repeated until only the significant predictors remained.
 After this, we applied an analogous procedure to the fixed effects until we were left only with the significant ones.
 As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$~ for fixed effects \cite{lmerTest}.
-This method lead to the same model as the one obtained with our primary AIC-based method described above.
+This method led to the same model as the one obtained with our primary AIC-based method described above.
 

### 2018-12-19 Rcodes/DecemberNew/KEN11d_finals2.R
writing

@@ -17,7 +17,7 @@ KEN11dK<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
     # quite similar, but not yet completely the same...
 
 anova(KEN11dK,type="marginal")
-
+exp(summary(KEN11dK)$coef[[1]])
 #---  -- --  --  -- -- -- -- -- -- -- - -- -- -- - -- - - - -- -- --  -- --  --  -- -- -- -- -- -- -- - -- -- -- - -- - - - -- -- --  -- --  --  -- -- -- -- -- -- -- - -- -- -- - -- - - - -- -- -
 # now ASAL and nonASAL separately
 
@@ -26,7 +26,7 @@ KEN11dK_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
              data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11dK_ASAL)
 
 anova(KEN11dK_ASAL,type="marginal")
-
+exp(summary(KEN11a)$coef[[1]])
 
 KEN11dK_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                   +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),

### 2018-12-19 writing/draft3/draft17.tex
working hard
writing
write tables
rewriting CV from SD (temperature)

@@
 \caption{\textit{\textbf{Mixed  effects model:} \\ Log of maize yield and weather, ARMA(1,1) errors}}
-%KEN11a  KEN11a_ASAL    KEN11a_nonASAL % in the file Models1.R in folder RCodes\NovemberNewVars												
+%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL % in the file KEN11d_finals2.R in folder RCodes\DecemberNew												
 \label{MainEst} 
@@
 \\
-\vspace{-0.2cm}Intercept&$\m0.259^{***}$&$19.916$&$\m0.243^{*}$&$5.230$&$\m0.344^{**}$&$10.061$\\
-  \\ \vspace{-0.2cm}Prec. total&$\m0.078^{*}$&$\05.402$&$\m0.006^{}$&$0.022$&$\m0.246^{***}$&$19.386$\\
+\vspace{-0.2cm}Intercept&$\m0.259^{***}$&$19.908$&$\m0.241^{*}$&$5.084$&$\m0.342^{**}$&$9.891$\\
+  \\ \vspace{-0.2cm}Prec. total&$\m0.078^{*}$&$\05.358$&$\m0.007^{}$&$0.028$&$\m0.244^{***}$&$19.140$\\
   \\
-  \vspace{-0.2cm}Prec. total sq.&$-0.028^{*}$&$\04.289$&$\m0.004$&$0.051$&$-0.128^{***}$&$23.747$\\
-    \\ \vspace{-0.2cm}Prec. c. of var.&$-0.079^{\bullet}$&$\03.277$&$-0.031$ &$0.246$&$-0.095$&$\02.231$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$-0.067^{*}$&$\04.810$&$-0.183^{**}$&$6.969$&$-0.012$&$\00.163$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.826$&$-0.157^{**}$&$8.065$&$-0.011$&$\00.096$\\
-  \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.127$&$-0.213^{*}$&$5.376$&$-0.130$&$\01.580$\\
-  \\  \vspace{-0.2cm}Temp. std. dev.&$\m0.042^{\bullet}$&$\03.125$&$\m0.038$&$0.558$&$\m0.057$ ${*}$&$\05.640$\\
+  \vspace{-0.2cm}Prec. total sq.&$-0.028^{*}$&$\04.263$&$\m0.003$&$0.037$&$-0.127^{***}$&$23.384$\\
+    \\ \vspace{-0.2cm}Prec. c. of var.&$-0.079^{\bullet}$&$\03.278$&$-0.031$ &$0.237$&$-0.097$&$\02.326$\\
+  \\  \vspace{-0.2cm}Dry spell -length&$-0.067^{*}$&$\04.800$&$-0.181^{**}$&$6.841$&$-0.012$&$\00.164$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.848$&$-0.156^{**}$&$8.032$&$-0.011$&$\00.087$\\
+  \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.010$&$-0.214^{*}$&$5.409$&$-0.128$&$\01.525$\\
+  \\  \vspace{-0.2cm}Temp. c. of var.&$\m0.042^{\bullet}$&$\02.961$&$\m0.032$&$0.355$&$\m0.058^{*}$&$\05.372$\\
   \\
@@
 
-\vspace{1cm}
 \textcolor{blue}{Maybe also a table with standard errors here}
 
-{
-\begin{threeparttable}
-\singlespacing
-\caption{\textit{\textbf{Mixed  effects model:} exponents of the coefficient estimates\\ Log of maize yield and weather, ARMA(1,1) errors}}
-%KEN11a  KEN11a_ASAL    KEN11a_nonASAL										
+{\begin{threeparttable}
+\normalsize
+\caption{\normalsize{\textbf{ {Mixed  effects model:}} exponents of the coefficient estimates Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL										
 \label{KenARe11_exponents} 
-\begin{footnotesize}
 \lineup
-\begin{indented}
-\item[]\begin{tabular}{@{}llrlllr} 
-\br   &\multicolumn{2}{c}{\textit{\textbf{All counties}}} &\multicolumn{2}{c}{\textit{\textbf{ASAL}}} &\multicolumn{2}{c}{\textit{\textbf{non-ASAL}}}\\
-    \multicolumn{1}{l}{\vspace{0.1cm}\textbf{Fixed effects:}}&{$exp(\beta)$}&F-value\tnote{a}% anova(KEN11a, type='marginal')
-    &{$exp(\beta)$}&F-value\tnote{a}&{$exp(\beta)$}&F-value\tnote{a}\\
+\begin{tabular}{@{}lllllll} 
+\br \\
+  \textbf{Fixed effects:}&\textit{\textbf{All counties}}&\textit{\textbf{ASAL}}&\textit{\textbf{non-ASAL}}\\
 \mr
 \\
-\vspace{-0.2cm}Intercept&$1.296^{***}$&$19.916$&$1.276^{*}$&$5.230$&$1.410^{**}$&$10.061$\\
-  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$\05.402$&$1.006^{}$&$0.022$&$1.278^{***}$&$19.386$\\
+\vspace{-0.2cm}Intercept&$1.296^{***}$&$1.272^{*}$&$1.408^{**}$\\
+  \\ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$1.007^{}$&$1.277^{***}$\\
   \\
-  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$\04.289$&$1.004$&$0.051$&$0.880^{***}$&$23.747$\\
-    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$\03.277$&$0.969$ &$0.246$&$0.909^{}$&$\02.231$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$\04.810$&$0.833^{**}$&$6.969$&$ 0.988^{}$&$\00.163$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$\04.826$&$0.855^{**}$&$8.065$&$0.989^{}$&$\00.096$\\
-  \\ \vspace{-0.2cm}Temp. - average&$0.819^{***}$&$12.127$&$0.808^{*}$&$5.376$&$0.878$ $^{}$&$\01.580$\\
-  \\  \vspace{-0.2cm}Temp. std. dev.&$1.043^{\bullet}$&$\03.125$&$1.039$&$0.558$&$1.059$ ${*}$&$\05.640$\\
+  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$1.003$&$0.881^{***}$\\
+    \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$0.970$ &$0.907^{}$\\
+  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
+  \\ \vspace{-0.2cm}Temp. - average&$0.820^{***}$&$0.808^{*}$&$0.880$ $^{}$\\
+  \\  \vspace{-0.2cm}Temp. c. of var.&$1.043^{\bullet}$&$1.032$&$1.060$ ${*}$\\
   \\
@@
 \end{tabular} 
-\end{indented}
-\end{footnotesize}
  \begin{tablenotes}
@@
     \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-        \begin{adjustwidth}{1cm}{} 
-    \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
-     \end{adjustwidth}
 \singlespacing
@@
    \includegraphics{Figure2a_2f.pdf}\label{MarEff2}
-\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and standard deviation (SD) of temperature (right column) on maize yields.
+\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and CV of temperature (right column) on maize yields.
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-CV of precipitation and SD of temperature (x-axis) are in multiples of their standard deviations.
+CV of precipitation and temperature (x-axis) are in multiples of their standard deviations.
 The effects are multiplicative as the models are in log-linear form.}

### 2018-12-19 Rcodes/DecemberNew/KEN11d_finals2.R

@@ -26,10 +26,12 @@ KEN11dK_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
              data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11dK_ASAL)
 
 anova(KEN11dK_ASAL,type="marginal")
-exp(summary(KEN11a)$coef[[1]])
+exp(summary(KEN11dK_ASAL)$coef[[1]])
+
 
 KEN11dK_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                   +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
                   data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11dK_nonASAL)
 
 anova(KEN11dK_nonASAL,type="marginal")
+exp(summary(KEN11dK_nonASAL)$coef[[1]])

### 2018-12-20 writing/draft3/draft17.tex
writing table of all measure

@@
 
-\textcolor{blue}{re-write the tables with estimates  UNITS KELVIN AND CV TIMES THOUSAND; possibly table of all available weather measure in the appendix, maybe a drop1 and add1 tables??}
-
 According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
@@
 
+\begin{table}[H]
+\caption{Precipitation and temperature measures considered}\label{vars}
+\begin{indented}
+\item[]\begin{tabular}{@{}lp{10cm}}
+\\[-1em]
+\br
+\\[-1em]
+$\bullet$& Total rainfall over the rainy season\\
+$\bullet$& Coefficient of variation of the rainfall during the rainy season\\
+$\bullet$& Maximum length of dry spell during the rainy season (in number of days)\\
+$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $4$ consecutive days without rain or more\\
+$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $10$ consecutive days without rain or more\\
+$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $20$ consecutive days without rain or more\\
+$\bullet$& Average temperature during the rainy season\\
+$\bullet$& Standard deviation of temperature during the rainy season\\
+$\bullet$& Cumulative degree days during the rainy season (excluding the days when maximum temperature above $30^{\circ}$C or below $10^{\circ}$C\\
+$\bullet$& Number of heatwave days during the rainy season when max. temperature $>35^{\circ}$C
+\\
+$\bullet$& Maximum daily rainfall  - to control for possible floods
+\\
+$\bullet$& Sum of precipitation amount on days where precipitation is above $90^{th}$ percentile of precip. of the whole period\\
+$\bullet$& Sum of precipitation amount on days where precipitation is above $95^{th}$ percentile of precip. of the whole period\\
+$\bullet$& Sum of precipitation amount on days where precipitation is above $99^{th}$ percentile of precip. of the whole period\\
+$\bullet$& Number of days where precipitation is above $90^{th}$ percentile of precip. of the whole period\\
+$\bullet$& Number of days where precipitation is above $95^{th}$ percentile of precip. of the whole period\\
+$\bullet$& Number of days where precipitation is above $99^{th}$ percentile of precip. of the whole period\\
+\br
+\end{tabular}
+\end{indented}
+
+\end{table}
+
 {\begin{center}

### 2018-12-20 writing/draft3/draft16.tex
\zvx\sgdca

@@
 
-In this study, we analyzed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
+In this study, we analysed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
 	We acquired the county level yearly yield data from the Famine Early Warning Systems Network (FEWS NET).
@@
 
-Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods.
+Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods (for the list of these measures see table~\ref{vars} in the appendix).
 The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, coefficient of variation of average seasonal temperature (converted into Kelvin because other scales of temperature such us Celsius are interval and coefficient of variation does not have any meaning for data measured on an interval scale),  maximum length of dry spell in number of days and the number of dry spells lasting for four days or more (we considered every day when precipitation didn't exceed $1$ mm as a dry day).

### 2018-12-20 writing/draft3/draft17.tex
\zvx\sgdca

--- 2018-12-19 writing/draft3/draft17.tex
@@
 This method led to the same model as the one obtained with our primary AIC-based method described above.
-
-\textcolor{blue}{re-write the tables with estimates  UNITS KELVIN AND CV TIMES THOUSAND; possibly table of all available weather measure in the appendix, maybe a drop1 and add1 tables??}
 
@@
 
+\begin{threeparttable}[H]
+\singlespacing
+\caption{Precipitation and temperature measures considered}\label{vars}
+\begin{tabular}{@{}lp{15cm}}
+\\[-1em]
+\br
+\\[-1em]
+$\bullet$& Total rainfall over the rainy season\\
+$\bullet$& Coefficient of variation of the rainfall during the rainy season\\
+$\bullet$& Maximum length of dry spell during the rainy season (in number of days)\\
+$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $4$ consecutive days without rain or more\tnote{a}\\
+$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $10$ consecutive days without rain or more\tnote{a}\\
+$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $20$ consecutive days without rain or more\tnote{a}\\
+$\bullet$& Average temperature during the rainy season\\
+$\bullet$& Standard deviation of temperature during the rainy season\\
+$\bullet$& Cumulative degree days during the rainy season (excluding the days when maximum temperature above $30^{\circ}$C or below $10^{\circ}$C\\
+$\bullet$& Number of heatwave days during the rainy season when max. temperature $>35^{\circ}$C
+\\
+$\bullet$& Maximum daily rainfall  - to control for possible floods
+\\
+$\bullet$& Sum of precipitation amount on days where precipitation is above $90^{th}$ percentile of precip. of the whole period\tnote{b}\\
+$\bullet$& Sum of precipitation amount on days where precipitation is above $95^{th}$ percentile of precip. of the whole period\tnote{c}\\
+$\bullet$& Sum of precipitation amount on days where precipitation is above $99^{th}$ percentile of precip. of the whole period\tnote{d}\\
+$\bullet$& Number of days where precipitation is above $90^{th}$ percentile of precip. of the whole period\tnote{b}\\
+$\bullet$& Number of days where precipitation is above $95^{th}$ percentile of precip. of the whole period\tnote{c}\\
+$\bullet$& Number of days where precipitation is above $99^{th}$ percentile of precip. of the whole period\tnote{d}\\
+\br
+\end{tabular}
+\begin{tablenotes}
+  \begin{footnotesize}
+  \item[a] Threshold for a dry day was considered $1$mm.
+  \item[b] $90^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
+\item[c] $95^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
+  \item[d] $99^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.
+  \end{footnotesize}
+\end{tablenotes}
+\end{threeparttable}
+
 {\begin{center}

### 2018-12-24 writing/draft3/draft17.tex
writing
correcting

@@
 
-\pdfminorversion=4
+% change the following at uni. 24.12.2018
+%\pdfminorversion=4
 \submitto{\ERL}
@@
 
-In this study, we analyzed the relationship between maize yields and climate. Our dataset consists of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
+In this study, we analysed the relationship between maize yields and climate. Our dataset consisted of an yearly panel of $47$ counties of Kenya describing the period of $1981-2017$. 
 	We acquired the county level yearly yield data from the Famine Early Warning Systems Network (FEWS NET).
@@
 In the non-ASAL counties, the single rainy season usually starts in March and lasts until August.
-Following the precipitation patterns and closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.
+Following the precipitation patterns and the closely related planting and harvesting calendar, we computed yearly values of various weather characteristics as follows: For the ASAL counties, we used daily data covering October, November and December of the previous year and March, April, May of the current year and for the non-ASAL counties we used daily data covering March to August of the current year.
 
-Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods (for the list of these measures see table~\ref{vars} in the appendix).
-The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, coefficient of variation of average seasonal temperature (converted into Kelvin because other scales of temperature such us Celsius are interval and coefficient of variation does not have any meaning for data measured on an interval scale),  maximum length of dry spell in number of days and the number of dry spells lasting for four days or more (we considered every day when precipitation didn't exceed $1$ mm as a dry day).
-We did not include squared values of the average seasonal temperature as it was not significant in our models and there is not much evidence of non-linear relationship between yield and temperature in the past literature.
+Following the procedure described above, we calculated a number of previously used characteristics of precipitation and temperature including indicators of floods.
+These measures are listed in table~\ref{vars} in the appendix.
+The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, coefficient of variation of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and the coefficient of variation does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$mm).
+We did not include squared values of the average seasonal temperature in our preferred models as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and temperature.
 \subsection{Statistical approach}\label{stats}
@@
 
-where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variable $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ a scale factor.
+where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variable $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ is a scale factor.
 The distribution of $\mathscr{B}$ can be written as:
@@
 It was apparent from the histogram of the maize yield data that the distribution of maize yield was closer to a log-normal than to a normal distribution.
-Therefore, we opted for a log-linear functional form of our models as is a common practice in agricultural economics.
+Therefore, we opted for a log-linear functional form of our models as is a common practice in the agricultural economics.
 Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.
@@
 The first step of the procedure was the estimation of the model with the complete set of our weather measures in the fixed effects and with the random intercepts.
-After this, we performed a search through the subsets of the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding predictors).
+After this, we performed a search through the subsets of the fixed effects minimizing the AIC and allowing for steps in both `backward' and `forward' directions (that is removing and adding the predictors).
 We decided not to include number of heatwave days and average maximum daily temperature although including them as fixed effects would slightly improve the AIC.
-The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
-Furthermore, including the heatwave days and maximum daily temperature would only lead to small change in AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for comparison\footnote{In general, REML is the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however, REML criterion is not meaningful for comparison of the models with different fixed effects.
-Therefore, the models had to be re-estimated by ML for comparison \cite{Zuur2009}.}).
+The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in the case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
+Furthermore, including heatwave days and maximum daily temperature would only lead to a small change in the AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for comparison\footnote{REML is generally the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however, the REML criterion is not meaningful for comparison of the models with different fixed effects.
+Therefore, we re-estimated the models by ML for the purpose of comparison \cite{Zuur2009}.}).
 In addition, the coefficients of both heatwave days and maximum daily temperature were positive, which is opposite to what we expected.
-Hence, we concluded that the decrease in AIC and the positive signs are likely to be results of the correlation with average temperature and we did not include them in our preferred model.
+Hence, we concluded that the decrease in the AIC and the positive signs are likely to be results of the high correlation with average temperature and we did not include them in our preferred specification.
 
-We further applied a series of likelihood ratio tests to test for significance of random slopes.
-No subset of the random slopes improved the fit, therefore besides the fixed effects, our final model only included the random intercepts.\footnote{Although adding the random slope of the coefficient of variation slightly improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
+We further applied a series of likelihood ratio tests to test for the significance of random slopes.
+No subset of the random slopes improved the fit; therefore, besides the fixed effects, our final model only included the random intercepts.\footnote{Although adding the random slope of the coefficient of variation improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
 The reason for not including the random slope is that the difference in AIC is negligible and the F-test of the comparison of the two models is insignificant favouring the simpler model (p-value$=0.128$).}
@@
 In the first part of the procedure, the insignificant random effects were removed one by one.
-In each iteration, the insignificant variable with the highest p-value was removed .
+In each iteration, the insignificant variable with the highest p-value was removed.
 This was repeated until only the significant predictors remained.
-After this, we applied an analogous procedure to the fixed effects until we were left only with the significant ones.
+After this, we applied an analogous procedure to the fixed effects until we were left only with the significant variables.
 As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$~ for fixed effects \cite{lmerTest}.

### 2018-12-24 writing/draft3/draft17.tex
writing
at least trying to

@@
 
-Besides our main model, we estimated the same specification for two additional subsamples of Kenyan counties.
-In particular, we estimated the model for the subsample of ASAL counties and for the subsample of non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate.
+Besides our main model, we estimated the same specification for two additional subsamples of the Kenyan counties.
+In particular, we estimated the model for the subsample of the ASAL counties and for the subsample of the non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with the significantly different climate.
 The estimates of all three models are summarised in table~\ref{MainEst}.
-The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of ASAL counties and the last two columns describe the estimates based on the subsample of non-ASAL counties.
+The first two columns represent the model for all counties, the third and fourth columns include estimates for the subsample of the ASAL counties and the last two columns describe the estimates based on the subsample of the non-ASAL counties.
 All predictors were standardised, therefore the estimates and the corresponding p-values can be interpreted as measures of relative importance.
-
 As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance (ANOVA) in table~\ref{MainEst}.
-The F-tests were conducted using the marginal rather than sequential sum of squares.
+The F-tests were conducted using the marginal rather than the sequential sum of squares.
 That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable.
 In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.
-The type~\textit{III} ANOVA does not depend on the order in which the effects are entered in the model.
+The main advantage of the type~\textit{III} ANOVA is that the test statistics do not depend on the order in which the effects are entered in the model.
 
-The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and the effects on yield are multiplicative.
-The exponents of the estimates (which can be interpreted as additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.
+The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects on yield are multiplicative.
+The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix.
+The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised.
+The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.
 
-First, we will focus on the estimates based on the sample of all counties, that is the first two columns in table~\ref{MainEst}.
-Consistently with our assumptions, seasonal precipitation has mostly positive and significant effect on maize yield while average temperature has negative effects on yields.
+In the following, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.
 
@@
 
-We can see in table~\ref{MainEst} that the squared seasonal precipitation is negative while the linear precipitation is posititve,
-
-affects the
+The estimates of the model based on all the Kenyan counties are summarised in the first two columns of table~\ref{MainEst}.
+Consistently with our assumptions, seasonal precipitation and average temperature have significant effect on yields.
+As we expected, the effect of temperature is negative.
+The linear term of precipitation is positive while the squared term is negative and as we can see in figure~\ref{MarEff1}(a), the marginal effect of precipitation is hill-shaped
 
@@
   \begin{figure}% new plots checked and good 17.dec.2018
-   \includegraphics{Figure1a_1f.pdf}\label{MarEff1}
-\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields.
+   \includegraphics{Figure1a_1f.pdf}
+\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-Precipitation and temperature (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
+Precipitation and temperature (x-axis) are in multiples of their standard deviations as the models were standardised.
+The effects are multiplicative as the models are in the log-linear form.}\label{MarEff1}
 \end{figure}
@@
   \begin{figure}
-   \includegraphics{Figure2a_2f.pdf}\label{MarEff2}
-\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and CV of temperature (right column) on maize yields.
+   \includegraphics{Figure2a_2f.pdf}
+\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and CV of temperature (right column) on maize yields assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-CV of precipitation and temperature (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
+CV of precipitation and temperature (x-axis) are in multiples of their standard deviations as the models were standardised.
+The effects are multiplicative because the models are in the log-linear form.}\label{MarEff2}
 \end{figure}
@@
   \begin{figure}
-   \includegraphics{Figure3a_3f.pdf}\label{MarEff3}
-\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yields.
+   \includegraphics{Figure3a_3f.pdf}
+\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yields assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
-Maximum length of dry spell and number of dry spells (x-axis) are in multiples of their standard deviations.
-The effects are multiplicative as the models are in log-linear form.}
+Maximum length of dry spell and number of dry spells (x-axis) are in multiples of their standard deviations as the models were standardised.
+The effects are multiplicative because the models are in the log-linear form.}\label{MarEff3}
 \end{figure}

### 2018-12-24 Rcodes/DecemberNew/KEN11d_finals2.R

@@ -1,5 +1,5 @@
 rm(list=ls())
-library(MASS);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 load("dataFS/Main/DaTS.RData")
 # load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo

### 2018-12-24 writing/draft3/draft17.tex
writing

@@
 In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.
-The main advantage of the type~\textit{III} ANOVA is that the test statistics do not depend on the order in which the effects are entered in the model.
+The main advantage of the type~\textit{III} ANOVA is that the test statistic does not depend on the order in which the effects are entered in the model.
 
-The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects on yield are multiplicative.
+The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects are multiplicative.
 The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix.
 The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised.
+If the values of the marginal effects are smaller than one, the effect of yield is negative while marginal effects higher than one can be interpreted as positive effect.
 The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.
 
-In the following, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.
+In the following few paragraphs, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.
 
@@
 As we expected, the effect of temperature is negative.
-The linear term of precipitation is positive while the squared term is negative and as we can see in figure~\ref{MarEff1}(a), the marginal effect of precipitation is hill-shaped
+The linear term of precipitation is positive while the squared term is negative and as we can see in figure~\ref{MarEff1}(a), the marginal effect of precipitation is hill-shaped.
+That is, the effect of precipitation is negative and increasing if its amount is below mean then the effect further increases with the amount of precipitation until it reaches $1.403$ times its standard deviation.
+At this point the effect starts to decrease with increasing precipitation and it became negative again once the amount reaches $2.805$ times its standard deviation.
 

### 2018-12-24 writing/draft3/draft18.tex
writing

--- 2018-12-24 writing/draft3/draft17.tex
@@
 \usepackage{harvard}
-
+\usepackage{soul}
 \usepackage[flushleft]{threeparttable}
@@
 
-\title{The Effects of weather on maize yields: New evidence from Kenya}
+\title{The Effects of weather on maize yield: New evidence from Kenya}
 
@@
 These measures are listed in table~\ref{vars} in the appendix.
-The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation of seasonal precipitation, coefficient of variation of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and the coefficient of variation does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$mm).
+The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation (CV) of seasonal precipitation, CV of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and CV does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$mm).
 We did not include squared values of the average seasonal temperature in our preferred models as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and temperature.
@@
 We further applied a series of likelihood ratio tests to test for the significance of random slopes.
-No subset of the random slopes improved the fit; therefore, besides the fixed effects, our final model only included the random intercepts.\footnote{Although adding the random slope of the coefficient of variation improved the AIC from $2122.226$ to $2122.110$, we decided to not to include the coefficient of variation of temperature in the random effects.
-The reason for not including the random slope is that the difference in AIC is negligible and the F-test of the comparison of the two models is insignificant favouring the simpler model (p-value$=0.128$).}
+No subset of the random slopes improved the fit; therefore, besides the fixed effects, our final model only included the random intercepts.\footnote{Although adding the random slope of CV improved the AIC from $2122.226$ to $2122.110$, we decided to not to include CV of temperature in the random effects.
+The reason for not including the random slope is that the difference in AIC was negligible and the F-test of the comparison of the two models was insignificant favouring the simpler model (p-value$=0.128$).}
 
@@
 The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised.
-If the values of the marginal effects are smaller than one, the effect of yield is negative while marginal effects higher than one can be interpreted as positive effect.
+For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as positive effect.
 The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.
@@
     \\ \vspace{-0.2cm}Prec. c. of var.&$-0.079^{\bullet}$&$\03.278$&$-0.031$ &$0.237$&$-0.097$&$\02.326$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$-0.067^{*}$&$\04.800$&$-0.181^{**}$&$6.841$&$-0.012$&$\00.164$\\
+  \\  \vspace{-0.2cm}Dry spell -length\tnote{b}&$-0.067^{*}$&$\04.800$&$-0.181^{**}$&$6.841$&$-0.012$&$\00.164$\\
   \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.848$&$-0.156^{**}$&$8.032$&$-0.011$&$\00.087$\\
@@
     \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
+  \item[b] In number of days.
      \end{adjustwidth}
@@
 
-The estimates of the model based on all the Kenyan counties are summarised in the first two columns of table~\ref{MainEst}.
-Consistently with our assumptions, seasonal precipitation and average temperature have significant effect on yields.
+The estimates of the model of the whole sample are summarised in the first two columns of table~\ref{MainEst}.
+Consistently with our assumptions, seasonal precipitation and average temperature have significant effects on yield.
 As we expected, the effect of temperature is negative.
-The linear term of precipitation is positive while the squared term is negative and as we can see in figure~\ref{MarEff1}(a), the marginal effect of precipitation is hill-shaped.
-That is, the effect of precipitation is negative and increasing if its amount is below mean then the effect further increases with the amount of precipitation until it reaches $1.403$ times its standard deviation.
-At this point the effect starts to decrease with increasing precipitation and it became negative again once the amount reaches $2.805$ times its standard deviation.
+The linear term of precipitation is positive while the squared term of precipitation is negative and as we can see in figure~\ref{MarEff1}(a), the plot of the marginal effects of precipitation is hill-shaped.
+That is, the effect of precipitation is negative and increasing if the amount of precipitation is below its mean and the marginal effect increases with the amount of precipitation until it reaches $1.403$ times its standard deviation.
+At this point, the effect starts to decrease with increasing precipitation and it becomes negative again once the amount of precipitation reaches $2.805$ times its standard deviation.
+Therefore, too  much rain is harmful for agricultural yield as well as not enough rain.
+This is intuitive as extreme rainfall can cause flooding which is damaging for yield.
 
-Besides seasonal precipitation and average temperature, the length and number of dry spells
+Besides total cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and the length and number of dry spells turned out to be significant (see the first column of table~\ref{MainEst}).
+The relationship of maize yield and each of these variables is linear as is apparent from the figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix.
+In accordance with our expectation, the yield decreases with increasing CV of precipitation.
+\hl{However, the result for CV of temperature is a bit peculiar as the yield seem to increase with increasing variability of temperature.
+This is probably because...}
+As it is apparent from figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with length of maximum dry spell linearly.
+This is in accordance with our expectations.
+
+According to both magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable of the model for all counties is average seasonal temperature.
+Average seasonal temperature has the lowest p-value of the predictors in this model and based on its magnitude and F-value it is much  more important than the second most important variable, which is precipitation.
+
+We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the third and fourth columns of table~\ref{MainEst}).
 
@@
    \includegraphics{Figure1a_1f.pdf}
-\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yields assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
+\caption{Predicted multiplicative marginal effects of seasonal precipitation (left column) and average seasonal temperature (right column) on maize yield assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
@@
 $\bullet$& Maximum length of dry spell during the rainy season (in number of days)\\
-$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $4$ consecutive days without rain or more\tnote{a}\\
-$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $10$ consecutive days without rain or more\tnote{a}\\
-$\bullet$& Number of drought spells during the rainy season: a drought spell defined as $20$ consecutive days without rain or more\tnote{a}\\
+$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $4$ consecutive days without rain or more\tnote{a}\\
+$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $10$ consecutive days without rain or more\tnote{a}\\
+$\bullet$& Number of dry spells during the rainy season: a dry spell defined as $20$ consecutive days without rain or more\tnote{a}\\
 $\bullet$& Average temperature during the rainy season\\
@@
    \includegraphics{Figure2a_2f.pdf}
-\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and CV of temperature (right column) on maize yields assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
+\caption{Predicted multiplicative marginal effects of coefficient of variation (CV) of precipitation (left column) and CV of temperature (right column) on maize yield assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.
@@
    \includegraphics{Figure3a_3f.pdf}
-\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yields assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
+\caption{Predicted multiplicative marginal effects of length of maximum dry spell in days (left column) and number of dry spells lasting for four days or more (right column) on maize yield assuming that all other variables are fixed at their means (which is zero, as the predictors were standardised).
 The first row (a, b) represents the model for all counties, the second row (c, d) is based on the subsample of arid and semi-arid counties (ASAL) and the third row (e, f) represents the model for the non-ASAL counties.

### 2018-12-24 Rcodes/December18/ggeffectsImages.R

@@ -101,7 +101,7 @@ nonASALSpell4[,c(2,4,5)]<-nonASALSpell4[,c(2,4,5)]/exp(summary(KEN11d_nonASAL)$t
 plot1<-plot(AllSpell)+ylab("All")+xlab(NULL)+xlab("(a)")+ggtitle("Effects of length of max dry spell on yields")+theme(plot.title=element_text(size=11))+scale_x_continuous(breaks = c(-2,0,2,4,6,8))+scale_y_continuous(breaks = c(0.4,0.6,0.8,1,1.2))
 plot2<-plot(AllSpell4)+ylab(NULL)+xlab(NULL)+xlab("(b)")+ggtitle("Effects of number of dry spells on yields")+theme(plot.title=element_text(size=11))+scale_x_continuous(breaks = c(-3,-2,-1,0,1,2))
 plot3<-plot(ASALSpell)+ylab("ASAL")+xlab(NULL)+xlab("(c)")+scale_x_continuous(breaks = c(-2,-1,0,1,2,3,4,5))+ggtitle(NULL)+ggtitle(NULL)+scale_y_continuous(breaks = c(0.5,1,1.5))
-plot4<-plot(ASALSpell4)+ylab(NULL)+xlab(NULL)+xlab("(d)")
+plot4<-plot(ASALSpell4)+ylab(NULL)+xlab(NULL)+ggtitle(NULL)+xlab("(d)")
 plot5<-plot(nonASALSpell)+ylab("non-ASAL")+xlab(NULL)+xlab("(e)")+ggtitle(NULL)+scale_x_continuous(breaks = c(-2,0,2,4,6,8))
 plot6<-plot(nonASALSpell4)+ylab(NULL)+xlab(NULL)+xlab("(f)")+ggtitle(NULL)+scale_x_continuous(breaks = c(-3,-2,-1,0,1,2))
 

### 2018-12-24 writing/draft3/draft18.tex
working

@@
 
-Besides total cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and the length and number of dry spells turned out to be significant (see the first column of table~\ref{MainEst}).
-The relationship of maize yield and each of these variables is linear as is apparent from the figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix.
+Besides cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and also length and number of dry spells were significant (see the first column of table~\ref{MainEst}).
+As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear.
 In accordance with our expectation, the yield decreases with increasing CV of precipitation.
-\hl{However, the result for CV of temperature is a bit peculiar as the yield seem to increase with increasing variability of temperature.
+\hl{However, we obtained a peculiar result for CV of temperature as the yield seem to increase with increasing variability of temperature.
 This is probably because...}
-As it is apparent from figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with length of maximum dry spell linearly.
+
+As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with length of maximum dry spell linearly.
 This is in accordance with our expectations.
@@
 According to both magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable of the model for all counties is average seasonal temperature.
-Average seasonal temperature has the lowest p-value of the predictors in this model and based on its magnitude and F-value it is much  more important than the second most important variable, which is precipitation.
+The temperature appears to be much more important than the second most important variable in this model which is cumulative precipitation.
+The value of the F-statistic corresponding to temperature is $12.010$ while the F-statistic of the linear term of cumulative precipitation is equal to $5.358$ and the F-statistic of its quadratic term is $4.263$.
+Also, according to its lowest p-value ($6\times10^{-4}$), temperature is the most important covariate.
 
-We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the third and fourth columns of table~\ref{MainEst}).
+We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
 

### 2018-12-25 writing/draft3/draft18.tex
write and editing...

@@
 In accordance with our expectation, the yield decreases with increasing CV of precipitation.
-\hl{However, we obtained a peculiar result for CV of temperature as the yield seem to increase with increasing variability of temperature.
+\hl{However, the result which we found for CV of temperature is a bit peculiar as the yield seem to increase with increasing variability of temperature.
 This is probably because...}
 
-As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with length of maximum dry spell linearly.
+As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with increasing length of maximum dry spell linearly.
 This is in accordance with our expectations.
 
-According to both magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable of the model for all counties is average seasonal temperature.
-The temperature appears to be much more important than the second most important variable in this model which is cumulative precipitation.
+According to the magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable is average seasonal temperature which appears to be much more important than the remaining predictors.
+The second most important variable is total seasonal precipitation.
 The value of the F-statistic corresponding to temperature is $12.010$ while the F-statistic of the linear term of cumulative precipitation is equal to $5.358$ and the F-statistic of its quadratic term is $4.263$.
-Also, according to its lowest p-value ($6\times10^{-4}$), temperature is the most important covariate.
+Similarly, according to the Satterthwaite's p-values, average temperature is the most important predictor as the corresponding Satterthwaite's p-value is equal to $6\times10^{-4}$ which is much lower than the p-values of the other explanatory variables.
 
 We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
+As for the subsample of the ASAL counties, the only significant fixed effects included average seasonal temperature and length and number of dry spells, all of which were negative.
+Interestingly, different predictors were  found significant for the subsample of the non-ASAL counties.
+These included seasonal precipitation and its square (the first one was positive and the second one negative..\textcolor {red}{here}.
+)
 
@@
 \\[-1em]
-$\bullet$& Total rainfall over the rainy season\\
-$\bullet$& Coefficient of variation of the rainfall during the rainy season\\
+$\bullet$& Total precipitation over the rainy season\\
+$\bullet$& Coefficient of variation of the precipitation during the rainy season\\
 $\bullet$& Maximum length of dry spell during the rainy season (in number of days)\\
@@
 \\
-$\bullet$& Maximum daily rainfall  - to control for possible floods
+$\bullet$& Maximum daily precipitation  - to control for possible floods
 \\

### 2018-12-25 writing/draft3/draft18.tex
writing

@@
 The units of all predictors in all three figures are multiples of their standard deviations since the predictors were standardised.
-For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as positive effect.
+For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as a positive effect.
+If, for example, a marginal effect of an explanatory variable $x$ was $1.1$ when $x$ is equal to its one standard deviation, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).
+
 The exponents of the estimates (which can be interpreted as the additive marginal effects on yield) can be found in table~\ref{KenARe11_exponents} in the appendix.
@@
 Therefore, too  much rain is harmful for agricultural yield as well as not enough rain.
-This is intuitive as extreme rainfall can cause flooding which is damaging for yield.
+This is intuitive mainly because extreme rainfall can cause flooding which is damaging for yield.
 
-Besides cumulative seasonal precipitation and average temperature, CV of both temperature and precipitation and also length and number of dry spells were significant (see the first column of table~\ref{MainEst}).
+Other fixed effects which we found significant include CV of temperature and precipitation and length and number of dry spells (see the first column of table~\ref{MainEst}).
 As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear.
 In accordance with our expectation, the yield decreases with increasing CV of precipitation.
-\hl{However, the result which we found for CV of temperature is a bit peculiar as the yield seem to increase with increasing variability of temperature.
+\hl{However, the result which we found for CV of temperature is a bit puzzling as the yield seem to increase with increasing variability of temperature.
 This is probably because...}
@@
 We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
-As for the subsample of the ASAL counties, the only significant fixed effects included average seasonal temperature and length and number of dry spells, all of which were negative.
-Interestingly, different predictors were  found significant for the subsample of the non-ASAL counties.
-These included seasonal precipitation and its square (the first one was positive and the second one negative..\textcolor {red}{here}.
-)
+Interestingly, different predictors were  found significant for the two subsamples.
+Regarding the subsample of the ASAL counties, the only significant fixed effects are average seasonal temperature and length and number of dry spells.
+All three variables have a negative sign.
+On the other hand, the variables which were found significant for the subsample of the non-ASAL counties include seasonal precipitation and its square with a positive and a negative sign, respectively and CV of temperature which has a positive sign as in the case of the entire sample.
+According to the F-statistic, the most important explanatory variables for the ASAL counties are number of dry spells with F-statistic equal to $8.032$ and length of the longest dry spell with F-value equal to $6.841$.
+The F-statistic of the average temperature is $5.409$ and it lower than zero for the remaining predictors.
+However, according to the magnitude of the standardised coefficient, the average temperature would be the most important as the coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
+With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its square.
+The F-statistic of the squared values of precipitation is $23.384$ and its value for the linear term of precipitation is $19.140$, while the variable with the third highest F-statistic is CV of temperature with the F-statistic equal to $5.372$, hence, the difference in importance of the precipitation and CV of temperature is relatively large.
+The standardised coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised coefficient of CV of temperature is $0.058$.
+In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of the coefficient is $-0.128$.
+However, average temperature is insignificant according to the Satterthwaite's p-value and its F-statistic is also relatively small, in particular $1.528$.
+
+To sum up, in the ASAL counties temperature is much more important than in the non-ASAL counties where temperature is not significant.
+On the other hand, precipitation is not significant in the ASAL counties while it is strongly significant in the non-ASAL counties.
+This is illustrated by figure~\ref{MarEff1}.
+The plot of the marginal effects of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(b)) while is steep and hill-shaped for the non-ASAL counties similarly as in the case of the whole sample.
+As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f), the marginal effects of temperature are linear and decreasing for the case of ASAL and non-ASAL counties.
+Marginal effects of the remaining predictors of the models for the two subsamples are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
 

### 2018-12-25 writing/draft3/draft18.tex
writing

@@
 
-The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and their effects are multiplicative.
+The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as the marginal effects and the coefficients are multiplicative.
 The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix.
@@
 As we expected, the effect of temperature is negative.
-The linear term of precipitation is positive while the squared term of precipitation is negative and as we can see in figure~\ref{MarEff1}(a), the plot of the marginal effects of precipitation is hill-shaped.
+The linear term of precipitation is positive while the squared term of precipitation is negative and as we can see in figure~\ref{MarEff1}(a), the plot of the marginal effect of precipitation is hill-shaped.
 That is, the effect of precipitation is negative and increasing if the amount of precipitation is below its mean and the marginal effect increases with the amount of precipitation until it reaches $1.403$ times its standard deviation.
@@
 
-To sum up, in the ASAL counties temperature is much more important than in the non-ASAL counties where temperature is not significant.
-On the other hand, precipitation is not significant in the ASAL counties while it is strongly significant in the non-ASAL counties.
-This is illustrated by figure~\ref{MarEff1}.
-The plot of the marginal effects of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(b)) while is steep and hill-shaped for the non-ASAL counties similarly as in the case of the whole sample.
-As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f), the marginal effects of temperature are linear and decreasing for the case of ASAL and non-ASAL counties.
-Marginal effects of the remaining predictors of the models for the two subsamples are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
+To sum up, for the ASAL counties temperature is much more important than for the non-ASAL counties where temperature is not significant.
+On the other hand, precipitation is not significant for the ASAL counties while it is strongly significant for the non-ASAL counties.
+This is illustrated in figure~\ref{MarEff1}(c)-(f) in this section.
+The plot of the marginal effect of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(c) in this section) while it is hill-shaped and steep for the non-ASAL counties (figure \ref{MarEff1}(e) in this section) similarly as in the case of the whole sample.
+As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the case of ASAL and non-ASAL counties.
+Marginal effects of the remaining predictors are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
 
@@
 \end{figure}
+
+Adjusted $R^2$ is not available for the linear mixed models.
+To get an approximation of the adjusted $R^2$, we estimated ordinary least squares (OLS) with the same explanatory variables which were included in our preferred linear mixed models as the fixed effects (see the list of variables in table~\ref{MainEst}).
+For each of the three samples (all counties, ASAL and non-ASAL) we estimated two variants of the OLS, one with the dummy variables for the counties and one without them.
+The values of the adjusted $R^2$ are summarasied in table~\ref{R2}.
+
+\begin{threeparttable}
+\singlespacing
+\caption{Alternative to adjusted R2}\label{R2}
+\begin{tabular}{@{}lll}
+\\[-1em]
+\br
+\\[-1em]
+&\multicolumn{2}{c}{\textbf{County level dummy variables:}}\\
+\textbf{Sample:}&Yes&No\\
+\textbf{All counties}& $0.507$&$0.307$\\
+\textbf{ASAL}& $0.379$&$0.170$\\
+\textbf{non-ASAL}&  $0.501$&$0.307$\\
+\br
+\end{tabular}
+
+\end{threeparttable}
+
+To check for presence of multicollinearity we calculated the variance inflation factors (VIF) of the explanatory variables.
+With only two exceptions, the VIF were below $2$ for all predictors in our final models.
+For CV of precipitation, the value of VIF was $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and $2.182$ in the model for the subsample of the non-ASAL counties.
+These values are still well below the threshold of problematic multicollinearity.
+The other exception was the seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably as a result of the presence of the squared of precipitation.
+The VIF is below $2$ in the other two models and even the value $2.407$ is below the threshold of the problematic multicollinearity.
+Therefore, we can conclude that the models do not suffer from presence of multicollinearity.
 

### 2018-12-25 Rcodes/December18/R2.R

@@ -1,0 +1,60 @@
+rm(list=ls())
+library(ggeffects);library(ggplot2);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+# I am oing to estimate linear regression to get some estimates of R^2
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d);vif(KEN11d)
+# and subsamples
+KEN11d_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                 +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11d_ASAL);vif(KEN11d_ASAL)
+
+KEN11d_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11d_nonASAL);vif(KEN11d_nonASAL)
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
[…]

### 2018-12-25 Rcodes/December18/ggeffectsImages.R

@@ -7,15 +7,15 @@ load("dataFS/Main/DaTS.RData")
 
 KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
-            data=ScaledTS,na.action=na.exclude); summary(KEN11d)
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d);vif(KEN11d)
 # and subsamples
 KEN11d_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                  +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11d_ASAL)
+                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11d_ASAL);vif(KEN11d_ASAL)
 
 KEN11d_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
                     +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
-                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11d_nonASAL)
+                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11d_nonASAL);vif(KEN11d_nonASAL)
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 exp(summary(KEN11d)$tTable[,1]); exp(summary(KEN11d_ASAL)$tTable[,1]);exp(summary(KEN11d_nonASAL)$tTable[,1])
 

### 2018-12-26 writing/draft3/draft19.tex
(re)writing

--- 2018-12-25 writing/draft3/draft18.tex
@@
 
+\begin{center}
 \begin{threeparttable}
-\singlespacing
-\caption{Alternative to adjusted R2}\label{R2}
-\begin{tabular}{@{}lll}
+\caption{\label{R2}Coefficient of determination $R^2$ for mixed models}
+\begin{indented}
+\item[]
+\begin{tabular}{@{}lcc}
 \\[-1em]
@@
 \\[-1em]
-&\multicolumn{2}{c}{\textbf{County level dummy variables:}}\\
-\textbf{Sample:}&Yes&No\\
-\textbf{All counties}& $0.507$&$0.307$\\
-\textbf{ASAL}& $0.379$&$0.170$\\
-\textbf{non-ASAL}&  $0.501$&$0.307$\\
+\textbf{Sample:}&Marginal\tnote{a}&Conditional\tnote{b}\\
+\hline
+{All counties}& $0.300$&$0.526$\\
+{ASAL}& $0.134$&$0.353$\\
+{non-ASAL}&  $0.153$&$0.436$\\
 \br
 \end{tabular}
+\begin{tablenotes}
+\begin{footnotesize}
+    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
+        \begin{adjustwidth}{1cm}{}
+    \item[a] The marginal coefficient is interpreted as the variance explained by the fixed effects
+  \item[b] while the conditional one is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts
+     \end{adjustwidth}
+     \end{footnotesize}
+\end{tablenotes}
 
+\end{indented}
 \end{threeparttable} 
+\end{center}
+\vspace{1cm}
 

### 2018-12-26 Rcodes/December18/R2better.R

@@ -1,0 +1,34 @@
+rm(list=ls())
+library(MuMIn)
+library(ggeffects);library(ggplot2);library(dplyr); library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+# I am oing to estimate linear regression to get some estimates of R^2
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1) ,
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d);vif(KEN11d)
+r.squaredGLMM(KEN11d)
+# test....
+
+        aa<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~1 | ID1 ,
+                    data=ScaledTS,na.action=na.exclude)
+        r.squaredGLMM(aa)
+
+# and subsamples
+KEN11d_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                 +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                 data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11d_ASAL);vif(KEN11d_ASAL)
+r.squaredGLMM(KEN11d_ASAL)
+
+
+KEN11d_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11d_nonASAL);vif(KEN11d_nonASAL)
[…]

### 2018-12-26 Rcodes/December18/playing.R

@@ -1,0 +1,2 @@
+aa<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1 

### 2018-12-26 writing/draft3/draft19.tex
writing..

@@
 
-Adjusted $R^2$ is not available for the linear mixed models.
-To get an approximation of the adjusted $R^2$, we estimated ordinary least squares (OLS) with the same explanatory variables which were included in our preferred linear mixed models as the fixed effects (see the list of variables in table~\ref{MainEst}).
-For each of the three samples (all counties, ASAL and non-ASAL) we estimated two variants of the OLS, one with the dummy variables for the counties and one without them.
-The values of the adjusted $R^2$ are summarasied in table~\ref{R2}.
+The AIC is an important criterion for comparison of competing models, however it does not provide any information about the absolute goodness of fit.
+In order to assess the absolute goodness of fit of a model, a different measure is needed.
+For a simple linear regression, the coeffiecient of determination $R^2$ or its adjusted variant is typically used for this purpose.
+Nevertheless, obtaining the $R^2$ or the adjusted $R^2$ is not straightforward for linear mixed models or for models estimated by the maximum likelihood in general.
+In this study, we used the coefficient of determination of generalized linear mixed-effects models introduced by \cite{Nakagawa2013}.
+For each of our samples (all counties, ASAL and non-ASAL), we present two versions of the statistic in table~\ref{R2}.
+These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects components and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects.
+In this case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts.
+As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties.
 

### 2018-12-26 writing/draft3/draft19.tex
writ

@@
 
-We will now discuss and compare the results of the models for the subsamples of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
-Interestingly, different predictors were  found significant for the two subsamples.
+We will now discuss and compare the results of the models based on the subsample of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the model based on the subsample of the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
+Interestingly, different predictors were found significant for each of the two subsamples.
 Regarding the subsample of the ASAL counties, the only significant fixed effects are average seasonal temperature and length and number of dry spells.
-All three variables have a negative sign.
-On the other hand, the variables which were found significant for the subsample of the non-ASAL counties include seasonal precipitation and its square with a positive and a negative sign, respectively and CV of temperature which has a positive sign as in the case of the entire sample.
-According to the F-statistic, the most important explanatory variables for the ASAL counties are number of dry spells with F-statistic equal to $8.032$ and length of the longest dry spell with F-value equal to $6.841$.
-The F-statistic of the average temperature is $5.409$ and it lower than zero for the remaining predictors.
-However, according to the magnitude of the standardised coefficient, the average temperature would be the most important as the coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
+According to our expectation, all three variables have a negative sign.
+On the other hand, the variables which were found significant for the subsample of the non-ASAL counties include seasonal precipitation (positive) and its square (negative) and CV of temperature (positive).
+According to the F-statistics, the most important explanatory variables for the ASAL counties are number of dry spells and length of the longest dry spell with values of the F-statistics equal to $8.032$ and $6.841$, respectively.
+The F-statistic of the average temperature is $5.409$ and the F-statistics of the remaining predictors are lower than zero.
+However, according to the magnitude of the standardised coefficients, the average temperature would be the most important as the coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
 With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its square.

### 2018-12-26 writing/draft3/draftMonika.tex
writing

--- 2018-12-26 writing/draft3/draft19.tex
@@
 \begin{indented}
-\item[]November 2018
+\item[]December 2018
 \end{indented}
@@
 The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation (CV) of seasonal precipitation, CV of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and CV does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$mm).
-We did not include squared values of the average seasonal temperature in our preferred models as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and temperature.
+We did not include squared values of the average seasonal temperature in our preferred models as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and average temperature.
 \subsection{Statistical approach}\label{stats}
@@
 
-where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variable $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ is a scale factor.
+where $\mathbf{X}$ is an $n \times p$ model matrix of fixed effects, $\mathbf{\beta}$ is a $p$-dimensional fixed-effects parameter, $\mathbf{Z}$ is an $n \times q$ model matrix for the $q$-dimensional vector of random-effects variables $\mathscr{B}$ evaluated at $\mathbf{b}$ and $\sigma$ is a scale factor.
 The distribution of $\mathscr{B}$ can be written as:
@@
 
-It was apparent from the histogram of the maize yield data that the distribution of maize yield was closer to a log-normal than to a normal distribution.
+It was apparent from the histogram of the maize yield that the distribution of the yield data was closer to a log-normal than to a normal distribution.
 Therefore, we opted for a log-linear functional form of our models as is a common practice in the agricultural economics.
-Furthermore, we scaled all predictors by subtracting mean and dividing them by standard deviation to avoid convergence problems.
+Furthermore, we scaled all predictors by subtracting their mean and dividing them by their standard deviation to avoid convergence problems during estimation procedures.
 
@@
 The reasons for not including them were their strong correlation with average seasonal temperature (the correlation coefficient was $0.981$ in the case of maximum daily temperature and $0.517$ for the number of heatwave days with the p-values smaller than $1\times10^{-6}$ in both cases), high variance inflation factor (VIF) of the maximum temperature ($5.397$) and the lack of significance of their individual t-tests (the p-values were $0.121$ for the maximum daily temperature and $0.075$ for the number of heatwave days).
-Furthermore, including heatwave days and maximum daily temperature would only lead to a small change in the AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were estimated using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for comparison\footnote{REML is generally the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however, the REML criterion is not meaningful for comparison of the models with different fixed effects.
+Furthermore, including heatwave days and maximum daily temperature would only lead to a small change in the AIC from $2080.877$ to $2076.439$ (these values differ from the AIC presented in table~\ref{ARMA} in the appendix because the final estimates reported in table~\ref{ARMA} were obtained using the restricted maximum likelihood (REML) whereas the models were re-estimated using the ML method for the comparison\footnote{REML is generally the preferred estimation method for the mixed effects models as the ML estimates of the variance component are biased, however, the REML criterion is not meaningful for comparison of the models with different fixed effects.
 Therefore, we re-estimated the models by ML for the purpose of comparison \cite{Zuur2009}.}).
@@
 
-We further applied a series of likelihood ratio tests to test for the significance of random slopes.
-No subset of the random slopes improved the fit; therefore, besides the fixed effects, our final model only included the random intercepts.\footnote{Although adding the random slope of CV improved the AIC from $2122.226$ to $2122.110$, we decided to not to include CV of temperature in the random effects.
+After we found a set of significant fixed effects, we applied a series of likelihood ratio tests to test for the significance of random slopes.
+No subset of the random slopes improved the fit; therefore, besides the fixed effects, our final model only includes the random intercepts.\footnote{Although adding the random slope of CV of temperature improved the AIC from $2122.226$ to $2122.110$, we decided to not to include this variable in the random effects.
 The reason for not including the random slope is that the difference in AIC was negligible and the F-test of the comparison of the two models was insignificant favouring the simpler model (p-value$=0.128$).}
@@
 In each iteration, the insignificant variable with the highest p-value was removed.
-This was repeated until only the significant predictors remained.
+This was repeated until only the significant random effects remained.
 After this, we applied an analogous procedure to the fixed effects until we were left only with the significant variables.
-As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$~ for fixed effects \cite{lmerTest}.
+As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}.
 This method led to the same model as the one obtained with our primary AIC-based method described above.
@@
 
-Besides our main model, we estimated the same specification for two additional subsamples of the Kenyan counties.
-In particular, we estimated the model for the subsample of the ASAL counties and for the subsample of the non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with the significantly different climate.
+Besides our main model estimated for all counties of Kenya, we estimated the same specification for two subsamples of the Kenyan counties.
+In particular, we estimated the model for the subsample of the ASAL counties and for the subsample of the non-ASAL counties separately to compare the effects of various weather characteristics across the two areas with significantly different climate.
 The estimates of all three models are summarised in table~\ref{MainEst}.
@@
 As an additional measure of relative importance we included F-values of the type \textit{III} analysis of variance (ANOVA) in table~\ref{MainEst}.
-The F-tests were conducted using the marginal rather than the sequential sum of squares.
+The type \textit{III} ANOVA is conducted using the marginal rather than the sequential sum of squares.
 That is, for each variable, the F-test is a test of significance of that part of explained sum of squares which can be contributed to the particular explanatory variable.
-In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst}) and the model without the explanatory variable in question but including all other predictors.
+In other worlds, the F-value is a test statistic of comparison of the preferred model (with all the variables in table~\ref{MainEst} in our case) and the model without the explanatory variable in question but including all other predictors.
 The main advantage of the type~\textit{III} ANOVA is that the test statistic does not depend on the order in which the effects are entered in the model.
 
-The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as the marginal effects and the coefficients are multiplicative.
+The preferred specification has a log-linear functional form, therefore the estimates in table~\ref{MainEst} are not directly interpretable as marginal effects and the coefficients are multiplicative.
 The predicted marginal effects of all coefficients are depicted in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix.
@@
 For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as a positive effect.
-If, for example, a marginal effect of an explanatory variable $x$ was $1.1$ when $x$ is equal to its one standard deviation, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figures~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).
+If, for example, a marginal effect of an explanatory variable $x$ being equal to its standard deviation is $1.1$, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figure~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).
 
@@
 
+\vspace{1cm}
+
 The estimates of the model of the whole sample are summarised in the first two columns of table~\ref{MainEst}.
@@
 Other fixed effects which we found significant include CV of temperature and precipitation and length and number of dry spells (see the first column of table~\ref{MainEst}).
-As it is apparent from figures~\ref{MarEff2}(a) and \ref{MarEff2}(b) in the appendix, the relationship of maize yield and each of these variables is linear.
+As it is apparent from figures~\ref{MarEff2}(a)-(b) and \ref{MarEff3}(a)-(b) in the appendix, the relationship of maize yield and each of these variables is linear.
 In accordance with our expectation, the yield decreases with increasing CV of precipitation.
-\hl{However, the result which we found for CV of temperature is a bit puzzling as the yield seem to increase with increasing variability of temperature.
-This is probably because...}
+\hl{However, the yield seems to increase with increasing variability of temperature which is counter-intuitive...}\textcolor{blue}{...\textit{reasoning}...}
 
@@
 
-According to the magnitude of the standardised coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable is average seasonal temperature which appears to be much more important than the remaining predictors.
+According to the magnitude of the standardised regression coefficients and F-values in the first two columns of table~\ref{MainEst}, the most important variable is average seasonal temperature which appears to be much more important than the remaining predictors.
 The second most important variable is total seasonal precipitation.
@@
 
-We will now discuss and compare the results of the models based on the subsample of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the model based on the subsample of the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
+We will now discuss and compare the results of the models estimated for the subsample of the ASAL counties (the third and fourth columns of table~\ref{MainEst}) and the model estimated for the subsample of the non-ASAL counties (the fifth and sixth columns of table~\ref{MainEst}).
 Interestingly, different predictors were found significant for each of the two subsamples.
@@
 According to our expectation, all three variables have a negative sign.
-On the other hand, the variables which were found significant for the subsample of the non-ASAL counties include seasonal precipitation (positive) and its square (negative) and CV of temperature (positive).
+On the other hand, the variables which were found significant for the subsample of the non-ASAL counties are seasonal precipitation (positive) and its square (negative) and CV of temperature (positive).
 According to the F-statistics, the most important explanatory variables for the ASAL counties are number of dry spells and length of the longest dry spell with values of the F-statistics equal to $8.032$ and $6.841$, respectively.
-The F-statistic of the average temperature is $5.409$ and the F-statistics of the remaining predictors are lower than zero.
-However, according to the magnitude of the standardised coefficients, the average temperature would be the most important as the coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
-With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its square.
-The F-statistic of the squared values of precipitation is $23.384$ and its value for the linear term of precipitation is $19.140$, while the variable with the third highest F-statistic is CV of temperature with the F-statistic equal to $5.372$, hence, the difference in importance of the precipitation and CV of temperature is relatively large.
-The standardised coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised coefficient of CV of temperature is $0.058$.
-In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of the coefficient is $-0.128$.
-However, average temperature is insignificant according to the Satterthwaite's p-value and its F-statistic is also relatively small, in particular $1.528$.
+The F-statistic of the average temperature is $5.409$ and the F-statistics of the remaining predictors are lower than one.
+However, according to the magnitude of the standardised regression coefficients, the average temperature is the most important as its coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
 
-To sum up, for the ASAL counties temperature is much more important than for the non-ASAL counties where temperature is not significant.
+With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its squared term.
+The F-statistic of the squared values of precipitation is $23.384$ and the F-value of the linear term of precipitation is $19.140$.
+The variable with the third highest value of the F-statistic is CV of temperature and the corresponding F-statistic is equal to $5.372$.
+Hence, the difference in importance of the precipitation and CV of temperature is relatively large.
+The standardised regression coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised regression coefficient of CV of temperature is $0.058$.
+In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of its coefficient is $-0.128$.
+However, average temperature is insignificant according to the Satterthwaite's p-value and also the value of its F-statistic is relatively low ($1.528$).
+
+To sum up, temperature is much more important for the ASAL counties than for the non-ASAL counties where temperature is isignificant.
 On the other hand, precipitation is not significant for the ASAL counties while it is strongly significant for the non-ASAL counties.
-This is illustrated in figure~\ref{MarEff1}(c)-(f) in this section.
+This is illustrated in figures~\ref{MarEff1}(c)-(f) in this section.
 The plot of the marginal effect of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(c) in this section) while it is hill-shaped and steep for the non-ASAL counties (figure \ref{MarEff1}(e) in this section) similarly as in the case of the whole sample.
-As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the case of ASAL and non-ASAL counties.
+As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the both subsamples .
 Marginal effects of the remaining predictors are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
@@
 In order to assess the absolute goodness of fit of a model, a different measure is needed.
-For a simple linear regression, the coeffiecient of determination $R^2$ or its adjusted variant is typically used for this purpose.
+For a simple linear regression, the coeffiecient of determination $R^2$ or its adjusted version is typically used for this purpose.
 Nevertheless, obtaining the $R^2$ or the adjusted $R^2$ is not straightforward for linear mixed models or for models estimated by the maximum likelihood in general.
@@
 For each of our samples (all counties, ASAL and non-ASAL), we present two versions of the statistic in table~\ref{R2}.
-These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects components and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects.
-In this case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts.
-As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties.
+These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects.
+In our case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts.
+As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties,  about $13\%$ of the variance in the model based on the ASAL counties and about $15\%$ of the variance in the model based on the non-ASAL counties.
 
@@
         \begin{adjustwidth}{1cm}{}
-    \item[a] The marginal coefficient is interpreted as the variance explained by the fixed effects
-  \item[b] while the conditional one is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts
+    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
+  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
      \end{adjustwidth}
@@
 
-To check for presence of multicollinearity we calculated the variance inflation factors (VIF) of the explanatory variables.
-With only two exceptions, the VIF were below $2$ for all predictors in our final models.
-For CV of precipitation, the value of VIF was $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and $2.182$ in the model for the subsample of the non-ASAL counties.
-These values are still well below the threshold of problematic multicollinearity.
-The other exception was the seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably as a result of the presence of the squared of precipitation.
-The VIF is below $2$ in the other two models and even the value $2.407$ is below the threshold of the problematic multicollinearity.
+Following a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity.
+With only two exceptions, the VIF is below $2$ for all predictors in our final models.
+For CV of precipitation, the value of VIF is $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it is equal to $2.182$ in the model based on the subsample of the non-ASAL counties.
+These values are well below the threshold of problematic multicollinearity.
+The other exception is the linear term of seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably due to presence of the squared term of precipitation.
+However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation is below $2$ in the other two models.
 Therefore, we can conclude that the models do not suffer from presence of multicollinearity.
 
-\begin{itemize}
-\color{blue}
-\item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.
-
-\item relative importance of the individual variables and its measures
 %\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
@@
 
-\item show VIF
+\begin{itemize}
+\color{blue}
 \item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data
@@
 \vspace{1cm}
-\textcolor{blue}{Possibly include a table of all values which I get from the lme or lme4 models summary in R, that is. correlation of the fixed effects etcetera}
+%\textcolor{blue}{Possibly include a table of all values which I get from the lme or lme4 models summary in R, that is. correlation of the fixed effects etcetera}
 
-\textcolor{blue}{Maybe also a table with standard errors here}
+%\textcolor{blue}{Maybe also a table with standard errors here}
 

### 2018-12-26 Rcodes/December18/playing.R

@@ -1,2 +1,14 @@
 aa<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
             +AvgTemp + CVTempK, random= ~1 | ID1 
+        
+        KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS,na.action=na.exclude); summary(KEN11d); exp(summary(KEN11d)$coef[[1]])
+        
+        extractAIC(update(KEN11d,.~.,method="ML"))
+        
+        
+        ff<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                    +AvgTemp + CVTempK, random= ~CVPrec+1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                    data=ScaledTS,na.action=na.exclude,method="ML"); summary(ff)
+        anova(ff,update(KEN11d,.~.,method="ML"))

### 2019-01-13 Rcodes/crossvalidation/MAPE_ASAL.R
MAPE ASAL and nonASAL

@@ -1,0 +1,33 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#
+
+KEN11dK_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                  +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                  data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11dK_ASAL)
+
+foo<-lme(Yield0~Year+ID1+SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+         +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+         data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.omit);
+
+
+frASAL<-model.frame(foo,data=ScaledTS[ScaledTS$ASAL==1,])[complete.cases(model.frame(foo,data=ScaledTS[ScaledTS$ASAL==1,])),]
+    #test:
+        bar<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+             +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+             data=frASAL,na.action=na.exclude);summary(bar)
+        #v.good
+        
+MAPE_ASAL<-cvFit(KEN11dK_ASAL,data=frASAL,y=log(frASAL$Yield0),K=698,cost=mape);MAPE_ASAL;summary(MAPE_ASAL)
+
+# and base
+
[…]

### 2019-01-13 Rcodes/crossvalidation/MAPE_nonASAL.R

@@ -1,0 +1,32 @@
+rm(list=ls())
+library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
+
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+#
+KEN11dK_nonASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                     +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+                     data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11dK_nonASAL)
+
+foo<-lme(Yield0~Year+ID1+SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+         +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+         data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.omit);
+
+
+frnonASAL<-model.frame(foo,data=ScaledTS[ScaledTS$ASAL==0,])[complete.cases(model.frame(foo,data=ScaledTS[ScaledTS$ASAL==0,])),]
+    #test:
+        bar<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+             +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+             data=frnonASAL,na.action=na.exclude);summary(bar)
+        #v.good
+        
+MAPE_nonASAL<-cvFit(KEN11dK_nonASAL,data=frnonASAL,y=log(frnonASAL$Yield0),K=602,cost=mape);MAPE_nonASAL;summary(MAPE_nonASAL)
+
+# and base
+
[…]

### 2019-01-13 Rcodes/crossvalidation/compareJustTotals.R

@@ -99,7 +99,10 @@ framik$Year<-as.numeric(framik$Year);framik$Year<-as.factor(framik$Year)
 framik$ID1<-as.numeric(framik$ID1);framik$ID1<-as.factor(framik$ID1)
 # maybe better check as csv:
 write.csv(framik, "framik.csv")
-
+#ok, looks good. now calculate MAPE and compare..
+mean(abs(log(framik$Yield0)-log(framik$Yield0lag)),na.rm=TRUE) # compare with:
+CVbase3 # and with:
+CVfit6 #P.E.R.F.E.C.T
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
 save.image("~/FoodSystemGitH/Rcodes/crossvalidation/compareJustTotals.RData")

### 2019-01-13 Rcodes/DecemberNew/KEN11d_finals2.R
WAD

@@ -1,5 +1,5 @@
 rm(list=ls())
-library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+library(cvTools);library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 load("dataFS/Main/DaTS.RData")
 # load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo

### 2019-01-13 Rcodes/crossvalidation/MAPE_ASAL.R

@@ -1,5 +1,5 @@
 rm(list=ls())
-library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+library(cvTools);library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 load("dataFS/Main/DaTS.RData")
 # load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
 

### 2019-01-13 writing/draft4/draft20.tex
MAPE asal non asal

--- 2018-12-26 writing/draft3/draft19.tex
@@
 According to our expectation, all three variables have a negative sign.
-On the other hand, the variables which were found significant for the subsample of the non-ASAL counties include seasonal precipitation (positive) and its square (negative) and CV of temperature (positive).
+On the other hand, the variables which were found significant for the subsample of the non-ASAL counties are seasonal precipitation (positive) and its square (negative) and CV of temperature (positive).
 According to the F-statistics, the most important explanatory variables for the ASAL counties are number of dry spells and length of the longest dry spell with values of the F-statistics equal to $8.032$ and $6.841$, respectively.
-The F-statistic of the average temperature is $5.409$ and the F-statistics of the remaining predictors are lower than zero.
-However, according to the magnitude of the standardised coefficients, the average temperature would be the most important as the coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
+The F-statistic of the average temperature is $5.409$ and the F-statistics of the remaining predictors are lower than one.
+However, according to the magnitude of the standardised regression coefficients, the average temperature is the most important as its coefficient is equal to $-0.214$ while it is equal to $-0.156$ and $-0.181$ for the number of dry spells and the length of maximum dry spell, respectively.
 With regard to the subsample of the non-ASAL counties, the most important explanatory variables based on the F-statistic are precipitation and its square.
-The F-statistic of the squared values of precipitation is $23.384$ and its value for the linear term of precipitation is $19.140$, while the variable with the third highest F-statistic is CV of temperature with the F-statistic equal to $5.372$, hence, the difference in importance of the precipitation and CV of temperature is relatively large.
-The standardised coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised coefficient of CV of temperature is $0.058$.
-In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of the coefficient is $-0.128$.
-However, average temperature is insignificant according to the Satterthwaite's p-value and its F-statistic is also relatively small, in particular $1.528$.
+The F-statistic of the squared values of precipitation is $23.384$ and its value for the linear term of precipitation is $19.140$, while the variable with the third highest value of the F-statistic is CV of temperature with the F-statistic equal to $5.372$, hence, the difference in importance of the precipitation and CV of temperature is relatively large.
+The standardised regression coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised regression coefficient of CV of temperature is $0.058$.
+In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of its coefficient is $-0.128$.
+However, average temperature is insignificant according to the Satterthwaite's p-value and the value of its F-statistic is also relatively low, in particular $1.528$.
 
-To sum up, for the ASAL counties temperature is much more important than for the non-ASAL counties where temperature is not significant.
+To sum up, temperature is much more important for the ASAL counties than for the non-ASAL counties where temperature is isignificant.
 On the other hand, precipitation is not significant for the ASAL counties while it is strongly significant for the non-ASAL counties.
-This is illustrated in figure~\ref{MarEff1}(c)-(f) in this section.
+This is illustrated in figures~\ref{MarEff1}(c)-(f) in this section.
 The plot of the marginal effect of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(c) in this section) while it is hill-shaped and steep for the non-ASAL counties (figure \ref{MarEff1}(e) in this section) similarly as in the case of the whole sample.
-As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the case of ASAL and non-ASAL counties.
+As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the both subsamples .
 Marginal effects of the remaining predictors are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
@@
 These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects components and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects.
-In this case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts.
-As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties.
+In our case, the only random effects are the intercepts, hence the difference between the two variants is the variance explained by the random intercepts.
+As is apparent from table~\ref{R2}, the fixed effects explain about $30\%$ of the variance in case of the model for all counties,  about $13\%$ of the variance in the model based on the ASAL counties and about $15\%$ of the variance in the model based on the non-ASAL counties.
 
@@
 
-To check for presence of multicollinearity we calculated the variance inflation factors (VIF) of the explanatory variables.
-With only two exceptions, the VIF were below $2$ for all predictors in our final models.
-For CV of precipitation, the value of VIF was $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and $2.182$ in the model for the subsample of the non-ASAL counties.
-These values are still well below the threshold of problematic multicollinearity.
-The other exception was the seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably as a result of the presence of the squared of precipitation.
-The VIF is below $2$ in the other two models and even the value $2.407$ is below the threshold of the problematic multicollinearity.
+According to a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity.
+With only two exceptions, the VIF is below $2$ for all predictors in our final models.
+For CV of precipitation, the value of VIF is $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it is equal to $2.182$ in the model based on the subsample of the non-ASAL counties.
+These values are well below the threshold of problematic multicollinearity.
+The other exception is the seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably due to presence of the squared term of precipitation.
+However, $2.407$ is below the threshold of the problematic multicollinearity.
+The VIF of squared precipitation is below $2$ in the other two models.
 Therefore, we can conclude that the models do not suffer from presence of multicollinearity.
 
-\begin{itemize}
-\color{blue}
-\item Verbal description and interpretation of the results. Discussing goodness of fit using various criteria such as AIC or alternatives to $R^2$.
-
-\item relative importance of the individual variables and its measures
 %\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.

### 2019-01-13 Rcodes/crossvalidation/MAPE_ASAL.R
yep, finilazing cross validaiton

@@ -2,6 +2,7 @@ rm(list=ls())
 library(cvTools);library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 load("dataFS/Main/DaTS.RData")
 # load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
+load("~/FoodSystemGitH/Rcodes/crossvalidation/MAPE_ASAL.RData")
 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 #
@@ -29,5 +30,18 @@ MAPE_ASAL<-cvFit(KEN11dK_ASAL,data=frASAL,y=log(frASAL$Yield0),K=698,cost=mape);
 KEN11dK_ASALb<-lme(log(Yield0)~SeasPr+AvgTempK , random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
                   data=ScaledTS[ScaledTS$ASAL==1,],na.action=na.exclude); summary(KEN11dK_ASALb)
 MAPE_ASAL_base<-cvFit(KEN11dK_ASALb,data=frASAL,y=log(frASAL$Yield0),K=698,cost=mape);MAPE_ASAL_base;summary(MAPE_ASAL_base)
+
+        frASAL<-pdata.frame(frASAL,index=c("ID1","Year"))
+        Yield0lag<-lag(frASAL$Yield0)
+        frASAL<-cbind.data.frame(frASAL,Yield0lag)
+        frASAL$Yield0lag[frASAL$Year==1981]<-NA
+        #ehm..malformed factor..:
+        frASAL$Year<-as.numeric(frASAL$Year);frASAL$Year<-as.factor(frASAL$Year)
+        frASAL$ID1<-as.numeric(frASAL$ID1);frASAL$ID1<-as.factor(frASAL$ID1)
+mean(abs(log(frASAL$Yield0)-log(frASAL$Yield0lag)),na.rm=TRUE) 
+
+MAPE_ASAL$cv/IQR(log(frASAL$Yield0))
+MAPE_ASAL_base$cv/IQR(log(frASAL$Yield0))
[…]

### 2019-01-13 Rcodes/crossvalidation/MAPE_nonASAL.R

@@ -2,6 +2,7 @@ rm(list=ls())
 library(cvTools);library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
 load("dataFS/Main/DaTS.RData")
 # load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
+load("~/FoodSystemGitH/Rcodes/crossvalidation/MAPE_nonASAL.RData")
 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 #
@@ -22,11 +23,22 @@ frnonASAL<-model.frame(foo,data=ScaledTS[ScaledTS$ASAL==0,])[complete.cases(mode
         #v.good
         
 MAPE_nonASAL<-cvFit(KEN11dK_nonASAL,data=frnonASAL,y=log(frnonASAL$Yield0),K=602,cost=mape);MAPE_nonASAL;summary(MAPE_nonASAL)
-
+        frnonASAL<-pdata.frame(frnonASAL,index=c("ID1","Year"))
+        Yield0lag<-lag(frnonASAL$Yield0)
+        frnonASAL<-cbind.data.frame(frnonASAL,Yield0lag)
+        frnonASAL$Yield0lag[frnonASAL$Year==1981]<-NA
+        #ehm..malformed factor..:
+        frnonASAL$Year<-as.numeric(frnonASAL$Year);frnonASAL$Year<-as.factor(frnonASAL$Year)
+        frnonASAL$ID1<-as.numeric(frnonASAL$ID1);frnonASAL$ID1<-as.factor(frnonASAL$ID1)
+        mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE) 
 # and base
 
 KEN11dK_nonASALb<-lme(log(Yield0)~SeasPr+AvgTempK , random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
                   data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11dK_nonASALb)
[…]

### 2019-01-13 Rcodes/crossvalidation/compareJustTotals.R

@@ -6,6 +6,8 @@ load("Rcodes/crossvalidation/compareJustTotals3.RData")
 # load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
 # load("Rcodes/crossvalidation/KEN11d_stepNice.RData")
 
+load("~/FoodSystemGitH/Rcodes/crossvalidation/compareJustTotals.RData")
+
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
 # from the file 'Rcodes/DecemberNew/'.. the so far final version (7.1.2019):
@@ -103,6 +105,12 @@ write.csv(framik, "framik.csv")
 mean(abs(log(framik$Yield0)-log(framik$Yield0lag)),na.rm=TRUE) # compare with:
 CVbase3 # and with:
 CVfit6 #P.E.R.F.E.C.T
+
+
+mean(abs(log(framik$Yield0)-log(framik$Yield0lag)),na.rm=TRUE)/IQR(log(framik$Yield0)) # compare with:
+CVbase3$cv/IQR(log(framik$Yield0)) # and with:
+CVfit6$cv/IQR(log(framik$Yield0)) #P.E.R.F.E.C.T
+
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
 save.image("~/FoodSystemGitH/Rcodes/crossvalidation/compareJustTotals.RData")

### 2019-01-15 writing/draft4/draft21.tex
yep, finilazing cross validaiton
writting correction

--- 2018-12-26 writing/draft3/draftMonika.tex
@@
 These measures are listed in table~\ref{vars} in the appendix.
-The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation (CV) of seasonal precipitation, CV of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and CV does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$mm).
-We did not include squared values of the average seasonal temperature in our preferred models as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and average temperature.
+The measures which we found significant were: Total seasonal precipitation and its squared values, average seasonal temperature, coefficient of variation (CV) of seasonal precipitation, CV of average seasonal temperature (converted into Kelvin because the Celsius scale is interval and CV does not have any meaning for data measured on an interval scale),  maximum length of dry spell measured as number of consecutive dry days and the number of dry spells lasting for four days or more (for the purpose of this study we defined a dry day as a day when precipitation didn't exceed $1$ mm).
+We did not include squared values of the average seasonal temperature in our preferred model as it was not significant and most of the past literature has not supported the evidence of the non-linear relationship between yield and average temperature.
 \subsection{Statistical approach}\label{stats}
@@
 We started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the first part of the procedure, the insignificant random effects were removed one by one.
+In the first part of the procedure, insignificant random effects were removed one by one.
 In each iteration, the insignificant variable with the highest p-value was removed.
-This was repeated until only the significant random effects remained.
-After this, we applied an analogous procedure to the fixed effects until we were left only with the significant variables.
+This was repeated until only significant random effects remained.
+After this, we applied an analogous procedure to the fixed effects until we were left only with significant variables.
 As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}.
-This method led to the same model as the one obtained with our primary AIC-based method described above.
+This method led us to the same results as those obtained with our primary AIC-based method described above.
 
@@
 
+\begin{center}
+\begin{threeparttable}
+\caption{\label{CVsc}Cross validation: Mean absolute prediction error scaled by IQR}
+% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
+\begin{indented}
+\item[]
+\begin{tabular}{@{}lccc}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
+\hline
+{All counties}& $0.3811$&$0.3889$&$0.4277$\\
+{ASAL}& $0.5300$&$0.5342$&$0.5835$\\
+{non-ASAL}&  $0.2845$&$0.3858$&$0.4510$\\
+\br
+\end{tabular}
+\begin{tablenotes}
+\begin{footnotesize}
+    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
+        \begin{adjustwidth}{1cm}{}
+    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
+  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
+     \end{adjustwidth}
+     \end{footnotesize}
+\end{tablenotes}
+
+\end{indented}
+\end{threeparttable} 
+\end{center}
+\vspace{1cm}
+
+\begin{center}
+\begin{threeparttable}
+\caption{\label{CV}Cross validation: Mean absolute prediction error}
+% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
+\begin{indented}
+\item[]
+\begin{tabular}{@{}lccc}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
+\hline
+{All counties}& $0.3599$&$0.3673$&$0.4040$\\
+{ASAL}& $0.134$&$0.353$\\
+{non-ASAL}&  $0.153$&$0.436$\\
+\br
+\end{tabular}
+\begin{tablenotes}
+\begin{footnotesize}
+    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
+        \begin{adjustwidth}{1cm}{}
+    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
+  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
+     \end{adjustwidth}
+     \end{footnotesize}
+\end{tablenotes}
+
+\end{indented}
+\end{threeparttable} 
+\end{center}
+\vspace{1cm}
+
 \begin{itemize}

### 2019-01-15 writing/draft4/draft21.tex
writing, or more like editing

@@
 After this, we applied an analogous procedure to the fixed effects until we were left only with significant variables.
-As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}.
+As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$ for fixed effects \cite{lmerTest}.
 This method led us to the same results as those obtained with our primary AIC-based method described above.
 
-According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
+According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991; 1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
 To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with an ARMA($p$,$q$) error structure.
 In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$.
-Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$).
+Comparing the AIC criteria and using the corresponding likelihood ratio test statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$).
 The value of the AIC criterion of the model with the ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.
@@
 For the values of the marginal effects smaller than one, the effect on yield is negative while marginal effects higher than one can be interpreted as a positive effect.
-If, for example, a marginal effect of an explanatory variable $x$ being equal to its standard deviation is $1.1$, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figure~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).
+If, for example, a size of a marginal effect of an explanatory variable $x$ was equal to $1.1$ for $x$ being equal to its standard deviation, one should interpret this as follows: if $x$ is equal to its standard deviation, the dependent variable (in our case maize yield) is $1.1$ times higher than it would be if the explanatory variable $x$ was equal to its mean (which is zero in figure~\ref{MarEff1} in this section and in figures~\ref{MarEff2} and~\ref{MarEff3} in the appendix as the predictors were standardised).
 
@@
 
-In the following few paragraphs, we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.
+In the following few paragraphs we will discuss the estimates based on the sample of all counties and after that we will focus on the estimates based on the two subsamples.
 
@@
   \\  \vspace{-0.2cm}Dry spell -length\tnote{b}&$-0.067^{*}$&$\04.800$&$-0.181^{**}$&$6.841$&$-0.012$&$\00.164$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$-0.063^{*}$&$\04.848$&$-0.156^{**}$&$8.032$&$-0.011$&$\00.087$\\
+  \\ \vspace{-0.2cm}Dry spells 	$\geq~4$ d.\tnote{c}&$-0.063^{*}$&$\04.848$&$-0.156^{**}$&$8.032$&$-0.011$&$\00.087$\\
   \\ \vspace{-0.2cm}Temp. - average&$-0.199^{***}$&$12.010$&$-0.214^{*}$&$5.409$&$-0.128$&$\01.525$\\
@@
     \item[a] Marginal (type III) sum of squares. The F-statistics correspond to the sum of squares attributable to each fixed effect.
-  \item[b] In number of days.
+  \item[b] Length of the longest dry spell in number of days.
+    \item[c] Number of dry spells lasting for four days or longer. 
      \end{adjustwidth}
@@
 
-As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with increasing length of maximum dry spell linearly.
+As we can see in figures~\ref{MarEff3}(a) and \ref{MarEff3}(b) in the appendix, yield declines with increasing number of dry spells and with increasing length of maximum dry spell roughly linearly.
 This is in accordance with our expectations.
@@
 The variable with the third highest value of the F-statistic is CV of temperature and the corresponding F-statistic is equal to $5.372$.
-Hence, the difference in importance of the precipitation and CV of temperature is relatively large.
+Hence, the difference in importance of precipitation and CV of temperature is relatively large.
 The standardised regression coefficient of squared precipitation is equal to $-0.127$, the coefficient of the linear term of precipitation is $0.244$ and the standardised regression coefficient of CV of temperature is $0.058$.
 In terms of the magnitude of the coefficients, the average temperature is also relatively important as the value of its coefficient is $-0.128$.
-However, average temperature is insignificant according to the Satterthwaite's p-value and also the value of its F-statistic is relatively low ($1.528$).
+However, average temperature is insignificant according to the Satterthwaite's p-value $(0.217)$ and also the value of its F-statistic is relatively low ($1.528$).
 
@@
 The plot of the marginal effect of precipitation is almost flat for the ASAL counties (figure \ref{MarEff1}(c) in this section) while it is hill-shaped and steep for the non-ASAL counties (figure \ref{MarEff1}(e) in this section) similarly as in the case of the whole sample.
-As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the both subsamples .
+As one can see in figures~\ref{MarEff1}(d) and~\ref{MarEff1}(f) in this section, the marginal effect of temperature is linear and decreasing for the both subsamples.
 Marginal effects of the remaining predictors are depicted in figures~\ref{MarEff2}(c)-(f) and~\ref{MarEff3}(c)-(f) in the appendix.
@@
 In this study, we used the coefficient of determination of generalized linear mixed-effects models introduced by \cite{Nakagawa2013}.
-For each of our samples (all counties, ASAL and non-ASAL), we present two versions of the statistic in table~\ref{R2}.
+For each of our samples (all counties, ASAL and non-ASAL), we present two versions of the generalised $R^2$ in table~\ref{R2}.
 These include marginal $R^2$ which can be interpreted as the variance explained by the fixed effects and the conditional $R^2$ which can be interpreted as the variance explained by both fixed effects and random effects.
@@
 \begin{footnotesize}
-    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
+    \item \textit{Notes:} The coefficient of determination for generalised linear mixed-effects models \cite{Nakagawa2013}.  \\
         \begin{adjustwidth}{1cm}{}
@@
 Following a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity.
-With only two exceptions, the VIF is below $2$ for all predictors in our final models.
-For CV of precipitation, the value of VIF is $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it is equal to $2.182$ in the model based on the subsample of the non-ASAL counties.
+With only two exceptions, the VIF was below $2$ for all predictors in our final models.
+For CV of precipitation, the value of VIF was $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it was equal to $2.182$ in the model based on the subsample of the non-ASAL counties.
 These values are well below the threshold of problematic multicollinearity.
-The other exception is the linear term of seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably due to presence of the squared term of precipitation.
-However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation is below $2$ in the other two models.
+The other exception was the linear term of seasonal precipitation as its VIF was equal to $2.407$ in the model for the non-ASAL counties, probably due to presence of its squared term in the model.
+However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation was below $2$ in the other two models.
 Therefore, we can conclude that the models do not suffer from presence of multicollinearity.
@@
     \\ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$0.970$ &$0.907^{}$\\
-  \\  \vspace{-0.2cm}Dry spell -length&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
-  \\ \vspace{-0.2cm}Dry spells 	$\geq$ 4 d.&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
+  \\  \vspace{-0.2cm}Dry spell -length\tnote{a}&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
+  \\ \vspace{-0.2cm}Dry spells $\geq~4$ d.\tnote{b}&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
   \\ \vspace{-0.2cm}Temp. - average&$0.820^{***}$&$0.808^{*}$&$0.880$ $^{}$\\
@@
     \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+            \begin{adjustwidth}{1cm}{} 
+  \item[a] Length of the longest dry spell in number of days.
+    \item[b] Number of dry spells lasting for four days or longer. 
+     \end{adjustwidth}
 \singlespacing

### 2019-01-15 writing/draft4/draft21.tex
tables jackknife

@@
 \begin{threeparttable}
-\caption{\label{CVsc}Cross validation: Mean absolute prediction error scaled by IQR}
+\caption{\label{CVsc}Cross validation: Mean absolute prediction error divided by IQR}
 % files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
@@
 \hline
-{All counties}& $0.3811$&$0.3889$&$0.4277$\\
-{ASAL}& $0.5300$&$0.5342$&$0.5835$\\
-{non-ASAL}&  $0.2845$&$0.3858$&$0.4510$\\
+{All counties}&$0.3811$&$0.3889$&$0.4277$\\
+{ASAL}&$0.5300$&$0.5342$&$0.5835$\\
+{non-ASAL}&$0.3858$&$0.4043$&$0.4510$\\
 \br
 \end{tabular}
-\begin{tablenotes}
-\begin{footnotesize}
-    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
-        \begin{adjustwidth}{1cm}{}
-    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
-  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
-     \end{adjustwidth}
-     \end{footnotesize}
-\end{tablenotes}
-
 \end{indented}
@@
 \vspace{1cm}
-
-\begin{center}
-\begin{threeparttable}
-\caption{\label{CV}Cross validation: Mean absolute prediction error}
-% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
-\begin{indented}
-\item[]
-\begin{tabular}{@{}lccc}
-\\[-1em]
-\br
-\\[-1em]
-\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
-\hline
-{All counties}& $0.3599$&$0.3673$&$0.4040$\\
-{ASAL}& $0.134$&$0.353$\\
-{non-ASAL}&  $0.153$&$0.436$\\
-\br
-\end{tabular}
-\begin{tablenotes}
-\begin{footnotesize}
-    \item \textit{Notes:} The coefficient of determination from generalised linear mixed-effects models. The statistic is based on \cite{Nakagawa2013}.  \\
-        \begin{adjustwidth}{1cm}{}
-    \item[a] The marginal $R^2$ is interpreted as the variance explained by the fixed effects
-  \item[b] The conditional $R^2$ is explained by both fixed affects and random effects (in our case by fixed effects and random intercepts)
-     \end{adjustwidth}
-     \end{footnotesize}
-\end{tablenotes}
-
-\end{indented}
-\end{threeparttable} 
-\end{center}
-\vspace{1cm}
-
-\begin{itemize}
-\color{blue}
-\item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data
-\end{itemize}
-
-\color{black}
 
@@
 
+\begin{center}
+\begin{threeparttable}
+\caption{\label{CV}Cross validation: Mean absolute prediction error}
+% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
+\begin{indented}
+\item[]
+\begin{tabular}{@{}lccc}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
+\hline
+{All counties}&$0.3599$&$0.3673$&$0.4040$\\
+{ASAL}&$0.4352$&$0.4385$&$0.4790$\\
+{non-ASAL}&$0.2715$&$0.2845$&$0.3173$\\
+\br
+\end{tabular}
+\end{indented}
+\end{threeparttable} 
+\end{center}
+\vspace{1cm}
+
+\clearpage
+
 \section*{References}
-
 \bibliography{referencesFS}

### 2019-01-15 Rcodes/crossvalidation/MAPE_ASAL.R

@@ -43,5 +43,9 @@ mean(abs(log(frASAL$Yield0)-log(frASAL$Yield0lag)),na.rm=TRUE)
 MAPE_ASAL$cv/IQR(log(frASAL$Yield0))
 MAPE_ASAL_base$cv/IQR(log(frASAL$Yield0))
 mean(abs(log(frASAL$Yield0)-log(frASAL$Yield0lag)),na.rm=TRUE)/IQR(log(frASAL$Yield0))
+
+MAPE_ASAL
+MAPE_ASAL_base
+mean(abs(log(frASAL$Yield0)-log(frASAL$Yield0lag)),na.rm=TRUE)
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 save.image("~/FoodSystemGitH/Rcodes/crossvalidation/MAPE_ASAL.RData")

### 2019-01-15 Rcodes/crossvalidation/MAPE_nonASAL.R

@@ -37,8 +37,12 @@ KEN11dK_nonASALb<-lme(log(Yield0)~SeasPr+AvgTempK , random= ~1 | ID1,correlation
                   data=ScaledTS[ScaledTS$ASAL==0,],na.action=na.exclude); summary(KEN11dK_nonASALb)
 MAPE_nonASAL_base<-cvFit(KEN11dK_nonASALb,data=frnonASAL,y=log(frnonASAL$Yield0),K=602,cost=mape);MAPE_nonASAL_base;summary(MAPE_nonASAL_base)
   
-  MAPE_nonASAL$cv/IQR(log(frnonASAL$Yield0))
-  KEN11dK_nonASALb$cv/IQR(log(frnonASAL$Yield0))
-  mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE) /IQR(log(frnonASAL$Yield0))
-  #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+MAPE_nonASAL$cv/IQR(log(frnonASAL$Yield0))
+MAPE_nonASAL_base$cv/IQR(log(frnonASAL$Yield0))
+mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE) /IQR(log(frnonASAL$Yield0))
+
+MAPE_nonASAL
+KEN11dK_nonASALb
+mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE)
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 save.image("~/FoodSystemGitH/Rcodes/crossvalidation/MAPE_nonASAL.RData")

### 2019-01-15 writing/draft4/draft22.tex
working MAPE writing

--- 2019-01-15 writing/draft4/draft21.tex
@@
 \end{tablenotes}     
-
 \end{indented}
@@
 \vspace{1cm}
+
+	%	\begin{itemize}%
+		%	\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
+This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
+	%\end{itemize}
 
@@
 
-%\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
-This would give us an approximation of $R^2$ which could be used to compare our models.
+To evaluate out of sample prediction accuracy of our models we implemented a procedure which is sometimes called the \textit{jackknife}.
+We re-estimated the preferred mixed effects models $n$-times, where $n$ is the number of observations (in our case $1300$ for the sample of all counties and $698$ and $602$ for the subsamples of ASAL and non-ASAL counties, respectively), each time we left out one observation.
+We subsequently used each of these $n$ estimated models to generate a predicted value
+of log of maize yield of the left-out observation.
+Then we calculated the absolute value of the difference between the predicted and real value of the left-out observation for each of the $n$ models.
+In the next step, we averaged the $n$ absolute values of the differences between predicted and real values to obtain the mean absolute prediction error (MAPE).
+Finally, we scaled the MAPE values by dividing them by the interquartile range (IQR) of the sample obtained as a difference between the third and first quartiles.
 
-	%	\begin{itemize}%
-		%	\item \textcolor{red}{Annemie has advised me that for a more precise approximation of $R^2$ we can estimate a simple regression of yields on the random intercepts (county level dummy variables) and then subtract the $R^2$ of this model from the $R^2$ of the simple regression with the explanatory variables of the preferred mixed models (the model described in the bullet point above).
-This would give us a percentage of yield variability explained by the weather measures not including the variability explained by the random intercepts.}
-	%\end{itemize}
+Besides obtaining the MAPE of our preferred specifications, we calculated MAPE of baseline models specified as mixed effects models with the set of fixed effects including two measures, in particular cumulative seasonal precipitation and seasonal average temperature.
+The only random effects of the baseline models are random intercepts as in our preferred specification and the baseline models have the same error structure ARMA($1$,$1$) as our main models.
+For the sake of comparison, we further calculated MAPE of naive forecasts defined as the last period's actual values of each county.
 
@@
 \begin{threeparttable}
-\caption{\label{CVsc}Cross validation: Mean absolute prediction error divided by IQR}
+\caption{\label{CV}	Leave-one-out cross validation: Mean absolute prediction error divided by IQR}
+\linespread{1.3}
 % files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
@@
 \\[-1em]
-\textbf{Sample:}&\textbf{All covariates}&\textbf{Only seasonal}&\textbf{Naive forecast}\\
+\textbf{Sample:}&\textbf{All covariates\tnote{a}}&\textbf{Only seasonal\tnote{b}}&\textbf{Naive forecast\tnote{c}}\\
 \hline
@@
 \end{indented}
+ \begin{tablenotes}
+  \begin{footnotesize}
+  \item[a] Mixed effects models: the preferred specification with all fixed effects presented in table~\ref{MainEst}.
+  \item[b] Mixed effects models: the baseline specification, only seasonal precipitation and average seasonal temperature in the fixed effects.
+  \item[c] Last period's actual values used as this period's forecast.
+  \end{footnotesize}
+\end{tablenotes}
 \end{threeparttable} 
@@
 \begin{threeparttable}
-\caption{\label{CV}Cross validation: Mean absolute prediction error}
+\caption{\label{CV}	Leave-one-out cross validation: Mean absolute prediction error}
+\linespread{1.3}
 % files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
@@
 \end{indented}
+
+ \begin{tablenotes}
+  \begin{footnotesize}
+  \item[a] Mixed effects models: the preferred specification with all fixed effects presented in table~\ref{MainEst}.
+  \item[b] Mixed effects models: the baseline specification, only seasonal precipitation and average seasonal temperature in the fixed effects.
+  \item[c] Last period's actual values used as this period's forecast.
+  \end{footnotesize}
+\end{tablenotes}
 \end{threeparttable} 
@@
 \clearpage
+
+\doublespacing
 

### 2019-01-15 Rcodes/crossvalidation/MAPE_nonASAL.R

@@ -42,7 +42,7 @@ MAPE_nonASAL_base$cv/IQR(log(frnonASAL$Yield0))
 mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE) /IQR(log(frnonASAL$Yield0))
 
 MAPE_nonASAL
-KEN11dK_nonASALb
-mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE)
+MAPE_nonASAL_base
+mean(abs(log(frnonASAL$Yield0)-log(frnonASAL$Yield0lag)),na.rm=TRUE) 
 #oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 save.image("~/FoodSystemGitH/Rcodes/crossvalidation/MAPE_nonASAL.RData")

### 2019-01-15 writing/draft4/draft22.tex
writing MAPE

@@
 
+The scaled MAPE of our preferred specification and the two baseline predictions can be found in table~\ref{CVdiv}.
+All three predictions were obtained for the sample of all counties and also for the models based on the subsamples of ASAL and non-ASAL counties.
+The values of MAPE not divided by IQR can be found in table~\ref{CV} in the appendix.
+
 \begin{center}
 \begin{threeparttable}
-\caption{\label{CV}	Leave-one-out cross validation: Mean absolute prediction error divided by IQR}
+\caption{\label{CVdiv}	Leave-one-out cross validation: Mean absolute prediction error divided by IQR}
 \linespread{1.3}
@@
 \\[-1em]
-\textbf{Sample:}&\textbf{All covariates\tnote{a}}&\textbf{Only seasonal\tnote{b}}&\textbf{Naive forecast\tnote{c}}\\
+\textbf{Sample}&\textbf{Preferred specification\tnote{a}}&\textbf{Baseline specification\tnote{b}}&\textbf{Naive forecast\tnote{c}}\\
 \hline
@@
   \begin{footnotesize}
-  \item[a] Mixed effects models: the preferred specification with all fixed effects presented in table~\ref{MainEst}.
+  \item[a] Mixed effects models: the preferred specification with all fixed effects listed in table~\ref{MainEst}.
   \item[b] Mixed effects models: the baseline specification, only seasonal precipitation and average seasonal temperature in the fixed effects.
@@
 
+According to MAPE, our preferred specifications perform better than the baseline mixed effects models and considerably better than the naive forecasts for all three samples (table~\ref{CVdiv}).
+For all three prediction methods, the smallest values of MAPE were achieved for the sample of all counties while the least precise forecasts with regard to MAPE were obtained for the subsample of ASAL counties (table~\ref{CVdiv}).
+
 \clearpage
-
+\section{Conclusions}\label{Conclusions}
+\clearpage
 \appendix

### 2019-01-15 writing/draft4/draft23.tex
writing MAPE description

--- 2019-01-15 writing/draft4/draft22.tex
@@
 \begin{indented}
-\item[]December 2018
+\item[]January 2019
 \end{indented}
@@
 These values are well below the threshold of problematic multicollinearity.
-The other exception was the linear term of seasonal precipitation as its VIF was equal to $2.407$ in the model for the non-ASAL counties, probably due to presence of its squared term in the model.
+The other exception was the linear term of precipitation as its VIF was equal to $2.407$ in the model for the non-ASAL counties, probably due to presence of its squared term in the model.
 However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation was below $2$ in the other two models.
@@
 of log of maize yield of the left-out observation.
-Then we calculated the absolute value of the difference between the predicted and real value of the left-out observation for each of the $n$ models.
-In the next step, we averaged the $n$ absolute values of the differences between predicted and real values to obtain the mean absolute prediction error (MAPE).
-Finally, we scaled the MAPE values by dividing them by the interquartile range (IQR) of the sample obtained as a difference between the third and first quartiles.
+After that, we determined the absolute value of the difference between the predicted and real value of the left-out observation for each of the $n$ models.
+In the next step, we calculated average of the $n$ absolute values of the differences between predicted and real values to obtain the mean absolute prediction error (MAPE).
+Finally, we scaled the MAPE values by dividing them by the interquartile range (IQR) of the sample obtained as a difference between the third and first sample quartiles.
 
-Besides obtaining the MAPE of our preferred specifications, we calculated MAPE of baseline models specified as mixed effects models with the set of fixed effects including two measures, in particular cumulative seasonal precipitation and seasonal average temperature.
-The only random effects of the baseline models are random intercepts as in our preferred specification and the baseline models have the same error structure ARMA($1$,$1$) as our main models.
+Besides obtaining the MAPE of our preferred specification, we worked out MAPE of a baseline model specified as a mixed effects model with the set of fixed effects including two measures, in particular cumulative seasonal precipitation and seasonal average temperature.
+The only random effects of the baseline model were random intercepts as in our preferred specification.
+The baseline model had the error structure ARMA($1$,$1$) as our main model.
 For the sake of comparison, we further calculated MAPE of naive forecasts defined as the last period's actual values of each county.
@@
 The scaled MAPE of our preferred specification and the two baseline predictions can be found in table~\ref{CVdiv}.
-All three predictions were obtained for the sample of all counties and also for the models based on the subsamples of ASAL and non-ASAL counties.
-The values of MAPE not divided by IQR can be found in table~\ref{CV} in the appendix.
+All three predictions were obtained for the full-sample model of all counties and also for the models based on the subsamples of ASAL and non-ASAL counties.
+The unscaled values of MAPE (not divided by the IQR) can be found in table~\ref{CV} in the appendix.
 
@@
 
-According to MAPE, our preferred specifications perform better than the baseline mixed effects models and considerably better than the naive forecasts for all three samples (table~\ref{CVdiv}).
-For all three prediction methods, the smallest values of MAPE were achieved for the sample of all counties while the least precise forecasts with regard to MAPE were obtained for the subsample of ASAL counties (table~\ref{CVdiv}).
+According to the MAPE, our preferred specifications perform better than the baseline mixed effects models and they perform considerably better than the naive forecasts for all three samples (table~\ref{CVdiv}).
+For all three prediction methods, the smallest values of MAPE were achieved for the sample of all counties while the least precise forecasts (with regard to MAPE) were obtained for the subsample of ASAL counties (table~\ref{CVdiv}).
 

### 2019-01-15 writing/draft4/draftMonika.tex
MAPE asal non asal
writing MAPE description

--- 2018-12-26 writing/draft3/draftMonika.tex
@@
 We started with a model which included the complete set of our weather measures in both fixed effects and random effects.
-In the first part of the procedure, the insignificant random effects were removed one by one.
+In the first part of the procedure, insignificant random effects were removed one by one.
 In each iteration, the insignificant variable with the highest p-value was removed.
-This was repeated until only the significant random effects remained.
-After this, we applied an analogous procedure to the fixed effects until we were left only with the significant variables.
-As recommended in literature, we considered the level of significance $\alpha=0.1$ for the random effects and $\alpha=0.05$~ for the fixed effects \cite{lmerTest}.
-This method led to the same model as the one obtained with our primary AIC-based method described above.
+This was repeated until only significant random effects remained.
+After this, we applied an analogous procedure to the fixed effects until we were left only with significant variables.
+As recommended in literature, we considered the level of significance $\alpha=0.1$ for random effects and $\alpha=0.05$ for fixed effects \cite{lmerTest}.
+This method led us to the same results as those obtained with our primary AIC-based method described above.
 
-According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991) and Baltagi \& Li (1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
+According to the conditional Lagrange multiplier (LM) test developed by Baltagi \& Li (1991; 1995), the errors exhibited a within group autcorrelation structure in our models (the p-value of the LM test was $1.1\times10^{-13}$).
 To further investigate the autocorrelation structure of the errors, we estimated a number of models (with the subset of predictors chosen as described above) with an ARMA($p$,$q$) error structure.
 In particular, we estimated all variants of our model with ARMA($p$,$q$) errors such that $p\leq2$ and $q\leq2$.
-Comparing the AIC criteria and using the corresponding likelihood ratio statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$).
+Comparing the AIC criteria and using the corresponding likelihood ratio test statistics, we found that the most appropriate error correlation structure was ARMA($1$,$1$).
 The value of the AIC criterion of the model with the ARMA($1$,$1$) error structure was $2122.2$ while the AIC was $2129.2$ for the model with ARMA($1$,$0$) errors.
@@
 \vspace{1cm}
-
-Following a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity.
-With only two exceptions, the VIF is below $2$ for all predictors in our final models.
-For CV of precipitation, the value of VIF is $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it is equal to $2.182$ in the model based on the subsample of the non-ASAL counties.
-These values are well below the threshold of problematic multicollinearity.
-The other exception is the linear term of seasonal precipitation as its VIF is $2.407$ in the model for the non-ASAL counties, probably due to presence of the squared term of precipitation.
-However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation is below $2$ in the other two models.
-Therefore, we can conclude that the models do not suffer from presence of multicollinearity.
-
-%\item Possibly estimate simple regressions with the same explanatory variables as the preferred mixed effects specification.
-This would give us an approximation of $R^2$ which could be used to compare our models.
 
@@
 
-\begin{itemize}
-\color{blue}
-\item If we get the yield data for the period from $2015$ onwards: Out of sample predictions and comparison with the real data
-\end{itemize}
+Following a standard practise, we calculated the variance inflation factors (VIF) of the explanatory variables to test for presence of multicollinearity.
+With only two exceptions, the VIF was below $2$ for all predictors in our final models.
+For CV of precipitation, the value of VIF was $2.172$ in the model of all counties, $2.147$ in the model for the subsample of the ASAL counties and it was equal to $2.182$ in the model based on the subsample of the non-ASAL counties.
+These values are well below the threshold of problematic multicollinearity.
+The other exception was the linear term of precipitation as its VIF was equal to $2.407$ in the model for the non-ASAL counties, probably due to presence of its squared term in the model.
+However, $2.407$ is below the threshold of the problematic multicollinearity and the VIF of precipitation was below $2$ in the other two models.
+Therefore, we can conclude that the models do not suffer from presence of multicollinearity.
 
-\color{black}
+To evaluate out of sample prediction accuracy of our models we implemented a procedure which is sometimes called the \textit{jackknife}.
+We re-estimated the preferred mixed effects models $n$-times, where $n$ is the number of observations (in our case $1300$ for the sample of all counties and $698$ and $602$ for the subsamples of ASAL and non-ASAL counties, respectively), each time we left out one observation.
+We subsequently used each of these $n$ estimated models to generate a predicted value
+of log of maize yield of the left-out observation.
+After that, we determined the absolute value of the difference between the predicted and real value of the left-out observation for each of the $n$ models.
+In the next step, we calculated average of the $n$ absolute values of the differences between predicted and real values to obtain the mean absolute prediction error (MAPE).
+Finally, we scaled the MAPE values by dividing them by the interquartile range (IQR) of the sample obtained as a difference between the third and first sample quartiles.
+
+Besides obtaining the MAPE of our preferred specification, we worked out MAPE of a baseline model specified as a mixed effects model with the set of fixed effects including two measures, in particular cumulative seasonal precipitation and seasonal average temperature.
+The only random effects of the baseline model were random intercepts as in our preferred specification.
+The baseline model had the error structure ARMA($1$,$1$) as our main model.
+For the sake of comparison, we further calculated MAPE of naive forecasts defined as the last period's actual values of each county.
+
+The scaled MAPE of our preferred specification and the two baseline predictions can be found in table~\ref{CVdiv}.
+All three predictions were obtained for the full-sample model of all counties and also for the models based on the subsamples of ASAL and non-ASAL counties.
+The unscaled values of MAPE (not divided by the IQR) can be found in table~\ref{CV} in the appendix.
+
+\begin{center}
+\begin{threeparttable}
+\caption{\label{CVdiv}	Leave-one-out cross validation: Mean absolute prediction error divided by IQR}
+\linespread{1.3}
+% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
+\begin{indented}
+\item[]
+\begin{tabular}{@{}lccc}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Sample}&\textbf{Preferred specification\tnote{a}}&\textbf{Baseline specification\tnote{b}}&\textbf{Naive forecast\tnote{c}}\\
+\hline
+{All counties}&$0.3811$&$0.3889$&$0.4277$\\
+{ASAL}&$0.5300$&$0.5342$&$0.5835$\\
+{non-ASAL}&$0.3858$&$0.4043$&$0.4510$\\
+\br
+\end{tabular}
+\end{indented}
+ \begin{tablenotes}
+  \begin{footnotesize}
+  \item[a] Mixed effects models: the preferred specification with all fixed effects listed in table~\ref{MainEst}.
+  \item[b] Mixed effects models: the baseline specification, only seasonal precipitation and average seasonal temperature in the fixed effects.
+  \item[c] Last period's actual values used as this period's forecast.
+  \end{footnotesize}
+\end{tablenotes}
+\end{threeparttable} 
+\end{center}
+\vspace{1cm}
+
+According to the MAPE, our preferred specifications perform better than the baseline mixed effects models and they perform considerably better than the naive forecasts for all three samples (table~\ref{CVdiv}).
+For all three prediction methods, the smallest values of MAPE were achieved for the sample of all counties while the least precise forecasts (with regard to MAPE) were obtained for the subsample of ASAL counties (table~\ref{CVdiv}).
 
 \clearpage
-
+\section{Conclusions}\label{Conclusions}
+\clearpage
 \appendix
@@
 
+\begin{center}
+\begin{threeparttable}
+\caption{\label{CV}	Leave-one-out cross validation: Mean absolute prediction error}
+\linespread{1.3}
+% files Rcodes/crossvalidation/compareJustTotals.R; Rcodes/crossvalidation/MAPE_ASAL.R;  Rcodes/crossvalidation/MAPE_nonASAL.R  
+\begin{indented}
+\item[]
+\begin{tabular}{@{}lccc}
+\\[-1em]
+\br
+\\[-1em]
+\textbf{Sample}&\textbf{Preferred specification\tnote{a}}&\textbf{Baseline specification\tnote{b}}&\textbf{Naive forecast\tnote{c}}\\
+\hline
+{All counties}&$0.3599$&$0.3673$&$0.4040$\\
+{ASAL}&$0.4352$&$0.4385$&$0.4790$\\
+{non-ASAL}&$0.2715$&$0.2845$&$0.3173$\\
+\br
+\end{tabular}
+\end{indented}
+
+ \begin{tablenotes}
+  \begin{footnotesize}
+  \item[a] Mixed effects models: the preferred specification with all fixed effects listed in table~\ref{MainEst}.
+  \item[b] Mixed effects models: the baseline specification, only seasonal precipitation and average seasonal temperature in the fixed effects.
+  \item[c] Last period's actual values used as this period's forecast.
+  \end{footnotesize}
+\end{tablenotes}
+\end{threeparttable} 
+\end{center}
+\vspace{1cm}
+
+\clearpage
+
+\doublespacing
+
 \section*{References}
-
 \bibliography{referencesFS}

### 2019-01-16 presentations/January2019/presentation4.tex
asal, nonASAL remove insig
asal, nonASAL remove insig

--- 2018-11-16 presentations/dec2017BMEcSeminar/presentation4.tex
@@
  \begin{itemize}
-
-\item Prolonged absence or marked deficiency of precipitation  
-\item Deficiency of precipitation that results in water shortage for some activity
-or for some group
-\item Period of abnormally dry weather sufficiently prolonged for the lack of precipitation to cause a serious hydrological imbalance 
-\end{itemize}
-
-(\citealp{heim2002,IPCCtrenberth})
-\end{frame}
-
-\note{Heim = a review of the 20tieth century indices, Trenberth = IPCC report, summary for policy makers \\
-
-}
-
-\begin{frame}\label{Definition of Drought}
-\frametitle{Categories of Definition of Drought} 
- \begin{itemize}
-
-\item \textbf{Conceptual definitions:} dictionary types, usually defining boundaries of the concept of drought \\ 
-e.g. \textit{An extended period - a season, a year, or several years of deficient rainfall relative to the statistical multi-year mean for a region \citep{schneider1996}}
-\item \textbf{Operational definitions:} Foundation for an effective early warning system \begin{itemize}
-\item e.g.  Standardized Precipitation Index (SPI), Palmer Drought Severity Index (PDSI)
-\end{itemize} 
-\end{itemize}
-
-(\citealp{wilhite1985,wilhite2000})
-\end{frame}
-
-\note{An example of operational definition of agricultural drought can be obtaining the rate of soil water depletion based on precipitation and evapotranspiration rates and expressing these relationships in terms of drought effects on plant behaviour \citep{wilhite2000} besides PDSI SPI\\
-\vspace{0.3cm}f  \\
-\vspace{0.3cm} f\\
-\vspace{0.3cm}
-
-}
-
-\begin{frame}
-\frametitle{Types of Drought}
-\textit{operational drought} \\
-  \begin{enumerate}
-      \item \textbf{Agricultural drought:} moister deficits in upper layer of
-soil up to about one meter depth
-      \item \textbf{Meteorological drought:} which refers to prolonged
-deficit of precipitation
-      \item \textbf{Hydrological drought:} relates to low stream flow, lake and levels of groundwater
-      \item \textbf{Socioeconomic drought:} associates the supply and demand of some economic good with elements of meteorological, agricultural and hydrological drought
-  \end{enumerate}
-  (\citealp{heim2002,IPCCtrenberth,AMS2013})
-\end{frame}
-
-\note{first three types in AMS 1997, then later the fourth added? \\
-\vspace{0.3cm}first three types physical, the last monetary..(KEYANTASH and dracup 2000, evaluation of drough indecies) \\
-\vspace{0.3cm} f\\
-\vspace{0.3cm}
-
-}
-
-\section{Drought Indices}
-
-\begin{frame}\label{Indices1}
-\frametitle{Drought Indices} 
-\begin{block}{Early measures of drought}
-\begin{itemize}
-\item Length of period without $24$-h precipitation of $1.27$mm \citep{munger1916}
-
-\item Length of drought in days, end of drought defined as $2.54$mm of precipitation in $48$ hours \citep{blumenstock1942}
-\item Measure of precipitation over a given time period \citep{wilhite1985}
-\item Antecedent Precipitation Index (API) based on amount and timing of precipitation, inverse drought index - for flood forecasting \citep{mcquigg1954}
-\end{itemize}
-\end{block}
-
-\end{frame}
-
-\note{these are some examples of early drought indices\\
-\vspace{0.3cm} A numerical measure is needed to compare severity of drought across different time periods
-or geographical locations.
-However, as a result of a large disagreement about a definition of drought, there is no single universal drought index.
-Instead of that a number of measures of drought has been developed\\
-\vspace{0.3cm} f\\
-\vspace{0.3cm}
-
-}
-
-\begin{frame}\label{Indices2}
-\frametitle{Drought Indices}
-
-\begin{itemize}
-\item \textbf{Palmer Drought Severity Index} (PDSI, \citealp{palmer1965})
-
-\begin{itemize}
-
-\item Significant milestone in history of drought severity quantification
-\item Based on a hydrological accounting system
-\item Incorporate antecedent precipitation, moisture supply and moisture demand
-\end{itemize}
-\item \textbf{Standardized Precipitation Index} (SPI, \citealp{SPI})
-\begin{itemize}
-\item Can be interpreted as the number of standard deviations by which the observed value differs from the long-term mean
-\item Standardized departure of observed precipitation from a chosen probability distribution function which models the precipitation data \citep{SPIonline}
-
-\end{itemize}
-
+\item Hi
 \end{itemize}
@@
 
-\note{On short timescales, the SPI is closely related to soil moisture, while at longer timescales, the SPI can be related to groundwater and reservoir storage.\\
-\vspace{0.3cm}SPI: usuallygamma or a Pearson Type III distribution \\
-\vspace{0.3cm} can be used across regions with very different climates\\
-\vspace{0.3cm}SPEI no assumptions about underlining system
-\\
-\vspace{0.3cm} requires more data than SPI
-\\
-\vspace{0.3cm} SPEI sensitive to method that calculate potential evapotranspiration
-\vspace{0.3cm}
-Evapotranspiration (ET) is the sum of evaporation and plant transpiration from the Earth's land and ocean surface to the atmosphere.
-Evaporation accounts for the movement of water to the air from sources such as the soil, canopy interception, and waterbodies.
-}
-
-\begin{frame}\label{Indices3}
-\frametitle{Drought Indices}
-
- \begin{itemize}
- \item \textbf{Standardized Precipitation Evapotranspiration Index} (SPEI, \citealp{SPEI})
-\begin{itemize}
-\item Extention of SPI
-\item Accounts for potential evapotranspiration (hence captures impacts of increased temperature on water demand)
-\end{itemize}
-\item Number of other drought indicators and indices exist
-\begin{itemize}
-\item E.g.: Percent of Normal Precipitation, Drought Area Index (DAI), Soil Moisture Anomaly (SMA), Standardized Water-level Index (SWI), Normalized Difference Vegetation Index (NDVI)
-\end{itemize}
-\item  \textit{For a detailed overview of drought indices see e.g.:} \cite{monacelli2005}, \cite{svoboda2016} or \cite{zargar2011}
-
-\item \cite{keyantash2002}: The quantification of drought: an evaluation of drought indices
-\end{itemize}
-
-\end{frame}
-
-\note{ the examples of indices are from the following categories:meteorology, soil moisture, hydrology, remote sensing \citep{svoboda2016}
-\vspace{0.3cm}f  \\
-\vspace{0.3cm} f\\
-\vspace{0.3cm}}
-
-\section{Effects of Droughts}
-
-\begin{frame}
-
-\frametitle{Effects of droughts on economy}\label{Effects} 
-\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
-
-\begin{itemize}
-
-\item Consists of equations describing model variables and a database (input-output tables, SAM matrix)
-\item Assuming optimizing behaviour (cost minimizing producers, optimal households demands)
-
-\vspace{0.3cm}
-
-\item \underline{\textbf{\cite{robinson2010}}}
-\begin{itemize}
-\item Ethiopia, Social Accounting Matrix (SAM)
-
-\item Model drought as a $20\%$ reduction in crop productivity and $20\%$ reduction of livestock capital
-\item Effects of shocks in production on cereal prices and food consumption
-\item Results in $2.3\%$ decrease in total GDP
-\item Includes a scenario with large-scaled inflow of wheat financed by rest of the world
-\end{itemize}
-\end{itemize}
-\end{block}
-\end{frame}
-
-\note{
-\vspace{0.3cm}the second study: robinson and willenbockel: effects of the 20 percent reductions caused by drought cause 2.3 percent decrease in Ethipia GDP.
-Also models price transition between international and domestic prices..
-\\
-\vspace{0.3cm}robinson: also model food aid (decrease prices of food>>good for all net food purchasers
-\\qe
-\vspace{0.3cm}a\\}
+\note{a\\
+b}
 
@@
 
-\begin{frame}
-
-\frametitle{Effects of droughts on economy}\label{Effects} 
-\begin{block}{\textbf{Computable General Equilibrium (CGE)}}
-\begin{itemize}
-
-\item[~] \underline{\textbf{\cite{OxfamIDS}}}
-\begin{itemize}
-\item Solves the within-country models and between-country trade
-relationships simultaneously
-\item 19 region, 12 sector/commodity group, 8 commodity groups
-\end{itemize}
-\begin{table}
-\begin{footnotesize}
-\begin{tabular}{lll} 
-\hline
-\rowcolor{PaleGreen}  & \textbf{Determined by model}  \\
-\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{(outputs)}  \\
- \hline
-Agricultural productivity growth	& Production volumes of \\
-&  \hspace{0.5cm}food commodities \\
-Commodity shares in hh. expenditure &Production vol. per capita \\
-Shares of import in demand (commodities) & World market  \\
-GDP growth rates&\hspace{0.5cm} food prices (change)\\
-Population growth& Domestic food price (change)\\
-& Volumes of global trade\\
-Shares of food in hh. expenditure  & Aggregate index of
-\\
- & \hspace{0.5cm}consumption p.capita
-\\
- \hline
-\end{tabular}
-\end{footnotesize}
-\end{table}
-
-\end{itemize}
-\end{block}
-\end{frame}
-
-\note{
-\vspace{0.3cm}A CGE model consists of equations describing model variables and a database (usually very detailed) consistent with these model equations\\
-\vspace{0.3cm}
-\vspace{0.3cm}
-}
-
-\begin{frame}
-
-\frametitle{Effects of droughts on economy}\label{Effects}
-
-\begin{itemize}
-
-\item \textbf{\cite{Pedram2016}}
-
-\begin{itemize}
-\item National production losses per disaster (droughts, floods and extreme temperatures)
-\item Worldwide, superposed epoch analysis
-\item On average $10.1\%$ reduction of cereal production can be linked to drought
-\end{itemize}
-\item \textbf{\cite{Mehrabi2017}}
-\begin{itemize}
-\item Cumulative production losses linked to extreme heat and drought events
-\item Per years (53), countries (131), commodities (6)
-\item Most severe losses in Botswana ($5.7\%$), Paraguay ($5.5\%$), Nigeria ($4.8\%$)
-\end{itemize}
-
-\end{itemize}
-
-\end{frame}
-\note{Lesk: 9.1 percent reduction is attributable to extreme heat.
-They did not find any significant effect of extreme cold and 
-oods on production.
-\vspace{0.3cm} mehrabi: estimate counter factual production in disaster years
-\vspace{0.3cm}mehrabi:units are a bit like a panel - years x commodity x country x disaster (disaster years and not disaster years
-\\mehrabi: difference in production of disaster years and non-disaster year average. Cumulative impact: sum per countries and disasters. Loss relatively to (divided by) production + loss of production
-\vspace{0.3cm}mehrabi: the most severe losses angola, botswana, USA, Australie, Paraguay, Angola\\
-\vspace{0.3cm}a
-\vspace{0.3cm}
-}
-
-\section{Effects of Climate}
-
-\begin{frame}
-
-\frametitle{Effects of climate and weather on economy}\label{EffectsCLimate}
-
-\begin{itemize}
-
-\item Effects on food prices
-
-\begin{itemize}
-
-\item \textbf{\cite{Brown2015}}
-\begin{itemize}
-\item  Climate measured by NDVI vegetation index 
-\item  Effects of NDVI and world food prices on local food prices
-\item $20\%$ prices affected by weather, $9\%$ by international prices
-\end{itemize}
-
-\end{itemize}
-
-\item Effects on production (production function approach)
-\begin{itemize}
-\item \textbf{\cite{Ochieng2016}}
-\begin{itemize}
-\item Effect of climate variability and change on crop revenue
-\item Augmented production function, Kenya
-\item  Household FE
-\end{itemize}
-\item \textbf{\cite{Deschenes2007Ric}:} US county level, positive effect of global warming
-\end{itemize}
-\item Effects on land prices or yields/acre (Ricardian analysis)
-\begin{itemize}
-\item  \textbf{\citet{kabubo2007}:} Kenya, warming harmful
-
-\item  \cite{SeoMendelsohn} and \cite{KMendelsohn2008}: mild and wet warming - positive, dry more severe warming - negative
-
-\end{itemize}
-
-\end{itemize}
-\end{frame}
-\note{brown, Khsingar - KALMAN FILTER
-
-\vspace{0.3cm}brown, Khsingar: developing world (Africa, South Asia, Latin America)	\\
-\vspace{0.3cm}Seo et. al: Africa- 16 agro-ecological zones\\
-\vspace{0.3cm}a\\
-\vspace{0.3cm}}
-% tady mozna pridat nejake ricardian studie a rozdelit to na dva slidy
-
 \section{My Suggestion}
@@
 
-\end{itemize}
-
-\end{frame}
-
-\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
-\vspace{0.3cm}a\\
-\vspace{0.3cm}}
-
-\begin{frame}
-
-\frametitle{My suggestion - panel estimation}\label{MySuggestion2}
-
-My interest: \textbf{Effects of drought on economy in Kenya}
-\begin{itemize}
-
-\item \textbf{Explanatory variable of interest}
-
-\begin{itemize}
-\item Dummy variable (0/1) drought occurred in a particular county and year or not
-\item Several varieties - various specifications of drought:
-
-\begin{itemize}
-\item Drought index monthly - NDVI, SPI, SPEI 
-\item Drought index below a specific value ($-1$, $-2$ or $-3$)
- \begin{description}[abc]  % for indentation of length of abc
- \item[$\bullet$] At least for one month during a specific year
- \item[$\bullet$] At least for one month during growing season in a specific year
- \item[$\bullet$] At least for one month during long rains in a specific year
-  \item[$\bullet$] At least for two (three) consecutive months during a growing season in a specific year
- \item[$\bullet$] At least for one month during two (three) consecutive growing seasons
- \end{description}
-
-\end{itemize}
-
-\end{itemize}
 \end{itemize}

### 2019-01-16 Rcodes/January2019/finalsASALnoInsig.R

@@ -1,0 +1,66 @@
+rm(list=ls())
+library(cvTools);library(MASS);library(dplyr); library(ggeffects);library(tseries); library(plm); library(nlme); library(lme4); library(lattice); library(car); library(lmerTest); library(optimx)
+load("dataFS/Main/DaTS.RData")
+# load("Rcodes/DecemberNew/KEN11d_stepNice.RData")
+#oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+# these models are the best ones for now:
+
+KEN11d<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+            +AvgTemp + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+            data=ScaledTS,na.action=na.exclude); summary(KEN11d)
+anova(KEN11d,type="marginal")
+
+# how does it look if I replace temperature in Celsius with temperature in Kelvin??
+KEN11dK<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+             +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
+             data=ScaledTS,na.action=na.exclude); summary(KEN11dK)
+# quite similar, but not yet completely the same...
+
+anova(KEN11dK,type="marginal")
+exp(summary(KEN11dK)$coef[[1]])
+#---  -- --  --  -- -- -- -- -- -- -- - -- -- -- - -- - - - -- -- --  -- --  --  -- -- -- -- -- -- -- - -- -- -- - -- - - - -- -- --  -- --  --  -- -- -- -- -- -- -- - -- -- -- - -- - - - -- -- -
+# now ASAL and nonASAL separately
+
+KEN11dK_ASAL<-lme(log(Yield0)~SeasPr+I(SeasPr^2)+CVPrec+Spell+Spell4
+                  +AvgTempK + CVTempK, random= ~1 | ID1,correlation=corARMA(form = ~ as.numeric(Year)|ID1, p=1,q=1),
[…]

### 2019-01-16 presentations/January2019/presentation4.tex
presentation seminar Annemie

@@
 
-\title{Drought and its Effects on Economy}  
-\author{Monika Novackova} 
-\date{December 2017} 
+\title{Effects of Weather on Maize Yield in Kenya}  
+\author{Monika Novackova \\ \footnotesize supervised by: \textcolor{darkblue}{Pedram Rowhani, Martin Todd and Annemie Maertens}}
+
+\date{January 2019} 
 \institute{University of Sussex}
@@
 
-\begin{frame}\label{My Interest}
-
-\large{My research interest:} \\
-\vspace{0.7cm}
-
-\large{\hspace{1cm}\textbf{Effects of droughts on economy in Kenya}}
-\end{frame}
-\note{A}
-
 \begin{frame}
@@
 
-\section{Definition of Drought}
+\section{Introduction}
 
-\begin{frame}\label{Definition of Drought}
-\frametitle{Definition of Drought} 
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}\label{Introduction}
+\frametitle{Introduction} 
  \begin{itemize}
-\item Hi
+\item SSRP project, related to ForPAc project
+\item Extreme weather $\rightarrow$ disasters \textcolor{red}{(maybe name some years of drought in Kenya??)}
+
+\begin{itemize}
+\item Early warning systems have been developed
+\end{itemize}
+\item Goals: Improve early warning systems in Kenya
+\item Shifting the disaster management from reactive to protective
+\end{itemize}
+
+\begin{itemize}
+\item Various approaches considered including:
+
+\begin{itemize}
+\item Relating NDMA Early warning phase classification to weather
+\item Relating markets and food prices to weather
+\item Relating maize yields to weather to weather
+\item Computable general equilibrium (CGE) models
+\end{itemize}
+\end{itemize}
+\end{frame}
+
+\note{Climate change has been increasing amplitudes and frequency of extreme weather events\\ Not every hazard turns into disaster\\
+Relating NDMA Early warning classification to weather also various precip and weather indexes
+r
+b}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}
+       \hspace{-0.8cm}  \textbf{NDMA Early Warning Phase\\ \hspace{-0.8cm} Classification}
+
+\vspace{-0.9cm}
+     \begin{figure}
+      \begin{columns}
+        \column{.4\linewidth}
+
+  \caption*{\textcolor{DarkGreen}{\textbf{\underline{\small{Example: Kitui county}}}}\\
+
+ \vspace{0.25cm}\hspace{-0.25cm} \textbf{\textit{\textcolor{blue}{Normal $\rightarrow$ Alert $\rightarrow$  Alarm}}} \\
+   \vspace{0.4cm}  
+ \hspace{-0.25cm}  $\bullet$ \small{Bulletins for ASAL counties} \\
+   \hspace{-0.25cm}  $\bullet$ \small{Online since 2013} \\
+     \vspace{0.1cm}  
+  \hspace{0.15cm}    \footnotesize{$\bullet$ But in pdf format} \\ 
+    \vspace{0.1cm}  
+\hspace{-0.25cm}   \small{$\bullet$ For county and month}
+
+ }
+        \column{.6\linewidth}
+
+    \vspace{-0.16cm}   \hspace{-0.1cm}   \includegraphics[width = 0.9\textwidth]{PhaseNA.png}  %565*238
+ \vspace{-0.3cm}  
+      \end{columns}
+    \end{figure}
+
+\vspace{-1.3cm}
+    \begin{figure}
+      \begin{columns}
+        \column{.5\linewidth }
+
+  \vspace{-1.8cm}        
+        \column{.5\linewidth}
+
+  \hspace{-2.6cm}  \includegraphics[width = 1.2\textwidth]{PrecNA.png}  %565*238
+      \end{columns}
+    \end{figure}
+
+\end{frame}
+
+\note{This is an example of one of the approaches that we eventually abandoned\\
+b}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}\label{Approach}
+\frametitle{Approach}
+
+\begin{itemize}
+\item Narrowing the research question:
+\begin{itemize}
+\item Relationship of maize yield and weather in Kenya
+
+\begin{itemize}
+\item What is it about weather that causes drought related disasters?
+\item Which particular characteristics of weather are the most 'responsible' for drought related disasters?
+\end{itemize}
+\end{itemize}
+\end{itemize}
+
+Sample:
+\begin{itemize}
+\item Panel of 47 counties of Kenya, $1981-2017$
 \end{itemize}
@@
 
-\note{a\\
-b}
+\note{After various problems with data (prices - not for every county, NDMA phases - not much of an objective criterion, probably political, various VCI values)
+\\ so the decision about the approach was partially driven by data (un)availability}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 
@@
 
-\frametitle{My suggestion - panel estimation}\label{MySuggestion3}
-
-My interest: \textbf{Effects of drought on economy in Kenya}
-\begin{itemize}
-
-\item \textbf{Explanatory variable of interest}
-
-\begin{itemize}
-\item Dummy variable (0/1) drought occurred in a particular county and year or not
-
-\end{itemize}
-
-\item \textbf{Units of analysis:} Counties in Kenya $\times$ year
-
-\item \textbf{Control variables ? - subject to availability}
-
-\begin{itemize}
-\item GDP, soil quality data, population,average land area, average value assets,  climate, degree days?
-\end{itemize}
-
-\item \textbf{Estimation methods}
-
-\begin{itemize}
-\item Fixed effects, SURE, Kalman filter, Box-Jenkins
-\end{itemize}
-
-\end{itemize}
-
-\end{frame}
-
-\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
-\vspace{0.3cm}a\\
-\vspace{0.3cm}}
-
-\begin{frame}
-
 \frametitle{My idea - panel estimation}\label{MySuggestion4}

### 2019-01-16 presentations/January2019/presentation4.tex
presentation first draft

@@
 \begin{itemize}
+\item Two predominant rainfall regimes in Kenya:
+
+\begin{enumerate}
+\item Arid and semi-Arid (ASAL): bi-modal
+	\begin{itemize}
+	\item MAM and OND
+	\item \textcolor{red}{Short and Long rains}
+	\end{itemize}
+\item non-ASAL: uni-modal
+	\begin{itemize}
+	\item March to August
+	\end{itemize}
+
+\end{enumerate}
+\item \textcolor{red}{Perhaps a map of ASAL and non-ASAL}
+
+\end{itemize}
+\end{frame}
+
+\note{Climate change has been increasing amplitudes and frequency of extreme weather events\\ Not every hazard turns into disaster
+r
+b}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}\label{Intro-Approaches}
+\frametitle{Intro-Approaches}
+
+\begin{itemize}
 \item Various approaches considered including:
@@
 
-\note{Climate change has been increasing amplitudes and frequency of extreme weather events\\ Not every hazard turns into disaster\\
+\note{
 Relating NDMA Early warning classification to weather also various precip and weather indexes
@@
 b}
-
 %oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
@@
 \item Relationship of maize yield and weather in Kenya
-
 \begin{itemize}
@@
 \\ so the decision about the approach was partially driven by data (un)availability}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}\label{Data}
+\frametitle{Data}
+
+\begin{itemize}
+\item Maize yields
+\begin{itemize}
+\item source: Famine Early Warning Systems Network (FEWS NET)
+\item County level, yearly, tonnes per hectare
+\end{itemize}
+\item Weather:
+\begin{itemize}
+\item Daily, $0.25^{\circ}$ resolution gridded data
+\item Precipitation
+	\begin{itemize}
+	\item Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS)
+	\end{itemize}
+\item Temperature
+	\begin{itemize}
+	\item Berkeley Earth
+	\end{itemize}
+	\item[$\blacksquare$] Aggregation needed to conform with the yield data
+\end{itemize}
+
+\end{itemize}
+
+\end{frame}
+
+\note{After various problems with data (prices - not for every county, NDMA phases - not much of an objective criterion, probably political, various VCI values)
+\\ so the decision about the approach was partially driven by data (un)availability \\ precip in mm, not sure if to put it into presentation as I would have to talk about units of temperature, mostly Kelvin but sometimes Celsius}
 

### 2019-01-16 presentations/January2019/presentation4.tex
presentation

@@
 \begin{enumerate}
-\item Arid and semi-Arid (ASAL): bi-modal
+\item Arid and semi-arid (ASAL): bi-modal
 	\begin{itemize}
-	\item MAM and OND
-	\item \textcolor{red}{Short and Long rains}
+	\item Short rains: March to May
+	\item Long rains: October to December
 	\end{itemize}
@@
 
-\begin{frame}\label{Intro-Approaches}
-\frametitle{Intro-Approaches}
+\begin{frame}\label{Approaches00}
+\frametitle{Approach}
 
@@
 
-\begin{frame}
+\begin{frame}\label{Approaches01}
        \hspace{-0.8cm}  \textbf{NDMA Early Warning Phase\\ \hspace{-0.8cm} Classification}
@@
         \column{.5\linewidth }
-
   \vspace{-1.8cm}        
         \column{.5\linewidth}
-
   \hspace{-2.6cm}  \includegraphics[width = 1.2\textwidth]{PrecNA.png}  %565*238
@@
 %oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
-
+\section{Data}
 \begin{frame}\label{Data}
@@
 	\begin{itemize}
-	\item Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS)
+	\item source: Climate Hazards Group InfraRed Precipitation with Station data (CHIRPS)
 	\end{itemize}
@@
 	\begin{itemize}
-	\item Berkeley Earth
+	\item source: Berkeley Earth
 	\end{itemize}
-	\item[$\blacksquare$] Aggregation needed to conform with the yield data
+	\item[$\blacksquare$] \textcolor{darkblue}{Aggregation needed to conform with the yield data}
 	\end{itemize}
@@
 \end{itemize}
+
+\begin{columns}
+\footnotesize
+	\begin{column}{0.4\textwidth}
+
+	\end{column}
+	\begin{column}{0.15\textwidth}
+  		$0.25^{\circ}$ grid \\ daily
+	\end{column}
+	\begin{column}{0.05\textwidth}
+  			$\rightarrow$ \\ $\rightarrow$
+	\end{column}
+	\begin{column}{0.25\textwidth} 
+   		county level \\  yearly
+   	\end{column}
+   		\begin{column}{0.4\textwidth} 
+   		\end{column}
+\end{columns}
 
@@
 \section*{}
-
 \frame[allowframebreaks]{ 
@@
 
-\frame[allowframebreaks]{ 
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+% 		---                	THE END       ---           ---
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\newgeometry{top=10mm, left=10mm, right=10mm, bottom=10mm}
+
+\frame[allowframebreaks, plain]{ 
 \small
-\frametitle{References}
-\bibliographystyle{apa}
-\bibliography{referencesFS}
-}
+\frametitle{Precipitation and temperature measures considered}
 
+\begin{itemize}
+
+\item Total precipitation over the rainy season\\
+\item Coefficient of variation of the precipitation during the rainy season\\
+\item Maximum length of dry spell during the rainy season (in number of days)\\
+\item Number of dry spells during the rainy season: a dry spell defined as $4$ consecutive days without rain or more\footnotemark[1]\\
+\item Number of dry spells during the rainy season: a dry spell defined as $10$ consecutive days without rain or more\footnotemark[1]\\
+\item Number of dry spells during the rainy season: a dry spell defined as $20$ consecutive days without rain or more\footnotemark[1]\\
+
+\end{itemize}
+
+\footnotetext[1]{Threshold for a dry day was considered $1$mm.}
+\end{frame}
+
+\frame[allowframebreaks,plain]{ 
+\small
+\frametitle{Precipitation and temperature measures considered}
+
+\begin{itemize}
+
+\item Average temperature during the rainy season\\
+\item Standard deviation of temperature during the rainy season\\
+\item Cumulative degree days during the rainy season (excluding the days when maximum temperature above $30^{\circ}$C or below $10^{\circ}$C\\
+\item Number of heatwave days during the rainy season when max. temperature $>35^{\circ}$C
+\\
+\item Maximum daily precipitation  - to control for possible floods
+\\
+\item Sum of precipitation amount on days where precipitation is above $90^{th}$ percentile of precip. of the whole period\footnotemark[2]\\
+\item Sum of precipitation amount on days where precipitation is above $95^{th}$ percentile of precip. of the whole period\footnotemark[3]\\
+\end{itemize}
+\footnotetext[2]{$90^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.}
+\footnotetext[3]{$95^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.}
+\end{frame}
+
+\frame[allowframebreaks,plain]{ 
+\small
+\frametitle{Precipitation and temperature measures considered}
+
+\begin{itemize}
+\item Sum of precipitation amount on days where precipitation is above $99^{th}$ percentile of precip. of the whole period\footnotemark[4]\\
+\item Number of days where precipitation is above $90^{th}$ percentile of precip. of the whole period\footnotemark[2]\\
+\item Number of days where precipitation is above $95^{th}$ percentile of precip. of the whole period\footnotemark[3]\\
+
+\item Number of days where precipitation is above $99^{th}$ percentile of precip. of the whole period\footnotemark[4]\\
+
+\end{itemize}
+
+\footnotetext[2]{$90^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.}
+\footnotetext[3]{$95^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.}
+\footnotetext[4]{ $99^{th}$ percentile was calculated for the subsample of wet days, that is the days where precipitation is above or equal to $1$mm.}
+\end{frame}
 \end{document}

### 2019-01-16 presentations/January2019/presentation4.tex
presentation

@@
 \end{columns}
+\textcolor{red}{add how averaged over seasons ASAL and non ASAL}
 
@@
 %oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}\label{measures} 
+\small
+\frametitle{Precipitation and temperature measures}
+\begin{itemize}
+\item Measures typically considered:
+
+	\begin{itemize}
+		\item Total precipitation over rainy seasons
+		\item Average temperature during rainy seasons
+		\item Average maximum daily temperature during rainy seasons
+	\end{itemize}
+
+\item{Less commonly used measures considered:}
+	\begin{itemize}
+		\item Coefficients of variation: precipitation and temperature
+		\item Maximum length of dry spell during rainy seasons
+		\item Number of dry spells during rainy seasons
+		\item Cumulative degree days over rainy seasons
+		\item Number of heatwave days
+		\item Cumulative precipitation on days when precip. $>90th$ percentile
+	\end{itemize}
+
+\end{itemize}
+\end{frame}
+
+\note{dry day defined when temperature below 1mm\\various definitions of dry spell (5days, 10days, 20days)\\ degree days between 10 and 30C\\ heatwave days above 35C}
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}
+
+\frametitle{ideas}
+\color{red}
+\begin{itemize}
+\item Plot of dependency of precip or some measure and yields..
+
+\end{itemize}
+\color{black}
+\end{frame}
 

### 2019-01-16 presentations/January2019/presentation1.tex
presentation

--- 2019-01-16 presentations/January2019/presentation4.tex
@@
 		\item Average temperature during rainy seasons
-		\item Average maximum daily temperature during rainy seasons
 	\end{itemize}
@@
 	\begin{itemize}
+		\item Maximum daily precipitation (floods)
+		\item Maximum daily temperature
 		\item Coefficients of variation: precipitation and temperature
@@
 		\item Number of dry spells during rainy seasons
-		\item Cumulative degree days over rainy seasons
 		\item Number of heatwave days
@@
 
+\section{Methods}
+
+\begin{frame}
+\frametitle{Linear mixed effects models}\label{Methods}
+
+\begin{itemize}
+\item Panel: County $\times$ Year
+\item Slopes allowed to vary over counties
+\end{itemize}
+
+\begingroup
+\everymath{\small}
+
+\begin{align*}
+\hspace{-0.5cm} \begin{array}{lclll}
+ log(y_{\scriptscriptstyle 1t} )&=&\sum_{\scriptscriptstyle m=1}^{\scriptscriptstyle p} \beta^{\scriptscriptstyle m}x^{\scriptscriptstyle m}_{\scriptscriptstyle 1t}+ \sum_{\scriptscriptstyle n=1}^{\scriptscriptstyle q}b^{\scriptscriptstyle n}_{\scriptscriptstyle 1}z^{\scriptscriptstyle n}_{\scriptscriptstyle 1t}+\epsilon_{\scriptscriptstyle 1t} \\[2mm]
+ log(y_{\scriptscriptstyle 2t} )&=&\sum_{\scriptscriptstyle m=1}^{\scriptscriptstyle p} \beta^{\scriptscriptstyle m}x^{\scriptscriptstyle m}_{\scriptscriptstyle 2t}+ &\sum_{\scriptscriptstyle n=1}^{\scriptscriptstyle q}b^{\scriptscriptstyle n}_{\scriptscriptstyle 2}z^{\scriptscriptstyle n}_{\scriptscriptstyle 2t}+\epsilon_{\scriptscriptstyle 2t}\\[1mm]
+  &\huge{\vdots}\\[-2mm]
+  & \huge{\vdots} \\[1mm]
+  log(y_{\scriptscriptstyle 47t} )&=&\sum_{\scriptscriptstyle m=1}^{\scriptscriptstyle p} \beta^{\scriptscriptstyle m}x^{\scriptscriptstyle m}_{\scriptscriptstyle 47t}+ &&\sum_{\scriptscriptstyle n=1}^{\scriptscriptstyle q}b^{\scriptscriptstyle n}_{\scriptscriptstyle 47}z^{\scriptscriptstyle n}_{\scriptscriptstyle 47t}+\epsilon_{\scriptscriptstyle 47t}
+\end{array}
+\end{align*}
+\endgroup
+
+\begin{table}
+\begin{small}
+\begin{tabular}{lll}
+
+$y_{it}$ &$=$& Maize yield t/hectares, county $i$ in year $t$\\
+$\boldsymbol{\beta}$ &$=$& Vector of effects of other covariates\\
+$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
+$\epsilon_{i,t}$ &$=$& Error term \\
+\end{tabular}
+\end{small}
+\end{table}
+
+\end{frame}
+
+\note{Maybe talk about ambigous methodology: the same words used for different terms//a}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}
+\frametitle{Linear mixed effects models}\label{Methods}
+
+\begin{itemize}
+\item Panel: County $\times$ Year
+\item Slopes allowed to vary over counties
+\end{itemize}
+
+\begin{columns}
+
+\begin{column}{0.5\textwidth} 
+\begingroup
+\everymath{\Large}
+\begin{align*}
+\begin{array}{lcl}
+ log(y_{\scriptscriptstyle it} )&=&\sum_{\scriptscriptstyle m=1}^{\scriptscriptstyle p} \beta^{\scriptscriptstyle m}x^{\scriptscriptstyle m}_{\scriptscriptstyle it}+ \sum_{\scriptscriptstyle n=1}^{\scriptscriptstyle q}b^{\scriptscriptstyle n}_{\scriptscriptstyle i}z^{\scriptscriptstyle n}_{\scriptscriptstyle it}+\epsilon_{\scriptscriptstyle it}
+\end{array}
+\end{align*}
+\endgroup
+\end{column}
+
+\begin{column}{0.5\textwidth} 
+\begingroup
+\everymath{\footnotesize}
+\begin{align*}
+\begin{array}{ll}
+i=1,...,47 & \text{counties}
+\end{array}
+\end{align*}
+\endgroup
+\end{column}
+\end{columns}
+
+\begin{table}
+\begin{small}
+\begin{tabular}{lll}
+
+$y_{it}$ &$=$& Maize yield t/hectares, county $i$ in year $t$\\
+$\alpha_i$ &$=$& Fixed effects), county $i$\\
+{\color{darkblue}$\delta$ }&{\color{darkblue}$=$}&{\color{darkblue}{Effect of drought on economy}}\\
+
+$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
+$\epsilon_{i,t}$ &$=$& Error term \\
+\end{tabular}
+\end{small}
+\end{table}
+
+\end{frame}
+
+\note{Maybe talk about ambigous methodology: the same words used for different terms//a}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
 \begin{frame}
@@
 \frametitle{ideas}
+
+\begin{itemize}
 \color{red}
-\begin{itemize}
 \item Plot of dependency of precip or some measure and yields..
-
+\item Maybe also general description of the mixed effects models as I have in the draft
+\item Look at the document on google docs 'Yield and climate' to see if I can use something
 \end{itemize}
@@
 \end{frame}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
 

### 2019-01-16 presentations/January2019/presentation1.tex
presssssssee

@@
 \usepackage{amssymb}
+ \usepackage{setspace}
 
@@
 \begin{frame}
-\frametitle{Linear mixed effects models}\label{Methods}
+\frametitle{\small Linear mixed effects models}\label{Methods}
 
 \begin{itemize}
-\item Panel: County $\times$ Year
-\item Slopes allowed to vary over counties
+\item \small Slopes allowed to vary across counties
 \end{itemize}
@@
 \begin{table}
-\begin{small}
+\begin{footnotesize}
+\setstretch{1.23}
 \begin{tabular}{lll}
 
-$y_{it}$ &$=$& Maize yield t/hectares, county $i$ in year $t$\\
-$\boldsymbol{\beta}$ &$=$& Vector of effects of other covariates\\
-$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
-$\epsilon_{i,t}$ &$=$& Error term \\
+$t$&$=$&index for year ($1981-2017$) \\
+$y_{\scriptscriptstyle it}$ &$=$&maize yield tones/hectares, county $i$, year $t$\\
+${\beta}^{m}$, \footnotesize{\textit{m=1,...p}}&$=$&\textit{p} fixed effects parameters\\
+$x^{m}_{it}$, \footnotesize{\textit{m=1,...,p}}, \footnotesize{\textit{i=1,...,47}} &$=$&fixed effects regressors, county \textit{i}, year \textit{t}\\
+$b^{n}_i$, \footnotesize{\textit{n=1,...q}}, \footnotesize{\textit{i=1,...,47}}&$=$&\textit{q} random effects parameters -\textbf{vary over county \textit{i}}\\
+$z^{n}_{it}$, \footnotesize{\textit{n=1,...,q}}, \footnotesize{\textit{i=1,...,47}} &$=$&random effects regressors, county \textit{i}, year \textit{t}\\
+$\epsilon_{it}$ &$=$&error term \\
 \end{tabular}
-\end{small}
+\end{footnotesize}
 \end{table}
@@
 
-\begin{column}{0.5\textwidth} 
+\begin{column}{0.6\textwidth} 
 \begingroup
@@
 
-\begin{column}{0.5\textwidth} 
+\begin{column}{0.4\textwidth} 
 \begingroup
@@
 \begin{table}
-\begin{small}
+\begin{footnotesize}
+\setstretch{1.23}
 \begin{tabular}{lll}
 
-$y_{it}$ &$=$& Maize yield t/hectares, county $i$ in year $t$\\
-$\alpha_i$ &$=$& Fixed effects), county $i$\\
-{\color{darkblue}$\delta$ }&{\color{darkblue}$=$}&{\color{darkblue}{Effect of drought on economy}}\\
-
-$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
-$\epsilon_{i,t}$ &$=$& Error term \\
+$t$&$=$&index for year ($1981-2017$) \\
+$y_{\scriptscriptstyle it}$ &$=$&maize yield tones/hectares, county $i$, year $t$\\
+${\beta}^{m}$, \footnotesize{\textit{m=1,...p}}&$=$&\textit{p} fixed effects parameters\\
+$x^{m}_{it}$, \footnotesize{\textit{m=1,...,p}}, \footnotesize{\textit{i=1,...,47}} &$=$&fixed effects regressors, county \textit{i}, year \textit{t}\\
+$b^{n}_i$, \footnotesize{\textit{n=1,...q}}, \footnotesize{\textit{i=1,...,47}}&$=$&\textit{q} random effects parameters -\textbf{vary over county \textit{i}}\\
+$z^{n}_{it}$, \footnotesize{\textit{n=1,...,q}}, \footnotesize{\textit{i=1,...,47}} &$=$&random effects regressors, county \textit{i}, year \textit{t}\\
+$\epsilon_{it}$ &$=$&error term \\
 \end{tabular}
-\end{small}
+\end{footnotesize}
 \end{table}

### 2019-01-17 presentations/January2019/presentation1.tex
presentation
work

@@
  \usepackage{setspace}
-
+\usepackage{changepage}
 \usetheme{Singapore}
@@
 \title{Effects of Weather on Maize Yield in Kenya}  
-\author{Monika Novackova \\ \footnotesize supervised by: \textcolor{darkblue}{Pedram Rowhani, Martin Todd and Annemie Maertens}}
+\author{Monika Novackova \\[2mm] \footnotesize supervised by: \textcolor{darkblue}{Pedram Rowhani, Martin Todd and Annemie Maertens}}
 
@@
 
+\frametitle{Model specification}\label{spec}
+
+Error structure, ARMA, AIC
+
+\end{frame}
+
+\note{
+
+\vspace{0.3cm}
+\vspace{0.3cm}a
+
+}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\section{Results}
+
+\begin{frame}
+
+\frametitle{Results}\label{Results} 
+{\begin{threeparttable}
+\caption{\normalsize{\textbf{ {Mixed  effects model:}} exponents of the coefficient estimates Log of maize yield and weather, ARMA(1,1) errors}}
+%KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL										
+\label{KenARe11_exponents} 
+\setstretch{1.33}
+\begin{tabular}{@{}lllllll} 
+ \\
+  \textbf{Fixed effects:}&\textit{\textbf{All counties}}&\textit{\textbf{ASAL}}&\textit{\textbf{non-ASAL}}\\
+\vspace{-0.2cm}Intercept&$1.296^{***}$&$1.272^{*}$&$1.408^{**}$\\
+ \vspace{-0.2cm}Prec. total&$1.081^{*}$&$1.007^{}$&$1.277^{***}$\\
+  \vspace{-0.2cm}Prec. total sq.&$0.973^{*}$&$1.003$&$0.881^{***}$\\
+ \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$0.970$ &$0.907^{}$\\
+ \vspace{-0.2cm}Dry spell -length\tnote{a}&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
+ \vspace{-0.2cm}Dry spells $\geq~4$ d.\tnote{b}&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
+ \vspace{-0.2cm}Temp. - average&$0.820^{***}$&$0.808^{*}$&$0.880$ $^{}$\\
+ \vspace{-0.2cm}Temp. c. of var.&$1.043^{\bullet}$&$1.032$&$1.060$ ${*}$\\
+  \hline
+\textit{Number of observations:}  &\multicolumn{2}{l}{$1300$}&\multicolumn{2}{l}{$698$}&\multicolumn{2}{l}{$602$}
+\\
+\end{tabular} 
+ \begin{tablenotes}
+  \begin{footnotesize}
+    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+            \begin{adjustwidth}{1cm}{} 
+  \item[a] Length of the longest dry spell in number of days.
+    \item[b] Number of dry spells lasting for four days or longer. 
+     \end{adjustwidth}
+  \end{footnotesize}
+\end{tablenotes}
+  \end{threeparttable}\par }
+\end{frame}
+
+\note{
+
+\vspace{0.3cm}
+\vspace{0.3cm}a
+
+}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}
+
 \frametitle{ideas}
@@
 \vspace{0.3cm}a
-\\
-\vspace{0.3cm}a
-\\
-\vspace{0.3cm}a
+
 }
-
-\begin{frame}
-
-\frametitle{Effects of droughts on economy}\label{Effects} 
-\begin{block}{\textbf{Computable General Equilibrium (CGE)} Models }
-\begin{itemize}
-
-\item \underline{\textbf{\cite{OxfamIDS}}}
-\begin{itemize}
-\item Exploring range of scenarios for food price increase in 2030
-\begin{itemize}
-\item 1. Baseline 2. Climate change 3. Climate change with adaptation 4. Adaptation only in sub-Saharan Africa
-\end{itemize}
-\item Global coverage, set of individual country models, linked through international trade
-\item Climate change (incl. drought) modelled as changes in factor productivity (usually negative)
-
-\end{itemize}
-
-\end{itemize}
-\end{block}
-\end{frame}
-
-\note{A}
-
-\section{My Suggestion}
-
-\begin{frame}
-
-\frametitle{My suggestion - panel estimation}\label{MySuggestion1}
-
-My interest: \textbf{Effects of drought on economy in Kenya}
-\begin{itemize}
-
-\item \textbf{Response variable}
-
-\begin{itemize}
-\item Volumes of production (crop specific, total)
-\item Profit per acre \normalsize{\citep{Deschenes2007Ric}}
-
-\begin{itemize}
-\item \footnotesize{(Value of agricultural products - prod. expanses)/acres (crops, pasture, grazing)}\\ 
-\end{itemize}
-\end{itemize}
-\item \textbf{Units of analysis}
-\begin{itemize}
-\item Counties in Kenya $\times$ year
-\
-\end{itemize}
-
-\item \textbf{Explanatory variable of interest}
-
-\begin{itemize}
-\item Dummy variable (0/1) drought occurred in a particular county and year or not
-\item Several varieties - various specifications of drought:
-
-\end{itemize}
-
-\end{itemize}
-
-\end{frame}
-
-\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
-\vspace{0.3cm}a\\
-\vspace{0.3cm}}
-
-\begin{frame}
-
-\frametitle{My idea - panel estimation}\label{MySuggestion4}
-
-\begin{align*}
-\begin{array}{lcl}
- Y_{i,t} &=& \alpha_i +\gamma_t + \delta D'_{i,t} +  \boldsymbol{\beta X_{i,t}}+ \epsilon_{i,t}
-\end{array}
-\end{align*}
-
-\begin{table}
-\begin{small}
-\begin{tabular}{lll}
-
-$Y_{i,t}$ &$=$& Response variable (food production/price), county $i$ in year $t$\\
-$\alpha_i$ &$=$& Fixed effects, county $i$\\
-{\color{darkblue}$\delta$ }&{\color{darkblue}$=$}&{\color{darkblue}{Effect of drought on economy}}\\
-$D_{i,t}$ &$=$& Indicator variable\\
-& &  $D=1$ if drought in county $i$ in year $t$, $D=0$ otherwise\\
-$\boldsymbol{\beta}$ &$=$& Vector of effects of other covariates\\
-$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
-$\epsilon_{i,t}$ &$=$& Error term \\
-{\color{DarkGreen}$\gamma_t$} & {\color{DarkGreen}$=$}& {\color{DarkGreen}Year specific indicator?}
-\end{tabular}
-\end{small}
-\end{table}
-
-\end{frame}
-
-\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc o}
-
-\note{hello}
 

### 2019-01-17 presentations/January2019/presentation1.tex
slides making

@@
 
-Error structure, ARMA, AIC
+\begin{itemize}
+\item AIC
+\item None of the random slopes significant
+\item Group autocorrelation structure present
+
+\begin{itemize}
+\item Tested by the Lagrange multipliear test developed by Baltagi \& Li ($1991$; $1995$)
+\item[$\boldsymbol{\rightarrow}$] The error structure modelled as ARMA($1$,$1$) 
+\begin{itemize}
+\item The preferred error structure found using AIC
+\end{itemize}
+\end{itemize}
+
+\end{itemize}
 
@@
 \note{
-
-\vspace{0.3cm}
-\vspace{0.3cm}a
+First set of fixed effects found- stepwise like procedure. Then test of significant of random slopes
+\vspace{0.3cm}to verify: alternative step-down model building approach
+using the Satterthwaites method to determine the p-values of the individual t-tests
 

### 2019-01-17 presentations/January2019/presentation1.tex
sllides presentation friday

@@
 \item Tested by the Lagrange multipliear test developed by Baltagi \& Li ($1991$; $1995$)
-\item[$\boldsymbol{\rightarrow}$] The error structure modelled as ARMA($1$,$1$) 
+\end{itemize}
+\color{darkblue}
+\rotatebox[origin=c]{270}{\Huge{$\boldsymbol{	\rightsquigarrow}$}}
+\color{black}
+\\[2mm]The error structure modelled as ARMA($1$,$1$) 
 \begin{itemize}
-\item The preferred error structure found using AIC
-\end{itemize}
+\item The preferred error structure found using the Likelihood Ratio statistics
+
 \end{itemize}
@@
 using the Satterthwaites method to determine the p-values of the individual t-tests
-
+\vspace{0.3cm} the error structure found as.. Likelihood ratio statistic. Comparison of AIC, BIC, loglik
 }
@@
 
-\frametitle{Results}\label{Results} 
+\frametitle{Mixed effects estimates:}\label{Results} 
 {\begin{threeparttable}
-\caption{\normalsize{\textbf{ {Mixed  effects model:}} exponents of the coefficient estimates Log of maize yield and weather, ARMA(1,1) errors}}
+\begin{footnotesize}
+\caption*{Yield as dependent variable\hfill exponents of the coefficients: $\exp{\beta}$}
 %KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL										
@@
 \setstretch{1.33}
-\begin{tabular}{@{}lllllll} 
- \\
+\begin{tabular}{llll} 
   \textbf{Fixed effects:}&\textit{\textbf{All counties}}&\textit{\textbf{ASAL}}&\textit{\textbf{non-ASAL}}\\
@@
  \vspace{-0.2cm}Prec. c. of var.&$0.924^{\bullet}$&$0.970$ &$0.907^{}$\\
- \vspace{-0.2cm}Dry spell -length\tnote{a}&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
- \vspace{-0.2cm}Dry spells $\geq~4$ d.\tnote{b}&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
+ \vspace{-0.2cm}Dry spell -length in days&$0.935^{*}$&$0.834^{**}$&$ 0.988^{}$\\
+ \vspace{-0.2cm}Dry spells $\geq~4$ days\tnote{b}&$0.939^{*}$&$0.855^{**}$&$0.989^{}$\\
  \vspace{-0.2cm}Temp. - average&$0.820^{***}$&$0.808^{*}$&$0.880$ $^{}$\\
- \vspace{-0.2cm}Temp. c. of var.&$1.043^{\bullet}$&$1.032$&$1.060$ ${*}$\\
+ \vspace{-0.2cm}Temp. c. of var.&$1.043^{\bullet}$&$1.032$&$1.060$ ${*}$\\[2mm]
   \hline
-\textit{Number of observations:}  &\multicolumn{2}{l}{$1300$}&\multicolumn{2}{l}{$698$}&\multicolumn{2}{l}{$602$}
+\textit{Number of observations:}  &\multicolumn{1}{c}{$1300$}&\multicolumn{1}{c}{$698$}&\multicolumn{1}{c}{$602$}
 \\
@@
     \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
-            \begin{adjustwidth}{1cm}{} 
-  \item[a] Length of the longest dry spell in number of days.
-    \item[b] Number of dry spells lasting for four days or longer. 
-     \end{adjustwidth}
   \end{footnotesize}
 \end{tablenotes}
+\end{footnotesize}
   \end{threeparttable}\par }

### 2019-01-17 presentations/January2019/presentation1.tex
slidy

@@
 \begin{footnotesize}
-\caption*{Yield as dependent variable\hfill exponents of the coefficients: $\exp{\beta}$}
+\caption*{\small Dependent variable: yield, coefficients in the exponents: $\mathit{e}^{\beta}$}
 %KEN11dK  KEN11dK_ASAL    KEN11dK_nonASAL										
@@
 \setstretch{1.33}
-\begin{tabular}{llll} 
+\begin{tabular}{llll}\hline
+&\multicolumn{3}{c}{\textit{Sample:}}\\
+&\multicolumn{1}{c}{\textit{(I)}}&\multicolumn{1}{c}{\textit{(II)}}&\multicolumn{1}{c}{\textit{(III)}}\\
   \textbf{Fixed effects:}&\textit{\textbf{All counties}}&\textit{\textbf{ASAL}}&\textit{\textbf{non-ASAL}}\\
+  \hline
 \vspace{-0.2cm}Intercept&$1.296^{***}$&$1.272^{*}$&$1.408^{**}$\\
@@
   \hline
-\textit{Number of observations:}  &\multicolumn{1}{c}{$1300$}&\multicolumn{1}{c}{$698$}&\multicolumn{1}{c}{$602$}
+\textit{Number of observations:}  &$1300$&$698$&$602$
 \\
@@
   \begin{footnotesize}
-    \item \textit{Notes:} Standard errors in brackets; \hfill $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
+  \item Standard errors in brackets \\
+    $^{\bullet}~p<0.1$; $^{*}~p<0.05$; $^{**}~p<0.01$; $^{***}~p<0.001$
   \end{footnotesize}
@@
   \end{threeparttable}\par }
+  \setstretch{1}
 \end{frame}
@@
 }
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}
+
+\frametitle{ideas}
+
+\begin{itemize}
+\color{red}
+\item Perhaps as a backup slide-> put a slide with the estimated coefficients, NOT exponents
+\item depending if time>I can copy it as image from the draft paper or copy the tables, separately each subsample
+\end{itemize}
+\color{black}
+\end{frame}
+
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+%oooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooooo
+
+\begin{frame}\label{MarginalEffects}
+
+\frametitle{Marginal Effects}
+\begin{figure}
+    \includegraphics[width = 0.6\textwidth]{Figure1a_1f.pdf}
+\end{figure}
+
+\end{frame}
 
